### 2011-01-07 MScDiplom_Lemenkova.tex
My first random commit message

\documentclass[10pt, a4paper]{article}
\usepackage{color}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{calc}
\usepackage{longtable}
\usepackage[svgnames,table]{xcolor}
%\usepackage[table]{xcolor}
\usepackage{paralist}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{fancyheadings}
\topmargin = 1pt
\headheight=10mm
\headsep=7pt
%\ensuremath
\usepackage{lastpage}
\usepackage{float}
\usepackage{subfig}
\usepackage{caption}
\usepackage{appendix}
\captionsetup{margin=10pt,font=small,labelfont=bf}
\usepackage{wrapfig}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\setcounter{tocdepth}{4}
\parindent 0pt
%\floatstyle{boxed} 
\restylefloat{figure}
\usepackage[a4paper=true, ]{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%% End Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% Begin Lemenkova Polina's MSc Thesis Document%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\pagenumbering{roman}

\title{Seagrass mapping and monitoring along the coasts of Crete, Greece}
\author{by Lemenkova Polina}
\date{\today}
\maketitle
\pagebreak

\textbf{Course Title:} \hfill{Geo-Information Science and Earth Observation for Environmental Modelling and Management}\\

\textbf{Level} \hfill{Master of Science (MSc)}\\
\textbf{Course Duration:} \hfill{September 2009 – March 2011}\\
\textbf{Consortium partners:} \\

University of Southampton (UK)\\
Lund University (Sweden)\\
University of Warsaw (Poland)\\
University of Twente, Faculty ITC (The Netherlands)

\pagebreak

\begin{center}
\textbf{Seagrass mapping and monitoring along the coasts of Crete, Greece}\\
 {by Lemenkova Polina}
\end{center}

Thesis submitted to the University of Twente, faculty ITC, in partial fulfilment of the requirements
for the degree of Master of Science in Geo-information Science and Earth Observation for
Environmental Modelling and Management.\\

Thesis Assessment Board:\\

Name Examiner 1 (Chair)
Name Examiner 2
External Examiner
Member
Supervisors: Venus V. (Department of Natural Resource, ITC)
Toxopeus A.G. (Department of Natural Resource, ITC)\\

\begin{figure}[b]
	\begin{center}
	\includegraphics[height=2cm,width=8cm]{ITC-logo.jpg}
	\end{center}
\end{figure}

\pagebreak

\begin{abstract}
\end{abstract}

\pagebreak
\textbf{Disclaimer}\\

This document describes work undertaken as part of a programme of study at the University of
Twente, Faculty ITC. All views and opinions expressed therein remain the sole responsibility of
the author, and do not necessarily represent those of the university.\\

\pagebreak

\section*{}
\markright{Table of Content}
\addcontentsline{toc}{section}{Table of Content}
\tableofcontents

\textbf{Software and utilities used in the current work}:\\
\begin{itemize}
\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
\item \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
\item WASI Water Color Simulator, for modelling optical water properties under changing environment
\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and GPS for data capture during the fieldwork
\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data preprocessing and preliminary statistical evalution
\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
\end{itemize}
\pagebreak

\section*{}
\markright{List of Figures}
\addcontentsline{toc}{section}{List of Figures}
\listoffigures
\pagebreak

\section*{}
\markright{List of Tables}
\addcontentsline{toc}{section}{List of Tables}
\listoftables
\pagebreak

\pagenumbering{arabic}
\rfoot{MSc Thesis\\ \emph{"Seagrass mapping and monitoring along the coasts of Crete, Greece"}}
\section{Introduction}
%\markright{I. Introduction}

\subsection{Summary}
The seagrasses, a unique group of aquatic plants growing submerged in the sea water, with root-like
structures (rhizoms) buried in the sediments and vertical elongate leaves. A flowering plant,
completely adapted to marine environment, they are however, more closely related to the lily family
(Liliaceae) than to true grass, despite their name “seagrass”, caused by the ribbon-like, grassy leaves
\cite{McKenzie08}\label{McKenzie08}.\\
Seagrasses create unique, complex, extremely diversified and productive ecological systems in the
littoral coastal zones between 0-50 meters in shallow waters all over the world \cite{Hogarth07}\label{Hogarth07}, and
serve as a valuable environmental indicators for the marine ecosystems health. Seagrasses are closely
connected and linked with complex interactions to other vegetation types, e.g. mangroves, coral reefs,
etc. An important constructing component of littoral ecosystems, seagrass contributes significantly to
their structure and functioning.\\
The adaptation to the salt waters is evidently influenced the global ditribution of the seagrasses,
limiting it to shallow coastal areas. A number of critical conditions determines grouth of seagrass,
including general climatic characteristics of the area, i.e.temperature, day length, geological and
geomorphological conditions, e.g. soft type of sediments (sand or mud), shallow depths, as well as
chemical and physical parameters of the water: salinity, waves \cite{McKenzie02}\label{McKenzie02}.
The seagrass Posidonia oceanica (further \textit{P.oceanica}) is a key species to inhabit littoral of the
Mediterranean Sea, see Fig. \ref{fig:1}

\begin{figure}[h]
	\centering
	\subfloat {\label{fig:1-1a}\includegraphics[width=0.4\textwidth]{Fig-1-1a.jpg}}
	\hspace{2mm}
	\subfloat {\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-1-1b.jpg}}
	\caption{Seagrass Posidonia Oceanica}
	\label{fig:1}
\end{figure}

and is widely spread along the coasts of Crete \cite{Dumay02} \label{Dumay02}. 
It plays an important role in a number
of geomorphological and ecological processes. Namely, it is a source of food for herbivorous fauna as
well as helter zones for fish and other marine organisms; it contributes to the nutrient recycling; it
provides sediments stability by reducing the degree of water movements, etc \cite{Francour99}\label{Francour99}.\\
The purpose of current MSc research work aims to apply methods of remote sensing and GIS
including processing and classification of satellite, aerial photos and videometric underwater imaging
as well as GIS-based spatial analysis towards mapping and environmental monitoring of \emph{P. oceanica}
along the coasts of Crete island, Greece. The technical implementation is based on ENVI, Erdas
Imagine, WASI and ArcGIS software using aerial and satellite images as well as results of the
underwater videometric measurements.

\subsection{Background}
\subsubsection{Global distribution of the seagrasses}
Globally, there are 58 recognized and described seagrass species \ref{fig:2}, belonging to two orders
(Hydrocharitales and Najadales), four families (Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
and Zosteraceae), and 12 genera (Enhalus, Thalassia, Halophila, Posidonia, Syringodium, Halodule,
Cymodocea, Amphibolis, Thalassodendron, Zostera, Heterozostera and Phyllospandix) \cite{Kuo89}\label{Kuo89}.
\begin{wrapfigure}{r}{0.4\textwidth}
\centering
	\includegraphics[scale=0.25]{Fig-1-2.jpg}
	\caption{Distribution of seagrasses in the world.\\ Source:\cite{Green03}\label{Green03} \label{fig:2}} 
\end{wrapfigure}
The distribution of the seagrasses is strongly influensed by several environmental factors, which
include climate (mostly, tropical and temperate areas), bathymetry (shallow shelf zones), hydrological
particularities (chemical content of water, nutritient availability and turbidity of waves), and geological 
characteristics - sedimentation and cover types of the seafloor \cite{McKenzie06}\label{McKenzie06}.\\
There are four European seagrass species in Mediterranean area \cite{Borum04}\label{Borum04}: Zostera marina,
Zostera noltii, Cymodocea nodosa and \textit{P.oceanica}. In Greece the common species are\textit{P.oceanica}
(L.) Delile, Cymodocea nodosa (Ucria) Ascherson, Zostera noltii Hornemann and Halophila
stipulacea \cite{Amoutzopoulou-Schina05}\label{Amoutzopoulou-Schina05}. 
These species differ in morphological and
phenological features (Fig. \ref{fig:3}) as well as in structure and dynamics. Thus, Cymodocea nodosa is
considered the pioneer species of \textit{P.oceanica} beds, the latter species forming the last stage. When P.
oceanica beds regresses, C. nodosa often replaces them \cite{DenHartog77}\label{DenHartog77}; as a result, \textit{P.oceanica},
C. nodosa, and Z. noltii do not form mixed persistent stands \cite{Buia91}\label{Buia91}.

\subsubsection{Ecological significance of the seagrasses}
Seagrass plays vital role in the marine ecosystems of the world ocean. Seagrasses are the only
flowering plant in the world that is able to live completely submerged. Seagrass is a habitat for
numerous marine fish species \cite{Nagelkerken00}\label{Nagelkerken00}, source of primary production and food for
fish, turtles and other organisms, which gives them special environmental value \cite{Noralez10}\label{Noralez10}.\\
Seagrass meadows produce enormous quantities of organic matter (leaves, epiphytes), which
constitutes the basis of the food web both within and outside the ecosystem \cite{Gobert06}\label{Gobert06}.

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
	\includegraphics[scale=0.25]{Fig-1-3.jpg}
	\caption{Morphology of different types of seagrasses.
	\label{fig:3}
Source: \url{http://www.ozcoasts.org.au}}
\end{wrapfigure}

Finally, seagrasses are an important component in the environmental "food chain" of the coastal
ecosystems, being the food source for dugongs, turtles, swans and various fish \cite{Cappo95}\label{Cappo95}.\\
Due to their wide distribution, meadows size, easy collection and abundance, sensitivity to the
modifications of the coastal zone and their important role in maintaining coastal water quality
and clarity, seagrass is perfect indicator and descriptor of the environmental health of marine
ecosystems, and is highly suitable for the environmental monitoring \cite{Pergent-Martini05}\label{Pergent-Martini05}.\\
Being often confused with marine “algae”, “seagrasses” are vastly different from them. There are
fundamental differences between both marine organisms, the major of them should be briefly
mentioned: first, seagrasses are true plants with root system and leaves which photosynthesise, have
complex structure and create landscape-similar vast formations on the colonised seafloor areas with
soft sediment, whereas algae are simple organisms that can only holdfast; secondly, seagrasses are
complex vascular plants with reproductive mechanism such as fruits, seeds and spores, while algae
have simple few cell structure with spores and gametes; finally, seagrasses uptake nutrient through
root system, while algae nourish directly from the water column \cite{Dixon05}\label{Dixon05}. There are other
differences between “seagrass” and “algae” but they go beyond the scope of this work.

\subsubsection{Environmental vulnerability of the seagrasses}
Meadows of P. oceanica are subjected to the human activities, as they occur in coastal areas, where
they can be affected both directly \cite{Meinesz91}\label{Meinesz91} or indirectly, through the impact on the
quality of waters and sediments \cite{Duarte02}\label{Duarte02}. As \textit{P.oceanica} is a long-living plant with a slow
growth rate, the anthropogenic modifications of the coastal zone, happening more rapidly than the
capacity of the plant to adapt to these changes, reduce its distribution area \cite{Micheli05}\label{Micheli05}. One
of the main drivers of seagrass decline is, for instance, the location of the fish farming near the
seagrass meadows. The negative effects of the sedimentation of waste particles in the farm vicinities
on \textit{P.oceanica} meadows are diverse and complex, and may cause benthic deterioration, accumulation
of organic matter and seagrass decline \cite{Holmer08}\label{Holmer08}. Seagrasses are subject to anthropogenic
nutrient (N and P) loading, which may occasionally casue morphological (e.g. leaf length) and
physiological (e.g. chlorophyll and nitrogen content of the leaves) responses towards changed
environmental conditions \cite{Leoni06,Leoni07}\label{Leoni06} \label{Leoni07}.

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[scale=0.20]{Fig-1-4.jpg}
\caption{Distribution of seagrass in relation to mean
ocean temperature. \\
Source: \cite{Orth06}\label{Orth06}}
\label{Fig.4}
\end{wrapfigure}

The detailed research of the fish
farm-induced decline of the seagrass meadows \cite{Diaz-Almela06}\label{Diaz-Almela06}reports the relationships of
fish farm organic and nutrient content in the sediments with dynamics of the key seagrass species (P.
oceanica) in the Mediterranean Sea. Nowadays P.oceanica is in the alarming state of regression due
to the deterioration of the environment in the Mediterranean Sea \cite{Ribed02}\label{Ribed02}. Due to these reasons,
\textit{P.oceanica} is a protected species since 1988 in some European countries \cite{Francour99}\label{Francour99}, and
its presence serves as an indicator of a stable healthy environment. Among other negative factors,
affecting both growth and status of the seagrasses the environmental contaminants can be mentioned,
e.g. thermal, sewage, dredging and chemical pollution as well as any other kind of maritime works,
e.g. trawling and anchoring of boats \cite{Ribed02}\label{Ribed02}. Other human activities that cause degrading of the
seagrass are recreational boating, commertial overexploitation of coastal resources, eutrophication
\cite{McKenzie06}\label{McKenzie06}.\\
Besides anthropogenic factors, various biochemical, climatic and environmental processes can cause
negative influence on seagrass distribution. Seagrass is exposed to threats from the global climate and
environmental change, i.e. increases in sea surface temperature; sea level rise; increased frequency
and intensity of storms and waves; local decrease of water quality, increased sedimentation,
contaminantion and nutrification; desiccation; salinity fluctuations; nutrient changes; suspended
sediments \cite{Blake00}\label{Blake00}. These stress-drivers can alone result in large-scale seagrass
degradation, but often seagrass undergo simultaneous affects from several of these factors together.
This naturally increase environmental pressing and leads to drastical loss of very large areas of
seagrass globally \cite{Orth06}\label{Orth06}.
In tropical areas, where most of seagrasses are located (Fig.\ref{Fig.4}), seagrasses are subject to catastrophic
extinction and loss, due to the cyclones, typhoons, storms, regular floods and increased rainfalls. Recovery from such
events can take up to several years and often it is only possible by means of the seed reserves from the local
environmental surveys \cite{McKenzie07}\label{McKenzie07}. Other threats for seagrasses in tropical areas are 
increased nutrient availability in the coastal zones, increased
eutrophication and invasive macroalgae. These processes have strong affect on the status of the
seagrass meadows, and often lead to their complete disappearance \cite{Holmer09}\label{Holmer09}.
Other environmental threat for the seagrasses arise as a result of the environmental struggle and
competition for existence among species. Thus, meadows of P.oceanica in the Mediterranean Sea are
presently facing invasion by alien algal species, partuicularly in areas where P.oceanica is already
degrading, stressed, have gaps and patchy structure in meadows and show other signs of regression
\cite{Montefalcone10}\label{Montefalcone10}.
Seagrasses are vulnerable fragile species, important for the marine coastal ecosystems, especially for
the protection of the beach structure. However, the facts about seagrass global degrading sound
worrying: about 54 percent of the total seagrass meadows have lost any part of their area; the areas, where
the seagrass ecosystems are degrading or lost, are not located in a specific area or continent, but
registered globally; since 1980s global losses of the seagrases on our planet is equal to two footbal
fields per hour \cite{Mellors09b}\label{Mellors09b}.

\subsubsection{General characteristics of Posidonia oceanica}
The endemic Mediterranean seagrass Posidonia oceanica (further \textit{P.oceanica}), is a main species in
marine coastal environment of Greece, forming, despite its slow growth, the largest, most widespread,
homogeneous and dense meadows in the Mediterranean between 5 and 40 m depth \cite{DenHartog70}\label{DenHartog70}.

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[scale=0.20]{Fig-1-5.jpg}
\caption{Geographical distribution of
\textit{P.oceanica}. Source: \cite{Borum04}\label{Borum04}}
\label{fig:5}
\end{wrapfigure}

The dominant and most productive coastal ecosystem of the Mediterranean, P.oceanica is
spatially restricted to the Mediterranean area (Fig.\ref{fig:5}), with its extention limited by the western part
of the Mediterranean Sea where cold Atlantic waters enter Gibraltar and mix with warm
Mediterranean waters, thus decreasing its temperature.
Morphological P.oceanica consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots
and short rhizomes (0.5-2.0 mm). The leaves are, perhaps, the most particular characteristics of

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[scale=0.15]{Fig-1-6.jpg}
\caption{\textit{P.oceanica}.\\
Source \cite{Luque04}\label{Luque04}}
\label{fig:6}
\end{wrapfigure}

P.oceanica, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:6}): having
usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}).
Growing P.oceanica make a meadows which, in turn, consisnt of smaller patches, called “matte”
(Fig.\ref{fig:7}), a monumental construction made by the growth of rhizomes and leaves with entangled roots
and entrapped sediment \cite{Francour06}\label{Francour06}.
Representing one of the most productive Mediterranean ecosystems, \textit{P.oceanica} usually serves as a
perfect biological indicator for the assessment of the quality of waters and environmental health
\cite{Boudouresque89}\label{Boudouresque89}. Some authors \cite{Guidetti08,Montefalcone09}\label{Guidetti08} \label{Montefalcone09} used status and population dynamics of \textit{P.oceanica} as indicators for the evaluation

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[scale=0.15]{Fig-1-7.jpg}
\caption{Scheme of matte structure of
P.Oceanica.\\ Source \cite{Pergent90}\label{Pergent90}}
\label{fig:7}
\end{wrapfigure}

of the meadow health status. \\
There are several environmental factors, determinating the growth of the \textit{P.oceanica}. The adaptation
to dry-summer subtropical climate reduces its extention to Mediterranean area only (Fig.\ref{fig:5}).
Besides, the distribution of the seagrasses changes with water depth: it is noticed \cite{Dural10}\label{Dural10} that
the highest flowering density is usually in the 4-7 m depth. 
\textit{P.oceanica} flowers appeared in shallow
stands in September while in November only in stands deeper than 15 m. This time delay is caused by
the different maximum summer temperatures at those depths \cite{Buia91}\label{Buia91}.

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[scale=0.15]{Fig-1-8.jpg}
\caption{Structure and components of
P.oceanica \\ Source \cite{DiCarlo04}\label{DiCarlo04}}
\label{fig:8}
\end{wrapfigure}

The phenology of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in
shallow waters plants have shorter and falciform leaves, compared to ones in the deeper and central
areas \cite{Dural10}\label{Dural10}. In \textit{P.oceanica} flower abundance is related to the structure of the meadow with
the maximal flower density in the densest stands, while the occurrence of flowering is regulated by
environmental factors \cite{Buia91}\label{Buia91}. Phenology of the \textit{P.oceanica} undergo
modifications with varying seasons during the year: during the flowering period (ca 3 months long)
the number of leaves on the flowering shoots decrease. Changes of the leaf growth also appear in the
flowering shoots with longer oldest leaves and shorter and narrower leaves induced during the
flowering \cite{Gobert01}\label{Gobert01}.

\subsection{Research problem}
Monitoring of the marine benthic ecosystems of seagrasses is essential for the environmental
assessment of the coastal zones. It increases our knowledge of the seagrass ecology, highlightes
threats to the seagrass and preventing them from possible losses and degrading and improves
techniques and methods of the underwater-based observations. Mapping the seagrass contributes to
the evaluating of the seagrass current distribution, analysis of its dynamics and changes over time, as
well as estimations of the degrading of seagrass meadows for the purpose of the coastal management.\\
Precise, correct and up-to-date information about the distribution of \textit{P.oceanica} is necessary for the
sustainable conservation of the marine environment and ecosystems in Mediterranean area, being an
important contribution to the environmental coastal zone management \cite{Pergent-Martini06}\label{Pergent-Martini06}.
However, mapping the sesgrass has limitations due to the specific location and characteristics of the
research object. The remote sensing techniques have traditionally been widely used for the seagrass
monitoring.

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[scale=0.25]{Fig-1-9.jpg}
\caption{Spectra of the seagrass on different
depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}}
\label{fig:9}
\end{wrapfigure}

The general overview of the application of various remote sensing data types (colour,
infrared, and black and white images) for the seagrass monitoring shows its high suitability and
potential as a research method \cite{Pasqualini01,Matarrese06}\label{Pasqualini01}\label{Matarrese06}. The using of the aerial
photographs as base maps for the seagrass meadows mapping is, perhaps, the most traditional
application \cite{McKenzie03,Kendrick00,Pasqualini99}\label{McKenzie03} \label{Kendrick00} \label{Pasqualini99}. The results of the image
processing of colour aerial photographs for the monitoring of littoral environment with seagrass beds
have been reported by several authors \cite{Kelly80,Walker89,Green96}\label{Kelly80} \label{Walker89} \label{Green96}.
Satellite imagery processing has also being used for the seagrass monitoring, due to their accuracy,
repeatability and informativeness as a source of data \cite{Dekker05b}\label{Dekker05b}, enabling regular temporal
coverage over the large remote areas and providing a cost-effective approach for the mapping of the
remotely located feature, such as underwater vegetation \cite{Jensen95}\label{Jensen95}. Satellite images provide
with detailed information on seagrass canopy and other environmental indicators \cite{Fyfe03}\label{Fyfe03}.
Various research papers report successful application of the image processing for the
seagrass mapping \cite{Calvo03,Dekker05a,Fornes06,Green96,Jackson07,Jensen95,Kendrick00,Lyzenga81,Malthus03,Matarrese08,Mount03,Pasqualini01,Pasqualini99,Pasqualini98a,Pergent-Martini06,Ralph05,Ribed02,Short01,Walker89}\label{Calvo03}\label{Dekker05a}\label{Fornes06}\label{Green96}\label{Jackson07}\label{Jensen95}\label{Kendrick00} \label{Lyzenga81}\label{Malthus03}\label{Matarrese08}\label{Mount03}\label{Pasqualini01}\label{Pasqualini99}\label{Pasqualini98a}\label{Pergent-Martini06}\label{Ralph05}\label{Ribed02}\label{Short01}\label{Walker89}. The application of the remote sensing data towards seagrass mapping is based on the spectral reflectance
 characteristics of the \textit{P.oceanica} seagrass, which enable its spectral discrimination from spectra of other seafloor types. 
It is proved \cite{Thorhaug07} \label{Thorhaug07} that the spectral signatures of different species of tropical seagrasses are well
distinguishable from each other. The application of the methods of images classification for seagrass
mapping is based on the classifying the pixels on the image according to their spectral reflectance
values \cite{Fornes06} \label{Fornes06}, so that the seafloor can be divided into several types: sand, rock, P.
oceanica, other vegetation, etc.
Other example of monitoring of \textit{P.oceanica} using remote sensing techniques \cite{Calvo03}\label{Calvo03}
reports the application of the CZCS images towards the case study of the Italian coast and shows
successful results of the neural-based classification using Isodata method of supervised classification.
In case of \textit{P.oceanica} meadows aerial and satellite images are particularly suited for the surveying
shallow waters \cite{Pasqualini98b}\label{Pasqualini98b} enabling to distinguish seagrass formations and dynamics of
the temporal evolution of seagrass meadows over the research area \cite{Pasqualini01}\label{Pasqualini01}.
However, using space borne satellite imagery for the seagrass mapping has certain limitations, due to
the uncertainties of the spectral signature of the seagrass at higher depths  (Fig.\ref{fig:9}), as well as some
optical particularities, e.g. light refraction under water, unevenness of the water surface, depths, etc.
Some problems can also arise in the images interpretations, as quite different objects may have
similar spectral reflectance, e.g. seagrass, dark-coloured bottom patches (mud), macroalgae.
The in-situ fieldwork including underwater videographic measurements is an important part of the
seagrass monitoring, and has been successfully applied towards seagrass mapping \cite{Haag08}\label{Haag08}.
The underwater measurements are used to validate the results and to receive detailed, accurate and
precise data for the selected locations. The underwater measurements cannot be applied for the whole
research area, however it provides with detailed monitoring along the route of the boat. Therefore, in
the selected locations it becomes a useful tool for the assessment of the distribution, density and
coverage of the seagrass along the tracklog. Besides, the underwater observations using scuba diving
equipment have been conducted for the measurements of depths.

\subsection{Research objective}
The current MSc research aims to explore the environmental conditions and dynamics of the spatial
distribution of \textit{P.oceanica} seagrass meadows along the Cretan coast, based on the remote sensing and
GIS techniques, knowledge about the coastal environment in Crete and integration of various data
from the following sources:
i) spectra of P.oceanica, carbonatic sand, silt and other seafloor types
ii) satellite imagery: Landsat, ASTER, MODIS, MERIS
iii) aerial photos: Google Earth
iv) in-situ fieldwork data of underwater videographic measurements
v) vector GIS layers.
The research aims to explore the dynamics of the environmental changes of the P.oceanica meadows
over time (2000-2010), as well as the extent of the spatial distribution over the whole island. The
main research objective is monitoring the seagrass \textit{P.oceanica} along the coasts of Crete, in order to
analyse the environmental changes of the P.oceanica meadows over the last 10 years (2000-2010)
and the spatial extent of its current distribution.
This research is supported by the in-situ measurements in two selected locations of the northern coast
of Crete (Ligaria beach in Agia Pelagia district and Xerocampos), using following methods of the
remote sensing techniques: underwater videometric measurements made by Olympus camera, aerial
Google Earth and satellite images from different sources, spatial GIS and statistical analysis.

\subsubsection{General objective}
The main objective of this study is to investigate the distribution of seagrass along the northern coast
of Crete using remote sensing techniques.
General objectives:
1) Mapping the spatial distribution of the seagrass \textit{P.oceanica} along the northern coasts of
Crete Island
2) Monitoring the dynamics of the environmental changes of the P.oceanica meadows over
time (2000-2010) in Ligaria beach, Agia Pelagia district.

\subsubsection{Specific objectives}
1. To apply remote sensing data (Google Earth aerial images, Landsat, SPOT satellie images)
for the monitoring of the seagrass meadows distribution
2. To study narrow-band spectral reflectance propoerties of \textit{P.oceanica} and other seafloor
cover types (sand and silt) using WASI swater colour simulation software
3. To use methods of the in situ diving observations and underwater videometric measurements
by Olympus camera in order to receive large-scale imagery of the P.oceanica mattes
4. To perform images classification (supervised and unsupervised) for the thematic mapping of
the P.oceanica seagrass distribution along the coasts of northern Crete.

\subsection{Research questions}
1. Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ
conditions ?
2. Whether broadband and hyperspectral remote sensing data can be used for the mapping of
\textit{P.oceanica} ?

\subsection{Hypotheses testing}
A statistical testing will be used to compare between the spectral responses of the different seagrass
types, whether it is spectrally distinct and at least one pair is statistically different at every spectral
band.
Hypothesis Ho: seagrass types are not spectrally distinct from other seafloor types with
varying in-situ conditions, which means
Ho: μ1 =μ2 =μ3=...= μn.
The alternative Hypothesis Ha claims the opposite statement: seagrass is spectrally
distinct with varying in-situ conditions, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn.
The distribution of the spectral responses at every spectral band is assumed to be normal as well as
the equality of the statistical variances.
The hypothesis testing is suggested to be carried out using the anova statistical test. The purpose of
anova test is to visualize in an effective and quick way the spectral differences between seagrass
species and their spatial distribution. Thus, the key hypotheses of the research will be tested to prove
whether the results of the research are accurate, reasonable and correct.

\subsection{Research scheme}

\subsection{Research approach}
Seagrass consistent monitoring and mappint is necessary and important for the sustainable coastal
development and conservation measures. Earlier, many seagrass meadows have been destroyed by
human activities in the coastal zone, mainly due to the ignorance of their existence, because
information on the seagrass bed exact location was not available \cite{Choo06}\label{Choo06}. Well-time seagrass
observations and mapping enables precise control of its spatial distribution, detection of any changes
in the seagrass landscapes, highlightes potential environmental threats in the coastal zone (e.g.
declining of meadows) before they become unmanageable for the coastal management services.
Choosing the right and most effective approach method for the seagrass monitoring is essential.
Remote sensing methods alone, though having evident advantages, are insufficient, because satellite
images of underwater habitats are notoriously difficult to identify and interpret. The best research
method should be based on the integrated approach, well described in various scientific works \cite{Brown02,Montefalcone06,Kirkman96}\label{Brown02} \label{Montefalcone06} \label{Kirkman96}, which includes combination of various
techniques of the seagrass monitoring, i.e. remote sensing imagery classification of aerial and satellite
images, GIS-based spatial analysis and ground in-situ surveys.
The current study is based on the application of the remote sensing data, broadband satellite imagery,
aerial images and the results of the underwater videographic measurements towards seagrass
mapping. Image classification is based on the principle of the differentiation between the spectral
signatures of various seafloor cover types. The spectrum of light coming up from the ocean surface in
shallow waters keeps information on the optical properties of the seawater components and benthic
substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. Spectral
reflectances, distinct for each seafloor cover type, enable their discrimination on the image. The preprocessing
of the images includes imagery corrections for atmospheric noises and effects of the water
column. Reflectance spectra of the seagrass canopy at different depths of the water-column are
analysed for the discrimination of their spectral signatures, enabling to separate various seafloor types
during classification. The results of the of imagery classification are analysed for the detection of the
dynamics in \textit{P.oceanica} seagrass distribution along the northern coasts of Crete. Aerial imagery from
Google Earth with high spatial resolution, suitable for the large-scale detailed mapping of seagrass
mattes, is used for the improvement of the accuracy of large seagrass meadows and separate mattes
within the meadows. The in situ underwater videometric measurements of the seafloor are collected
during the fieldwork in Crete, for the validation of the classification results and to determine the exact
current distribution of the \textit{P.oceanica} meadows. The image processing includes steps of the remote
sensing techniques, i.e. calibration, masking from land and cloud, atmospheric correction, sea surface
glint and depth effects correction as recommended \cite{Matarrese08}\label{Matarrese08}. During the image classification
working step the training sites for the supervised classification methods are designed, as well as its
control and trials of different classification approaches (Unsupervised, K-means or Isodata;
Supervised, Maximum Likelihood).
\pagebreak

\section[Overview of literature]{Seagrass monitoring: overview of literature and research resources}
%\markright{II. Seagrass monitoring: overview of literature and research resources}
%\renewcommand{\headrulewidth}{0.4pt}

\subsection{Seagrass global monitoring: history and perspectives}
Mapping and monitoring the seagrass is important for the environmental assessment of the marine
ecosystemsn in coastal areas. Regular tracking of current distribution of seagrass meadows, based on
correct information and cartographic visualization of seagrasses, is a preventive environmental
management, which helps to analyse potential environmental risks of coastal areas, decrease of the
number of species, loss of meadows and patches of the seagrasses.
The tradition of global seagrass mapping though has not a very long history comparing to the
terrestrial cartography, due to the technical difficulties of underwater observations, yet nowadays is
has become a successful, rapidly developping, increasingly popular and challenging research branch.
Seagrasses can be mapped using various approaches, with the most popular method of the remote
sensing application, enabling to detect and discriminate the substratum and various vegetation types
growing on the seafloor.
Several global seagrass surveys organise and provide regular monitoring of the segrass distribution,
health and species sustainability, amongst which the most world-known are, perhaps, the following
ones: \href{http://www.seagrassnet.org/}{Global Seagrass Monitoring project}, \href{http://www.seagrasswatch.org/}{Seagrasswatch}, 
located in Australia, \href{http://mediterranean.seagrassonline.org/}{the Mediterranean association Seagrass-2000},
\href{http://www.imedea.uib.es/index.php}{the Mediterranean Institute for Advanced Studies},
\href{http://www2.fiu.edu/~seagrass/}{Seagrass Ecosystems Research Laboratory in South Florida},
\href{http://www.flseagrass.org/}{Florida Seagrass organisation}, \href{http://www.seagrasses.org/}{the Seagrasses.org} and the 
\href{http://wsa.seagrassonline.org/}{World Seagrass Association}.
 These organizations aim at global seagrass monitoring, providing with
research results and reporting guidelines and manuals with standartized methods and
recommendations, specific for the seagrass research and monitoring. There are also university marine
centres and research institutes conducting seagrass monitoring and as a particular part of their
research. Their reports and guidelines were used for references in the current research.
Regular obserbvations and monitoring of the seagrasses are known since 1960s, mainly in tropical
regions (Australia). Since that time traditional methods of the seagrass monitoring and common
recommendations are being elaborated. The development of the underwater SCUBA diving
equipment and devices enabled to conduct underwater detailed measurements and observations
largely contributed to the improvement of the traditional in-situ observations of seagrass. From the
other side, development of the remote sensing methods and data acquisition from space contributed to
the new methods of seagrass mapping, using distance approach and generally based on images
classification.

\subsection{\textit{In-situ} observations of the seagrass meadows}
The traditional methods of in-situ seagrass monitoring include in general the following standard
scheme \cite{McKenzieetal03} \label{McKenzieetal03}. The seagrass is being sampled on the selected sites using transect lines,
quadrant frame, single point markers, markers, GPS and other equipment. The seagrass sampling is
taken on the regular way with observation polints covering the study area with normal distribution.
During the measurement process, the vertical photograph of the measurements frame is taken, and the
following points are traditionally estimated: percentage of the seagrass cover withing the quadrate,
species composition, sediment composition, canopy height, epyphyte abundance, algae percent cover,
count of microfauna and a specimen of seagrass is being taken. This scheme, well described by
McKenzie \cite{McKenzieetal03}\label{McKenzieetal03} is widely used and well-known among the marine biologists and
seagrass researchers. Applications of the in-situ seagrass observations of the structuring epiphyte
community composition in the P.oceanica ecosystems in Mediterranean Sea is, for example,
described by Villegas \cite{Villegas06}\label{Villegas06}. Realization of traditional methods for mapping seagrass usually
involves intensive and time-consuming in-situ observations during the fieldwork, as, for example,
reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. Other methods of seagrass in-situ monitoring are based on
using the active hydroacoustic sonar sensors that send towards a sea floor a signal of energy and then
collect the return echoes for the analysis, yet this requires specialised equipment and is mostly used in
the depth waters, combined with bathymetric measurements. Another limitation of the acoustic
techniques is that they are not effective at identifying the composition or even the presence of
submerged aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}, because
initially acoustic methods were intended for the bathymetric surveying, i.e. they are not adjusted for
the benthic habitat discrimination. The results of the in-situ measurements and observation are usually
managed and treated using integrated GIS approach, as, for example, reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.

\subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}
The application of the remote sensing data for seagrass mapping is based on our knowledge of the
spectral reflectance properties of the target objects, and using it for the classification of these objects
on the image. In case of seagrasses it is spectral reflectance of the seafloor cover types, which can be
analysed using measurements of optical properties of sea water: radiance and irradiance.
Optical remote sensing methods can get through the clear waters to approximately 15–30 m \cite{Mumby04}\label{Mumby04}.
 When sunlight enters the waters and goes down into the water column, parts of the
electromagnetic energy are absorbed and scattered, which is determined by the optical and physical
properties of the water, e.g. concentration of suspended particles, chlorofill, coloured dissolved
organic matter (gelbstoff) that make up the water content \cite{Ackleson2003}\label{Ackleson, 2003}. Besides, light is strongly
dependent on wavelengths, i.e. it is greater in blue wavelengths (400 nm) than in others.

\begin{figure}
	\centering
	\subfloat [RAMSES-ACC-UV - Hyperspectral UVA/UVB Irradiance Sensor: 280-500 nm]{\label{fig:1-1a}\includegraphics[width=0.4\textwidth]{Fig-2-1.jpg}}
	\subfloat [RAMSES-ARC - Hyperspectral UV-VIS Radiance Sensor: 320-950 nm.]{\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-2-2.jpg}}
	\caption{RAMSES Hyperspectral Sensors. Source: \href{http://www.trios.de}{Trios}}
	\label{fig:10}
\end{figure}

The optical properties of the sea water vary with different environmental conditions and reflect
current chemical content and physical specifics of the water, revealing the variability and distribution
of color of the sea waters, determined by the material in the water, e.g. chlorophyll, as well as its
physical properties, e.g. water absorption, attenuation, backscattering \cite{Maffione01a}\label{Maffione01a}.
Shallow waters generally contain more dissolved substances and suspended particles, which directly
influnces the transparency and colour of the waters of shelf zones \cite{Jerlov52}\label{Jerlov52}. Being highly
dynamic environments, coastal waters experience a variety of processes which alter their optical
properties incessantly. The effects of these processes influence application of the hyperspectral
remote sensing and reinforce other proceses \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides, increase
sedimentation processes which in turn may change microrelief properties and topology. 
The optical properties of the water are best reflected in the values of its radiance and irradiance,
which can be converted into spectral reflectance, or reflectivity. The spectral irradiance (E) is a
radiant flux of the electromagnetic solar radiation energy, received per surface unit area in a given
time (W·m−2·nm−1), while radiance (L) characterizes total emission or reflection that passes through
or is emitted from a particular area (W·sr−1·m−2 ·nm−1). Therefore, the spectral reflectance, or the
reflectivity of the object, can be estimated by the direct mathematical division of these first two
characteristics, and is expressed in persentage.
Therefore, the irradiance and radiance of the water should be measured in order to estimate spectral
reflectance of the various seafloor cover types. The optical measurements of the irradiance and
radiance of the sea water and bottom cover types of the seafloor can be received by the means of the
optical sensors spectroradiometers. There are several companies producing radiometers with various
characteristics and adjusted for defferent purposes, e.g. portable and miniature spectrometers from
\href{http://www.stellarnet.us}{StellarNet}, Hyperspectral Ocean Colour Radiometer (HyperOCR) sensor by
\href{http://www.satlantic.com}{Satlantic}, traceable spectroradiometers by \href{http://www.orboptronix.com}{Orbotronix} 
and lots of others. Among other radiometers there are ones designed by the \href{http://www.trios.de/}{Trios company} producing
optical sensors, GER-Series Field portable spectroradiometers from
\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. Trios-RAMSES hyperspectral radiometers
(Trios-RAMSES Hyperspectral UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-
VIS Radiance Sensor) are small-sized, low power-consumpting, flexible for fieldwork yet with high
level of precision, specially calibrated for air and for water application as well as colour
measurements  (Fig.\ref{fig:10}). These products have been used for the radiance and irradiance
measurements in Agia Pelagia bay, Crete island, 2009.

\subsection[Application of remote sensing data]{Application of remote sensing data towards seagrass mapping}
Various methods and approaches have been applied towards mapping of the seagrasses, based on
digitized aerial photographs, GPS data, remote sensing and SCUBA-based fieldwork measurements.
SCUBA-based in-situ observations, though providing high resolution and accuracy results in seagrass
mapping, is limited in application, because of their time consumption, weather-dependancy and
unsuitability for the case of monitoring large areas of water for small-scale mapping. Methods of
underwater videography with GPS is a tool of seagrass monitoring which has certain advantages, i.e.
high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
collection in the field \cite{Schultz08}\label{Schultz08}. However, it cannot cover large areas for small-scale mapping.
Remote sensing techniques offer clear advantages over other methods of in-situ field measurements
and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
their weather-independency, cost-effectiveness, accuracy and spatial coverage, which enables
periodic monitoring of the seagrass mewdoes and gives access to the distant and unapproachable
areas. Integrated together with GIS vector layers and maps, remote sensing data enable historical
mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.\\
However, application of the remote sensing techniques for mapping of submerged vegetation,
seafloor cover types and benthic vegetation, inter alia seagrasses, are still in their development.

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[scale=0.30]{Fig-11.jpg}
\caption{In situ optical reflectance spectra of
seagrass. Shaded areas - percent of spectra lying within
the range of reflectance. White lines - mean spectra.
Source: \cite{Hochberg03b}\label{Hochberg03b}}
\label{fig:11}
\end{wrapfigure}

Approaches and methods for seagrass protection and monitoring still remain location specific or, at
least, nation specific, depending to large extent on the tools available for researchers \cite{Mellors09a} \label{Mellors09a}.
Universal, international, standardised methods for seagrass directly for seagrasses
as such still should be developed. Various case studies have been performed, yet their mostly report
methods adjusted for particular areas, without evaluating starndard general algorithms that could be
extrapolated towards other regions. \\ Application of the remote sensing towards seagrass mapping is
generally based on the assumption that various types of the seafloor bottom have different
characteristics of the reflectivity, which is visually expressed in distinct colors of the objects. In its
turn, reflectivity of the sediments is affected by the water optical properties and content. For
example, Stephens \cite{Stephens03}\label{Stephens03} prove that microalgal biomass and community structure affect
hyperspectral reflectance of sediments, which enable to estimate total microalgal biomass from
measurements of hyperspectral reflectance. Spectral measurements of the target objects are made
my the means of the radiometers, which receive and register the amounts of enegry (radiance and
irradiance) from the objects. \\ Measuring optical properties of water allow to calculate spectra of the
objects and to discriminate them on the aerial and satellite images. Thus, various scientists report
success in spectral discrimination of submerged vegetation and other seafloor cover types on imagery
using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01} \label{Louchard03} \label{Dehouck08} \label{Werdell03}.
Studies of spectral reflectances of the different seagrass species comparing to the spectra of sand and
other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
macroalgae differ from each other, as well as from sand and deep water reflectance spectra. These
differences are well detectable by the means of the remote sensing research methods. Comparing to
the terrestrial plants, aquatic vegtation inter alia seagrass cannot be detected using red edge of the
spectrum, as these wavelengths are significantly absorbed by water \cite{Kirk94}\label{Kirk94}, as well as by
scattering and absorption by phytoplankton. Some authors \cite{Dierssen03}\label{Dierssen03} also use
spectrally based radiative transfer approach to quantitatively estimate shallow-water bathymetry and
leaf area index (LAI) of the seagrass. The spectral reflectances in general are the
result of the spectral absorption in different bands, typical for each target object.
Spectral reflectance of the seagrass  (Fig.\ref{fig:11}) is largely influenced by the water
depth where it is located, and is generally decreasing in values by increasing depths.
The most important spectral diapason for marine mapping of submerged vegetation
and, particularly, for seagrass, is 350-800 nm \cite{Barille09}\label{Barille09}.
Using airborne imagery for retrospective data (before 1970s) together with the most
recent imagery allows to detect changes in seagrass distribution on over different years
and to analyse dynamics of the seagrass distribution\cite{Ball09}\label{Ball09}. Another important advantage
of the application of the remote sensing data for mapping of submerged aquatic vegetation has
commertial nature: using remote sensing data and methods enables more low-cost and up-to-date
seagrass mapping \cite{Mumby99}\label{Mumby99}, and is especially useful for the areas where the fieldwork data
capturing is unavailable.

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[scale=0.25]{Fig-12.jpg}
\caption{Difference between broaddand
multispectral and hyperspectral resolution of
spectral signatures. Source: \cite{Clark99}\label{Clark99}}
\label{fig:12}
\end{wrapfigure}

Seagrass meadows may reach spatial scales from several up to hundreds of metres, therefore they are
susceptible by the means of satellite imagery from remote sensors, both with moderate resolution
(e.g., \href{http://landsat.gsfc.nasa.gov/}{Landsat MSS, Landsat TM, Landsat ETM+}, \href{http://envisat.esa.int/instruments/meris/}{MERIS}, \href{http://asterweb.jpl.nasa.gov/}{ASTER}, \href{http://modis.gsfc.nasa.gov/}{MODIS}) and high resolution as
well (e.g., \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird}, \href{http://www.spotimage.fr/}{SPOT}, \href{http://www.itres.com/products/imagers/casi550}{CASI}). The possibility of their application towards seagrass mapping
vary and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:12}). In the next subparagraphs
we briefly discuss the limitations and research experience of the using of various imagery
for the seagrass mapping.

\begin{table}[htbp]
\caption{\textbf{Characteristics of selected ocean-colour sensors}}
\begin{center}
\rowcolors{1}{OldLace}{FloralWhite}
\begin{tabular}{|p{2cm}|p{20mm}|p{15mm}|p{17mm}|p{15mm}|p{15mm}|p{15mm}|}
\hline\hline
\textbf{Satellite Sensor:} & {\textbf{Landsat}} & {\textbf{SPOT}} & {\textbf{ENVISAT}} & {\textbf{Terra}} & {\textbf{GeoEye}} & {\textbf{Nimbus 7}} \\ \hline\hline
	\textbf{Origin} & USA & France & Europe (ESA) & USA (NASA) & USA & USA \\ \hline
	\textbf{Instrument} & \textit{TM, ETM+} & \textit{Pan, XS} & \textit{MERIS} & \textit{MODIS} & \textit{SeaWiFS} & \textit{CZCS}\\ \hline
	\textbf{Number of channels} & 8 & 6 & 15 & 36 & 8 & 5 \\ \hline
	\textbf{Wavelength coverage} & 185-1700 & 500-1750 & 412.5- 900 & 469-1640 & 412-865 & 443-750 \\ \hline
	\textbf{Ground resolution (nadir), m} & 15-pan,30(MS),60 (TIR) & 2.5, 5, 10,20 & 1.2 km/300 m & 1.0 km & 1.13 km & 825 m \\ \hline
	\textbf{Band Width (nm)} & 0.45-10.4 & 10,20 & 2.5,7.5,9,10 14,20 & 0.18,0.3,0.5,10 - 50 & 20,40 & 20,100 \\ \hline
	\textbf{Launched} & 1972, 1999 & 1986 & 2002 & 1999 & 1997 & 1978 \\ \hline
	\textbf{Recurrent period, days} & 16 & 24 & 35 & 16 & 16 & 16 \\ \hline
\end{tabular}
\end{center}
\label{tab:1}
\end{table}

\subsubsection{Review of multispectral imagery used for seagrass mapping}
Seagrass mapping using remotely sensed data from multispectral sensors is based on the classification
and discrimination of the seafloor types using their spectral characteristics in different wavebands.
Perhaps, the most known imagery, widely used in the remote sensing mapping, is received from the
group of Landsat sensors, known for its historical, pioneeer role in the satellite industry.
Being the longest running satellite system, lunched in 1972, Landsat is the only source of archival
data going back to 1984 at a sufficient spatial resolution \cite{Dekker05}\label{Dekker05}, which makes its data
desirable for historial mapping or environmental analysis of change detection of the seagrass
landscapes.
The Landsat TM and Landsat ETM+ data, with the most recent from sensor Landsat 7, an advanced
and multispectral scanning, launched in 1999, proved to be feasibile and useful for the mapping of
submerged vegetation, such as seagrasses or coral reefs. The successful applications of the imagery
Landsat TM towards the seagrass mapping, were reported in numerous research works \cite{Palandro03,Gullstroom06,Wabnitz08,Bierwirth93,Ferguson97,Rasib97}\label{Palandro03} \label{Gullstroom06} \label{Wabnitz08} \label{Bierwirth93} \label{Ferguson97} \label{Rasib97}. The Landsat data are particularly suitable for the
case of change detection of seagrass landscapes at a decadal scale, because being the main sensor
onboard the Landsat satellites, Landsat Thematic Mapper (TM) provides the longest time series
available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[scale=0.35]{Fig-13.jpg}
\caption{Band coverage of ASTER and Landsat channels on
the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
\label{fig:13}
\end{wrapfigure}

Another well-known multispectral sensor, SPOT provides multispectral imagery with a spatial
resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
orbital repeat cycle for nadir viewing and imagery with a spatial resolution 20 - 2.5 m
\cite{SPOT}\label{SPOT}. \href{http://www.spotimage.fr/}{SPOT} imagery was used for mapping beds of Posidonia oceanica in the
Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}. The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
available high-resolution imagery at 1- and 4-meter resolution from \href{http://www.geoeye.com}{Geoeye}. IKONOS
imagery has been applied for the seagrass mapping due to its high resolution and accessibility. Thus,
the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
made using IKONOS, Landsat TM, and CASI, show that in the blue part of the
spectrum, the best results are achieved by IKONOS and CASI, while Landsat TM has not high
enough resolution. It may be caused to some extent by the loss of the radiance contrast, atmospheric
Rayleigh scattering and defects of scattering. However, comparing between CASI and IKONOS, the
same authors prove that CASI enable to receive still more accurate results of the the classification
than IKONOS \cite{Mumby02}\label{Mumby02}.
Another comparative analysis of the application of \href{http://www.itres.com/products/imagers/casi550}{CASI}, Landsat and \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird} imagery \cite{Phinn08}\label{Phinn08}
 demonstrates the high suitability of CASI images for the fine-scale mapping of seagrass
landscapes. Thus, CASI and Quickbird-2 images enable to identify even separate seagrass species
with small width and heterogeneous nature of the seagrass patches, which could not be detected using
Landsat TM images with their 30*30m resolution.
Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}), launched in
1999 onboard Terra sensor, provides high-resolution images of the Earth in 15 different bands of the
electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:13}). It has diverse
subsystems for the visible near infrared (VNIR) with 15 m resolution, shortwave infrared (SWIR), and
thermal infrared (TIR) wavelength regions \cite{Kalinowski04}\label{Kalinowski04}. For each channel there is separate
onboard calibration (OBC) system, telescope with independent pointing and different detector technsology \cite{Arai05}\label{Arai05}. 
This makes ASTER imagery expecially suitable for the application towards detailed
mapping of surface temperature, emissivity and reflectance of objects and bathymetric elevations as
well. Successful application of the ASTER imagery towards mapping of submerged vegetation
reported, for example, by Hirose \cite{Hirose04} \label{Hirose04}. Other multispectra images have also been used for the
seagrass interpretation. For example, the success of using imagery from the multispectral airborne
scanner Daedalus AADS1268 is reported by Heege et al \cite{Heege03}\label{Heege03} where they aim at classification of
macrophytes in shallow waters of the Lake Constance.
In regard to the methods chosen for the image interpretation, the supervised classifications proves to
be the most worthy in a majority of case studies \cite{Palandro03,Peneva08}\label{Palandro03} \label{Peneva08}. While
often methods of the unsupervised classification are used as a tool for classifying submerged object
and features on the multispectral images and aerial imagery \cite{Fletcher09}\label{Fletcher09}, it still does not
provide common algorithms that can be applied to other images and regions. Thus, the strength of the
optical signal coming from the various types of the seafloor is strongly influenced by the effects of the
water column, its depth and chemical properties. Yet methods of the unsupervised classification
overleap mixed spectral effects of the water column which shift the real values of the spectra, with the
pure reflectance from the benthos; as a result, it may cause significant errors \cite{Dierssen03}\label{Dierssen03}.
 Therefore, for the accurate assessments of various seafloor cover types, water
column depth and its optical properties, methods of supervised classification should be preferably
used as being more suitable in classification and interpretations of imagery.
While this and other studies \cite{Phinn08}\label{Phinn08}demonstrated the advantages and success of the
application of multispectral imagery for the spectral discrimination of the seafloor cover types and
mapping the submerged landscapes on the basis of pixels classification, the application of data from
hyperspectral sensors has better potential due to their higher resolution.
\pagebreak

\subsubsection{Review of hyperspectral imagery used for seagrass mapping}
The application of the hyperspectral sensors is most effective and provide more accurate classification
results (Fig.2.1.),

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[scale=0.30]{Fig-14.jpg}
\caption{Multispectral vs. hyperspectral band
coverage. Source: \cite{Sabins97}\label{Sabins97}}
\label{fig:14}
\end{wrapfigure}

due to their higher spectral resolution \cite{Bharathi03}\label{Bharathi03} with interval
narrows to 10 nanometers, while broadband sensors are limited to the spectral width of ca
150 nm (Fig.\ref{fig:14}). 
Hyperspectral imagery is acquired through the simultaneous acquisition of images in many
narrow, contiguous spectral bands from hyperspectral scanners (mostly) cover the 400-
to 2500-nm spectral bands \cite{Schmidt03}\label{Schmidt03}.
Perhaps, the most advantageous and general characteristics of hyperspectral imagery is its
high spectral resolution, desirable for the case of seagrass monitoring.

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[scale=0.15]{Fig-15.jpg}
\caption{Reflectance spectra of seagrass; Rrs(0-)
- subsurface RS reflectance; Rrs b - the bottom
reflectance. Source:\cite{Yang10}\label{Yang10}}
\label{fig:15}
\end{wrapfigure}

The most suitable
scanner fit for detailed seagrass mapping would cover bands of 550-750 nm and have a spectrum
resolution of 5-15 nm \cite{Fyfe04}\label{Fyfe04}, which exactly characterizes typical airborne and
hyperspectral satellite scanners. Important features may be detected in the narrow
wavelengths of hyperspectral imagery, while this information can be lost in the broader
wavelengths of other sensors. With its 126 spectral bands, HyMap imagery enables to
distinguish features of interest, i.e. seagrass types \cite{Peneva07}\label{Peneva07}, which is the major
advantage of the of hyperspectral data for mapping landscapes of the seagrasses.
While comparing multispectral imagery with airborne hyperspectral, the last one showed higher
overall accuracies \cite{Phinn08}\label{Phinn08}. Several, regular and narrow (10 nm) spectral bands, specific for
hyperspectral imagery, are strong tool enabling to detect even slight and subtle differences in the
spectral reflectances between various seafloor types, e.g. different seagrass species at diverse depths,
algae, corals, dark-coloured sands or other types of sediments \cite{Hochberg03a}\label{Hochberg03a}. 
Therefore, there is great potential of the application of hyperspectral remote sensing imagery towards seagrass
mapping at species level, as long as they are distinguishable spectrally (Fig.\ref{fig:15}) which, for example, has been
tested in the case study of Australian marine ecosystems by Fyfe \cite{Fyfe04}\label{Fyfe04}. Application of various
classifications methods, inter alia maximum likelihood, minimum distance and means, towards
hyperspectral imagery \cite{Peneva08}\label{Peneva08}, combined with fieldwork measurements ensures accurate
mapping results with the maximum likelihood methos producing the best results.
Therefore, accurate mapping of the seagrass landscapes and other seafloor types using remote sensing
approaches requires application of high-spatial resolution (higher than 5 m) or hyperspectral imagery.
Comparative analysis of the application of hyperspectral and multispectral imagery towards the
seafloor types classification \cite{Hochberg03a}\label{Hochberg03a} demonstrates that coral, seagrasses and
sand wery vell distinguishable in their spectra with an overall classification accuracy of 98 percent.
However, the use of data from various sensors, both hyperspectral and multispectral, are possible and
reasonable, as soon as it meets the research specific objective. 
Thus, the use of multispectral imagery
with high spatial resolution is preferable to using hyperspectral medium resolution data in case of
mapping benthic vegetation in areas where the spatial heterogeneity is very high \cite{Vahtmae07}\label{Vahtmae07}.

\section{Materials and methods}
%\markright{III. Materials and methods}

The research is technically based on WASI, for the analysis of spectral signatures, \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} and \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software for the processing and analysing of aerial and broadband images, Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat}.
\href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0} software is used for the spatial analysis, general mapping and cartographic layout
presentation; the raster processing techniques are applied for the detection of seagrass spatial
distribution using supervised classification.
The research data include Google Earth aerial images and scenes from the \href{http://landsat.gsfc.nasa.gov/}{Landsat TM and ETM+}
covering research period of 10 years in the same year time, taken from \href{http://www.usgs.gov/pubprod/}{USGS}, \href{http://glovis.usgs.gov/}{GloVis}, \href{http://www.osdpd.noaa.gov/ml/index.html}{NOAA} and \href{http://glcfapp.glcf.umd.edu:8080/esdi/index.jsp}{The Earth Science Data Interface}. 
The satellite imagery provides
vital information of the most recent changes in \textit{P.oceanica} within the coastal areas, as well as the
condition (poor or destroyed).
The raster processing includes making mosaic-like covering for the whole research area. The Google
Earth images are most appropriate for the detailed mapping than stellite than \href{http://landsat.gsfc.nasa.gov/}{Landsat} or \href{http://asterweb.jpl.nasa.gov/}{ASTER} enabling to produce accurate maps with correct results. Therefore, the main set of images for the
current work is Google Earth aerial images. The  \href{http://landsat.gsfc.nasa.gov/}{Landsat} images are used for the general overview.
Besides aerial and satellite photographs, data acquired during the fieldwork (21 Sep – 11 Oct) are
necessary addition to the mapping helping to solve problems of interpretation during the images
classification \cite{Pasqualini98a}\label{Pasqualini98a}. Therefore, this work includes sampling of the in-situ
measurements of the seagrass distribution.
The sampling stations were located in two candidate places on the northern (Ligaria) and southern
(Xerocampus) parts of Crete island, as these regions are well suitable for the seagrass, due to the
annual mean water temperatures and geological factors, i.e. seafloor conditions and sediments. The
field campaign has been carried out during the September-October period 2010. 
The information about the location of the seagrass (mostly represented by the P.oceanica species) is useful for the
understanding of the relationship between the spatial distribution of the seagrass and the environment
of the selected areas of Cretan beaches. The results of the videographic measurements are used for the
seafloor types detection, because the objects represented on the photos can be well distinguished and
classified according to the following well-known characteristics \cite{Butler87}\label{Butler87}: size (yet some
measurements are necessary for the similar-looking objects); shape (the general form is most reliable
evidence for identification); colour (common and reliable object indicator); texture (when changes in
tone are too small to be distinguishable, texture may assist identification, e.g., stippled, granular,
rough, smooth, etc.); associated features (those usually found near other objects, e.g., rocks and soil).

\pagebreak

\subsection{Study area}
General area: Island of Crete, Greece (Fig.\ref{fig:16}).\\
The study area of the current MSc research is located in the shallow areas of the Ligaria beach, Agia
Pelagia and Xerocampos, Crete island, Greece (Fig.\ref{fig:17}). These shelf areas have maximal depth of four meters.
Seagrass sampling has been performed at two stations at a depth of 4 meters, in the following selected
areas:
1. Ligaria beach (Agia Pelagia district), 36°20′N 22°59′E
2. Xerokampos, 35°12′N 26°18′E
3. Agia Pelagia, 36°20′N 22°59′E

\begin{figure}[h]
\centering
\includegraphics[scale=0.10]{Fig-16.jpg}
\caption{Study area: Crete island}
\label{fig:16}
\end{figure}

\subsection{Fieldwork data collection}
Seagrass sampling has been performed at two research stations at Crete island - Ligaria beach
(36°20′N 22°59′E) and Xerokampos (35°12′N 26°18′E), at depths lesser than 4 meters.

\begin{figure}[h]
\centering
\includegraphics[scale=0.25]{Fig-17.jpg}
\caption{Study area: Ligaria beach, Crete island}
\label{fig:17}
\end{figure}

The Ligaria Beach is a narrow, sandy and pebble beach (Fig.12), located ca 15 km north-west from Heraklion.
The in-situ measurements were conducted during the fieldwork in the period 21.09.2010-11.10.2010.
The fieldwork on seagrass monitoring included visual estimations and photo- and video footage of the
above-ground seagrass patches, sediment seafloor cover types, species compositions, water depth and
geograpic locations recorded using GPS.

\subsubsection{Fieldwork equipment}
The research materials and equipment were provided by \href{http://www.nhmc.uoc.gr/index-2.htm}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:18}): 1) three iPAQs; 2)
three GPS; 3) Three underwater video cameras, Olympus ST 8000, suitable for
photographing up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
Markers and cords for the depths measurements; 5) Waterproof plasctic Otterbox for keeping the
iPAQ dry; 6) \href{http://www.padi.com/scuba/}{SCUBA Diving equipment}, taken from the Ligaria Diving Centre; 7) Boat (Fig.\ref{fig:19});

\begin{figure}[h]
	\centering
	\subfloat [iPAQ, HP]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-19a.jpg}}
	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-19b.jpg}}
	\subfloat [GPS]{\label{fig:19a}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
	\subfloat [Olympus waterproof camera]{\label{fig:19a}\includegraphics[width=0.25\textwidth]{Fig-19d.jpg}}
	\caption{Fieldwork equipment}
	\label{fig:18}
\end{figure}

A GPS and iPAQ have been used for detection of the geodetic coordinates and keeping the tracklogs
along the boat route for GIS project.

\begin{figure}[h]
\centering
\includegraphics[scale=0.70]{Fig-18.jpg}
\caption{Boat used for the fieldwork measurements on Crete island. 2010}
\label{fig:19}
\end{figure}

\newpage

\subsubsection{Sampling design}
The sampling design of the fieldwork was aimed at surveying of the spatial distribution of the
meadows of \textit{P.oceanica}, and spatial pattern of the seagrass meadows consisted from separate
patches. The fieldwork included several routes of the boat in the Ligaria beach sampling site, nine
routes in total, in the directions parallel to the coastline, ca 180-200 m long each, thus enabling the
course plot to cover the area of growing seagrass: shelf areas not deeper than four meters.
The measurements of the seafloor cover types have been made using underwater video cameras
Olympus ST 8010, mounted under the boat to capture video footage and imagery. The data records
were made along each path using these cameras. A videographic approach, tested previously \cite{Norris97}\label{Norris97},
 was applied during the fieldwork on Crete island, for collecting information on benthic cover types and
distribution of the seagrass patches from photo transects, in order to use for the calibration of mapping approaches.
Three underwater video cameras, located on the bottom of the boat, provided videometric measurements of the
seafloor during the track, and resulted in a series of consequent overlapping images of the sea bottom under the
boat path. The general locations of the sampling sites and routes were selected on the basis of the visual examination
of the seagrass beds during snorkelling and SCUBA diving, recommendations from the Greek collaborators of the Natural Nistory Museum of Crete, and available maps covering the research area. To ensure even and objective selection of sampling sites
we used transect sampling method, i.e. photographs were taken along the research path. Transect
method is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
enabling the occurrences of the seagrass meadows to be recorded and counted in a systematic and
accurate way, detect spatial distribution of the single seagrass mattes to be properly identified,
without location bias.

\begin{figure}[h]
	\centering
	\subfloat [view from below]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
	\subfloat [view \textit{en-face}]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
	\subfloat [view from above]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
	\caption{Scheme of placements of the Olimpus cameras during the measurements}
	\label{fig:20}
\end{figure}

The videographic measurements and photos, enabling to detect types of the
seafloor, were captured by means of three underwater digital video cameras Olympus ST 8000 (Fig.\ref{fig:20}), at
five minutes interval along the tracklog path (Fig.\ref{fig:21}).
research sampling design also included measurements of depths, because bathymetry is one of the
most determinating factors for the seagrass locations. Measurements of depths were performed during
the fieldwork using cord, iPAQs, and markers in order to assure that the videometric measurements
are taken at depths not more than four meters, in the shelf area.
As a result, nine transects were established of one m wide and 20 m long track, to cover seagrass beds
with videographic measurements. The GPS allowed to capture measurement locations on the iPAQ,
encapsulated in a plastic waterproof Otterbox. Camera were adjusted horizontally by a leveller and
were mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{app:20}). The data
were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state. The
locations of the route were randomly selected in the areas of the Ligaria beach, to ensure most dense
coverage of the seagrass meadows in the research area. The underwater measurements of the seagrass
coverage were carried out by taking video footage and photos of ca 0.5 m2 size each.
The results of the underwater videometric measurements include series of digital images helping to
classify seafloor cover types and seagrass meadows, according to the differences in the structure,
colour, texture and shapes of the depicted objects.

\begin{figure}[h]
\centering
\includegraphics[scale=0.40]{Fig-21.jpg}
\caption{Locations of the video measurements and GPS tracklogs, Ligaria}
\label{fig:21}
\end{figure}

Along the Ligaria beach there are several types of the seagrass landscapes, namely, segrass meadows
continuously covering vast areas (Fig.\ref{app:18}), aggregated seagrass patches, represented by separate
mattes with short irregular channels between them (Fig.\ref{app:18}) and isolated seagrass patches, or mattes,
which located separately from broad meadows (Fig.\ref{app:18}).
The results of the underwater videometric measurements of the Olympus cameras made during the
ship route include nine total tracklog routes in the selected research area, including series of
consequent images, completely covering the area under the boat path. The received data contain
information on seagrass presence within the study area, distribution of seagrass P.oceanica meadows
and nature of the seafloor cover types (rocks, sandy, mixed). Seagrass species on the Ligaria and Agia
Pelagia beach consist of P.oceanica. The types of sediment on the Ligaria vary from coarse sand,
P.oceanica patches, sand, mud, rock gravel and fine sand (Fig.\ref{app:20}).

\subsection{Data of spectral measurements}

\subsection{Google Earth aerial imagery for seagrass mapping}
Apart from the satellite imagery, the aerial photographs from the Google Earth provide
a powerful tool for seagrass mapping, because they are important, reliable, detailed
and up-to-date source of imagery.
\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[scale=0.27]{Fig-22.jpg}
\caption{Capturing aerial imagery from the Google Earth: grabbing process}
\label{fig:22}
\end{wrapfigure}

Perhaps, the most clear advantage of the Google
Earth imagery is high resolution (15 m in land areas and lower in the oceans).
Obtained from the airborne platforms, Google Earth images have general spatial
resolution of several meters (though vatying in different areas), which allows very
detailed habitat and seafloor types discrimination, comparing with images
acquired from the spaceborne satellite platforms.
The spatial coverage of the Google imagery is lesser comparing to spaceborne images, but this can be solved as
well: while in general providing smaller area coverage than satellites images, the Google Earth
images can be stitched to the composite maps of the acceptable spatial extent, using Google grabbing
process which allows multiple overlapping of single images over the flight paths, and generates
mosaics.

\subsection{Application of satellite imagery for seagrass mapping (Landsat, ASTER, MODIS)}

\pagebreak
\section{Results}
%\markright{IV. Results}

\subsection{Review of the collected data}
The remote sensing imagery were collected for the Crete island area covering the research period and
area, enabling the field data to be used for the calibration and validation. The available aerial and
satellite imagery are commonly used for the mapping of seagrass landscapes and their application is
proven by various research papers. The imagery include satellite multi-spectral imagery (Landsat-TM,
ETM+) and aerial imagery from the Google Earth.
The overview of the collected data enables to summarize their following types:
1. Spectra of P.oceanica and other seafloor types at different environmental conditions
2. Aerial imagery from the Google Earth
3. Satellite images from the open sources (Landsat)
4. Underwater videometric measurements of the Olympus cameras made during the ship route
The available satellite imagery (Fig.\ref{app:20}) and aerial images are read into the ArcGIS project.
The available broadband and hyperspectral remote sensing data are used for the mapping of the
seagrass in shelf areas no deeper than 4.0 meters, for the environmental monitoring in order to detect
the dynamics in changes of the seagrass distribution along Crete during the past 10 years using
different satellite images for 2000-2010.
The spatial resolution of Landsat ETM+ image is 30 m in the visible and near infrared bands (bands
1-5 and 7); the spatial resolution of ASTER 15 m for the visible and near-infrared bands. IKONOS
acquires data in 3 visible channels and NIR, with spatial resolution of 1-4 meters. ASTER and
IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
research includes fieldwork in-situ measurements of the seagrass distribution along the northern and
south-eastern coasts of Crete in chosen locations.
The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
the seafloor have been received by the means of the optical sensors Trios-RAMSES Hyperspectral
UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.2.3
and 2.4), both adjusted for the measurements of the irradiance and radiance.
The highest measured values are located in the diapason of 410-730 nm for the water irradiance and 430-650 for the water radiance (Fig.\ref{fig:23}).

%\begin{figure}[h]
	%\centering
	%\subfloat [Irradiance]{\label{fig:23a}\includegraphics[width=0.8\textwidth]{Fig-23a.jpg}}\\
	%\subfloat [Radiance]{\label{fig:23b}\includegraphics[width=0.8\textwidth]{G-Radiance_Bezier.jpg}}
	%\caption{Optical properties of the sea water with sediments}
	%\label{fig:23}
%\end{figure}

Afterwards, the measured values of the
radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
properties of the sea water and bottom cover types (Fig.\ref{fig:23}). The spectral range of radiance
cover diapason of 320-950 nm, and ittadiance measurements are covered in the interval of 280-500
nm, which is suitable for a characteristics of seagrass reflectance.
Different curves on the radiance and irradiance graphs (Fig.\ref{fig:23}) represent several series of
the measurements. The values of the spectral reflectance are received from the
computations of these values using Excel mathematical formulae. The graphs shown on Fig.\ref{fig:23} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and
without sediments and suspended particles, respectively.
\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{GNU-17.jpg}
\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete}
\label{fig:25}
\end{figure}

\subsection{Data pre-processing}
A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonatic sand - were
tested for the spectral variability and separability under varying conditions of dfifferent environmental
constituents (e.g. depth, water content, sun angle etc), in order to determine the
potential that it may have on the approaches for further images processing and classification.The raw
initial data of measurements of the spectral reflectance have been preprocessed and analysed. These data
have been measured using different step of the wavebands: some measurements
were made with 3 nm interval, while others – using 4 and 1 nm step.

\begin{wrapfigure}{r}{0.6\textwidth}
\centering
\includegraphics[scale=0.23]{Fig-24.jpg}
\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
\label{fig:24}
\end{wrapfigure}

Therefore, these data had to be interpolated and standardized to one
format, which is values of spectral reflectancese with one nm step. For the
interpolation we used script written on Python programming language that allowed to receive more 
detailed data by interpolating them from 3 and 4 nm step up to 2 nm (Fig.\ref{fig:24}).
The interpolated data contained spectral measurements of the seagrass P.oceanica (measured at Agia
Pelagia beach, Crete), sand (measured at Agia Pelagia beach, Crete), silt and default artificial
spectrum of constant albedo at WASI (Fig.\ref{fig:25}). We included measurements of the Thalassia seagrass
(measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectances of
different seagrass speciea under various environmental conditions.

\begin{figure}[h]
\centering
\includegraphics[scale=0.25]{Fig-28.jpg}
\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
\label{fig:25}
\end{figure}

The most suitable wavebands for the seagrass monitroring usually lay between 400 and 700 nm,
which can be concluded by the visual examination, comparison and analysis of the different spectra of
the seagrasses. Therefore, we chosen the spectra 400-750 as the most appropriate range for
further research experiment.The results of the linear interpolation (Fig.\ref{fig:25}) demonstrate values of the
sand spectral reflectances with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
The initial measured data were stored in raw-oriented format, so that re-formatting them into the
column-based layout was done using “transpose” command in Open Office. The next step included
calculation of the median, mean, quartiles and other statistical values at every data set (see Appendix).
After the preliminary analysis, the measured data were visualised using GNUplot programm, which
enables fine plotting of various data. The most acceptable method of interpolation was Bézier curve ,
as it has trend-friendly graph better showing the general behavior of the curves at different wavebands
comparing to splines (see Appendix for more results). Therefore, after several experiment with
various interpolation techniques, we have choosen Bézier curves interpolation which contain convex
hulls made on its control points, and therefore is best suitable for our case: analysis of optical
properties of seawater.
\subsection{Analysis of spectral signatures}
The distinguishing spectral signatures for various seafloor types (e.g. seagrass species, coral reefs,
various types of sand, mud, other sediments, (Fig.\ref{fig:26})) exist in well-defined and narrow (10–20 nm)
wavelength ranges.

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{GNU-18.jpg}
\caption{Multiplot graph showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete. Gnuplot. \\Two complimentray graphs below show the results of the statistical analysis}
\label{fig:26}
\end{figure}

The values of their spectral reflectance are accepted as constant. The results of
spectral measurements enable to analyse, whether P.Oceanica is spectrally distinct from other sea
floor types with changing environmental conditions, using the differences in their spectral signatures
on the graphs in a WASI, the Water Colour Simulator software. \\The Water Colour Simulator WASI, a
software tool for analysing and simulating the most common types of spectra \cite{Gege05}\label{Gege05}, is highly
suitable for the seagrass spectral analysis.
There are several environmental characteristics, included in WASI interface (Fig.\ref{fig:29}), which influence the
results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
in water column, water temperature, sun angle, concentration of gelbstoff (coloured dissolved organic
matter), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small
particle \cite{Gege04}\label{Gege04}. The backward-scattering coefficient (bb), also included in WASI, is a
fundamental optical property which plays a central role in the ocean-color remote sensing, providing
the remotely sensed optical signal, as well as suspended particle distributions, water clarity, and
underwater visibility \cite{Maffione97}\label{Maffione97}. WASI enables simulation of backscattering of pure
water, large and small particles. The values of all these parameters can be redacted and changed
manuall. However, the most important, major factors affecting the in-situ conditions are water depth
and cover fraction of the seafloor types: P.oceanica and carbonatic sand (Fig.\ref{fig:27}).

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{GNU-21.jpg}
\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete. Gnuplot. \\Two complimentray graphs below show the results of the statistical analysis}
\label{fig:26}
\end{figure}

Seagrass \textit{P.oceanica} can be mapped using remotely acquired spectral imagery, if it has distinctive
reflectance signatures at different depths. Therefore, the depths of the shelf area are the first variable
condition for this research question. The depths values chosen for the current research are lesser than
3.5 meters, covering shelf zone, and providing the best enironmental conditions for the seagrass
\textit{P.oceanica}: 0.5, 2.0 and 3.5 meters with an interval of 1.5 meters.

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{GNU-20.jpg}
\caption{Statitical comparison of spectral reflectances of \textit{P.oceanica} and carbonate sand. Selected measurements sets}
\label{fig:26}
\end{figure}

The 3.5 m depth limit was chosen based on the analysis of the separability of seagrass reflectance signatures, received by the means of
previous in-situ measurements (year 2009) of the radiance and irradiance of water in Agia Pelagia
bay, which indicated that \textit{P.oceanica} seagrass could be only well separated within depths of 3.5 m.
A statistical analysis of WASI-simulated spectral reflectance has been used in order to answer the first
research question: whether the P.oceanica spectra is spectrally distinct at varying environmental insitu
conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
To answer this research question, different seafloor cover types are discriminated using data of the
broadband remote sensing.

\begin{table}[htbp]
\rowcolors{1}{LightSteelBlue}{Honeydew}
\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, with average values (for sets 1 - 350). Generalisation up to step 20 nm}
\begin{center}
%\rowcolors{1}{lightgray}{white}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Wavelength, nm & mean & Q1 & min & Q3 & max & median \\ \hline
\textbf{318.23354} & 0.030467 & 0.038907 & 0.029764 & 0.069762 & 0.116100 & 0.048984 \\ \hline
\textbf{338.29900} & 0.033560 & 0.043030 & 0.032847 & 0.076252 & 0.128290 & 0.054334 \\ \hline
\textbf{358.37932} & 0.039897 & 0.051320 & 0.039131 & 0.091143 & 0.154409 & 0.066354 \\ \hline
\textbf{378.47247} & 0.045786 & 0.058849 & 0.045039 & 0.105993 & 0.177150 & 0.078339 \\ \hline
\textbf{398.57640} & 0.052346 & 0.066859 & 0.051625 & 0.122328 & 0.201038 & 0.091905 \\ \hline
\textbf{418.68905} & 0.058942 & 0.074842 & 0.058240 & 0.138984 & 0.226028 & 0.106207 \\ \hline
\textbf{438.80838} & 0.070690 & 0.089761 & 0.069999 & 0.167825 & 0.267712 & 0.130580 \\ \hline
\textbf{458.93235} & 0.080107 & 0.102089 & 0.079392 & 0.191424 & 0.304933 & 0.149762 \\ \hline
\textbf{479.05891} & 0.086278 & 0.110348 & 0.085546 & 0.207940 & 0.332934 & 0.162881 \\ \hline
\textbf{499.18600} & 0.098498 & 0.126592 & 0.097689 & 0.239161 & 0.380603 & 0.188192 \\ \hline
\textbf{519.31159} & 0.110222 & 0.143976 & 0.109022 & 0.265948 & 0.421370 & 0.210237 \\ \hline
\textbf{539.43363} & 0.125694 & 0.166156 & 0.124310 & 0.302678 & 0.475842 & 0.240492 \\ \hline
\textbf{559.55006} & 0.131844 & 0.175865 & 0.130257 & 0.319385 & 0.503907 & 0.253726 \\ \hline
\textbf{579.65885} & 0.121915 & 0.166084 & 0.120186 & 0.302078 & 0.485661 & 0.237873 \\ \hline
\textbf{599.75794} & 0.077236 & 0.111113 & 0.075669 & 0.200557 & 0.343876 & 0.154016 \\ \hline
\textbf{619.84529} & 0.057038 & 0.085386 & 0.055588 & 0.154781 & 0.281737 & 0.117003 \\ \hline
\textbf{639.91885} & 0.050841 & 0.077438 & 0.049432 & 0.141810 & 0.265181 & 0.106188 \\ \hline
\textbf{659.97657} & 0.038118 & 0.059971 & 0.036997 & 0.110774 & 0.223366 & 0.082411 \\ \hline
\textbf{680.01641} & 0.032892 & 0.052265 & 0.032012 & 0.095789 & 0.198389 & 0.071686 \\ \hline
\textbf{700.03633} & 0.034756 & 0.058426 & 0.033643 & 0.101504 & 0.215413 & 0.075789 \\ \hline
\textbf{720.03426} & 0.026287 & 0.047084 & 0.025162 & 0.082940 & 0.193535 & 0.061184 \\ \hline
\textbf{740.00817} & 0.009740 & 0.022355 & 0.010145 & 0.042861 & 0.132244 & 0.029414 \\ \hline
\textbf{759.95601} & 0.008651 & 0.020234 & 0.009191 & 0.039371 & 0.128165 & 0.026558 \\ \hline
\textbf{779.87573} & 0.006019 & 0.014157 & 0.006397 & 0.028320 & 0.100314 & 0.019322 \\ \hline
\textbf{799.76528} & 0.008607 & 0.020030 & 0.009050 & 0.037323 & 0.118123 & 0.026535 \\ \hline
\textbf{819.62263} & 0.008777 & 0.020295 & 0.009288 & 0.037598 & 0.118055 & 0.026799 \\ \hline
\textbf{839.44571} & 0.002562 & 0.008010 & 0.002864 & 0.017472 & 0.076515 & 0.011432 \\ \hline
\textbf{859.23249} & 0.002653 & 0.006921 & 0.002897 & 0.016121 & 0.070377 & 0.009889 \\ \hline
\textbf{878.98091} & 0.002327 & 0.008080 & 0.002821 & 0.018625 & 0.082462 & 0.011617 \\ \hline
\textbf{898.68893} & 0.004077 & 0.013098 & 0.004709 & 0.031217 & 0.101146 & 0.019767 \\ \hline
\textbf{918.35451} & 0.006215 & 0.020162 & 0.007229 & 0.045213 & 0.199711 & 0.029950 \\ \hline
\textbf{937.97559} & 0.005783 & 0.026132 & 0.007307 & 0.065475 & 0.195072 & 0.040848 \\ \hline
\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
\end{tabular}
\end{center}
\label{fig:5}
\end{table}

The results enable to study reflectance properties of the seagrass and other
seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
remote sensing sensors (MODIS, ASTER, MERIS and SeaWiFS), (Fig.\ref{fig:28}).

\begin{figure}[h]
	\centering
	\subfloat [CZCS]{\label{fig:34a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
	\subfloat [MERIS]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
	\subfloat [MODIS]{\label{fig:34c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
	\subfloat [SeaWiFS]{\label{fig:34d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
	\caption{Simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
	\label{fig:28}
\end{figure}

In order to focus on the factors of primar importance, other and less influencing factors are excluded,
i.e. sun angle, concentration of the suspected particles in the water column, content of gelbstoff, etc.
For these factors default values of WASI are accepted.
Under normal conditions by independent water colour sampling, values of the remote-sensing
reflectance can vary by 12–24 per cent \cite{Toole00}\label{Toole00}, and these variations in the radiometric
determinations are mainly caused by the variety of the environmental factors.
Different factors influence colour and spectral reflectance of water, among which different bottom
depths, concentration of suspended particles in water column, water temperature, sun angle,
concentration of gelbstoff (coloured dissolved organic matter), concentration of phytoplankton,
aerosol scattering, exponent of backscattering by small particles, cloudness, viewing geometry and
wind speed (which is, however, not the major source of uncertainty). All these environmental
components increase the absorbtion and scattering of light which, in its turn, results in a complex
relationship between their concentrations and the radiance of water that finally influence its spectral
reflectance.
However, although there are a variety of environmental factors that contribute to the spectral
reflectance, the most important ones from the mentioned above are water depth and seafloor fraction.
It is because spectra of the seagrass \textit{P.oceanica} vary qualitatively over the depths interval of 0.5–4 m,
and secondly, the content and cover fraction of the seafloor have the most distinctive effect on the
spectral reflectance of the water.
Therefore, in current MSc research we focus on these two major factors, and study the response of the
water reflectance towards changing conditions of the water column depth and seafloor bottom cover
fraction (seagrass P.oceanica, carbonate sand and silt).
The analysis of the spectral signatures of the seagrass and sand clearly shows (Fig.\ref{fig:27}) that seagrass
has spectral reflectance much lesser than that of a carbonate sand, in general not increasing values of
10 percent reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33 percent in its
highest values.

\subsubsection{Water color spectral simulation using WASI}
The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
the seagrass \textit{P.oceanica}meadows, silt and carbonate sand differ and can be clearly discriminated while mapping. A study is
based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
types, where the spectral signatures were examined. WASI software (Fig.\ref{fig:29}) is used to simulate
spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
environmental conditions which influence its colour.

\begin{figure}[h]
\centering
\includegraphics[scale=0.30]{Fig-30.jpg}
\caption{Main interface of WASI water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-
4.0m}
\label{fig:29}
\end{figure}

The data provided with the model was determined at freshwater of the Boddensee \cite{Gege05}\label{Gege05}, 
yet the model was adjasted for the marine
environment, so that its parameters (concentration of chlorophyll, concentration of small particles,
yellow substance, etc) now perfectly simulate the Mediterranean water conditions.
The remote sensing reflectance has been compared under the conditions of different water depths and
cover fraction of the seafloor, in order to assess spectral signatures of the seagrass and carbonate sand
as major seafloor types. WASI enables to use forward or inverse calculation for the spectrum types at
a diapason of 350-800 nm with a 1nm spectral resolution. For the spectral analysis we applied
forward calculations, which is a computing and plotting of series of spectra according to specified
parameter settings, with exactly defined depths and cover fraction.
The specific parameters chosen for the simulation of the environmengal conditions where seagrass
grow. The adaptations to life in salt sea water requires various phisical and chemical parameters
which include salinity, temperature of 17-25°C, light requirements with 10-20 percent on average, ranging
from 4.4percent  minimal up to 29 percent \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as 35-
45° and reflection factor 0.0201. P.oceanica. 
The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:30}).

\begin{figure}[h]
	\centering
	\subfloat [Concentration of Gelbstoff at depths 0.5 – 8.0 m]{\label{fig:31a}\includegraphics[width=0.4\textwidth]{Fig-31a.jpg}}
	\subfloat [Bottom reflectance at depth 0.5-4.0 meters]{\label{fig:31b}\includegraphics[width=0.4\textwidth]{Fig-31b.jpg}}
	\caption{Sea water physical properties, modelled by WASI}
	\label{fig:30}
\end{figure}
\paragraph{Model parameters: depth and bottom cover fraction}
Although seagrass \textit{P.oceanica} can be found until depth limits down to 40 m depth (Hartog
1970), the most preferrable limits of its distribution in the Mediterranean Sea are shallow waters until
4 meters of depth.

\begin{table}
\rowcolors{1}{Aquamarine}{Azure}
\caption{Model-specific parameters of water WASI adjusted to simulate environment of the Mediterranean Sea along Crete}
\centering
%\rowcolors{1}{lightgray}{white}
% \scalebox{0.7}
  \begin{tabular}{|  p{2cm} | p{6cm} | p{25mm} | }
    \hline
    \textbf{Parameter, WASI} & \textbf{Name and description} & \textbf{Values} \\ \hline\hline
    C-L & Concentration of large suspended particles & 8 \\ \hline
   C(i), i=0..5 & Concentration of Phytoplankton & 0.035 -0.089 ug -l \\ \hline
    bbS & Specific backscattering for small particles & 0.005 m2g-1 \\ \hline
   T-W & Temperature of water & 17-25 °C \\ \hline
    n & Exponent of Backscattering by small particles & 0.005 m2g-1 \\ \hline
    Q &Anisotropy factor of upwelling radiation ("Q-factor") & 5.00 \\ \hline
    sigma-L & Reflection factor of sky radiance & 0.0201 \\ \hline
    b1 & Backscattering coefficient of saline waters & 0.00144 m-1 \\ \hline
    λ0 & Reference Wavelength for Gelbstoff absorption & 440 \\ \hline
    sun & Sun zenith angle & 45.0 \\ \hline
    zB & Bottom depth & 4.00 \\ \hline
   f(i), i=0...5 & Areal fraction of bottom surface type number n & 01/10/10 \\ \hline
    K0 & Coefficient of Attentuation & 1.0546 \\ \hline
   view & Viewing angle (0 = nadir) & 0\\ \hline
    C-X & Concentration of non-chlorophyllous particles (absorption at λ0) & 0 \\ \hline
   λS & Reference wavelength for scattering of small particles & 500 \\ \hline
    C-S & Concentration of small suspended particles & 0 \\ \hline
   S & Exponent of Gelbstoff absorption & 0.0140 \\ \hline
    C-Y & Concentration of Gelbstoff (absorption at λ0) & 0.400 \\ \hline
    Bn & BDRF of bottom reflectance (sand) & 0.318 sr -1\\
    \hline
  \end{tabular}
   \label{table:1}
\end{table}

The increase of depths (zB) influences weakening of light and thus directly affects
production of chlorophyll, because when light passes through the water and suspended particles, it is
being largely modifed through the absorption and scattering before it finally reaches plant canopy of
the seagrass. Therefore, the most healthy and suitable areas for the seagrass grow are located at depths
lesser than 4 meters (Fig.\ref{fig:31}).

\begin{figure}[h]
	\centering
	\subfloat [depth of 0.5 m]{\label{fig:32a}\includegraphics[width=0.4\textwidth]{Fig-32a.jpg}}
	\subfloat [depth of 2.0 m]{\label{fig:32b}\includegraphics[width=0.4\textwidth]{Fig-32b.jpg}}\\
	\subfloat [depth of 3.5 m]{\label{fig:32c}\includegraphics[width=0.4\textwidth]{Fig-32c.jpg}}
	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d.jpg}}
	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths}
	\label{fig:31}
\end{figure}

An analysis of the spectral reflectances of \textit{P.oceanica}.is done using the WASI simulations in order to
determine, which wavebands can be still used to identify \textit{P.oceanica}. The analysis of spectra shows
that the appropriate wavebands for seagrass mapping lay between 500 and 600 nm and has
also peaks at around 700 nm, ca between 680 and 710 nm (Fig.\ref{fig:31}). The highest values of the bottom
reflectance are at spectra of 500-600 nm. The most appropriate depths at which the spectral signatues
could still be discrtiminated for seagrass are lesser than 2.5 meters.
The remote sensing properties of the broad-band satellite sensors are often limited and do not allow
easy and unique discrimination of seagrass from other seafloor tyeps. Therefore, we focus on narrowband
spectra with the resolution of 1 nm.

\paragraph{Spectral discrimination of \textit{P.oceanica}}
The patches of white sandy bottoms of the seafloor, studied in the fieldwork in Ligaria beach, are
much brighter than mattes of P.Oceanica ((Fig.\ref{app:30}): bottom types in Ligaria), which can be clearly seen
at the graph comparing values of the spectral reflectance of the carbonate sand and than that of P.
oceanica (Fig.\ref{fig:33}).

The graph is received in excel spreadsheet using mean values of spectral reflectances of
sand (Fig.\ref{fig:34}) and seagrass (Fig.\ref{fig:32}), respectively, which have been calculated from measurements of radiance and
irradiance received in Agia Pelagia bay.

\begin{figure}[h]
\centering
\includegraphics[scale=0.40]{GNU-19.jpg}
\caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and measured values (dotted): multiplot of measurement sets 200-300}
\label{fig:29}
\end{figure}

\begin{figure}[h]
	\centering
	\subfloat [Sand]{\label{fig:27a}\includegraphics[width=0.4\textwidth]{Fig-27a.jpg}}
	\subfloat [\textit{P.oceanica}]{\label{fig:27b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
	\label{fig:33}
\end{figure}

\paragraph{Spectral discrimination of sand}

\subsubsection{Statistical analysis of the observational data and hypothesis testing}
The raw data of measurements of spectral reflectances of P.oceanica have been tested statistically. We
analysed 400 datasets made on two different data.
The total amount of measured data was large and included following datasets made using
hyperspectral radiometer Ramses: 350 measuremetn sets of P.oceanica reflectance for 14th Oct, 400
sets of P.oceanica reflectance for 15th Oct, 84 datasets for seawater reflectance with sediments, 105
datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonat sand.
Fig. 4.11. Spectral reflectance of water with sediments
Therefore, a statistical approach is necessary for the processing of such data. Statistical pre-processing
of large sets of serial data enables to generalyse data by using the most typical and predicted values of
spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
The statistical calculators were made by means of Open Office and included following computations
at each data set (table 1): mean, median, Q1 and Q3, standard deviation, min and max values. The
statistical pre-processing and analysis were made shown the mean values of spectral reflectances,
which were used for comparison with reflectance data of various seafloor cover types as well as
seawater with and without sediment.
The visualization of the data plotting was made by means of Gnuplot software, which enabels fine
drawing and advanced displaying of large amounts of serial data, ultimate control over graph properties, 
the simplicity of plotting and the ease of scripting. (see Fig.15 and Appendix).
The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
areas of quartiles (Q1, Q3, shaded areas on graphs), which cover most probable data distribution
(spectral reflectance) for each spectral band. In x-axis I displayed the areas of 400 – 950 nm in
spectra, where the measurements were done; the y-axis was adjusted for the better visualisation of the
P.oceanica spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
more than 0.20 nm, except for borders), I did not extend the y-scale up to 100 percent, in order to
focus on the most necessary area of spectral values (Fig. 15 and Appendices). The mean values are
highligted using bold points, as I used these values for plotting the final graph of P.oceanica spectra in
Ligaria beach, Agia Pelagia.
\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{GNU-13.jpg}
\caption{Multiplot of spectral reflectance of P.oceanica. Series 1-100. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
mean value within the range (red bold dots). Gnuplot visualization­}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{GNU-16.jpg}
\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
mean value within the range (red bold dots). Gnuplot visualization­}
\label{app:1}
\end{center}
\end{figure}

A statistical hypothesis test has been applied for the making decision and controlling the wealth of the
observational data of of the hyperspectral measurements of the water reflectances.
The received results are statistically significant if they are unlikely to have occurred by chance alone,
according to a pre-determined threshold probability, the significance level. Therefore, we applied
critical tests of significance to analyse the measured data according to their significant value.
Answering the first research question, we suggest the following statement.
Assuming that the null hypothesis, Hypothesis Ho, is true, which is “the spectral distinguishability of
the seagrass P.oceanica from other seafloor types is not changing with varying in-situ conditions,
Ho: μ1 =μ2 =μ3=...= μn.”, what is the probability of observing the values of spectral reflectance of
P.oceanica at various depths, for the test statistic, that is at least as extreme as the values of spectral
reflectance of P.oceanica at these depths, that was actually observed? The alternative Hypothesis Ha
Fig.15. Multiplot display of spectral reflectances of P.oceanica. Measurement sets 1-100, 15th
October. Bold red dots show the mean values within each dataset; vertical areas: quartiles Q1-Q2.
claims the opposite statement: “the spectral discernibility of the seagrass P.oceanica is distinctly
changing with varying in-situ conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.

\paragraph{ANOVA-testing}
The ANOVA (ANalysis Of Variance) testing is applied for the analysis of the probability of the
spectral reflectance of P.oceanica to be more or less spectrally disticnt from other seafloor cover
types with changing depth. W are checking the statement Hypothesis Ho, is true, which is “the
spectral distinguishability of the seagrass P.oceanica from other seafloor types is not changing with
varying in-situ conditions, Ho: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
P(reject H0|H0 is valid)= P(X>c|p=)=.05
where c comes for critical values and p=probability.
The result P=.05 is, hence, very small, which makes the statement of Hypothesis Ho highly unlikely
(less than 1 in a 10 chance). The one-way ANOVA highlighted a significant difference between data
of spectral reflectances of P.oceanica at different depths, i.e. true is the opposite statement:
Hypothesis Ha=“the spectral discernibility of the seagrass P.oceanica is distinctly changing with
varying in-situ conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.

\paragraph{Student-t testing}
The Student-t test is one of the most commonly used techniques for testing a hypothesis on the basis
of a difference between sample means. In our case the Student t-test demonstrates if the variation
between two analysed groups – values of spectral reflectances of seagrass P.oceanica and sand - is
significant. Therefore, we use Student t-test to compare two sets of quantitative data of spectral
reflectances of P.oceanica and sand, respectively, with samples collected independently of one
another. The Student-t test can be performed knowing just the means, standard deviation, and number
of data points. Therefore, we used the data (Appendix 11) of means of spectral reflectances of both
bottom cover types within data sets, their standard deviation, and the number of data points.
The results show the following outcome.
As a result of the statistical testing, we came to the following conclusion.
The Hypothesis Ha is true, which claims that the spectral discernibility of the seagrass P.oceanica is
distinctly changing with varying in-situ conditions, i.e. increasing depth, Ho: μ1≠μ2≠μ3≠...≠ μn .
Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
(Box Plots, Fig..).
The graphs werer plotted on the basis of the following statistical data calculated from the sampling
measurements data: Table 2.

\begin{table}
\rowcolors{1}{Khaki}{LightYellow}
\caption{Statistical analysis of the measurements of spectral reflectance of P.oceanica (fragment).}
\centering
  \begin{tabular}{| c | c | c | c | c |}
    \hline
    \textbf{Statistical values} & \textbf{318.19} & \textbf{418.3448} & \textbf{518.8361} & \textbf{619.3662} \\ \hline \hline
    Mean & 0.034521 & 0.049235 & 0.076088 & 0.085287 \\ \hline
    St Dev & 0.019849 & 0.030338 & 0.042587 & 0.046569 \\ \hline
   Median & 0.033035 & 0.046197 & 0.07704 & 0.080277 \\ \hline
   Q1 & 0.019293 & 0.023706 & 0.045606& 0.049468 \\ \hline
   Q3 & 0.042477 & 0.057934 & 0.092289 & 0.114191 \\ \hline
   Minimum & 0.007898 & 0.010087 & 0.021373 & 0.022691 \\ \hline
   Maximum & 0.076835 & 0.110906 & 0.18938 & 0.216231 \\ \hline
   25th Pct & 0.019293 & 0.023706 & 0.045606 & 0.049468 \\ \hline
   50th Pct & 0.013743 & 0.022491 & 0.031435 & 0.030808 \\ \hline
   75th Pct & 0.009441 & -0.01174 & -0.01525 & -0.03391 \\ \hline
   Min & 0.011395 & 0.013618 & 0.024232 & 0.026777 \\ \hline
   Max & 0.034358 & 0.052972 & 0.097091 & 0.1 \\ \hline
  \end{tabular}
   \label{table:1}
\end{table}

where the abbreviations stand for the following values:
Mean: Average of the data to be plotted, AVERAGE {data}
St Dev: Standard deviation, STDEV{data}
Median: Median of the data, MEDIAN{data}
Calculating interquartile ranges: Q1=First quartile, PERCENTILE({data}*0.25) and Q3=Third
quartile, PERCENTILE({data}*0.75)
Minimum: Minimum value,MIN{data}
Maximum: Maximum value, MAX{data}
25th Pct: Plotting value of first quartile = Q1
50th Pct: Plotting value of median = Median-Q1
75th Pct: Plotting value of third quartile = Median-Q3
Min: Lower error bar lenth=Q1 - Minimum
Max: Upper error bar lenth=Maximum - Q3
The statistical analysis is displayed on the Graph 1.1 which compares the spread in sets of
measurement data of spectral reflectances of under various environmental conditions (depth).

\subsubsection{Results of the remote sensing application}
The in situ spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
types as the different sensors would percept them: MODIS, ASTER, MERIS, SeaWiFS and CZCS.
These models simulate spectral views of the chosen sensors, how these sensors will "see" seagrass (Fig.\ref{fig:28}) and sand as
pixels, with accepted default atmosphere and water column effects (given by WASI software). In such
a way we defined an empirical upper limit to the discriminative potential of these sensors. To create
these sensor-specific reflectance spectra, spectral responses of each of these sensor were applied for
the spectral signature of seafloor cover types in full-resolution spectra.

\pagebreak

\subsection{Image processing using Erdas Imagine}
The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
investigate the distribution of the seagrass P.oceanica within the research area.
In the image processing part of the research supervised classification has been applied to the aerial
and satellite images. Seagrass meadows and other seafloor cover types were evaluated through a
detailed examination of the imagery. \\
Seagrass beds are clearly visible in color aerial Google Earth
photographs (Fig.\ref{fig:35}), contrasting against slightly-yellow and browny sand bottom. The seagrass areas are
detected using different bands combinations, masked and studied for the estimation of the changes in
the areas. The classification is based on the spectral properties of the \textit{P.oceanica}, such as brightness,
colour, texture and structure of the seagrass mattes (Fig.\ref{fig:35}a). The raster-based mapping
includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
for each photograph by classification a series of polygons characteristic of each of the sea floor
bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:36}).

\begin{figure}
	\centering
	\subfloat [small-scaled (1:54,000)]{\label{fig:35a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
	\subfloat [enlarged (1:10,000)]{\label{fig:35b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
	\caption{Seafloor types of the Arina beach: \textit{P.oceanica} sand, rocks}
	\label{fig:35}
\end{figure}

On the basis of the image processing and classification, applied to the northern coats of Crete, the
marine bottom types between 10 m depth has been mapped, including the limit depths of the of the
seagrass meadows provided on the basis of the available and collected field data. The classification
has been completed using ENVI and Erdas Imagine software.

\begin{figure}
\begin{center}
\includegraphics[scale=0.40]{Fig-36.jpg}
\caption{Results of the image classification in Erdas Imagine: seagrass distribution in Bali area, Crete}
\label{fig:36}
\end{center}
\end{figure}

\subsection{Seagrass mapping using ArcGIS}

\subsubsection{Data integration}
The integrated approach used in this research work has high potential as a means to monitor changes
in seagrass landscape occurring in a shallow waters over Crete area.

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{Fig-37.jpg}
\caption{Google aerial images incorporated into the GIS project: fragment of ArcMap layout}
\label{fig:37}
\end{center}
\end{figure}

It encompasses the integration of
high resolution aerial color Google Earth photography, spaceborne satellite imgery, assessment of
spectral signatures using WASI software, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
based mapping. The use of GIS for data incorporation (Fig.\ref{fig:37}), storage, analyses, visualizing and mapping
enables to analyse environmental changes within seagrass landscapes based on data from various
sources: aerial and satellite images, geographically referenced maps of Crete island and results of
images classification showing areas of seagrass distribution.
The final mapping has been supported in ArcGIS through the data exporting, conversion and
integration of various data in one GIS-project. Data collected during the fieldwork, imagery
of the seagrass distribution are added into a GIS dataset for the assessment and spatial analysis.

\subsubsection{Upscale mapping of the seagrass landscapes}
Although ocean pelagic landscapes have a high degree of spatial variance and less structurally
complex, comparing to the terrestrial ones, the landscape-level phenomena have similar features, and
there are accepted definitions of landscapes elements within the seagrass meadows \cite{Robbins94}\label{Robbins94}. 
In general, the structure of seagrass landscape is more simplier than that of the terrestrial
ecosystems in biodiversity and complexity, however, seagrass landscapes show variation in spatial
patterns over different special scales. Complexity of the landscape of seagrass meadow show the
measure of patches in size and shape, expressed in ratio of patch perimeter to area. The fragmentation
of the seagrass landscapes is expressed in contiguity displaying patch aggregation within meadows.\\
The general principles of the hierarchy within the seagrass landscapes are based upon the quantitative
analysis of the spatial patterns, consisted by components and separate elements. Thus, the bunches of
individual shoots construct patches, the first hierarchical level. Patches are arranged into discrete
clumps of mattes (at a scale of centimeters to meters) which, in turn, make up beds with l-100m in
diameter. Finally, seagrass beds, are arranged into meadows that may extend over kilometer-wide
areas, historically defined as landscapes \cite{Robbins94}\label{Robbins94}.

\begin{figure}[h]
	\centering
	\subfloat [Meadows of the seagrass. Source: Googel Earth]{\label{fig:38a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
	\subfloat [Patches (mattes) of the seagrass. Source: in-situ videometric measurements]{\label{fig:38b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
	\caption{Variations in spatial structure of the seagrass landscapes}
	\label{fig:38}
\end{figure}

Besides spatial structure of the seagrass meadows, there are strong and complex patterns of depth zonation, specific fo individual
seagrass species, but such detailed clasiification goes beyond the scope of the current MSc research.
Unique feature of seagrass meadows, most characteristic for shallow areas, is dynamics of their
landscapes. Homogeneous, continuous seagrass meadows can reach in size up to 40 m2 (Fig.\ref{fig:38} a),
however they are often interrupted by gaps and channels of open spaces generated by the complex,
disturbin sedimentation processes, turbulence and turbidity regime of waters and landscape dynamics,
which form gaps and channels within meadows, making them “patchy-looking”, i.e. diversified by
separate mattes (Fig.\ref{fig:38} b). The traditional definition for gap in vegetation cover is “disturbancegenerated
openings in either floral or faunal cover” \cite{Connell78}\label{Connell78}.

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{Fig-39.jpg}
\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the ArcGIS project}
\label{fig:39}
\end{center}
\end{figure}

Formation and increase of these
gaps within the mosaic of seagrass meadows is caused by different reasons. The most probable
drivers for the process of gaps formation within seagrass meadows are removal of interior vegetation,
differential growth of seagrass meadows and increased sedimentation. Thus, storms lead to severe
deposition of sediments, burying parts of seagrass meadow, the same effect has movements and
deposition of sediments during and after floods\cite{Bell99}\label{Bell99}.
Increased nutrient sedimentation, expecially phosphorus, were explored \cite{Jensen01}\label{Jensen01} as a
potential mechanism for increasing patch dynamics and morphological plastisity within seagrass
meadows. Finally, increasing the degree of fragmentation of the landscapes of P.oceanica meadows
can be caused by the invasion of alien species, such as Cymodocea nodosa, Caulerpa prolifera,
Caulerpa taxifolia. Invaders are in general strong colonizers comparing to native P.oceanica: they
occupy much greater habitat space within the regressed meadows of stressed native seagrass
\cite{Montefalcone10}\label{Montefalcone10}. However, on the northern coasts of Crete the only dominating seagrass
species is P.oceanica.
Morphological differences in scale of seagrass landscape formations, discussed above, cause need for
the different-scale mapping. Therefore, the investigation of the seagrass meadows at different levels is
performed using underwater videometric measurements, aerial and satellite imagery.

\pagebreak

\subsubsection{Results of GIS mapping}
This research part results in series of thematic maps showing seagrass spatial distribution over the Island of
Crete. The research is based on the aerial images classification and analysis using ArcGIS and Envi
software (Fig.24). The map of the seagrass locations within the research area of Ligaria beach illustrates the seagrass
distribution at a more detailed scale.

\pagebreak
\section{Discussion and conclusions}
%\markright{V. Discussion and conclusions}
The suggested methodology of the analysis of seagrass adaptation towards different environments of
the coastal zones can be applied to other marine coastal areas with growing seagrass.\\
More than 50 percent of the world population lives within one km of the coast which results in continued
anthropogenic pressures in coastal regions. \\Therefore, management of coastal resources and shelf
zone protection in the Mediterranean Sea become increasingly important and require large-scale
monitoring and mapping of the shelf zone as vital instrument for the environmental assessment.
\pagebreak

\section{Keywords}
\markright{Keywords}
Seagrass, \textit{\textit{P.oceanica}}, analysis of spectral reflectance, remote sensing, monitoring, Google Earth, Crete
\pagebreak

\begin{appendices}
\pagenumbering{roman}
\appendix
\appendixpage
\addappheadtotoc
\appendixname{\textbf{1:\\Results of the measurements of seafloor spectral reflectance}}
%\setcounter{figure}{0}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.20]{App-1.jpg}
\caption{Values of spectral reflectances of seagrass P.oceanica (fragment)}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.20]{GNU-401-420.jpg}
\caption{Spectral reflectance of P.oceanica. Measurement series 401-420}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.20]{GNU-401-420-candles.jpg}
\caption{Statistical analysis of the measurement data: spectral reflectance of the P.oceanica. 
Visualisation of the interquartile ranging. Example of data set 401-420. Shown midspread of  statistical quartiles Q1 and Q3, min and max values within the range.}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.20]{GNU-enlarged-M-401-420.jpg}
\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
Visualisation of the interquartile ranging and plotted together with measurement data.}
\label{app:1}
\end{center}
\end{figure}
\pagebreak
%%%

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textbf{P.oceanica}. SPSS}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline\hline
\textbf{R} & \textbf{R square} & \textbf{Adjusted R square} & \textbf{Std. Error of the Estimate} \\ \hline\hline
.277 1 & .077 & .068126692 & .462 \\ \hline
\end{tabular}
\end{center}
the independent variable is wavelength. 
\label{fig:5}
\end{table}

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textbf{P.oceanica}. SPSS}
\begin{center}
\begin{tabular}{|c|c|c|c| c|c|}
\hline\hline
& \textbf{Sum of squares} & \textbf{df} & \textbf{Mean Square} & F & Sig.\\ \hline\hline
Regression & 1.810 & 1 & 1.810 & 8.477 & .004\\ \hline
Residual & 21.781 & 102 & .214 & & \\ \hline
Total & 23.592 & 103 & & & \\ \hline
\end{tabular}
\end{center}
the independent variable is wavelength. 
\label{fig:5}
\end{table}

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. SPSS}
\begin{center}
\begin{tabular}{| c | c | p{2cm}| c | c | c |}
\hline\hline
& \multicolumn{2}{|c|}{Unstandardized Coefficients}\\
\cline{2-3}
& B & Std.Error & Stand. Coef. Beta & t & Sig.\\ \hline\hline
wavelength & -.001 & .000 & -.277 & -2.912 & .004 \\ \hline
constant & .464 & .121 & & 3.821 & .000 \\ \hline
\end{tabular}
\end{center}
a. The underlying process assumed is independence (white noise).\\
b. Based on the asymptotic chi-square approximation.
\label{fig:5}
\end{table}
\pagebreak

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.20]{GNU-Stat-M-401-420.jpg}
\caption{Fragment of the statistical analysis of the P.oceanica reflectance. Example of data set 401-420. 
Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means.}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{GNU-Radiance_Bezier.jpg}
\caption{Radiance of the seawater with sediments. Measured in 
aquarium tank. Bezier Interpolation}
\label{fig:23}
\end{figure}

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline\hline
\textbf{Source of Variation} & \textbf{SS} & \textbf{df} & \textbf{MS} & \textbf{F} & \textbf{P value} & \textbf{F crit} \\ \hline
Between Groups & 373841.7048 & 2 & 186920.8524 & 407.85359 & 1.11677 & 3.01153 \\ \hline
Within Groups & 261233.1668 & 570 & 458.3038014 & & & \\ \hline
Total &  635074.8716 & 572 & & & & \\ \hline
\end{tabular}
\end{center}
P>0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
\label{fig:5}
\end{table}

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters}
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline\hline
\textbf{Groups} & \textbf{Count} & \textbf{Sum} & \textbf{Average} & \textbf{Variance} \\ \hline\hline
Column 1 & 191 &	760.126692 & 3.9797209 & 18.6948046 \\ \hline
Column 2 & 191 &	4527.214218 & 23.7026922  & 251.4863978 \\ \hline
Column 3 & 191 &	12465.19972 & 65.2628257 & 1104.7302015 \\ \hline
\end{tabular}
\end{center}
P>0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
\label{fig:5}
\end{table}
\pagebreak

\rowcolors{1}{GhostWhite}{LightCyan}
\begin{longtable}{|c|c|c|c|c|c|c|}
\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October} \label{tab:6} \\
\hline
  \multicolumn{1}{|l|}{\textbf{wl}} &
   \multicolumn{1}{l|}{\textbf{mean}} & 
   \multicolumn{1}{l|}{\textbf{Q1}} & 
   \multicolumn{1}{l|}{\textbf{min}} & 
   \multicolumn{1}{l|}{\textbf{Q3}} & 
   \multicolumn{1}{l|}{\textbf{max}} & 
   \multicolumn{1}{l|}{\textbf{median}} \\ \hline
\endfirsthead

\multicolumn{7}{c}{Table 4: Spectral reflectance of \textbf{P.oceanica} \textit{(continued from previous page)}}\\
\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
\endhead

\textbf{318.23354} & 0.030467 & 0.038907 & 0.029764 & 0.069762 & 0.116100 & 0.048984 \\ \hline
\textbf{321.57666} & 0.028129 & 0.037310 & 0.027965 & 0.066572 & 0.110918 & 0.046450 \\ \hline
\textbf{324.92025} & 0.030163 & 0.038570 & 0.029502 & 0.068216 & 0.114120 & 0.047710 \\ \hline
\textbf{328.26428} & 0.030457 & 0.038936 & 0.029789 & 0.069098 & 0.116080 & 0.048574 \\ \hline
\textbf{331.60876} & 0.031407 & 0.040192 & 0.030731 & 0.071247 & 0.119681 & 0.050299 \\ \hline
\textbf{334.95367} & 0.032480 & 0.041545 & 0.031781 & 0.073494 & 0.123651 & 0.051941 \\ \hline
\textbf{338.29900} & 0.033560 & 0.043030 & 0.032847 & 0.076252 & 0.128290 & 0.054334 \\ \hline
\textbf{341.64474} & 0.034067 & 0.043938 & 0.033356 & 0.077587 & 0.131152 & 0.055506 \\ \hline
\textbf{344.99089} & 0.035346 & 0.045272 & 0.034616 & 0.080212 & 0.135395 & 0.057653 \\ \hline
\textbf{348.33743} & 0.036199 & 0.046719 & 0.035465 & 0.082716 & 0.139829 & 0.059655 \\ \hline
\textbf{351.68436} & 0.037336 & 0.048027 & 0.036587 & 0.085062 & 0.144101 & 0.061509 \\ \hline
\textbf{355.03166} & 0.038611 & 0.049598 & 0.037852 & 0.087963 & 0.149193 & 0.063835 \\ \hline
\textbf{358.37932} & 0.039897 & 0.051320 & 0.039131 & 0.091143 & 0.154409 & 0.066354 \\ \hline
\textbf{361.72735} & 0.040712 & 0.052408 & 0.039951 & 0.093252 & 0.157952 & 0.068034 \\ \hline
\textbf{365.07572} & 0.041541 & 0.053517 & 0.040782 & 0.095491 & 0.161650 & 0.069682 \\ \hline
\textbf{368.42442} & 0.042378 & 0.054625 & 0.041622 & 0.097634 & 0.164944 & 0.071524 \\ \hline
\textbf{371.77346} & 0.043682 & 0.056205 & 0.042926 & 0.100782 & 0.169458 & 0.074006 \\ \hline
\textbf{375.12281} & 0.044525 & 0.057191 & 0.043779 & 0.102916 & 0.172533 & 0.075888 \\ \hline
\textbf{378.47247} & 0.045786 & 0.058849 & 0.045039 & 0.105993 & 0.177150 & 0.078339 \\ \hline
\textbf{381.82243} & 0.047361 & 0.060698 & 0.046615 & 0.109681 & 0.182421 & 0.081219 \\ \hline
\textbf{385.17269} & 0.048457 & 0.062067 & 0.047717 & 0.112447 & 0.186097 & 0.083586 \\ \hline
\textbf{388.52322} & 0.049503 & 0.063533 & 0.048767 & 0.115361 & 0.190440 & 0.086143 \\ \hline
\textbf{391.87402} & 0.051271 & 0.065536 & 0.050529 & 0.119310 & 0.196618 & 0.089165 \\ \hline
\textbf{395.22508} & 0.052033 & 0.066436 & 0.051298 & 0.121231 & 0.199547 & 0.090884 \\ \hline
\textbf{398.57640} & 0.052346 & 0.066859 & 0.051625 & 0.122328 & 0.201038 & 0.091905 \\ \hline
\textbf{401.92796} & 0.053246 & 0.067963 & 0.052517 & 0.124750 & 0.204926 & 0.093940 \\ \hline
\textbf{405.27975} & 0.054472 & 0.069505 & 0.053749 & 0.127847 & 0.209718 & 0.096463 \\ \hline
\textbf{408.63176} & 0.055492 & 0.070692 & 0.054773 & 0.130382 & 0.213406 & 0.098620 \\ \hline
\textbf{411.98399} & 0.056331 & 0.071703 & 0.055621 & 0.132628 & 0.216692 & 0.100545 \\ \hline
\textbf{415.33642} & 0.057682 & 0.073235 & 0.056976 & 0.135758 & 0.221309 & 0.103354 \\ \hline
\textbf{418.68905} & 0.058942 & 0.074842 & 0.058240 & 0.138984 & 0.226028 & 0.106207 \\ \hline
\textbf{422.04186} & 0.060504 & 0.076815 & 0.059805 & 0.142833 & 0.231707 & 0.109487 \\ \hline
\textbf{425.39485} & 0.062547 & 0.079376 & 0.061847 & 0.147773 & 0.238892 & 0.113584 \\ \hline
\textbf{428.74801} & 0.064817 & 0.082141 & 0.064115 & 0.153137 & 0.246681 & 0.118126 \\ \hline
\textbf{432.10132} & 0.066664 & 0.084493 & 0.065967 & 0.157743 & 0.253230 & 0.122093 \\ \hline
\textbf{435.45478} & 0.068616 & 0.086926 & 0.067925 & 0.162414 & 0.259765 & 0.126092 \\ \hline
\textbf{438.80838} & 0.070690 & 0.089761 & 0.069999 & 0.167825 & 0.267712 & 0.130580 \\ \hline
\textbf{442.16211} & 0.072701 & 0.092360 & 0.072010 & 0.172808 & 0.275027 & 0.134721 \\ \hline
\textbf{445.51596} & 0.074550 & 0.094801 & 0.073855 & 0.177420 & 0.282203 & 0.138517 \\ \hline
\textbf{448.86992} & 0.076196 & 0.096995 & 0.075496 & 0.181512 & 0.288621 & 0.141853 \\ \hline
\textbf{452.22397} & 0.077581 & 0.098795 & 0.076876 & 0.184975 & 0.294309 & 0.144634 \\ \hline
\textbf{455.57812} & 0.079009 & 0.100624 & 0.078296 & 0.188516 & 0.300251 & 0.147486 \\ \hline
\textbf{458.93235} & 0.080107 & 0.102089 & 0.079392 & 0.191424 & 0.304933 & 0.149762 \\ \hline
\textbf{462.28665} & 0.081264 & 0.103609 & 0.080545 & 0.194453 & 0.309937 & 0.152238 \\ \hline
\textbf{465.64102} & 0.082366 & 0.105087 & 0.081658 & 0.197331 & 0.314862 & 0.154601 \\ \hline
\textbf{468.99543} & 0.083316 & 0.106273 & 0.082589 & 0.199757 & 0.319033 & 0.156521 \\ \hline
\textbf{472.34990} & 0.084192 & 0.107427 & 0.083464 & 0.202075 & 0.323294 & 0.158350 \\ \hline
\textbf{475.70439} & 0.085122 & 0.108714 & 0.084392 & 0.204674 & 0.327802 & 0.160310 \\ \hline
\textbf{479.05891} & 0.086278 & 0.110348 & 0.085546 & 0.207940 & 0.332934 & 0.162881 \\ \hline
\textbf{482.41344} & 0.088011 & 0.112712 & 0.087275 & 0.212533 & 0.340090 & 0.166595 \\ \hline
\textbf{485.76798} & 0.090064 & 0.115295 & 0.089321 & 0.217585 & 0.348041 & 0.170644 \\ \hline
\textbf{489.12252} & 0.091822 & 0.117678 & 0.091073 & 0.222210 & 0.355070 & 0.174440 \\ \hline
\textbf{492.47704} & 0.093593 & 0.120058 & 0.092835 & 0.226915 & 0.361882 & 0.178261 \\ \hline
\textbf{495.83154} & 0.095851 & 0.123152 & 0.095072 & 0.232757 & 0.370879 & 0.182979 \\ \hline
\textbf{499.18600} & 0.098498 & 0.126592 & 0.097689 & 0.239161 & 0.380603 & 0.188192 \\ \hline
\textbf{502.54043} & 0.100497 & 0.129243 & 0.099648 & 0.243918 & 0.387852 & 0.192036 \\ \hline
\textbf{505.89480} & 0.101834 & 0.131069 & 0.100939 & 0.246962 & 0.392456 & 0.194588 \\ \hline
\textbf{509.24911} & 0.103094 & 0.133022 & 0.102139 & 0.249966 & 0.397001 & 0.196947 \\ \hline
\textbf{512.60336} & 0.104998 & 0.136069 & 0.103967 & 0.254523 & 0.404017 & 0.200556 \\ \hline
\textbf{515.95752} & 0.107481 & 0.139807 & 0.106364 & 0.259986 & 0.412790 & 0.205238 \\ \hline
\textbf{519.31159} & 0.110222 & 0.143976 & 0.109022 & 0.265948 & 0.421370 & 0.210237 \\ \hline
\textbf{522.66557} & 0.112891 & 0.148090 & 0.111650 & 0.272217 & 0.430485 & 0.215339 \\ \hline
\textbf{526.01943} & 0.115799 & 0.152266 & 0.114534 & 0.278942 & 0.440382 & 0.220946 \\ \hline
\textbf{529.37318} & 0.118594 & 0.156156 & 0.117302 & 0.285384 & 0.450045 & 0.226407 \\ \hline
\textbf{532.72680} & 0.121111 & 0.159690 & 0.119794 & 0.291491 & 0.458964 & 0.231403 \\ \hline
\textbf{536.08029} & 0.123519 & 0.163063 & 0.122169 & 0.297338 & 0.467779 & 0.236144 \\ \hline
\textbf{539.43363} & 0.125694 & 0.166156 & 0.124310 & 0.302678 & 0.475842 & 0.240492 \\ \hline
\textbf{542.78681} & 0.127313 & 0.168524 & 0.125897 & 0.306828 & 0.482311 & 0.243819 \\ \hline
\textbf{546.13983} & 0.128393 & 0.170255 & 0.126944 & 0.309747 & 0.487098 & 0.246131 \\ \hline
\textbf{549.49267} & 0.129369 & 0.171844 & 0.127887 & 0.312256 & 0.491208 & 0.248155 \\ \hline
\textbf{552.84533} & 0.130192 & 0.173167 & 0.128678 & 0.314539 & 0.495323 & 0.249936 \\ \hline
\textbf{556.19780} & 0.131149 & 0.174662 & 0.129597 & 0.317126 & 0.499790 & 0.251984 \\ \hline
\textbf{559.55006} & 0.131844 & 0.175865 & 0.130257 & 0.319385 & 0.503907 & 0.253726 \\ \hline
\textbf{562.90211} & 0.132098 & 0.176510 & 0.130479 & 0.320731 & 0.506658 & 0.254511 \\ \hline
\textbf{566.25394} & 0.131944 & 0.176786 & 0.130273 & 0.321450 & 0.508802 & 0.254761 \\ \hline
\textbf{569.60554} & 0.131139 & 0.176171 & 0.129456 & 0.320617 & 0.508657 & 0.253650 \\ \hline
\textbf{572.95690} & 0.129337 & 0.174497 & 0.127627 & 0.317523 & 0.505482 & 0.250781 \\ \hline
\textbf{576.30800} & 0.126457 & 0.171261 & 0.124730 & 0.311678 & 0.498350 & 0.245834 \\ \hline
\textbf{579.65885} & 0.121915 & 0.166084 & 0.120186 & 0.302078 & 0.485661 & 0.237873 \\ \hline
\textbf{583.00943} & 0.116464 & 0.159695 & 0.114741 & 0.289985 & 0.469210 & 0.227889 \\ \hline
\textbf{586.35972} & 0.110124 & 0.152038 & 0.108417 & 0.275585 & 0.449536 & 0.216026 \\ \hline
\textbf{589.70973} & 0.102719 & 0.142758 & 0.101038 & 0.258779 & 0.425248 & 0.201868 \\ \hline
\textbf{593.05945} & 0.094713 & 0.132871 & 0.093063 & 0.240383 & 0.399506 & 0.186687 \\ \hline
\textbf{596.40885} & 0.086028 & 0.121987 & 0.084416 & 0.220594 & 0.372134 & 0.170438 \\ \hline
\textbf{599.75794} & 0.077236 & 0.111113 & 0.075669 & 0.200557 & 0.343876 & 0.154016 \\ \hline
\textbf{603.10670} & 0.070071 & 0.102079 & 0.068539 & 0.183962 & 0.320865 & 0.140543 \\ \hline
\textbf{606.45513} & 0.064972 & 0.095380 & 0.063465 & 0.172189 & 0.304973 & 0.131080 \\ \hline
\textbf{609.80321} & 0.061747 & 0.091191 & 0.060257 & 0.164863 & 0.295132 & 0.125235 \\ \hline
\textbf{613.15094} & 0.059707 & 0.088656 & 0.058235 & 0.160355 & 0.289194 & 0.121614 \\ \hline
\textbf{616.49830} & 0.058283 & 0.086915 & 0.056821 & 0.157314 & 0.285336 & 0.119161 \\ \hline
\textbf{619.84529} & 0.057038 & 0.085386 & 0.055588 & 0.154781 & 0.281737 & 0.117003 \\ \hline
\textbf{623.19189} & 0.055996 & 0.084129 & 0.054556 & 0.152697 & 0.278854 & 0.115243 \\ \hline
\textbf{626.53811} & 0.055200 & 0.083282 & 0.053763 & 0.151068 & 0.277016 & 0.113879 \\ \hline
\textbf{629.88392} & 0.054368 & 0.082234 & 0.052931 & 0.149414 & 0.275056 & 0.112433 \\ \hline
\textbf{633.22932} & 0.053299 & 0.080860 & 0.051869 & 0.147176 & 0.271850 & 0.110512 \\ \hline
\textbf{636.57430} & 0.052033 & 0.079147 & 0.050611 & 0.144415 & 0.268135 & 0.108222 \\ \hline
\textbf{639.91885} & 0.050841 & 0.077438 & 0.049432 & 0.141810 & 0.265181 & 0.106188 \\ \hline
\textbf{643.26296} & 0.049737 & 0.075977 & 0.048336 & 0.139466 & 0.263239 & 0.104485 \\ \hline
\textbf{646.60662} & 0.048467 & 0.074330 & 0.047076 & 0.136488 & 0.260596 & 0.102251 \\ \hline
\textbf{649.94982} & 0.046762 & 0.072058 & 0.045397 & 0.132376 & 0.255761 & 0.099096 \\ \hline
\textbf{653.29255} & 0.044482 & 0.068866 & 0.043159 & 0.126637 & 0.248247 & 0.094730 \\ \hline
\textbf{656.63481} & 0.041366 & 0.064447 & 0.040108 & 0.118961 & 0.236639 & 0.088686 \\ \hline
\textbf{659.97657} & 0.038118 & 0.059971 & 0.036997 & 0.110774 & 0.223366 & 0.082411 \\ \hline
\textbf{663.31784} & 0.035847 & 0.056353 & 0.034750 & 0.104541 & 0.212843 & 0.077616 \\ \hline
\textbf{666.65861} & 0.034385 & 0.054006 & 0.033356 & 0.100476 & 0.205602 & 0.074622 \\ \hline
\textbf{669.99886} & 0.033576 & 0.052732 & 0.032604 & 0.098106 & 0.201320 & 0.072881 \\ \hline
\textbf{673.33858} & 0.033122 & 0.052033 & 0.032194 & 0.096638 & 0.198744 & 0.072021 \\ \hline
\textbf{676.67777} & 0.032897 & 0.051894 & 0.032003 & 0.095851 & 0.197694 & 0.071600 \\ \hline
\textbf{680.01641} & 0.032892 & 0.052265 & 0.032012 & 0.095789 & 0.198389 & 0.071686 \\ \hline
\textbf{683.35451} & 0.033331 & 0.053301 & 0.032432 & 0.097053 & 0.201578 & 0.072478 \\ \hline
\textbf{686.69203} & 0.033823 & 0.054940 & 0.032870 & 0.098577 & 0.206179 & 0.073481 \\ \hline
\textbf{690.02899} & 0.034458 & 0.056468 & 0.033426 & 0.099736 & 0.209895 & 0.074120 \\ \hline
\textbf{693.36536} & 0.034436 & 0.056844 & 0.033329 & 0.100112 & 0.211708 & 0.074967 \\ \hline
\textbf{696.70114} & 0.034695 & 0.057874 & 0.033554 & 0.100594 & 0.213198 & 0.075367 \\ \hline
\textbf{700.03633} & 0.034756 & 0.058426 & 0.033643 & 0.101504 & 0.215413 & 0.075789 \\ \hline
\textbf{703.37090} & 0.034209 & 0.058339 & 0.033206 & 0.101601 & 0.215696 & 0.075467 \\ \hline
\textbf{706.70485} & 0.033056 & 0.057120 & 0.031999 & 0.100262 & 0.213853 & 0.073792 \\ \hline
\textbf{710.03817} & 0.031680 & 0.055368 & 0.030599 & 0.097286 & 0.209716 & 0.071411 \\ \hline
\textbf{713.37085} & 0.030077 & 0.053314 & 0.028958 & 0.093925 & 0.206402 & 0.068982 \\ \hline
\textbf{716.70288} & 0.028935 & 0.050808 & 0.027789 & 0.089359 & 0.202502 & 0.066018 \\ \hline
\textbf{720.03426} & 0.026287 & 0.047084 & 0.025162 & 0.082940 & 0.193535 & 0.061184 \\ \hline
\textbf{723.36497} & 0.023116 & 0.041962 & 0.022192 & 0.075244 & 0.182863 & 0.055350 \\ \hline
\textbf{726.69500} & 0.019755 & 0.037276 & 0.019223 & 0.067583 & 0.171617 & 0.049522 \\ \hline
\textbf{730.02435} & 0.016355 & 0.032668 & 0.016168 & 0.059609 & 0.159449 & 0.042677 \\ \hline
\textbf{733.35299} & 0.013247 & 0.028377 & 0.013325 & 0.052370 & 0.147592 & 0.036934 \\ \hline
\textbf{736.68094} & 0.010923 & 0.024633 & 0.011200 & 0.046727 & 0.138634 & 0.032350 \\ \hline
\textbf{740.00817} & 0.009740 & 0.022355 & 0.010145 & 0.042861 & 0.132244 & 0.029414 \\ \hline
\textbf{743.33467} & 0.008848 & 0.020756 & 0.009299 & 0.040084 & 0.127367 & 0.027450 \\ \hline
\textbf{746.66045} & 0.008789 & 0.019615 & 0.009272 & 0.038166 & 0.123288 & 0.025956 \\ \hline
\textbf{749.98548} & 0.007817 & 0.018631 & 0.008305 & 0.036577 & 0.119290 & 0.024701 \\ \hline
\textbf{753.30975} & 0.007533 & 0.017951 & 0.008013 & 0.035542 & 0.116242 & 0.023967 \\ \hline
\textbf{756.63327} & 0.007777 & 0.018797 & 0.008273 & 0.036601 & 0.119191 & 0.024621 \\ \hline
\textbf{759.95601} & 0.008651 & 0.020234 & 0.009191 & 0.039371 & 0.128165 & 0.026558 \\ \hline
\textbf{763.27797} & 0.008355 & 0.020033 & 0.008883 & 0.038779 & 0.126589 & 0.026439 \\ \hline
\textbf{766.59914} & 0.007232 & 0.017224 & 0.007708 & 0.033862 & 0.113056 & 0.022658 \\ \hline
\textbf{769.91951} & 0.005575 & 0.015235 & 0.005970 & 0.029884 & 0.103584 & 0.020196 \\ \hline
\textbf{773.23907} & 0.006016 & 0.014264 & 0.006407 & 0.028277 & 0.100504 & 0.019270 \\ \hline
\textbf{776.55781} & 0.005801 & 0.013966 & 0.006189 & 0.027876 & 0.099265 & 0.019096 \\ \hline
\textbf{779.87573} & 0.006019 & 0.014157 & 0.006397 & 0.028320 & 0.100314 & 0.019322 \\ \hline
\textbf{783.19280} & 0.006136 & 0.014643 & 0.006524 & 0.028946 & 0.101978 & 0.019978 \\ \hline
\textbf{786.50903} & 0.006559 & 0.015475 & 0.006967 & 0.030261 & 0.104960 & 0.021034 \\ \hline
\textbf{789.82440} & 0.007038 & 0.016900 & 0.007440 & 0.031880 & 0.107764 & 0.022351 \\ \hline
\textbf{793.13891} & 0.007598 & 0.017699 & 0.008021 & 0.033662 & 0.111651 & 0.023996 \\ \hline
\textbf{796.45254} & 0.008137 & 0.018912 & 0.008556 & 0.035621 & 0.114760 & 0.025299 \\ \hline
\textbf{799.76528} & 0.008607 & 0.020030 & 0.009050 & 0.037323 & 0.118123 & 0.026535 \\ \hline
\textbf{803.07713} & 0.009078 & 0.020935 & 0.009544 & 0.038749 & 0.120699 & 0.027598 \\ \hline
\textbf{806.38808} & 0.009299 & 0.021600 & 0.009772 & 0.039865 & 0.122278 & 0.028669 \\ \hline
\textbf{809.69811} & 0.009526 & 0.022082 & 0.009994 & 0.040470 & 0.124400 & 0.029071 \\ \hline
\textbf{813.00722} & 0.009575 & 0.022239 & 0.010086 & 0.040394 & 0.124681 & 0.029112 \\ \hline
\textbf{816.31539} & 0.010059 & 0.021517 & 0.010572 & 0.039929 & 0.124390 & 0.028547 \\ \hline
\textbf{819.62263} & 0.008777 & 0.020295 & 0.009288 & 0.037598 & 0.118055 & 0.026799 \\ \hline
\textbf{822.92891} & 0.007647 & 0.017661 & 0.008133 & 0.034147 & 0.112749 & 0.024239 \\ \hline
\textbf{826.23423} & 0.006897 & 0.015734 & 0.007343 & 0.030561 & 0.104462 & 0.021328 \\ \hline
\textbf{829.53858} & 0.005932 & 0.013286 & 0.006340 & 0.026347 & 0.095980 & 0.018276 \\ \hline
\textbf{832.84195} & 0.004547 & 0.010824 & 0.005261 & 0.022895 & 0.087424 & 0.015291 \\ \hline
\textbf{836.14433} & 0.003285 & 0.009222 & 0.003567 & 0.019407 & 0.080997 & 0.013090 \\ \hline
\textbf{839.44571} & 0.002562 & 0.008010 & 0.002864 & 0.017472 & 0.076515 & 0.011432 \\ \hline
\textbf{842.74608} & 0.002580 & 0.007522 & 0.002094 & 0.016318 & 0.073901 & 0.010511 \\ \hline
\textbf{846.04544} & 0.003121 & 0.007283 & 0.003442 & 0.015689 & 0.080570 & 0.010195 \\ \hline
\textbf{849.34376} & 0.002719 & 0.006935 & 0.002961 & 0.016016 & 0.076860 & 0.009926 \\ \hline
\textbf{852.64105} & 0.002742 & 0.006733 & 0.002994 & 0.015949 & 0.070643 & 0.009824 \\ \hline
\textbf{855.93730} & 0.002840 & 0.007120 & 0.003106 & 0.015919 & 0.074262 & 0.010103 \\ \hline
\textbf{859.23249} & 0.002653 & 0.006921 & 0.002897 & 0.016121 & 0.070377 & 0.009889 \\ \hline
\textbf{862.52661} & 0.002825 & 0.006846 & 0.003153 & 0.016677 & 0.078344 & 0.009943 \\ \hline
\textbf{865.81966} & 0.002508 & 0.007308 & 0.002710 & 0.016406 & 0.073265 & 0.010239 \\ \hline
\textbf{869.11162} & 0.002322 & 0.007310 & 0.002532 & 0.017118 & 0.081434 & 0.010382 \\ \hline
\textbf{872.40249} & 0.001873 & 0.007289 & 0.002542 & 0.017948 & 0.086487 & 0.010688 \\ \hline
\textbf{875.69226} & 0.001769 & 0.007640 & 0.001769 & 0.017981 & 0.074717 & 0.010912 \\ \hline
\textbf{878.98091} & 0.002327 & 0.008080 & 0.002821 & 0.018625 & 0.082462 & 0.011617 \\ \hline
\textbf{882.26844} & 0.001661 & 0.008815 & 0.002284 & 0.019452 & 0.079564 & 0.011996 \\ \hline
\textbf{885.55484} & 0.002940 & 0.008588 & 0.003400 & 0.021344 & 0.080998 & 0.012489 \\ \hline
\textbf{888.84010} & 0.002827 & 0.009126 & 0.003467 & 0.022626 & 0.089420 & 0.013547 \\ \hline
\textbf{892.12421} & 0.002523 & 0.010281 & 0.002997 & 0.023568 & 0.129197 & 0.015324 \\ \hline
\textbf{895.40716} & 0.001578 & 0.010879 & 0.001578 & 0.025648 & 0.091357 & 0.016328 \\ \hline
\textbf{898.68893} & 0.004077 & 0.013098 & 0.004709 & 0.031217 & 0.101146 & 0.019767 \\ \hline
\textbf{901.96953} & 0.004560 & 0.014295 & 0.005228 & 0.032824 & 0.141000 & 0.020699 \\ \hline
\textbf{905.24894} & 0.004167 & 0.014798 & 0.004249 & 0.034899 & 0.135645 & 0.022265 \\ \hline
\textbf{908.52716} & 0.003891 & 0.015616 & 0.002986 & 0.037403 & 0.150828 & 0.024065 \\ \hline
\textbf{911.80416} & 0.003641 & 0.018706 & 0.004306 & 0.043272 & 0.115375 & 0.027328 \\ \hline
\textbf{915.07995} & 0.003120 & 0.018093 & 0.003995 & 0.045688 & 0.135804 & 0.027234 \\ \hline
\textbf{918.35451} & 0.006215 & 0.020162 & 0.007229 & 0.045213 & 0.199711 & 0.029950 \\ \hline
\textbf{921.62783} & 0.004669 & 0.021630 & 0.005940 & 0.047785 & 0.245588 & 0.032131 \\ \hline
\textbf{924.89991} & 0.003051 & 0.019678 & 0.003957 & 0.051741 & 0.127093 & 0.032278 \\ \hline
\textbf{928.17074} & 0.002955 & 0.021413 & 0.004277 & 0.053042 & 0.266342 & 0.035277 \\ \hline
\textbf{931.44030} & 0.002679 & 0.023349 & 0.004029 & 0.057574 & 0.225003 & 0.036074 \\ \hline
\textbf{934.70859} & 0.003446 & 0.023229 & 0.004285 & 0.059098 & 0.549439 & 0.037722 \\ \hline
\textbf{937.97559} & 0.005783 & 0.026132 & 0.007307 & 0.065475 & 0.195072 & 0.040848 \\ \hline
\textbf{941.24130} & 0.005209 & 0.024892 & 0.006147 & 0.058814 & 0.281685 & 0.039602 \\ \hline
\textbf{944.50571} & 0.003354 & 0.025994 & 0.004299 & 0.064521 & 0.232462 & 0.041118 \\ \hline
\textbf{947.76881} & 0.003156 & 0.023476 & 0.003813 & 0.059536 & 0.246006 & 0.037035 \\ \hline
\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
\end{longtable}
\pagebreak

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{GNU-Irr-Splines.jpg}
\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in GNUplot­}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{GNU-Irr.jpg}
\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in GNUplot}
\label{fig:21}
\end{figure}
\pagebreak
\newpage

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline\hline
& & & \multicolumn{3}{c|}{Box-Ljung Statistic}\\
\cline{4-6}
Lag & Autocorrelation & Std error(a) & Value & df & Sig (b)\\ \hline\hline
1 & .844 & .071 & 141.324 & 1 & .000 \\ \hline
2 & .816 & .071 & 273.208 & 2 & .000 \\ \hline
3 & .799 & .071 & 400.323 & 3 & .000 \\ \hline
4 & .780 & .071 & 522.217 & 4 & .000 \\ \hline
5 & .723 & .070 & 627.372 & 5 & .000 \\ \hline
6 & .700 & .070 & 726.581 & 6 & .000 \\ \hline
7 & .745 & .070 & 839.623 & 7 & .000 \\ \hline
8 & .624 & .070 & 919.336 & 8 & .000 \\ \hline
9 & .584 & .070 & 989.540 & 9 & .000 \\ \hline
10 & .562 & .069 & 1054.920 & 10 & .000 \\ \hline
11 & .522 & .069 & 1111.681 & 11 & .000 \\ \hline
12& .464 & .069 & 1156.693 & 12 & .000 \\ \hline
13 & .475 & .069 & 1204.314 & 13 & .000 \\ \hline
14 & .463 & .069 & 1249.726 & 14 & .000 \\ \hline
15 & .401 & .069 & 1283.916 & 15 & .000 \\ \hline
16 & .389 & .068 & 1316.427 & 16 & .000 \\ \hline
\end{tabular}
\end{center}
a. The underlying process assumed is independence (white noise).\\
b. Based on the asymptotic chi-square approximation.
\label{fig:5}
\end{table}

\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{Autocorr.jpg}
\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
\label{fig:21}
\end{figure}
\pagebreak
\newpage

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline\hline
Lag & Partial Autocorrelation & Std error\\ \hline\hline
1 & .844 & .072 \\ \hline
1 &.358 & .072 \\ \hline
3 & .217 &.072 \\ \hline
4 & .125 &.072 \\ \hline
5 & -.093 & .072 \\ \hline
6 & .003 & .072 \\ \hline
7 & .310 & .072 \\ \hline
8 & -.403 & .072 \\ \hline
9 & -.117 & .072 \\ \hline
10 & .033 & .072 \\ \hline
11 & -.104 & .072 \\ \hline
12 & .023 &.072 \\ \hline
13 & .246 &.072 \\ \hline
14 & -.164 & .072 \\ \hline
15 & .084 & .072 \\ \hline
16 & .152 & .072 \\ \hline
\end{tabular}
\end{center}
\label{fig:5}
\end{table}

\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{Partial_Corr.jpg}
\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
\label{fig:21}
\end{figure}
\pagebreak
\newpage

\appendixname{\textbf{2: \\Graphs of the measurements of the spectral reflectance of P.oceanica}}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{GNU-09.jpg}
\caption{Spectral reflectance of seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle (GNUplot)­}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{GNU-10.jpg}
\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. GNUplot display­}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.25]{GNU-11.jpg}
\caption{Spectral reflectance of P.oceanica. Measurement series 25-50. GNUplot display­}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.25]{GNU-12.jpg}
\caption{RS reflectance of P.oceanica. Series 1-25. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
mean value within the range.­}
\label{app:1}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.4]{GNU-14.jpg}
\caption{Multiplot of spectral reflectance of P.oceanica. Series 126-150. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
mean value within the range (red bold dots). GNUplot visualization­}
\label{app:1}
\end{center}
\end{figure}
\pagebreak

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline\hline
& \multicolumn{2}{|c|}{Cluster} &  & \multicolumn{2}{|c|}{Cluster}\\
%\cline{2,3,5,6}
 & 1 & 2 &  & 1 & 2 \\ \hline\hline
V2 &	.353010 &	.000644 &	V16 &	.240091 &	.000374 \\ \hline
V3 &	 .271519 &	.000394 &	V17 &	.270306 &	.000435 \\ \hline
V4 &	.273594 &	.000644 &	V18 &	.196188 &	.000211 \\ \hline
V5 &	.406525 &	.000527 &	V19 &	.301486 &	.000370 \\ \hline
V6 &	.344973 &	.000509 &	V20 &	.178575 &	.000328 \\ \hline
V7 &	.395117 &	.000646 &	V21 &	.231708 &	.000440 \\ \hline
V8 &	.254642 &	.000352 &	V22 &	.186493 &	.000365 \\ \hline
V9 &	.288881 &	.000267 &	V23 &	.261761 &	.000418 \\ \hline
V10 &	.191850 &	.000255 &	V24 &	.282032 &	.000264 \\ \hline
V11 &	.249462 &	.000323 &	V25 &	.190709 &	.000375 \\ \hline
V12 &	.209622 &	.000414 &	V26 &	.177432 &	.000206 \\ \hline
V13 &	.218312 &	.000206 &	V27 &	.195104 &	.000268 \\ \hline
V14 &	.177432 &	.000268 &	V28 &	.256630 &	.000386 \\ \hline
V15 &	.184836 &	.000223 &	V29 &	.283744 &	.000436 \\ \hline
	 &	 &	 &		V30 &	.406525 &	.000646 \\ \hline
\end{tabular}
\end{center}
\label{fig:5}
\end{table}

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline\hline
& \multicolumn{2}{|c|}{Change in Cluster} \\
 Iteration & 1 & 2 \\ \hline\hline
1 &	.326 &	.206 \\ \hline
2 &	.024 & 	.010 \\ \hline
3 &	.008 &	.003 \\ \hline
4 &	.000 &	.000 \\ \hline
\end{tabular}
\end{center}
a. Convergence achieved due to no or small change in cluster centers. The maximum absolute coordinate change for any center is .000. The current iteration is 4. The minimum distance between initial centers is 1.435.
\label{fig:5}
\end{table}

\begin{table}[H]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
\begin{center}
\begin{tabular}{|c|c|}
\hline\hline
Cluster 1 & 55.000 \\ \hline
Cluster 2 & 136.000 \\ \hline
Valid &	191.000 \\ \hline
Missing & .000 \\ \hline
\end{tabular}
\end{center}
\label{fig:5}
\end{table}
\pagebreak

\begin{table}[htbp]
\rowcolors{1}{Ivory}{GhostWhite}
\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline\hline
& \multicolumn{2}{|c|}{Cluster} &  & \multicolumn{2}{|c|}{Cluster}\\
%\cline{2,3,5,6}
 & 1 & 2 &  & 1 & 2 \\ \hline\hline
V2  &	0.265141 &	0.046966 &	V16 &	0.181395 &	0.033191 \\ \hline
V3 &	0.202643 &	0.036088 &	V17 &	0.202698 &	0.036614 \\ \hline
V4 &	0.20699 &	0.037789 &	V18 &	0.149364 &	0.028217 \\ \hline
V5 &	0.298975 &	0.052158 &	V19 &	0.225452 &	0.040193 \\ \hline
V6 &	0.25915 &	0.045968 &	V20 &	0.135163 &	0.025506 \\ \hline
V7 &	0.293508 &	0.049314 &	V21 &	0.175176 &	0.031866 \\ \hline
V8 &	0.193292 &	0.036402 &	V22 &	0.142295 &	0.027148 \\ \hline
V9 &	0.215848 &	0.038738 &	V23 &	0.19514 &	0.034553 \\ \hline
V10 &	0.147798 &	0.027489 &	V24 &	0.214946 &	0.038321 \\ \hline
V11 &	0.188955 &	0.034031 &	V25 &	0.14585 &	0.027665 \\ \hline
V12 &	0.157947 &	0.029319 &	V26 &	0.134787 &	0.024912 \\ \hline
V13 &	0.165877 &	0.029961 &	V27 &	0.149217 &	0.028326 \\ \hline
V14 &	0.137149 &	0.02626 &	V28 &	0.193388 &	0.034999 \\ \hline
V15 &	0.140563 &	0.026211 &	V29 &	0.214436 &	0.038433 \\ \hline
 &	 &	 &	V30	 & 0.299183	 & 0.053451 \\ \hline
\end{tabular}
\end{center}
\label{fig:5}
\end{table}

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.4]{GNU-15.jpg}
\caption{Multiplot of spectral reflectance of P.oceanica. Series 151-200. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
mean value within the range (red bold dots). GNUplot visualization­}
\label{app:1}
\end{center}
\end{figure}
\pagebreak

\rowcolors{1}{OldLace}{FloralWhite}
\begin{longtable}{|r|r|r|r|r|r|r|}
\caption{Results of the statistical analysis of spectral reflectance of \textbf{carbonate sand}, with average values (for sets 1-3). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 14th October} \label{fig:7} \\
\hline
  \multicolumn{1}{|l|}{\textbf{wl}} &
   \multicolumn{1}{l|}{\textbf{min}} & 
   \multicolumn{1}{l|}{\textbf{Q1}} & 
   \multicolumn{1}{l|}{\textbf{mean}} & 
   \multicolumn{1}{l|}{\textbf{Q3}} & 
   \multicolumn{1}{l|}{\textbf{max}} & 
   \multicolumn{1}{c|}{\textbf{median}} \\ \hline
\endfirsthead

\multicolumn{7}{c}{Table 5: Spectral reflectance of \textbf{carbonate sand} \textit{(continued from previous page)}}\\
\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
\endhead

402 & 0.091522 & 0.108252 & 0.119264 & 0.125199 & 0.156262 & 0.120855 \\ \hline
405 & 0.094129 & 0.111507 & 0.126390 & 0.139560 & 0.174126 & 0.124623 \\ \hline
408 & 0.096608 & 0.114474 & 0.129816 & 0.143488 & 0.178780 & 0.127983 \\ \hline
412 & 0.098703 & 0.117396 & 0.133108 & 0.147164 & 0.183246 & 0.131282 \\ \hline
415 & 0.101803 & 0.120880 & 0.137375 & 0.151792 & 0.189039 & 0.135571 \\ \hline
418 & 0.105012 & 0.124746 & 0.141910 & 0.156937 & 0.195073 & 0.140136 \\ \hline
422 & 0.108640 & 0.129140 & 0.147168 & 0.162721 & 0.202021 & 0.145337 \\ \hline
425 & 0.113279 & 0.134815 & 0.153679 & 0.169999 & 0.210668 & 0.151906 \\ \hline
428 & 0.118544 & 0.141149 & 0.160985 & 0.178059 & 0.220194 & 0.159402 \\ \hline
432 & 0.122987 & 0.146844 & 0.167659 & 0.185518 & 0.228841 & 0.166154 \\ \hline
435 & 0.127911 & 0.152289 & 0.174248 & 0.192960 & 0.237370 & 0.172836 \\ \hline
438 & 0.132310 & 0.158310 & 0.181427 & 0.201169 & 0.246705 & 0.180044 \\ \hline
442 & 0.136621 & 0.163674 & 0.187881 & 0.208310 & 0.255343 & 0.186472 \\ \hline
445 & 0.140316 & 0.168266 & 0.193508 & 0.214558 & 0.262974 & 0.192106 \\ \hline
448 & 0.143322 & 0.172017 & 0.198207 & 0.219703 & 0.269568 & 0.196709 \\ \hline
452 & 0.145645 & 0.174991 & 0.201955 & 0.223770 & 0.274973 & 0.200367 \\ \hline
455 & 0.148101 & 0.178031 & 0.205747 & 0.227926 & 0.280378 & 0.204052 \\ \hline
459 & 0.149944 & 0.180646 & 0.208895 & 0.230917 & 0.284950 & 0.207175 \\ \hline
462 & 0.151978 & 0.183204 & 0.212228 & 0.234461 & 0.289839 & 0.210556 \\ \hline
465 & 0.136264 & 0.184985 & 0.214436 & 0.238727 & 0.294286 & 0.213490 \\ \hline
469 & 0.155248 & 0.187804 & 0.217729 & 0.240237 & 0.298004 & 0.215926 \\ \hline
472 & 0.156554 & 0.189708 & 0.220128 & 0.242769 & 0.301445 & 0.218291 \\ \hline
475 & 0.158239 & 0.192110 & 0.223000 & 0.246279 & 0.305677 & 0.221116 \\ \hline
479 & 0.160611 & 0.195242 & 0.226849 & 0.250810 & 0.310961 & 0.224984 \\ \hline
482 & 0.164012 & 0.200152 & 0.232478 & 0.257229 & 0.318555 & 0.230599 \\ \hline
485 & 0.168472 & 0.205823 & 0.238938 & 0.264573 & 0.327001 & 0.237010 \\ \hline
489 & 0.172449 & 0.211132 & 0.245045 & 0.271446 & 0.334981 & 0.243087 \\ \hline
492 & 0.176260 & 0.216078 & 0.250975 & 0.278269 & 0.342800 & 0.249133 \\ \hline
495 & 0.180950 & 0.222148 & 0.258178 & 0.286469 & 0.352048 & 0.256243 \\ \hline
499 & 0.186009 & 0.228774 & 0.265843 & 0.295200 & 0.361896 & 0.263884 \\ \hline
502 & 0.189596 & 0.233599 & 0.271417 & 0.301563 & 0.368882 & 0.269139 \\ \hline
505 & 0.191587 & 0.236300 & 0.274828 & 0.305379 & 0.372817 & 0.272281 \\ \hline
509 & 0.193259 & 0.238538 & 0.277796 & 0.309446 & 0.376347 & 0.275357 \\ \hline
512 & 0.196153 & 0.242620 & 0.282441 & 0.315177 & 0.381948 & 0.279890 \\ \hline
515 & 0.200015 & 0.248183 & 0.288410 & 0.322126 & 0.389945 & 0.285937 \\ \hline
519 & 0.204139 & 0.253981 & 0.294768 & 0.329524 & 0.398796 & 0.291707 \\ \hline
522 & 0.208548 & 0.259760 & 0.301397 & 0.337066 & 0.407969 & 0.298231 \\ \hline
526 & 0.213382 & 0.266241 & 0.308821 & 0.345426 & 0.418276 & 0.305397 \\ \hline
529 & 0.218205 & 0.272715 & 0.316136 & 0.353749 & 0.428329 & 0.312514 \\ \hline
532 & 0.222568 & 0.278565 & 0.322872 & 0.361518 & 0.437755 & 0.319122 \\ \hline
536 & 0.226789 & 0.284495 & 0.329470 & 0.369010 & 0.446804 & 0.325558 \\ \hline
539 & 0.230711 & 0.289790 & 0.335482 & 0.376082 & 0.455017 & 0.331317 \\ \hline
542 & 0.233458 & 0.293802 & 0.340066 & 0.381589 & 0.461345 & 0.335918 \\ \hline
546 & 0.235193 & 0.296532 & 0.343224 & 0.385560 & 0.465598 & 0.338854 \\ \hline
549 & 0.236680 & 0.299381 & 0.345918 & 0.388909 & 0.469140 & 0.341256 \\ \hline
552 & 0.237769 & 0.301722 & 0.348197 & 0.391858 & 0.472296 & 0.343234 \\ \hline
556 & 0.239247 & 0.304604 & 0.350951 & 0.395307 & 0.476036 & 0.345633 \\ \hline
559 & 0.240209 & 0.306984 & 0.353309 & 0.398254 & 0.479382 & 0.347674 \\ \hline
562 & 0.240654 & 0.308165 & 0.354596 & 0.400054 & 0.481185 & 0.348907 \\ \hline
566 & 0.240416 & 0.308874 & 0.355162 & 0.401224 & 0.481897 & 0.348929 \\ \hline
569 & 0.238748 & 0.307064 & 0.353688 & 0.400427 & 0.480263 & 0.347042 \\ \hline
572 & 0.235026 & 0.302640 & 0.349433 & 0.395980 & 0.475648 & 0.342958 \\ \hline
576 & 0.212260 & 0.294628 & 0.340050 & 0.386378 & 0.466085 & 0.334751 \\ \hline
579 & 0.219835 & 0.283193 & 0.329664 & 0.375185 & 0.450895 & 0.324215 \\ \hline
583 & 0.208955 & 0.268530 & 0.314599 & 0.359047 & 0.432720 & 0.308769 \\ \hline
586 & 0.196370 & 0.252380 & 0.296960 & 0.338330 & 0.411550 & 0.290283 \\ \hline
589 & 0.181828 & 0.233326 & 0.276049 & 0.316609 & 0.386565 & 0.267500 \\ \hline
593 & 0.166276 & 0.212916 & 0.253544 & 0.293790 & 0.359268 & 0.243094 \\ \hline
596 & 0.149323 & 0.191055 & 0.229373 & 0.267668 & 0.331353 & 0.217166 \\ \hline
599 & 0.132182 & 0.169324 & 0.205072 & 0.241088 & 0.303079 & 0.191137 \\ \hline
603 & 0.117895 & 0.151271 & 0.185181 & 0.219420 & 0.280528 & 0.170074 \\ \hline
606 & 0.107449 & 0.138360 & 0.171146 & 0.203897 & 0.264634 & 0.155317 \\ \hline
609 & 0.101008 & 0.130206 & 0.162388 & 0.194168 & 0.254754 & 0.146269 \\ \hline
613 & 0.097084 & 0.125265 & 0.157073 & 0.188174 & 0.248485 & 0.140671 \\ \hline
616 & 0.094383 & 0.121987 & 0.153592 & 0.184349 & 0.244702 & 0.137070 \\ \hline
619 & 0.092105 & 0.119159 & 0.150566 & 0.180848 & 0.241217 & 0.134371 \\ \hline
623 & 0.090017 & 0.116792 & 0.147899 & 0.177856 & 0.238265 & 0.131097 \\ \hline
626 & 0.088364 & 0.114636 & 0.145734 & 0.175448 & 0.236222 & 0.128638 \\ \hline
629 & 0.086562 & 0.112565 & 0.143418 & 0.173135 & 0.233625 & 0.126183 \\ \hline
633 & 0.084448 & 0.109785 & 0.140489 & 0.170056 & 0.230251 & 0.123296 \\ \hline
636 & 0.082208 & 0.106766 & 0.137216 & 0.166497 & 0.226502 & 0.119990 \\ \hline
639 & 0.079925 & 0.103852 & 0.134122 & 0.162951 & 0.222721 & 0.116886 \\ \hline
643 & 0.077879 & 0.101395 & 0.131468 & 0.159978 & 0.219899 & 0.114099 \\ \hline
646 & 0.075563 & 0.098621 & 0.128409 & 0.156518 & 0.216573 & 0.110812 \\ \hline
649 & 0.072328 & 0.095004 & 0.124254 & 0.151735 & 0.212143 & 0.106456 \\ \hline
653 & 0.068455 & 0.089876 & 0.118539 & 0.145085 & 0.205574 & 0.100544 \\ \hline
656 & 0.063138 & 0.083138 & 0.110767 & 0.135900 & 0.196204 & 0.092816 \\ \hline
660 & 0.057899 & 0.076562 & 0.102919 & 0.126826 & 0.186379 & 0.085030 \\ \hline
663 & 0.054237 & 0.071651 & 0.097286 & 0.120346 & 0.179417 & 0.079514 \\ \hline
666 & 0.051717 & 0.068781 & 0.093827 & 0.116423 & 0.175195 & 0.076101 \\ \hline
670 & 0.050172 & 0.066690 & 0.091590 & 0.113854 & 0.172561 & 0.073923 \\ \hline
673 & 0.048823 & 0.064991 & 0.089610 & 0.111550 & 0.170358 & 0.071985 \\ \hline
676 & 0.047230 & 0.063165 & 0.087474 & 0.109078 & 0.167774 & 0.069807 \\ \hline
680 & 0.045818 & 0.061108 & 0.085168 & 0.106421 & 0.165049 & 0.067486 \\ \hline
683 & 0.044262 & 0.058946 & 0.082783 & 0.103630 & 0.162579 & 0.065041 \\ \hline
686 & 0.041986 & 0.055955 & 0.079361 & 0.099441 & 0.158648 & 0.061599 \\ \hline
690 & 0.040081 & 0.051396 & 0.074010 & 0.092741 & 0.152034 & 0.055896 \\ \hline
693 & 0.034783 & 0.045555 & 0.067241 & 0.084820 & 0.143195 & 0.050116 \\ \hline
696 & 0.030094 & 0.039507 & 0.060070 & 0.076166 & 0.133838 & 0.043600 \\ \hline
700 & 0.025499 & 0.033281 & 0.052660 & 0.066929 & 0.124164 & 0.036937 \\ \hline
703 & 0.020752 & 0.027050 & 0.045008 & 0.057197 & 0.113843 & 0.030173 \\ \hline
706 & 0.016243 & 0.021080 & 0.037529 & 0.047681 & 0.103394 & 0.023549 \\ \hline
710 & 0.012303 & 0.015914 & 0.030791 & 0.038860 & 0.093244 & 0.017834 \\ \hline
713 & 0.007533 & 0.011641 & 0.025113 & 0.031338 & 0.084822 & 0.013233 \\ \hline
716 & 0.006371 & 0.008236 & 0.020172 & 0.024328 & 0.077202 & 0.009214 \\ \hline
720 & 0.004247 & 0.005386 & 0.015711 & 0.018057 & 0.069074 & 0.006121 \\ \hline
723 & 0.002773 & 0.003524 & 0.012159 & 0.012999 & 0.061638 & 0.003836 \\ \hline
726 & 0.001812 & 0.002194 & 0.009537 & 0.009237 & 0.055408 & 0.002543 \\ \hline
730 & 0.001248 & 0.001453 & 0.007627 & 0.006800 & 0.049434 & 0.001739 \\ \hline
733 & 0.000801 & 0.001018 & 0.006264 & 0.004722 & 0.044850 & 0.001208 \\ \hline
736 & 0.000192 & 0.000855 & 0.005279 & 0.003743 & 0.042097 & 0.000954 \\ \hline
740 & 0.000471 & 0.000718 & 0.004811 & 0.003055 & 0.039099 & 0.000816 \\ \hline
743 & 0.000506 & 0.000644 & 0.004419 & 0.002901 & 0.037332 & 0.000740 \\ \hline
746 & 0.000211 & 0.000582 & 0.004282 & 0.002898 & 0.036089 & 0.000715 \\ \hline
750 & 0.000318 & 0.000557 & 0.003937 & 0.002175 & 0.033962 & 0.000667 \\ \hline
\end{longtable}

\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{Fig-26.jpg}
\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES}
\label{fig:34}
\end{figure}
\pagebreak
\newpage

\begin{table}
\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}
\centering
\rowcolors{1}{Ivory}{LightGoldenrodYellow}
  \begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
    \hline
    \textbf{Parameters} & \textbf{15 October, set 1} & \textbf{14 October, set 4} \\ \hline \hline
     Version &1 & 1 \\ \hline
     IDData & 78B1-2009-10-15-08-56-22-280-272 & 78B1-2009-10-14-15-15-59-342-726 \\ \hline
     IDDevice & SAM-820C & SAM-8204\\ \hline
     IDDataType & SPECTRUM & SPECTRUM \\ \hline
     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
      DateTime & 2009-10-15 08:56:22 & 10/14/2009, 15:15:59 \\ \hline
     PositionLatitude & 35.2451=35°14'29.4354” & 35.24752 = 35° 14' 51.072" \\ \hline
     PositionLongitude & 25.01269=25°0'45.6834" & 25.0098 = 25°0'35.2794" \\ \hline
     Comment & P.oceanica & P.oceanica \\ \hline
     CommentSub1 & 0.5m depth & 2.5m depth \\ \hline
     CommentSub2 & diffuse & diffuse \\ \hline
     CommentSub3 & Agia Pelagia & Agia Pelagia site 2, close to rocks \\ \hline
     IDMethodType & SAM Control & SAM Control \\ \hline
    MethodName & SAM-820C & SAM-8204 \\ \hline
    Mission & No Mission & No Mission \\ \hline
    MissionSub & 1 & 1 \\ \hline
    RecordType & 0 & 0 \\ \hline
    CalFactor & 1 & 1 \\ \hline
   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
   IntegrationTime & 512 & 1024 \\ \hline
    P31 & -1 & -1 \\ \hline
   P31e & 0 & 0 \\ \hline
   PathLength & +INF & +INF \\ \hline
   RAWDynamic & 65535 & 65535 \\ \hline
   Temperature & +NAN & +NAN \\ \hline
   Unit1 & 0101 Wavelength nm &  1 1 Wavelength nm \\ \hline
   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
   Unit3 & f0-06 Error mW/(m2 nm) & 3-3 Error mW/(m2 nm Sr) \\ \hline
  Unit4 & f1-00 Status & f1-00 Status \\ \hline
  \end{tabular}
   \label{table:1}
\end{table}

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.20]{Locations_Meas.jpg}
\caption{Locations of selected measurements, visualization on Google Earth­}
\label{app:1}
\end{center}
\end{figure}

\begin{table}
\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}
\centering
\rowcolors{1}{Lavender}{GhostWhite}
  \begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
    \hline
    \textbf{Parameters} & \textbf{14 October, set 1} & \textbf{15 October, set 4} \\ \hline \hline
     Version &1 & 1 \\ \hline
     IDData & 78B1-2009-10-14-15-32-02-078-323 & 78B1-2009-10-15-10-36-02-217-136 \\ \hline
     IDDevice & SAM-820C & SAM-8204\\ \hline
     IDDataType & SPECTRUM & SPECTRUM \\ \hline
     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
      DateTime & 2009-10-14 15:32:02 & 2009-10-15 10:36:02 \\ \hline
     PositionLatitude & 35.24752= 35°14'51.072" & 35.245071=35°14'42.2556" \\ \hline
     PositionLongitude & 25.0098= 25°0'35.2794" & 25.012661=25°0'45.5796" \\ \hline
     Comment & P.oceanica & P.oceanica \\ \hline
     CommentSub1 & 3.5m depth & 1.5m depth \\ \hline
     CommentSub2 & diffuse & diffuse \\ \hline
     CommentSub3 & Agia Pelagia & Agia Pelagia site 1, close to rocks \\ \hline
     IDMethodType & SAM Control & SAM Control \\ \hline
    MethodName & SAM-820C & SAM-8204 \\ \hline
    Mission & No Mission & No Mission \\ \hline
    MissionSub & 1 & 1 \\ \hline
    RecordType & 0 & 0 \\ \hline
    CalFactor & 1 & 1 \\ \hline
   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
   IntegrationTime & 64 & 32 \\ \hline
    P31 & -1 & -1 \\ \hline
   P31e & 0 & 0 \\ \hline
   PathLength & +INF & +INF \\ \hline
   RAWDynamic & 65535 & 65535 \\ \hline
   Temperature & +NAN & +NAN \\ \hline
   Unit1 & 01-01 Wavelength nm &  01-01 Wavelength nm \\ \hline
   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
   Unit3 & f0-06 Error mW/(m2 nm) & f0-03 Error mW/(m2 nm Sr) \\ \hline
  Unit4 & f1-00 Status & f1-00 Status \\ \hline
  \end{tabular}
   \label{table:1}
\end{table}

\appendixname{\textbf{3:\\ Google Earth aerial images covering research area of Crete island}}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{App-8.jpg}
\caption{Spectral reflectance of P.oceanica. Measurement series 1-16}
\label{app:8}
\end{center}
\end{figure}

\appendixname{\textbf{4: \\Capturing aerial imagery from the Google Earth: grabbing process}}
Script command of FWTools2.4.7 enabling to reduce the size of the aerial images, from .tif to .ecw
format:
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{Script_Gdal.jpg}
\caption{Script command of FWTools}
\label{app:9}
\end{center}
\end{figure}
\pagebreak
\newpage

\appendixname{\textbf{5: \\Types of seagrass structural patterns, Ligaria beach, Crete}}
\begin{figure}[h]
	\centering
	\subfloat [Fragment of meadow]{\includegraphics[width=0.3\textwidth]{Fragment_meadow.jpg}}
	\hspace{1mm}
	\subfloat [Aggregated seagrass patch]{\includegraphics[width=0.3\textwidth]{aggreg_patch.jpg}}
	\hspace{1mm}
	\subfloat [Isolated patch of seagrass]{\includegraphics[width=0.3\textwidth]{isolated_patch.jpg}}
	\caption{Types of seagrass structural patterns, Ligaria beach, Crete}
	\label{app:30}
\end{figure}

\appendixname{\textbf{6: \\Sampling design of the depths measurements. Ligaria beach, Crete}}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.35]{Sampling_design.jpg}
\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}
\label{app:9}
\end{center}
\end{figure}
\pagebreak
\newpage

\appendixname{\textbf{7: \\Results of videographic measurements: seafloor types on Crete island}}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.45]{Fig-3-14.jpg}
\caption{Ligaria beach, Crete: seafloor types}
\label{app:9}
\end{center}
\end{figure}

\appendixname{\textbf{8: \\Various seafloor types on Crete island}}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.25]{Fig-3-15.jpg}
\caption{Various seafloor types; Ligaria beach, Crete}
\label{app:9}
\end{center}
\end{figure}
\pagebreak

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.35]{Fig-3-16.jpg}
\caption{Measurement underwater equipment}
\label{app:9}
\end{center}
\end{figure}
\pagebreak
\newpage

\appendixname{\textbf{9: \\Data pre-processing: interpolation of the results of spectral measurements}}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.30]{Interpolation.jpg}
\caption{Interpolation of the spectral measurements}
\label{app:10}
\end{center}
\end{figure}

\begin{table}
\caption{Available broadband satellite images covering the research area of Crete island}
\centering
\rowcolors{1}{Ivory}{GhostWhite}
  \begin{tabular}{| p{1cm} | p{3cm} | p{2cm} | p{5cm} |}
    \hline
    \textbf{No} & \textbf{Image cource} & \textbf{Data} & \textbf{Name} \\ \hline \hline
     1 & Landsat ETM+ & 2005/May/04 & WRS2p181r035L71181035-035-20050504-ETM-GLS2005 \\ \hline
     2 & Landsat TM & 2006/Nov/07 & WRS2p181r036L5181036-036-20061107-TM-GLS2005 \\ \hline
     3 & Landsat ETM+ & 2005/Apr/25 & WRS2p182r035L71182035-035-20050425-ETM-GLS2005 \\ \hline
     4 & Landsat ETM+ & 2000/Jul/09 & WRS2p181r036-7dx-20000709-ETM-GLS2000 \\ \hline
     5 & Landsat TM & 1987/Jun/10 & LandsatWRS2p183r035p183r035-5dx-19870610-TM-GLS1990 \\ \hline
     6 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-261Mosaic-LandsatN-35N-35-35ETM-EarthSat-MrSID-19990808-20020624 \\ \hline
     7 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-260Mosaic-LandsatN-35N-35-30ETM-EarthSat-MrSID-19990808-20020617 \\ \hline
     8 & Landsat ETM+ & 2000/Jun/30 & WRS2p182r036-7x-20000630-ETM-EarthSat \\ \hline
     9 & Landsat MSS / Earth Sat & 1975/Jul/26 & LandsatWRS1p196r35-2m-19750726-MSS-EarthSat \\ \hline
    10 & Landsat TM / Earth Sat & 1987/Jun/10 & 012-807LandsatWRS2p183r35-5t-19870610-TM-EarthSat \\ \hline
    11 & Landsat ETM+ & 2000/Jun/30 & LandsatWRS2p182r036-7dx-20000630-ETM-GLS2000 \\ \hline
    12 & Landsat ETM+ & 2005/Apr/09 & LandsatWRS2p182r036L71182036-036-20050409-ETM-GLS2005 \\ \hline
  \end{tabular}
   \label{table:1}
\end{table}

\begin{figure}
	\centering
	\subfloat [Landsat 2006-11-07-TM]{\includegraphics[width=0.3\textwidth]{Landsat1.jpg}}
	\hspace{1mm}
	\subfloat [Landsat-2005-04-09]{\includegraphics[width=0.3\textwidth]{Landsat2.jpg}}
	\hspace{1mm}
	\subfloat [Mosaic-Landsat- ETM-1999-08-08]{\includegraphics[width=0.3\textwidth]{Landsat3.jpg}}
	\hspace{1mm}
	\subfloat[Landsat 2000-07-09 ETM]{\includegraphics[width=0.3\textwidth]{Landsat4.jpg}}
	\hspace{1mm}
	\subfloat [Landsat 2000-06-30 ETM]{\includegraphics[width=0.3\textwidth]{Landsat5.jpg}}
	\hspace{1mm}
	\subfloat [Landsat 1987-06-10-TM]{\includegraphics[width=0.3\textwidth]{Landsat6.jpg}}
	\hspace{1mm}
	\subfloat [Landsat-2005-05-04-ETM]{\includegraphics[width=0.3\textwidth]{Landsat7.jpg}}
	\hspace{1mm}
	\subfloat [Landsat-1975-07-26]{\includegraphics[width=0.3\textwidth]{Landsat8.jpg}}
	\hspace{1mm}
	\subfloat [Landsat TM]{\includegraphics[width=0.3\textwidth]{Landsat9.jpg}}
	\hspace{1mm}
	\subfloat [Mosaic-Landsat-ETM-1999-08-08]{\includegraphics[width=0.3\textwidth]{Landsat10.jpg}}
	\caption{Landsat imagery, Crete island. Previews}
	\label{app:30}
\end{figure}
\end{appendices}
\pagebreak
\clearpage

\begin{thebibliography}{999}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{plainnat}
\pagenumbering{roman}
\markright{Bibliography}

\bibitem{Ackleson2003}Ackleson S. 2003. \emph{Light in shallow waters: A brief research review}. Limnol. Oceanogr.,
48(1, part 2), 2003, 323–328. American Society of Limnology and Oceanography, Inc. \pageref{Ackleson, 2003}
\bibitem{Alcoverro1997} %
Alcoverro T., Romero J., Duarte C.M., Lopez N. 1997. \emph{Spatial and temporal variations
in nutrient limitation of seagrass Posidonia oceanic growth in the NW Mediterranean}.
Marine ecology Progress Series.Vol.146: 155-161. \pageref{}
\bibitem{Amoutzopoulou-Schina05}Amoutzopoulou-Schina H and S.Haritonidis. 2005. \emph{Distribution and phenology of the
marine phanerogam \textit{Posidonia oceanica} in the Pagassitikos Gulf, Greece}. Journal of
Biological Research 4: 203 – 211, 2005 J. \pageref{Amoutzopoulou-Schina05}
\bibitem{Arai05}
Arai K. and Tonooka H. 2005. \emph{Radiometric Performance Evaluation of ASTER VNIR, SWIR and TIR} 
IEEE Transactions on Geoscience and Remote Sensing, Vol. 43, No. 12, Dec. 2005. \pageref{Arai05}
\bibitem{Ardizzone06}
Ardizzone G., Belluscio A. and Maiorano L. 2006. \emph{Long-term change in the structure of
a \textit{Posidonia oceanica} landscape and its reference for a monitoring plan}. Marine Ecology.
ISSN 0173-9565. \pageref{Ardizzone06}
\bibitem{Ball09}
Ball D., Soto-Berelov M., Young P. and Coots A. 2009. \emph{Baywide Seagrass Monitoring
Program - Historical Seagrass Mapping}. Fisheries Victoria Technical Report Series No. 70. \pageref{Ball09}
\bibitem{Barille09}
Barille L., Robin M., Harin N., Bargain A., Patrick L. 2009. \emph{Increase in seagrass
distribution at Bourgneuf Bay (France) detected by spatial remote sensing}. Aquatic Botany. \pageref{Barille09}
\bibitem{Bakran-Petricioli06} %
Bakran-Petricioli T., Antonic O., Bukovec D., Petricioli D., Janekovic I., Krizan J.,
Kusan V. and Dujmovic S. 2006. \emph{Modelling spatial distribution of the Croatian marine
benthic habitats}. Ecological Modelling 191, 96–105. \pageref{Bakran-Petricioli06}
\bibitem{Bell99}
Bell S.S., Robbins B.D. and Jensen S.L. 1999. \emph{Gap Dynamics in a Seagrass Landscape}.
Springer-Verlag. Ecosystems (1999) 2: 493–504. \pageref{Bell99}
\bibitem{Bierwirth93}
Bierwirth P. N., Lee T. J. and Burne R. V. 1993. \emph{Shalow sea-floor reflectance and water
depth derived by unmixing multispectral imagery}. Photogrammetric Engineering and Remote
Sensing, vol. 59, No. 3, March, pp. 331-338. \pageref{Bierwirth93}
\bibitem{Blake00}
Blake S., Roob R. and Patterson E. 2000. \emph{Seagrass mapping of Victoria's Minor Inlets}.
Marine and Freshwater Resources Institute Report No. 28. \pageref{Blake00}
\bibitem{Borum04}
Borum J., Duarte C.M., Krause-Jensen D. and Greve T.M. 2004. \emph{European seagrasses:
an introduction to monitoring and management}. The M and MS project. ISBN: 87-89143-21-3.
Internet version: Available at http://www.seagrasses.org \pageref{Borum04}
\bibitem{Brown02}
Brown C.J., Cooper K.M., Meadows W.J., Limpenny D.S., Rees H.L. 2002. \emph{Small scale
mapping of sea bed assemblages in the eastern English Channel using sonar and remote
sampling techniques}. Estuarine Coastal and Shelf Science, 54, 263-278. \pageref{Brown02}
\bibitem{Boudouresque89}
Boudouresque, C.F., Meinesz, A., Fresi, E. and Gravez, V., 1989. \emph{2nd International
Workshop on \textit{Posidonia oceanica} Beds}. G.I.S. Posidonie, Oceanica, 2, 322 pp. \pageref{Boudouresque89}
\bibitem{Bharathi03}
Bharathi A.M, Kumar S.C, Kalliappan S. 2003. \emph{Target Identification in multispectral images
using in-situ hyperspectral reflectance}. The Geospatial Resource Portal. Remote Sensing.
www.gisdevelopment.net \pageref{Bharathi03}
\bibitem{Buia91}
Buia M.C. and Mazzella L. 1991. \emph{Reproductive phenology of the Mediterranean
seagrasses \textit{Posidonia oceanica} (L.) Delile, Cymodocea nodosa ( Ucria ) Aschers., and
Zostera noltii Hornem}. Aquatic Botany, 40 (1991) 343-362. Elsevier Science Publishers
B.V., Amsterdam. \pageref{Buia91}
\bibitem{Butler87}
Butler 1987. \emph{Marine resource mapping: an introductory manual}. \pageref{Butler87}
\bibitem{Cappo95}
Cappo, M., Alongi, D.M., Williams, D. and Duke, N. 1995. \emph{A Review and Synthesis of
Australian Fisheries Habitat Research: Major threats, issues and gaps in knowledge of
coastal and marine fisheries habitats}. Fisheries Research and Development Corporation. \pageref{Cappo95}
\bibitem{Carter08}
Carter G.A., Criss G.A., Blossom G.A. and Biber P.D. 2008. \emph{Decadal-Scale Changes in
Coverage on the Mississippi Barrier Islands, Northern Gulf of Mexico}.
http://precedings.nature.com/documents/3669/version/1/files/npre20093669-1.pdf \pageref{Carter08}
\bibitem{Calvo03}
Calvo S., Ciraolo G. and Loggia G. la. 2003. \emph{Monitoring \textit{Posidonia oceanica} meadows
in a Mediterranean coastal lagoon (Stagnone, Italy) by means of neural network and
ISODATA classification methods}. INT. J. REMOTE SENSING, 2003, VOL. 24, NO. 13,
2703–2716 \pageref{Calvo03}
\bibitem{Cambridge09}
Cambridge M.L., Kendrick G.A. 2009. \emph{Contrasting responses of seagrass transplants
(Posidonia australis) to nitrogen, phosphorus and iron addition in an estuary and a coastal
embayment}. Journal of Experimental Marine Biology and Ecology. 371, 34–41. \pageref{Cambridge09}
\bibitem{Choo06}
Choo, C.K. 2006. \emph{Save our Seahorses}. First Edition. www.sosmalaysia.org \pageref{Choo06}
\bibitem{Clark99}
Clark, R.N., 1999. \emph{Spectroscopy of rocks and minerals and principles of spectroscopy}.
In: A.N. Rencz (ed.). Manual of Remote Sensing, Chapter 1, John Wiley and Sons, New
York. 3-58. \pageref{Choo06}
\bibitem{Connell78}
Connell J.H. 1978. \emph{Diversity in tropical rain forests and coral reefs}. Science 199:1302–
10. \pageref{Connell78}
\bibitem{Dehouck08}
Dehouck A., Martiny N., Froidefond J.-M., Sénéchal N. 2008. \emph{In-Water Reflectance
Spectra Measured On Board a Jet-Ski Across a Complex Nearshore Zone of Case-2 Waters
during the ECORS Experiment}. Manuscript published in "Ocean Optics XIX, Barga: Italie
(2008)" \pageref{Dehouck08}
\bibitem{Dekker05a}
Dekker A.G., Brando V.E., Anstee J.M. 2005. \emph{Retrospective seagrass change detection
in a shallow coastal tidal Australian lake}. Remote Sensing of Environment 97:415-433. \pageref{Dekker05a}
\bibitem{Dekker05b}
Dekker A., Brando V., Anstee J. 2005. \emph{Remote Sensing of Seagrass Ecosystems: Use of
Spaceborne and Airborne Sensors}. SVNL083-Larkum. \pageref{Dekker05b}
\bibitem{DiCarlo04}
Di Carlo G. 2004. \emph{The natural recolonisation process of the seagrass Posidonia
oceanica (L.) Delile after the introduction of the Italo-Algerian methane pipeline in the SW
Mediterranean sea}. 
\pageref{DiCarlo04}
\bibitem{Diaz-Almela06}
Diaz-Almela E., Marba N., Alvarez E., Santiago R., Holmer M., Grau A., Danovaro R.,
Argyrou M., Karakassis I, Duarte C.M. 2006. \emph{Benthic inputs as predictors of seagrass 1
(Posidonia 2 oceanica) fish farm-induced decline}. http://arxiv.org/ftp/qbio/
papers/0611/0611006.pdf \pageref{Diaz-Almela06}
\bibitem{Dierssen03}
Dierssen H and Zimmerman R.C. 2003. \emph{Ocean color remote sensing of seagrass and
bathymetry in the Bahamas Banks by high-resolution airborne imagery}. Limnol. Oceanogr.,
48(1, part 2), 2003, 444–455, by the American Society of Limnology and Oceanography. \pageref{Dierssen03}
\bibitem{Dixon05}
Dixon R, editor. 2005. \emph{Watching the Seagrass Grow – a guide for community seagrass
monitoring in NSW (2nd Ed)}. Reference manual . The Community Environment Network,
Ourimbah. \pageref{Dixon05}
\bibitem{Duarte02}
Duarte, C.M. 2002. \emph{The future of seagrass meadows}. Environmental Conservation 29
(2), 192–206. \pageref{Duarte02}
\bibitem{Dumay02}
Dumay O., Fernandez C. and Pergent G. 2002. \emph{Primary production and vegetative cycle
in \textit{Posidonia oceanica} when in competition with the green algae Caulerpa taxifolia nd
Caulerpa racemosa}. J.Mar. Biol. Ass. UK, 82, 379-387. \pageref{Dumay02}
\bibitem{Dural10}
Dural B. 2010. \emph{Phenological observations on \textit{Posidonia oceanica} (L.) Delile medows
along the coast of Akkum (Sıgacık Bay, Aegean Sea, Turkey)}. J. Black Sea/Mediterranean
Environment, Vol. 16 (1): 133-144. \pageref{Dural10}
\bibitem{Edwards00}
Edwards A.J. (ed.) 2000. \emph{Remote Sensing handbook for tropical coastal management}.
http://www.unesco.org/csi/pub/source/rs.htm \pageref{Edwards00}
\bibitem{Ferguson97}
Ferguson R.L. and Karl Korfmacher K. 1997 \emph{Remote sensing and GIS analysis of seagrass
meadows in North Carolina, USA}. Aquatic Botany 58 (1997) 241-258. \pageref{Ferguson97}
\bibitem{Fletcher09}
Fletcher R.S., Pulich Jr. W. and Hardegree B. 2009. \emph{A Semiautomated Approach for
Monitoring Landscape Changes in Texas Seagrass Beds from Aerial Photography}. Coastal
Education and Research Foundation doi: 10.2112/07-0882.1 \pageref{Fletcher09}
\bibitem{Fornes06}
Fornes A., Basterretxea G., Orfila A., Jordi A., Alvarez A. and Tintore J. 2006. \emph{Mapping
\textit{Posidonia oceanica} from IKONOS}. ISPRS Journal of Photogrammetry and Remote Sensing
60 (2006) 315–322. \pageref{Fornes06}
\bibitem{Francour06}
Francour, P., Magreau, J.F., Mannoni, P.A., Cottalorda, J.M., Gratiot, J., 2006.
\emph{Management guide for Marine Protected Areas of the Mediterranean Sea, Permanent
Ecological Moorings}. Universite de Nice-Sophia Antipolis and Parc National de Port-Cros,
Nice, 68 pp. \pageref{Francour06}
\bibitem{Francour99}
Francour P., Ganteaume A. and M. Poulain. 1999. \emph{Effects of boats anchoring in
\textit{Posidonia oceanica} seagrass beds in the Port-Cross National Park (north-western
Mediterranean Sea)}. Aquatic conservation: Marine and Freshwater Ecosystems: 9: 391-400. \pageref{Francour99}
\bibitem{Fyfe03}
Fyfe S.K. 2003. \emph{Spatial and temporal variation in spectral reflectance: Are seagrass
species spectrally distinct?} Limnology and Oceanography 48: 464–479. \pageref{Fyfe03}
\bibitem{Fyfe04}
Fyfe S.K. 2004. \emph{Hyperspectral studies of New South Wales with particular emphasis on
the detection of light stress in Eelgrass Zostera capricorni. PhD Thesis}. The University of
Wollongong. Research Online, http://ro.uow.edu.au/theses/472/ \pageref{Fyfe03}
\bibitem{Gege04}
Gege P. 2004. \emph{The water color simulator WASI: an integrating software tool for analysis
and simulation of optical in situ spectra}. Computers and Geosciences 30 (2004) 523–532. \pageref{Gege04}
\bibitem{Gege05}
Gege, P. 2005. \emph{The Water Colour Simulator WASI. User manual for version 3}. DLR
Internal Report IB 564-01/05, 83 pp. \pageref{Gege05}
\bibitem{Gobert06}
Gobert S., Cambridge, M. L. Velimirov, B. Pergent, G. Lepoint, G. Bouquegneau, J. M.
Dauby, P. Pergent-Martini, C. Walker, D. I. 2006. Biology of Posidonia. In Larkum A.W.D.,
Orth, R.J., Duarte C.M. (eds.), \emph{Seagrasses: Biology, ecology and conservation}. Springer
Verlag, Berlin, Pages 387–408. \pageref{Gobert06}
\bibitem{Gobert01}
Gobert S., Defawe O, Lepoint G, Demoulin V and Bouquegneau J-M. 2001. \emph{Anthesis
effects on \textit{Posidonia oceanica} (L.) Delile phenology in the Bay of Calvi (Corsica,
Mediterranean Sea)}. Hydrobiologia 455: 121–125. \pageref{Gobert01}
\bibitem{Green96}
Green, E.P., P.J. Mumby, A.J. Edwards, and C.D. Clark, 1996. \emph{A review of remote
sensing for the assessment and management of tropical coastal resources}. Coast. Manag.,
24(1):1-40. \pageref{Green96}
\bibitem{Green03}
Green E.P., Short F.T. 2003. \emph{World Atlas of Seagrasses}. University of California Press,
Berkeley, USA. \pageref{Green03}
\bibitem{Guidetti08}
Guidetti P. 2008. \emph{Population dynamics of the seagrass \textit{Posidonia oceanica} for
evaluating meadow health status: a study case using reconstruction techniques}. Laboratory
of Zoology and Marine Biology, DiSTeBA, CoNISMa, University of Lecce. \pageref{Guidetti08}
\bibitem{Gullstroom06}
Gullstroom M.,, Lunden B., Bodin M., Kangwe J., Ohman M.C., Mtolera M.S.P. and
Bjork M. 2006. \emph{Assessment of changes in the seagrass-dominated submerged vegetation of
tropical Chwaka Bay (Zanzibar) using satellite remote sensing}. Estuarine, Coastal and Shelf
Science 67 (2006) 399-408. \pageref{Gullstroom06}
\bibitem{Haag08}
Haag S.M., Kennish M., Sakowicz G.P. 2008. \emph{Seagrass Habitat Characterization in
Estuarine Waters of the Jacques Cousteau National Estuarine Research Reserve Using
Underwater Videographic Imaging Techniques}. Journal of Coastal Research, SI(55), 171–
179. West Palm Beach (Florida), ISSN 0749-0208. \pageref{Haag08}
\bibitem{DenHartog70}
Den Hartog, C. 1970. \emph{The Seagrasses of the World}. North Holland Publishing Company,
Amsterdam, 275 pp. 31 plates. \pageref{DenHartog70}
\bibitem{DenHartog77}
Den Hartog, C., 1977. \emph{Structure, function, and classification in seagrass communities}.
In: C.P. McRoy and C. Helfferich (Editors), Seagrass Ecosystems: a Scientific Perspective.
Marcel Dekker, New York, pp. 89-121. \pageref{DenHartog77}
\bibitem{Heege03}
Heege T., Bogner A., Pinnel N. 2003. \emph{Mapping of submerged aquatic vegetation with a
physically based process chain}. Proc. Remote Sensing, September 8-12 2--3/ Barcelona,
Spain. Ed. by Kramer E. The International Society for Optical Engineering, Vol.5233. \pageref{Heege03}
\bibitem{Hirose04}
Hirose K., Syoji M., Hang H.T.M., Anh N.H., Triet T. and Nam V.N. 2004. \emph{Satellite
data application for mangrove management}. International Symposium on Geoinfomatics for
Spatial Infrastructure Development in Earth and Allied Sciences. \pageref{Hirose04} 
\bibitem{Hochberg03a}
Hochberg, E.J., Atkinson, M.J. 2003a. \emph{Capabilities of remote sensors to classify coral,
algae, and sand as pure and mixed spectra}. Remote Sensing of Environment 85, 174–189. \pageref{Hochberg03a}
\bibitem{Hochberg03b}
Hochberg, E.J., Atkinson, M.J., Andrefouet, S., 2003b. \emph{Spectral reflectance of coral reef
bottom-types worldwide and implications for coral reef remote sensing}. Remote Sensing of
Environment 85, 159–173. \pageref{Hochberg03b}
\bibitem{Hogarth07}
Hogarth P. 2007. \emph{The Biology of Mangroves and Seagrasses}. Oxford University Press. \pageref{Hogarth07}
\bibitem{Holmer09}
Holmer M., Marba N, Lamote M and Duarte C. 2009\emph{Deterioration of Sediment Quality in
Seagrass Meadows (\textit{Posidonia oceanica}) Invaded by Macroalgae (Caulerpa sp.)}. Estuaries
and Coasts (2009) 32:456–466 DOI 10.1007/s12237-009-9133-4 \pageref{Hogarth07}
\bibitem{Holmer08}
Holmer M, Argyrou M, Dalsgaard T., Danovaro R, Diaz-Almela E, Duarte C,
Frederiksen M, Grau A, Karakassis I, Marbà N, Mirto S, Pérez M, Pusceddu A, Tsapakis M.2008 \pageref{}
\emph{Effects of fish farm waste on \textit{Posidonia oceanica} meadows: Synthesis and provision of
monitoring and management tools}. Marine Pollution Bulletin 56 (2008) 1618–1629. \pageref{Holmer08}
\bibitem{Iverson86}
Iverson R.L. and Bittaker H.F. 1986. \emph{Seagrass Distribution and Abundance in Eastern
Gulf of Mexico Coastal Waters}. Estuarine, Coastal and Shelf Science, 22, 577-602. \pageref{Iverson86}
\bibitem{Jackson07}
Jackson J., Nemeth D.J. 2007. \emph{A New Method to Describe Seagrass Habitat Sampled
During Fisheries-Independent Monitoring}. Estuaries and Coasts Vol. 30, No. 1,
p. 171–178 February 2007 \pageref{Jackson07}
\bibitem{Jensen01}
Jensen S. and Bell S. 2001. \emph{Seagrass growth and patch dynamics: cross-scale
morphological plasticity}. Plant Ecology 155: 201–217, 2001. \pageref{Jensen01}
\bibitem{Jensen95}
Jensen J.R., Rutchey K, Koch M.S, Narumalani S. 1995. \emph{Inland wetland change
detection in the everglades water conservation area 2a using a time-series of normalized
remotely-sensed data}. Photogrammetric Eng Remote Sensing 61: 199–209 \pageref{Jensen95}
\bibitem{Jerlov52}
Jerlov N.G. 1952. \emph{Influence of suspended and dissolved matter on the transparency of
sea water}. Wiley online Library. TellusVolume 5, Issue 1, Published online: 2010. \pageref{Jerlov52}
\bibitem{Kalinowski04}
Kalinowski A. and Oliver S. \emph{2004 ASTER Mineral Index Processing Manual}. Remote
Sensing Applications. Geoscience Australia \pageref{Kalinowski04}
\bibitem{Kelly80}
Kelly, M.G., 1980. \emph{Remote sensing of seagrass beds, Handbook of Seagrass Biology: An
Ecosystem Perspective} (R.C. Phillips and C.P. Mc Roy, editors), Garland STPM Press, New
York, N.Y., pp. 69-86. \pageref{Kelly80}
\bibitem{Kendrick00}
Kendrick GA, Hegge BJ, Wyllie A, Davidson A, Lord DA. 2000. \emph{Changes in Seagrass
Cover on Success and Parmelia Banks, Western Australia Between 1965 and 1995}.
Estuarine, Coastal and Shelf Science, 50:341-353. \pageref{Kendrick00}
\bibitem{Kirk94}
Kirk J.T.O. 1994. \emph{Light and photosynthesis in aquatic ecosystems}, pp 1–509. University
Press, Cambridge, UK. \pageref{Kirk94}
\bibitem{Kirkman96}
Kirkman H. 1996. \emph{Baseline and Monitoring Methods for Seagrass Meadows}. Journal of
Environmental Management (1996) 47, 191–201. \pageref{Kirkman96}
\bibitem{Kuo89}
Kuo J, McComb AJ, 1989. \emph{Seagrass taxonomy, structure and development. In: Larkum
AWD, McComb AJ, Shepherd, eds. Biology of seagrasses: a treatise on the biology of
seagrasses with special reference to the Australian region}. Elsevier, The Netherlands: 6-73. \pageref{Kuo89}
\bibitem{Leoni06}
Leoni V., Pasqualini V., Pergent-Martini Ch., Vela A., Pergent G. 2006. \emph{Morphological
responses of \textit{Posidonia oceanica} to experimental nutrient enrichment of the canopy water}.
Journal of Experimental Marine Biology and Ecology xx (2006) xxx–xxx. \pageref{Leoni06}
\bibitem{Leoni07}
Leoni V., Pasqualini V., Pergent-Martini Ch., Vela A., Pergent G. 2007. \emph{Physiological
responses of \textit{Posidonia oceanica} to experimental nutrient enrichment of the canopy water}.
Journal of Experimental Marine Biology and Ecology 349, 73–83. \pageref{Leoni07}
\bibitem{Leoni08}
Leoni V., Vela A., Pasqualini V., Pergent-Martini Ch., Pergent G. 2008. \emph{Effects of
experimental reduction of light and nutrient enrichments (N and P) on seagrasses: a review}.
Aquatic Conservation: Marine and Freshwater Ecosystems, 18: 202–220. \pageref{Leoni08}
\bibitem{Lewis01}
Lewis M. 2001. \emph{Discriminating Vegetation with Hyperspectral Imagery - What is
Possible?} http://www.ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=00978199 Proceedings,
International Geoscience and Remote Sensing Symposium, 9-13 July 2001, Sydney,
Australia. ISBN 0-7803-7031-7 \pageref{Lewis01}
\bibitem{Louchard03}
Louchard E.M., Reid R.P., Stephens F.C., Davis C.O., Leathers R.A., Downes T.V.
2003. \emph{Optical remote sensing of benthic habitats and bathymetry in coastal environments at
Lee Stocking Island, Bahamas: A comparative spectral classification approach}. Limnology
and Oceanography (2003) 48:511-521. \pageref{Louchard03}
\bibitem{Luque04}
Luque A. and Templado J.A. 2004. \emph{Praderas y bosques marinos de Andalucia
Consejeria de Medio Ambiente, Junta de Andalucia} (in press). \pageref{Luque04}
\bibitem{Lyzenga81}
Lyzenga D.R. 1981. \emph{Remote sensing of bottom reflectance and water attenuation in
shallow water using aircraft and Landsat data}. International Journal of Remote Sensing 2,
71-82. \pageref{Lyzenga81}
\bibitem{Maffione97}
Maffione R.A. and Dana D.R. 1997 \emph{Instruments and methods for measuring the
backward-scattering coefficient of ocean waters}. Optical Society of America \pageref{Maffione97}
\bibitem{Maffione01a}
Maffione R.A. 2001a. \emph{Evolution and Revolution in Measuring Ocean Optical Properties}.
Oceanography, Vol.14, No.3. \pageref{Maffione01a}
\bibitem{Maffione01b}
Maffione R.A. 2001b.\emph{Hyperspectral Optical Properties, Remote Sensing, and
Underwater Visibility}. Hydro-Optics, Biology, and Instrumentation Laboratories. \pageref{Maffione01b}
\bibitem{Malthus03}
Malthus T.J. and Karpouzli E. 2003. \emph{Integrating field and high spatial resolution
satellite based methods for monitoring shallow submersed aquatic habitats in the Sound of
Eriskay, Scotland, UK}. Int Journal Remote Sensing 24: 2585–2593. \pageref{Malthus03}
\bibitem{Matarrese08}
Matarrese R., Acquaro M., Morea A., Tijani K., Chiaradia M.T. 2008. \emph{Applications of
remote sensing techniques for mapping \textit{Posidonia oceanica} meadows}. Geoscience and
Remote Sensing Symposium. ISBN:978-1-4244-2807-6. \pageref{Matarrese08}
\bibitem{Matarrese06}
Matarrese R., Acquaro M., Morea A., Khalid T., Chiaradia M.T. 2006 \emph{Applications of remote
sensing techniques for mapping \textit{Posidonia oceanica} meadows}.
http://www.igarss08.org/Abstracts/pdfs/3888.pdf \pageref{Matarrese06}
\bibitem{McKenzie02}
McKenzie L.J., Campbell S.J. 2002. \emph{Manual for Community (citizen) Monitoring of
Seagrass Habitat}. Western Pacific Edition. Marine Plant Ecology Group, QDPI, Northern
Fisheries Centre, Cairns. 43 pp. \pageref{McKenzie02}
\bibitem{McKenzie03}
McKenzie L.J., Campbell S.J. and Roder C.A. 2003. \emph{Seagrass-Watch: Manual for
Mapping and Monitoring Seagrass Resources by Community (citizen) Volunteers}. 2nd
Edition. QFS, NFC, Cairns. 100pp. \pageref{McKenzie03}
\bibitem{McKenzieetal03}
McKenzie L.J. 2003. \emph{Guidelines for the rapid assessment of seagrass habitats in the
Western Pacific}. Marine Plant Ecology Group, QDPI, Northern Fisheries Centre, Cairns,
http://www.seagrasswatch.org \pageref{McKenzieetal03}
\bibitem{McKenzie06}
McKenzie L.J., Campbell S.J. and F.Lasi 2006. \emph{Seagrasses and Mangroves}. In: Green
A.P. Lokani W., Atu P., Ramonia P., Thomas P. and Almanmy J (eds). 2006 \emph{Solomon
Islands Marine Assessment. Technical report of survey conducted May 13 to June 17 2004}.
The Nature Conservancy, Indo-Pacific Resource Centre. Report No 1/06 \pageref{McKenzie06}
\bibitem{McKenzie07}
McKenzie L.J. and Yoshida R. 2007. \emph{Seagrass-Watch: Guidelines for Monitoring
Seagrass Habitats in the Fiji Islands. Proceeding of a training workshop. Corpus Christi
Teachers College Laucala Bay, Suva, Fiji, 16th June 2007}. Seagrass-Watch HQ Cairns, 42pp. \pageref{McKenzie07}
\bibitem{McKenzie08}
McKenzie L.J. 2008. \emph{Seagrass-Watch. Proceedings of a Workshop for Mapping and
Monitoring Seagrass Habitats in North East Arnhem Land, Northern Territory. 18-20
October 2008}. Seagrass-Watch HQ, Cairns, 49pp. \pageref{McKenzie08}
\bibitem{McKenzie09}
McKenzie L.J., Yoshida R. 2009. Seagrass Watch. Proceedings of a Workshop for
Monitoring Seagrass Habitats in Indonesia. The Nature Conservancy, Coral Triangle
Centre, Sanur, Bali, Indonesia. \pageref{McKenzie09}
\bibitem{Meinesz91}
Meinesz, A., Lefevre, J.R., Astier, J.M., 1991. \emph{Impact of coastal development on the
infralittoral zone along the south-eastern Mediterranean shore of continental France}.
Marine Pollution Bullettin 23, 343–347. \pageref{Meinesz91}
\bibitem{Mellors09a}
Mellors J. and McKenzie L. 2009(a). \emph{Seagrass-Watch: Guidelines for Monitoring
Seagrass Habitats. Proceedings of a workshop held on Thursday Island, Torres Strait 4-
8.03.2009}. \pageref{Mellors09a}
\bibitem{Mellors09b}
Mellors J and McLenzie L. 2009(b). \emph{Seagrass-Watch: Seagrass Monitoring Guidelines
for Torres Strait Communities. Meer Island. Proceedings of a workshop held on Meer
Island, Torres Strait 14-15 September 2009}. 58 pp. \pageref{Mellors09b}
\bibitem{Micheli05}
Micheli C., Paganin P, Peirano A., G. Caye, Meinesz A. and Bianchi C.N. 2005. \emph{Genetic
variability of \textit{Posidonia oceanica} (L.) Delile in relation to local factors and biogeographic
patterns}. Aquatic Botany 82 (2005) 210–221 \pageref{Micheli05}
\bibitem{Montefalcone09}
Montefalcone M. 2009.\emph{Ecosystem health assessment using the Mediterranean seagrass
\textit{Posidonia oceanica}: A review}. Ecological indicators 595 – 604 \pageref{Montefalcone09}
\bibitem{Montefalcone10}
Montefalcone M., Albertelli G., Morri C. and Bianchi C.N. 2010. \emph{Patterns of wide-scale
substitution within meadows of the seagrass \textit{Posidonia oceanica} in NW Mediterranean Sea:
invaders are stronger than natives}. Aquatic Conserv: Mar. Freshw. Ecosyst. 20: 507–515. \pageref{Montefalcone10}
\bibitem{Montefalcone06}
Montefalcone M., Albertelli G., Bianchi C.N., Mariani M., Morri C. 2006. \emph{A new
synthetic index and a protocol for monitoring the status of \textit{Posidonia oceanica} meadows: a
case study at Sanremo (Ligurian Sea, NW Mediterranean)}. Aquatic Conservation: Marine and
Freshwater Ecosystems, 16, 29–42. \pageref{Montefalcone06}
\bibitem{Mount03}
Mount R. 2003. \emph{The application of digital aerial photography to shallow water seabed
mapping and monitoring – how deep can you see? Coastal GIS 2003: and integrated
approach to Australian coastal issues, 7th - 8th July 2003}. Wollongong, University of
Wollongong, pp. 139-156. \pageref{Mount03}
\bibitem{Mumby99}
Mumby P.G., Green E.P., Edwards A.J. and Clark C.D. 1999. \emph{The cost-effectiveness of
remote sensing for tropical coastal resources assessment and management}. Journal of
Environmental Management Volume 55, Issue 3, March 1999, Pages 157-166 \pageref{Mumby99}
\bibitem{Mumby02}
Mumby, P. J. and Edwards, A. J. 2002. \emph{Mapping marine environments with IKONOS
imagery: Enhanced spatial resolution can deliver greater thematic accuracy}. Remote
Sensing of Environment, 82, 248–257 \pageref{Mumby02}
\bibitem{Mumby04}
Mumby P.J. Skirving W., Strong A.E., Hardy J.T., LeDrew E.F., Hochberg E.J., Stumpf
R.P. and David L.T. 2004. \emph{Remote sensing of coral reefs and their physical environment}.
Marine Pollution Bulletin, 48:219-228. \pageref{Mumby04}
\bibitem{Nagelkerken00}
Nagelkerken I., van der Velde G., Gorissen M.W., Meijer G.J., van’t Hofc T. and den
Hartog C. 2000. \emph{Importance of Mangroves, Seagrass Beds and the Shallow Coral Reef as a
Nursery for Important Coral Reef Fishes, Using a Visual Census Technique}. Estuarine,
Coastal and Shelf Science 51, 31–44 doi:10.1006/ecss.2000.0617, online:
http://www.idealibrary.com \pageref{Nagelkerken00}
\bibitem{Noralez10}
Noralez S. 2010. \emph{Exploring the Limits of Mapping Seagrass. A case study for P.
oceanica}. MSc Thesis. ITC. Netherlands. \pageref{Noralez10}
\bibitem{Norris97}
Norris J.G., Wyllie-Echeverria S., Mumford T., Bailey A. and Turner T. 1997.
\emph{Estimating basal area coverage of subtidal seagrass beds using underwater videography}.
Aquatic Botany 58 (1997) 269-287 \pageref{Noralez10}
\bibitem{SPOT} 
\emph{SPOT, Official Website} http://www.spotimage.com/web/en/172-spot-images.php \pageref{SPOT}
\bibitem{Orth06}
Orth R. J., Carruthers T.J.B., Dennison W.C., Duarte C. M., Fourqurean J.W. , Heck Jr.
K. L. , Hughes A. R., Kendrick G. A., Kenworthy W. J., Olyarnik S., Short F.T., Waycott
M. and Williams S. L. 2006. \emph{A Global Crisis for Seagrass Ecosystems}. Vol. 56 No. 12,
BioScience 987. \pageref{Orth06}
\bibitem{Palandro03}
Palandro D., Andréfouët S., Muller-Karger F. E., Dustan Ph., Hu Ch., and Hallock P.
2003. \emph{Detection of changes in coral reef communities using Landsat-5 TM and Landsat-7
ETM+ data}. Can. J. Remote Sensing, Vol. 29, No. 2, pp. 201–209. \pageref{Palandro03}
\bibitem{Pasqualini98a}
Pasqualini V., Pergent-Martini C. and Pergent, G. 1998(a). \emph{Use of remote sensing for the
characterization of the Mediterranean coastal environment – the case of Posidonia
oceanica}. Journal of Coastal Conservation 4: 59-66, 1998, EUCC; Opulus Press Uppsala. \pageref{Pasqualini98a}
\bibitem{Pasqualini98b}
Pasqualini V., Pergent-Martini C., Clabaut P., Pergent G. 1998(b). \emph{Mapping of
\textit{Posidonia oceanica} using aerial photographs and side scan sonar: Application off the
island of Corsica (France)}. Estuarine Coastal Shelf Sci 47:359–367. \pageref{Pasqualini98b}
\bibitem{Pasqualini99}
Pasqualini V, Pergent-Martini C, Pergent G. 1999. \emph{Environmental impact identification
along the Corsican coast (Mediterranean Sea) using image processing}. Aquatic Botany,
65:311-320. \pageref{Pasqualini99}
\bibitem{Pasqualini01}
Pasqualini V., Pergent-Martini C, Clabaut P, Marteel H., and Pergent G. 2001.
\emph{Integration of Aerial Remote Sensing, Photogrammetry, and GIS Technologies in Seagrass
Mapping}. Photogrammetric Engineering and Remote Sensing Vol. 67, No. 1, January 2001,
pp. 99-105. \pageref{Pasqualini01}
\bibitem{Pasqualini05}
Pasqualini V., Pergent-Martini C, Pergent G., Agreila M., Skoufas G., Sourbes L. and
Tsirika A. 2005. \emph{Use of SPOT 5 for mapping seagrasses: An application to Posidonia
oceanica}. Remote Sensing of Environment, 94:39-45. \pageref{Pasqualini05}
\bibitem{Peneva08}
Peneva E., Griffith J.A. and Carter G.A. 2008. \emph{Seagrass Mapping in the Northern Gulf
of Mexico using Airborne Hyperspectral Imagery: A Comparison of Classification Methods}.
Journal of Coastal Research, Number 244:850-856. 2008. Published By: Coastal Education
and Research Foundation DOI: 10.2112/06-0764.1 \pageref{Peneva08}
\bibitem{Peneva07}
Peneva E.I., Griffith J.A. and Carter G.A. 2007. \emph{Mapping seagrass distribution on Horn
Island}, MS using hyperspectral imagery.
ftp://ftp.ecn.purdue.edu/jshan/proceedings/asprs2005/Files/0134.pdf \pageref{Peneva07}
\bibitem{Pergent90}
Pergent G. 1990. \emph{Lepidochronological analysis of the seagrass \textit{Posidonia oceanica} (L.)
Delile: a standardized approach}. Aquatic Botany, 37 (1990) 39-54 39 Elsevier Science
Publishers B.V., Amsterdam -- Printed in The Netherlands \pageref{Pergent90}
\bibitem{Pergent-Martini06}
Pergent-Martini C., Pasqualini V., Ferrat L and Pergent G. 2006. \emph{Ecological Data in
Integrated Coastal Zone Management: Case Study of \textit{Posidonia oceanica} Meadows Along
the Corsican Coastline (Mediterranean Sea)}. Environ Manage (2006) 38:889–895 DOI
10.1007/s00267-005-0056-y. \pageref{Pergent-Martini06}
\bibitem{Pergent-Martini05}
Pergent-Martini C, Leoni V., Pasqualini V., Ardizzone G.D., Balestri E., Bedini R.,
Belluscio A., Belsher T., Borg J., Boudouresque C.F., Boumaza S., Bouquegneau J.M., Buia
M.C., Calvo S., Cebrian J., Charbonnel E., Cinelli F., Cossu A., Di Maida G., Dural B.,
Francour P., Gobert S., Lepoint G., Meinesz A., Molenaar H., Mansour H.M., Panayotidis
P., Peirano A., Pergent G., Piazzi L, Pirrotta M, Relini G, Romero G, Sanchez-Lizaso J.L,
Semroud R., Shembri P., Shili A., Tomasello A., Velimirov B. 2005. \emph{Descriptors of
\textit{Posidonia oceanica} meadows: Use and application}. Ecological Indicators 5, 213–230. \pageref{Pergent-Martini05}
\bibitem{Phinn08}
Phinn S, Roelfsema C, Dekker A, Brando V, Anstee J. 2008. \emph{Mapping seagrass species,
cover and biomass in shallow waters: An assessment of satellite multi-spectral and airborne
hyperspectral imaging systems in Moreton Bay (Australia)}. Remote Sensing of
Environment, 112:3413-3425. \pageref{Phinn08}
\bibitem{Ralph05}
Ralph P.J., Macinnis-Ng C.M.O., Frankart C. 2005. \emph{Fluorescence imaging application:
effect of leaf age on seagrass photokinetics}. Aquatic Botany 81 (2005) 69–84. \pageref{Ralph05}
\bibitem{Rasib97}
Rasib A.W. and Hashim M. 1997. \emph{Mapping Seagrass From Satellite Remote Sensing
Data. Proceedings of Asian Conference on Remote Sensing}. http://eprints.utm.my/2161/ \pageref{Rasib97}
\bibitem{Ribed02}
Ribed P.S. 2002. \emph{Submarine vegetation studies using remote sensing: importance in
beach measurements}. Journal of Environmental Protection and Ecology, 3 No 4, 953-958
(2002) \pageref{Ribed02}
\bibitem{Richardson06}
Richardson L.L., LeDrew Ellesworth F. 2006. \emph{Remote Sensing of Aquatic Costal
Ecosystems Processes}. Springer. Science and Management Applications. \pageref{Richardson06}
\bibitem{Robbins94}
Robbins B.D., Bell S.S. 1994. \emph{Seagrass landscapes: a terrestrial approach to the marine
subtidal environment}. TREE vol.9, no 8, doi:10.1016/0169-5347(94)90041-8 \pageref{Robbins94}
\bibitem{Sabins97}
Sabins F. 1997. \emph{Remote Sensing, Principles and Interpretation}. W.H. Freeman and
Company, New York. \pageref{Sabins97}
\bibitem{Schmidt03}
Schmidt K. and Skidmore A.K. 2003. \emph{Spectral discrimination of vegetation types in a
coastal wetland}. Remote Sensing of Environment. 85:92-108. \pageref{Schmidt03}
\bibitem{Schmieder97}
Schmieder K. 1997. \emph{Littoral zone - GIS of Lake Constance: a useful tool in lake
monitoring and autecological studies with submersed macrophytes}. Aquatic Botany 58
(1997) 333-346 \pageref{Schmieder97}
\bibitem{Schultz08}
Schultz S.T. 2008. \emph{Seagrass monitoring by underwater videography: Disturbance
regimes, sampling design, and statistical power}. Aquatic Botany 88 (2008) 228–238. \pageref{Schultz08}
\bibitem{Short01}
Short F.T. and Coles R.G. (eds). 2001. \emph{Global Seagrass Research Methods}, p 482.
Elsevier Publishing Co., Amsterdam \pageref{Short01}
\bibitem{Shortis07}
Shortis M., Harvey E., Seager J. 2007. \emph{A Review of the Status and Trends in Underwater
Videometric Measurement}. SPIE Conference 6491, Videometrics IX,IS and T/SPIE Electronic
Imaging, San Jose, California, U.S.A., 28 January–1 February 2007 \pageref{Shortis07}
\bibitem{Stephens03}
Stephens F.C., Louchard E.M., Reid R.P., Maffione R.A. 2003. \emph{Effects of microalgal
communities on reflectance spectra of carbonate sediments in subtidal optically shallow
marine environments}. Limnology and Oceanography, 48:535-546. \pageref{Stephens03}
\bibitem{Thorhaug07}
Thorhaug, A., Richardson, A. D. and Berlyn, G. P. 2007. \emph{Spectral reflectance of the
seagrasses: Thalassia testudinum, Halodule wrightii, Syringodium filiforme and five marine
algae}. International Journal of Remote Sensing, 28: 7, 1487 — 1501. \pageref{Thorhaug07}
\bibitem{Toole00}
Toole D.A., Siegel D.A., Menzies D.W., Neumann M.J. and Smith R.C. 2000. 
\emph{Remote-sensing reflectance determinations in the coastal ocean
environment: impact of instrumental characteristics and environmental variability}. Applied
Optics 39(3), 456-469. \pageref{Toole00}
\bibitem{Vahtmae06}
Vahtmae E, Kutser T, Martin G, Kotta J. 2006. \emph{Feasibility of hyperspectral remote
sensing for mapping benthic macroalgal cover in turbid coastal waters - a Baltic Sea case
study. Remote Sensing of Environment} (2006) 101:342-351. \pageref{Vahtmae06}
\bibitem{Vahtmae07}
Vahtmae E. and Kutser T. 2007. \emph{Mapping bottom type and water depth in shallow
coastal waters with satellite remote sensing}. Journal of Coastal Research, SI 50 (Proceedings
of the 9th International Coastal Symposium), 185 – 189. Gold Coast, Australia, ISSN
0749.0208 \pageref{Vahtmae07}
\bibitem{Villegas06}
Villegas P. P. 2006. \emph{Magnitude of herbivory in \textit{Posidonia oceanica} (L.) Delile and
factors responsible for spatial variation}. Available on web: http://www.tesisenxarxa.net/ \pageref{Villegas06}
\bibitem{Wabnitz08}
Wabnitz C.C., Andréfouët S., Torres-Pulliza D., Müller-Karger F. E. and Kramer P.A.
2008. \emph{Regional-scale seagrass habitat mapping in the Wider Caribbean region using
Landsat sensors: Applications to conservation and ecology}. Remote Sensing of
Environment Vol. 112, Issue 8, 15 August 2008, Pages 3455-3467 Earth Observations for
Marine and Coastal Biodiversity and Ecosystems Special Issue. \pageref{Wabnitz08}
\bibitem{Walker89}
Walker, D.I., 1989. \emph{Methods for monitoring seagrass habitat}. Proceedings Rep.
Workshop, 20-22 June 1988, Melbourne, Australia, VIMS Working Paper 18, pp. 1-26. \pageref{Walker89}
\bibitem{Werdell03}
Werdell P.J. and Roesler C.S. 2003. \emph{Remote assessment of benthic substrate composition
in shallow waters using multispectral reflectance}. Limnol. Oceanogr., 48 (1, part 2), 557–
567 \pageref{Werdell03}
\bibitem{Yang10}
Yang Ch., Yang D., Cao W., Zhao J., Wang G., Sun Zh., Xu Zh., Kumar M. S. R. 2010. 
\emph{Analysis of seagrass reflectivity by using a water column correction algorithm}. International
Journal of Remote Sensing. 31: 17, 4595 — 4608. \pageref{Yang10}

\end{thebibliography}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%% End MSc Thesis Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-01-09 MScDiplom_Lemenkova.tex
remove leftover
added tables capturing and updated figures referencing

@@
 
-\section[Overview of literature]{Seagrass monitoring: overview of literature and research resources}
+\section[Overview of literature]{Seagrass monitoring: overview of literature and research resources} \label{sec:2}
 %\markright{II. Seagrass monitoring: overview of literature and research resources}
@@
 Landsat TM images with their 30*30m resolution.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+\centering
+\includegraphics[scale=0.30]{Fig-14.jpg}
+\caption{Multispectral vs. hyperspectral band
+coverage. Source: \cite{Sabins97}\label{Sabins97}}
+\label{fig:14}
+\end{wrapfigure}
+
 Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}), launched in
@@
 pure reflectance from the benthos; as a result, it may cause significant errors \cite{Dierssen03}\label{Dierssen03}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+\centering
+\includegraphics[scale=0.15]{Fig-15.jpg}
+\caption{Reflectance spectra of seagrass; Rrs(0-)
+- subsurface RS reflectance; Rrs b - the bottom
+reflectance. Source:\cite{Yang10}\label{Yang10}}
+\label{fig:15}
+\end{wrapfigure}
+
 Therefore, for the accurate assessments of various seafloor cover types, water
@@
 hyperspectral sensors has better potential due to their higher resolution.
-\pagebreak
 
@@
 The application of the hyperspectral sensors is most effective and provide more accurate classification
-results (Fig.2.1.),
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-\centering
-\includegraphics[scale=0.30]{Fig-14.jpg}
-\caption{Multispectral vs. hyperspectral band
-coverage. Source: \cite{Sabins97}\label{Sabins97}}
-\label{fig:14}
-\end{wrapfigure}
-
-due to their higher spectral resolution \cite{Bharathi03}\label{Bharathi03} with interval
+results (Fig.2.1.), due to their higher spectral resolution \cite{Bharathi03}\label{Bharathi03} with interval
 narrows to 10 nanometers, while broadband sensors are limited to the spectral width of ca
@@
 high spectral resolution, desirable for the case of seagrass monitoring. 
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-\centering
-\includegraphics[scale=0.15]{Fig-15.jpg}
-\caption{Reflectance spectra of seagrass; Rrs(0-)
-- subsurface RS reflectance; Rrs b - the bottom
-reflectance. Source:\cite{Yang10}\label{Yang10}}
-\label{fig:15}
-\end{wrapfigure}
-
 The most suitable
@@
 
-The research is technically based on WASI, for the analysis of spectral signatures, \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} and \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software for the processing and analysing of aerial and broadband images, Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat}.
+In section \nameref{sec:2} we briefly discussed main existing cartographic methods and tools for seagrss mapping.
+In current  research we try to combine various techniques for assessment of seagrass in Mediterranean environment, in the cse study of Crete.
+Therefore, logically separated research sections are based on various software and tools.
+thus, spectral analysis and assessment of spectral signaturesis technically based on WASI, sttistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using\href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
 \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0} software is used for the spatial analysis, general mapping and cartographic layout
@@
 current work is Google Earth aerial images. The  \href{http://landsat.gsfc.nasa.gov/}{Landsat} images are used for the general overview.
-Besides aerial and satellite photographs, data acquired during the fieldwork (21 Sep – 11 Oct) are
+Besides aerial and satellite photographs, data acquired during the fieldwork (three weeks in September - October) are
 necessary addition to the mapping helping to solve problems of interpretation during the images
@@
 
-\pagebreak
-
 \subsection{Study area}
@@
 geograpic locations recorded using GPS.
+\pagebreak
 
@@
 The videographic measurements and photos, enabling to detect types of the
-seafloor, were captured by means of three underwater digital video cameras Olympus ST 8000 (Fig.\ref{fig:20}), at
+seafloor, were captured by means of three underwater digital video cameras Olympus ST 8000 (Fig.\ref{fig:18}), at
 five minutes interval along the tracklog path (Fig.\ref{fig:21}).
-research sampling design also included measurements of depths, because bathymetry is one of the
+research sampling design also included measurements of depths (Fig. \ref{fig:63}), because bathymetry is one of the
 most determinating factors for the seagrass locations. Measurements of depths were performed during
@@
 with videographic measurements. The GPS allowed to capture measurement locations on the iPAQ,
-encapsulated in a plastic waterproof Otterbox. Camera were adjusted horizontally by a leveller and
-were mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{app:20}). The data
+encapsulated in a plastic waterproof Otterbox (Fig \ref{}). Camera were adjusted horizontally by a leveller and
+were mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:20}). The data
 were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state. The
@@
 coverage of the seagrass meadows in the research area. The underwater measurements of the seagrass
-coverage were carried out by taking video footage and photos of ca 0.5 m2 size each.
+coverage were carried out by taking video footage and photos of ca 0.5 m2 size each (Fig \ref{66}).
 The results of the underwater videometric measurements include series of digital images helping to
@@
 Along the Ligaria beach there are several types of the seagrass landscapes, namely, segrass meadows
-continuously covering vast areas (Fig.\ref{app:18}), aggregated seagrass patches, represented by separate
-mattes with short irregular channels between them (Fig.\ref{app:18}) and isolated seagrass patches, or mattes,
-which located separately from broad meadows (Fig.\ref{app:18}).
+continuously covering vast areas (Fig.\ref{fig:62} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:62} (b)) and isolated seagrass patches, or mattes,
+which located separately from broad meadows (Fig.\ref{fig:62} (c)).
 The results of the underwater videometric measurements of the Olympus cameras made during the
@@
 information on seagrass presence within the study area, distribution of seagrass P.oceanica meadows
-and nature of the seafloor cover types (rocks, sandy, mixed). Seagrass species on the Ligaria and Agia
+and nature of the seafloor cover types: rocks, sandy, mixed (Fig. \ref{fig:65}). Seagrass species on the Ligaria and Agia
 Pelagia beach consist of P.oceanica. The types of sediment on the Ligaria vary from coarse sand,
-P.oceanica patches, sand, mud, rock gravel and fine sand (Fig.\ref{app:20}).
+P.oceanica patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:64}).
 
@@
 The overview of the collected data enables to summarize their following types:
-1. Spectra of P.oceanica and other seafloor types at different environmental conditions
-2. Aerial imagery from the Google Earth
-3. Satellite images from the open sources (Landsat)
-4. Underwater videometric measurements of the Olympus cameras made during the ship route
-The available satellite imagery (Fig.\ref{app:20}) and aerial images are read into the ArcGIS project.
+\begin{itemize}
+\item[] 1.
+Optical spectra of P.oceanica, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig.
+\ref{fig:23})
+\item[]2. Aerial imagery from the \href{http://www.google.com/earth/index.html}{Google Earth}
+\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:68})
+\item[]4. Results of underwater videometric measurements of the Olympus cameras made during the ship route
+\end{itemize}
+The available satellite imagery (Fig.\ref{fig:68}) and aerial images are read into the ArcGIS project.
 The available broadband and hyperspectral remote sensing data are used for the mapping of the
@@
 research includes fieldwork in-situ measurements of the seagrass distribution along the northern and
-south-eastern coasts of Crete in chosen locations.
+south-eastern coasts of Crete in chosen locations.\\
 The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
 the seafloor have been received by the means of the optical sensors Trios-RAMSES Hyperspectral
-UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.2.3
-and 2.4), both adjusted for the measurements of the irradiance and radiance.
-The highest measured values are located in the diapason of 410-730 nm for the water irradiance and 430-650 for the water radiance (Fig.\ref{fig:23}).
-
-%\begin{figure}[h]
-	%\centering
-	%\subfloat [Irradiance]{\label{fig:23a}\includegraphics[width=0.8\textwidth]{Fig-23a.jpg}}\\
-	%\subfloat [Radiance]{\label{fig:23b}\includegraphics[width=0.8\textwidth]{G-Radiance_Bezier.jpg}}
-	%\caption{Optical properties of the sea water with sediments}
-	%\label{fig:23}
-%\end{figure}
-
+UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.
+\ref{fig:10}), both adjusted for the measurements of the irradiance and radiance (see Appendices: Tab.
+\ref{tab:18}, Tab.
+\ref{tab:19}).
+The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:49}) and 430-650 for the water radiance (Fig.\ref{fig:47}). 
 Afterwards, the measured values of the
@@
 nm, which is suitable for a characteristics of seagrass reflectance.
-Different curves on the radiance and irradiance graphs (Fig.\ref{fig:23}) represent several series of
+Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:43}, or \ref{fig:45}, enlarged) represent several series of
 the measurements. The values of the spectral reflectance are received from the
-computations of these values using Excel mathematical formulae. The graphs shown on Fig.\ref{fig:23} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and
-without sediments and suspended particles, respectively.
+computations of these values using Excel mathematical formulae.
+The graphs shown on Fig.\ref{fig:47}, Fig.
+\ref{fig:48}, Fig.
+\ref{fig:49} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively.
+Graph on Fig \ref{fig:44} displays statistuical analysis of the measured sets of observations, i.e.
+shows the midspread of the statistical quartiles (Q1-Q3), mean and extreme values.
 \begin{figure}[h]
@@
 \caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete}
-\label{fig:25}
-\end{figure}
+\label{fig:23}
+\end{figure}
+\pagebreak
 
 \subsection{Data pre-processing}
-A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonatic sand - were
-tested for the spectral variability and separability under varying conditions of dfifferent environmental
-constituents (e.g. depth, water content, sun angle etc), in order to determine the
-potential that it may have on the approaches for further images processing and classification.The raw
-initial data of measurements of the spectral reflectance have been preprocessed and analysed. These data
+A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonatic sand - consist of multiple measurement sets of P.oceanica made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October, 1-445 for 14th October for P.oceanica (Tab.
+\ref{tab:10}); 1-106 for water without sedimnts, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
+\ref{tab:17}).
+All these data have been analysed and statistically tested.
+We measured mean, extreme values (min-max), median and statistical quartiles, in order to visualise distribution of the values at varuious spectral wavelengths.
+Carbonate sand was measured at wavelengths 402-750 nm, while P.oceanica - at the interval 318-951 nm.
+We visualised the behavour of different spectra on relevant graphs (Fig.
+\ref{fig:43}, \ref{fig:44}, \ref{fig:45}).
+The available collected data were tested for the spectral variability and separability under varying conditions of dfifferent environmental
+constituents (e.g. depth, water content, sun angle), in order to determine the
+potential that it may have on the approaches for further images processing and classification.
+Behavour of the spectra of P.oceanica have been tested using datasets for various depths: 0.5, 1.5 and 2.5 meters (Tab.
+\ref{tab:8}, Tab.
+\ref{tab:9}).
+The raw
+initial data of measurements of the spectral reflectance have been preprocessed and statistically analysed thereafter withgraphs visualising mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:33}, \ref{fig:28}).
+These data
 have been measured using different step of the wavebands: some measurements
@@
 
-\begin{wrapfigure}{r}{0.6\textwidth}
-\centering
-\includegraphics[scale=0.23]{Fig-24.jpg}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.30]{Fig-24.jpg}
 \caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
 \label{fig:24}
-\end{wrapfigure}
+\end{figure}
 
@@
 Pelagia beach, Crete), sand (measured at Agia Pelagia beach, Crete), silt and default artificial
-spectrum of constant albedo at WASI (Fig.\ref{fig:25}). We included measurements of the Thalassia seagrass
+spectrum of constant albedo at WASI (Fig.\ref{fig:25}). 
+We included measurements of the Thalassia seagrass
 (measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectances of
 different seagrass speciea under various environmental conditions.
-
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.25]{Fig-28.jpg}
-\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
-seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
-\label{fig:25}
-\end{figure}
-
 The most suitable wavebands for the seagrass monitroring usually lay between 400 and 700 nm,
@@
 The initial measured data were stored in raw-oriented format, so that re-formatting them into the
-column-based layout was done using “transpose” command in Open Office. The next step included
-calculation of the median, mean, quartiles and other statistical values at every data set (see Appendix).
+column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:42}). The next step included
+calculation of the median, mean, quartiles and other statistical values at every data set (see Appendix, Fig. \ref{fig:54}, Fig. \ref{fig:55}).
 After the preliminary analysis, the measured data were visualised using GNUplot programm, which
-enables fine plotting of various data. The most acceptable method of interpolation was Bézier curve ,
+enables fine plotting of various datasets together: Fig \ref{fig:26}, Fig.\ref{fig:27}, Fig.\ref{fig:28}.
+The most acceptable method of interpolation was B\'ezier curve (Fig.
+\ref{fig:47}, Fig.
+\ref{fig:49}),
 as it has trend-friendly graph better showing the general behavior of the curves at different wavebands
-comparing to splines (see Appendix for more results). Therefore, after several experiment with
-various interpolation techniques, we have choosen Bézier curves interpolation which contain convex
+comparing to splines (see Appendix for more results).
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{Fig-28.jpg}
+\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
+seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
+\label{fig:25}
+\end{figure}
+
+Therefore, after several experiments with
+various interpolation techniques (Fig.\ref{fig:48}), we have choosen Bézier curves interpolation (Fig.\ref{fig:49}) which contain convex
 hulls made on its control points, and therefore is best suitable for our case: analysis of optical
 properties of seawater.
+%
 \subsection{Analysis of spectral signatures}
@@
 \\Two complimentray graphs below show the results of the statistical analysis}
-\label{fig:26}
+\label{fig:27}
 \end{figure}
@@
 \caption{Statitical comparison of spectral reflectances of \textit{P.oceanica} and carbonate sand. Selected measurements sets}
-\label{fig:26}
+\label{fig:28}
 \end{figure}
@@
 	\caption{Simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
-	\label{fig:28}
+	\label{fig:29}
 \end{figure}
@@
 4.0m}
-\label{fig:29}
+\label{fig:30}
 \end{figure}
@@
 	\caption{Sea water physical properties, modelled by WASI}
-	\label{fig:30}
+	\label{fig:31}
 \end{figure}
@@
   \end{tabular}
-   \label{table:1}
+   \label{tab:3}
 \end{table}
@@
 	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths}
-	\label{fig:31}
+	\label{fig:32}
 \end{figure}
@@
 The patches of white sandy bottoms of the seafloor, studied in the fieldwork in Ligaria beach, are
-much brighter than mattes of P.Oceanica ((Fig.\ref{app:30}): bottom types in Ligaria), which can be clearly seen
-at the graph comparing values of the spectral reflectance of the carbonate sand and than that of P.
-oceanica (Fig.\ref{fig:33}).
+much brighter than mattes of P.Oceanica ((Fig.\ref{fig:64}): seafloor types in Ligaria), which can be clearly seen
+at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:65}) and that of P.
+oceanica (Fig.\ref{fig:62}).
 
@@
 \caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and measured values (dotted): multiplot of measurement sets 200-300}
-\label{fig:29}
+\label{fig:33}
 \end{figure}
@@
 	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
-	\label{fig:33}
-\end{figure}
+	\label{fig:34}
+\end{figure}
+\pagebreak
 
@@
 mean value within the range (red bold dots). Gnuplot visualization­}
-\label{app:1}
+\label{fig:35}
 \end{center}
@@
 mean value within the range (red bold dots). Gnuplot visualization­}
-\label{app:1}
+\label{fig:36}
 \end{center}
@@
 Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.
+To analyse average values of spectral reflectance of P.oceanica by means of the measured sets we calculated 
+mean values for the total set of emasurements 1 – 350 (15th October): mean of min and max values, 
+mean of average for each waveband, mean for statistical quartiles for each wavelength and mean for median, respectively. 
+The adjusted averages evaluate mean values, needed for statistical analysis of the datasets. 
+The outcome of this calculations is shown in the table 5 (Appendices). 
+As a result, we have mean values of spectral reflectances of P.oceanica for extreme, average, median and both quartiles,
+independent of the individual sets 1-350.
 
 \paragraph{ANOVA-testing}
-The ANOVA (ANalysis Of Variance) testing is applied for the analysis of the probability of the
+We applied ANOVA (ANalysis Of Variance) testing for the analysis of the probability of the
 spectral reflectance of P.oceanica to be more or less spectrally disticnt from other seafloor cover
@@
 \paragraph{Student-t testing}
-The Student-t test is one of the most commonly used techniques for testing a hypothesis on the basis
-of a difference between sample means. In our case the Student t-test demonstrates if the variation
+The Student-t test, one of the most commonly used techniques for testing a hypothesis on the basis
+of a difference between sample means, is used for the data analysis. In our case the Student t-test demonstrates, if the variation
 between two analysed groups – values of spectral reflectances of seagrass P.oceanica and sand - is
@@
   \end{tabular}
-   \label{table:1}
+   \label{tab:4}
 \end{table}
@@
 
-\pagebreak
-
 \subsection{Image processing using Erdas Imagine}
@@
 	\caption{Seafloor types of the Arina beach: \textit{P.oceanica} sand, rocks}
-	\label{fig:35}
+	\label{fig:37}
 \end{figure}
@@
 \caption{Results of the image classification in Erdas Imagine: seagrass distribution in Bali area, Crete}
-\label{fig:36}
+\label{fig:38}
 \end{center}
@@
 \caption{Google aerial images incorporated into the GIS project: fragment of ArcMap layout}
-\label{fig:37}
+\label{fig:39}
 \end{center}
@@
 	\caption{Variations in spatial structure of the seagrass landscapes}
-	\label{fig:38}
+	\label{fig:40}
 \end{figure}
@@
 \caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the ArcGIS project}
-\label{fig:39}
+\label{fig:41}
 \end{center}
@@
 \caption{Values of spectral reflectances of seagrass P.oceanica (fragment)}
-\label{app:1}
+\label{fig:42}
 \end{center}
@@
 \caption{Spectral reflectance of P.oceanica. Measurement series 401-420}
-\label{app:1}
+\label{fig:43}
 \end{center}
@@
 Visualisation of the interquartile ranging. Example of data set 401-420. Shown midspread of  statistical quartiles Q1 and Q3, min and max values within the range.}
-\label{app:1}
+\label{fig:44}
 \end{center}
@@
 Visualisation of the interquartile ranging and plotted together with measurement data.}
-\label{app:1}
+\label{fig:45}
 \end{center}
@@
 SPSS}
+\label{tab:5}
 \begin{center}
@@
 \caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textbf{P.oceanica}. SPSS}
+\label{tab:6}
 \begin{center}
@@
 \caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. SPSS}
+\label{tab:7}
 \begin{center}
@@
 Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means.}
-\label{app:1}
+\label{fig:46}
 \end{center}
@@
 aquarium tank. Bezier Interpolation}
-\label{fig:23}
+\label{fig:47}
 \end{figure}
@@
 \caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters}
+\label{tab:8}
 \begin{center}
@@
 \end{center}
-P>0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
+P more than .05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
 \label{fig:5}
@@
 \caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters}
+\label{tab:9}
 \begin{center}
@@
 \begin{longtable}{|c|c|c|c|c|c|c|}
-\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October} \label{tab:6} \\
+\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October} 
+\label{tab:10} \\
 \hline
@@
 
-\multicolumn{7}{c}{Table 4: Spectral reflectance of \textbf{P.oceanica} \textit{(continued from previous page)}}\\
+\multicolumn{7}{c}{Table 10: Spectral reflectance of \textbf{P.oceanica} \textit{(continued from previous page)}}\\
 \hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
@@
 \caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in GNUplot­}
-\label{app:1}
+\label{fig:48}
 \end{center}
@@
 \caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in GNUplot}
-\label{fig:21}
+\label{fig:49}
 \end{figure}
@@
 \caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
+\label{tab:11}
 \begin{center}
@@
 \caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
-\label{fig:21}
+\label{fig:50}
 \end{figure}
@@
 \caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
+\label{tab:12}
 \begin{center}
@@
 \end{center}
-\label{fig:5}
 \end{table}
@@
 \caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
-\label{fig:21}
+\label{fig:51}
 \end{figure}
@@
 \caption{Spectral reflectance of seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle (GNUplot)­}
-\label{app:1}
+\label{fig:52}
 \end{center}
@@
 \caption{Spectral reflectance of the seawater without sediments. Interpolation graph. GNUplot display­}
-\label{app:1}
+\label{fig:53}
 \end{center}
@@
 \caption{Spectral reflectance of P.oceanica. Measurement series 25-50. GNUplot display­}
-\label{app:1}
+\label{fig:54}
 \end{center}
@@
 mean value within the range.­}
-\label{app:1}
+\label{fig:55}
 \end{center}
@@
 mean value within the range (red bold dots). GNUplot visualization­}
-\label{app:1}
+\label{fig:56}
 \end{center}
@@
 \caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+\label{tab:13}
 \begin{center}
@@
 \caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+\label{tab:14}
 \begin{center}
@@
 \caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+\label{tab:15}
 \begin{center}
@@
 \caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+\label{tab:16}
 \begin{center}
@@
 mean value within the range (red bold dots). GNUplot visualization­}
-\label{app:1}
+\label{fig:57}
 \end{center}
@@
 Wavelength step: 3 nm.
-Measured on Agia Pelagia beach, 14th October} \label{fig:7} \\
+Measured on Agia Pelagia beach, 14th October}
+\label{tab:17} \\
 \hline
@@
 \caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES}
-\label{fig:34}
+\label{fig:58}
 \end{figure}
@@
 
-\begin{table}
+\begin{table}[H]
 \caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}
+\label{tab:18}
 \centering
@@
   \end{tabular}
-   \label{table:1}
 \end{table}
-
+\pagebreak
+
+\appendixname{\textbf{3:\\ Google Earth aerial images covering research area of Crete island}}
 \begin{figure}[H]
@@
 \caption{Locations of selected measurements, visualization on Google Earth­}
-\label{app:1}
-\end{center}
-\end{figure}
+\label{fig:59}
+\end{center}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[scale=0.25]{App-8.jpg}
+\caption{Random mosaic of selected aerial Google Earth images}
+\label{fig:60}
+\end{center}
+\end{figure}
+\pagebreak
 
@@
 \caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}
+\label{tab:19}
 \centering
@@
   \end{tabular}
-   \label{table:1}
 \end{table}
-
-\appendixname{\textbf{3:\\ Google Earth aerial images covering research area of Crete island}}
-\begin{figure}[h]
-\begin{center}
-\includegraphics[scale=0.25]{App-8.jpg}
-\caption{Spectral reflectance of P.oceanica. Measurement series 1-16}
-\label{app:8}
-\end{center}
-\end{figure}
 
@@
 \caption{Script command of FWTools}
-\label{app:9}
+\label{fig:61}
 \end{center}
@@
 	\caption{Types of seagrass structural patterns, Ligaria beach, Crete}
-	\label{app:30}
+	\label{fig:62}
 \end{figure}
@@
 \caption{Scheme of the depths measurements; fieldwork on Ligaria beach}
-\label{app:9}
+\label{fig:63}
 \end{center}
@@
 \caption{Ligaria beach, Crete: seafloor types}
-\label{app:9}
+\label{fig:64}
 \end{center}
@@
 \caption{Various seafloor types; Ligaria beach, Crete}
-\label{app:9}
+\label{fig:65}
 \end{center}
@@
 \caption{Measurement underwater equipment}
-\label{app:9}
+\label{fig:66}
 \end{center}
@@
 \caption{Interpolation of the spectral measurements}
-\label{app:10}
+\label{fig:67}
 \end{center}
@@
   \end{tabular}
-   \label{table:1}
+\label{tab:20}
 \end{table}
@@
 	\caption{Landsat imagery, Crete island. Previews}
-	\label{app:30}
+	\label{fig:68}
 \end{figure}

### 2011-01-14 MScDiplom_Lemenkova.tex
added statistics in Gretle, R

@@
 \usepackage{calc}
+\usepackage{amsmath}
 \usepackage{longtable}
@@
 \item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
+\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
 \end{itemize}
@@
 \end{figure}
+\pagebreak
+
+\begin{table}[htbp]
+\rowcolors{1}{Ivory}{GhostWhite}
+\caption{Summary of the table with values of the spectral reflectance of \textbf{P.oceanica} measured on three different depths, R}
+\label{tab:13}
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline\hline
+wl & X0.5m & X1.5m & X2.5m\\ \hline\hline
+Min.   :318.2   &  Min.   :0.00000     &   Min.   : 0.7792     &   Min.   :0.005096  \\ \hline
+ 1st Qu.:476.9   &  1st Qu.:0.03398    &  1st Qu.:14.4084    &  1st Qu.:0.091878  \\ \hline
+ Median :636.1   &   Median :0.10955    &  Median :22.4621   &  Median :0.269328 \\ \hline 
+ Mean   :635.6   &   Mean   :0.11392   &   Mean   :19.6899   &  Mean   :0.301121  \\ \hline
+ 3rd Qu.:794.5   &   3rd Qu.:0.17198   &   3rd Qu.:25.8785    &  3rd Qu.:0.431055 \\ \hline  
+ Max.   :950.9     &  Max.   :0.32937    &  Max.   :30.9573   &   Max.   :0.981314  \\ \hline
+\end{tabular}
+\end{center}
+From: DepthData $<$- read.table(file="Three-Depths.dat", sep="", header=T); summary(DepthData)
+\end{table}
+
 \begin{figure}[H]
@@
 \end{figure}
+\pagebreak
+
 \begin{figure}[H]
@@
 \end{figure}
+
+%%% Gretl
+
+\begin{table}[htbp]
+\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Gretl.}
+\begin{center}
+Model 1: TSLS, using observations 1--191\\
+Dependent variable: const\\
+Instrumented: v17 \\
+Instruments: v12 \\
+
+\vspace{1em}
+
+\begin{tabular}{lr@{.}lr@{.}lr@{.}lr@{.}l}
+  &
+ \multicolumn{2}{c}{Coefficient} &
+  \multicolumn{2}{c}{Std.\ Error} &
+   \multicolumn{2}{c}{$z$} &
+    \multicolumn{2}{c}{p-value} \\[1ex]
+v17 &  6 & 88502 & 0 & 527503 & 13 & 0521 & 0 & 0000 \\
+\end{tabular}
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Mean dependent var &  1.000000 & S.D. dependent var &  0.000000 \\
+Sum squared resid &  97.51535 & S.E. of regression &  0.716407 \\
+Log-likelihood & $-$237.5799 & Akaike criterion &  477.1598 \\
+Schwarz criterion &  480.4121 & Hannan--Quinn &  478.4771 \\
+\end{tabular}
+
+\vspace{1em}
+\begin{raggedright}
+Hausman test --\\
+\quad Null hypothesis: OLS estimates are consistent\\
+\quad Asymptotic test statistic: $\chi^2(1)$ = 13.2834\\
+\quad with p-value = 0.000267768\\
+\vspace{1ex}
+Weak instrument test -- \\
+\quad First-stage $F(1, 190)$ = 16670.2 \\
+\end{raggedright}
+
+\end{center}
+\label{tab:14}
+\end{table}
+%%% Gretl
+
 \begin{figure}[H]
@@
 \end{center}
-\label{fig:5}
+\label{tab:14}
 \end{table}

### 2011-01-15 MScDiplom_Lemenkova.tex
added 16 more new tables: statistics using GRETL

@@
 \subsubsection{Global distribution of the seagrasses}
-Globally, there are 58 recognized and described seagrass species \ref{fig:2}, belonging to two orders
+Globally, there are 58 recognized and described seagrass species (Fig. \ref{fig:2}), belonging to two orders
 (Hydrocharitales and Najadales), four families (Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
@@
 \subsubsection{Statistical analysis of the observational data and hypothesis testing}
-The raw data of measurements of spectral reflectances of P.oceanica have been tested statistically. We
+In the statistical anlysis of the raw observed data pool of the values of spectral reflectances (about 400 single measurements for P.oceanica), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
+parameters.
+Examples are the mean and variance of a probability distribution of measurements dataset (Fig.).
+We divided the total data pool into datasets of 25 measurements for more convenience and ease of use, both for P.oceanica and carbonate sand.
+The mean of a probability distribution of spectral reflectance describes the average value of the random variable over all of its possible realizations (see for example, Fig.
+Appendix).
+Conceptually, there are an infinite number of such realizations, therefore parameters are not known to us.
+However, in the statistical analysis of the observed data our goal is to estimate these parameters using a finite amount of information available to us.
+Thus, we collect a number of realizations (a sample of 25 measurements) and then estimate the unknown parameters using a statistical analysis (Appendices, see Tables ). \\
+Finally we
 analysed 400 datasets made on two different data. 
+Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random. 
+Probability thus expresses uncertainty about the possible values of the values of spectral reflectances  of P.oceanica.
+Each time we observe the outcome of a random variable of spectral reflectance, we obtain an observation. 
+Once observed, its value is known and hence it is no longer random. 
+So, there is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
+We applied various methods of statistical probability analysis towards data pool, in order to analyse the distribution of values of spectral reflectances of P.oceanica (see Appendices, Tables). 
+A probability distribution, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
 The total amount of measured data was large and included following datasets made using
@@
 spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
-The statistical calculators were made by means of Open Office and included following computations
+The statistical calculators were mainly made by means of Gretle and SPSS. The Open Office was used for preliminary data view and preprocessing, and included following computations
 at each data set (table 1): mean, median, Q1 and Q3, standard deviation, min and max values. The
@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[scale=0.4]{G-gamma-rad.jpg}
+\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
+\label{fig:48}
+\end{center}
+\end{figure}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.4]{G-normal-dist.jpg}
+\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
+\label{fig:49}
+\end{figure}
+
+%%% Gretle start
+\begin{table}
+\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16. Gretl}
+\begin{center}
+Model 2: Logistic, using observations 1:01--8:23 ($T$ = 191)\\
+Dependent variable: v16\\
+$\hat{y} = 1 / (1 + e^{-X\hat{\beta}})$
+
+\vspace{1em}
+
+\begin{tabular}{|c|c|c|c|c|}
+  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
+const &  1.58324 &     0.407258 &       3.8876 &         0.0001 \\
+v1 &   $-$0.00512630 &     0.000615503 &       $-$8.3286 &         0.0000 \\
+\end{tabular}
+
+\vspace{1em}Statistics based on the transformed data:
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Sum squared resid &  462.8481 & S.E. of regression &  1.564906 \\
+$R^2$ &  0.268481 & Adjusted $R^2$ &  0.264610 \\
+$F(1, 189)$ &  69.36634 & P-value($F$) &  1.63\textrm{e--14} \\
+Log-likelihood & $-$355.5467 & Akaike criterion &  715.0935 \\
+Schwarz criterion &  721.5980 & Hannan--Quinn &  717.7281 \\
+$\hat{\rho}$ &  0.980662 & Durbin--Watson &  0.020351 \\
+\end{tabular}
+
+\vspace{1em}Statistics based on the original data:
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Mean dependent var &  0.262542 & S.D. dependent var &  0.205031 \\
+Sum squared resid &  10.80565 & S.E. of regression &  0.239108 \\
+\end{tabular}
+\end{center}
+\end{table}
+%%% end Gretl
+
+%%%
+\begin{table}
+\rowcolors{1}{Lavender}{LightYellow}
+\caption{Robust estimation of the seawater radiance (measured in aquarium tank, Heraklion): results of the \textit{least absolute deviation} (LAD), Series:V16. Gretl}
+\begin{center}
+Model 6: LAD, using observations 1:01--8:23 ($T$ = 191)\\
+Dependent variable: v16\\
+
+\vspace{1em}
+
+\begin{tabular}{|c c c c c|}
+  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
+const &   0.738911 &     0.0862011 &       8.5719 &         0.0000 \\
+v1 &   $-$0.000775186 &     0.000114010 &       $-$6.7993 &         0.0000 \\
+\end{tabular}
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Median depend. var &  0.227857 & S.D. dependent var &  0.205031 \\
+Sum absolute resid &  30.22342 & Sum squared resid &  8.059632 \\
+Log-likelihood &  28.74727 & Akaike criterion & $-$53.49454 \\
+Schwarz criterion & $-$46.99000 & Hannan--Quinn & $-$50.85990 \\
+\end{tabular}
+\end{center}
+\end{table}
+%%%
+\pagebreak
+\newpage
+
 \begin{table}[htbp]
@@
 \includegraphics[scale=0.25]{GNU-09.jpg}
-\caption{Spectral reflectance of seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle (GNUplot)­}
+\caption{Spectral reflectance of seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle (Gnuplot)­}
 \label{fig:52}
@@
 \end{figure}
+
+%%% Gretl start
+\begin{table}
+\rowcolors{1}{Bisque}{Ivory}
+\begin{center}
+\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Gretl.}
+Model 2: OLS, using observations 326--350\\
+Dependent variable: v20 \\
+\vspace{1em}
+\begin{tabular}{|c c c c c|}
+  & {Coefficient} &  {Std.\ Error} &   {$t$-ratio} &    {p-value} \\[1ex]
+const &  0.414957 &    0.0299342 &      13.8623 &        0.0000 \\
+v1 &   $-$0.000441803 &     4.52185\textrm{e--005} &       $-$9.7704 &         0.0000 \\ \hline\hline
+\end{tabular}
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Mean dependent var &  0.134005 & S.D. dependent var &  0.140653 \\
+Sum squared resid &  2.497416 & S.E. of regression &  0.114951 \\
+$R^2$ &  0.335585 & Adjusted $R^2$ &  0.332070 \\
+$F(1, 189)$ &  95.46076 & P-value($F$) &  1.65\textrm{e--18} \\
+Log-likelihood &  143.1678 & Akaike criterion & $-$282.3357 \\
+Schwarz criterion & $-$275.8311 & Hannan--Quinn & $-$279.7010 \\
+\end{tabular}
+\end{center}
+\end{table}
+\pagebreak
+%%% Gretl end
+
 \begin{figure}[h]
 \begin{center}
-\includegraphics[scale=0.25]{GNU-10.jpg}
-\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. GNUplot display­}
+\includegraphics[scale=0.22]{GNU-10.jpg}
+\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. Gnuplot display­}
 \label{fig:53}
@@
 \end{figure}
+
+%%% Gretl start
+\begin{table}[H]
+\rowcolors{1}{Thistle}{MintCream}
+\begin{center}
+\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. Gretl.­}
+Summary Statistics, using the observations 301--325\\
+\vspace{8pt}
+\begin{tabular}{|c|c|c|c|c|}
+Variable & {Mean}& {Median}  & {Minimum}    & {Maximum} \\[1ex]
+v12 & 0.221475 & 0.139032 & 0.00101700 & 0.780730\\
+v13 & 0.217074 & 0.139285 & 0.00108700 & 0.754430\\
+v14 & 0.167442 & 0.110878 & 0.000348000 & 0.575242\\
+v15 & 0.155908 & 0.111737 & 0.0125650 & 0.483602\\
+v16 & 0.158329 & 0.111647 & 0.0101590 & 0.498096\\
+v17 & 0.167575 & 0.111349 & 0.000836000 & 0.579004\\
+v18 & 0.125868 & 0.0863670 & 0.000662000 & 0.417701\\
+v19 & 0.179261 & 0.117898 & 0.000935000 & 0.612021\\
+v20 & 0.141302 & 0.0932400 & 0.000714000 & 0.478343\\
+v21 & 0.228822 & 0.143105 & 0.00149500 & 0.800529\\
+v22 & 0.174892 & 0.114994 & 0.000343000 & 0.592146\\ \hline\hline
+\end{tabular}
+\begin{tabular}{|c|c|c|c|c|}
+\rowcolors{1}{Lavender}{MintCream}
+Variable &  {Std.\ Dev.}  & {Coef.Var.}   & {Skewness}    & {Ex.\ kurtosis} \\[1ex]
+v12 & 0.246134 & 1.11134 & 0.930598 & $-$0.393370\\
+v13 & 0.239324 & 1.10250 & 0.910879 & $-$0.440017\\
+v14 & 0.181797 & 1.08573 & 0.884089 & $-$0.466508\\
+v15 & 0.151811 & 0.973718 & 0.796534 & $-$0.646866\\
+v16 & 0.155834 & 0.984241 & 0.832126 & $-$0.587890\\
+v17 & 0.182383 & 1.08837 & 0.901570 & $-$0.435120\\
+v18 & 0.134253 & 1.06662 & 0.820573 & $-$0.603688\\
+v19 & 0.195217 & 1.08901 & 0.873615 & $-$0.515800\\
+v20 & 0.152787 & 1.08128 & 0.863700 & $-$0.525357\\
+v21 & 0.252696 & 1.10434 & 0.924172 & $-$0.408867\\
+v22 & 0.189797 & 1.08523 & 0.858809 & $-$0.546215\\
+\end{tabular}
+\end{center}
+\end{table}
+%%% Gretle end
+
 \pagebreak
@@
 
-%%% Gretl
-
+%%% Gretl start
 \begin{table}[htbp]
@@
 
-\begin{tabular}{lr@{.}lr@{.}lr@{.}lr@{.}l}
-  &
- \multicolumn{2}{c}{Coefficient} &
-  \multicolumn{2}{c}{Std.\ Error} &
-   \multicolumn{2}{c}{$z$} &
-    \multicolumn{2}{c}{p-value} \\[1ex]
-v17 &  6 & 88502 & 0 & 527503 & 13 & 0521 & 0 & 0000 \\
+\begin{tabular}{|c c c c c|}
+  &  {Coefficient} &  {Std.\ Error} &  {$z$} &    {p-value} \\[1ex]
+v17 &  6.88502 & 0.527503 & 13.0521 & 0.0000 \\
 \end{tabular}
@@
 \end{table}
-%%% Gretl
+%%% Gretl end
 
@@
 \end{figure}
+
+%%% Gretl start
+\begin{table}[htbp]
+\rowcolors{1}{Ivory}{GhostWhite}
+\caption{Basic mathematical formulae used for statistical analysis of the measurement set of the spectral reflectance.}
+\label{tab:13}
+\begin{center}
+\begin{tabular}{|c|c|}
+\hline\hline
+Basic Statistics & Math Formulae \\ \hline\hline
+\hline
+Mean & $\Sigma(x_i/n=\bar x)$ \\ \hline
+Variance & $1 \over n-1$ $\Sigma(x_i-\bar x)=s^2_x$  \\ \hline
+Standard Deviation & $s=s^2$ \\ \hline
+Coefficient of Variation & $s/\bar x$ \\ \hline
+Skewness & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^3$  \\ \hline
+Excess Kurtosis & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^4-3$  \\ \hline
+
+\end{tabular}
+\end{center}
+\label{tab:14}
+\end{table}
+%%% Gretl end
+
 \pagebreak
@@
 
+%%%Gretl begin
+\begin{table}[H]
+\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Gretl.}
+\begin{center}
+Eigenanalysis of the Correlation Matrix\\
+\vspace{1em}
+
+\begin{tabular}{|c|c|c|c|}
+{Component } &  {Eigenvalue } & {Proportion} & {Cumulative} \\
+1 &   6.4848    &   0.9264    &   0.9264 \\
+2 & 0.4949 &      0.0707  &     0.9971 \\
+3 &  0.0140    &   0.0020    &   0.9991 \\
+4 &    0.0052  &     0.0007    &   0.9998 \\
+5 &   0.0008   &    0.0001     &  1.0000 \\
+6 &  0.0002    &   0.0000      & 1.0000 \\
+7 &   0.0001   &    0.0000     &  1.0000 \\
+\end{tabular}
+
+Eigenvectors (component loadings)
+\vspace{1ex}
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+v1        &	     -0.308  &	  -0.880  &	   0.334   &	 -0.126  &	   0.033  &	  -0.042   &	  0.006\\ \hline
+v2      &	      0.383  &	  -0.295  &	  -0.512   &	 -0.230  &	  -0.365  &	  -0.370   &	 -0.425\\ \hline
+v3      &	      0.383  &	  -0.300  &	  -0.465   &	 -0.010  &	   0.562  &	   0.307   &	  0.370\\ \hline
+v4      &	      0.390  &	  -0.145  &	   0.203   &	  0.714  &	  -0.129  &	  -0.415   &	  0.293\\ \hline
+v5      &	      0.392  &	  -0.091  &	   0.300   &	  0.262  &	   0.020  &	   0.568   &	 -0.597\\ \hline
+v6      &	      0.392  &	  -0.011  &	   0.259   &	 -0.410  &	  -0.543  &	   0.299   &	  0.476\\ \hline
+v7      &	      0.390  &	   0.138  &	   0.460   &	 -0.429  &	   0.488  &	  -0.427   &	 -0.113\\ \hline
+\end{tabular}
+\end{center}
+\label{tab:14}
+\end{table}
+%%%Gretl end
+\pagebreak
+
+%%%Gretl begin
+\begin{table}[H]
+\begin{center}
+\rowcolors{1}{AntiqueWhite}{LightGoldenrodYellow}
+\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textbf{P.oceanica}. Gretl.}
+Model 2: Tobit, using observations 1:01--8:23 ($T$ = 191)\\
+Dependent variable: v12\\
+\vspace{1em}
+\begin{tabular}{|c c c c c|}
+  &  {Coefficient} &    {Std.\ Error} & {$z$} & {p-value} \\[1ex]
+const &   $-$0.00413795 &     0.0502867 &     $-$0.0823 &         0.9344 \\
+v2 &   0.890117 &     0.957730 &       0.9294 &         0.3527 \\
+\end{tabular}
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Mean dependent var &  0.035372 & S.D. dependent var &  0.053185 \\
+Censored obs &         1 & sigma &  0.050920 \\
+Log-likelihood &  295.1357 & Akaike criterion & $-$584.2715 \\
+Schwarz criterion & $-$574.5147 & Hannan--Quinn & $-$580.3195 \\
+$\hat{\rho}$ & $-$0.032865 & Durbin--Watson &  2.065699 \\
+\end{tabular}
+
+\vspace{1em}
+\begin{raggedright}
+Test for normality of residual --\\
+\quad Null hypothesis: error is normally distributed\\
+\quad Test statistic: $\chi^2(2)$ = 190.404\\
+\quad with p-value = 4.51226e-042\\
+\vspace{1ex}
+\end{raggedright}
+\end{center}
+\end{table}
+%%%Gretl end
 \begin{figure}[H]
 \begin{center}
-\includegraphics[scale=0.4]{GNU-15.jpg}
+\includegraphics[scale=0.38]{GNU-15.jpg}
 \caption{Multiplot of spectral reflectance of P.oceanica. Series 151-200. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
@@
 \pagebreak
+
+%%%Gretl begin
+\begin{table}[H]
+\rowcolors{1}{Gainsboro}{MintCream}
+\begin{center}
+\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica}. Gretl.}
+Model 1: WLS, using observations 1--95\\
+Dependent variable: v20\\
+Variable used as weight: v12
+
+\vspace{1em}
+\begin{tabular}{|c|c|c|c|c|}
+  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
+const &   0.239525 &     0.0771338 & 3.1053 &         0.0025 \\
+v1 &   0.000120809 &     0.000147131 &       0.8211 &         0.4137 \\
+\end{tabular}
+
+\vspace{1em}Statistics based on the weighted data:
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Sum squared resid &  0.290924 & S.E. of regression &  0.055930 \\
+$R^2$ &  0.007197 & Adjusted $R^2$ & -0.003478 \\
+$F(1, 93)$ &  0.674200 & P-value($F$) &  0.413691 \\
+Log-likelihood &  140.1579 & Akaike criterion & $-$276.3158 \\
+Schwarz criterion & $-$271.2081 & Hannan--Quinn & $-$274.2519 \\
+\end{tabular}
+
+\vspace{1em}Statistics based on the original data:
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Mean dependent var &  0.136154 & S.D. dependent var &  0.149193 \\
+Sum squared resid &  5.689669 & S.E. of regression &  0.247344 \\
+\end{tabular}
+\end{center}
+\end{table}
+%%%Gretl end
+
+%%% one more Gretl start
+\begin{table}
+\rowcolors{1}{BlanchedAlmond}{MintCream}
+\begin{center}
+\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{P.oceanica}. Gretl.}
+Model 3: Prais--Winsten, using observations 1:01--8:23 ($T$ = 191)\\
+Dependent variable: v12\\
+$\hat{\rho}$ = -0.0343656
+
+\vspace{1em}
+
+\begin{tabular}{|c c c c c|}
+  & {Coefficient} &  {Std.\ Error} & {$t$-ratio} &   {p-value} \\[1ex]
+const &   $-$0.00498439 &     0.00977709 &       $-$0.5098 &         0.6108 \\
+v2 &   0.912742 &     0.205904 &      4.4328 &         0.0000 \\
+\end{tabular}
+
+\vspace{1em}Statistics based on the rho-differenced data:
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Mean dependent var &  0.035372 & S.D. dependent var &  0.053185 \\
+Sum squared resid &  0.490951 & S.E. of regression &  0.050967 \\
+$R^2$ &  0.086511 & Adjusted $R^2$ &  0.081678 \\
+$F(1, 189)$ &  19.67163 & P-value($F$) &  0.000016 \\
+$\hat{\rho}$ & $-$0.001888 & Durbin--Watson &  2.003741 \\
+\end{tabular}
+\end{center}
+\end{table}
+%%%one more Gretl end
+\pagebreak
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[scale=0.4]{G-autocorr-sand51-75.jpg}
+\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand. Series 51-75, variable 27. Gretl visualization­}
+\label{fig:57}
+\end{center}
+\end{figure}
+
+%%% Gretl start
+\begin{table}[htbp]
+\rowcolors{1}{Ivory}{GhostWhite}
+\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
+\label{tab:14}
+\begin{center}
+\begin{tabular}{|c c c c c c c||}
+\hline\hline
+LAG & ACF & & PACF & & Q-stat & [p-value] \\ \\ \hline\hline
+1  & 0.9851 & ***  & 0.9851& ***   & 104.8293 & [0.000]\\ \hline
+    2  & 0.9657  &***  &-0.1598  &   ***  &  206.5423  &[0.000]\\ \hline
+    3  & 0.9420  &***  &-0.1337  &  ***   &  304.2787  &[0.000]\\ \hline
+    4  & 0.9144  &***  &-0.1107  &  ***   &  397.2949  &[0.000]\\ \hline
+    5  & 0.8835  &***  &-0.0905  &  ***   &  484.9833  &[0.000]\\ \hline
+    6  & 0.8495 & ***  &-0.0732  &  ***   &  566.8791  &[0.000]\\ \hline
+    7  & 0.8131 & ***  &-0.0565  &  ***   &  642.6707  &[0.000]\\ \hline
+    8  & 0.7748 & ***  &-0.0423  &  ***   &  712.1933  &[0.000]\\ \hline
+    9  & 0.7350 & ***  &-0.0317  &  ***   &  775.4114  &[0.000]\\ \hline
+   10 &  0.6941&  *** & -0.0280 &  ***   &   832.3807 & [0.000]\\ \hline
+   11 &  0.6524 & *** & -0.0221 &  ***   &   883.2508 & [0.000]\\ \hline
+   12 &  0.6103 & *** & -0.0197 &  ***   &   928.2405 & [0.000]\\ \hline
+   13 &  0.5679 & *** & -0.0170 &  ***   &   967.6289 & [0.000]\\ \hline
+   14 &  0.5256 & *** & -0.0162 &  ***   &  1001.7381&  [0.000]\\ \hline
+   15 &  0.4835 & *** & -0.0154 &  ***   &  1030.9237 & [0.000]\\ \hline
+   16 & 0.4418 & *** & -0.0172  &  ***  &  1055.5584 & [0.000]\\ \hline
+   17  & 0.4v004 & *** & -0.0196 &***     &  1076.0231 & [0.000]\\ \hline
+   18 & 0.35v94 & *** & -0.0216  & ***   &  1092.7041 & [0.000]\\ \hline
+   19 &  0.3189 & *** & -0.0224   & ***  &  1105.9911 & [0.000]\\ \hline
+   20 &  0.2788 & *** & -0.0269   & ***  &  1116.2679 & [0.000]\\ \hline
+\end{tabular}
+\end{center}
+\label{fig:5}
+\end{table}
+%%% Gretl end
+\pagebreak
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[scale=0.4]{G-QQplot-sand.jpg}
+\caption{Normal Q-Q plot: estimated versus observed values of the measurement of carbonate sand, variable 27. Series 51-75. Gretl modelling visualization­}
+\label{fig:57}
+\end{center}
+\end{figure}
+
+%%% Gretl start
+\begin{table}[H]
+\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
+\begin{center}
+Model 2: Quantile estimates, using observations 1:01--5:09 ($T$ = 105)\\
+Dependent variable: v27\\
+$\tau$ = 0.5\\
+Asymptotic standard errors assuming IID errors\\
+
+\vspace{1em}
+
+\begin{tabular}{c c c c c}
+***   & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
+const &  0.837827 &    0.0386417 &      21.6820 &         0.0000 \\
+v1 &   $-$0.00111120 &     6.60949\textrm{e--005} &       $-$16.8122 &         0.0000 \\ \hline \hline
+\end{tabular}
+
+\vspace{1ex}
+\begin{tabular}{lrlr}
+Median depend. var &  0.191024 & S.D. dependent var &  0.135439 \\
+Sum absolute resid &  8.625039 & Sum squared resid &  1.370360 \\
+Log-likelihood &  84.64508 & Akaike criterion & $-$165.2902 \\
+Schwarz criterion & $-$159.9822 & Hannan--Quinn & $-$163.1393 \\ \hline \hline
+\end{tabular}
+\end{center}
+\end{table}
+%%% Gretl end
+\pagebreak
+
+%%%Gretl begin
+\begin{table}[H]
+\begin{small}
+\rowcolors{1}{Ivory}{LightGoldenrodYellow}
+\caption{Results of the \textit{periodogramm} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
+\label{tab:14}
+\begin{center}
+\begin{tabular}{|p{15mm} p{15mm} p{15mm} p{15mm} || p{15mm} p{15mm}p{15mm} p{15mm}|}
+omega & scaled frequency & periods & log spectral density & omega & scaled frequency & periods & log spectral density \\ \\ \hline\hline
+0.05984      & 1    &      105.00  &      -1.2466  &  1.61568   &   27    &        3.89     &    -11.111\\ 
+ 0.11968     &  2   &        52.50   &      -5.8761 &   1.67552   &   28    &        3.75     &    -11.170\\ 
+ 0.17952     &  3   &        35.00   &      -4.0657 &  1.73536   &   29    &        3.62     &    -11.208\\ 
+ 0.23936     &  4   &        26.25   &      -7.1031 &  1.79520   &   30    &        3.50     &    -11.461\\ 
+ 0.29920     &  5   &        21.00   &      -6.1742 &  1.85504   &   31    &        3.39     &    -11.380\\ 
+ 0.35904     &  6   &        17.50   &      -7.2772  & 1.91488    & 32     &       3.28      &   -11.686\\ 
+ 0.41888     &  7     &      15.00   &      -8.7802 &  1.97472    &  33    &        3.18     &    -11.624\\ 
+ 0.47872     &  8     &      13.13   &      -7.6806 & 2.03456    &  34    &        3.09         &-11.724\\ 
+ 0.53856     &  9     &      11.67   &      -8.4848  & 2.09440    &  35    &        3.00    &     -11.818\\ 
+ 0.59840     & 10    &       10.50  &       -8.5764 &    2.15423    &  36    &        2.92    &     -11.805\\ 
+ 0.65824     & 11    &        9.55   &      -9.4643 &  2.21407    &  37    &        2.84    &     -12.002\\ 
+ 0.71808     & 12    &        8.75   &      -9.0868 &  2.27391    &  38    &        2.76    &     -11.930\\ 
+ 0.77792     & 13    &        8.08   &      -9.3608 &  2.33375    &  39    &        2.69    &     -12.010\\ 
+ 0.83776     & 14    &        7.50     &    -9.9332 &  2.39359    &  40    &        2.63    &     -12.124\\ 
+ 0.89760     & 15    &        7.00     &    -9.6836 &  2.45343    &  41    &        2.56    &     -12.190\\ 
+ 0.95744     & 16    &        6.56     &    -10.00  & 2.51327    &  42           & 2.50    &     -12.248\\ 
+ 1.01728     & 17    &        6.18     &    -9.8994 & 2.57311    &  43      &      2.44    &     -12.306\\ 
+ 1.07712     & 18    &        5.83     &    -10.188 & 2.63295    &  44      &      2.39    &     -12.323\\ 
+ 1.13696     & 19    &        5.53     &    -10.197 & 2.69279    &  45      &      2.33    &     -12.353\\ 
+ 1.19680     & 20    &        5.25     &    -10.467 &  2.75263    &  46      &      2.28    &     -12.420\\ 
+ 1.25664     & 21    &        5.00     &    -10.505 &  2.81247    &  47      &      2.23    &     -12.382\\ 
+ 1.31648     & 22    &        4.77    &    -10.641 &  2.87231    &  48      &      2.19    &     -12.487\\ 
+ 1.37632     & 23    &        4.57     &    -10.906 &  2.93215    &  49      &    2.14      &   -12.433\\ 
+ 1.43616     & 24    &        4.38     &    -10.957 &  2.99199    &  50        &    2.10    &     -12.609\\ 
+ 1.49600      &25    &        4.20     &    -10.785 &   3.05183    &  51      &      2.06    &     -12.410\\ 
+ 3.11167    &  52      &      2.02    &     -12.477 & & & & \\   \hline \hline
+\end{tabular}
+\end{center}
+\end{small}
+Number of observations = 105
+\label{fig:5}
+\end{table}
+%%% Gretl end
+\begin{figure}[H]
+\begin{center}
+\includegraphics[scale=0.25]{G-period-sand30.jpg}
+\caption{Plot illustrating periodogramm for v30 from measurements of carbonate sand, series 51-75. Gretl modelling visualization­}
+\label{fig:57}
+\end{center}
+\end{figure}
+\pagebreak
+
+%%%Begin Plot Gretl
+\begin{figure}[H]
+\begin{center}
+\includegraphics[scale=0.32]{G-polynom-trend-wo.jpg}
+\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
+\label{fig:57}
+\end{center}
+\end{figure}
+\begin{figure}[H]
+\begin{center}
+\includegraphics[scale=0.32]{G-polynom-trend-with.jpg}
+\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
+\label{fig:57}
+\end{center}
+\end{figure}
+\begin{figure}[H]
+\begin{center}
+\includegraphics[scale=0.32]{G-exp-mov-aver.jpg}
+\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
+\label{fig:57}
+\end{center}
+\end{figure}
+\pagebreak
+%%%end Plot Gretl
 

### 2011-01-16 MScDiplom_Lemenkova.tex
formatted LaTeX document

@@
 
+\vspace{5em}
+
 \textbf{Software and utilities used in the current work}:\\
@@
 	\includegraphics[scale=0.20]{Fig-1-4.jpg}
-\caption{Distribution of seagrass in relation to mean
-ocean temperature. \\
+	\caption{Distribution of seagrass in relation to mean ocean temperature. \\
 	Source: \cite{Orth06}\label{Orth06}}
@@
 	\includegraphics[scale=0.15]{Fig-1-7.jpg}
-\caption{Scheme of matte structure of
-P.Oceanica.\\ Source \cite{Pergent90}\label{Pergent90}}
+	\caption{Scheme of matte structure of \textit{	P.Oceanica}.\\ Source \cite{Pergent90}\label{Pergent90}}
 	\label{fig:7}
@@
 	\includegraphics[scale=0.15]{Fig-1-8.jpg}
-\caption{Structure and components of
-P.oceanica \\ Source \cite{DiCarlo04}\label{DiCarlo04}}
+	\caption{Structure and components of P.oceanica \\ Source \cite{DiCarlo04}\label{DiCarlo04}}
 	\label{fig:8}
@@
 	\includegraphics[scale=0.25]{Fig-1-9.jpg}
-\caption{Spectra of the seagrass on different
-depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}}
+	\caption{Spectra of the seagrass on different depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}}
 	\label{fig:9}
@@
 	\includegraphics[scale=0.30]{Fig-11.jpg}
-\caption{In situ optical reflectance spectra of
-seagrass. Shaded areas - percent of spectra lying within
-the range of reflectance. White lines - mean spectra.
-Source: \cite{Hochberg03b}\label{Hochberg03b}}
+	\caption{In situ optical reflectance spectra of seagrass. Shaded areas - percent of spectra lying within the range of reflectance. 
+		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
 	\label{fig:11}
@@
 	\includegraphics[scale=0.25]{Fig-12.jpg}
-\caption{Difference between broaddand
-multispectral and hyperspectral resolution of
+	\caption{Difference between broaddand multispectral and hyperspectral resolution of
 		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
@@
 	\includegraphics[scale=0.35]{Fig-13.jpg}
-\caption{Band coverage of ASTER and Landsat channels on
-the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
+ 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
 	\label{fig:13}
@@
 	\includegraphics[scale=0.30]{Fig-14.jpg}
-\caption{Multispectral vs. hyperspectral band
-coverage. Source: \cite{Sabins97}\label{Sabins97}}
+	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
 	\label{fig:14}
@@
 	\includegraphics[scale=0.15]{Fig-15.jpg}
-\caption{Reflectance spectra of seagrass; Rrs(0-)
-- subsurface RS reflectance; Rrs b - the bottom
+	\caption{Reflectance spectra of seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
 		reflectance. Source:\cite{Yang10}\label{Yang10}}
@@
 and up-to-date source of imagery.
+
 \begin{wrapfigure}{r}{0.5\textwidth}
@@
 shows the midspread of the statistical quartiles (Q1-Q3), mean and extreme values.
+
 \begin{figure}[h]
@@
 	\begin{center}
-%\rowcolors{1}{lightgray}{white}
 	\begin{tabular}{|c|c|c|c|c|c|c|}
@@
 	\end{center}
-\label{fig:5}
+	\label{tab:2}
 \end{table}
@@
 	\includegraphics[scale=0.30]{Fig-30.jpg}
-\caption{Main interface of WASI water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-
-4.0m}
+	\caption{Main interface of WASI water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m}
 	\label{fig:30}
@@
 \end{figure}
+
 \paragraph{Model parameters: depth and bottom cover fraction}
@@
 	\centering
-%\rowcolors{1}{lightgray}{white}
-% \scalebox{0.7}
 	 \begin{tabular}{|  p{2cm} | p{6cm} | p{25mm} | }
@@
 oceanica (Fig.\ref{fig:62}).
-
 The graph is received in excel spreadsheet using mean values of spectral reflectances of
@@
 Thus, we collect a number of realizations (a sample of 25 measurements) and then estimate the unknown parameters using a statistical analysis (Appendices, see Tables ). \\
-Finally we
-analysed 400 datasets made on two different data. 
+Finally we analysed 400 datasets made on two different data. 
 Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random. 
@@
 Ligaria beach, Agia Pelagia.
+
 \begin{figure}
@@
 SPSS}
-\label{tab:5}
 	\begin{center}
@@
 	\end{center}
-the independent variable is wavelength. 
-\label{fig:5}
+	The independent variable is wavelength. 
+	\label{tab:5}
 \end{table}
@@
 	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{P.oceanica}. SPSS}
-\label{tab:6}
 	\begin{center}
@@
 	\end{center}
-the independent variable is wavelength. 
-\label{fig:5}
+	The independent variable is wavelength. 
+	\label{tab:6}
 \end{table}
@@
 	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. SPSS}
-\label{tab:7}
 	\begin{center}
@@
 	\end{center}
-a. The underlying process assumed is independence (white noise).\\
-b. Based on the asymptotic chi-square approximation.
-\label{fig:5}
+	The underlying process assumed is independence (white noise).\\
+	 Based on the asymptotic chi-square approximation.
+	\label{tab:7}
 \end{table}
@@
 	\includegraphics[scale=0.25]{GNU-Radiance_Bezier.jpg}
-\caption{Radiance of the seawater with sediments. Measured in 
-aquarium tank. Bezier Interpolation}
+	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation}
 	\label{fig:47}
 \end{figure}
-
 \begin{table}[htbp]
@@
 	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 0.5, 		1.5 and 2.5 meters}
-\label{tab:8}
 	\begin{center}
@@
 	P more than .05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
-\label{fig:5}
+	\label{tab:8}
 \end{table}
-
 \begin{table}[htbp]
@@
 	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters}
-\label{tab:9}
 	\begin{center}
@@
 	\end{center}
-P>0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
-\label{fig:5}
+	P$>$0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5). 
+ 	\label{tab:9}
 \end{table}
@@
 \begin{longtable}{|c|c|c|c|c|c|c|}
-\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October} 
-\label{tab:10} \\
-\hline
+	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 	\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
+\label{tab:10} 
 \end{longtable}
@@
 \newpage
-
 \begin{figure}[h]
@@
 		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
-\label{fig:48}
+		\label{fig:50}
 	\end{center}
@@
 	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
-\label{fig:49}
+	\label{fig:51}
 \end{figure}
@@
 	\vspace{1em}
-
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{center}
+	\label{tab:11}
 \end{table}
@@
 		Dependent variable: v16\\
-
 	\vspace{1em}
@@
 	\end{tabular}
+	\label{tab:12}
 	\end{center}
@@
 	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
-\label{tab:11}
 	\begin{center}
@@
 	\end{center}
-a. The underlying process assumed is independence (white noise).\\
-b. Based on the asymptotic chi-square approximation.
-\label{fig:5}
+	The underlying process assumed is independence (white noise).\\
+	Based on the asymptotic chi-square approximation.
+	\label{tab:13}
 \end{table}
@@
 	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
-\label{fig:50}
+	\label{fig:52}
 \end{figure}
@@
 	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
-\label{tab:12}
 	\begin{center}
@@
 	\end{tabular}
+	\label{tab:14}
 	\end{center}
@@
 	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
-\label{fig:51}
+	\label{fig:53}
 \end{figure}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 	\begin{center}
 		\includegraphics[scale=0.25]{GNU-09.jpg}
-\caption{Spectral reflectance of seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle (Gnuplot)­}
-\label{fig:52}
+		\caption{Spectral reflectance of the seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle (Gnuplot)­}
+		\label{fig:54}
 	\end{center}
@@
 %%% Gretl start
-\begin{table}
+\begin{table}[H]
 	\rowcolors{1}{Bisque}{Ivory}
@@
 		Dependent variable: v20 \\
+
 		\vspace{1em}
-\begin{tabular}{|c c c c c|}
+		\begin{tabular}{c c c c c}
 		  & {Coefficient} &  {Std.\ Error} &   {$t$-ratio} &    {p-value} \\[1ex]
@@
 	\end{tabular}
-\vspace{1ex}
+
+	\vspace{2em}
 	\begin{tabular}{lrlr}
@@
 	\end{tabular}
+	\label{tab:15}
 	\end{center}
@@
 		\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. Gnuplot display­}
-\label{fig:53}
+		\label{fig:55}
 	\end{center}
@@
 		Summary Statistics, using the observations 301--325\\
-\vspace{8pt}
+
+	\vspace{2em}
+
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{center}
+	\label{tab:16}
 \end{table}
@@
 	\caption{Summary of the table with values of the spectral reflectance of \textbf{P.oceanica} measured on three different depths, R}
-\label{tab:13}
 	\begin{center}
@@
 		From: DepthData $<$- read.table(file="Three-Depths.dat", sep="", header=T); summary(DepthData)
+	\label{tab:17}
 \end{table}
@@
 		\caption{Spectral reflectance of P.oceanica. Measurement series 25-50. GNUplot display­}
-\label{fig:54}
+		\label{fig:56}
 	\end{center}
@@
 		\includegraphics[scale=0.25]{GNU-12.jpg}
-\caption{RS reflectance of P.oceanica. Series 1-25. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
-mean value within the range.­}
-\label{fig:55}
+		\caption{RS reflectance of P.oceanica. Series 1-25. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range.­}
+		\label{fig:57}
 	\end{center}
@@
 	\end{center}
-\label{tab:14}
+	\label{tab:18}
 \end{table}
@@
 		\includegraphics[scale=0.4]{GNU-14.jpg}
-\caption{Multiplot of spectral reflectance of P.oceanica. Series 126-150. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
-mean value within the range (red bold dots). GNUplot visualization­}
-\label{fig:56}
+		\caption{Multiplot of spectral reflectance of P.oceanica.
+Series 126-150.
+Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
+GNUplot visualization­}
+		\label{fig:58}
 	\end{center}
@@
 		\hline
+
+		\vspace{1ex}
+
 		Mean & $\Sigma(x_i/n=\bar x)$ \\ \hline
@@
 		Excess Kurtosis & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^4-3$  \\ \hline
-
-\end{tabular}
-\end{center}
-\label{tab:14}
+	\end{tabular}
+	\end{center}
+	\label{tab:19}
 \end{table}
 %%% Gretl end
-
 \pagebreak
@@
 		& \multicolumn{2}{|c|}{Cluster} &  & \multicolumn{2}{|c|}{Cluster}\\
-%\cline{2,3,5,6}
 		 & 1 & 2 &  & 1 & 2 \\ \hline\hline
@@
 	\end{center}
-\label{tab:14}
+	\label{tab:20}
 \end{table}
@@
 	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
-\label{tab:14}
 	\begin{center}
@@
 	\end{center}
-a.
 	Convergence achieved due to no or small change in cluster centers.
@@
 The minimum distance between initial centers is 1.435.
-\label{fig:5}
+\label{tab:21}
 \end{table}
@@
 	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
-\label{tab:15}
 		\begin{center}
@@
 		\end{center}
-\label{fig:5}
+	\label{tab:22}
 \end{table}
@@
 	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
-\label{tab:16}
 	\begin{center}
@@
 	\end{center}
-\label{fig:5}
+	\label{tab:23}
 \end{table}
@@
 	Eigenvectors (component loadings)
+
 	\vspace{1ex}
+
 		\begin{tabular}{|c|c|c|c|c|c|c|c|}
@@
 	\end{center}
-\label{tab:14}
+	\label{tab:24}
 \end{table}
@@
 	\end{raggedright}
+	\label{tab:25}
 	\end{center}
@@
 %%%Gretl end
+
 \begin{figure}[H]
@@
 		\includegraphics[scale=0.38]{GNU-15.jpg}
-\caption{Multiplot of spectral reflectance of P.oceanica. Series 151-200. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
-mean value within the range (red bold dots). GNUplot visualization­}
-\label{fig:57}
-\end{center}
-\end{figure}
-\pagebreak
+		\caption{Multiplot of spectral reflectance of P.oceanica.
+Series 151-200.
+Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
+GNUplot visualization­}
+		\label{fig:59}
+	\end{center}
+\end{figure}
+\pagebreak
+
+%%% one more Gretl start
+\begin{table}[H]
+	\rowcolors{1}{BlanchedAlmond}{MintCream}
+	\begin{center}
+	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{P.oceanica}. Gretl.}
+		Model 3: Prais--Winsten, using observations 1:01--8:23 ($T$ = 191)\\
+		Dependent variable: v12\\
+	$\hat{\rho}$ = -0.0343656
+
+	\vspace{1em}
+
+	\begin{tabular}{|c c c c c|}
+		  & {Coefficient} &  {Std.\ Error} & {$t$-ratio} &   {p-value} \\[1ex]
+		const &   $-$0.00498439 &     0.00977709 &       $-$0.5098 &         0.6108 \\
+		v2 &   0.912742 &     0.205904 &      4.4328 &         0.0000 \\
+	\end{tabular}
+
+	\vspace{1em}Statistics based on the rho-differenced data:
+
+	\vspace{1ex}
+	\begin{tabular}{lrlr}
+		Mean dependent var &  0.035372 & S.D. dependent var &  0.053185 \\
+		Sum squared resid &  0.490951 & S.E. of regression &  0.050967 \\
+		$R^2$ &  0.086511 & Adjusted $R^2$ &  0.081678 \\
+		$F(1, 189)$ &  19.67163 & P-value($F$) &  0.000016 \\
+		$\hat{\rho}$ & $-$0.001888 & Durbin--Watson &  2.003741 \\
+	\end{tabular}
+	\end{center}
+	\label{tab:26}
+\end{table}
+%%%one more Gretl end
 
@@
 	\end{tabular}
+	\label{tab:27}
 	\end{center}
@@
 %%%Gretl end
-
-%%% one more Gretl start
-\begin{table}
-\rowcolors{1}{BlanchedAlmond}{MintCream}
-\begin{center}
-\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{P.oceanica}. Gretl.}
-Model 3: Prais--Winsten, using observations 1:01--8:23 ($T$ = 191)\\
-Dependent variable: v12\\
-$\hat{\rho}$ = -0.0343656
-
-\vspace{1em}
-
-\begin{tabular}{|c c c c c|}
-  & {Coefficient} &  {Std.\ Error} & {$t$-ratio} &   {p-value} \\[1ex]
-const &   $-$0.00498439 &     0.00977709 &       $-$0.5098 &         0.6108 \\
-v2 &   0.912742 &     0.205904 &      4.4328 &         0.0000 \\
-\end{tabular}
-
-\vspace{1em}Statistics based on the rho-differenced data:
-
-\vspace{1ex}
-\begin{tabular}{lrlr}
-Mean dependent var &  0.035372 & S.D. dependent var &  0.053185 \\
-Sum squared resid &  0.490951 & S.E. of regression &  0.050967 \\
-$R^2$ &  0.086511 & Adjusted $R^2$ &  0.081678 \\
-$F(1, 189)$ &  19.67163 & P-value($F$) &  0.000016 \\
-$\hat{\rho}$ & $-$0.001888 & Durbin--Watson &  2.003741 \\
-\end{tabular}
-\end{center}
-\end{table}
-%%%one more Gretl end
 \pagebreak
@@
 	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand. Series 51-75, variable 27. Gretl visualization­}
-\label{fig:57}
+	\label{fig:60}
 	\end{center}
@@
 	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
-\label{tab:14}
 	\begin{center}
@@
 	\end{center}
-\label{fig:5}
+	\label{tab:28}
 \end{table}
@@
 		\caption{Normal Q-Q plot: estimated versus observed values of the measurement of carbonate sand, variable 27. Series 51-75. Gretl modelling visualization­}
-\label{fig:57}
+	\label{fig:61}
 	\end{center}
@@
 	\end{center}
+\label{tab:29}
 \end{table}
@@
 	\caption{Results of the \textit{periodogramm} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
-\label{tab:14}
 	\begin{center}
@@
 	Number of observations = 105
-\label{fig:5}
+	\label{tab:30}
 \end{table}
 %%% Gretl end
+
 \begin{figure}[H]
@@
 		\caption{Plot illustrating periodogramm for v30 from measurements of carbonate sand, series 51-75. Gretl modelling visualization­}
-\label{fig:57}
-\end{center}
+	\end{center}
+	\label{fig:62}
 \end{figure}
@@
 		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
-\label{fig:57}
-\end{center}
-\end{figure}
+	\end{center}
+	\label{fig:63}
+\end{figure}
+
 \begin{figure}[H]
@@
 		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
-\label{fig:57}
-\end{center}
+	\end{center}
+	\label{fig:64}
 \end{figure}
@@
 		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
-\label{fig:57}
-\end{center}
+	\end{center}
+	\label{fig:65}
 \end{figure}
@@
 Wavelength step: 3 nm.
-Measured on Agia Pelagia beach, 14th October}
-\label{tab:17} \\
-\hline
+Measured on Agia Pelagia beach, 14th October} \\
  		 \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 	\multicolumn{7}{c}{Table 5: Spectral reflectance of \textbf{carbonate sand} \textit{(continued from previous page)}}\\
-\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
+	\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}				{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
 	\endhead
@@
 	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES}
-\label{fig:58}
+	\label{fig:66}
 \end{figure}
@@
 	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}
-\label{tab:18}
 	\centering
@@
 	  \end{tabular}
+	\label{tab:32}
 \end{table}
@@
 		\caption{Locations of selected measurements, visualization on Google Earth­}
-\label{fig:59}
-\end{center}
+	\end{center}
+	\label{fig:67}
 \end{figure}
@@
 		\caption{Random mosaic of selected aerial Google Earth images}
-\label{fig:60}
-\end{center}
-\end{figure}
-\pagebreak
-
-\begin{table}
+	\end{center}
+	\label{fig:68}	
+\end{figure}
+\pagebreak
+
+\begin{table}[H]
 	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}
-\label{tab:19}
 	\centering
@@
 	\end{tabular}
+	\label{tab:33}
 \end{table}
+\pagebreak
 
@@
 Script command of FWTools2.4.7 enabling to reduce the size of the aerial images, from .tif to .ecw
-format:
+format.
 \begin{figure}[h]
 	\begin{center}
-\includegraphics[scale=0.25]{Script_Gdal.jpg}
+		\includegraphics[scale=0.4]{Script_Gdal.jpg}
 		\caption{Script command of FWTools}
-\label{fig:61}
-\end{center}
+	\end{center}
+	\label{fig:69}
 \end{figure}
@@
 	\caption{Types of seagrass structural patterns, Ligaria beach, Crete}
-	\label{fig:62}
+	\label{fig:70}
 \end{figure}
@@
 		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}
-\label{fig:63}
-\end{center}
-\end{figure}
-\pagebreak
-\newpage
+	\end{center}
+	\label{fig:71}
+\end{figure}
+\pagebreak
 
@@
 		\caption{Ligaria beach, Crete: seafloor types}
-\label{fig:64}
-\end{center}
+	\end{center}
+	\label{fig:72}
 \end{figure}
@@
 		\caption{Various seafloor types; Ligaria beach, Crete}
-\label{fig:65}
-\end{center}
+	\end{center}
+	\label{fig:73}
 \end{figure}
@@
 		\caption{Measurement underwater equipment}
-\label{fig:66}
-\end{center}
-\end{figure}
-\pagebreak
-\newpage
+		\label{fig:74}
+	\end{center}
+\end{figure}
+\pagebreak
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 	\begin{center}
@@
 		\caption{Interpolation of the spectral measurements}
-\label{fig:67}
+		\label{fig:75}
 	\end{center}
@@
 	  \end{tabular}
-\label{tab:20}
+	\label{tab:34}
 \end{table}
@@
 	\caption{Landsat imagery, Crete island. Previews}
-	\label{fig:68}
+	\label{fig:76}
 \end{figure}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%% End MSc Thesis Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%% End Polina Lemenkova's MSc Thesis Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-01-16 MScDiplom_Lemenkova.tex
briefly described methods of stat analysis (Results), referenced Tables

@@
 Cymodocea, Amphibolis, Thalassodendron, Zostera, Heterozostera and Phyllospandix) \cite{Kuo89}\label{Kuo89}.
+
 \begin{wrapfigure}{r}{0.4\textwidth}
@@
 	\includegraphics[scale=0.25]{Fig-1-2.jpg}
-	\caption{Distribution of seagrasses in the world.\\ Source:\cite{Green03}\label{Green03} \label{fig:2}} 
+	\caption{Distribution of seagrasses in the world.\\ Source:\cite{Green03}\label{Green03} 
+	\label{fig:2}} 
 \end{wrapfigure}
+
 The distribution of the seagrasses is strongly influensed by several environmental factors, which
@@
 	\includegraphics[scale=0.25]{Fig-1-3.jpg}
-	\caption{Morphology of different types of seagrasses.
+	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
 	\label{fig:3}
-Source: \url{http://www.ozcoasts.org.au}}
 \end{wrapfigure}
@@
 from 4.4percent  minimal up to 29 percent \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as 35-
-45° and reflection factor 0.0201. P.oceanica. 
+45° and reflection factor 0.0201. P.oceanica. These values, simyulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) on the WASI user-defined parameters.
 The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:30}).
@@
 \subsubsection{Statistical analysis of the observational data and hypothesis testing}
-In the statistical anlysis of the raw observed data pool of the values of spectral reflectances (about 400 single measurements for P.oceanica), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
+The total amount of measured data was large and included following datasets made using
+hyperspectral radiometer Ramses: 350 measuremetn sets of P.oceanica reflectance for 14th Oct (Tab.\ref{tab:32}), 400
+sets of P.oceanica reflectance for 15th Oct (Tab.\ref{tab:33}), 84 datasets for seawater reflectance with sediments, 105
+datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonat sand.
+A statistical approach is evidently necessary for the proper processing of such amounts of data. 
+Statistical pre-processing of large sets of serial data enables to generalyse data by using the most typical and predicted values of
+spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
+The statistical calculators were mainly made by means of Gretle and SPSS. 
+The Open Office was used for preliminary data view and preprocessing, and included following computations
+at each data set (for example, Tab.\ref{tab:31}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). \\
+The statistical pre-processing and analysis were made to display the mean values of spectral reflectances,
+which were used for comparison with reflectance data of various seafloor cover types as well as
+seawater with and without sediment (e.g. Tab.\ref{tab:2}).
+In the statistical anlysis of the raw observed data pool of the values of spectral reflectances (about 400 single measurements for P.oceanica, e.g.
+Tab.\ref{tab:10}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
 parameters.
-Examples are the mean and variance of a probability distribution of measurements dataset (Fig.).
+Examples are the mean and variance of a probability distribution of measurements dataset (for example, Tab.\ref{tab:16}).
 We divided the total data pool into datasets of 25 measurements for more convenience and ease of use, both for P.oceanica and carbonate sand.
-The mean of a probability distribution of spectral reflectance describes the average value of the random variable over all of its possible realizations (see for example, Fig.
-Appendix).
+The mean of a probability distribution of spectral reflectance describes the average value of the random variable over all of its possible realizations (Appendices).
 Conceptually, there are an infinite number of such realizations, therefore parameters are not known to us.
-However, in the statistical analysis of the observed data our goal is to estimate these parameters using a finite amount of information available to us.
-Thus, we collect a number of realizations (a sample of 25 measurements) and then estimate the unknown parameters using a statistical analysis (Appendices, see Tables ). \\
-Finally we analysed 400 datasets made on two different data. 
+However, in the statistical analysis of the observed data our goal is to estimate these parameters using a finite amount of information available to us: fieldwork observations.
+\\
+Thus, we collect a number of realizations (a sample of 25 measurements for both sand and P.oceanica) and then estimate the statistical parameters (Appendix). \\
+Finally we analysed abount 400 datasets totally (example of carbonate sand on Fig.\ref{tab:31}) made during different days. 
 Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random. 
-Probability thus expresses uncertainty about the possible values of the values of spectral reflectances  of P.oceanica.
-Each time we observe the outcome of a random variable of spectral reflectance, we obtain an observation. 
-Once observed, its value is known and hence it is no longer random. 
-So, there is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
-We applied various methods of statistical probability analysis towards data pool, in order to analyse the distribution of values of spectral reflectances of P.oceanica (see Appendices, Tables). 
+Thus, we analysed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implmented in Gretl software (Tab.\ref{tab:19}). 
+Their probability expresses uncertainty about the possible values of the spectral reflectances  of P.oceanica and sand, respectively.
+There is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
+Each time we observe the outcome of a random variable of spectral reflectance, we obtain an observation, which is hence no longer random. 
+We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the Appendix: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyse the distribution of values of spectral reflectances of \textit{P.oceanica}.
 A probability distribution, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
-The total amount of measured data was large and included following datasets made using
-hyperspectral radiometer Ramses: 350 measuremetn sets of P.oceanica reflectance for 14th Oct, 400
-sets of P.oceanica reflectance for 15th Oct, 84 datasets for seawater reflectance with sediments, 105
-datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonat sand.
-Fig. 4.11. Spectral reflectance of water with sediments
-Therefore, a statistical approach is necessary for the processing of such data. Statistical pre-processing
-of large sets of serial data enables to generalyse data by using the most typical and predicted values of
-spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
-The statistical calculators were mainly made by means of Gretle and SPSS. The Open Office was used for preliminary data view and preprocessing, and included following computations
-at each data set (table 1): mean, median, Q1 and Q3, standard deviation, min and max values. The
-statistical pre-processing and analysis were made shown the mean values of spectral reflectances,
-which were used for comparison with reflectance data of various seafloor cover types as well as
-seawater with and without sediment.
+\\
+The \textit{least absolute deviations (LAD)}, a popular optimization technique, was used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
+Non-linear logistic anlysis is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
+the value of spectral reflectance for each single observation) is not continuous, but discrete and independent for each case (Tab.\ref{tab:11}).
+\\
+As the observations were made in a repeated way, we also analysed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand.
+The \textit{autocorrelation}, a cross-correlation of an observed value with itself, made on two different days (14th and 15th October),
+ shows the similarity between observations as a function of the time separation between them.
+As can be seen on Fig.\ref{fig:52}, Fig.\ref{fig:53} and Fig.\ref{fig:60}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
+The autocorrelation analysis has been performed using both SPSS and Gretl software.
+\\
+The method of \textit{least squares} minimizes the sum of the squares of the errors made in solving every single observation made.
+We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}.\\
+To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectances at several depths: Tab.\ref{tab:5}, Tab.\ref{tab:6}, Tab.\ref{tab:7}.
+The summary of the values of spectral reflectance properties on various depths is presented on the Fig.\ref{fig:17}.
+The \textit{k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
+It enables to highlight the main areas of the location of values of spectral reflectances along the spectra.\\
+To estimate spectral density of a spectral reflectancfe measurements we used periodogram function (Tab.\ref{tab:30}), 
+which corresponds to the general spectrum of the observations with representation of a variable quantity.  
+Thus, the periodogram of the series of observation highlights so-called “dead spots” of low power between the frequencies we wishe to
+exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:43} ) 
+and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:45}, enlarged part of 500-660nm).\\
 The visualization of the data plotting was made by means of Gnuplot software, which enabels fine
 drawing and advanced displaying of large amounts of serial data, ultimate control over graph properties, 
-the simplicity of plotting and the ease of scripting. (see Fig.15 and Appendix).
+the simplicity of plotting and the ease of scripting. (see Fig.\ref{fig:54}, Appendices).
+
 The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
-areas of quartiles (Q1, Q3, shaded areas on graphs), which cover most probable data distribution
-(spectral reflectance) for each spectral band. In x-axis I displayed the areas of 400 – 950 nm in
+areas of quartiles (Q1, Q3, shaded areas on graph, Fig.\ref{fig:55}), which cover most probable data distribution
+(spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
 spectra, where the measurements were done; the y-axis was adjusted for the better visualisation of the
 P.oceanica spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
-more than 0.20 nm, except for borders), I did not extend the y-scale up to 100 percent, in order to
-focus on the most necessary area of spectral values (Fig. 15 and Appendices). The mean values are
-highligted using bold points, as I used these values for plotting the final graph of P.oceanica spectra in
+more than 0.20 nm, except for borders), we did not extend the y-scale up to 100 percent, in order to
+focus on the most necessary area of spectral values (Fig.\ref{fig:35} and Appendices). The mean values are
+highligted using bold red points, as we used these values for plotting the final graph of P.oceanica spectra in
 Ligaria beach, Agia Pelagia.
@@
 The adjusted averages evaluate mean values, needed for statistical analysis of the datasets. 
-The outcome of this calculations is shown in the table 5 (Appendices). 
+The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (Appendices). 
 As a result, we have mean values of spectral reflectances of P.oceanica for extreme, average, median and both quartiles,
@@
 varying in-situ conditions, Ho: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
-P(reject H0|H0 is valid)= P(X>c|p=)=.05
+P(reject H0|H0 is valid)= P(X$>$c|p=)=.05
 where c comes for critical values and p=probability.
-The result P=.05 is, hence, very small, which makes the statement of Hypothesis Ho highly unlikely
+The result P=.05 is (Tab.\ref{tab:8}), hence, very small, which makes the statement of Hypothesis Ho highly unlikely
 (less than 1 in a 10 chance). The one-way ANOVA highlighted a significant difference between data
-of spectral reflectances of P.oceanica at different depths, i.e. true is the opposite statement:
+of spectral reflectances of P.oceanica at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
 Hypothesis Ha=“the spectral discernibility of the seagrass P.oceanica is distinctly changing with
@@
 The in situ spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
-types as the different sensors would percept them: MODIS, ASTER, MERIS, SeaWiFS and CZCS.
+types as the different sensors would percept them: MODIS, ASTER, MERIS, SeaWiFS and CZCS.\\
+These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
 These models simulate spectral views of the chosen sensors, how these sensors will "see" seagrass (Fig.\ref{fig:28}) and sand as
@@
 \pagenumbering{roman}
-\appendix
+\appendix \label{app}
 \appendixpage
@@
 \end{figure}
+
 \begin{figure}[H]
@@
 \end{figure}
+
 \begin{table}[htbp]
@@
 \end{table}
+
 \begin{table}[htbp]
@@
 \begin{longtable}{|c|c|c|c|c|c|c|}
-	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October} \\
+	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October}\label{tab:10} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 	\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
-\label{tab:10} 
 \end{longtable}
@@
 \newpage
-\begin{figure}[h]
+\begin{figure}[H]
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
+	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}\label{tab:14}
 	\begin{center}
@@
 	\end{tabular}
-	\label{tab:14}
 	\end{center}
@@
 	\begin{center}
-		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. Gretl.­}
+		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. Gretl.­}\label{tab:16}
 		Summary Statistics, using the observations 301--325\\
@@
 	\end{center}
-	\label{tab:16}
+
 \end{table}
@@
 Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-GNUplot visualization­}
+Gnuplot visualization­}
 		\label{fig:59}
@@
 	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{P.oceanica}. Gretl.}
+	\label{tab:26}
 		Model 3: Prais--Winsten, using observations 1:01--8:23 ($T$ = 191)\\
@@
 	\end{center}
-	\label{tab:26}
 \end{table}
@@
 	\caption{Results of the \textit{periodogramm} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
+	\label{tab:30}
 	\begin{center}
@@
 	Number of observations = 105
-	\label{tab:30}
 \end{table}
@@
 Wavelength step: 3 nm.
-Measured on Agia Pelagia beach, 14th October} \\
+Measured on Agia Pelagia beach, 14th October}\label{tab:31} \\
  		 \multicolumn{1}{|l|}{\textbf{wl}} &

### 2011-01-17 MScDiplom_Lemenkova.tex
added work flowchart diagram for stat analysis

@@
 
-\vspace{5em}
+\vspace{4em}
 
@@
 	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
+	\item \href{http://staruml.sourceforge.net/en/}{StarUML}: modelling flowchart diagrams for work algorithms
 \end{itemize}
@@
 \\
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.45, angle=90]{UML_StatAn.jpg}
+	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
+	\label{fig:34}
+\end{figure}
+
 Thus, we collect a number of realizations (a sample of 25 measurements for both sand and P.oceanica) and then estimate the statistical parameters (Appendix). \\

### 2011-01-18 MScDiplom_Lemenkova.tex
added flowchart, p.25

@@
 	\item \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
-	\item WASI Water Color Simulator, for modelling optical water properties under changing environment
+	\item \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}, for modelling optical water properties under changing environment
 	\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and GPS for data capture during the fieldwork
@@
 	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
-	\item \href{http://staruml.sourceforge.net/en/}{StarUML}: modelling flowchart diagrams for work algorithms
+	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia}: modelling flowchart diagrams for work algorithms
 \end{itemize}
@@
 different satellite images for 2000-2010.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.40]{UML_Dia.jpg}
+	\caption{Workflow for the data aquisition, \href{http://live.gnome.org/Dia}{Dia} display}
+	\label{fig:21}
+\end{figure}
+
 The spatial resolution of Landsat ETM+ image is 30 m in the visible and near infrared bands (bands
@@
 	\includegraphics[scale=0.5]{GNU-17.jpg}
-	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete}
+	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot}}
 	\label{fig:23}
@@
 	\includegraphics[scale=0.30]{GNU-18.jpg}
-	\caption{Multiplot graph showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-Gnuplot.
-\\Two complimentray graphs below show the results of the statistical analysis}
+	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
+\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentray graphs below show the results of the statistical analysis.}
 	\label{fig:26}
@@
 	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-Gnuplot.
-\\Two complimentray graphs below show the results of the statistical analysis}
+\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentray graphs below show the results of the statistical analysis}
 	\label{fig:27}
@@
 	\includegraphics[scale=0.30]{GNU-20.jpg}
-	\caption{Statitical comparison of spectral reflectances of \textit{P.oceanica} and carbonate sand. Selected measurements sets}
+	\caption{Statitical comparison of spectral reflectances of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
 	\label{fig:28}
@@
 	\includegraphics[scale=0.40]{GNU-19.jpg}
-	\caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and 			measured values (dotted): multiplot of measurement sets 200-300}
+	\caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and 			measured values (dotted): multiplot of measurement sets 200-300.
+\href{http://www.gnuplot.info/}{Gnuplot}}
 	\label{fig:33}
@@
 		\caption{Multiplot of spectral reflectance of P.oceanica. Series 1-100. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
-			mean value within the range (red bold dots). Gnuplot visualization­}
+			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
 		\label{fig:35}
@@
 		\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
-			mean value within the range (red bold dots). Gnuplot visualization­}
+			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
 		\label{fig:36}
@@
 	\subfloat [enlarged (1:10,000)]{\label{fig:35b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
-	\caption{Seafloor types of the Arina beach: \textit{P.oceanica} sand, rocks}
+	\caption{Seafloor types of the Arina beach: \textit{P.oceanica,} sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
 	\label{fig:37}
@@
 		\includegraphics[scale=0.40]{Fig-36.jpg}
-		\caption{Results of the image classification in Erdas Imagine: seagrass distribution in Bali area, Crete}
+		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }
 		\label{fig:38}
@@
 		\includegraphics[scale=0.25]{Fig-37.jpg}
-		\caption{Google aerial images incorporated into the GIS project: fragment of ArcMap layout}
+		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout}
 		\label{fig:39}
@@
 	\centering
-	\subfloat [Meadows of the seagrass. Source: Googel Earth]{\label{fig:38a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
+	\subfloat [Meadows of the seagrass. Source: \href{http://www.google.com/earth/index.html}{Google Earth}]{\label{fig:38a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
 	\subfloat [Patches (mattes) of the seagrass. Source: in-situ videometric measurements]{\label{fig:38b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
@@
 		\includegraphics[scale=0.25]{Fig-39.jpg}
-		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the ArcGIS project}
+		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project}
 		\label{fig:41}
@@
 		\includegraphics[scale=0.20]{GNU-401-420.jpg}
-		\caption{Spectral reflectance of P.oceanica. Measurement series 401-420}
+		\caption{Spectral reflectance of P.oceanica. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
 		\label{fig:43}
@@
 		\caption{Statistical analysis of the measurement data: spectral reflectance of the P.oceanica. 
-		Visualisation of the interquartile ranging. Example of data set 401-420. Shown midspread of  statistical quartiles Q1 and Q3, min and max values within the range.}
+		Visualisation of the interquartile ranging.
+Example of data set 401-420.
+Shown midspread of  statistical quartiles Q1 and Q3, min and max values within the range.
+\href{http://www.gnuplot.info/}{Gnuplot}}
 		\label{fig:44}
@@
 		\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
-		Visualisation of the interquartile ranging and plotted together with measurement data.}
+		Visualisation of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
 		\label{fig:45}
@@
 	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral 			reflectance of \textbf{P.oceanica}.
-SPSS}
+\href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{P.oceanica}. SPSS}
+	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{P.oceanica}.
+\href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. SPSS}
+	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\caption{Fragment of the statistical analysis of the P.oceanica reflectance. Example of data set 401-420. 
-		Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means.}
+		Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
 	\label{fig:46}
@@
 	\includegraphics[scale=0.25]{GNU-Radiance_Bezier.jpg}
-	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation}
+	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}}
 	\label{fig:47}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters}
+	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters.
+\href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of\textbf{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters}
+	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 		0.5, 1.5 and 2.5 meters.
+\href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 		\includegraphics[scale=0.25]{GNU-Irr-Splines.jpg}
-		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in GNUplot­}
+		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}­}
 		\label{fig:48}
@@
 	\includegraphics[scale=0.25]{GNU-Irr.jpg}
-	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in GNUplot}
+	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}}
 	\label{fig:49}
@@
 		\includegraphics[scale=0.4]{G-gamma-rad.jpg}
-		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
+		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}}
 		\label{fig:50}
@@
 	\includegraphics[scale=0.4]{G-normal-dist.jpg}
-	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in Gretl.­}
+	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}­}
 	\label{fig:51}
@@
 \begin{table}
-	\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16. Gretl}
+	\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16. \href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 	\rowcolors{1}{Lavender}{LightYellow}
-	\caption{Robust estimation of the seawater radiance (measured in aquarium tank, Heraklion): results of the \textit{least absolute deviation} (LAD), Series:V16. Gretl}
+	\caption{Robust estimation of the seawater radiance (measured in aquarium tank, Heraklion): results of the \textit{least absolute deviation} (LAD), Series:V16.
+\href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\includegraphics[scale=0.45]{Autocorr.jpg}
-	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
+	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
 	\label{fig:52}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. SPSS}\label{tab:14}
+	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3.
+\href{http://www.spss.com/}{SPSS}}\label{tab:14}
 	\begin{center}
@@
 	\includegraphics[scale=0.45]{Partial_Corr.jpg}
-	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in SPSS}
+	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
 	\label{fig:53}
@@
 		\includegraphics[scale=0.25]{GNU-09.jpg}
-		\caption{Spectral reflectance of the seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle (Gnuplot)­}
+		\caption{Spectral reflectance of the seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle. \href{http://www.gnuplot.info/}{Gnuplot}­}
 		\label{fig:54}
@@
 	\begin{center}
-		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Gretl.}
+		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+\href{http://gretl.sourceforge.net/}{Gretl}}
 		Model 2: OLS, using observations 326--350\\
@@
 		\includegraphics[scale=0.22]{GNU-10.jpg}
-		\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. Gnuplot display­}
+		\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. \href{http://www.gnuplot.info/}{Gnuplot} display­}
 		\label{fig:55}
@@
 	\begin{center}
-		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. Gretl.­}\label{tab:16}
+		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. \href{http://gretl.sourceforge.net/}{Gretl}­}\label{tab:16}
 		Summary Statistics, using the observations 301--325\\
@@
 		\includegraphics[scale=0.25]{GNU-11.jpg}
-		\caption{Spectral reflectance of P.oceanica. Measurement series 25-50. GNUplot display­}
+		\caption{Spectral reflectance of P.oceanica. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
 		\label{fig:56}
@@
 		\includegraphics[scale=0.25]{GNU-12.jpg}
-		\caption{RS reflectance of P.oceanica. Series 1-25. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range.­}
+		\caption{Remote sensing reflectance of P.oceanica.
+Series 1-25.
+\\Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range.
+\href{http://www.gnuplot.info/}{Gnuplot}­}
 		\label{fig:57}
@@
 \begin{table}[htbp]
-	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Gretl.}
+	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+\href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-GNUplot visualization­}
+\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
 		\label{fig:58}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+\href{http://www.spss.com/}{SPSS}}
 	\label{tab:13}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+\href{http://www.spss.com/}{SPSS}}
 		\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.). SPSS}
+	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 \begin{table}[H]
-	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Gretl.}
+	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+\href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 		\rowcolors{1}{AntiqueWhite}{LightGoldenrodYellow}
-		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textbf{P.oceanica}. Gretl.}
+		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textbf{P.oceanica}.
+\href{http://gretl.sourceforge.net/}{Gretl}}
 			Model 2: Tobit, using observations 1:01--8:23 ($T$ = 191)\\
@@
 Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-Gnuplot visualization­}
+\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
 		\label{fig:59}
@@
 	\begin{center}
-	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{P.oceanica}. Gretl.}
+	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{P.oceanica}.
+\href{http://gretl.sourceforge.net/}{Gretl}}
 	\label{tab:26}
@@
 	\begin{center}
-	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica}. Gretl.}
+	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica}.
+\href{http://gretl.sourceforge.net/}{Gretl}}
 		Model 1: WLS, using observations 1--95\\
@@
 	\includegraphics[scale=0.4]{G-autocorr-sand51-75.jpg}
-	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand. Series 51-75, variable 27. Gretl visualization­}
+	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand. Series 51-75, variable 27. \href{http://gretl.sourceforge.net/}{Gretl} visualization­}
 	\label{fig:60}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 		\includegraphics[scale=0.4]{G-QQplot-sand.jpg}
-		\caption{Normal Q-Q plot: estimated versus observed values of the measurement of carbonate sand, variable 27. Series 51-75. Gretl modelling visualization­}
+		\caption{Normal Q-Q plot: estimated versus observed values of the measurement of carbonate sand, variable 27. Series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
 	\label{fig:61}
@@
 \begin{table}[H]
-	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
+	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
-	\caption{Results of the \textit{periodogramm} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, Gretl.}
+	\caption{Results of the \textit{periodogramm} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
 	\label{tab:30}
@@
 		\includegraphics[scale=0.25]{G-period-sand30.jpg}
-		\caption{Plot illustrating periodogramm for v30 from measurements of carbonate sand, series 51-75. Gretl modelling visualization­}
+		\caption{Plot illustrating periodogramm for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
 	\end{center}
@@
 		\includegraphics[scale=0.32]{G-polynom-trend-wo.jpg}
-		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
+		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15.
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
 	\end{center}
@@
 		\includegraphics[scale=0.32]{G-polynom-trend-with.jpg}
-		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
+		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
 	\end{center}
@@
 		\includegraphics[scale=0.32]{G-exp-mov-aver.jpg}
-		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. Gretl modelling visualization­}
+		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
 	\end{center}
@@
 	\includegraphics[scale=0.35]{Fig-26.jpg}
-	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES}
+	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES. \href{http://www.gnuplot.info/}{Gnuplot}}
 	\label{fig:66}

### 2011-01-20 MScDiplom_Lemenkova.tex
added flowchart, p.9

@@
 	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
-	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia}: modelling flowchart diagrams for work algorithms
+	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia} , \href{http://argouml.tigris.org/}{AgroUML}: diagramming flowcharts for work algorithms
 \end{itemize}
@@
 
-\subsection{Research scheme}
-
 \subsection{Research approach}
@@
 shallow waters keeps information on the optical properties of the seawater components and benthic
-substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. Spectral
-reflectances, distinct for each seafloor cover type, enable their discrimination on the image. The preprocessing
+substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. \\
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
+	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{AgroUML}}
+	\label{fig:16}
+\end{figure}
+
+Spectral reflectances, distinct for each seafloor cover type, enable their discrimination on the image. The preprocessing
 of the images includes imagery corrections for atmospheric noises and effects of the water
@@
 Supervised, Maximum Likelihood).
+
+\subsection{Research scheme}
+
 \pagebreak

### 2011-01-21 MScDiplom_Lemenkova.tex
added collage of general research approach, Inkscape

@@
 	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
-	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia} , \href{http://argouml.tigris.org/}{AgroUML}: diagramming flowcharts for work algorithms
+	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia} , \href{http://argouml.tigris.org/}{ArgoUML}: diagramming flowcharts for work algorithms
+	\item \href{http://inkscape.org/}{Inkscape}: drawing
 \end{itemize}
@@
 	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
-	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{AgroUML}}
+	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}}
 	\label{fig:16}
@@
 \subsection{Research scheme}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{UML_GIS.png}
+	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
+	\label{fig:16}
+\end{figure}
 
@@
 \rowcolors{1}{Khaki}{LightYellow}
-\caption{Statistical analysis of the measurements of spectral reflectance of P.oceanica (fragment).}
+\caption{Statistical analysis of the measurements of spectral reflectance of P.oceanica (fragment)}
 \centering

### 2011-01-23 MScDiplom_Lemenkova.tex
updated Diploma directory, added new files, renumerated Figures and Tables

@@
 \usepackage[svgnames,table]{xcolor}
+\usepackage[subfigure]{tocloft}
 %\usepackage[table]{xcolor}
@@
 \usepackage{caption}
+\usepackage{chngcntr}
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
 \usepackage{appendix}
@@
 
-\begin{figure}[b]
+\begin{figure}[h]
 	\begin{center}
@@
 
-\vspace{4em}
+\vspace{2em}
 
@@
 	\item \href{http://inkscape.org/}{Inkscape}: drawing
+	\item \href{http://www.notepad-plus-plus.org/}{Notepad++}: auxiliary text editing and raw data formatting
 \end{itemize}
@@
 \listoffigures
+\setcounter{lofdepth}{2}
 \pagebreak
@@
 \section{Introduction}
-%\markright{I. Introduction}
+%%%%%%%%%%%%%%%%%%Section1%%%%%%%%%%%%%%%%%%
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+
+\markright{I. Introduction}
 
@@
 The distribution of the seagrasses is strongly influensed by several environmental factors, which
-include climate (mostly, tropical and temperate areas), bathymetry (shallow shelf zones), hydrological
+include climate (mostly, tropical and temperate are--as), bathymetry (shallow shelf zones), hydrological
 particularities (chemical content of water, nutritient availability and turbidity of waves), and geological 
@@
 	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}}
-	\label{fig:16}
+	\label{fig:10}
 \end{figure}
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.5]{UML_GIS.png}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.6]{UML_GIS.png}
 	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
-	\label{fig:16}
-\end{figure}
-
-\pagebreak
-
+	\label{fig:11}
+\end{figure}
+
+\pagebreak
+%%%%%%%%%%%%%%%%%Section2%%%%%%%%%%%%%%%%%%%%%%
 \section[Overview of literature]{Seagrass monitoring: overview of literature and research resources} \label{sec:2}
-%\markright{II. Seagrass monitoring: overview of literature and research resources}
-%\renewcommand{\headrulewidth}{0.4pt}
+\markright{II. Seagrass monitoring: overview of literature and research resources}
+\renewcommand{\headrulewidth}{0.4pt}
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
 
@@
 	\caption{RAMSES Hyperspectral Sensors. Source: \href{http://www.trios.de}{Trios}}
-	\label{fig:10}
+	\label{fig:12}
 \end{figure}
@@
 		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
-	\label{fig:11}
+	\label{fig:13}
 \end{wrapfigure}
@@
 		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
-	\label{fig:12}
+	\label{fig:14}
 \end{wrapfigure}
@@
  	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
-	\label{fig:13}
+	\label{fig:15}
 \end{wrapfigure}
@@
 	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
-	\label{fig:14}
+	\label{fig:16}
 \end{wrapfigure}
@@
 		reflectance. Source:\cite{Yang10}\label{Yang10}}
-	\label{fig:15}
+	\label{fig:17}
 \end{wrapfigure}
@@
 
+%%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%
 \section{Materials and methods}
-%\markright{III. Materials and methods}
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
 
@@
 	\caption{Study area: Crete island}
-	\label{fig:16}
+	\label{fig:18}
 \end{figure}
@@
 	\caption{Study area: Ligaria beach, Crete island}
-	\label{fig:17}
+	\label{fig:19}
 \end{figure}
@@
 	\centering
-	\subfloat [iPAQ, HP]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-19a.jpg}}
-	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-19b.jpg}}
-	\subfloat [GPS]{\label{fig:19a}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
-	\subfloat [Olympus waterproof camera]{\label{fig:19a}\includegraphics[width=0.25\textwidth]{Fig-19d.jpg}}
+	\subfloat [iPAQ, HP]{\label{fig:20a}\includegraphics[width=0.3\textwidth]{Fig-19a.jpg}}
+	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:20b}\includegraphics[width=0.3\textwidth]{Fig-19b.jpg}}
+	\subfloat [GPS]{\label{fig:20c}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
+	\subfloat [Olympus waterproof camera]{\label{fig:20d}\includegraphics[width=0.25\textwidth]{Fig-19d.jpg}}
 	\caption{Fieldwork equipment}
-	\label{fig:18}
+	\label{fig:20}
 \end{figure}
@@
 	\caption{Boat used for the fieldwork measurements on Crete island. 2010}
-	\label{fig:19}
+	\label{fig:21}
 \end{figure}
@@
 	\centering
-	\subfloat [view from below]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
-	\subfloat [view \textit{en-face}]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
-	\subfloat [view from above]{\label{fig:19a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
+	\subfloat [view from below]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
+	\subfloat [view \textit{en-face}]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
+	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
 	\caption{Scheme of placements of the Olimpus cameras during the measurements}
-	\label{fig:20}
+	\label{fig:22}
 \end{figure}
@@
 	\caption{Locations of the video measurements and GPS tracklogs, Ligaria}
-	\label{fig:21}
+	\label{fig:23}
 \end{figure}
@@
 and up-to-date source of imagery.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.27]{Fig-22.jpg}
-	\caption{Capturing aerial imagery from the Google Earth: grabbing process}
-	\label{fig:22}
-\end{wrapfigure}
-
 Perhaps, the most clear advantage of the Google
@@
 
+%%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
 \pagebreak
 \section{Results}
-%\markright{IV. Results}
+\markright{IV. Results}
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
 
@@
 	\caption{Workflow for the data aquisition, \href{http://live.gnome.org/Dia}{Dia} display}
-	\label{fig:21}
+	\label{fig:24}
 \end{figure}
@@
 	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:23}
+	\label{fig:25}
 \end{figure}
@@
 were made with 3 nm interval, while others – using 4 and 1 nm step.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.30]{Fig-24.jpg}
-	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
-	\label{fig:24}
-\end{figure}
-
 Therefore, these data had to be interpolated and standardized to one
@@
 comparing to splines (see Appendix for more results). 
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.25]{Fig-28.jpg}
-	\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
-		seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
-	\label{fig:25}
-\end{figure}
-
 Therefore, after several experiments with
@@
 	\centering
-	\subfloat [CZCS]{\label{fig:34a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
-	\subfloat [MERIS]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
-	\subfloat [MODIS]{\label{fig:34c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
-	\subfloat [SeaWiFS]{\label{fig:34d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
+	\subfloat [CZCS]{\label{fig:29a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
+	\subfloat [MERIS]{\label{fig:29b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
+	\subfloat [MODIS]{\label{fig:29c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
+	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
 	\caption{Simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 	\centering
@@
 	\centering
-	\subfloat [Sand]{\label{fig:27a}\includegraphics[width=0.4\textwidth]{Fig-27a.jpg}}
-	\subfloat [\textit{P.oceanica}]{\label{fig:27b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
+	\subfloat [Sand]{\label{fig:34a}\includegraphics[width=0.4\textwidth]{Fig-27a.jpg}}
+	\subfloat [\textit{P.oceanica}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
 	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
@@
 	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
-	\label{fig:34}
+	\label{fig:35}
 \end{figure}
@@
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
-		\label{fig:35}
+		\label{fig:36}
 	\end{center}
@@
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
-		\label{fig:36}
+		\label{fig:37}
 	\end{center}
@@
 	\centering
-	\subfloat [small-scaled (1:54,000)]{\label{fig:35a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
-	\subfloat [enlarged (1:10,000)]{\label{fig:35b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
+	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
+	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
 	\caption{Seafloor types of the Arina beach: \textit{P.oceanica,} sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
-	\label{fig:37}
+	\label{fig:38}
 \end{figure}
@@
 		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }
-		\label{fig:38}
+		\label{fig:39}
 	\end{center}
@@
 		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout}
-		\label{fig:39}
+		\label{fig:40}
 	\end{center}
@@
 	\centering
-	\subfloat [Meadows of the seagrass. Source: \href{http://www.google.com/earth/index.html}{Google Earth}]{\label{fig:38a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
-	\subfloat [Patches (mattes) of the seagrass. Source: in-situ videometric measurements]{\label{fig:38b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
+	\subfloat [Meadows of the seagrass. Source: \href{http://www.google.com/earth/index.html}{Google Earth}]{\label{fig:41a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
+	\subfloat [Patches (mattes) of the seagrass. Source: in-situ videometric measurements]{\label{fig:41b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
 	\caption{Variations in spatial structure of the seagrass landscapes}
-	\label{fig:40}
+	\label{fig:41}
 \end{figure}
@@
 		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project}
-		\label{fig:41}
+		\label{fig:42}
 	\end{center}
@@
 
-\pagebreak
-
 \subsubsection{Results of GIS mapping}
@@
 \section{Discussion and conclusions}
-%\markright{V. Discussion and conclusions}
 The suggested methodology of the analysis of seagrass adaptation towards different environments of
@@
 
+%%%%%%%%%%%%%%%%%%%  Appendix  %%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{appendices}
@@
 \addappheadtotoc
+%%%
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+%\renewcommand{\thefigure}{A-\arabic{figure}}
+%\renewcommand{\thetable}{A-\arabic{table}}
+
+\section{appendices}
 \appendixname{\textbf{1:\\Results of the measurements of seafloor spectral reflectance}}
-%\setcounter{figure}{0}
+%%%
 
@@
 		\caption{Values of spectral reflectances of seagrass P.oceanica (fragment)}
-		\label{fig:42}
+		\label{fig:A-1}
 	\end{center}
@@
 		\caption{Spectral reflectance of P.oceanica. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
-		\label{fig:43}
+		\label{fig:44}
 	\end{center}
@@
 \href{http://www.gnuplot.info/}{Gnuplot}}
-		\label{fig:44}
+		\label{fig:45}
 	\end{center}
@@
 		Visualisation of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
-		\label{fig:45}
+		\label{fig:46}
 	\end{center}
@@
 		Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:46}
+	\label{fig:47}
 	\end{center}
@@
 	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:47}
+	\label{fig:48}
 \end{figure}
@@
 format.
-\begin{figure}[h]
+\begin{figure}[H]
 	\begin{center}
@@
 \end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{Fig-24.jpg}
+	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
+	\label{fig:24}
+\end{figure}
+
 \pagebreak
@@
 	\end{center}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{Fig-28.jpg}
+	\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
+		seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
+	\label{fig:25}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.6]{Fig-22.jpg}
+	\caption{Capturing aerial imagery from the Google Earth: grabbing process}
+	\label{fig:22}
 \end{figure}

### 2011-01-23 MScDiplom_Lemenkova.tex
added new sections into ToC (List of Tools, Abstract, etc)

@@
 
-\textbf{Course Title:} \hfill{Geo-Information Science and Earth Observation for Environmental Modelling and Management}\\
+\textbf{Course Title:} \\
+{Geo-Information Science and Earth Observation for Environmental Modelling and Management}\\
 
@@
 
-University of Southampton (UK)\\
-Lund University (Sweden)\\
-University of Warsaw (Poland)\\
-University of Twente, Faculty ITC (The Netherlands)
-
+\begin{itemize}
+	\item[*] University of Southampton (UK)\\
+	\item[*] Lund University (Sweden)\\
+	\item[*] University of Warsaw (Poland)\\
+	\item[*] University of Twente, Faculty ITC (The Netherlands)
+\end{itemize}
 \pagebreak
@@
 
-Thesis submitted to the University of Twente, faculty ITC, in partial fulfilment of the requirements
+Thesis submitted to the University of Twente, Faculty ITC, in partial fulfilment of the requirements
 for the degree of Master of Science in Geo-information Science and Earth Observation for
@@
 
-Thesis Assessment Board:\\
-
-Name Examiner 1 (Chair)
-Name Examiner 2
-External Examiner
-Member
-Supervisors: Venus V. (Department of Natural Resource, ITC)
-Toxopeus A.G. (Department of Natural Resource, ITC)\\
+\textbf{Thesis Assessment Board}:
+\begin{itemize}
+	\item Name Examiner 1 (Chair)
+	\item Name Examiner 2
+	\item External Examiner
+\end{itemize}
+
+\vspace{2em}
+
+\textbf{Member Supervisors}: 
+\begin{itemize}
+	\item[*] Venus V. (Department of Natural Resource, ITC)
+	\item[*] Toxopeus A.G. (Department of Natural Resource, ITC)
+\end{itemize}
 
@@
 \end{figure}
-
-\pagebreak
-
-\begin{abstract}
-\end{abstract}
-
-\pagebreak
-\textbf{Disclaimer}\\
-
-This document describes work undertaken as part of a programme of study at the University of
-Twente, Faculty ITC. All views and opinions expressed therein remain the sole responsibility of
-the author, and do not necessarily represent those of the university.\\
-
 \pagebreak
@@
 	\tableofcontents 
+\pagebreak
+
+\section*{Acknowledgements}
+	\markright{Acknowledgements}
+	\addcontentsline{toc}{section}{Acknowledgements}
+\pagebreak
+
+\section*{Disclaimer}
+	\markright{Disclaimer}
+	\addcontentsline{toc}{section}{Disclaimer}
+		This document describes work undertaken as part of a programme of study at the University of
+		Twente, Faculty ITC. All views and opinions expressed therein remain the sole responsibility of
+		the author, and do not necessarily represent those of the university.\\
+\pagebreak
+
+\section*{Abstract}
+	\markright{Abstract}
+	\addcontentsline{toc}{section}{Abstract}
+\pagebreak
+
+\section*{List of Tools}\label{sec:0}
+	\markright{List of Tools}
+	\addcontentsline{toc}{section}{List of Tools}
 
@@
 		\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
-	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia} , \href{http://argouml.tigris.org/}{ArgoUML}: diagramming flowcharts for work algorithms
+		\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia} , \href{http://argouml.tigris.org/}						{ArgoUML}: diagramming flowcharts for work algorithms
 		\item \href{http://inkscape.org/}{Inkscape}: drawing
@@
 aerial images and the results of the underwater videographic measurements towards seagrass
-mapping. Image classification is based on the principle of the differentiation between the spectral
-signatures of various seafloor cover types. The spectrum of light coming up from the ocean surface in
+mapping  (Fig.\ref{fig:1.11}). Image classification is based on the principle of the differentiation between the spectral
+signatures of various seafloor cover types  (Fig.\ref{fig:1.10}). The spectrum of light coming up from the ocean surface in
 shallow waters keeps information on the optical properties of the seawater components and benthic
@@
 	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}}
-	\label{fig:10}
+	\label{fig:1.10}
 \end{figure}
@@
 	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
-	\label{fig:11}
+	\label{fig:1.11}
 \end{figure}
@@
 ones: \href{http://www.seagrassnet.org/}{Global Seagrass Monitoring project}, \href{http://www.seagrasswatch.org/}{Seagrasswatch}, 
-located in Australia, \href{http://mediterranean.seagrassonline.org/}{the Mediterranean association Seagrass-2000},
-\href{http://www.imedea.uib.es/index.php}{the Mediterranean Institute for Advanced Studies},
-\href{http://www2.fiu.edu/~seagrass/}{Seagrass Ecosystems Research Laboratory in South Florida},
-\href{http://www.flseagrass.org/}{Florida Seagrass organisation}, \href{http://www.seagrasses.org/}{the Seagrasses.org} and the 
-\href{http://wsa.seagrassonline.org/}{World Seagrass Association}.
- These organizations aim at global seagrass monitoring, providing with
+located in Australia, two European-located institutions - \href{http://mediterranean.seagrassonline.org/}{the Mediterranean association Seagrass-2000} and \href{http://www.imedea.uib.es/index.php}{the Mediterranean Institute for Advanced Studies}, US American
+\href{http://www2.fiu.edu/~seagrass/}{Seagrass Ecosystems Research Laboratory in South Florida} and
+\href{http://www.flseagrass.org/}{Florida Seagrass organisation}, the \href{http://wsa.seagrassonline.org/}{World Seagrass Association} and  
+the \href{http://www.seagrasses.org/}{Seagrasses.org}.
+ These organizations aim at the global seagrass monitoring, providing with
 research results and reporting guidelines and manuals with standartized methods and
@@
 	\caption{RAMSES Hyperspectral Sensors. Source: \href{http://www.trios.de}{Trios}}
-	\label{fig:12}
+	\label{fig:2.1}
 \end{figure}
@@
 level of precision, specially calibrated for air and for water application as well as colour
-measurements  (Fig.\ref{fig:10}). These products have been used for the radiance and irradiance
+measurements  (Fig.\ref{fig:2.1}). These products have been used for the radiance and irradiance
 measurements in Agia Pelagia bay, Crete island, 2009.
@@
 		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
-	\label{fig:14}
+	\label{fig:2.3}
 \end{wrapfigure}
@@
 The possibility of their application towards seagrass mapping
-vary and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:12}). In the next subparagraphs
+vary and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:2.3}). In the next subparagraphs
 we briefly discuss the limitations and research experience of the using of various imagery
@@
 Being the longest running satellite system, lunched in 1972, Landsat is the only source of archival
-data going back to 1984 at a sufficient spatial resolution \cite{Dekker05}\label{Dekker05}, which makes its data
+data going back to 1984 at a sufficient spatial resolution \cite{Dekker05b}\label{Dekker05b}, which makes its data
 desirable for historial mapping or environmental analysis of change detection of the seagrass
@@
 submerged vegetation, such as seagrasses or coral reefs. The successful applications of the imagery
-Landsat TM towards the seagrass mapping, were reported in numerous research works \cite{Palandro03,Gullstroom06,Wabnitz08,Bierwirth93,Ferguson97,Rasib97}\label{Palandro03} \label{Gullstroom06} \label{Wabnitz08} \label{Bierwirth93} \label{Ferguson97} \label{Rasib97}.
-The Landsat data are particularly suitable for the
+Landsat TM towards the seagrass mapping, were reported in numerous research works \cite{Palandro03,Gullstroom06,Wabnitz08,Bierwirth93,Ferguson97,Rasib97}\label{Palandro03}\label{Gullstroom06}\label{Wabnitz08}\label{Bierwirth93}\label{Ferguson97}\label{Rasib97}.\\ The Landsat data are particularly suitable for the
 case of change detection of seagrass landscapes at a decadal scale, because being the main sensor
-onboard the Landsat satellites, Landsat Thematic Mapper (TM) provides the longest time series
+onboard the Landsat satellites, the Thematic Mapper (TM) provides the longest time series
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
@@
  	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
-	\label{fig:15}
+	\label{fig:2.4}
 \end{wrapfigure}
@@
 	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
-	\label{fig:16}
+	\label{fig:2.5}
 \end{wrapfigure}
@@
 1999 onboard Terra sensor, provides high-resolution images of the Earth in 15 different bands of the
-electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:13}). It has diverse
+electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:2.4}). It has diverse
 subsystems for the visible near infrared (VNIR) with 15 m resolution, shortwave infrared (SWIR), and
@@
 		reflectance. Source:\cite{Yang10}\label{Yang10}}
-	\label{fig:17}
+	\label{fig:2.6}
 \end{wrapfigure}
@@
 mapping the submerged landscapes on the basis of pixels classification, the application of data from
-hyperspectral sensors has better potential due to their higher resolution.
+hyperspectral sensors has better potential due to their higher resolution (Fig.\ref{fig:2.5}).
 
@@
 The application of the hyperspectral sensors is most effective and provide more accurate classification
-results (Fig.2.1.), due to their higher spectral resolution \cite{Bharathi03}\label{Bharathi03} with interval
+results  (Fig.\ref{fig:2.5}), due to their higher spectral resolution \cite{Bharathi03}\label{Bharathi03} with interval
 narrows to 10 nanometers, while broadband sensors are limited to the spectral width of ca
-150 nm (Fig.\ref{fig:14}). 
+150 nm  (Fig.\ref{fig:2.3}). 
 Hyperspectral imagery is acquired through the simultaneous acquisition of images in many
@@
 Therefore, there is great potential of the application of hyperspectral remote sensing imagery towards seagrass
-mapping at species level, as long as they are distinguishable spectrally (Fig.\ref{fig:15}) which, for example, has been
+mapping at species level, as long as they are distinguishable spectrally (Fig.\ref{fig:2.6}) which, for example, has been
 tested in the case study of Australian marine ecosystems by Fyfe \cite{Fyfe04}\label{Fyfe04}. Application of various
@@
 
-In section \nameref{sec:2} we briefly discussed main existing cartographic methods and tools for seagrss mapping.
-In current  research we try to combine various techniques for assessment of seagrass in Mediterranean environment, in the cse study of Crete.
-Therefore, logically separated research sections are based on various software and tools.
-thus, spectral analysis and assessment of spectral signaturesis technically based on WASI, sttistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using\href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
+In the section \nameref{sec:2} we briefly discussed main existing cartographic methods and tools for seagrss mapping.
+In the current  research we try to combine various techniques from mentioned above for the assessment of the seagrass distribution in the Mediterranean environment, for the case study of Crete.
+We mentioned in \nameref{sec:0} the software that we used for logically separated research sections, which required various approach.
+Thus, for example,  spectral analysis and assessment of spectral signaturesis technically based on WASI, sttistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using\href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
 \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0} software is used for the spatial analysis, general mapping and cartographic layout
@@
 
-\section{appendices}
+\section*{}
 \appendixname{\textbf{1:\\Results of the measurements of seafloor spectral reflectance}}

### 2011-01-23 MScDiplom_Lemenkova.tex
corrected references to figures, added acknowledgements

@@
 	\addcontentsline{toc}{section}{Acknowledgements}
+
+This research was carried out at the University of Twente, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)}, Enschede, the Netherlands, under the financial support of the EU \href{http://ec.europa.eu/education/external-relation-programmes/doc72_en.htm}{Erasmus Mundus Scholarship}.
+\vspace{1mm}\\
+I thank my scientific supervisors, Dr.
+\textit{Albertus G.Toxopeus}  and Mr.
+\textit{Valentijn Venus} from the ITC, Department of Natural Resources, for our successful scientific collaboration, help and discussions of my research work.
+\\
+I also express them my most sincere appreciations for our fieldwork on Crete.
+\\ Besides scientific routine, we had an unforgettable time spent together.
+\\ Personally, I would like to mention that I have visited Greece the first time in my life, and I will always remember my Cretan period-2010 and Ligaria beach.\vspace{1mm}\\
+I extend my gratitude to the staff of the Museum of Natural History, the University of Crete, especially to Dr.\textit{Petros Limberakis} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g.
+SCUBA gear diving, instruments for underwater measurements, etc).
+I also appretiate their very warm, personal attitude and interest towards our research.
+\vspace{1mm}\\
+My sincere gratitude goes again to Mr.\textit{Valentijn Venus} for his generous providing me with a lot of spatial data   relevant to my research (e.g.
+such as digital vector layers of Crete island, raster georeferenced maps for the whole island for my GIS project, Excel spreadsheets, etc).
+Without these vital data the implementation of this project would have hardly been possible.
+I am also thankful to him for the technical help in Google Earth imagery capturing, using Gdal technologies and assistance during the fieldwork (videometric underwater footage, consultancies in seagrass observation and monitoring, driving boat, etc).
+\\ My special thank to his personal, friendly support in various situations on Crete.\vspace{1mm}\\
+I am thankful to Dr.rer.nat.\textit{Peter Gege}, the software developer from the Earth Observation Centre of the German Space Agency, for his generous consultancies concerning the technical question about WASI that arose during my work.
+I am thankful for all his explanations and articles given to me via e-mail \vspace{1mm}\\
+My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throuout my career.
+He encouraged me to use various very useful software and tools, always gave me necessary recommendations and precious help, thus highly contributing to my professional growth.
+\\ My special thank for his help with Python scripting.
+\\
+I am also heartfelt thankful to him for his personal friendship and support. 
+\vspace{1mm}\\
+I am greatful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery.
+\vspace{1mm}\\
+I thank Dr.
+Ir.
+\textit{Cornelis A.J.M.
+de Bie} from the Department of Natural Resource, ITC, for his patient sharing of iPAQ knowledge during my Polish fieldwork in summer 2010.
+I used this skills on Crete.
+\vspace{1mm}\\
+Many thanks for Dr.\textit{Dingtian Yang} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectances of Thalassia seagrass, which enabled to compare spectra of various seagrass types.
+
 \pagebreak
@@
 		the author, and do not necessarily represent those of the university.\\
+
+\vspace{3em}
+
+		I herewith certify that although I confered with other researchers while preparing this document, and used a range of various research and literature 			resources, cited in the text, the content of this MSc Thesis document remains my own, original work, written by me personally.
+\\
+
+\vspace{1em}
+
+		Signed.................................(Lemenkova Polina) \\
+
+\vspace{1em}
+
+		Enschede, The Netherlands. \\ 18th February 2011.
 \pagebreak
@@
 \section{Introduction}
-%%%%%%%%%%%%%%%%%%Section1%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%  Section 1   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \counterwithin{figure}{section}
@@
 \pagebreak
-%%%%%%%%%%%%%%%%%Section2%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%  Section2   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section[Overview of literature]{Seagrass monitoring: overview of literature and research resources} \label{sec:2}
@@
 
-%%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Materials and methods}
@@
 We mentioned in \nameref{sec:0} the software that we used for logically separated research sections, which required various approach.
-Thus, for example,  spectral analysis and assessment of spectral signaturesis technically based on WASI, sttistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using\href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
+Thus, for example,  spectral analysis and assessment of spectral signaturesis technically based on WASI, sttistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://gretl.sourceforge.net/}{Gretl} and \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using\href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
 \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0} software is used for the spatial analysis, general mapping and cartographic layout
@@
 (Xerocampus) parts of Crete island, as these regions are well suitable for the seagrass, due to the
-annual mean water temperatures and geological factors, i.e. seafloor conditions and sediments. The
+annual mean water temperatures and geological factors, i.e. seafloor conditions and sediments. \\ The
 field campaign has been carried out during the September-October period 2010. 
@@
 \subsection{Study area}
-General area: Island of Crete, Greece (Fig.\ref{fig:16}).\\
+General area: Island of Crete, Greece (Fig.\ref{fig:3.1}).\\
 The study area of the current MSc research is located in the shallow areas of the Ligaria beach, Agia
-Pelagia and Xerocampos, Crete island, Greece (Fig.\ref{fig:17}). These shelf areas have maximal depth of four meters.
+Pelagia and Xerocampos, Crete island, Greece (Fig.\ref{fig:3.2}). These shelf areas have maximal depth of four meters.
 Seagrass sampling has been performed at two stations at a depth of 4 meters, in the following selected
@@
 	\caption{Study area: Crete island}
-	\label{fig:18}
+	\label{fig:3.1}
 \end{figure}
@@
 	\caption{Study area: Ligaria beach, Crete island}
-	\label{fig:19}
-\end{figure}
-
-The Ligaria Beach is a narrow, sandy and pebble beach (Fig.12), located ca 15 km north-west from Heraklion.
+	\label{fig:3.2}
+\end{figure}
+
+The Ligaria Beach is a narrow, sandy and pebble beach (Fig.\ref{fig:3.2}), located ca 15 km north-west from Heraklion.
 The in-situ measurements were conducted during the fieldwork in the period 21.09.2010-11.10.2010.
@@
 \subsubsection{Fieldwork equipment}
-The research materials and equipment were provided by \href{http://www.nhmc.uoc.gr/index-2.htm}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:18}): 1) three iPAQs; 2)
-three GPS; 3) Three underwater video cameras, Olympus ST 8000, suitable for
+The research materials and equipment were provided by \href{http://www.nhmc.uoc.gr/index-2.htm}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:3.3}): 1) three iPAQs (Fig.\ref{fig:3.3a}); 2)
+three GPS (Fig.\ref{fig:3.3c}); 3) Three underwater video cameras, Olympus ST 8000  (Fig.\ref{fig:3.3d}), suitable for
 photographing up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
-Markers and cords for the depths measurements; 5) Waterproof plasctic Otterbox for keeping the
-iPAQ dry; 6) \href{http://www.padi.com/scuba/}{SCUBA Diving equipment}, taken from the Ligaria Diving Centre; 7) Boat (Fig.\ref{fig:19});
+Markers and cords for the depths measurements; 5) Waterproof plasctic Otterbox  (Fig.\ref{fig:3.3b})for keeping the
+iPAQ dry; 6) \href{http://www.padi.com/scuba/}{SCUBA Diving equipment}, taken from the Ligaria Diving Centre; 7) Boat (Fig.\ref{fig:3.4});
 
@@
 	\centering
-	\subfloat [iPAQ, HP]{\label{fig:20a}\includegraphics[width=0.3\textwidth]{Fig-19a.jpg}}
-	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:20b}\includegraphics[width=0.3\textwidth]{Fig-19b.jpg}}
-	\subfloat [GPS]{\label{fig:20c}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
-	\subfloat [Olympus waterproof camera]{\label{fig:20d}\includegraphics[width=0.25\textwidth]{Fig-19d.jpg}}
+	\subfloat [iPAQ, HP]{\label{fig:3.3a}\includegraphics[width=0.3\textwidth]{Fig-19a.jpg}}
+	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:3.3b}\includegraphics[width=0.3\textwidth]{Fig-19b.jpg}}
+	\subfloat [GPS]{\label{fig:3.3c}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
+	\subfloat [Olympus waterproof camera]{\label{fig:3.3d}\includegraphics[width=0.25\textwidth]{Fig-19d.jpg}}
 	\caption{Fieldwork equipment}
-	\label{fig:20}
+	\label{fig:3.3}
 \end{figure}
@@
 	\caption{Boat used for the fieldwork measurements on Crete island. 2010}
-	\label{fig:21}
+	\label{fig:3.4}
 \end{figure}
@@
 	\caption{Scheme of placements of the Olimpus cameras during the measurements}
-	\label{fig:22}
+	\label{fig:3.5}
 \end{figure}
@@
 The videographic measurements and photos, enabling to detect types of the
-seafloor, were captured by means of three underwater digital video cameras Olympus ST 8000 (Fig.\ref{fig:18}), at
-five minutes interval along the tracklog path (Fig.\ref{fig:21}).
-research sampling design also included measurements of depths (Fig. \ref{fig:63}), because bathymetry is one of the
+seafloor, were captured by means of three underwater digital video cameras Olympus ST 8000  (Fig.\ref{fig:3.3d}), at
+five minutes interval along the tracklog path (Fig.\ref{fig:3.6}).
+research sampling design also included measurements of depths (Fig. \ref{fig:A.4}), because bathymetry is one of the
 most determinating factors for the seagrass locations. Measurements of depths were performed during
@@
 with videographic measurements. The GPS allowed to capture measurement locations on the iPAQ,
-encapsulated in a plastic waterproof Otterbox (Fig \ref{}). Camera were adjusted horizontally by a leveller and
-were mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:20}). The data
+encapsulated in a plastic waterproof Otterbox (Fig \ref{fig:3.3b}). Camera were adjusted horizontally by a leveller and
+were mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:3.5}). The data
 were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state. The
@@
 coverage of the seagrass meadows in the research area. The underwater measurements of the seagrass
-coverage were carried out by taking video footage and photos of ca 0.5 m2 size each (Fig \ref{66}).
+coverage were carried out by taking video footage and photos of ca 0.5 m2 size each (Fig \ref{fig:A.7}).
 The results of the underwater videometric measurements include series of digital images helping to
@@
 	\caption{Locations of the video measurements and GPS tracklogs, Ligaria}
-	\label{fig:23}
+	\label{fig:3.6}
 \end{figure}
@@
 Along the Ligaria beach there are several types of the seagrass landscapes, namely, segrass meadows
-continuously covering vast areas (Fig.\ref{fig:62} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:62} (b)) and isolated seagrass patches, or mattes,
-which located separately from broad meadows (Fig.\ref{fig:62} (c)).
+continuously covering vast areas (Fig.\ref{fig:A.3} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:A.3} (b)) and isolated seagrass patches, or mattes,
+which located separately from broad meadows (Fig.\ref{fig:A.3} (c)).
 The results of the underwater videometric measurements of the Olympus cameras made during the
@@
 information on seagrass presence within the study area, distribution of seagrass P.oceanica meadows
-and nature of the seafloor cover types: rocks, sandy, mixed (Fig. \ref{fig:65}). Seagrass species on the Ligaria and Agia
+and nature of the seafloor cover types: rocks, sandy, mixed (Fig. \ref{fig:A5}). Seagrass species on the Ligaria and Agia
 Pelagia beach consist of P.oceanica. The types of sediment on the Ligaria vary from coarse sand,
-P.oceanica patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:64}).
+P.oceanica patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:A.6}).
 
@@
 Optical spectra of P.oceanica, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig.
-\ref{fig:23})
+\ref{fig:A.23})
 	\item[]2. Aerial imagery from the \href{http://www.google.com/earth/index.html}{Google Earth}
-	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:68})
+	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:A.37})
 	\item[]4. Results of underwater videometric measurements of the Olympus cameras made during the ship route
 \end{itemize}
-The available satellite imagery (Fig.\ref{fig:68}) and aerial images are read into the ArcGIS project.
+The available satellite imagery (Fig.\ref{fig:A.38}) and aerial images are read into the ArcGIS project.
 The available broadband and hyperspectral remote sensing data are used for the mapping of the
@@
 UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.
-\ref{fig:10}), both adjusted for the measurements of the irradiance and radiance (see Appendices: Tab.
-\ref{tab:18}, Tab.
-\ref{tab:19}).
-The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:49}) and 430-650 for the water radiance (Fig.\ref{fig:47}). 
+\ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see Appendices: Tab.
+\ref{tab:A.1}, Tab.
+\ref{tab:A.2}).
+The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:A.18}) and 430-650 for the water radiance (Fig.\ref{fig:A.16}). 
 Afterwards, the measured values of the
 radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
-properties of the sea water and bottom cover types (Fig.\ref{fig:23}). The spectral range of radiance
+properties of the sea water and bottom cover types (Fig.\ref{fig:A.15}). The spectral range of radiance
 cover diapason of 320-950 nm, and ittadiance measurements are covered in the interval of 280-500
 nm, which is suitable for a characteristics of seagrass reflectance.
-Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:43}, or \ref{fig:45}, enlarged) represent several series of
+Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:A.12}, or \ref{fig:A.14}, enlarged) represent several series of
 the measurements. The values of the spectral reflectance are received from the
-computations of these values using Excel mathematical formulae.
-The graphs shown on Fig.\ref{fig:47}, Fig.
-\ref{fig:48}, Fig.
-\ref{fig:49} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively.
-Graph on Fig \ref{fig:44} displays statistuical analysis of the measured sets of observations, i.e.
+computations of these values using mathematical formulae.
+The graphs shown on Fig.
+\ref{fig:4.2}, Fig.
+\ref{fig:A.23}, Fig.
+\ref{fig:A.24}, Fig.
+\ref{fig:A.17} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively.
+Graph on Fig \ref{fig:A.13} displays statistuical analysis of the measured sets of observations, i.e.
 shows the midspread of the statistical quartiles (Q1-Q3), mean and extreme values.
@@
 	\includegraphics[scale=0.5]{GNU-17.jpg}
-	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:25}
+	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}				{Gnuplot}}\label{fig:4.2}
 \end{figure}
@@
 \subsection{Data pre-processing}
-A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonatic sand - consist of multiple measurement sets of P.oceanica made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October, 1-445 for 14th October for P.oceanica (Tab.
-\ref{tab:10}); 1-106 for water without sedimnts, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
+A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonatic sand - consist of multiple measurement sets of P.oceanica made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
+\ref{tab:A.8}), 1-445 for 14th October for P.oceanica; 1-106 for water without sedimnts, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
 \ref{tab:17}).
@@
 We visualised the behavour of different spectra on relevant graphs (Fig.
-\ref{fig:43}, \ref{fig:44}, \ref{fig:45}).
+\ref{fig:4.4}, \ref{fig:A.35}, \ref{fig:A.28}).
 The available collected data were tested for the spectral variability and separability under varying conditions of dfifferent environmental
@@
 \ref{tab:8}, Tab.
-\ref{tab:9}).
+\ref{tab:A.7}).
 The raw
-initial data of measurements of the spectral reflectance have been preprocessed and statistically analysed thereafter withgraphs visualising mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:33}, \ref{fig:28}).
+initial data of measurements of the spectral reflectance have been preprocessed and statistically analysed thereafter with graphs visualising mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:4.10}, \ref{fig:28}).
 These data
@@
 were made with 3 nm interval, while others – using 4 and 1 nm step.
-Therefore, these data had to be interpolated and standardized to one
+Therefore, these data had to be interpolated (Fig. \ref{fig:A.10}) and standardized to one
 format, which is values of spectral reflectancese with one nm step. For the
 interpolation we used script written on Python programming language that allowed to receive more 
-detailed data by interpolating them from 3 and 4 nm step up to 2 nm (Fig.\ref{fig:24}).
+detailed data by interpolating them from 3 and 4 nm step up to 2 nm (Fig.\ref{fig:A.8}).
 The interpolated data contained spectral measurements of the seagrass P.oceanica (measured at Agia
 Pelagia beach, Crete), sand (measured at Agia Pelagia beach, Crete), silt and default artificial
-spectrum of constant albedo at WASI (Fig.\ref{fig:25}). 
+spectrum of constant albedo at WASI (Fig.\ref{fig:A.9}). 
 We included measurements of the Thalassia seagrass
@@
 the seagrasses. Therefore, we chosen the spectra 400-750 as the most appropriate range for
-further research experiment.The results of the linear interpolation (Fig.\ref{fig:25}) demonstrate values of the
+further research experiment.The results of the linear interpolation (Fig.\ref{fig:A.35}) demonstrate values of the
 sand spectral reflectances with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
 The initial measured data were stored in raw-oriented format, so that re-formatting them into the
-column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:42}). The next step included
-calculation of the median, mean, quartiles and other statistical values at every data set (see Appendix, Fig. \ref{fig:54}, Fig. \ref{fig:55}).
-After the preliminary analysis, the measured data were visualised using GNUplot programm, which
-enables fine plotting of various datasets together: Fig \ref{fig:26}, Fig.\ref{fig:27}, Fig.\ref{fig:28}.
+column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:A.11}). The next step included
+calculation of the median, mean, quartiles and other statistical values at every data set (see Appendix, Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
+After the preliminary analysis, the measured data were visualised using Gnuplot programm, which
+enables fine plotting of various datasets together: Fig \ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:28}.
 The most acceptable method of interpolation was B\'ezier curve (Fig.
-\ref{fig:47}, Fig.
-\ref{fig:49}),
+\ref{fig:A.18}, Fig.
+\ref{fig:A.16}),
 as it has trend-friendly graph better showing the general behavior of the curves at different wavebands
@@
 Therefore, after several experiments with
-various interpolation techniques (Fig.\ref{fig:48}), we have choosen Bézier curves interpolation (Fig.\ref{fig:49}) which contain convex
+various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have choosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contain convex
 hulls made on its control points, and therefore is best suitable for our case: analysis of optical
@@
 The distinguishing spectral signatures for various seafloor types (e.g. seagrass species, coral reefs,
-various types of sand, mud, other sediments, (Fig.\ref{fig:26})) exist in well-defined and narrow (10–20 nm)
+various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})) exist in well-defined and narrow (10–20 nm)
 wavelength ranges.
@@
 \href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentray graphs below show the results of the statistical analysis.}
-	\label{fig:26}
+	\label{fig:4.4}
 \end{figure}
@@
 suitable for the seagrass spectral analysis.
-There are several environmental characteristics, included in WASI interface (Fig.\ref{fig:29}), which influence the
+There are several environmental characteristics, included in WASI interface (Fig.\ref{fig:4.7}), which influence the
 results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
@@
 water, large and small particles. The values of all these parameters can be redacted and changed
-manuall. However, the most important, major factors affecting the in-situ conditions are water depth
-and cover fraction of the seafloor types: P.oceanica and carbonatic sand (Fig.\ref{fig:27}).
+manuall. However, the most important, major factors affecting the in-situ conditions are water depth (Fig.\ref{fig:4.9})
+and cover fraction of the seafloor types: P.oceanica and carbonatic sand .
 
@@
 	\rowcolors{1}{LightSteelBlue}{Honeydew}
-	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, with average values (for sets 1 - 350). Generalisation up to step 20 nm}
+	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, with average values (for sets 1 - 350). Generalisation up to step 20 nm}\label{tab:4.1}
 	\begin{center}
@@
 	\end{center}
-	\label{tab:2}
 \end{table}
@@
 seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
-remote sensing sensors (MODIS, ASTER, MERIS and SeaWiFS), (Fig.\ref{fig:28}).
+remote sensing sensors (MODIS, ASTER, MERIS and SeaWiFS), (Fig.\ref{fig:4.6}).
 
@@
 	\caption{Simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
-	\label{fig:29}
+	\label{fig:4.6}
 \end{figure}
@@
 based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
-types, where the spectral signatures were examined. WASI software (Fig.\ref{fig:29}) is used to simulate
+types, where the spectral signatures were examined. WASI software (Fig.\ref{fig:4.7}) is used to simulate
 spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
@@
 	\caption{Main interface of WASI water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m}
-	\label{fig:30}
+	\label{fig:4.7}
 \end{figure}
@@
 45° and reflection factor 0.0201. P.oceanica. These values, simyulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) on the WASI user-defined parameters.
-The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:30}).
+The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:4.8}).
 
@@
 	\caption{Sea water physical properties, modelled by WASI}
-	\label{fig:31}
+	\label{fig:4.8}
 \end{figure}
@@
 the seagrass. Therefore, the most healthy and suitable areas for the seagrass grow are located at depths
-lesser than 4 meters (Fig.\ref{fig:31}).
+lesser than 4 meters (Fig.\ref{fig:4.9}).
 
@@
 	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths}
-	\label{fig:32}
+	\label{fig:4.9}
 \end{figure}
@@
 that the appropriate wavebands for seagrass mapping lay between 500 and 600 nm and has
-also peaks at around 700 nm, ca between 680 and 710 nm (Fig.\ref{fig:31}). The highest values of the bottom
+also peaks at around 700 nm, ca between 680 and 710 nm (Fig.\ref{fig:4.10}). The highest values of the bottom
 reflectance are at spectra of 500-600 nm. The most appropriate depths at which the spectral signatues
@@
 The patches of white sandy bottoms of the seafloor, studied in the fieldwork in Ligaria beach, are
-much brighter than mattes of P.Oceanica ((Fig.\ref{fig:64}): seafloor types in Ligaria), which can be clearly seen
-at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:65}) and that of P.
-oceanica (Fig.\ref{fig:62}).
+much brighter than mattes of P.Oceanica (Fig.\ref{fig:A.6}, \ref{fig:A.5}): seafloor types in Ligaria), which can be clearly seen
+at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:A.35}) and that of P.
+oceanica (Fig.\ref{fig:A.28}).
 The graph is received in excel spreadsheet using mean values of spectral reflectances of
-sand (Fig.\ref{fig:34}) and seagrass (Fig.\ref{fig:32}), respectively, which have been calculated from measurements of radiance and
+sand (Fig.\ref{fig:A.27}) and seagrass (Fig.\ref{fig:A.35}), respectively, which have been calculated from measurements of radiance and
 irradiance received in Agia Pelagia bay.
@@
 \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:33}
+	\label{fig:4.10}
 \end{figure}
@@
 	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
-	\label{fig:34}
+	\label{fig:4.11}
 \end{figure}
@@
 The total amount of measured data was large and included following datasets made using
-hyperspectral radiometer Ramses: 350 measuremetn sets of P.oceanica reflectance for 14th Oct (Tab.\ref{tab:32}), 400
-sets of P.oceanica reflectance for 15th Oct (Tab.\ref{tab:33}), 84 datasets for seawater reflectance with sediments, 105
+hyperspectral radiometer Ramses: 350 measuremetn sets of P.oceanica reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
+sets of P.oceanica reflectance for 15th Oct (Tab.\ref{tab:A.8}), 84 datasets for seawater reflectance with sediments, 105
 datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonat sand.
 A statistical approach is evidently necessary for the proper processing of such amounts of data. 
+The schematic view of the statistical approach used for the data processing  can be seen on Fig. \ref{fig:4.12}.
 Statistical pre-processing of large sets of serial data enables to generalyse data by using the most typical and predicted values of
@@
 The Open Office was used for preliminary data view and preprocessing, and included following computations
-at each data set (for example, Tab.\ref{tab:31}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). \\
+at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). 
 The statistical pre-processing and analysis were made to display the mean values of spectral reflectances,
 which were used for comparison with reflectance data of various seafloor cover types as well as
-seawater with and without sediment (e.g. Tab.\ref{tab:2}).
+seawater with and without sediment (e.g. Fig..\ref{fig:A.16}).
 In the statistical anlysis of the raw observed data pool of the values of spectral reflectances (about 400 single measurements for P.oceanica, e.g.
-Tab.\ref{tab:10}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
+Tab.\ref{tab:A.8}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
 parameters.
-Examples are the mean and variance of a probability distribution of measurements dataset (for example, Tab.\ref{tab:16}).
+Examples are the mean and variance of a probability distribution of measurements dataset (for example, Tab.\ref{tab:4.1}).
 We divided the total data pool into datasets of 25 measurements for more convenience and ease of use, both for P.oceanica and carbonate sand.
@@
 	\centering
-	\includegraphics[scale=0.45, angle=90]{UML_StatAn.jpg}
+	\includegraphics[scale=0.43, angle=90]{UML_StatAn.jpg}
 	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
-	\label{fig:35}
-\end{figure}
+	\label{fig:4.12}
+\end{figure}
+\pagebreak
 
 Thus, we collect a number of realizations (a sample of 25 measurements for both sand and P.oceanica) and then estimate the statistical parameters (Appendix). \\
-Finally we analysed abount 400 datasets totally (example of carbonate sand on Fig.\ref{tab:31}) made during different days. 
+Finally we analysed abount 400 datasets totally (example of carbonate sand on Fig.\ref{fig:4.14}, P.oceanica - on  Fig.\ref{fig:4.13}) made during different days. 
 Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random. 
@@
 A probability distribution, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
-\\
 The \textit{least absolute deviations (LAD)}, a popular optimization technique, was used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
@@
  shows the similarity between observations as a function of the time separation between them.
-As can be seen on Fig.\ref{fig:52}, Fig.\ref{fig:53} and Fig.\ref{fig:60}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
+As can be seen on Fig.\ref{fig:A.21}, Fig.\ref{fig:A.22} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
 The autocorrelation analysis has been performed using both SPSS and Gretl software.
@@
 To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectances at several depths: Tab.\ref{tab:5}, Tab.\ref{tab:6}, Tab.\ref{tab:7}.
-The summary of the values of spectral reflectance properties on various depths is presented on the Fig.\ref{fig:17}.
+The summary of the values of spectral reflectance properties on various depths is presented on the Tab.\ref{tab:A.7}.
 The \textit{k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
@@
 which corresponds to the general spectrum of the observations with representation of a variable quantity.  
-Thus, the periodogram of the series of observation highlights so-called “dead spots” of low power between the frequencies we wishe to
-exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:43} ) 
-and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:45}, enlarged part of 500-660nm).\\
+Thus, the periodogram (Fig. \ref{fig:A.31}) of the series of observation highlights so-called “dead spots” of low power between the frequencies we wishe to
+exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:A.12} ) 
+and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:A.14}, enlarged part of 500-660nm).\\
 The visualization of the data plotting was made by means of Gnuplot software, which enabels fine
 drawing and advanced displaying of large amounts of serial data, ultimate control over graph properties, 
-the simplicity of plotting and the ease of scripting. (see Fig.\ref{fig:54}, Appendices).
-
+the simplicity of plotting and the ease of scripting. (see Fig.\ref{fig:4.13}, Appendices).
 The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
-areas of quartiles (Q1, Q3, shaded areas on graph, Fig.\ref{fig:55}), which cover most probable data distribution
+areas of quartiles (Q1, Q3, shaded areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution
 (spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
@@
 more than 0.20 nm, except for borders), we did not extend the y-scale up to 100 percent, in order to
-focus on the most necessary area of spectral values (Fig.\ref{fig:35} and Appendices). The mean values are
+focus on the most necessary area of spectral values (Fig.\ref{fig:4.14}, Fig.\ref{fig:4.4} and Appendices). The mean values are
 highligted using bold red points, as we used these values for plotting the final graph of P.oceanica spectra in
@@
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
-		\label{fig:36}
+		\label{fig:4.13}
 	\end{center}
@@
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
-		\label{fig:37}
+		\label{fig:4.14}
 	\end{center}
@@
 Seagrass beds are clearly visible in color aerial Google Earth
-photographs (Fig.\ref{fig:35}), contrasting against slightly-yellow and browny sand bottom. The seagrass areas are
+photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and browny sand bottom. The seagrass areas are
 detected using different bands combinations, masked and studied for the estimation of the changes in
 the areas. The classification is based on the spectral properties of the \textit{P.oceanica}, such as brightness,
-colour, texture and structure of the seagrass mattes (Fig.\ref{fig:35}a). The raster-based mapping
+colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping
 includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
@@
 bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
-seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:36}).
+seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).
 
@@
 		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout}
-		\label{fig:40}
+		\label{fig:4.17}
 	\end{center}
@@
 spectral signatures using WASI software, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
-based mapping. The use of GIS for data incorporation (Fig.\ref{fig:37}), storage, analyses, visualizing and mapping
+based mapping. The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
 enables to analyse environmental changes within seagrass landscapes based on data from various
@@
 This research part results in series of thematic maps showing seagrass spatial distribution over the Island of
-Crete. The research is based on the aerial images classification and analysis using ArcGIS and Envi
-software (Fig.24). The map of the seagrass locations within the research area of Ligaria beach illustrates the seagrass
+Crete. The research is based on the aerial images classification and analysis using ArcGIS and Erdas Imagine
+software. The map of the seagrass locations within the research area of Ligaria beach illustrates the seagrass
 distribution at a more detailed scale.
@@
 
-%%%%%%%%%%%%%%%%%%%  Appendix  %%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%  Appendix  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{appendices}
@@
 \appendixpage
-\addappheadtotoc
-%%%
 \counterwithin{figure}{section}
@@
 
-\section*{}
-\appendixname{\textbf{1:\\Results of the measurements of seafloor spectral reflectance}}
-%%%
+\section{appendices}
+
+\subsection{\textit{Appendix-1}. Capturing aerial imagery from the Google Earth: grabbing process}
+%\appendixname{\textbf{1: \\Capturing aerial imagery from the Google Earth: grabbing process}}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.6]{Fig-22.jpg}
+	\caption{Capturing aerial imagery from the Google Earth: grabbing process}
+	\label{fig:A.1}
+\end{figure}
+\pagebreak
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.4]{Script_Gdal.jpg}
+		\caption{Script command of FWTools2.4.7 enabling to reduce the size of the aerial images, from .tif to .ecw format.}
+	\end{center}
+	\label{fig:A.2}
+\end{figure}
+\pagebreak
+
+\subsection{\textit{Appendix-2}. Instrumental adjustment and tuning (Trios-RAMSES setup)}
+%\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}
+
+\begin{table}[H]
+	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}\label{tab:A.1}
+	\centering
+	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
+	  \begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
+		\hline
+		    \textbf{Parameters} & \textbf{15 October, set 1} & \textbf{14 October, set 4} \\ \hline \hline
+		     Version &1 & 1 \\ \hline
+		     IDData & 78B1-2009-10-15-08-56-22-280-272 & 78B1-2009-10-14-15-15-59-342-726 \\ \hline
+		     IDDevice & SAM-820C & SAM-8204\\ \hline
+		     IDDataType & SPECTRUM & SPECTRUM \\ \hline
+		     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
+		     DateTime & 2009-10-15 08:56:22 & 10/14/2009, 15:15:59 \\ \hline
+		     PositionLatitude & 35.2451=35°14'29.4354” & 35.24752 = 35° 14' 51.072" \\ \hline
+		     PositionLongitude & 25.01269=25°0'45.6834" & 25.0098 = 25°0'35.2794" \\ \hline
+		     Comment & P.oceanica & P.oceanica \\ \hline
+		     CommentSub1 & 0.5m depth & 2.5m depth \\ \hline
+		     CommentSub2 & diffuse & diffuse \\ \hline
+		     CommentSub3 & Agia Pelagia & Agia Pelagia site 2, close to rocks \\ \hline
+		     IDMethodType & SAM Control & SAM Control \\ \hline
+		    MethodName & SAM-820C & SAM-8204 \\ \hline
+		    Mission & No Mission & No Mission \\ \hline
+		    MissionSub & 1 & 1 \\ \hline
+		    RecordType & 0 & 0 \\ \hline
+		    CalFactor & 1 & 1 \\ \hline
+		   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
+		   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
+		   IntegrationTime & 512 & 1024 \\ \hline
+		    P31 & -1 & -1 \\ \hline
+		   P31e & 0 & 0 \\ \hline
+		   PathLength & +INF & +INF \\ \hline
+		   RAWDynamic & 65535 & 65535 \\ \hline
+		   Temperature & +NAN & +NAN \\ \hline
+		   Unit1 & 0101 Wavelength nm &  1 1 Wavelength nm \\ \hline
+		   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
+		   Unit3 & f0-06 Error mW/(m2 nm) & 3-3 Error mW/(m2 nm Sr) \\ \hline
+		  Unit4 & f1-00 Status & f1-00 Status \\ \hline
+	  \end{tabular}	
+\end{table}
+\pagebreak
+
+\begin{table}[H]
+	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}\label{tab:A.2}	\centering
+	\rowcolors{1}{Lavender}{GhostWhite}
+	\begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
+	\hline
+	    \textbf{Parameters} & \textbf{14 October, set 1} & \textbf{15 October, set 4} \\ \hline \hline
+	     Version &1 & 1 \\ \hline
+	     IDData & 78B1-2009-10-14-15-32-02-078-323 & 78B1-2009-10-15-10-36-02-217-136 \\ \hline
+	     IDDevice & SAM-820C & SAM-8204\\ \hline
+	     IDDataType & SPECTRUM & SPECTRUM \\ \hline
+	     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
+	      DateTime & 2009-10-14 15:32:02 & 2009-10-15 10:36:02 \\ \hline
+	     PositionLatitude & 35.24752= 35°14'51.072" & 35.245071=35°14'42.2556" \\ \hline
+	     PositionLongitude & 25.0098= 25°0'35.2794" & 25.012661=25°0'45.5796" \\ \hline
+	     Comment & P.oceanica & P.oceanica \\ \hline
+	     CommentSub1 & 3.5m depth & 1.5m depth \\ \hline
+	     CommentSub2 & diffuse & diffuse \\ \hline
+	     CommentSub3 & Agia Pelagia & Agia Pelagia site 1, close to rocks \\ \hline
+	     IDMethodType & SAM Control & SAM Control \\ \hline
+	    MethodName & SAM-820C & SAM-8204 \\ \hline
+	    Mission & No Mission & No Mission \\ \hline
+	    MissionSub & 1 & 1 \\ \hline
+	    RecordType & 0 & 0 \\ \hline
+	    CalFactor & 1 & 1 \\ \hline
+	   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
+	   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
+	   IntegrationTime & 64 & 32 \\ \hline
+	    P31 & -1 & -1 \\ \hline
+	   P31e & 0 & 0 \\ \hline
+	   PathLength & +INF & +INF \\ \hline
+	   RAWDynamic & 65535 & 65535 \\ \hline
+	   Temperature & +NAN & +NAN \\ \hline
+	   Unit1 & 01-01 Wavelength nm &  01-01 Wavelength nm \\ \hline
+	   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
+	   Unit3 & f0-06 Error mW/(m2 nm) & f0-03 Error mW/(m2 nm Sr) \\ \hline
+	  Unit4 & f1-00 Status & f1-00 Status \\ \hline
+	\end{tabular}
+\end{table}
+\pagebreak
+
+\subsection{\textit{Appendix-3}. Types of seagrass structural patterns, Ligaria beach, Crete}
+%\appendixname{\textbf{3: \\Types of seagrass structural patterns, Ligaria beach, Crete}}
+
+\begin{figure}[h]
+	\centering
+	\subfloat [Fragment of meadow]{\includegraphics[width=0.3\textwidth]{Fragment_meadow.jpg}}
+	\hspace{1mm}
+	\subfloat [Aggregated seagrass patch]{\includegraphics[width=0.3\textwidth]{aggreg_patch.jpg}}
+	\hspace{1mm}
+	\subfloat [Isolated patch of seagrass]{\includegraphics[width=0.3\textwidth]{isolated_patch.jpg}}
+	\caption{Types of seagrass structural patterns, Ligaria beach, Crete}
+	\label{fig:A.3}
+\end{figure}
+
+\subsection{\textit{Appendix-4}. Sampling design of the depths measurements. Ligaria beach, Crete}
+%\appendixname{\textbf{4: \\Sampling design of the depths measurements. Ligaria beach, Crete}}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[scale=0.35]{Sampling_design.jpg}
+		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}\label{fig:A.4}
+	\end{center}	
+\end{figure}
+\pagebreak
+
+\subsection{\textit{Appendix-5}. Results of videographic measurements: seafloor types on Crete island}
+%\appendixname{\textbf{5: \\Results of videographic measurements: seafloor types on Crete island}}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[scale=0.45]{Fig-3-14.jpg}
+		\caption{Ligaria beach, Crete: seafloor types}\label{fig:A.5}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[scale=0.25]{Fig-3-15.jpg}
+		\caption{Various seafloor types; Ligaria beach, Crete}\label{fig:A.6}
+	\end{center}
+\end{figure}
+\pagebreak
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[scale=0.35]{Fig-3-16.jpg}
+		\caption{Measurement underwater equipment}\label{fig:A.7}		
+	\end{center}
+\end{figure}
+\pagebreak
+
+\subsection{\textit{Appendix-6}. Data pre-processing}
+%\appendixname{\textbf{6: \\Data pre-processing}}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{Fig-24.jpg}
+	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
+	\label{fig:A.8}
+\end{figure}
+\pagebreak
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{Fig-28.jpg}
+	\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
+		seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
+	\label{fig:A.9}
+\end{figure}
+\pagebreak
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.35]{Interpolation.jpg}
+		\caption{Interpolation of the spectral measurements by means of Open Office}
+		\label{fig:A.10}
+	\end{center}
+\end{figure}
+\pagebreak
+
+\subsection{\textit{Appendix-7}. Data processing and statistical evaluation: Results of the measurements of the seafloor optical properties.}
+%\appendixname{\textbf{7:\\ Data processing and statistical evaluation: \\ Results of the measurements of the seafloor optical properties.}}
 
@@
 		\caption{Values of spectral reflectances of seagrass P.oceanica (fragment)}
-		\label{fig:A-1}
+		\label{fig:A.11}
 	\end{center}
@@
 		\caption{Spectral reflectance of P.oceanica. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
-		\label{fig:44}
+		\label{fig:A.12}
 	\end{center}
@@
 \href{http://www.gnuplot.info/}{Gnuplot}}
-		\label{fig:45}
+		\label{fig:A.13}
 	\end{center}
@@
 		Visualisation of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
-		\label{fig:46}
+		\label{fig:A.14}
 	\end{center}
@@
 		Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:47}
+	\label{fig:A.15}
 	\end{center}
@@
 	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:48}
+	\label{fig:A.16}
 \end{figure}
@@
 	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
-\href{http://www.spss.com/}{SPSS}}
+\href{http://www.spss.com/}{SPSS}}\label{tab:8}
 	\begin{center}
@@
 	P more than .05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5).
-	\label{tab:8}
+
 \end{table}
@@
 	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
-\href{http://www.spss.com/}{SPSS}}
+\href{http://www.spss.com/}{SPSS}}\label{tab:A.7}
 	\begin{center}
@@
 	P$>$0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5).
- 	\label{tab:9}
+
 \end{table}
@@
 \begin{longtable}{|c|c|c|c|c|c|c|}
-	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October}\label{tab:10} \\
+	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October}\label{tab:A.8} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 
-	\multicolumn{7}{c}{Table 10: Spectral reflectance of \textbf{P.oceanica} \textit{(continued from previous page)}}\\
+	\multicolumn{7}{c}{Table A.8: Spectral reflectance of \textbf{P.oceanica} \textit{(continued from previous page)}}\\
 	\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
@@
 		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}­}
-		\label{fig:48}
+		\label{fig:A.17}
 	\end{center}
@@
 	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:49}
+	\label{fig:A.18}
 \end{figure}
@@
 		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}}
-		\label{fig:50}
+		\label{fig:A.19}
 	\end{center}
@@
 	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}­}
-	\label{fig:51}
+	\label{fig:A.20}
 \end{figure}
@@
 	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
-	\label{fig:52}
+	\label{fig:A.21}
 \end{figure}
@@
 	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
-	\label{fig:53}
+	\label{fig:A.22}
 \end{figure}
@@
 \newpage
-
-\appendixname{\textbf{2: \\Graphs of the measurements of the spectral reflectance of P.oceanica}}
 
@@
 		\caption{Spectral reflectance of the seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle. \href{http://www.gnuplot.info/}{Gnuplot}­}
-		\label{fig:54}
+		\label{fig:A.23}
 	\end{center}
@@
 		\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. \href{http://www.gnuplot.info/}{Gnuplot} display­}
-		\label{fig:55}
+		\label{fig:A.24}
 	\end{center}
@@
 %%% Gretle end
-
 \pagebreak
@@
 		\caption{Spectral reflectance of P.oceanica. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
-		\label{fig:56}
+		\label{fig:A.25}
 	\end{center}
@@
 \href{http://www.gnuplot.info/}{Gnuplot}­}
-		\label{fig:57}
+		\label{fig:A.26}
 	\end{center}
@@
 \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
-		\label{fig:58}
+		\label{fig:A.27}
 	\end{center}
@@
 \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
-		\label{fig:59}
+		\label{fig:A.28}
 	\end{center}
@@
 	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand. Series 51-75, variable 27. \href{http://gretl.sourceforge.net/}{Gretl} visualization­}
-	\label{fig:60}
+	\label{fig:A.29}
 	\end{center}
@@
 		\caption{Normal Q-Q plot: estimated versus observed values of the measurement of carbonate sand, variable 27. Series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\label{fig:61}
+	\label{fig:A.30}
 	\end{center}
@@
 		\includegraphics[scale=0.25]{G-period-sand30.jpg}
-		\caption{Plot illustrating periodogramm for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\end{center}
-	\label{fig:62}
+		\caption{Plot illustrating periodogramm for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
+	\end{center}	
 \end{figure}
@@
 	\end{center}
-	\label{fig:63}
+	\label{fig:A.32}
 \end{figure}
@@
 	\end{center}
-	\label{fig:64}
+	\label{fig:A.33}
 \end{figure}
@@
 	\end{center}
-	\label{fig:65}
+	\label{fig:A.34}
 \end{figure}
@@
 Wavelength step: 3 nm.
-Measured on Agia Pelagia beach, 14th October}\label{tab:31} \\
+Measured on Agia Pelagia beach, 14th October}\label{tab:A.29} \\
  		 \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 
-	\multicolumn{7}{c}{Table 5: Spectral reflectance of \textbf{carbonate sand} \textit{(continued from previous page)}}\\
+	\multicolumn{7}{c}{Table A.29: Spectral reflectance of \textbf{carbonate sand} \textit{(continued from previous page)}}\\
 	\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}				{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
@@
 	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES. \href{http://www.gnuplot.info/}{Gnuplot}}
-	\label{fig:66}
+	\label{fig:A.35}
 \end{figure}
@@
 
-\begin{table}[H]
-	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}
-	\centering
-	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
-	  \begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
-		\hline
-		    \textbf{Parameters} & \textbf{15 October, set 1} & \textbf{14 October, set 4} \\ \hline \hline
-		     Version &1 & 1 \\ \hline
-		     IDData & 78B1-2009-10-15-08-56-22-280-272 & 78B1-2009-10-14-15-15-59-342-726 \\ \hline
-		     IDDevice & SAM-820C & SAM-8204\\ \hline
-		     IDDataType & SPECTRUM & SPECTRUM \\ \hline
-		     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
-		     DateTime & 2009-10-15 08:56:22 & 10/14/2009, 15:15:59 \\ \hline
-		     PositionLatitude & 35.2451=35°14'29.4354” & 35.24752 = 35° 14' 51.072" \\ \hline
-		     PositionLongitude & 25.01269=25°0'45.6834" & 25.0098 = 25°0'35.2794" \\ \hline
-		     Comment & P.oceanica & P.oceanica \\ \hline
-		     CommentSub1 & 0.5m depth & 2.5m depth \\ \hline
-		     CommentSub2 & diffuse & diffuse \\ \hline
-		     CommentSub3 & Agia Pelagia & Agia Pelagia site 2, close to rocks \\ \hline
-		     IDMethodType & SAM Control & SAM Control \\ \hline
-		    MethodName & SAM-820C & SAM-8204 \\ \hline
-		    Mission & No Mission & No Mission \\ \hline
-		    MissionSub & 1 & 1 \\ \hline
-		    RecordType & 0 & 0 \\ \hline
-		    CalFactor & 1 & 1 \\ \hline
-		   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
-		   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
-		   IntegrationTime & 512 & 1024 \\ \hline
-		    P31 & -1 & -1 \\ \hline
-		   P31e & 0 & 0 \\ \hline
-		   PathLength & +INF & +INF \\ \hline
-		   RAWDynamic & 65535 & 65535 \\ \hline
-		   Temperature & +NAN & +NAN \\ \hline
-		   Unit1 & 0101 Wavelength nm &  1 1 Wavelength nm \\ \hline
-		   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
-		   Unit3 & f0-06 Error mW/(m2 nm) & 3-3 Error mW/(m2 nm Sr) \\ \hline
-		  Unit4 & f1-00 Status & f1-00 Status \\ \hline
-	  \end{tabular}
-	\label{tab:32}
-\end{table}
-\pagebreak
-
-\appendixname{\textbf{3:\\ Google Earth aerial images covering research area of Crete island}}
+\subsection{\textit{Appendix-8}. Satellite and aerial images covering research area of Crete island: selected examples}
+%\appendixname{\textbf{8:\\ Satellite and aerial images covering research area of Crete island: selected examples}}
+
 \begin{figure}[H]
@@
 		\includegraphics[scale=0.20]{Locations_Meas.jpg}
-		\caption{Locations of selected measurements, visualization on Google Earth­}
-	\end{center}
-	\label{fig:67}
+		\caption{Locations of selected measurements, visualization on Google Earth­}\label{fig:A.36}
+	\end{center}	
 \end{figure}
@@
 		\includegraphics[scale=0.25]{App-8.jpg}
-		\caption{Random mosaic of selected aerial Google Earth images}
-	\end{center}
-	\label{fig:68}	
-\end{figure}
-\pagebreak
-
-\begin{table}[H]
-	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}
-	\centering
-	\rowcolors{1}{Lavender}{GhostWhite}
-	\begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
-	\hline
-	    \textbf{Parameters} & \textbf{14 October, set 1} & \textbf{15 October, set 4} \\ \hline \hline
-	     Version &1 & 1 \\ \hline
-	     IDData & 78B1-2009-10-14-15-32-02-078-323 & 78B1-2009-10-15-10-36-02-217-136 \\ \hline
-	     IDDevice & SAM-820C & SAM-8204\\ \hline
-	     IDDataType & SPECTRUM & SPECTRUM \\ \hline
-	     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
-	      DateTime & 2009-10-14 15:32:02 & 2009-10-15 10:36:02 \\ \hline
-	     PositionLatitude & 35.24752= 35°14'51.072" & 35.245071=35°14'42.2556" \\ \hline
-	     PositionLongitude & 25.0098= 25°0'35.2794" & 25.012661=25°0'45.5796" \\ \hline
-	     Comment & P.oceanica & P.oceanica \\ \hline
-	     CommentSub1 & 3.5m depth & 1.5m depth \\ \hline
-	     CommentSub2 & diffuse & diffuse \\ \hline
-	     CommentSub3 & Agia Pelagia & Agia Pelagia site 1, close to rocks \\ \hline
-	     IDMethodType & SAM Control & SAM Control \\ \hline
-	    MethodName & SAM-820C & SAM-8204 \\ \hline
-	    Mission & No Mission & No Mission \\ \hline
-	    MissionSub & 1 & 1 \\ \hline
-	    RecordType & 0 & 0 \\ \hline
-	    CalFactor & 1 & 1 \\ \hline
-	   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
-	   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
-	   IntegrationTime & 64 & 32 \\ \hline
-	    P31 & -1 & -1 \\ \hline
-	   P31e & 0 & 0 \\ \hline
-	   PathLength & +INF & +INF \\ \hline
-	   RAWDynamic & 65535 & 65535 \\ \hline
-	   Temperature & +NAN & +NAN \\ \hline
-	   Unit1 & 01-01 Wavelength nm &  01-01 Wavelength nm \\ \hline
-	   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
-	   Unit3 & f0-06 Error mW/(m2 nm) & f0-03 Error mW/(m2 nm Sr) \\ \hline
-	  Unit4 & f1-00 Status & f1-00 Status \\ \hline
-	\end{tabular}
-	\label{tab:33}
-\end{table}
-\pagebreak
-
-\appendixname{\textbf{4: \\Capturing aerial imagery from the Google Earth: grabbing process}}
-Script command of FWTools2.4.7 enabling to reduce the size of the aerial images, from .tif to .ecw
-format.
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.4]{Script_Gdal.jpg}
-		\caption{Script command of FWTools}
-	\end{center}
-	\label{fig:69}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{Fig-24.jpg}
-	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
-	\label{fig:24}
-\end{figure}
-
-\pagebreak
-\newpage
-
-\appendixname{\textbf{5: \\Types of seagrass structural patterns, Ligaria beach, Crete}}
-\begin{figure}[h]
-	\centering
-	\subfloat [Fragment of meadow]{\includegraphics[width=0.3\textwidth]{Fragment_meadow.jpg}}
-	\hspace{1mm}
-	\subfloat [Aggregated seagrass patch]{\includegraphics[width=0.3\textwidth]{aggreg_patch.jpg}}
-	\hspace{1mm}
-	\subfloat [Isolated patch of seagrass]{\includegraphics[width=0.3\textwidth]{isolated_patch.jpg}}
-	\caption{Types of seagrass structural patterns, Ligaria beach, Crete}
-	\label{fig:70}
-\end{figure}
-
-\appendixname{\textbf{6: \\Sampling design of the depths measurements. Ligaria beach, Crete}}
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.35]{Sampling_design.jpg}
-		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}
-	\end{center}
-	\label{fig:71}
-\end{figure}
-\pagebreak
-
-\appendixname{\textbf{7: \\Results of videographic measurements: seafloor types on Crete island}}
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.45]{Fig-3-14.jpg}
-		\caption{Ligaria beach, Crete: seafloor types}
-	\end{center}
-	\label{fig:72}
-\end{figure}
-
-\appendixname{\textbf{8: \\Various seafloor types on Crete island}}
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.25]{Fig-3-15.jpg}
-		\caption{Various seafloor types; Ligaria beach, Crete}
-	\end{center}
-	\label{fig:73}
-\end{figure}
-\pagebreak
-
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.35]{Fig-3-16.jpg}
-		\caption{Measurement underwater equipment}
-		\label{fig:74}
-	\end{center}
-\end{figure}
-\pagebreak
-
-\appendixname{\textbf{9: \\Data pre-processing: interpolation of the results of spectral measurements}}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.30]{Interpolation.jpg}
-		\caption{Interpolation of the spectral measurements}
-		\label{fig:75}
-	\end{center}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.4]{Fig-28.jpg}
-	\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
-		seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
-	\label{fig:25}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.6]{Fig-22.jpg}
-	\caption{Capturing aerial imagery from the Google Earth: grabbing process}
-	\label{fig:22}
-\end{figure}
+		\caption{Random mosaic of selected aerial Google Earth images}\label{fig:A.37}	
+	\end{center}	
+\end{figure}
+\pagebreak
 
 \begin{table}
-	\caption{Available broadband satellite images covering the research area of Crete island}
+	\caption{Available broadband Landsat satellite images covering the research area of Crete island}\label{tab:34}
 	\centering
@@
 	  \end{tabular}	
-	\label{tab:34}
 \end{table}
@@
 	\caption{Landsat imagery, Crete island. Previews}
-	\label{fig:76}
+	\label{fig:A.38}
 \end{figure}
@@
 \begin{thebibliography}{999}
-\addcontentsline{toc}{chapter}{Bibliography}
+\addcontentsline{toc}{chapter}{\textbf{Bibliography}}
 \bibliographystyle{plainnat}

### 2011-01-25 MScDiplom_Lemenkova.tex
tracked down and corrected misprints

@@
 SCUBA gear diving, instruments for underwater measurements, etc).
-I also appretiate their very warm, personal attitude and interest towards our research.
+I also appreciate their very warm, personal attitude and interest towards our research.
 \vspace{1mm}\\
@@
 I am thankful for all his explanations and articles given to me via e-mail \vspace{1mm}\\
-My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throuout my career.
+My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throughout my career.
 He encouraged me to use various very useful software and tools, always gave me necessary recommendations and precious help, thus highly contributing to my professional growth.
@@
 \vspace{1mm}\\
-I am greatful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery.
+I am grateful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery.
 \vspace{1mm}\\
@@
 	\textbf{Software and utilities used in the current work}:\\
+
+	\vspace{1ex}
+	\textbf{Main Tools}:
 	\begin{itemize}
@@
 		\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
-		\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data preprocessing and preliminary statistical evalution
-		\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
+		\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data pre-processing and preliminary statistical evaluation
+
 		\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
@@
 		\item \href{http://inkscape.org/}{Inkscape}: drawing
+	\end{itemize}
+
+	\vspace{1ex}
+	\textbf{Auxiliary Tools}:
+	\begin{itemize}
+		\item \href{http://www.lyx.org/}{LyX} for exporting \LaTeX document to the Word .doc format via HTML
+		\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
+		\item \href{http://openev.sourceforge.net/}{OpenEV}: used for raster Google Earth imagery viewing, visualisation and preliminary analysis
 		\item \href{http://www.notepad-plus-plus.org/}{Notepad++}: auxiliary text editing and raw data formatting
 	\end{itemize}
+
 \pagebreak
@@
 completely adapted to marine environment, they are however, more closely related to the lily family
-(Liliaceae) than to true grass, despite their name “seagrass”, caused by the ribbon-like, grassy leaves
+(Liliaceae) than to the true grass, despite their name “seagrass”, caused by the ribbon-like, grassy leaves
 \cite{McKenzie08}\label{McKenzie08}.\\
@@
 littoral coastal zones between 0-50 meters in shallow waters all over the world \cite{Hogarth07}\label{Hogarth07}, and
-serve as a valuable environmental indicators for the marine ecosystems health. Seagrasses are closely
+serve as valuable environmental indicators for the marine ecosystems health. Seagrasses are closely
 connected and linked with complex interactions to other vegetation types, e.g. mangroves, coral reefs,
@@
 their structure and functioning.\\
-The adaptation to the salt waters is evidently influenced the global ditribution of the seagrasses,
-limiting it to shallow coastal areas. A number of critical conditions determines grouth of seagrass,
+The adaptation to the salt waters is evidently influenced the global distribution of the seagrasses,
+limiting it to shallow coastal areas. A number of critical conditions determine growth of the seagrass,
 including general climatic characteristics of the area, i.e.temperature, day length, geological and
@@
 of geomorphological and ecological processes. Namely, it is a source of food for herbivorous fauna as
-well as helter zones for fish and other marine organisms; it contributes to the nutrient recycling; it
+well as shelter zones for fish and other marine organisms; it contributes to the nutrient recycling; it
 provides sediments stability by reducing the degree of water movements, etc \cite{Francour99}\label{Francour99}.\\
@@
 
-The distribution of the seagrasses is strongly influensed by several environmental factors, which
+The distribution of the seagrasses is strongly influenced by several environmental factors, which
 include climate (mostly, tropical and temperate are--as), bathymetry (shallow shelf zones), hydrological
-particularities (chemical content of water, nutritient availability and turbidity of waves), and geological 
+particularities (chemical content of water, nutrient availability and turbidity of waves), and geological 
 characteristics - sedimentation and cover types of the seafloor \cite{McKenzie06}\label{McKenzie06}.\\
 There are four European seagrass species in Mediterranean area \cite{Borum04}\label{Borum04}: Zostera marina,
-Zostera noltii, Cymodocea nodosa and \textit{P.oceanica}. In Greece the common species are\textit{P.oceanica}
+Zostera noltii, Cymodocea nodosa and \textit{P.oceanica}. In Greece the common species are \textit{P.oceanica}
 (L.) Delile, Cymodocea nodosa (Ucria) Ascherson, Zostera noltii Hornemann and Halophila
@@
 fundamental differences between both marine organisms, the major of them should be briefly
-mentioned: first, seagrasses are true plants with root system and leaves which photosynthesise, have
-complex structure and create landscape-similar vast formations on the colonised seafloor areas with
+mentioned: first, seagrasses are true plants with root system and leaves which photosynthesis, have
+complex structure and create landscape-similar vast formations on the colonized seafloor areas with
 soft sediment, whereas algae are simple organisms that can only holdfast; secondly, seagrasses are
@@
 e.g. trawling and anchoring of boats \cite{Ribed02}\label{Ribed02}. Other human activities that cause degrading of the
-seagrass are recreational boating, commertial overexploitation of coastal resources, eutrophication
+seagrass are recreational boating, commercial overexploitation of coastal resources, eutrophication
 \cite{McKenzie06}\label{McKenzie06}.\\
@@
 and intensity of storms and waves; local decrease of water quality, increased sedimentation,
-contaminantion and nutrification; desiccation; salinity fluctuations; nutrient changes; suspended
+contamination and nitrification; desiccation; salinity fluctuations; nutrient changes; suspended
 sediments \cite{Blake00}\label{Blake00}. These stress-drivers can alone result in large-scale seagrass
 degradation, but often seagrass undergo simultaneous affects from several of these factors together.
-This naturally increase environmental pressing and leads to drastical loss of very large areas of
+This naturally increase environmental pressing and leads to drastic loss of very large areas of
 seagrass globally \cite{Orth06}\label{Orth06}.
@@
 seagrass meadows, and often lead to their complete disappearance \cite{Holmer09}\label{Holmer09}.
-Other environmental threat for the seagrasses arise as a result of the environmental struggle and
+Other environmental threat for the seagrasses arises as a result of the environmental struggle and
 competition for existence among species. Thus, meadows of P.oceanica in the Mediterranean Sea are
-presently facing invasion by alien algal species, partuicularly in areas where P.oceanica is already
+presently facing invasion by alien algal species, particularly in areas where P.oceanica is already
 degrading, stressed, have gaps and patchy structure in meadows and show other signs of regression
@@
 the seagrass ecosystems are degrading or lost, are not located in a specific area or continent, but
-registered globally; since 1980s global losses of the seagrases on our planet is equal to two footbal
+registered globally; since 1980s global losses of the seagrasses on our planet is equal to two football
 fields per hour \cite{Mellors09b}\label{Mellors09b}.
@@
 The dominant and most productive coastal ecosystem of the Mediterranean, P.oceanica is
-spatially restricted to the Mediterranean area (Fig.\ref{fig:5}), with its extention limited by the western part
+spatially restricted to the Mediterranean area (Fig.\ref{fig:5}), with its extension limited by the western part
 of the Mediterranean Sea where cold Atlantic waters enter Gibraltar and mix with warm
@@
 usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}).
-Growing P.oceanica make a meadows which, in turn, consisnt of smaller patches, called “matte”
+Growing P.oceanica make a meadows which, in turn, consist of smaller patches, called “matte”
 (Fig.\ref{fig:7}), a monumental construction made by the growth of rhizomes and leaves with entangled roots
@@
 of the meadow health status. \\
-There are several environmental factors, determinating the growth of the \textit{P.oceanica}. The adaptation
-to dry-summer subtropical climate reduces its extention to Mediterranean area only (Fig.\ref{fig:5}).
+There are several environmental factors, determining the growth of the \textit{P.oceanica}. The adaptation
+to dry-summer subtropical climate reduces its extension to Mediterranean area only (Fig.\ref{fig:5}).
 Besides, the distribution of the seagrasses changes with water depth: it is noticed \cite{Dural10}\label{Dural10} that
@@
 the maximal flower density in the densest stands, while the occurrence of flowering is regulated by
-environmental factors \cite{Buia91}\label{Buia91}. Phenology of the \textit{P.oceanica} undergo
+environmental factors \cite{Buia91}\label{Buia91}. Phenology of the \textit{P.oceanica} undergoes 
 modifications with varying seasons during the year: during the flowering period (ca 3 months long)
@@
 Monitoring of the marine benthic ecosystems of seagrasses is essential for the environmental
-assessment of the coastal zones. It increases our knowledge of the seagrass ecology, highlightes
+assessment of the coastal zones. It increases our knowledge of the seagrass ecology, highlights 
 threats to the seagrass and preventing them from possible losses and degrading and improves
@@
 important contribution to the environmental coastal zone management \cite{Pergent-Martini06}\label{Pergent-Martini06}.
-However, mapping the sesgrass has limitations due to the specific location and characteristics of the
+However, mapping the seagrass has limitations due to the specific location and characteristics of the
 research object. The remote sensing techniques have traditionally been widely used for the seagrass
@@
 Satellite imagery processing has also being used for the seagrass monitoring, due to their accuracy,
-repeatability and informativeness as a source of data \cite{Dekker05b}\label{Dekker05b}, enabling regular temporal
+repeatability and information value as a source of data \cite{Dekker05b}\label{Dekker05b}, enabling regular temporal
 coverage over the large remote areas and providing a cost-effective approach for the mapping of the
@@
 the selected locations it becomes a useful tool for the assessment of the distribution, density and
-coverage of the seagrass along the tracklog. Besides, the underwater observations using scuba diving
+coverage of the seagrass along the track log. Besides, the underwater observations using scuba diving
 equipment have been conducted for the measurements of depths.
@@
 from the following sources:
-i) spectra of P.oceanica, carbonatic sand, silt and other seafloor types
+i) spectra of P.oceanica, carbonate sand, silt and other seafloor types
 ii) satellite imagery: Landsat, ASTER, MODIS, MERIS
@@
 \subsubsection{Specific objectives}
-1. To apply remote sensing data (Google Earth aerial images, Landsat, SPOT satellie images)
+1. To apply remote sensing data (Google Earth aerial images, Landsat, SPOT satellite images)
 for the monitoring of the seagrass meadows distribution
-2. To study narrow-band spectral reflectance propoerties of \textit{P.oceanica} and other seafloor
-cover types (sand and silt) using WASI swater colour simulation software
+2. To study narrow-band spectral reflectance properties of \textit{P.oceanica} and other seafloor
+cover types (sand and silt) using WASI water colour simulation software
 3. To use methods of the in situ diving observations and underwater videometric measurements
@@
 the equality of the statistical variances.
-The hypothesis testing is suggested to be carried out using the anova statistical test. The purpose of
-anova test is to visualize in an effective and quick way the spectral differences between seagrass
+The hypothesis testing is suggested to be carried out using the ANOVA statistical test. The purpose of
+ANOVA test is to visualize in an effective and quick way the spectral differences between seagrass
 species and their spatial distribution. Thus, the key hypotheses of the research will be tested to prove
@@
 \subsection{Research approach}
-Seagrass consistent monitoring and mappint is necessary and important for the sustainable coastal
+Seagrass consistent monitoring and mapping is necessary and important for the sustainable coastal
 development and conservation measures. Earlier, many seagrass meadows have been destroyed by
@@
 observations and mapping enables precise control of its spatial distribution, detection of any changes
-in the seagrass landscapes, highlightes potential environmental threats in the coastal zone (e.g.
+in the seagrass landscapes, highlights potential environmental threats in the coastal zone (e.g.
 declining of meadows) before they become unmanageable for the coastal management services.
@@
 
-Spectral reflectances, distinct for each seafloor cover type, enable their discrimination on the image. The preprocessing
+Spectral reflectance, distinct for each seafloor cover type, enables their discrimination on the image. The pre-processing
 of the images includes imagery corrections for atmospheric noises and effects of the water
@@
 	\centering
-	\includegraphics[scale=0.6]{UML_GIS.png}
+	\includegraphics[scale=0.57]{UML_GIS.png}
 	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
@@
 Mapping and monitoring the seagrass is important for the environmental assessment of the marine
-ecosystemsn in coastal areas. Regular tracking of current distribution of seagrass meadows, based on
+ecosystems in coastal areas. Regular tracking of current distribution of seagrass meadows, based on
 correct information and cartographic visualization of seagrasses, is a preventive environmental
@@
 terrestrial cartography, due to the technical difficulties of underwater observations, yet nowadays is
-has become a successful, rapidly developping, increasingly popular and challenging research branch.
+has become a successful, rapidly developing and increasingly popular and challenging research branch.
 Seagrasses can be mapped using various approaches, with the most popular method of the remote
@@
 growing on the seafloor.
-Several global seagrass surveys organise and provide regular monitoring of the segrass distribution,
+Several global seagrass surveys organise and provide regular monitoring of the seagrass distribution,
 health and species sustainability, amongst which the most world-known are, perhaps, the following
@@
  These organizations aim at the global seagrass monitoring, providing with
-research results and reporting guidelines and manuals with standartized methods and
+research results and reporting guidelines and manuals with standardized methods and
 recommendations, specific for the seagrass research and monitoring. There are also university marine
@@
 research. Their reports and guidelines were used for references in the current research.
-Regular obserbvations and monitoring of the seagrasses are known since 1960s, mainly in tropical
+Regular observations and monitoring of the seagrasses are known since 1960s, mainly in tropical
 regions (Australia). Since that time traditional methods of the seagrass monitoring and common
@@
 quadrant frame, single point markers, markers, GPS and other equipment. The seagrass sampling is
-taken on the regular way with observation polints covering the study area with normal distribution.
+taken on the regular way with observation points covering the study area with normal distribution.
 During the measurement process, the vertical photograph of the measurements frame is taken, and the
-following points are traditionally estimated: percentage of the seagrass cover withing the quadrate,
+following points are traditionally estimated: percentage of the seagrass cover within the quadrate,
 species composition, sediment composition, canopy height, epyphyte abundance, algae percent cover,
@@
 reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. Other methods of seagrass in-situ monitoring are based on
-using the active hydroacoustic sonar sensors that send towards a sea floor a signal of energy and then
-collect the return echoes for the analysis, yet this requires specialised equipment and is mostly used in
+using the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then
+collect the return echoes for the analysis, yet this requires specialized equipment and is mostly used in
 the depth waters, combined with bathymetric measurements. Another limitation of the acoustic
@@
 on the image. In case of seagrasses it is spectral reflectance of the seafloor cover types, which can be
-analysed using measurements of optical properties of sea water: radiance and irradiance.
+analyzed using measurements of optical properties of sea water: radiance and irradiance.
 Optical remote sensing methods can get through the clear waters to approximately 15–30 m \cite{Mumby04}\label{Mumby04}.
@@
 electromagnetic energy are absorbed and scattered, which is determined by the optical and physical
-properties of the water, e.g. concentration of suspended particles, chlorofill, coloured dissolved
+properties of the water, e.g. concentration of suspended particles, chlorofill, colored dissolved
 organic matter (gelbstoff) that make up the water content \cite{Ackleson2003}\label{Ackleson, 2003}. Besides, light is strongly
@@
 Shallow waters generally contain more dissolved substances and suspended particles, which directly
-influnces the transparency and colour of the waters of shelf zones \cite{Jerlov52}\label{Jerlov52}. Being highly
+influences the transparency and color of the waters of shelf zones \cite{Jerlov52}\label{Jerlov52}. Being highly
 dynamic environments, coastal waters experience a variety of processes which alter their optical
 properties incessantly. The effects of these processes influence application of the hyperspectral
-remote sensing and reinforce other proceses \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides, increase
-sedimentation processes which in turn may change microrelief properties and topology. 
+remote sensing and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
+sedimentation processes, which in turn, may change micro relief properties and topology. 
 The optical properties of the water are best reflected in the values of its radiance and irradiance,
@@
 reflectivity of the object, can be estimated by the direct mathematical division of these first two
-characteristics, and is expressed in persentage.
+characteristics, and is expressed in percentage.
 Therefore, the irradiance and radiance of the water should be measured in order to estimate spectral
@@
 optical sensors spectroradiometers. There are several companies producing radiometers with various
-characteristics and adjusted for defferent purposes, e.g. portable and miniature spectrometers from
+characteristics and adjusted for different purposes, e.g. portable and miniature spectrometers from
 \href{http://www.stellarnet.us}{StellarNet}, Hyperspectral Ocean Colour Radiometer (HyperOCR) sensor by
@@
 (Trios-RAMSES Hyperspectral UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-
-VIS Radiance Sensor) are small-sized, low power-consumpting, flexible for fieldwork yet with high
-level of precision, specially calibrated for air and for water application as well as colour
+VIS Radiance Sensor) are small-sized, low power-consuming, flexible for fieldwork yet with high
+level of precision, specially calibrated for air and for water application as well as color
 measurements  (Fig.\ref{fig:2.1}). These products have been used for the radiance and irradiance
@@
 SCUBA-based in-situ observations, though providing high resolution and accuracy results in seagrass
-mapping, is limited in application, because of their time consumption, weather-dependancy and
+mapping, is limited in application, because of their time consumption, weather-dependency and
 unsuitability for the case of monitoring large areas of water for small-scale mapping. Methods of
@@
 and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
-their weather-independency, cost-effectiveness, accuracy and spatial coverage, which enables
-periodic monitoring of the seagrass mewdoes and gives access to the distant and unapproachable
+their weather-independence, cost-effectiveness, accuracy and spatial coverage, which enables
+periodic monitoring of the seagrass meadows and gives access to the distant and unapproachable
 areas. Integrated together with GIS vector layers and maps, remote sensing data enable historical
@@
 least, nation specific, depending to large extent on the tools available for researchers \cite{Mellors09a} \label{Mellors09a}.
-Universal, international, standardised methods for seagrass directly for seagrasses
+Universal, international, standardized methods for seagrass directly for seagrasses
 as such still should be developed. Various case studies have been performed, yet their mostly report
-methods adjusted for particular areas, without evaluating starndard general algorithms that could be
+methods adjusted for particular areas, without evaluating standard general algorithms that could be
 extrapolated towards other regions. \\ Application of the remote sensing towards seagrass mapping is
@@
 turn, reflectivity of the sediments is affected by the water optical properties and content. For
-example, Stephens \cite{Stephens03}\label{Stephens03} prove that microalgal biomass and community structure affect
-hyperspectral reflectance of sediments, which enable to estimate total microalgal biomass from
+example, Stephens \cite{Stephens03}\label{Stephens03} prove that microalgae biomass and community structure affect
+hyperspectral reflectance of sediments, which enable to estimate total microalgae biomass from
 measurements of hyperspectral reflectance. Spectral measurements of the target objects are made
-my the means of the radiometers, which receive and register the amounts of enegry (radiance and
+by means of the radiometers, which receive and register the amounts of energy (radiance and
 irradiance) from the objects. \\ Measuring optical properties of water allow to calculate spectra of the
@@
 using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01} \label{Louchard03} \label{Dehouck08} \label{Werdell03}.
-Studies of spectral reflectances of the different seagrass species comparing to the spectra of sand and
+Studies of spectral reflectance of the different seagrass species comparing to the spectra of sand and
 other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
@@
 differences are well detectable by the means of the remote sensing research methods. Comparing to
-the terrestrial plants, aquatic vegtation inter alia seagrass cannot be detected using red edge of the
+the terrestrial plants, aquatic vegetation, \textit{inter alia} seagrass, cannot be detected using red edge of the
 spectrum, as these wavelengths are significantly absorbed by water \cite{Kirk94}\label{Kirk94}, as well as by
@@
 spectrally based radiative transfer approach to quantitatively estimate shallow-water bathymetry and
-leaf area index (LAI) of the seagrass. The spectral reflectances in general are the
+leaf area index (LAI) of the seagrass. The spectral reflectance in general is the
 result of the spectral absorption in different bands, typical for each target object.
-Spectral reflectance of the seagrass  (Fig.\ref{fig:11}) is largely influenced by the water
+Spectral reflectance of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
 depth where it is located, and is generally decreasing in values by increasing depths.
@@
 recent imagery allows to detect changes in seagrass distribution on over different years
-and to analyse dynamics of the seagrass distribution\cite{Ball09}\label{Ball09}. Another important advantage
+and to analyse dynamics of the seagrass distribution \cite{Ball09}\label{Ball09}. Another important advantage
 of the application of the remote sensing data for mapping of submerged aquatic vegetation has
-commertial nature: using remote sensing data and methods enables more low-cost and up-to-date
+commercial nature: using remote sensing data and methods enables more low-cost and up-to-date
 seagrass mapping \cite{Mumby99}\label{Mumby99}, and is especially useful for the areas where the fieldwork data
@@
 	\includegraphics[scale=0.25]{Fig-12.jpg}
-	\caption{Difference between broaddand multispectral and hyperspectral resolution of
+	\caption{Difference between broadband multispectral and hyperspectral resolution of
 		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
@@
 well (e.g., \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird}, \href{http://www.spotimage.fr/}{SPOT}, \href{http://www.itres.com/products/imagers/casi550}{CASI}).
-The possibility of their application towards seagrass mapping
-vary and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:2.3}). In the next subparagraphs
+The possibility of their application towards seagrass mapping varies and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:2.3}).
+In the next subparagraphs
 we briefly discuss the limitations and research experience of the using of various imagery
@@
 Perhaps, the most known imagery, widely used in the remote sensing mapping, is received from the
-group of Landsat sensors, known for its historical, pioneeer role in the satellite industry.
+group of Landsat sensors, known for its historical, pioneer role in the satellite industry.
 Being the longest running satellite system, lunched in 1972, Landsat is the only source of archival
 data going back to 1984 at a sufficient spatial resolution \cite{Dekker05b}\label{Dekker05b}, which makes its data
-desirable for historial mapping or environmental analysis of change detection of the seagrass
+desirable for historical mapping or environmental analysis of change detection of the seagrass
 landscapes.
-The Landsat TM and Landsat ETM+ data, with the most recent from sensor Landsat 7, an advanced
-and multispectral scanning, launched in 1999, proved to be feasibile and useful for the mapping of
+The Landsat TM and Landsat ETM+ data, with the most recent from sensor Landsat 7 - an advanced
+and multispectral scanning, launched in 1999, prove to be feasible and useful for the mapping of
 submerged vegetation, such as seagrasses or coral reefs. The successful applications of the imagery
 Landsat TM towards the seagrass mapping, were reported in numerous research works \cite{Palandro03,Gullstroom06,Wabnitz08,Bierwirth93,Ferguson97,Rasib97}\label{Palandro03}\label{Gullstroom06}\label{Wabnitz08}\label{Bierwirth93}\label{Ferguson97}\label{Rasib97}.\\ The Landsat data are particularly suitable for the
-case of change detection of seagrass landscapes at a decadal scale, because being the main sensor
+case of change detection of seagrass landscapes at a decade scale, because being the main sensor
 onboard the Landsat satellites, the Thematic Mapper (TM) provides the longest time series
@@
 made using IKONOS, Landsat TM, and CASI, show that in the blue part of the
-spectrum, the best results are achieved by IKONOS and CASI, while Landsat TM has not high
+spectrum, the best results are achieved by the IKONOS and CASI, while Landsat TM has not high
 enough resolution. It may be caused to some extent by the loss of the radiance contrast, atmospheric
 Rayleigh scattering and defects of scattering. However, comparing between CASI and IKONOS, the
-same authors prove that CASI enable to receive still more accurate results of the the classification
+same authors prove that CASI enable to receive still more accurate results of the classification
 than IKONOS \cite{Mumby02}\label{Mumby02}.
@@
 Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}), launched in
-1999 onboard Terra sensor, provides high-resolution images of the Earth in 15 different bands of the
+1999 on board Terra sensor, provides high-resolution images of the Earth in 15 different bands of the
 electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:2.4}). It has diverse
@@
 thermal infrared (TIR) wavelength regions \cite{Kalinowski04}\label{Kalinowski04}. For each channel there is separate
-onboard calibration (OBC) system, telescope with independent pointing and different detector technsology \cite{Arai05}\label{Arai05}. 
-This makes ASTER imagery expecially suitable for the application towards detailed
+on board calibration (OBC) system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}. 
+This makes ASTER imagery especially suitable for the application towards detailed
 mapping of surface temperature, emissivity and reflectance of objects and bathymetric elevations as
 well. Successful application of the ASTER imagery towards mapping of submerged vegetation
-reported, for example, by Hirose \cite{Hirose04} \label{Hirose04}. Other multispectra images have also been used for the
+reported, for example, by Hirose \cite{Hirose04} \label{Hirose04}. Other multispectral images have also been used for the
 seagrass interpretation. For example, the success of using imagery from the multispectral airborne
@@
 used as being more suitable in classification and interpretations of imagery.
-While this and other studies \cite{Phinn08}\label{Phinn08}demonstrated the advantages and success of the
+While this and other studies \cite{Phinn08}\label{Phinn08} demonstrated the advantages and success of the
 application of multispectral imagery for the spectral discrimination of the seafloor cover types and
@@
 hyperspectral imagery, are strong tool enabling to detect even slight and subtle differences in the
-spectral reflectances between various seafloor types, e.g. different seagrass species at diverse depths,
+spectral reflectance between various seafloor types, e.g. different seagrass species at diverse depths,
 algae, corals, dark-coloured sands or other types of sediments \cite{Hochberg03a}\label{Hochberg03a}. 
@@
 hyperspectral imagery \cite{Peneva08}\label{Peneva08}, combined with fieldwork measurements ensures accurate
-mapping results with the maximum likelihood methos producing the best results.
+mapping results with the maximum likelihood methods producing the best results.
 Therefore, accurate mapping of the seagrass landscapes and other seafloor types using remote sensing
@@
 seafloor types classification \cite{Hochberg03a}\label{Hochberg03a} demonstrates that coral, seagrasses and
-sand wery vell distinguishable in their spectra with an overall classification accuracy of 98 percent.
-However, the use of data from various sensors, both hyperspectral and multispectral, are possible and
+sand very well distinguishable in their spectra with an overall classification accuracy of 98 percent.
+However, the use of data from various sensors, both hyperspectral and multispectral, is possible and
 reasonable, as soon as it meets the research specific objective. 
@@
 
-In the section \nameref{sec:2} we briefly discussed main existing cartographic methods and tools for seagrss mapping.
+In the section \nameref{sec:2} we briefly discussed main existing cartographic methods and tools for seagrass mapping.
 In the current  research we try to combine various techniques from mentioned above for the assessment of the seagrass distribution in the Mediterranean environment, for the case study of Crete.
 We mentioned in \nameref{sec:0} the software that we used for logically separated research sections, which required various approach.
-Thus, for example,  spectral analysis and assessment of spectral signaturesis technically based on WASI, sttistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://gretl.sourceforge.net/}{Gretl} and \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using\href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
+Thus, for example,  spectral analysis and assessment of spectral signatures is technically based on WASI, statistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://gretl.sourceforge.net/}{Gretl} and \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
 \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0} software is used for the spatial analysis, general mapping and cartographic layout
@@
 The raster processing includes making mosaic-like covering for the whole research area. The Google
-Earth images are most appropriate for the detailed mapping than stellite than \href{http://landsat.gsfc.nasa.gov/}{Landsat} or \href{http://asterweb.jpl.nasa.gov/}{ASTER} enabling to produce accurate maps with correct results.
+Earth images are most appropriate for the detailed mapping than satellite than \href{http://landsat.gsfc.nasa.gov/}{Landsat} or \href{http://asterweb.jpl.nasa.gov/}{ASTER} enabling to produce accurate maps with correct results.
 Therefore, the main set of images for the
@@
 The sampling stations were located in two candidate places on the northern (Ligaria) and southern
-(Xerocampus) parts of Crete island, as these regions are well suitable for the seagrass, due to the
+(Xerocampos) parts of Crete island, as these regions are well suitable for the seagrass, due to the
 annual mean water temperatures and geological factors, i.e. seafloor conditions and sediments. \\ The
@@
 \subsection{Fieldwork data collection}
-Seagrass sampling has been performed at two research stations at Crete island - Ligaria beach
-(36°20′N 22°59′E) and Xerokampos (35°12′N 26°18′E), at depths lesser than 4 meters.
+Seagrass sampling has been performed at two research stations at Crete Island - Ligaria beach
+(36°20′N 22°59′E) and Xerocampos (35°12′N 26°18′E), at depths lesser than 4 meters.
 
@@
 above-ground seagrass patches, sediment seafloor cover types, species compositions, water depth and
-geograpic locations recorded using GPS.
+geographic locations recorded using GPS.
 \pagebreak
@@
 photographing up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
-Markers and cords for the depths measurements; 5) Waterproof plasctic Otterbox  (Fig.\ref{fig:3.3b})for keeping the
-iPAQ dry; 6) \href{http://www.padi.com/scuba/}{SCUBA Diving equipment}, taken from the Ligaria Diving Centre; 7) Boat (Fig.\ref{fig:3.4});
+Markers and cords for the depths measurements; 5) Waterproof plastic Otterbox  (Fig.\ref{fig:3.3b})for keeping the
+iPAQ dry; 6) \href{http://www.padi.com/scuba/}{SCUBA Diving equipment}, taken from the Ligaria Diving Center; 7) Boat (Fig.\ref{fig:3.4});
 
@@
 	\includegraphics[scale=0.70]{Fig-18.jpg}
-	\caption{Boat used for the fieldwork measurements on Crete island. 2010}
+	\caption{Boat used for the fieldwork measurements on Crete Island. 2010}
 	\label{fig:3.4}
@@
 patches. The fieldwork included several routes of the boat in the Ligaria beach sampling site, nine
-routes in total, in the directions parallel to the coastline, ca 180-200 m long each, thus enabling the
+routes in total, in the directions parallel to the coastline, ca 180-200 m long each one, thus enabling the
 course plot to cover the area of growing seagrass: shelf areas not deeper than four meters.
@@
 boat path. The general locations of the sampling sites and routes were selected on the basis of the visual examination
-of the seagrass beds during snorkelling and SCUBA diving, recommendations from the Greek collaborators of the Natural Nistory Museum of Crete, and available maps covering the research area.
+of the seagrass beds during snorkeling and SCUBA diving, recommendations from the Greek collaborators of the Natural Nistory Museum of Crete, and available maps covering the research area.
 To ensure even and objective selection of sampling sites
@@
 research sampling design also included measurements of depths (Fig. \ref{fig:A.4}), because bathymetry is one of the
-most determinating factors for the seagrass locations. Measurements of depths were performed during
+most determining factors for the seagrass locations. Measurements of depths were performed during
 the fieldwork using cord, iPAQs, and markers in order to assure that the videometric measurements
@@
 with videographic measurements. The GPS allowed to capture measurement locations on the iPAQ,
-encapsulated in a plastic waterproof Otterbox (Fig \ref{fig:3.3b}). Camera were adjusted horizontally by a leveller and
-were mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:3.5}). The data
+encapsulated in a plastic waterproof Otterbox (Fig \ref{fig:3.3b}). The camera were adjusted horizontally by a leveler and
+ mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:3.5}). The data
 were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state. The
@@
 
-Along the Ligaria beach there are several types of the seagrass landscapes, namely, segrass meadows
+Along the Ligaria beach there are several types of the seagrass landscapes, namely, seagrass meadows
 continuously covering vast areas (Fig.\ref{fig:A.3} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:A.3} (b)) and isolated seagrass patches, or mattes,
@@
 and up-to-date source of imagery.
-Perhaps, the most clear advantage of the Google
-Earth imagery is high resolution (15 m in land areas and lower in the oceans).
+Perhaps, the clearest advantage of the Google
+Earth imagery is its high resolution (15 m in land areas and lower in the oceans).
 Obtained from the airborne platforms, Google Earth images have general spatial
-resolution of several meters (though vatying in different areas), which allows very
+resolution of several meters (though varying in different areas), which allows very
 detailed habitat and seafloor types discrimination, comparing with images
-acquired from the spaceborne satellite platforms.
+acquired from the space-borne satellite platforms.
 The spatial coverage of the Google imagery is lesser comparing to spaceborne images, but this can be solved as
@@
 \subsection{Review of the collected data}
-The remote sensing imagery were collected for the Crete island area covering the research period and
+The remote sensing imagery was collected for the Crete island area covering the research period and
 area, enabling the field data to be used for the calibration and validation. The available aerial and
 satellite imagery are commonly used for the mapping of seagrass landscapes and their application is
-proven by various research papers. The imagery include satellite multi-spectral imagery (Landsat-TM,
+proven by various research papers. The imagery includes satellite multi-spectral imagery (Landsat-TM,
 ETM+) and aerial imagery from the Google Earth.
@@
 	\includegraphics[scale=0.40]{UML_Dia.jpg}
-	\caption{Workflow for the data aquisition, \href{http://live.gnome.org/Dia}{Dia} display}
+	\caption{Work flow for the data acquisition, \href{http://live.gnome.org/Dia}{Dia} display}
 	\label{fig:24}
@@
 UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.
-\ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see Appendices: Tab.
+\ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see \nameref{app}: Tab.
 \ref{tab:A.1}, Tab.
@@
 properties of the sea water and bottom cover types (Fig.\ref{fig:A.15}). The spectral range of radiance
-cover diapason of 320-950 nm, and ittadiance measurements are covered in the interval of 280-500
-nm, which is suitable for a characteristics of seagrass reflectance.
+cover diapason of 320-950 nm, and irradiance measurements are covered in the interval of 280-500
+nm, which is suitable for characteristics of seagrass reflectance.
 Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:A.12}, or \ref{fig:A.14}, enlarged) represent several series of
@@
 \ref{fig:A.17} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively.
-Graph on Fig \ref{fig:A.13} displays statistuical analysis of the measured sets of observations, i.e.
-shows the midspread of the statistical quartiles (Q1-Q3), mean and extreme values.
+Graph on Fig \ref{fig:A.13} displays statistical analysis of the measured sets of observations, i.e.
+shows the mid-spread of the statistical quartiles (Q1-Q3), mean and extreme values.
 
@@
 \subsection{Data pre-processing}
-A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonatic sand - consist of multiple measurement sets of P.oceanica made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
-\ref{tab:A.8}), 1-445 for 14th October for P.oceanica; 1-106 for water without sedimnts, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
+A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of P.oceanica made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
+\ref{tab:A.8}), 1-445 for 14th October for P.oceanica; 1-106 for water without sediments, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
 \ref{tab:17}).
-All these data have been analysed and statistically tested.
-We measured mean, extreme values (min-max), median and statistical quartiles, in order to visualise distribution of the values at varuious spectral wavelengths.
+All these data have been analyzed and statistically tested.
+We measured mean, extreme values (min-max), median and statistical quartiles, in order to visualize distribution of the values at various spectral wavelengths.
 Carbonate sand was measured at wavelengths 402-750 nm, while P.oceanica - at the interval 318-951 nm.
-We visualised the behavour of different spectra on relevant graphs (Fig.
+We visualized the behavior of different spectra on relevant graphs (Fig.
 \ref{fig:4.4}, \ref{fig:A.35}, \ref{fig:A.28}).
-The available collected data were tested for the spectral variability and separability under varying conditions of dfifferent environmental
+The available collected data were tested for the spectral variability and separability under varying conditions of different environmental
 constituents (e.g. depth, water content, sun angle), in order to determine the
 potential that it may have on the approaches for further images processing and classification.
-Behavour of the spectra of P.oceanica have been tested using datasets for various depths: 0.5, 1.5 and 2.5 meters (Tab.
+Behavior of the spectra of P.oceanica have been tested using datasets for various depths: 0.5, 1.5 and 2.5 meters (Tab.
 \ref{tab:8}, Tab.
@@
 The raw
-initial data of measurements of the spectral reflectance have been preprocessed and statistically analysed thereafter with graphs visualising mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:4.10}, \ref{fig:28}).
+initial data of measurements of the spectral reflectance have been preprocessed and statistically analyzed thereafter with graphs visualizing mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:4.10}, \ref{fig:28}).
 These data
@@
 Therefore, these data had to be interpolated (Fig. \ref{fig:A.10}) and standardized to one
-format, which is values of spectral reflectancese with one nm step. For the
-interpolation we used script written on Python programming language that allowed to receive more 
+format, which is values of spectral reflectance with one nm step. For the
+interpolation we used script written on Python programming language that allowed receiving more 
 detailed data by interpolating them from 3 and 4 nm step up to 2 nm (Fig.\ref{fig:A.8}).
@@
 We included measurements of the Thalassia seagrass
-(measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectances of
-different seagrass speciea under various environmental conditions.
-The most suitable wavebands for the seagrass monitroring usually lay between 400 and 700 nm,
+(measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectance of
+different seagrass species under various environmental conditions.
+The most suitable wavebands for the seagrass monitoring usually lay between 400 and 700 nm,
 which can be concluded by the visual examination, comparison and analysis of the different spectra of
 the seagrasses. Therefore, we chosen the spectra 400-750 as the most appropriate range for
-further research experiment.The results of the linear interpolation (Fig.\ref{fig:A.35}) demonstrate values of the
-sand spectral reflectances with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
+further research experiment. The results of the linear interpolation (Fig.\ref{fig:A.35}) demonstrate values of the
+sand spectral reflectance with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
 The initial measured data were stored in raw-oriented format, so that re-formatting them into the
 column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:A.11}). The next step included
-calculation of the median, mean, quartiles and other statistical values at every data set (see Appendix, Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
-After the preliminary analysis, the measured data were visualised using Gnuplot programm, which
+calculation of the median, mean, quartiles and other statistical values at every data set (see \nameref{app}, Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
+After the preliminary analysis, the measured data were visualized using Gnuplot program, which
 enables fine plotting of various datasets together: Fig \ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:28}.
@@
 as it has trend-friendly graph better showing the general behavior of the curves at different wavebands
-comparing to splines (see Appendix for more results). 
+comparing to splines (see \nameref{app} for more results). 
 Therefore, after several experiments with
-various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have choosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contain convex
+various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have chosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contains convex
 hulls made on its control points, and therefore is best suitable for our case: analysis of optical
@@
 	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentray graphs below show the results of the statistical analysis.}
+\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis.}
 	\label{fig:4.4}
@@
 The values of their spectral reflectance are accepted as constant. The results of
-spectral measurements enable to analyse, whether P.Oceanica is spectrally distinct from other sea
+spectral measurements enable to analyze, whether P.Oceanica is spectrally distinct from other sea
 floor types with changing environmental conditions, using the differences in their spectral signatures
 on the graphs in a WASI, the Water Colour Simulator software. \\The Water Colour Simulator WASI, a
-software tool for analysing and simulating the most common types of spectra \cite{Gege05}\label{Gege05}, is highly
+software tool for analyzing and simulating the most common types of spectra \cite{Gege05}\label{Gege05}, is highly
 suitable for the seagrass spectral analysis.
@@
 water, large and small particles. The values of all these parameters can be redacted and changed
-manuall. However, the most important, major factors affecting the in-situ conditions are water depth (Fig.\ref{fig:4.9})
-and cover fraction of the seafloor types: P.oceanica and carbonatic sand .
+manually. However, the most important, major factors affecting the in-situ conditions are water depth (Fig.\ref{fig:4.9})
+and cover fraction of the seafloor types: P.oceanica and carbonate sand .
 
@@
 	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentray graphs below show the results of the statistical analysis}
+\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis}
 	\label{fig:27}
@@
 condition for this research question. The depths values chosen for the current research are lesser than
-3.5 meters, covering shelf zone, and providing the best enironmental conditions for the seagrass
+3.5 meters, covering shelf zone, and providing the best environmental conditions for the seagrass
 \textit{P.oceanica}: 0.5, 2.0 and 3.5 meters with an interval of 1.5 meters.
@@
 	\includegraphics[scale=0.30]{GNU-20.jpg}
-	\caption{Statitical comparison of spectral reflectances of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
+	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
 	\label{fig:28}
@@
 A statistical analysis of WASI-simulated spectral reflectance has been used in order to answer the first
-research question: whether the P.oceanica spectra is spectrally distinct at varying environmental insitu
+research question: whether the P.oceanica spectra is spectrally distinct at varying environmental in-situ
 conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
@@
 
-In order to focus on the factors of primar importance, other and less influencing factors are excluded,
+In order to focus on the factors of primary importance, other and less influencing factors are excluded,
 i.e. sun angle, concentration of the suspected particles in the water column, content of gelbstoff, etc.
@@
 concentration of gelbstoff (coloured dissolved organic matter), concentration of phytoplankton,
-aerosol scattering, exponent of backscattering by small particles, cloudness, viewing geometry and
+aerosol scattering, exponent of backscattering by small particles, cloudiness, viewing geometry and
 wind speed (which is, however, not the major source of uncertainty). All these environmental
-components increase the absorbtion and scattering of light which, in its turn, results in a complex
+components increase the absorption and scattering of light which, in its turn, results in a complex
 relationship between their concentrations and the radiance of water that finally influence its spectral
@@
 The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
-the seagrass \textit{P.oceanica}meadows, silt and carbonate sand differ and can be clearly discriminated while mapping. A study is
+the seagrass \textit{P.oceanica} meadows, silt and carbonate sand differ and can be clearly discriminated while mapping. A study is
 based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
@@
 The data provided with the model was determined at freshwater of the Boddensee \cite{Gege05}\label{Gege05}, 
-yet the model was adjasted for the marine
-environment, so that its parameters (concentration of chlorophyll, concentration of small particles,
+yet the model was adjusted for the marine environment, so that its parameters (concentration of chlorophyll, concentration of small particles,
 yellow substance, etc) now perfectly simulate the Mediterranean water conditions.
@@
 a diapason of 350-800 nm with a 1nm spectral resolution. For the spectral analysis we applied
-forward calculations, which is a computing and plotting of series of spectra according to specified
+forward calculations, i.e. a computing and plotting of series of spectra according to specified
 parameter settings, with exactly defined depths and cover fraction.
-The specific parameters chosen for the simulation of the environmengal conditions where seagrass
-grow. The adaptations to life in salt sea water requires various phisical and chemical parameters
+The specific parameters have been chosen for the simulation of the environmental conditions where seagrass
+grow. The adaptations to life in salt sea water requires various physical and chemical parameters
 which include salinity, temperature of 17-25°C, light requirements with 10-20 percent on average, ranging
 from 4.4percent  minimal up to 29 percent \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as 35-
-45° and reflection factor 0.0201. P.oceanica. These values, simyulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) on the WASI user-defined parameters.
+45° and reflection factor 0.0201. P.oceanica. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
 The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:4.8}).
@@
 	\subfloat [Bottom reflectance at depth 0.5-4.0 meters]{\label{fig:31b}\includegraphics[width=0.4\textwidth]{Fig-31b.jpg}}
-	\caption{Sea water physical properties, modelled by WASI}
+	\caption{Sea water physical properties, modeled by WASI}
 	\label{fig:4.8}
@@
 Although seagrass \textit{P.oceanica} can be found until depth limits down to 40 m depth (Hartog
-1970), the most preferrable limits of its distribution in the Mediterranean Sea are shallow waters until
+1970), the most preferable limits of its distribution in the Mediterranean Sea are shallow waters until
 4 meters of depth.
@@
 production of chlorophyll, because when light passes through the water and suspended particles, it is
-being largely modifed through the absorption and scattering before it finally reaches plant canopy of
+being largely modified through the absorption and scattering before it finally reaches plant canopy of
 the seagrass. Therefore, the most healthy and suitable areas for the seagrass grow are located at depths
@@
 
-An analysis of the spectral reflectances of \textit{P.oceanica}.is done using the WASI simulations in order to
+An analysis of the spectral reflectance of \textit{P.oceanica} is done using the WASI simulations in order to
 determine, which wavebands can be still used to identify \textit{P.oceanica}. The analysis of spectra shows
@@
 also peaks at around 700 nm, ca between 680 and 710 nm (Fig.\ref{fig:4.10}). The highest values of the bottom
-reflectance are at spectra of 500-600 nm. The most appropriate depths at which the spectral signatues
-could still be discrtiminated for seagrass are lesser than 2.5 meters.
+reflectance are at spectra of 500-600 nm. The most appropriate depths at which the spectral signatures of the seagrass
+could still be discriminated are lesser than 2.5 meters.
 The remote sensing properties of the broad-band satellite sensors are often limited and do not allow
-easy and unique discrimination of seagrass from other seafloor tyeps. Therefore, we focus on narrowband
+easy and unique discrimination of seagrass from other seafloor types. Therefore, we focus on narrow-band
 spectra with the resolution of 1 nm.
@@
 oceanica (Fig.\ref{fig:A.28}).
-The graph is received in excel spreadsheet using mean values of spectral reflectances of
+The graph is received in excel spreadsheet using mean values of spectral reflectance of
 sand (Fig.\ref{fig:A.27}) and seagrass (Fig.\ref{fig:A.35}), respectively, which have been calculated from measurements of radiance and
@@
 The total amount of measured data was large and included following datasets made using
-hyperspectral radiometer Ramses: 350 measuremetn sets of P.oceanica reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
+hyperspectral radiometer Ramses: 350 measurement sets of P.oceanica reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
 sets of P.oceanica reflectance for 15th Oct (Tab.\ref{tab:A.8}), 84 datasets for seawater reflectance with sediments, 105
-datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonat sand.
+datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonate sand.
 A statistical approach is evidently necessary for the proper processing of such amounts of data. 
 The schematic view of the statistical approach used for the data processing  can be seen on Fig. \ref{fig:4.12}.
-Statistical pre-processing of large sets of serial data enables to generalyse data by using the most typical and predicted values of
+Statistical pre-processing of large sets of serial data enables to generalize data by using the most typical and predicted values of
 spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
@@
 at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). 
-The statistical pre-processing and analysis were made to display the mean values of spectral reflectances,
+The statistical pre-processing and analysis were made to display the mean values of spectral reflectance,
 which were used for comparison with reflectance data of various seafloor cover types as well as
 seawater with and without sediment (e.g. Fig..\ref{fig:A.16}).
-In the statistical anlysis of the raw observed data pool of the values of spectral reflectances (about 400 single measurements for P.oceanica, e.g.
+In the statistical analysis of the raw observed data pool of the values of spectral reflectance (about 400 single measurements for P.oceanica, e.g.
 Tab.\ref{tab:A.8}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
@@
 We divided the total data pool into datasets of 25 measurements for more convenience and ease of use, both for P.oceanica and carbonate sand.
-The mean of a probability distribution of spectral reflectance describes the average value of the random variable over all of its possible realizations (Appendices).
+The mean of a probability distribution of spectral reflectance describes the average value of the random variable over all of its possible realizations (\nameref{app}).
 Conceptually, there are an infinite number of such realizations, therefore parameters are not known to us.
@@
 
-Thus, we collect a number of realizations (a sample of 25 measurements for both sand and P.oceanica) and then estimate the statistical parameters (Appendix). \\
-Finally we analysed abount 400 datasets totally (example of carbonate sand on Fig.\ref{fig:4.14}, P.oceanica - on  Fig.\ref{fig:4.13}) made during different days. 
+Thus, we collect a number of realizations (a sample of 25 measurements for both sand and P.oceanica) and then estimate the statistical parameters (\nameref{app}). \\
+Finally we analyzed about 400 datasets totally (example of carbonate sand on Fig.\ref{fig:4.14}, P.oceanica - on  Fig.\ref{fig:4.13}) made during different days. 
 Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random. 
-Thus, we analysed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implmented in Gretl software (Tab.\ref{tab:19}). 
-Their probability expresses uncertainty about the possible values of the spectral reflectances  of P.oceanica and sand, respectively.
+Thus, we analyzed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implemented in Gretl software (Tab.\ref{tab:19}). 
+Their probability expresses uncertainty about the possible values of the spectral reflectance of P.oceanica and sand, respectively.
 There is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
 Each time we observe the outcome of a random variable of spectral reflectance, we obtain an observation, which is hence no longer random. 
-We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the Appendix: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyse the distribution of values of spectral reflectances of \textit{P.oceanica}.
+We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the \nameref{app}: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyze the distribution of values of spectral reflectance of \textit{P.oceanica}.
 A probability distribution, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
-The \textit{least absolute deviations (LAD)}, a popular optimization technique, was used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
-Non-linear logistic anlysis is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
+The \textit{least absolute deviations (LAD)}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
+Non-linear logistic analysis is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
 the value of spectral reflectance for each single observation) is not continuous, but discrete and independent for each case (Tab.\ref{tab:11}).
 \\
-As the observations were made in a repeated way, we also analysed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand.
+As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand.
 The \textit{autocorrelation}, a cross-correlation of an observed value with itself, made on two different days (14th and 15th October),
@@
 We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}.\\
-To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectances at several depths: Tab.\ref{tab:5}, Tab.\ref{tab:6}, Tab.\ref{tab:7}.
+To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectance at several depths: Tab.\ref{tab:5}, Tab.\ref{tab:6}, Tab.\ref{tab:7}.
 The summary of the values of spectral reflectance properties on various depths is presented on the Tab.\ref{tab:A.7}.
 The \textit{k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
-It enables to highlight the main areas of the location of values of spectral reflectances along the spectra.\\
-To estimate spectral density of a spectral reflectancfe measurements we used periodogram function (Tab.\ref{tab:30}), 
+It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
+To estimate spectral density of a spectral reflectance measurements we used periodogram function (Tab.\ref{tab:30}), 
 which corresponds to the general spectrum of the observations with representation of a variable quantity.  
-Thus, the periodogram (Fig. \ref{fig:A.31}) of the series of observation highlights so-called “dead spots” of low power between the frequencies we wishe to
+Thus, the periodogram (Fig. \ref{fig:A.31}) of the series of observation highlights so-called “dead spots” of low power between the frequencies we wish to
 exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:A.12} ) 
 and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:A.14}, enlarged part of 500-660nm).\\
-The visualization of the data plotting was made by means of Gnuplot software, which enabels fine
+The visualization of the data plotting was made by means of Gnuplot software, which enables fine
 drawing and advanced displaying of large amounts of serial data, ultimate control over graph properties, 
-the simplicity of plotting and the ease of scripting. (see Fig.\ref{fig:4.13}, Appendices).
+the simplicity of plotting and the ease of scripting (see Fig.\ref{fig:4.13}, \nameref{app}).
 The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
@@
 (spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
-spectra, where the measurements were done; the y-axis was adjusted for the better visualisation of the
+spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
 P.oceanica spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
 more than 0.20 nm, except for borders), we did not extend the y-scale up to 100 percent, in order to
-focus on the most necessary area of spectral values (Fig.\ref{fig:4.14}, Fig.\ref{fig:4.4} and Appendices). The mean values are
-highligted using bold red points, as we used these values for plotting the final graph of P.oceanica spectra in
+focus on the most necessary area of spectral values (Fig.\ref{fig:4.14}, Fig.\ref{fig:4.4} and \nameref{app}). The mean values are
+highlighted using bold red points, as we used these values for plotting the final graph of P.oceanica spectra in
 Ligaria beach, Agia Pelagia.
@@
 		\includegraphics[scale=0.45]{GNU-13.jpg}
-		\caption{Multiplot of spectral reflectance of P.oceanica. Series 1-100. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
+		\caption{Multiplot of spectral reflectance of P.oceanica. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
@@
 		\includegraphics[scale=0.5]{GNU-16.jpg}
-		\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and
+		\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown mid spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
@@
 A statistical hypothesis test has been applied for the making decision and controlling the wealth of the
-observational data of of the hyperspectral measurements of the water reflectances.
+observational data of the hyperspectral measurements of the water reflectance.
 The received results are statistically significant if they are unlikely to have occurred by chance alone,
 according to a pre-determined threshold probability, the significance level. Therefore, we applied
-critical tests of significance to analyse the measured data according to their significant value.
+critical tests of significance to analyze the measured data according to their significant value.
 Answering the first research question, we suggest the following statement.
@@
 reflectance of P.oceanica at these depths, that was actually observed? The alternative Hypothesis Ha
-Fig.15. Multiplot display of spectral reflectances of P.oceanica. Measurement sets 1-100, 15th
+Fig.15. Multiplot display of spectral reflectance of P.oceanica. Measurement sets 1-100, 15th
 October. Bold red dots show the mean values within each dataset; vertical areas: quartiles Q1-Q2.
@@
 Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.
-To analyse average values of spectral reflectance of P.oceanica by means of the measured sets we calculated 
-mean values for the total set of emasurements 1 – 350 (15th October): mean of min and max values, 
+To analyze average values of spectral reflectance of P.oceanica by means of the measured sets we calculated 
+mean values for the total set of measurements 1 – 350 (15th October): mean of min and max values, 
 mean of average for each waveband, mean for statistical quartiles for each wavelength and mean for median, respectively. 
 The adjusted averages evaluate mean values, needed for statistical analysis of the datasets. 
-The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (Appendices). 
-As a result, we have mean values of spectral reflectances of P.oceanica for extreme, average, median and both quartiles,
+The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (\nameref{app}). 
+As a result, we have mean values of spectral reflectance of P.oceanica for extreme, average, median and both quartiles,
 independent of the individual sets 1-350.
@@
 of a difference between sample means, is used for the data analysis. In our case the Student t-test demonstrates, if the variation
-between two analysed groups – values of spectral reflectances of seagrass P.oceanica and sand - is
+between two analyzed groups – values of spectral reflectance of seagrass P.oceanica and sand - is
 significant. Therefore, we use Student t-test to compare two sets of quantitative data of spectral
-reflectances of P.oceanica and sand, respectively, with samples collected independently of one
+reflectance of P.oceanica and sand, respectively, with samples collected independently of one
 another. The Student-t test can be performed knowing just the means, standard deviation, and number
-of data points. Therefore, we used the data (Appendix 11) of means of spectral reflectances of both
+of data points. Therefore, we used the data (\nameref{app}) of means of spectral reflectance of both
 bottom cover types within data sets, their standard deviation, and the number of data points.
@@
 (Box Plots, Fig..).
-The graphs werer plotted on the basis of the following statistical data calculated from the sampling
+The graphs were plotted on the basis of the following statistical data calculated from the sampling
 measurements data: Table 2.
@@
 75th Pct: Plotting value of third quartile = Median-Q3
-Min: Lower error bar lenth=Q1 - Minimum
-Max: Upper error bar lenth=Maximum - Q3
+Min: Lower error bar length=Q1 - Minimum
+Max: Upper error bar length=Maximum - Q3
 The statistical analysis is displayed on the Graph 1.1 which compares the spread in sets of
-measurement data of spectral reflectances of under various environmental conditions (depth).
+measurement data of spectral reflectance of under various environmental conditions (depth).
 
@@
 a way we defined an empirical upper limit to the discriminative potential of these sensors. To create
-these sensor-specific reflectance spectra, spectral responses of each of these sensor were applied for
+these sensor-specific reflectance spectra, spectral responses of each of these sensors were applied for
 the spectral signature of seafloor cover types in full-resolution spectra.
@@
 Seagrass beds are clearly visible in color aerial Google Earth
-photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and browny sand bottom. The seagrass areas are
+photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
 detected using different bands combinations, masked and studied for the estimation of the changes in
@@
 The integrated approach used in this research work has high potential as a means to monitor changes
-in seagrass landscape occurring in a shallow waters over Crete area.
+in seagrass landscape occurring in shallow waters over Crete area.
 
@@
 It encompasses the integration of
-high resolution aerial color Google Earth photography, spaceborne satellite imgery, assessment of
+high resolution aerial color Google Earth photography, spaceborne satellite imagery, assessment of
 spectral signatures using WASI software, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
 based mapping. The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
-enables to analyse environmental changes within seagrass landscapes based on data from various
+enables to analyze environmental changes within seagrass landscapes based on data from various
 sources: aerial and satellite images, geographically referenced maps of Crete island and results of
@@
 there are accepted definitions of landscapes elements within the seagrass meadows \cite{Robbins94}\label{Robbins94}. 
-In general, the structure of seagrass landscape is more simplier than that of the terrestrial
-ecosystems in biodiversity and complexity, however, seagrass landscapes show variation in spatial
-patterns over different special scales. Complexity of the landscape of seagrass meadow show the
+In general, the structure of seagrass landscape is simpler than that of the terrestrial
+ecosystems in biodiversity and complexity; however, seagrass landscapes show variation in spatial
+patterns over different special scales. The complexity of the landscape of seagrass meadow is shown by the
 measure of patches in size and shape, expressed in ratio of patch perimeter to area. The fragmentation
@@
 The general principles of the hierarchy within the seagrass landscapes are based upon the quantitative
-analysis of the spatial patterns, consisted by components and separate elements. Thus, the bunches of
+analysis of the spatial patterns, consisted by components and separate elements. Thus, bunches of
 individual shoots construct patches, the first hierarchical level. Patches are arranged into discrete
 clumps of mattes (at a scale of centimeters to meters) which, in turn, make up beds with l-100m in
-diameter. Finally, seagrass beds, are arranged into meadows that may extend over kilometer-wide
+diameter. Finally, seagrass beds are arranged into meadows that may extend over kilometer-wide
 areas, historically defined as landscapes \cite{Robbins94}\label{Robbins94}.
@@
 
-Besides spatial structure of the seagrass meadows, there are strong and complex patterns of depth zonation, specific fo individual
-seagrass species, but such detailed clasiification goes beyond the scope of the current MSc research.
+Besides spatial structure of the seagrass meadows, there are strong and complex patterns of depth zoning, specific to individual
+seagrass species, but such detailed classification goes beyond the scope of the current MSc research.
 Unique feature of seagrass meadows, most characteristic for shallow areas, is dynamics of their
@@
 however they are often interrupted by gaps and channels of open spaces generated by the complex,
-disturbin sedimentation processes, turbulence and turbidity regime of waters and landscape dynamics,
+disturbing sedimentation processes, turbulence and turbidity regime of waters and landscape dynamics,
 which form gaps and channels within meadows, making them “patchy-looking”, i.e. diversified by
-separate mattes (Fig.\ref{fig:38} b). The traditional definition for gap in vegetation cover is “disturbancegenerated
+separate mattes (Fig.\ref{fig:38} b). The traditional definition for gap in vegetation cover is “disturbance generated
 openings in either floral or faunal cover” \cite{Connell78}\label{Connell78}.
@@
 deposition of sediments during and after floods\cite{Bell99}\label{Bell99}.
-Increased nutrient sedimentation, expecially phosphorus, were explored \cite{Jensen01}\label{Jensen01} as a
-potential mechanism for increasing patch dynamics and morphological plastisity within seagrass
+Increased nutrient sedimentation, especially phosphorus, were explored \cite{Jensen01}\label{Jensen01} as a
+potential mechanism for increasing patch dynamics and morphological plasticity within seagrass
 meadows. Finally, increasing the degree of fragmentation of the landscapes of P.oceanica meadows
@@
 
-\section{appendices}
+\section{appendices}\label{app}
 
@@
 
-\subsection{\textit{Appendix-2}. Instrumental adjustment and tuning (Trios-RAMSES setup)}
+\subsection[\textit{Appendix-2}. Instrumental adjustment and tuning]{\textit{Appendix-2}. Instrumental adjustment and tuning \\ (Trios-RAMSES setup)}
 %\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}
@@
 
-\subsection{\textit{Appendix-3}. Types of seagrass structural patterns, Ligaria beach, Crete}
+\subsection[\textit{Appendix-3}. Types of seagrass structural patterns]{\textit{Appendix-3}. Types of seagrass structural patterns, Ligaria beach, Crete}
 %\appendixname{\textbf{3: \\Types of seagrass structural patterns, Ligaria beach, Crete}}
@@
 		\caption{Statistical analysis of the measurement data: spectral reflectance of the P.oceanica. 
-		Visualisation of the interquartile ranging.
+		Visualization of the interquartile ranging.
 Example of data set 401-420.
-Shown midspread of  statistical quartiles Q1 and Q3, min and max values within the range.
+Shown mid-spread of  statistical quartiles Q1 and Q3, min and max values within the range.
 \href{http://www.gnuplot.info/}{Gnuplot}}
@@
 		\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
-		Visualisation of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
+		Visualization of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
 		\label{fig:A.14}
@@
 	\caption{Fragment of the statistical analysis of the P.oceanica reflectance. Example of data set 401-420. 
-		Visualisation of the measured data together with styatistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
+		Visualization of the measured data together with statistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
 	\label{fig:A.15}
@@
 Series 126-150.
-Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
+Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
 \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
@@
 Series 151-200.
-Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
+Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
 \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
@@
 	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
-	\caption{Results of the \textit{periodogramm} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
 	\label{tab:30}
@@
 		\includegraphics[scale=0.25]{G-period-sand30.jpg}
-		\caption{Plot illustrating periodogramm for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
+		\caption{Plot illustrating periodogram for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
 	\end{center}	
@@
 
-\begin{table}
+\begin{table}[H]
 	\caption{Available broadband Landsat satellite images covering the research area of Crete island}\label{tab:34}
@@
 \end{table}
-
-\begin{figure}
+\pagebreak
+\newpage
+
+\begin{figure}[H]
 	\centering
@@
 \end{figure}
+\pagebreak
+\newpage
+
+\subsection{\textit{Appendix-9}. Analysis and classification of the satellite and aerial images}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.23]{OpenEV-2.jpg}
+		\caption{The point querying shows the selected points and their coordinates within the area of seagrass meadow (green). \href{http://openev.sourceforge.net/}{OpenEV}}\label{fig:A.39}	
+	\end{center}	
+\end{figure}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.23]{OpenEV-3.jpg}
+	\caption{\textit{Logarithmic Enhancement to Raster}, applied to the aerial Google Earth image: the seagrass meadow can now be easily seen as a bright spot of purple color.
+\href{http://openev.sourceforge.net/}{OpenEV}}\label{fig:A.40}
+	\end{center}	
+\end{figure}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.25]{OpenEV-1.jpg}
+		\caption{Raster properties dialog: visualization and spatial info about the image (projection UTM, zone 35, datum WGS-84, etc.)  \href{http://openev.sourceforge.net/}{OpenEV})}\label{fig:A.41}	
+	\end{center}	
+\end{figure}
+
 \end{appendices}

### 2011-01-26 MScDiplom_Lemenkova.tex
added first ILWIS screenshot

@@
 
-Thesis submitted to the University of Twente, Faculty ITC, in partial fulfilment of the requirements
+Thesis submitted to the University of Twente, Faculty ITC, in partial fulfillment of the requirements
 for the degree of Master of Science in Geo-information Science and Earth Observation for
@@
 \vspace{1mm}\\
-Many thanks for Dr.\textit{Dingtian Yang} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectances of Thalassia seagrass, which enabled to compare spectra of various seagrass types.
+Many thanks for Dr.\textit{Dingtian Yang} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.
 
@@
 
-		I herewith certify that although I confered with other researchers while preparing this document, and used a range of various research and literature 			resources, cited in the text, the content of this MSc Thesis document remains my own, original work, written by me personally.
+		I herewith certify that although I conferred with other researchers while preparing this document, and used a range of various research and literature 			resources, cited in the text, the content of this MSc Thesis document remains my own, original work, written by me personally.
 \\
@@
 		\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
-		\item \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and 					classification
+		\item \href{http://www.ilwis.org/}{ILWIS} , \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
 		\item \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}, for modelling optical water 					properties under changing environment
@@
 		\caption{Raster properties dialog: visualization and spatial info about the image (projection UTM, zone 35, datum WGS-84, etc.)  \href{http://openev.sourceforge.net/}{OpenEV})}\label{fig:A.41}	
+	\end{center}	
+\end{figure}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.25]{Ilwis-2.jpg}
+		\caption{Color composite image composed of 3 images of Cretan shelf, Google Earth. \href{http://www.ilwis.org/}{ILWIS}}\label{fig:A.42}	
 	\end{center}	

### 2011-01-27 MScDiplom_Lemenkova.tex
moved 'Data pre-processing' to Methods; changed latin words to Italic font

@@
 chemical and physical parameters of the water: salinity, waves \cite{McKenzie02}\label{McKenzie02}.
-The seagrass Posidonia oceanica (further \textit{P.oceanica}) is a key species to inhabit littoral of the
+The seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}) is a key species to inhabit littoral of the
 Mediterranean Sea, see Fig. \ref{fig:1}
@@
 	\subfloat {\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-1-1b.jpg}}
-	\caption{Seagrass Posidonia Oceanica}
+	\caption{Seagrass \textit{Posidonia Oceanica}}
 	\label{fig:1}
@@
 Globally, there are 58 recognized and described seagrass species (Fig. \ref{fig:2}), belonging to two orders
-(Hydrocharitales and Najadales), four families (Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
-and Zosteraceae), and 12 genera (Enhalus, Thalassia, Halophila, Posidonia, Syringodium, Halodule,
-Cymodocea, Amphibolis, Thalassodendron, Zostera, Heterozostera and Phyllospandix) \cite{Kuo89}\label{Kuo89}.
+(\textit{Hydrocharitales} and \textit{Najadales}), four families (\textit{Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
+and Zosteraceae}), and 12 genera (\textit{Enhalus, Thalassia, Halophila, Posidonia, Syringodium, Halodule,
+Cymodocea, Amphibolis, Thalassodendron, Zostera, Heterozostera} and \textit{Phyllospandix}) \cite{Kuo89}\label{Kuo89}.
 
@@
 \subsubsection{Environmental vulnerability of the seagrasses}
-Meadows of P. oceanica are subjected to the human activities, as they occur in coastal areas, where
+Meadows of \textit{P. oceanica} are subjected to the human activities, as they occur in coastal areas, where
 they can be affected both directly \cite{Meinesz91}\label{Meinesz91} or indirectly, through the impact on the
@@
 farm-induced decline of the seagrass meadows \cite{Diaz-Almela06}\label{Diaz-Almela06}reports the relationships of
-fish farm organic and nutrient content in the sediments with dynamics of the key seagrass species (P.
-oceanica) in the Mediterranean Sea. Nowadays P.oceanica is in the alarming state of regression due
+fish farm organic and nutrient content in the sediments with dynamics of the key seagrass species (\textit{P.
+oceanica}) in the Mediterranean Sea. Nowadays P.oceanica is in the alarming state of regression due
 to the deterioration of the environment in the Mediterranean Sea \cite{Ribed02}\label{Ribed02}. Due to these reasons,
@@
 competition for existence among species. Thus, meadows of P.oceanica in the Mediterranean Sea are
-presently facing invasion by alien algal species, particularly in areas where P.oceanica is already
+presently facing invasion by alien algal species, particularly in areas where \textit{P.oceanica} is already
 degrading, stressed, have gaps and patchy structure in meadows and show other signs of regression
@@
 
-\subsubsection{General characteristics of Posidonia oceanica}
-The endemic Mediterranean seagrass Posidonia oceanica (further \textit{P.oceanica}), is a main species in
+\subsubsection{General characteristics of \textit{P.oceanica}}
+The endemic Mediterranean seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}), is a main species in
 marine coastal environment of Greece, forming, despite its slow growth, the largest, most widespread,
@@
 
-The dominant and most productive coastal ecosystem of the Mediterranean, P.oceanica is
+The dominant and most productive coastal ecosystem of the Mediterranean, \textit{P.oceanica} is
 spatially restricted to the Mediterranean area (Fig.\ref{fig:5}), with its extension limited by the western part
@@
 Mediterranean waters, thus decreasing its temperature.
-Morphological P.oceanica consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots
+Morphological \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots
 and short rhizomes (0.5-2.0 mm). The leaves are, perhaps, the most particular characteristics of
@@
 	\includegraphics[scale=0.15]{Fig-1-6.jpg}
-	\caption{\textit{P.oceanica}.\\
+	\caption{\textit{	extit{P.oceanica}}.\\
 	Source \cite{Luque04}\label{Luque04}}
@@
 
-P.oceanica, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:6}): having
+\textit{P.oceanica}, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:6}): having
 usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}).
-Growing P.oceanica make a meadows which, in turn, consist of smaller patches, called “matte”
+Growing \textit{P.oceanica} make a meadows which, in turn, consist of smaller patches, called “matte”
 (Fig.\ref{fig:7}), a monumental construction made by the growth of rhizomes and leaves with entangled roots
@@
 	\includegraphics[scale=0.15]{Fig-1-7.jpg}
-	\caption{Scheme of matte structure of \textit{	P.Oceanica}.\\ Source \cite{Pergent90}\label{Pergent90}}
+	\caption{Scheme of matte structure of \textit{	P.oceanica}.\\ Source \cite{Pergent90}\label{Pergent90}}
 	\label{fig:7}
@@
 	\includegraphics[scale=0.15]{Fig-1-8.jpg}
-	\caption{Structure and components of P.oceanica \\ Source \cite{DiCarlo04}\label{DiCarlo04}}
+	\caption{Structure and components of \textit{P.oceanica} \\ Source \cite{DiCarlo04}\label{DiCarlo04}}
 	\label{fig:8}
@@
 from the following sources:
-i) spectra of P.oceanica, carbonate sand, silt and other seafloor types
+i) spectra of 	extit{P.oceanica}, carbonate sand, silt and other seafloor types
 ii) satellite imagery: Landsat, ASTER, MODIS, MERIS
@@
 v) vector GIS layers.
-The research aims to explore the dynamics of the environmental changes of the P.oceanica meadows
+The research aims to explore the dynamics of the environmental changes of the \textit{P.oceanica} meadows
 over time (2000-2010), as well as the extent of the spatial distribution over the whole island. The
 main research objective is monitoring the seagrass \textit{P.oceanica} along the coasts of Crete, in order to
-analyse the environmental changes of the P.oceanica meadows over the last 10 years (2000-2010)
+analyse the environmental changes of the 	extit{P.oceanica} meadows over the last 10 years (2000-2010)
 and the spatial extent of its current distribution.
@@
 Crete Island
-2) Monitoring the dynamics of the environmental changes of the P.oceanica meadows over
+2) Monitoring the dynamics of the environmental changes of the \textit{P.oceanica} meadows over
 time (2000-2010) in Ligaria beach, Agia Pelagia district.
@@
 3. To use methods of the in situ diving observations and underwater videometric measurements
-by Olympus camera in order to receive large-scale imagery of the P.oceanica mattes
+by Olympus camera in order to receive large-scale imagery of the 	extit{P.oceanica} mattes
 4. To perform images classification (supervised and unsupervised) for the thematic mapping of
-the P.oceanica seagrass distribution along the coasts of northern Crete.
+the 	extit{P.oceanica} seagrass distribution along the coasts of northern Crete.
 
@@
 seagrass researchers. Applications of the in-situ seagrass observations of the structuring epiphyte
-community composition in the P.oceanica ecosystems in Mediterranean Sea is, for example,
+community composition in the \textit{P.oceanica} ecosystems in Mediterranean Sea is, for example,
 described by Villegas \cite{Villegas06}\label{Villegas06}. Realization of traditional methods for mapping seagrass usually
@@
 orbital repeat cycle for nadir viewing and imagery with a spatial resolution 20 - 2.5 m
-\cite{SPOT}\label{SPOT}. \href{http://www.spotimage.fr/}{SPOT} imagery was used for mapping beds of Posidonia oceanica in the
+\cite{SPOT}\label{SPOT}. \href{http://www.spotimage.fr/}{SPOT} imagery was used for mapping beds of \textit{P.oceanica} in the
 Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}.
@@
 In the current  research we try to combine various techniques from mentioned above for the assessment of the seagrass distribution in the Mediterranean environment, for the case study of Crete.
-We mentioned in \nameref{sec:0} the software that we used for logically separated research sections, which required various approach.
+We mentioned in the \nameref{sec:0} the software that we used for logically separated research sections, which required various approach.
 Thus, for example,  spectral analysis and assessment of spectral signatures is technically based on WASI, statistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://gretl.sourceforge.net/}{Gretl} and \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
@@
 field campaign has been carried out during the September-October period 2010. 
-The information about the location of the seagrass (mostly represented by the P.oceanica species) is useful for the
+The information about the location of the seagrass (mostly represented by the 	extit{P.oceanica} species) is useful for the
 understanding of the relationship between the spatial distribution of the seagrass and the environment
@@
 consequent images, completely covering the area under the boat path. The received data contain
-information on seagrass presence within the study area, distribution of seagrass P.oceanica meadows
+information on seagrass presence within the study area, distribution of seagrass \textit{P.oceanica} meadows
 and nature of the seafloor cover types: rocks, sandy, mixed (Fig. \ref{fig:A5}). Seagrass species on the Ligaria and Agia
-Pelagia beach consist of P.oceanica. The types of sediment on the Ligaria vary from coarse sand,
-P.oceanica patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:A.6}).
-
-\subsection{Data of spectral measurements}
-
-\subsection{Google Earth aerial imagery for seagrass mapping}
+Pelagia beach consist of 	extit{P.oceanica}. The types of sediment on the Ligaria vary from coarse sand,
+\textit{P.oceanica} patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:A.6}).
+
+\subsection{Data pre-processing}
+A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of \textit{P.oceanica} made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
+\ref{tab:A.8}), 1-445 for 14th October for 	\textit{P.oceanica}; 1-106 for water without sediments, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
+\ref{tab:17}).
+All these data have been analyzed and statistically tested.
+We measured mean, extreme values (min-max), median and statistical quartiles, in order to visualize distribution of the values at various spectral wavelengths.
+Carbonate sand was measured at wavelengths 402-750 nm, while \textit{P.oceanica} - at the interval 318-951 nm.
+We visualized the behavior of different spectra on relevant graphs (Fig.
+\ref{fig:4.4}, \ref{fig:A.35}, \ref{fig:A.28}).
+The available collected data were tested for the spectral variability and separability under varying conditions of different environmental
+constituents (e.g. depth, water content, sun angle), in order to determine the
+potential that it may have on the approaches for further images processing and classification.
+Behavior of the spectra of \textit{P.oceanica} have been tested using datasets for various depths: 0.5, 1.5 and 2.5 meters (Tab.
+\ref{tab:8}, Tab.
+\ref{tab:A.7}).
+The raw
+initial data of measurements of the spectral reflectance have been preprocessed and statistically analyzed thereafter with graphs visualizing mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:4.10}, \ref{fig:28}).
+These data
+have been measured using different step of the wavebands: some measurements
+were made with 3 nm interval, while others – using 4 and 1 nm step.
+Therefore, these data had to be interpolated (Fig. \ref{fig:A.10}) and standardized to one
+format, which is values of spectral reflectance with one nm step. For the
+interpolation we used script written on Python programming language that allowed receiving more 
+detailed data by interpolating them from 3 and 4 nm step up to 2 nm (Fig.\ref{fig:A.8}).
+The interpolated data contained spectral measurements of the seagrass 	\textit{P.oceanica} (measured at Agia
+Pelagia beach, Crete), sand (measured at Agia Pelagia beach, Crete), silt and default artificial
+spectrum of constant albedo at WASI (Fig.\ref{fig:A.9}). 
+We included measurements of the Thalassia seagrass
+(measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectance of
+different seagrass species under various environmental conditions.
+The most suitable wavebands for the seagrass monitoring usually lay between 400 and 700 nm,
+which can be concluded by the visual examination, comparison and analysis of the different spectra of
+the seagrasses. Therefore, we chosen the spectra 400-750 as the most appropriate range for
+further research experiment. The results of the linear interpolation (Fig.\ref{fig:A.35}) demonstrate values of the
+sand spectral reflectance with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
+The initial measured data were stored in raw-oriented format, so that re-formatting them into the
+column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:A.11}). The next step included
+calculation of the median, mean, quartiles and other statistical values at every data set (see \nameref{app}, Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
+After the preliminary analysis, the measured data were visualized using Gnuplot program, which
+enables fine plotting of various datasets together: Fig \ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:28}.
+The most acceptable method of interpolation was B\'ezier curve (Fig.
+\ref{fig:A.18}, Fig.
+\ref{fig:A.16}),
+as it has trend-friendly graph better showing the general behavior of the curves at different wavebands
+comparing to splines (see \nameref{app} for more results). 
+Therefore, after several experiments with
+various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have chosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contains convex
+hulls made on its control points, and therefore is best suitable for our case: analysis of optical
+properties of seawater.
+
+\subsection[Google Earth aerial imagery]{Google Earth aerial imagery for seagrass mapping}
 Apart from the satellite imagery, the aerial photographs from the Google Earth provide
@@
 	\item[] 1.
-Optical spectra of P.oceanica, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig.
+Optical spectra of \textit{P.oceanica}, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig.
 \ref{fig:A.23})
@@
 
-\subsection{Data pre-processing}
-A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of P.oceanica made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
-\ref{tab:A.8}), 1-445 for 14th October for P.oceanica; 1-106 for water without sediments, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
-\ref{tab:17}).
-All these data have been analyzed and statistically tested.
-We measured mean, extreme values (min-max), median and statistical quartiles, in order to visualize distribution of the values at various spectral wavelengths.
-Carbonate sand was measured at wavelengths 402-750 nm, while P.oceanica - at the interval 318-951 nm.
-We visualized the behavior of different spectra on relevant graphs (Fig.
-\ref{fig:4.4}, \ref{fig:A.35}, \ref{fig:A.28}).
-The available collected data were tested for the spectral variability and separability under varying conditions of different environmental
-constituents (e.g. depth, water content, sun angle), in order to determine the
-potential that it may have on the approaches for further images processing and classification.
-Behavior of the spectra of P.oceanica have been tested using datasets for various depths: 0.5, 1.5 and 2.5 meters (Tab.
-\ref{tab:8}, Tab.
-\ref{tab:A.7}).
-The raw
-initial data of measurements of the spectral reflectance have been preprocessed and statistically analyzed thereafter with graphs visualizing mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:4.10}, \ref{fig:28}).
-These data
-have been measured using different step of the wavebands: some measurements
-were made with 3 nm interval, while others – using 4 and 1 nm step.
-Therefore, these data had to be interpolated (Fig. \ref{fig:A.10}) and standardized to one
-format, which is values of spectral reflectance with one nm step. For the
-interpolation we used script written on Python programming language that allowed receiving more 
-detailed data by interpolating them from 3 and 4 nm step up to 2 nm (Fig.\ref{fig:A.8}).
-The interpolated data contained spectral measurements of the seagrass P.oceanica (measured at Agia
-Pelagia beach, Crete), sand (measured at Agia Pelagia beach, Crete), silt and default artificial
-spectrum of constant albedo at WASI (Fig.\ref{fig:A.9}). 
-We included measurements of the Thalassia seagrass
-(measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectance of
-different seagrass species under various environmental conditions.
-The most suitable wavebands for the seagrass monitoring usually lay between 400 and 700 nm,
-which can be concluded by the visual examination, comparison and analysis of the different spectra of
-the seagrasses. Therefore, we chosen the spectra 400-750 as the most appropriate range for
-further research experiment. The results of the linear interpolation (Fig.\ref{fig:A.35}) demonstrate values of the
-sand spectral reflectance with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
-The initial measured data were stored in raw-oriented format, so that re-formatting them into the
-column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:A.11}). The next step included
-calculation of the median, mean, quartiles and other statistical values at every data set (see \nameref{app}, Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
-After the preliminary analysis, the measured data were visualized using Gnuplot program, which
-enables fine plotting of various datasets together: Fig \ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:28}.
-The most acceptable method of interpolation was B\'ezier curve (Fig.
-\ref{fig:A.18}, Fig.
-\ref{fig:A.16}),
-as it has trend-friendly graph better showing the general behavior of the curves at different wavebands
-comparing to splines (see \nameref{app} for more results). 
-Therefore, after several experiments with
-various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have chosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contains convex
-hulls made on its control points, and therefore is best suitable for our case: analysis of optical
-properties of seawater.
 %
@@
 manually. However, the most important, major factors affecting the in-situ conditions are water depth (Fig.\ref{fig:4.9})
-and cover fraction of the seafloor types: P.oceanica and carbonate sand .
+and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand .
 
@@
 A statistical analysis of WASI-simulated spectral reflectance has been used in order to answer the first
-research question: whether the P.oceanica spectra is spectrally distinct at varying environmental in-situ
+research question: whether the 	extit{P.oceanica} spectra is spectrally distinct at varying environmental in-situ
 conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
@@
 water reflectance towards changing conditions of the water column depth and seafloor bottom cover
-fraction (seagrass P.oceanica, carbonate sand and silt).
+fraction (seagrass 	extit{P.oceanica}, carbonate sand and silt).
 The analysis of the spectral signatures of the seagrass and sand clearly shows (Fig.\ref{fig:27}) that seagrass
@@
 from 4.4percent  minimal up to 29 percent \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as 35-
-45° and reflection factor 0.0201. P.oceanica. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
+45° and reflection factor 0.0201. \textit{P.oceanica}. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
 The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:4.8}).
@@
 	\subfloat [Sand]{\label{fig:34a}\includegraphics[width=0.4\textwidth]{Fig-27a.jpg}}
-	\subfloat [\textit{P.oceanica}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
+	\subfloat [\textit{	extit{P.oceanica}}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
 	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
@@
 The total amount of measured data was large and included following datasets made using
-hyperspectral radiometer Ramses: 350 measurement sets of P.oceanica reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
-sets of P.oceanica reflectance for 15th Oct (Tab.\ref{tab:A.8}), 84 datasets for seawater reflectance with sediments, 105
+hyperspectral radiometer Ramses: 350 measurement sets of 	extit{P.oceanica} reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
+sets of 	extit{P.oceanica} reflectance for 15th Oct (Tab.\ref{tab:A.8}), 84 datasets for seawater reflectance with sediments, 105
 datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonate sand.
@@
 seawater with and without sediment (e.g. Fig..\ref{fig:A.16}).
-In the statistical analysis of the raw observed data pool of the values of spectral reflectance (about 400 single measurements for P.oceanica, e.g.
+In the statistical analysis of the raw observed data pool of the values of spectral reflectance (about 400 single measurements for \textit{P.oceanica}, e.g.
 Tab.\ref{tab:A.8}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
@@
 Examples are the mean and variance of a probability distribution of measurements dataset (for example, Tab.\ref{tab:4.1}).
-We divided the total data pool into datasets of 25 measurements for more convenience and ease of use, both for P.oceanica and carbonate sand.
+We divided the total data pool into datasets of 25 measurements for more convenience and ease of use, both for \textit{P.oceanica} and carbonate sand.
 The mean of a probability distribution of spectral reflectance describes the average value of the random variable over all of its possible realizations (\nameref{app}).
@@
 
-Thus, we collect a number of realizations (a sample of 25 measurements for both sand and P.oceanica) and then estimate the statistical parameters (\nameref{app}). \\
-Finally we analyzed about 400 datasets totally (example of carbonate sand on Fig.\ref{fig:4.14}, P.oceanica - on  Fig.\ref{fig:4.13}) made during different days. 
+Thus, we collect a number of realizations (a sample of 25 measurements for both sand and \textit{P.oceanica}) and then estimate the statistical parameters (\nameref{app}). \\
+Finally we analyzed about 400 datasets totally (example of carbonate sand on Fig.\ref{fig:4.14}, \textit{P.oceanica} - on  Fig.\ref{fig:4.13}) made during different days. 
 Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random. 
 Thus, we analyzed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implemented in Gretl software (Tab.\ref{tab:19}). 
-Their probability expresses uncertainty about the possible values of the spectral reflectance of P.oceanica and sand, respectively.
+Their probability expresses uncertainty about the possible values of the spectral reflectance of 	extit{P.oceanica} and sand, respectively.
 There is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
@@
 \\
-As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand.
+As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{	extit{P.oceanica}} and Tab.\ref{tab:28} for carbonate sand.
 The \textit{autocorrelation}, a cross-correlation of an observed value with itself, made on two different days (14th and 15th October),
@@
 spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
-P.oceanica spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
+	extit{P.oceanica} spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
 more than 0.20 nm, except for borders), we did not extend the y-scale up to 100 percent, in order to
 focus on the most necessary area of spectral values (Fig.\ref{fig:4.14}, Fig.\ref{fig:4.4} and \nameref{app}). The mean values are
-highlighted using bold red points, as we used these values for plotting the final graph of P.oceanica spectra in
+highlighted using bold red points, as we used these values for plotting the final graph of 	extit{P.oceanica} spectra in
 Ligaria beach, Agia Pelagia.
@@
 		\includegraphics[scale=0.45]{GNU-13.jpg}
-		\caption{Multiplot of spectral reflectance of P.oceanica. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
+		\caption{Multiplot of spectral reflectance of 	extit{P.oceanica}. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
@@
 Assuming that the null hypothesis, Hypothesis Ho, is true, which is “the spectral distinguishability of
-the seagrass P.oceanica from other seafloor types is not changing with varying in-situ conditions,
+the seagrass 	extit{P.oceanica} from other seafloor types is not changing with varying in-situ conditions,
 Ho: μ1 =μ2 =μ3=...= μn.”, what is the probability of observing the values of spectral reflectance of
-P.oceanica at various depths, for the test statistic, that is at least as extreme as the values of spectral
-reflectance of P.oceanica at these depths, that was actually observed? The alternative Hypothesis Ha
-Fig.15. Multiplot display of spectral reflectance of P.oceanica. Measurement sets 1-100, 15th
+	extit{P.oceanica} at various depths, for the test statistic, that is at least as extreme as the values of spectral
+reflectance of 	extit{P.oceanica} at these depths, that was actually observed? The alternative Hypothesis Ha
+Fig.15. Multiplot display of spectral reflectance of 	extit{P.oceanica}. Measurement sets 1-100, 15th
 October. Bold red dots show the mean values within each dataset; vertical areas: quartiles Q1-Q2.
-claims the opposite statement: “the spectral discernibility of the seagrass P.oceanica is distinctly
+claims the opposite statement: “the spectral discernibility of the seagrass 	extit{P.oceanica} is distinctly
 changing with varying in-situ conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
 Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.
-To analyze average values of spectral reflectance of P.oceanica by means of the measured sets we calculated 
+To analyze average values of spectral reflectance of 	extit{P.oceanica} by means of the measured sets we calculated 
 mean values for the total set of measurements 1 – 350 (15th October): mean of min and max values, 
@@
 The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (\nameref{app}). 
-As a result, we have mean values of spectral reflectance of P.oceanica for extreme, average, median and both quartiles,
+As a result, we have mean values of spectral reflectance of 	extit{P.oceanica} for extreme, average, median and both quartiles,
 independent of the individual sets 1-350.
@@
 We applied ANOVA (ANalysis Of Variance) testing for the analysis of the probability of the
-spectral reflectance of P.oceanica to be more or less spectrally disticnt from other seafloor cover
+spectral reflectance of 	extit{P.oceanica} to be more or less spectrally disticnt from other seafloor cover
 types with changing depth. W are checking the statement Hypothesis Ho, is true, which is “the
-spectral distinguishability of the seagrass P.oceanica from other seafloor types is not changing with
+spectral distinguishability of the seagrass 	extit{P.oceanica} from other seafloor types is not changing with
 varying in-situ conditions, Ho: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
@@
 (less than 1 in a 10 chance). The one-way ANOVA highlighted a significant difference between data
-of spectral reflectances of P.oceanica at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
-Hypothesis Ha=“the spectral discernibility of the seagrass P.oceanica is distinctly changing with
+of spectral reflectances of 	extit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
+Hypothesis Ha=“the spectral discernibility of the seagrass 	extit{P.oceanica} is distinctly changing with
 varying in-situ conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
@@
 of a difference between sample means, is used for the data analysis. In our case the Student t-test demonstrates, if the variation
-between two analyzed groups – values of spectral reflectance of seagrass P.oceanica and sand - is
+between two analyzed groups – values of spectral reflectance of seagrass 	extit{P.oceanica} and sand - is
 significant. Therefore, we use Student t-test to compare two sets of quantitative data of spectral
-reflectance of P.oceanica and sand, respectively, with samples collected independently of one
+reflectance of 	extit{P.oceanica} and sand, respectively, with samples collected independently of one
 another. The Student-t test can be performed knowing just the means, standard deviation, and number
@@
 As a result of the statistical testing, we came to the following conclusion.
-The Hypothesis Ha is true, which claims that the spectral discernibility of the seagrass P.oceanica is
+The Hypothesis Ha is true, which claims that the spectral discernibility of the seagrass 	extit{P.oceanica} is
 distinctly changing with varying in-situ conditions, i.e. increasing depth, Ho: μ1≠μ2≠μ3≠...≠ μn .
@@
 \rowcolors{1}{Khaki}{LightYellow}
-\caption{Statistical analysis of the measurements of spectral reflectance of P.oceanica (fragment)}
+\caption{Statistical analysis of the measurements of spectral reflectance of 	extit{P.oceanica} (fragment)}
 \centering
@@
 \subsubsection{Results of the remote sensing application}
-The in situ spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
+The in situ spectral reflectance data of \textit{	extit{P.oceanica}} and sand were used to model these seafloor cover
 types as the different sensors would percept them: MODIS, ASTER, MERIS, SeaWiFS and CZCS.\\
@@
 The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
-investigate the distribution of the seagrass P.oceanica within the research area.
+investigate the distribution of the seagrass 	extit{P.oceanica} within the research area.
 In the image processing part of the research supervised classification has been applied to the aerial
@@
 detected using different bands combinations, masked and studied for the estimation of the changes in
-the areas. The classification is based on the spectral properties of the \textit{P.oceanica}, such as brightness,
+the areas. The classification is based on the spectral properties of the \textit{	extit{P.oceanica}}, such as brightness,
 colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping
@@
 for each photograph by classification a series of polygons characteristic of each of the sea floor
-bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
+bottom types: sandy surface, seagrass bed for each species including \textit{	extit{P.oceanica}} (meadows), patchy
 seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).
@@
 	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
-	\caption{Seafloor types of the Arina beach: \textit{P.oceanica,} sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
+	\caption{Seafloor types of the Arina beach: \textit{	extit{P.oceanica},} sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
 	\label{fig:38}
@@
 potential mechanism for increasing patch dynamics and morphological plasticity within seagrass
-meadows. Finally, increasing the degree of fragmentation of the landscapes of P.oceanica meadows
+meadows. Finally, increasing the degree of fragmentation of the landscapes of 	extit{P.oceanica} meadows
 can be caused by the invasion of alien species, such as Cymodocea nodosa, Caulerpa prolifera,
-Caulerpa taxifolia. Invaders are in general strong colonizers comparing to native P.oceanica: they
+Caulerpa taxifolia. Invaders are in general strong colonizers comparing to native 	extit{P.oceanica}: they
 occupy much greater habitat space within the regressed meadows of stressed native seagrass
 \cite{Montefalcone10}\label{Montefalcone10}. However, on the northern coasts of Crete the only dominating seagrass
-species is P.oceanica.
+species is 	extit{P.oceanica}.
 Morphological differences in scale of seagrass landscape formations, discussed above, cause need for
@@
 \markright{Keywords}
-Seagrass, \textit{\textit{P.oceanica}}, analysis of spectral reflectance, remote sensing, monitoring, Google Earth, Crete
+Seagrass, \textit{\textit{	extit{P.oceanica}}}, analysis of spectral reflectance, remote sensing, monitoring, Google Earth, Crete
 \pagebreak
@@
 		     PositionLongitude & 25.01269=25°0'45.6834" & 25.0098 = 25°0'35.2794" \\ \hline
-		     Comment & P.oceanica & P.oceanica \\ \hline
+		     Comment & 	extit{P.oceanica} & 	extit{P.oceanica} \\ \hline
 		     CommentSub1 & 0.5m depth & 2.5m depth \\ \hline
@@
 	     PositionLongitude & 25.0098= 25°0'35.2794" & 25.012661=25°0'45.5796" \\ \hline
-	     Comment & P.oceanica & P.oceanica \\ \hline
+	     Comment & 	extit{P.oceanica} & 	extit{P.oceanica} \\ \hline
 	     CommentSub1 & 3.5m depth & 1.5m depth \\ \hline
@@
 		\includegraphics[scale=0.20]{App-1.jpg}
-		\caption{Values of spectral reflectances of seagrass P.oceanica (fragment)}
+		\caption{Values of spectral reflectances of seagrass 	extit{P.oceanica} (fragment)}
 		\label{fig:A.11}
@@
 		\includegraphics[scale=0.20]{GNU-401-420.jpg}
-		\caption{Spectral reflectance of P.oceanica. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
+		\caption{Spectral reflectance of 	extit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
 		\label{fig:A.12}
@@
 		\includegraphics[scale=0.20]{GNU-401-420-candles.jpg}
-		\caption{Statistical analysis of the measurement data: spectral reflectance of the P.oceanica. 
+		\caption{Statistical analysis of the measurement data: spectral reflectance of the 	extit{P.oceanica}. 
 		Visualization of the interquartile ranging.
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral 			reflectance of \textbf{P.oceanica}.
+	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral 			reflectance of \textbf{	extit{P.oceanica}}.
 \href{http://www.spss.com/}{SPSS}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{P.oceanica}.
+	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{	extit{P.oceanica}}.
 \href{http://www.spss.com/}{SPSS}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
+	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{	extit{P.oceanica}}. \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\includegraphics[scale=0.20]{GNU-Stat-M-401-420.jpg}
-	\caption{Fragment of the statistical analysis of the P.oceanica reflectance. Example of data set 401-420. 
+	\caption{Fragment of the statistical analysis of the 	extit{P.oceanica} reflectance. Example of data set 401-420. 
 		Visualization of the measured data together with statistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
+	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{	extit{P.oceanica}} at various depths: 0.5, 1.5 and 2.5 meters.
 \href{http://www.spss.com/}{SPSS}}\label{tab:8}
@@
 	\end{center} 
-	P more than .05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5).
+	P more than .05, which means that there is a significant difference in radiance of 	extit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).
 
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
+	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{	extit{P.oceanica}} at various depths: 0.5, 1.5 and 2.5 meters.
 \href{http://www.spss.com/}{SPSS}}\label{tab:A.7}
@@
 	\end{center}
-	P$>$0.05, which means that there is a significant difference in radiance of P.oceanica at three different depth (0.5, 1.5 and 2.5).
+	P$>$0.05, which means that there is a significant difference in radiance of 	extit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).
 
@@
 \begin{longtable}{|c|c|c|c|c|c|c|}
-	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October}\label{tab:A.8} \\
+	\caption{Results of the statistical analysis of spectral reflectance of \textbf{	extit{P.oceanica}}, sets 1-350).
+Wavelength step: 3 nm.
+Measured on Agia Pelagia beach, 15th October}\label{tab:A.8} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 
-	\multicolumn{7}{c}{Table A.8: Spectral reflectance of \textbf{P.oceanica} \textit{(continued from previous page)}}\\
+	\multicolumn{7}{c}{Table A.8: Spectral reflectance of \textbf{	extit{P.oceanica}} \textit{(continued from previous page)}}\\
 	\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\includegraphics[scale=0.45]{Autocorr.jpg}
-	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
+	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
 	\label{fig:A.21}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3.
+	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{	extit{P.oceanica}} (15.X.), Series:V3.
 \href{http://www.spss.com/}{SPSS}}\label{tab:14}
@@
 	\includegraphics[scale=0.45]{Partial_Corr.jpg}
-	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
+	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{	extit{P.oceanica}} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
 	\label{fig:A.22}
@@
 	\begin{center}
-		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textit{	extit{P.oceanica}} (15.X.).
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Summary of the table with values of the spectral reflectance of \textbf{P.oceanica} measured on three different depths, R}
+	\caption{Summary of the table with values of the spectral reflectance of \textbf{	extit{P.oceanica}} measured on three different depths, R}
 	\begin{center}
@@
 		\includegraphics[scale=0.25]{GNU-11.jpg}
-		\caption{Spectral reflectance of P.oceanica. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
+		\caption{Spectral reflectance of 	extit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
 		\label{fig:A.25}
@@
 		\includegraphics[scale=0.25]{GNU-12.jpg}
-		\caption{Remote sensing reflectance of P.oceanica.
+		\caption{Remote sensing reflectance of 	extit{P.oceanica}.
 Series 1-25.
@@
 \begin{table}[htbp]
-	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 		\includegraphics[scale=0.4]{GNU-14.jpg}
-		\caption{Multiplot of spectral reflectance of P.oceanica.
+		\caption{Multiplot of spectral reflectance of 	extit{P.oceanica}.
 Series 126-150.
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
 \href{http://www.spss.com/}{SPSS}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
+	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
+\href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
 \href{http://www.spss.com/}{SPSS}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
+	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
+\href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 \begin{table}[H]
-	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica} (15.X.).
+	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 		\rowcolors{1}{AntiqueWhite}{LightGoldenrodYellow}
-		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textbf{P.oceanica}.
+		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textbf{	extit{P.oceanica}}.
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 		\includegraphics[scale=0.38]{GNU-15.jpg}
-		\caption{Multiplot of spectral reflectance of P.oceanica.
+		\caption{Multiplot of spectral reflectance of 	extit{P.oceanica}.
 Series 151-200.
@@
 	\begin{center}
-	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{P.oceanica}.
+	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{	extit{P.oceanica}}.
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 	\begin{center}
-	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textbf{P.oceanica}.
+	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica}.
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 \bibitem{Amoutzopoulou-Schina05}Amoutzopoulou-Schina H and S.Haritonidis. 2005. \emph{Distribution and phenology of the
-marine phanerogam \textit{Posidonia oceanica} in the Pagassitikos Gulf, Greece}. Journal of
+marine phanerogam \textit{	extit{Posidonia oceanica}} in the Pagassitikos Gulf, Greece}. Journal of
 Biological Research 4: 203 – 211, 2005 J. \pageref{Amoutzopoulou-Schina05}
@@
 Ardizzone G., Belluscio A. and Maiorano L. 2006. \emph{Long-term change in the structure of
-a \textit{Posidonia oceanica} landscape and its reference for a monitoring plan}. Marine Ecology.
+a \textit{	extit{Posidonia oceanica}} landscape and its reference for a monitoring plan}. Marine Ecology.
 ISSN 0173-9565. \pageref{Ardizzone06}
@@
 Argyrou M., Karakassis I, Duarte C.M. 2006. \emph{Benthic inputs as predictors of seagrass 1
-(Posidonia 2 oceanica) fish farm-induced decline}. http://arxiv.org/ftp/qbio/
+(\textit{Posidonia oceanica}) fish farm-induced decline}. http://arxiv.org/ftp/qbio/
 papers/0611/0611006.pdf \pageref{Diaz-Almela06}
@@
 Francour P., Ganteaume A. and M. Poulain. 1999. \emph{Effects of boats anchoring in
-\textit{Posidonia oceanica} seagrass beds in the Port-Cross National Park (north-western
+\textit{	extit{Posidonia oceanica}} seagrass beds in the Port-Cross National Park (north-western
 Mediterranean Sea)}. Aquatic conservation: Marine and Freshwater Ecosystems: 9: 391-400. \pageref{Francour99}
@@
 Leoni V., Pasqualini V., Pergent-Martini Ch., Vela A., Pergent G. 2006. \emph{Morphological
-responses of \textit{Posidonia oceanica} to experimental nutrient enrichment of the canopy water}.
+responses of \textit{	extit{Posidonia oceanica}} to experimental nutrient enrichment of the canopy water}.
 Journal of Experimental Marine Biology and Ecology xx (2006) xxx–xxx. \pageref{Leoni06}
@@
 Leoni V., Pasqualini V., Pergent-Martini Ch., Vela A., Pergent G. 2007. \emph{Physiological
-responses of \textit{Posidonia oceanica} to experimental nutrient enrichment of the canopy water}.
+responses of \textit{{Posidonia oceanica}} to experimental nutrient enrichment of the canopy water}.
 Journal of Experimental Marine Biology and Ecology 349, 73–83. \pageref{Leoni07}
@@
 Matarrese R., Acquaro M., Morea A., Tijani K., Chiaradia M.T. 2008. \emph{Applications of
-remote sensing techniques for mapping \textit{Posidonia oceanica} meadows}. Geoscience and
+remote sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}. Geoscience and
 Remote Sensing Symposium. ISBN:978-1-4244-2807-6. \pageref{Matarrese08}
@@
 Matarrese R., Acquaro M., Morea A., Khalid T., Chiaradia M.T. 2006 \emph{Applications of remote
-sensing techniques for mapping \textit{Posidonia oceanica} meadows}.
+sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}.
 http://www.igarss08.org/Abstracts/pdfs/3888.pdf \pageref{Matarrese06}
@@
 Micheli C., Paganin P, Peirano A., G. Caye, Meinesz A. and Bianchi C.N. 2005. \emph{Genetic
-variability of \textit{Posidonia oceanica} (L.) Delile in relation to local factors and biogeographic
+variability of \textit{{Posidonia oceanica}} (L.) Delile in relation to local factors and biogeographic
 patterns}. Aquatic Botany 82 (2005) 210–221 \pageref{Micheli05}
@@
 Montefalcone M. 2009.\emph{Ecosystem health assessment using the Mediterranean seagrass
-\textit{Posidonia oceanica}: A review}. Ecological indicators 595 – 604 \pageref{Montefalcone09}
+\textit{{Posidonia oceanica}}: A review}. Ecological indicators 595 – 604 \pageref{Montefalcone09}
 \bibitem{Montefalcone10}
 Montefalcone M., Albertelli G., Morri C. and Bianchi C.N. 2010. \emph{Patterns of wide-scale
-substitution within meadows of the seagrass \textit{Posidonia oceanica} in NW Mediterranean Sea:
+substitution within meadows of the seagrass \textit{{Posidonia oceanica}} in NW Mediterranean Sea:
 invaders are stronger than natives}. Aquatic Conserv: Mar. Freshw. Ecosyst. 20: 507–515. \pageref{Montefalcone10}
@@
 Montefalcone M., Albertelli G., Bianchi C.N., Mariani M., Morri C. 2006. \emph{A new
-synthetic index and a protocol for monitoring the status of \textit{Posidonia oceanica} meadows: a
+synthetic index and a protocol for monitoring the status of \textit{{Posidonia oceanica}} meadows: a
 case study at Sanremo (Ligurian Sea, NW Mediterranean)}. Aquatic Conservation: Marine and
@@
 Pasqualini V., Pergent-Martini C., Clabaut P., Pergent G. 1998(b). \emph{Mapping of
-\textit{Posidonia oceanica} using aerial photographs and side scan sonar: Application off the
+\textit{{Posidonia oceanica}} using aerial photographs and side scan sonar: Application off the
 island of Corsica (France)}. Estuarine Coastal Shelf Sci 47:359–367. \pageref{Pasqualini98b}
@@
 \bibitem{Pergent90}
-Pergent G. 1990. \emph{Lepidochronological analysis of the seagrass \textit{Posidonia oceanica} (L.)
+Pergent G. 1990. \emph{Lepidochronological analysis of the seagrass \textit{{Posidonia oceanica}} (L.)
 Delile: a standardized approach}. Aquatic Botany, 37 (1990) 39-54 39 Elsevier Science
@@
 Pergent-Martini C., Pasqualini V., Ferrat L and Pergent G. 2006. \emph{Ecological Data in
-Integrated Coastal Zone Management: Case Study of \textit{Posidonia oceanica} Meadows Along
+Integrated Coastal Zone Management: Case Study of \textit{{Posidonia oceanica}} Meadows Along
 the Corsican Coastline (Mediterranean Sea)}. Environ Manage (2006) 38:889–895 DOI
@@
 Semroud R., Shembri P., Shili A., Tomasello A., Velimirov B. 2005. \emph{Descriptors of
-\textit{Posidonia oceanica} meadows: Use and application}. Ecological Indicators 5, 213–230. \pageref{Pergent-Martini05}
+\textit{{Posidonia oceanica}} meadows: Use and application}. Ecological Indicators 5, 213–230. \pageref{Pergent-Martini05}
 \bibitem{Phinn08}

### 2011-01-28 MScDiplom_Lemenkova.tex
updated Valentijn's comments: 1,2,4,6,7,9,11,12 (8 from total 15), revised research objectives and aims, Latin words, structure of paragraphs

@@
 \textbf{Seagrass mapping and monitoring along the coasts of Crete, Greece}\\
+\vspace{1em}
  {by Lemenkova Polina}
@@
 	\item Name Examiner 2
-	\item External Examiner
+	\item External Examiner \hfill{Dr. \textit{Petter Pilesj\"o} (Lund University, GIS Centre)}
 \end{itemize}
@@
 
-\textbf{Member Supervisors}: 
+\textbf{Member Supervisors}: \\
 \begin{itemize}
-	\item[*] Venus V. (Department of Natural Resource, ITC)
-	\item[*] Toxopeus A.G. (Department of Natural Resource, ITC)
+	\item[* ] Mr. \textit{Valentijn Venus} (Department of Natural Resource, ITC)
+	\item[* ] Dr. \textit{Albertus G. Toxopeus} (Department of Natural Resource, ITC)
 \end{itemize}
@@
 \vspace{1mm}\\
-I am grateful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery.
+I am grateful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources, for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery.
 \vspace{1mm}\\
@@
 structures (rhizoms) buried in the sediments and vertical elongate leaves. A flowering plant,
-completely adapted to marine environment, they are however, more closely related to the lily family
-(Liliaceae) than to the true grass, despite their name “seagrass”, caused by the ribbon-like, grassy leaves
-\cite{McKenzie08}\label{McKenzie08}.\\
-Seagrasses create unique, complex, extremely diversified and productive ecological systems in the
+completely adapted to the marine environment, they are however, more closely related to the lily family
+(\textit{Liliaceae}) than to the true grass, despite their name \textit{seagrass}, caused by the ribbon-like, grassy leaves
+\cite{McKenzie08}\label{McKenzie08}. Seagrasses create unique, complex, extremely diversified and productive ecological systems in the
 littoral coastal zones between 0-50 meters in shallow waters all over the world \cite{Hogarth07}\label{Hogarth07}, and
@@
 provides sediments stability by reducing the degree of water movements, etc \cite{Francour99}\label{Francour99}.\\
-The purpose of current MSc research work aims to apply methods of remote sensing and GIS
-including processing and classification of satellite, aerial photos and videometric underwater imaging
-as well as GIS-based spatial analysis towards mapping and environmental monitoring of \emph{P. oceanica}
-along the coasts of Crete island, Greece. The technical implementation is based on ENVI, Erdas
-Imagine, WASI and ArcGIS software using aerial and satellite images as well as results of the
-underwater videometric measurements.
+The purpose of current MSc research work aims to apply methods of remote sensing analysis, including Radiative Transfer Models (RTM), 
+GIS-based spatial analysis, processing and classification of satellite and aerial photos, as well as videometric underwater footage, towards mapping and environmental monitoring of seagrass \textit{P.
+oceanica} along the selected locations of the northern coasts of Crete island, Greece.
+The technical implementation is based on WASI RTM software, GIS (ILWIS, Erdas Imagine and ArcGIS), using aerial and satellite images and the results of the underwater videometric measurements.
 
@@
 The distribution of the seagrasses is strongly influenced by several environmental factors, which
-include climate (mostly, tropical and temperate are--as), bathymetry (shallow shelf zones), hydrological
+include climate (mostly, tropical and temperate areas), bathymetry (shallow shelf zones), hydrological
 particularities (chemical content of water, nutrient availability and turbidity of waves), and geological 
 characteristics - sedimentation and cover types of the seafloor \cite{McKenzie06}\label{McKenzie06}.\\
-There are four European seagrass species in Mediterranean area \cite{Borum04}\label{Borum04}: Zostera marina,
-Zostera noltii, Cymodocea nodosa and \textit{P.oceanica}. In Greece the common species are \textit{P.oceanica}
-(L.) Delile, Cymodocea nodosa (Ucria) Ascherson, Zostera noltii Hornemann and Halophila
-stipulacea \cite{Amoutzopoulou-Schina05}\label{Amoutzopoulou-Schina05}. 
-These species differ in morphological and
-phenological features (Fig. \ref{fig:3}) as well as in structure and dynamics. Thus, Cymodocea nodosa is
-considered the pioneer species of \textit{P.oceanica} beds, the latter species forming the last stage. When P.
-oceanica beds regresses, C. nodosa often replaces them \cite{DenHartog77}\label{DenHartog77}; as a result, \textit{P.oceanica},
-C. nodosa, and Z. noltii do not form mixed persistent stands \cite{Buia91}\label{Buia91}.
+There are four European seagrass species in Mediterranean area \cite{Borum04}\label{Borum04}: \textit{Zostera marina,
+Zostera noltii, Cymodocea nodosa} and \textit{P.oceanica}. In Greece the common species are \textit{P.oceanica}
+\textit{(L.) Delile, Cymodocea nodosa (Ucria) Ascherson, Zostera noltii Hornemann} and \textit{Halophila
+stipulacea} \cite{Amoutzopoulou-Schina05}\label{Amoutzopoulou-Schina05}. 
+These species differ in morphological and phenological features (Fig. \ref{fig:3}) as well as in structure and dynamics. Thus, \textit{Cymodocea nodosa} is
+considered the pioneer species of \textit{P.oceanica} beds, the latter species forming the last stage.
+When \textit{P.oceanica} beds regresses, \textit{C.nodosa} often replaces them \cite{DenHartog77}\label{DenHartog77}; as a result, \textit{P.oceanica},
+\textit{C.nodosa}, and \textit{Z.noltii} do not form mixed persistent stands \cite{Buia91}\label{Buia91}.
 
 \subsubsection{Ecological significance of the seagrasses}
-Seagrass plays vital role in the marine ecosystems of the world ocean. Seagrasses are the only
-flowering plant in the world that is able to live completely submerged. Seagrass is a habitat for
+Seagrass plays vital role in the marine ecosystems of the world ocean. 
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-1-3.jpg}
+	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
+	\label{fig:3}
+\end{wrapfigure}
+Seagrasses are the only flowering plant in the world that is able to live completely submerged. Seagrass is a habitat for
 numerous marine fish species \cite{Nagelkerken00}\label{Nagelkerken00}, source of primary production and food for
@@
 constitutes the basis of the food web both within and outside the ecosystem \cite{Gobert06}\label{Gobert06}.
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-\centering
-	\includegraphics[scale=0.25]{Fig-1-3.jpg}
-	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
-	\label{fig:3}
-\end{wrapfigure}
-
 Finally, seagrasses are an important component in the environmental "food chain" of the coastal
@@
 degradation, but often seagrass undergo simultaneous affects from several of these factors together.
-This naturally increase environmental pressing and leads to drastic loss of very large areas of
+It naturally increases the environmental pressing  to drastic loss of very large areas of
 seagrass globally \cite{Orth06}\label{Orth06}.
@@
 eutrophication and invasive macroalgae. These processes have strong affect on the status of the
-seagrass meadows, and often lead to their complete disappearance \cite{Holmer09}\label{Holmer09}.
+seagrass meadows, and often lead to their complete disappearance \cite{Holmer09}\label{Holmer09}.\\
 Other environmental threat for the seagrasses arises as a result of the environmental struggle and
@@
 \cite{Montefalcone10}\label{Montefalcone10}.
-Seagrasses are vulnerable fragile species, important for the marine coastal ecosystems, especially for
-the protection of the beach structure. However, the facts about seagrass global degrading sound
-worrying: about 54 percent of the total seagrass meadows have lost any part of their area; the areas, where
-the seagrass ecosystems are degrading or lost, are not located in a specific area or continent, but
-registered globally; since 1980s global losses of the seagrasses on our planet is equal to two football
-fields per hour \cite{Mellors09b}\label{Mellors09b}.
-
-\subsubsection{General characteristics of \textit{P.oceanica}}
-The endemic Mediterranean seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}), is a main species in
-marine coastal environment of Greece, forming, despite its slow growth, the largest, most widespread,
-homogeneous and dense meadows in the Mediterranean between 5 and 40 m depth \cite{DenHartog70}\label{DenHartog70}.
 
@@
 
+Seagrasses are vulnerable fragile species, important for the marine coastal ecosystems, especially for
+the protection of the beach structure. However, the facts about seagrass global degrading sound
+worrying: about 54 \% of the total seagrass meadows have lost any part of their area; the areas, where
+the seagrass ecosystems are degrading or lost, are not located in a specific area or continent, but
+registered globally; since 1980s global losses of the seagrasses on our planet is equal to two football
+fields per hour \cite{Mellors09b}\label{Mellors09b}.
+
+\subsubsection{General characteristics of \textit{P.oceanica}}
+The endemic Mediterranean seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}), is a main species in
+marine coastal environment of Greece, forming, despite its slow growth, the largest, most widespread,
+homogeneous and dense meadows in the Mediterranean between 5 and 40 m depth \cite{DenHartog70}\label{DenHartog70}.
 The dominant and most productive coastal ecosystem of the Mediterranean, \textit{P.oceanica} is
@@
 	\includegraphics[scale=0.15]{Fig-1-6.jpg}
-	\caption{\textit{	extit{P.oceanica}}.\\
+	\caption{\textit{P.oceanica}.\\
 	Source \cite{Luque04}\label{Luque04}}
@@
 similar spectral reflectance, e.g. seagrass, dark-coloured bottom patches (mud), macroalgae.
-The in-situ fieldwork including underwater videographic measurements is an important part of the
+The \textit{in-situ} fieldwork including underwater videographic measurements is an important part of the
 seagrass monitoring, and has been successfully applied towards seagrass mapping \cite{Haag08}\label{Haag08}.
@@
 \subsection{Research objective}
-The current MSc research aims to explore the environmental conditions and dynamics of the spatial
-distribution of \textit{P.oceanica} seagrass meadows along the Cretan coast, based on the remote sensing and
-GIS techniques, knowledge about the coastal environment in Crete and integration of various data
+The current MSc research aims to explore the environmental conditions for the spatial distribution of \textit{P.oceanica} seagrass along the northern coast of Crete Island, based on the remote sensing and GIS techniques, knowledge about the coastal environment in Crete and integration of various data
 from the following sources:
-i) spectra of 	extit{P.oceanica}, carbonate sand, silt and other seafloor types
-ii) satellite imagery: Landsat, ASTER, MODIS, MERIS
-iii) aerial photos: Google Earth
-iv) in-situ fieldwork data of underwater videographic measurements
-v) vector GIS layers.
-The research aims to explore the dynamics of the environmental changes of the \textit{P.oceanica} meadows
-over time (2000-2010), as well as the extent of the spatial distribution over the whole island. The
-main research objective is monitoring the seagrass \textit{P.oceanica} along the coasts of Crete, in order to
-analyse the environmental changes of the 	extit{P.oceanica} meadows over the last 10 years (2000-2010)
-and the spatial extent of its current distribution.
-This research is supported by the in-situ measurements in two selected locations of the northern coast
+	\begin{itemize}
+		\item[] i) spectra of \textit{P.oceanica}, carbonate sand, silt and other seafloor types
+		\item[] ii) satellite imagery: Landsat TM, Landsat ETM+
+		\item[] iii) aerial photos: Google Earth
+		\item[] iv) \textit{in-situ} fieldwork data of underwater videographic measurements
+		\item[] v) vector GIS layers.
+	\end{itemize}
+The research aims to explore the limitations of the application of the Radiative Transfer Models (further RTMs) and remote sensing techniques 
+towards the study of the environmental properties of the \textit{P.oceanica} meadows distribution
+over the selected locations of northern Crete.
+The main research objective is monitoring the seagrass \textit{P.oceanica} in selected areas of Crete and analyzing the environmental conditions for the growth of \textit{P.oceanica}.
+\\
+This research is supported by the \textit{in-situ} measurements in two selected locations of the northern coast
 of Crete (Ligaria beach in Agia Pelagia district and Xerocampos), using following methods of the
-remote sensing techniques: underwater videometric measurements made by Olympus camera, aerial
-Google Earth and satellite images from different sources, spatial GIS and statistical analysis.
+remote sensing techniques: spectral modelling by means of the RTM Water Colour Simulator WASI,  underwater videometric measurements made by Olympus camera, aerial Google Earth and satellite images from different sources, spatial GIS and statistical analysis.
 
 \subsubsection{General objective}
-The main objective of this study is to investigate the distribution of seagrass along the northern coast
-of Crete using remote sensing techniques.
+The main objective of this study is to analyse the optical properties of the seagrass \textit{P.oceanica} and other seafloor types, and to apply the remote sensing techniques towards the investigation of the seagrass distribution in the selected locations along the northern coast of Crete.
 General objectives:
-1) Mapping the spatial distribution of the seagrass \textit{P.oceanica} along the northern coasts of
-Crete Island
-2) Monitoring the dynamics of the environmental changes of the \textit{P.oceanica} meadows over
-time (2000-2010) in Ligaria beach, Agia Pelagia district.
+	\begin{itemize}
+		\item[*]Analysing spectral reflectance of the \textit{P.oceanica} and other seafloor cover types means of radiative transfer model tools (RTMs), using WASI.
+		\item[*]Mapping spatial distribution of the seagrass \textit{P.oceanica} over selected locations along the northern coasts of Crete Island 
+	\end{itemize}
 
 \subsubsection{Specific objectives}
-1. To apply remote sensing data (Google Earth aerial images, Landsat, SPOT satellite images)
-for the monitoring of the seagrass meadows distribution
-2. To study narrow-band spectral reflectance properties of \textit{P.oceanica} and other seafloor
-cover types (sand and silt) using WASI water colour simulation software
-3. To use methods of the in situ diving observations and underwater videometric measurements
-by Olympus camera in order to receive large-scale imagery of the 	extit{P.oceanica} mattes
-4. To perform images classification (supervised and unsupervised) for the thematic mapping of
-the 	extit{P.oceanica} seagrass distribution along the coasts of northern Crete.
+	\begin{itemize}	
+		\item[]1.	To study narrow-band spectral reflectance properties of \textit{P.oceanica} and other seafloor cover types (sand and silt) using WASI water colour simulation software 
+		\item[]2.	To use methods of the in situ diving observations and underwater videometric measurements by Olympus camera in order to receive large-scale imagery of the P.oceanica mattes 
+		\item[]3.	To apply remote sensing data (Google Earth aerial images, Landsat TM and ETM+ satellite images) for the monitoring of the seagrass meadows distribution 
+		\item[]4.	To perform supervised images classification for the thematic mapping of the \textit{P.oceanica} seagrass distribution along the selected locations over the coasts of northern Crete. 
+	\end{itemize}
 
 \subsection{Research questions}
-1. Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ
-conditions ?
-2. Whether broadband and hyperspectral remote sensing data can be used for the mapping of
-\textit{P.oceanica} ?
+	\begin{itemize}	
+		\item Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying \textit{in-situ} conditions ?
+		\item Whether broadband and hyperspectral remote sensing data can be used for the mapping of \textit{P.oceanica} ?
+	\end{itemize}
 
 \subsection{Hypotheses testing}
-A statistical testing will be used to compare between the spectral responses of the different seagrass
-types, whether it is spectrally distinct and at least one pair is statistically different at every spectral
+A statistical testing will be used to compare between the spectral responses of the different seafloor cover types (i.e. sand and \textit{P. oceanica}),
+ whether it is spectrally distinct and at least one pair is statistically different at every spectral
 band.
 Hypothesis Ho: seagrass types are not spectrally distinct from other seafloor types with
-varying in-situ conditions, which means
+varying \textit{in-situ} conditions, which means
 Ho: μ1 =μ2 =μ3=...= μn.
 The alternative Hypothesis Ha claims the opposite statement: seagrass is spectrally
-distinct with varying in-situ conditions, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn.
+distinct with varying \textit{in-situ} conditions, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn.
 The distribution of the spectral responses at every spectral band is assumed to be normal as well as
@@
 techniques of the seagrass monitoring, i.e. remote sensing imagery classification of aerial and satellite
-images, GIS-based spatial analysis and ground in-situ surveys.
+images, GIS-based spatial analysis and ground \textit{in-situ} surveys.
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
+	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}}
+	\label{fig:1.10}
+\end{figure}
 The current study is based on the application of the remote sensing data, broadband satellite imagery,
@@
 substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. \\
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
-	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}}
-	\label{fig:1.10}
-\end{figure}
-
-Spectral reflectance, distinct for each seafloor cover type, enables their discrimination on the image. The pre-processing
+The spectral reflectance, distinct for each seafloor cover type, enables their discrimination on the image. The pre-processing
 of the images includes imagery corrections for atmospheric noises and effects of the water
@@
 
-\subsection{Research scheme}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.57]{UML_GIS.png}
-	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
-	\label{fig:1.11}
-\end{figure}
-
-\pagebreak
 %%%%%%%%%%%%%%%%%  Section2   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 terrestrial cartography, due to the technical difficulties of underwater observations, yet nowadays is
-has become a successful, rapidly developing and increasingly popular and challenging research branch.
+has become a successful, rapidly developing and increasingly popular and challenging research branch.\\
 Seagrasses can be mapped using various approaches, with the most popular method of the remote
 sensing application, enabling to detect and discriminate the substratum and various vegetation types
-growing on the seafloor.
-Several global seagrass surveys organise and provide regular monitoring of the seagrass distribution,
+growing on the seafloor.\\
+Several global seagrass surveys organize and provide regular monitoring of the seagrass distribution,
 health and species sustainability, amongst which the most world-known are, perhaps, the following
@@
 \href{http://www2.fiu.edu/~seagrass/}{Seagrass Ecosystems Research Laboratory in South Florida} and
-\href{http://www.flseagrass.org/}{Florida Seagrass organisation}, the \href{http://wsa.seagrassonline.org/}{World Seagrass Association} and  
-the \href{http://www.seagrasses.org/}{Seagrasses.org}.
+\href{http://www.flseagrass.org/}{Florida Seagrass organisation}, the \href{http://www.seagrasses.org/}{Seagrasses.org} and the \href{http://wsa.seagrassonline.org/}{World Seagrass Association}.
 These organizations aim at the global seagrass monitoring, providing with
@@
 centres and research institutes conducting seagrass monitoring and as a particular part of their
-research. Their reports and guidelines were used for references in the current research.
+research. Their reports and guidelines were used for references in the current research.\\
 Regular observations and monitoring of the seagrasses are known since 1960s, mainly in tropical
@@
 equipment and devices enabled to conduct underwater detailed measurements and observations
-largely contributed to the improvement of the traditional in-situ observations of seagrass. From the
+largely contributed to the improvement of the traditional \textit{in-situ} observations of seagrass. From the
 other side, development of the remote sensing methods and data acquisition from space contributed to
@@
 McKenzie \cite{McKenzieetal03}\label{McKenzieetal03} is widely used and well-known among the marine biologists and
-seagrass researchers. Applications of the in-situ seagrass observations of the structuring epiphyte
+seagrass researchers. Applications of the \textit{in-situ} seagrass observations of the structuring epiphyte
 community composition in the \textit{P.oceanica} ecosystems in Mediterranean Sea is, for example,
 described by Villegas \cite{Villegas06}\label{Villegas06}. Realization of traditional methods for mapping seagrass usually
-involves intensive and time-consuming in-situ observations during the fieldwork, as, for example,
-reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. Other methods of seagrass in-situ monitoring are based on
+involves intensive and time-consuming \textit{in-situ} observations during the fieldwork, as, for example,
+reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. \\
+Other methods of seagrass \textit{in-situ} monitoring are based on
 using the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then
@@
 initially acoustic methods were intended for the bathymetric surveying, i.e. they are not adjusted for
-the benthic habitat discrimination. The results of the in-situ measurements and observation are usually
+the benthic habitat discrimination. The results of the \textit{in-situ} measurements and observation are usually
 managed and treated using integrated GIS approach, as, for example, reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
@@
 remote sensing and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
-sedimentation processes, which in turn, may change micro relief properties and topology. 
+sedimentation processes, which in turn, may change micro relief properties and topology. \\
 The optical properties of the water are best reflected in the values of its radiance and irradiance,
@@
 radiant flux of the electromagnetic solar radiation energy, received per surface unit area in a given
-time (W·m−2·nm−1), while radiance (L) characterizes total emission or reflection that passes through
-or is emitted from a particular area (W·sr−1·m−2 ·nm−1). Therefore, the spectral reflectance, or the
+time ($W*m^{−2}*nm^{−1}$), while radiance (L) characterizes total emission or reflection that passes through
+or is emitted from a particular area ($W*sr^{−1}*m^{−2}*nm^{−1}$). Therefore, the spectral reflectance, or the
 reflectivity of the object, can be estimated by the direct mathematical division of these first two
 characteristics, and is expressed in percentage.
-Therefore, the irradiance and radiance of the water should be measured in order to estimate spectral
-reflectance of the various seafloor cover types. The optical measurements of the irradiance and
-radiance of the sea water and bottom cover types of the seafloor can be received by the means of the
+The irradiance and radiance of the water thus should be measured, in order to estimate spectral
+reflectance of the various seafloor cover types. \\
+The optical measurements of the irradiance and radiance of the sea water and bottom cover types of the seafloor can be received by the means of the
 optical sensors spectroradiometers. There are several companies producing radiometers with various
@@
 measurements  (Fig.\ref{fig:2.1}). These products have been used for the radiance and irradiance
-measurements in Agia Pelagia bay, Crete island, 2009.
+measurements in Agia Pelagia bay, Crete Island, 2009.
 
@@
 digitized aerial photographs, GPS data, remote sensing and SCUBA-based fieldwork measurements.
-SCUBA-based in-situ observations, though providing high resolution and accuracy results in seagrass
+SCUBA-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
 mapping, is limited in application, because of their time consumption, weather-dependency and
-unsuitability for the case of monitoring large areas of water for small-scale mapping. Methods of
-underwater videography with GPS is a tool of seagrass monitoring which has certain advantages, i.e.
+unsuitability for the case of monitoring large areas of water for small-scale mapping.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.30]{Fig-11.jpg}
+	\caption{In situ optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
+		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
+	\label{fig:13}
+\end{wrapfigure}
+
+The underwater videography with a GPS is a tool of seagrass monitoring which has certain advantages, i.e.
 high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
 collection in the field \cite{Schultz08}\label{Schultz08}. However, it cannot cover large areas for small-scale mapping.
-Remote sensing techniques offer clear advantages over other methods of in-situ field measurements
+Remote sensing techniques offer clear advantages over other methods of \textit{in-situ} field measurements
 and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
@@
 However, application of the remote sensing techniques for mapping of submerged vegetation,
-seafloor cover types and benthic vegetation, inter alia seagrasses, are still in their development.
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-11.jpg}
-	\caption{In situ optical reflectance spectra of seagrass. Shaded areas - percent of spectra lying within the range of reflectance. 
-		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
-	\label{fig:13}
-\end{wrapfigure}
-
-Approaches and methods for seagrass protection and monitoring still remain location specific or, at
-least, nation specific, depending to large extent on the tools available for researchers \cite{Mellors09a} \label{Mellors09a}.
+seafloor cover types and benthic vegetation, \textit{inter alia} seagrasses, are still in their development.
+Approaches and methods for the seagrass protection and monitoring still remain location-specific or, at
+least, nation-specific, depending to large extent on the tools available for the researchers \cite{Mellors09a}\label{Mellors09a}.
 Universal, international, standardized methods for seagrass directly for seagrasses
@@
 hyperspectral reflectance of sediments, which enable to estimate total microalgae biomass from
-measurements of hyperspectral reflectance. Spectral measurements of the target objects are made
+measurements of hyperspectral reflectance. \\
+Spectral measurements of the target objects are made
 by means of the radiometers, which receive and register the amounts of energy (radiance and
-irradiance) from the objects. \\ Measuring optical properties of water allow to calculate spectra of the
+irradiance) from the objects. Measuring optical properties of of the seawater allows to calculate spectra of the
 objects and to discriminate them on the aerial and satellite images. Thus, various scientists report
 success in spectral discrimination of submerged vegetation and other seafloor cover types on imagery
-using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01} \label{Louchard03} \label{Dehouck08} \label{Werdell03}.
+using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01}\label{Louchard03}\label{Dehouck08}\label{Werdell03}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-12.jpg}
+	\caption{Difference between broadband multispectral and hyperspectral resolution of
+		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
+	\label{fig:2.3}
+\end{wrapfigure}
+
 Studies of spectral reflectance of the different seagrass species comparing to the spectra of sand and
@@
 recent imagery allows to detect changes in seagrass distribution on over different years
-and to analyse dynamics of the seagrass distribution \cite{Ball09}\label{Ball09}. Another important advantage
-of the application of the remote sensing data for mapping of submerged aquatic vegetation has
+and to analyse dynamics of the seagrass distribution \cite{Ball09}\label{Ball09}. \\
+Another important advantage of the application of the remote sensing data for mapping of submerged aquatic vegetation has
 commercial nature: using remote sensing data and methods enables more low-cost and up-to-date
@@
 capturing is unavailable.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-12.jpg}
-	\caption{Difference between broadband multispectral and hyperspectral resolution of
-		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
-	\label{fig:2.3}
-\end{wrapfigure}
-
 Seagrass meadows may reach spatial scales from several up to hundreds of metres, therefore they are
@@
 The possibility of their application towards seagrass mapping varies and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:2.3}).
-In the next subparagraphs
-we briefly discuss the limitations and research experience of the using of various imagery
-for the seagrass mapping.
 
@@
 
-\subsubsection{Review of multispectral imagery used for seagrass mapping}
+In the next paragraphs (\nameref{sec:2.4.1} and \nameref{sec:2.4.2}) we will briefly discuss research experience of the application of satellite imagery towards seagrass mapping, its advantages and limitations, reported by various researchers.
+
+\subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}
 Seagrass mapping using remotely sensed data from multispectral sensors is based on the classification
@@
 landscapes.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.35]{Fig-13.jpg}
+ 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
+	\label{fig:2.4}
+\end{wrapfigure}
+
 The Landsat TM and Landsat ETM+ data, with the most recent from sensor Landsat 7 - an advanced
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.35]{Fig-13.jpg}
- 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
-	\label{fig:2.4}
-\end{wrapfigure}
-
 Another well-known multispectral sensor, SPOT provides multispectral imagery with a spatial
@@
 Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.30]{Fig-14.jpg}
+	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
+	\label{fig:2.5}
+\end{wrapfigure}
+
 The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
@@
 spectrum, the best results are achieved by the IKONOS and CASI, while Landsat TM has not high
-enough resolution. It may be caused to some extent by the loss of the radiance contrast, atmospheric
+enough resolution. 
+It may be caused to some extent by the loss of the radiance contrast, atmospheric
 Rayleigh scattering and defects of scattering. However, comparing between CASI and IKONOS, the
@@
 Another comparative analysis of the application of \href{http://www.itres.com/products/imagers/casi550}{CASI}, Landsat and \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird} imagery \cite{Phinn08}\label{Phinn08}
- demonstrates the high suitability of CASI images for the fine-scale mapping of seagrass
+ demonstrates high suitability of CASI images for the fine-scale mapping of the seagrass
 landscapes. Thus, CASI and Quickbird-2 images enable to identify even separate seagrass species
@@
 Landsat TM images with their 30*30m resolution.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-14.jpg}
-	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
-	\label{fig:2.5}
-\end{wrapfigure}
-
 Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}), launched in
@@
 scanner Daedalus AADS1268 is reported by Heege et al \cite{Heege03}\label{Heege03} where they aim at classification of
-macrophytes in shallow waters of the Lake Constance.
+macrophytes in shallow waters of the Lake Constance. \\
 In regard to the methods chosen for the image interpretation, the supervised classifications proves to
@@
 
-\subsubsection{Review of hyperspectral imagery used for seagrass mapping}
+\subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}
 The application of the hyperspectral sensors is most effective and provide more accurate classification
@@
 mapping at species level, as long as they are distinguishable spectrally (Fig.\ref{fig:2.6}) which, for example, has been
-tested in the case study of Australian marine ecosystems by Fyfe \cite{Fyfe04}\label{Fyfe04}. Application of various
-classifications methods, inter alia maximum likelihood, minimum distance and means, towards
+tested in the case study of Australian marine ecosystems by Fyfe \cite{Fyfe04}\label{Fyfe04}. 
+Application of various classifications methods, \textit{inter alia} maximum likelihood, minimum distance and means, towards
 hyperspectral imagery \cite{Peneva08}\label{Peneva08}, combined with fieldwork measurements ensures accurate
@@
 seafloor types classification \cite{Hochberg03a}\label{Hochberg03a} demonstrates that coral, seagrasses and
-sand very well distinguishable in their spectra with an overall classification accuracy of 98 percent.
+sand very well distinguishable in their spectra with an overall classification accuracy of 98 \%.
 However, the use of data from various sensors, both hyperspectral and multispectral, is possible and
@@
 necessary addition to the mapping helping to solve problems of interpretation during the images
-classification \cite{Pasqualini98a}\label{Pasqualini98a}. Therefore, this work includes sampling of the in-situ
+classification \cite{Pasqualini98a}\label{Pasqualini98a}. Therefore, this work includes sampling of the \textit{in-situ}
 measurements of the seagrass distribution.
@@
 The Ligaria Beach is a narrow, sandy and pebble beach (Fig.\ref{fig:3.2}), located ca 15 km north-west from Heraklion.
-The in-situ measurements were conducted during the fieldwork in the period 21.09.2010-11.10.2010.
+The \textit{in-situ} measurements were conducted during the fieldwork in the period 21.09.2010-11.10.2010.
 The fieldwork on seagrass monitoring included visual estimations and photo- and video footage of the
@@
 boat path. The general locations of the sampling sites and routes were selected on the basis of the visual examination
-of the seagrass beds during snorkeling and SCUBA diving, recommendations from the Greek collaborators of the Natural Nistory Museum of Crete, and available maps covering the research area.
+of the seagrass beds during snorkeling and SCUBA diving, recommendations from the Greek collaborators of the \textit{Natural History Museum of Crete}, and available maps covering the research area.
 To ensure even and objective selection of sampling sites
-we used transect sampling method, i.e. photographs were taken along the research path. Transect
-method is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
+we used transect sampling method, i.e. photographs were taken along the research path. \\
+Transect method is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
 enabling the occurrences of the seagrass meadows to be recorded and counted in a systematic and
@@
 
-\subsection[Google Earth aerial imagery]{Google Earth aerial imagery for seagrass mapping}
+\subsection[Google Earth aerial imagery]{Google Earth aerial imagery for the seagrass mapping}
 Apart from the satellite imagery, the aerial photographs from the Google Earth provide
@@
 detailed habitat and seafloor types discrimination, comparing with images
-acquired from the space-borne satellite platforms.
+acquired from the space-borne satellite platforms.\\
 The spatial coverage of the Google imagery is lesser comparing to spaceborne images, but this can be solved as
@@
 
-\subsection{Application of satellite imagery for seagrass mapping (Landsat, ASTER, MODIS)}
-
 %%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
-\pagebreak
 \section{Results}
@@
 IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
-research includes fieldwork in-situ measurements of the seagrass distribution along the northern and
+research includes fieldwork \textit{in-situ} measurements of the seagrass distribution along the northern and
 south-eastern coasts of Crete in chosen locations.\\
@@
 \ref{tab:A.2}).
+\\
 The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:A.18}) and 430-650 for the water radiance (Fig.\ref{fig:A.16}). 
-Afterwards, the measured values of the
-radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
+Afterwards, the measured values of the radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
 properties of the sea water and bottom cover types (Fig.\ref{fig:A.15}). The spectral range of radiance
@@
 wavelength ranges. 
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.30]{GNU-18.jpg}
-	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis.}
-	\label{fig:4.4}
-\end{figure}
-
 The values of their spectral reflectance are accepted as constant. The results of
@@
 results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
-in water column, water temperature, sun angle, concentration of gelbstoff (coloured dissolved organic
+in water column, water temperature, sun angle, concentration of Gelbstoff (coloured dissolved organic
 matter), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small
@@
 water, large and small particles. The values of all these parameters can be redacted and changed
-manually. However, the most important, major factors affecting the in-situ conditions are water depth (Fig.\ref{fig:4.9})
-and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand .
+manually. However, the most important, major factors affecting the \textit{in-situ} conditions are water depth (Fig.\ref{fig:4.9})
+and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand.
 
@@
 	\centering
-	\includegraphics[scale=0.30]{GNU-21.jpg}
-	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis}
-	\label{fig:27}
-\end{figure}
-
-Seagrass \textit{P.oceanica} can be mapped using remotely acquired spectral imagery, if it has distinctive
+	\includegraphics[scale=0.27]{GNU-18.jpg}
+	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
+\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis.}
+	\label{fig:4.4}
+\end{figure}
+
+Seagrass \textit{P.oceanica} can be mapped using remotely acquired spectral imagery, but only if it has distinctive
 reflectance signatures at different depths. Therefore, the depths of the shelf area are the first variable
@@
 	\centering
-	\includegraphics[scale=0.30]{GNU-20.jpg}
+	\includegraphics[scale=0.3]{GNU-20-new.jpg}
 	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
@@
 The 3.5 m depth limit was chosen based on the analysis of the separability of seagrass reflectance signatures, received by the means of
-previous in-situ measurements (year 2009) of the radiance and irradiance of water in Agia Pelagia
+previous \textit{in-situ} measurements (year 2009) of the radiance and irradiance of water in Agia Pelagia
 bay, which indicated that \textit{P.oceanica} seagrass could be only well separated within depths of 3.5 m.
 A statistical analysis of WASI-simulated spectral reflectance has been used in order to answer the first
-research question: whether the 	extit{P.oceanica} spectra is spectrally distinct at varying environmental in-situ
+research question: whether the 	extit{P.oceanica} spectra is spectrally distinct at varying environmental \textit{in-situ}
 conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
@@
 Under normal conditions by independent water colour sampling, values of the remote-sensing
-reflectance can vary by 12–24 per cent \cite{Toole00}\label{Toole00}, and these variations in the radiometric
+reflectance can vary by 12–24 \% \cite{Toole00}\label{Toole00}, and these variations in the radiometric
 determinations are mainly caused by the variety of the environmental factors.
@@
 depths, concentration of suspended particles in water column, water temperature, sun angle,
-concentration of gelbstoff (coloured dissolved organic matter), concentration of phytoplankton,
+concentration of Gelbstoff (coloured dissolved organic matter), concentration of phytoplankton,
 aerosol scattering, exponent of backscattering by small particles, cloudiness, viewing geometry and
@@
 relationship between their concentrations and the radiance of water that finally influence its spectral
-reflectance.
+reflectance.\\
 However, although there are a variety of environmental factors that contribute to the spectral
-reflectance, the most important ones from the mentioned above are water depth and seafloor fraction.
+reflectance, the most important ones from the mentioned above are the water column (depth) and seafloor fraction.
 It is because spectra of the seagrass \textit{P.oceanica} vary qualitatively over the depths interval of 0.5–4 m,
 and secondly, the content and cover fraction of the seafloor have the most distinctive effect on the
-spectral reflectance of the water.
-Therefore, in current MSc research we focus on these two major factors, and study the response of the
+spectral reflectance of the water. \\
+Therefore, in the current research we focus on these two major factors, and study the response of the
 water reflectance towards changing conditions of the water column depth and seafloor bottom cover
-fraction (seagrass 	extit{P.oceanica}, carbonate sand and silt).
-The analysis of the spectral signatures of the seagrass and sand clearly shows (Fig.\ref{fig:27}) that seagrass
+fraction (seagrass \textit{P.oceanica}, carbonate sand and silt).\\
+The analysis of the spectral signatures of the seagrass \textit{P.oceanica} and sand clearly shows (Fig.\ref{fig:27}) that seagrass
 has spectral reflectance much lesser than that of a carbonate sand, in general not increasing values of
-10 percent reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33 percent in its
+10 \% reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33\% in its
 highest values.
@@
 forward calculations, i.e. a computing and plotting of series of spectra according to specified
-parameter settings, with exactly defined depths and cover fraction.
+parameter settings, with exactly defined depths and cover fraction. \\
 The specific parameters have been chosen for the simulation of the environmental conditions where seagrass
 grow. The adaptations to life in salt sea water requires various physical and chemical parameters
-which include salinity, temperature of 17-25°C, light requirements with 10-20 percent on average, ranging
-from 4.4percent  minimal up to 29 percent \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as 35-
-45° and reflection factor 0.0201. \textit{P.oceanica}. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
+which include salinity, temperature of $17-25\,^{\circ}C$, light requirements with 10-20 \% on average, ranging
+from 4.4 \%  minimal up to 29 \% \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as $35-45\,^{\circ}$
+and reflection factor 0.0201. \textit{P.oceanica}. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
 The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:4.8}).
@@
 
-\paragraph{Spectral discrimination of sand}
-
 \subsubsection{Statistical analysis of the observational data and hypothesis testing}
 The total amount of measured data was large and included following datasets made using
-hyperspectral radiometer Ramses: 350 measurement sets of 	extit{P.oceanica} reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
-sets of 	extit{P.oceanica} reflectance for 15th Oct (Tab.\ref{tab:A.8}), 84 datasets for seawater reflectance with sediments, 105
+hyperspectral radiometer Ramses: 350 measurement sets of \textit{P.oceanica} reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
+sets of \textit{P.oceanica} reflectance for 15th Oct (Tab.\ref{tab:A.8}), 84 datasets for seawater reflectance with sediments, 105
 datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonate sand.
@@
 Thus, we analyzed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implemented in Gretl software (Tab.\ref{tab:19}). 
-Their probability expresses uncertainty about the possible values of the spectral reflectance of 	extit{P.oceanica} and sand, respectively.
+Their probability expresses uncertainty about the possible values of the spectral reflectance of \textit{P.oceanica} and sand, respectively.
 There is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
@@
 spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
-	extit{P.oceanica} spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
+\textit{P.oceanica} spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
 more than 0.20 nm, except for borders), we did not extend the y-scale up to 100 percent, in order to
 focus on the most necessary area of spectral values (Fig.\ref{fig:4.14}, Fig.\ref{fig:4.4} and \nameref{app}). The mean values are
-highlighted using bold red points, as we used these values for plotting the final graph of 	extit{P.oceanica} spectra in
+highlighted using bold red points, as we used these values for plotting the final graph of \textit{P.oceanica} spectra in
 Ligaria beach, Agia Pelagia.
@@
 		\includegraphics[scale=0.45]{GNU-13.jpg}
-		\caption{Multiplot of spectral reflectance of 	extit{P.oceanica}. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
+		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
 			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
@@
 Assuming that the null hypothesis, Hypothesis Ho, is true, which is “the spectral distinguishability of
-the seagrass 	extit{P.oceanica} from other seafloor types is not changing with varying in-situ conditions,
+the seagrass \textit{P.oceanica} from other seafloor types is not changing with varying \textit{in-situ} conditions,
 Ho: μ1 =μ2 =μ3=...= μn.”, what is the probability of observing the values of spectral reflectance of
-	extit{P.oceanica} at various depths, for the test statistic, that is at least as extreme as the values of spectral
-reflectance of 	extit{P.oceanica} at these depths, that was actually observed? The alternative Hypothesis Ha
-Fig.15. Multiplot display of spectral reflectance of 	extit{P.oceanica}. Measurement sets 1-100, 15th
+\textit{P.oceanica} at various depths, for the test statistic, that is at least as extreme as the values of spectral
+reflectance of \textit{P.oceanica} at these depths, that was actually observed? The alternative Hypothesis Ha
+Fig.15. Multiplot display of spectral reflectance of \textit{P.oceanica}. Measurement sets 1-100, 15th
 October. Bold red dots show the mean values within each dataset; vertical areas: quartiles Q1-Q2.
-claims the opposite statement: “the spectral discernibility of the seagrass 	extit{P.oceanica} is distinctly
-changing with varying in-situ conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
+claims the opposite statement: “the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly
+changing with varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
 Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.
-To analyze average values of spectral reflectance of 	extit{P.oceanica} by means of the measured sets we calculated 
+To analyze average values of spectral reflectance of \textit{P.oceanica} by means of the measured sets we calculated 
 mean values for the total set of measurements 1 – 350 (15th October): mean of min and max values, 
@@
 The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (\nameref{app}). 
-As a result, we have mean values of spectral reflectance of 	extit{P.oceanica} for extreme, average, median and both quartiles,
+As a result, we have mean values of spectral reflectance of \textit{P.oceanica} for extreme, average, median and both quartiles,
 independent of the individual sets 1-350.
@@
 We applied ANOVA (ANalysis Of Variance) testing for the analysis of the probability of the
-spectral reflectance of 	extit{P.oceanica} to be more or less spectrally disticnt from other seafloor cover
+spectral reflectance of \textit{P.oceanica} to be more or less spectrally distinct from other seafloor cover
 types with changing depth. W are checking the statement Hypothesis Ho, is true, which is “the
-spectral distinguishability of the seagrass 	extit{P.oceanica} from other seafloor types is not changing with
-varying in-situ conditions, Ho: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
+spectral distinguishability of the seagrass \textit{P.oceanica} from other seafloor types is not changing with
+varying \textit{in-situ} conditions, Ho: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
 P(reject H0|H0 is valid)= P(X$>$c|p=)=.05
@@
 (less than 1 in a 10 chance). The one-way ANOVA highlighted a significant difference between data
-of spectral reflectances of 	extit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
-Hypothesis Ha=“the spectral discernibility of the seagrass 	extit{P.oceanica} is distinctly changing with
-varying in-situ conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
+of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
+Hypothesis Ha=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with
+varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
 
@@
 of a difference between sample means, is used for the data analysis. In our case the Student t-test demonstrates, if the variation
-between two analyzed groups – values of spectral reflectance of seagrass 	extit{P.oceanica} and sand - is
+between two analyzed groups – values of spectral reflectance of seagrass \textit{P.oceanica} and sand - is
 significant. Therefore, we use Student t-test to compare two sets of quantitative data of spectral
-reflectance of 	extit{P.oceanica} and sand, respectively, with samples collected independently of one
+reflectance of \textit{P.oceanica} and sand, respectively, with samples collected independently of one
 another. The Student-t test can be performed knowing just the means, standard deviation, and number
-of data points. Therefore, we used the data (\nameref{app}) of means of spectral reflectance of both
-bottom cover types within data sets, their standard deviation, and the number of data points.
+of data points. Therefore, we used the data (\nameref{app}) of means of spectral reflectance of the both
+ cover types within data sets, their standard deviation, and the number of data points.
 The results show the following outcome.
 As a result of the statistical testing, we came to the following conclusion.
-The Hypothesis Ha is true, which claims that the spectral discernibility of the seagrass 	extit{P.oceanica} is
-distinctly changing with varying in-situ conditions, i.e. increasing depth, Ho: μ1≠μ2≠μ3≠...≠ μn .
+The Hypothesis Ha is true, which claims that the spectral discernibility of the seagrass 	\textit{P.oceanica} is
+distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ1≠μ2≠μ3≠...≠ μn .
 Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
-(Box Plots, Fig..).
+(Box Plots).
 The graphs were plotted on the basis of the following statistical data calculated from the sampling
@@
 The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
-investigate the distribution of the seagrass 	extit{P.oceanica} within the research area.
+investigate the distribution of the seagrass \textit{P.oceanica} within the research area.
 In the image processing part of the research supervised classification has been applied to the aerial
@@
 detected using different bands combinations, masked and studied for the estimation of the changes in
-the areas. The classification is based on the spectral properties of the \textit{	extit{P.oceanica}}, such as brightness,
+the areas. The classification is based on the spectral properties of the \textit{P.oceanica}, such as brightness,
 colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping
@@
 for each photograph by classification a series of polygons characteristic of each of the sea floor
-bottom types: sandy surface, seagrass bed for each species including \textit{	extit{P.oceanica}} (meadows), patchy
+bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
 seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).
@@
 	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
-	\caption{Seafloor types of the Arina beach: \textit{	extit{P.oceanica},} sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
+	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
 	\label{fig:38}
@@
 	\subfloat [Meadows of the seagrass. Source: \href{http://www.google.com/earth/index.html}{Google Earth}]{\label{fig:41a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
-	\subfloat [Patches (mattes) of the seagrass. Source: in-situ videometric measurements]{\label{fig:41b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
+	\subfloat [Patches (mattes) of the seagrass. Source: \textit{in-situ} videometric measurements]{\label{fig:41b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
 	\caption{Variations in spatial structure of the seagrass landscapes}
@@
 can be caused by the invasion of alien species, such as Cymodocea nodosa, Caulerpa prolifera,
-Caulerpa taxifolia. Invaders are in general strong colonizers comparing to native 	extit{P.oceanica}: they
+Caulerpa taxifolia. Invaders are in general strong colonizers comparing to native \textit{P.oceanica}: they
 occupy much greater habitat space within the regressed meadows of stressed native seagrass
 \cite{Montefalcone10}\label{Montefalcone10}. However, on the northern coasts of Crete the only dominating seagrass
-species is 	extit{P.oceanica}.
+species is \textit{P.oceanica}.
 Morphological differences in scale of seagrass landscape formations, discussed above, cause need for
@@
 the coastal zones can be applied to other marine coastal areas with growing seagrass.\\
-More than 50 percent of the world population lives within one km of the coast which results in continued
+More than 50 \% of the world population lives within one km of the coast which results in continued
 anthropogenic pressures in coastal regions. \\Therefore, management of coastal resources and shelf
@@
 \markright{Keywords}
-Seagrass, \textit{\textit{	extit{P.oceanica}}}, analysis of spectral reflectance, remote sensing, monitoring, Google Earth, Crete
+Seagrass, \textit{P.oceanica}, analysis of spectral reflectance, remote sensing, monitoring, Google Earth, Crete
 \pagebreak
@@
 	     PositionLongitude & 25.0098= 25°0'35.2794" & 25.012661=25°0'45.5796" \\ \hline
-	     Comment & 	extit{P.oceanica} & 	extit{P.oceanica} \\ \hline
+	     Comment & \textit{P.oceanica} & \textit{P.oceanica} \\ \hline
 	     CommentSub1 & 3.5m depth & 1.5m depth \\ \hline
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.30]{GNU-21.jpg}
+	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete.
+\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis}
+	\label{fig:27}
+\end{figure}
 \begin{figure}[h]
@@
 		\includegraphics[scale=0.20]{App-1.jpg}
-		\caption{Values of spectral reflectances of seagrass 	extit{P.oceanica} (fragment)}
+		\caption{Values of spectral reflectances of seagrass \textit{P.oceanica} (fragment)}
 		\label{fig:A.11}
@@
 		\includegraphics[scale=0.20]{GNU-401-420.jpg}
-		\caption{Spectral reflectance of 	extit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
+		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
 		\label{fig:A.12}
@@
 		\includegraphics[scale=0.20]{GNU-401-420-candles.jpg}
-		\caption{Statistical analysis of the measurement data: spectral reflectance of the 	extit{P.oceanica}. 
+		\caption{Statistical analysis of the measurement data: spectral reflectance of the \textit{P.oceanica}. 
 		Visualization of the interquartile ranging.
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral 			reflectance of \textbf{	extit{P.oceanica}}.
+	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.
 \href{http://www.spss.com/}{SPSS}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of 		\textbf{	extit{P.oceanica}}.
+	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.
 \href{http://www.spss.com/}{SPSS}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textbf{	extit{P.oceanica}}. \href{http://www.spss.com/}{SPSS}}
+	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textit{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\includegraphics[scale=0.20]{GNU-Stat-M-401-420.jpg}
-	\caption{Fragment of the statistical analysis of the 	extit{P.oceanica} reflectance. Example of data set 401-420. 
+	\caption{Fragment of the statistical analysis of the \textit{P.oceanica} reflectance. Example of data set 401-420. 
 		Visualization of the measured data together with statistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{	extit{P.oceanica}} at various depths: 0.5, 1.5 and 2.5 meters.
+	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
 \href{http://www.spss.com/}{SPSS}}\label{tab:8}
@@
 	\end{center} 
-	P more than .05, which means that there is a significant difference in radiance of 	extit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).
+	P more than .05, which means that there is a significant difference in radiance of \textit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).
 
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{	extit{P.oceanica}} at various depths: 0.5, 1.5 and 2.5 meters.
+	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
 \href{http://www.spss.com/}{SPSS}}\label{tab:A.7}
@@
 	\end{center}
-	P$>$0.05, which means that there is a significant difference in radiance of 	extit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).
+	P$>$0.05, which means that there is a significant difference in radiance of \textit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).
 
@@
 \begin{longtable}{|c|c|c|c|c|c|c|}
-	\caption{Results of the statistical analysis of spectral reflectance of \textbf{	extit{P.oceanica}}, sets 1-350).
-Wavelength step: 3 nm.
-Measured on Agia Pelagia beach, 15th October}\label{tab:A.8} \\
+	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October}\label{tab:A.8} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 
-	\multicolumn{7}{c}{Table A.8: Spectral reflectance of \textbf{	extit{P.oceanica}} \textit{(continued from previous page)}}\\
+	\multicolumn{7}{c}{Table A.8: Spectral reflectance of \textit{P.oceanica} \textit{(continued from previous page)}}\\
 	\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\includegraphics[scale=0.45]{Autocorr.jpg}
-	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
+	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
 	\label{fig:A.21}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{	extit{P.oceanica}} (15.X.), Series:V3.
+	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3.
 \href{http://www.spss.com/}{SPSS}}\label{tab:14}
@@
 	\includegraphics[scale=0.45]{Partial_Corr.jpg}
-	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{	extit{P.oceanica}} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
+	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
 	\label{fig:A.22}
@@
 	\begin{center}
-		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textit{	extit{P.oceanica}} (15.X.).
+		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textit{P.oceanica} (15.X.).
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Summary of the table with values of the spectral reflectance of \textbf{	extit{P.oceanica}} measured on three different depths, R}
+	\caption{Summary of the table with values of the spectral reflectance of \textit{P.oceanica} measured on three different depths, R}
 	\begin{center}
@@
 		\includegraphics[scale=0.25]{GNU-11.jpg}
-		\caption{Spectral reflectance of 	extit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
+		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
 		\label{fig:A.25}
@@
 		\includegraphics[scale=0.25]{GNU-12.jpg}
-		\caption{Remote sensing reflectance of 	extit{P.oceanica}.
+		\caption{Remote sensing reflectance of \textit{P.oceanica}.
 Series 1-25.
@@
 		\includegraphics[scale=0.4]{GNU-14.jpg}
-		\caption{Multiplot of spectral reflectance of 	extit{P.oceanica}.
+		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
 Series 126-150.
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
-\href{http://www.spss.com/}{SPSS}}
+	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
+	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).
 \href{http://www.spss.com/}{SPSS}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
-\href{http://www.spss.com/}{SPSS}}
+	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
 	\begin{center}
@@
 \begin{table}[H]
-	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
+	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica} (15.X.).
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 		\rowcolors{1}{AntiqueWhite}{LightGoldenrodYellow}
-		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textbf{	extit{P.oceanica}}.
+		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textit{P.oceanica}.
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 		\includegraphics[scale=0.38]{GNU-15.jpg}
-		\caption{Multiplot of spectral reflectance of 	extit{P.oceanica}.
+		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
 Series 151-200.
@@
 	\begin{center}
-	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textbf{	extit{P.oceanica}}.
+	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textit{P.oceanica}.
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 \begin{table}[H]
-	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
-	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textbf{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
 	\label{tab:30}
@@
 \end{figure}
+\pagebreak
+
+\subsection{\textit{Appendix-10}. Research general workflow}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.57]{UML_GIS.png}
+	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
+	\label{fig:1.11}
+\end{figure}
 
@@
 Bharathi A.M, Kumar S.C, Kalliappan S. 2003. \emph{Target Identification in multispectral images
-using in-situ hyperspectral reflectance}. The Geospatial Resource Portal. Remote Sensing.
+using \textit{in-situ} hyperspectral reflectance}. The Geospatial Resource Portal. Remote Sensing.
 www.gisdevelopment.net \pageref{Bharathi03}

### 2011-01-30 MScDiplom_Lemenkova.tex
added Section 'Radiative Transfer Models (RTM)' - comment 5 of Valentijn

@@
 Without these vital data the implementation of this project would have hardly been possible.
-I am also thankful to him for the technical help in Google Earth imagery capturing, using Gdal technologies and assistance during the fieldwork (videometric underwater footage, consultancies in seagrass observation and monitoring, driving boat, etc).
+I am also thankful to him for the technical help in Google Earth imagery capturing, using GDAL technologies and assistance during the fieldwork (videometric underwater footage, consultancies in seagrass observation and monitoring, driving boat, etc).
 \\ My special thank to his personal, friendly support in various situations on Crete.\vspace{1mm}\\
@@
 \section*{}
+\setcounter{lofdepth}{2}
 \markright{List of Figures}
@@
 \listoffigures
-\setcounter{lofdepth}{2}
 \pagebreak
@@
 
-\subsection{\textit{In-situ} observations of the seagrass meadows}
-The traditional methods of in-situ seagrass monitoring include in general the following standard
-scheme \cite{McKenzieetal03} \label{McKenzieetal03}. The seagrass is being sampled on the selected sites using transect lines,
-quadrant frame, single point markers, markers, GPS and other equipment. The seagrass sampling is
-taken on the regular way with observation points covering the study area with normal distribution.
-During the measurement process, the vertical photograph of the measurements frame is taken, and the
-following points are traditionally estimated: percentage of the seagrass cover within the quadrate,
-species composition, sediment composition, canopy height, epyphyte abundance, algae percent cover,
-count of microfauna and a specimen of seagrass is being taken. This scheme, well described by
-McKenzie \cite{McKenzieetal03}\label{McKenzieetal03} is widely used and well-known among the marine biologists and
-seagrass researchers. Applications of the \textit{in-situ} seagrass observations of the structuring epiphyte
-community composition in the \textit{P.oceanica} ecosystems in Mediterranean Sea is, for example,
-described by Villegas \cite{Villegas06}\label{Villegas06}. Realization of traditional methods for mapping seagrass usually
-involves intensive and time-consuming \textit{in-situ} observations during the fieldwork, as, for example,
-reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. \\
-Other methods of seagrass \textit{in-situ} monitoring are based on
-using the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then
-collect the return echoes for the analysis, yet this requires specialized equipment and is mostly used in
-the depth waters, combined with bathymetric measurements. Another limitation of the acoustic
-techniques is that they are not effective at identifying the composition or even the presence of
-submerged aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}, because
-initially acoustic methods were intended for the bathymetric surveying, i.e. they are not adjusted for
-the benthic habitat discrimination. The results of the \textit{in-situ} measurements and observation are usually
-managed and treated using integrated GIS approach, as, for example, reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
-
 \subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}
@@
 
-\subsection[Application of remote sensing data]{Application of remote sensing data towards seagrass mapping}
+\subsection[Radiative Transfer Models (RTM)]{Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing tools}
+
+Understanding radiance distribution within a water column is necessary for the studies of the underwater visibility, because scattering properties of the water body naturally vary with changing depth, wavelength and environmental conditions.
+Simulation of the radiance quantities for natural water bodies enables to analyse seafloor color remote sensing properties.
+Therefore, the artificial modelling of the seawater optical properties by means of the Radiative Transfer Modelling (further RTM) is used when water optics is studied under changing environment (e.g., depths, sun angle, suspended particles in water column).
+\\ In such cases a retrieval of water optical parameters from the remote measurements should be tested and analyzed.
+\\
+\pagebreak
+\newpage
+We briefly discuss below the most effective RTM software and algorithms from a range of various best-known up-to-date tools. 
+	\begin{itemize}
+		\item [*] The \textit{Second Simulation of a Satellite Signal in the Solar Spectrum}, version 1 \href{http://6s.ltdri.org/}{(6SV1)} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \href{http://modis.gsfc.nasa.gov/}{NASA MODIS} satellite imagery lookup tables.
+The 6S algorithm has been implemented by the \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{GRASS} software for the atmospheric correction
+		\item [*] Another example of the RTM adjusted for a specific imagery is a German \href{http://www.imk-asf.kit.edu/english/312.php}{\textit{KOPRA RTM}}, fitted for \href{http://envisat.esa.int/instruments/meris/}{MIPAS/ENVISAT} imagery.
+		\item [*]  The \href{http://www.sequoiasci.com/products/Hydrolight.aspx}{\textit{HydroLight RTM}} commercial software is an advanced model for oceanic radiative transfer calculations, developed by the Sequoia Scientific, Inc., USA.
+It is designed to solve a wide range of problems in optical oceanography and limnology.
+		\item [*] The British \href{http://research.metoffice.gov.uk/research/interproj/nwpsaf/rtm/}{\textit{RTTOV-9 RTM}} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers.
+It includes a number of useful tolls, e.g.
+can compute sea-surface emissivity for each channel, enables, to specify cloudiness for radiance calculations, etc.
+This program runs under Unix/Linux platforms and is open source.
+		\item [*] The \href{http://rtweb.aer.com/}{\textit{Rapid Radiative Transfer Model (RRTM)}} is one more RTM freeware, designed in the USA, Massachusetts. 
+		\item [*] The \href{http://www.modtran.org/}{\textit{MODTRAN}}, a commercial software from Spectral Sciences Inc., is another example of the RTM.
+Among its newest updates is generation of atmospheric correction data; it also has an option to write spectral output in binary, and convert it to ASCII.
+		\item [*] A \textit{Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium} \href{ftp://climate1.gsfc.nasa.gov/wiscombe/Multiple_Scatt/DISORT1.2/}{(DISORT)}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
+		\item [*] The \textit{Community Radiative Transfer Model}, \href{http://www.jcsda.noaa.gov/projects_crtm.php}{CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA, NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
+		\item [*] An \textit{Automatized Atmospheric Absorption Atlas Operational Release} \href{http://www.noveltis.fr/4AOP/}{(4AOP)} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{CNES} (the French Space Agency).
+Being a fine program which enables accurate computations of the radiance, defining other emissivity functions and other set up, it is however, particularly designed for the infrared waves (thus, not quite suitable in our case).
+To the advantages of this freeware we would count its user friendly interface and open source accessibility for the academic purposes.
+	\end{itemize}
+For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface enabling us to adjust various environmental parameters.
+
+\subsection[\textit{In-situ} observations of the seagrass]{The \textit{in-situ} observations of the seagrass}
+The traditional methods of \textit{in-situ} seagrass monitoring include in general the following standard
+scheme \cite{McKenzieetal03} \label{McKenzieetal03}. The seagrass is being sampled on the selected sites using transect lines,
+quadrant frame, single point markers, markers, GPS and other equipment. The seagrass sampling is
+taken on the regular way with observation points covering the study area with normal distribution.
+During the measurement process, the vertical photograph of the measurements frame is taken, and the
+following points are traditionally estimated: percentage of the seagrass cover within the quadrate,
+species composition, sediment composition, canopy height, epyphyte abundance, algae percent cover,
+count of microfauna and a specimen of seagrass is being taken. This scheme, well described by
+McKenzie \cite{McKenzieetal03}\label{McKenzieetal03} is widely used and well-known among the marine biologists and
+seagrass researchers. Applications of the \textit{in-situ} seagrass observations of the structuring epiphyte
+community composition in the \textit{P.oceanica} ecosystems in Mediterranean Sea is, for example,
+described by Villegas \cite{Villegas06}\label{Villegas06}. Realization of traditional methods for mapping seagrass usually
+involves intensive and time-consuming \textit{in-situ} observations during the fieldwork, as, for example,
+reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. \\
+Other methods of seagrass \textit{in-situ} monitoring are based on
+using the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then
+collect the return echoes for the analysis, yet this requires specialized equipment and is mostly used in
+the depth waters, combined with bathymetric measurements. Another limitation of the acoustic
+techniques is that they are not effective at identifying the composition or even the presence of
+submerged aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}, because
+initially acoustic methods were intended for the bathymetric surveying, i.e. they are not adjusted for
+the benthic habitat discrimination. The results of the \textit{in-situ} measurements and observation are usually
+managed and treated using integrated GIS approach, as, for example, reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
+
+\subsection[Application of the remote sensing data]{Application of the remote sensing data towards seagrass mapping}
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
@@
 unsuitability for the case of monitoring large areas of water for small-scale mapping. 
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-11.jpg}
-	\caption{In situ optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
-		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
-	\label{fig:13}
-\end{wrapfigure}
-
 The underwater videography with a GPS is a tool of seagrass monitoring which has certain advantages, i.e.
@@
 areas. Integrated together with GIS vector layers and maps, remote sensing data enable historical
-mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.\\
+mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.30]{Fig-11.jpg}
+	\caption{\textit{In situ} optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
+		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
+	\label{fig:13}
+\end{wrapfigure}
+
 However, application of the remote sensing techniques for mapping of submerged vegetation,
@@
 hyperspectral reflectance of sediments, which enable to estimate total microalgae biomass from
-measurements of hyperspectral reflectance. \\
+measurements of hyperspectral reflectance.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-12.jpg}
+	\caption{Difference between broadband multispectral and hyperspectral resolution of
+		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
+	\label{fig:2.3}
+\end{wrapfigure}
+
 Spectral measurements of the target objects are made
@@
 using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01}\label{Louchard03}\label{Dehouck08}\label{Werdell03}.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-12.jpg}
-	\caption{Difference between broadband multispectral and hyperspectral resolution of
-		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
-	\label{fig:2.3}
-\end{wrapfigure}
-
 Studies of spectral reflectance of the different seagrass species comparing to the spectra of sand and
@@
 result of the spectral absorption in different bands, typical for each target object.
+
+\begin{table}[htbp]
+	\caption{Characteristics of selected ocean-colour sensors}
+	\begin{center}
+		\rowcolors{1}{OldLace}{FloralWhite}
+		\begin{tabular}{|p{2cm}|p{20mm}|p{15mm}|p{17mm}|p{15mm}|p{15mm}|p{15mm}|}
+			\hline\hline
+			\textbf{Satellite Sensor:} & {\textbf{Landsat}} & {\textbf{SPOT}} & {\textbf{ENVISAT}} & {\textbf{Terra}} & {\textbf{GeoEye}} & {\textbf{Nimbus 7}} \\ \hline\hline
+				\textbf{Origin} & USA & France & Europe (ESA) & USA (NASA) & USA & USA \\ \hline
+				\textbf{Instrument} & \textit{TM, ETM+} & \textit{Pan, XS} & \textit{MERIS} & \textit{MODIS} & \textit{SeaWiFS} & \textit{CZCS}\\ \hline
+			\textbf{Number of channels} & 8 & 6 & 15 & 36 & 8 & 5 \\ \hline
+			\textbf{Wavelength coverage} & 185-1700 & 500-1750 & 412.5- 900 & 469-1640 & 412-865 & 443-750 \\ \hline
+			\textbf{Ground resolution (nadir), m} & 15-pan,30(MS),60 (TIR) & 2.5, 5, 10,20 & 1.2 km/300 m & 1.0 km & 1.13 km & 825 m \\ \hline
+			\textbf{Band Width (nm)} & 0.45-10.4 & 10,20 & 2.5,7.5,9,10 14,20 & 0.18,0.3,0.5,10 - 50 & 20,40 & 20,100 \\ \hline
+			\textbf{Launched} & 1972, 1999 & 1986 & 2002 & 1999 & 1997 & 1978 \\ \hline
+			\textbf{Recurrent period, days} & 16 & 24 & 35 & 16 & 16 & 16 \\ \hline
+		\end{tabular}
+	\end{center}
+	\label{tab:1}
+\end{table}
+
 Spectral reflectance of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
@@
 The possibility of their application towards seagrass mapping varies and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:2.3}).
-
-\begin{table}[htbp]
-	\caption{\textbf{Characteristics of selected ocean-colour sensors}}
-	\begin{center}
-		\rowcolors{1}{OldLace}{FloralWhite}
-		\begin{tabular}{|p{2cm}|p{20mm}|p{15mm}|p{17mm}|p{15mm}|p{15mm}|p{15mm}|}
-			\hline\hline
-			\textbf{Satellite Sensor:} & {\textbf{Landsat}} & {\textbf{SPOT}} & {\textbf{ENVISAT}} & {\textbf{Terra}} & {\textbf{GeoEye}} & {\textbf{Nimbus 7}} \\ \hline\hline
-				\textbf{Origin} & USA & France & Europe (ESA) & USA (NASA) & USA & USA \\ \hline
-				\textbf{Instrument} & \textit{TM, ETM+} & \textit{Pan, XS} & \textit{MERIS} & \textit{MODIS} & \textit{SeaWiFS} & \textit{CZCS}\\ \hline
-			\textbf{Number of channels} & 8 & 6 & 15 & 36 & 8 & 5 \\ \hline
-			\textbf{Wavelength coverage} & 185-1700 & 500-1750 & 412.5- 900 & 469-1640 & 412-865 & 443-750 \\ \hline
-			\textbf{Ground resolution (nadir), m} & 15-pan,30(MS),60 (TIR) & 2.5, 5, 10,20 & 1.2 km/300 m & 1.0 km & 1.13 km & 825 m \\ \hline
-			\textbf{Band Width (nm)} & 0.45-10.4 & 10,20 & 2.5,7.5,9,10 14,20 & 0.18,0.3,0.5,10 - 50 & 20,40 & 20,100 \\ \hline
-			\textbf{Launched} & 1972, 1999 & 1986 & 2002 & 1999 & 1997 & 1978 \\ \hline
-			\textbf{Recurrent period, days} & 16 & 24 & 35 & 16 & 16 & 16 \\ \hline
-		\end{tabular}
-	\end{center}
-	\label{tab:1}
-\end{table}
-
 In the next paragraphs (\nameref{sec:2.4.1} and \nameref{sec:2.4.2}) we will briefly discuss research experience of the application of satellite imagery towards seagrass mapping, its advantages and limitations, reported by various researchers.
@@
 Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}. 
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-14.jpg}
-	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
-	\label{fig:2.5}
-\end{wrapfigure}
-
 The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
@@
 than IKONOS \cite{Mumby02}\label{Mumby02}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.30]{Fig-14.jpg}
+	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
+	\label{fig:2.5}
+\end{wrapfigure}
+
 Another comparative analysis of the application of \href{http://www.itres.com/products/imagers/casi550}{CASI}, Landsat and \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird} imagery \cite{Phinn08}\label{Phinn08}
@@
 well. Successful application of the ASTER imagery towards mapping of submerged vegetation
-reported, for example, by Hirose \cite{Hirose04} \label{Hirose04}. Other multispectral images have also been used for the
+reported, for example, by Hirose \cite{Hirose04}\label{Hirose04}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.15]{Fig-15.jpg}
+	\caption{Reflectance spectra of seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
+		reflectance. Source:\cite{Yang10}\label{Yang10}}
+	\label{fig:2.6}
+\end{wrapfigure}
+
+Other multispectral images have also been used for the
 seagrass interpretation. For example, the success of using imagery from the multispectral airborne
 scanner Daedalus AADS1268 is reported by Heege et al \cite{Heege03}\label{Heege03} where they aim at classification of
-macrophytes in shallow waters of the Lake Constance. \\
+macrophytes in shallow waters of the Lake Constance. 
 In regard to the methods chosen for the image interpretation, the supervised classifications proves to
-be the most worthy in a majority of case studies \cite{Palandro03,Peneva08}\label{Palandro03} \label{Peneva08}. While
+be the most worthy in a majority of case studies \cite{Palandro03,Peneva08}\label{Palandro03}\label{Peneva08}. While
 often methods of the unsupervised classification are used as a tool for classifying submerged object
@@
 pure reflectance from the benthos; as a result, it may cause significant errors \cite{Dierssen03}\label{Dierssen03}.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-15.jpg}
-	\caption{Reflectance spectra of seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
-		reflectance. Source:\cite{Yang10}\label{Yang10}}
-	\label{fig:2.6}
-\end{wrapfigure}
-
 Therefore, for the accurate assessments of various seafloor cover types, water
@@
 areas:
-1. Ligaria beach (Agia Pelagia district), 36°20′N 22°59′E
-2. Xerokampos, 35°12′N 26°18′E
-3. Agia Pelagia, 36°20′N 22°59′E
+\begin{itemize}
+	\item [1.] Ligaria beach (Agia Pelagia district), $36\,^{\circ}20\textquoteright N, 22\,^{\circ}59\textquoteright E$ 
+	\item [2.] Xerokampos, $35\,^{\circ}12\textquoteright N, 26\,^{\circ}18\textquoteright E$
+	\item [3.] Agia Pelagia, $36\,^{\circ}20\textquoteright N, 22\,^{\circ}59\textquoteright E$.
+\end{itemize}
 
@@
 Seagrass sampling has been performed at two research stations at Crete Island - Ligaria beach
-(36°20′N 22°59′E) and Xerocampos (35°12′N 26°18′E), at depths lesser than 4 meters.
+($36\,^{\circ}20\textquoteright N, 22\,^{\circ}59\textquoteright E$) and Xerocampos ($35\,^{\circ}12\textquoteright N, 26\,^{\circ}18\textquoteright E$), at depths lesser than 4 meters.
 
@@
 geographic locations recorded using GPS.
-\pagebreak
 
@@
 along the boat route for GIS project.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.70]{Fig-18.jpg}
-	\caption{Boat used for the fieldwork measurements on Crete Island. 2010}
-	\label{fig:3.4}
-\end{figure}
-
-\newpage
 
@@
 To ensure even and objective selection of sampling sites
-we used transect sampling method, i.e. photographs were taken along the research path. \\
+we used transect sampling method, i.e. photographs were taken along the research path.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.70]{Fig-18.jpg}
+	\caption{Boat used for the fieldwork measurements on Crete Island. 2010}
+	\label{fig:3.4}
+\end{figure}
+
 Transect method is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
@@
 without location bias. 
-
-\begin{figure}[H]
-	\centering
-	\subfloat [view from below]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
-	\subfloat [view \textit{en-face}]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
-	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
-	\caption{Scheme of placements of the Olimpus cameras during the measurements}
-	\label{fig:3.5}
-\end{figure}
-
 The videographic measurements and photos, enabling to detect types of the
@@
  mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:3.5}). The data
-were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state. The
-locations of the route were randomly selected in the areas of the Ligaria beach, to ensure most dense
+were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state.
+
+\begin{figure}[H]
+	\centering
+	\subfloat [view from below]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
+	\subfloat [view \textit{en-face}]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
+	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
+	\caption{Scheme of placements of the Olimpus cameras during the measurements}
+	\label{fig:3.5}
+\end{figure}
+The locations of the route were randomly selected in the areas of the Ligaria beach, to ensure most dense
 coverage of the seagrass meadows in the research area. The underwater measurements of the seagrass
@@
 	\rowcolors{1}{LightSteelBlue}{Honeydew}
-	\caption{Results of the statistical analysis of spectral reflectance of \textbf{P.oceanica}, with average values (for sets 1 - 350). Generalisation up to step 20 nm}\label{tab:4.1}
+	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, with average values (for sets 1 - 350). Generalisation up to step 20 nm}\label{tab:4.1}
 	\begin{center}
@@
 \begin{table}[htbp]
-	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
+	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textit{P.oceanica} (15.X.).
 \href{http://gretl.sourceforge.net/}{Gretl}}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textbf{	extit{P.oceanica}} (15.X.).
+	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).
 \href{http://www.spss.com/}{SPSS}}

### 2011-01-30 MScDiplom_Lemenkova.tex
added 'List of Acronyms', 'Document Outline'

@@
 \usepackage{color}
+\usepackage{setspace}
 \usepackage{graphics}
@@
 \section*{Abstract}
-	\markright{Abstract}
+	\markright{Summary Abstract of the MSc Thesis}
 	\addcontentsline{toc}{section}{Abstract}
@@
 
+\section*{Document Outline}
+	\markright{Structure of the MSc Thesis}
+	\addcontentsline{toc}{section}{Document Outline}
+\textit{Chapter 1}: \nameref{sec:1} is an introductory section.
+It describes general background of the research problem, outlines the need and actuality of the problem (including the environmental vulnerability of the seagrasses), and highlights limitations and possibilities of the remote sensing application for the current work.
+In this chapter we also set up research objectives, put research questions for hypothesis testing and sketch research approach for the proposed work.\\
+In \textit{Chapter 2}: \nameref{sec:2} we review the existing literature and reported research experience on the similar problem: studies of spectral separability of various seafloor types, limitations and advantages of the remote sensing techniques applied for seagrass mapping seagrass ecology and mapping seagrass environment.
+We considered not only the Mediterranean environment, but also papers from Australian and Chinese scientists, because seagrass monitoring is most actively developed in the southern regions of our planet.
+The review of the existing RTM algorithms for the retrieval of the optical parameters, as well as description of various tools for the spectrometric measurements - are given in the same Chapter as well.
+More close attention has been given to the imaging hyperspectral radiometer Trios RAMSES, used for the data collection.\\
+\textit{Chapter 3}: \nameref{sec:3} deals with the methods and materials used in current work.
+We start our discussion from the fieldwork area location and describe tools and instruments used during the fieldwork, as well as sampling design.
+Procedures of the data pre-processing and capturing imagery from the Google Earth are also presented in the same chapter.
+\\
+\textit{Chapter 4}: \nameref{sec:4} presents to the reader the main results of the current MSc work, obtained during the processing fieldwork and other collected data.
+It starts from the review of the collected data from different sources, then describes the modelling component of the research, namely WASI colour simulation and plotted resulting graphs.
+We also discuss here the particularities of various sensors, suitable enough for deriving radiometric information to discriminate sand from the seagrass.
+Furthermore, this chapter includes analysis of the spectral signatures of various seafloor cover types in conditions of changing environment, possibility and limitations for \textit{P.oceanica} spectral discrimination, statistical analysis of the data sets and mapping based on images processing.\\
+\textit{Chapter 5}: \nameref{sec:5} summarizes and briefly discusses again the main results, reported in detail in a previous chapter.
+This chapter proposes some final discussions to the reader, and comes to the main research outcomes.\\
+Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \textit{Chapter 6}: \nameref{sec:6}.\\
+A lot of plotted graphs, statistical outcomes, tables, auxiliary (yet relevant to our research) pictures and images, illustrating our work - are collected in the \nameref{app}.
+Due to the standard editorial limitations of the current MSc work, it was not possible to include all them to the main chapters.
+However, we tried to make the structure of the Appendices most clear and easy-readable as possible, by dividing the Appendices into 10 various sub-sections and referencing to them from the main text where necessary.
+\\
+The \nameref{bib} section includes literature and internet resources used for the current work.
+The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.
+
+\pagebreak
+
+\section*{List of Acronyms}
+	\markright{Acronyms}
+	\addcontentsline{toc}{section}{List of Acronyms}
+\onehalfspace
+
+\textbf{4AOP}: Automatized Atmospheric Absorption Atlas Operational Release \\
+\textbf{6SV1}:  Second Simulation of a Satellite Signal in the Solar Spectrum, Version1\\
+\textbf{ANOVA}: ANalysis of VAriance\\
+\textbf{ASTER}: Advanced Spaceborne Thermal Emission and Reflection Radiometer\\
+\textbf{BRDF}:  Bidirectional Reflectance Distribution Function\\
+\textbf{CASI}:		Compact Airborne Spectrographic Imager\\
+\textbf{CNES}: 		Centre National d'Etudes Spatiales (of France)\\
+\textbf{CRTM}:		Community Radiative Transfer Model\\
+\textbf{CZCS}:		 Coastal Zone Color Scanner\\
+\textbf{DISORT}:	 Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium\\
+\textbf{ENVISAT}:	ENVIronmental SATellite\\
+\textbf{ESA}:		European Space Agency\\
+\textbf{ERS}: 		European Remote Sensing satellite\\
+\textbf{GIS}:		Geographic Information System\\
+\textbf{GNU}:		GNU's Not Unix \\
+\textbf{GPS}:		 Global Positioning System\\
+\textbf{GRASS}:	 Geographic Resources Analysis Support System\\
+\textbf{GRETL}:		GNU Regression, Econometrics and Time-series Library\\
+\textbf{HyMap}:	Hyperspectral Mapper\\
+\textbf{HyperOCR}:	Hyperspectral Ocean Colour Radiometer\\
+\textbf{IKONOS}: 	derived its name from the Greek term \textit{eikon} for image\\
+\textbf{ILWIS}:		Integrated Land and Water Information System\\
+\textbf{iPAQ}:	internet Pocket PC by CompAQ\\
+\textbf{KOPRA}:	 Karlsruhe Optimized and Precise Radiative transfer Algorithm\\
+\textbf{LaTeX}:		Lamport Tau Epsilon Chi (Greek for \textit{techne}), modified by L.Lamport\\
+\textbf{LAI}:	 Leaf Area Index\\
+\textbf{Landsat TM}:	 Landsat Thematic Mapper \\
+\textbf{Landsat ETM+}:	Landsat Enhanced Thematic Mapper Plus\\
+\textbf{LyX}: 		LyriX-files\\
+\textbf{MERIS}:		MEdium Resolution Imaging Spectrometer\\
+\textbf{MIPAS}:		Michelson Interferometer for Passive Atmospheric Sounding\\
+\textbf{MODIS}:	Moderate Resolution Imaging Spectroradiometer\\
+\textbf{NASA}:		National Aeronautics and Space Administration\\
+\textbf{NIR}:		 Near-Infra-Red\\
+\textbf{NOAA}:		 National Oceanic and Atmospheric Administration\\
+\textbf{OBC}:		On Board Calibration\\
+\textbf{Olympus ST}: 	 Olympus Stylus Tough\\
+\textbf{Open EV}: 	Open Earth View\\
+\textbf{P.oceanica}:	Posidonia oceanica\\
+\textbf{RAMSES}:	RAdiance Measurement SEnSor\\
+\textbf{RRTM}:		Rapid Radiative Transfer Model\\
+\textbf{RTTOV}: 	Radiative Transfer for TIROS Operational Vertical Sounder\\
+\textbf{RTM}:		Radiative Transfer Model\\
+\textbf{SAM}:		Scanning Acoustic Microscope\\
+\textbf{SCUBA}:	Self Contained Underwater Breathing Apparatus\\
+\textbf{SeaWiFS}: 	Sea-viewing Wide Field-of-view Sensor\\
+\textbf{SPOT}:		Satellite Pour l'Observation de la Terre\\
+\textbf{SPSS	}:	Statistical Package for the Social Sciences\\
+\textbf{STDEV}:		Standard Deviation\\
+\textbf{SWIR}: 		Short Wave InfraRed\\
+\textbf{TIR}: 		Thermal InfraRed\\
+\textbf{TIROS}: 	Television and InfraRed Observation Satellite\\
+\textbf{UML}:		Unified Modelling Language\\
+\textbf{UV}:		UltraViolet (10 nm to 400 nm)\\
+\textbf{UVA}:		UltraViolet A, or long waves (315 nm-400 nm)\\
+\textbf{UVB}:		 UltraViolet B, or medium waves (280 nm-315 nm)\\
+\textbf{VIS}:		VISible\\
+\textbf{VNIR}:		Visible and Near-InfraRed\\
+\textbf{WASI}:		WAter colour SImulator\\
+
+\pagebreak
+\singlespace
 \section*{List of Tools}\label{sec:0}
@@
 \rfoot{MSc Thesis\\ \emph{"Seagrass mapping and monitoring along the coasts of Crete, Greece"}}
-\section{Introduction}
+\section{Introduction}\label{sec:1}
 %%%%%%%%%%%%%%%%%%  Section 1   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\end{itemize}
-For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface enabling us to adjust various environmental parameters.
+For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
 
@@
 
-\subsection[Application of the remote sensing data]{Application of the remote sensing data towards seagrass mapping}
+\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
@@
 %%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and methods}
+\section{Materials and methods}\label{sec:3}
 \counterwithin{figure}{section}
@@
 %%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
+\section{Results}\label{sec:4}
 \markright{IV. Results}
@@
 \pagebreak
-\section{Discussion and conclusions}
+\section{Discussion and conclusions}\label{sec:5}
 The suggested methodology of the analysis of seagrass adaptation towards different environments of
@@
 \pagebreak
+
+\section{Recommendations}\label{sec:6}
 
@@
 \begin{thebibliography}{999}
-\addcontentsline{toc}{chapter}{\textbf{Bibliography}}
+\addcontentsline{toc}{chapter}{\textbf{Bibliography}}\label{bib}
 \bibliographystyle{plainnat}

### 2011-01-31 MScDiplom_Lemenkova.tex
written Abstract of MSc Thesis

@@
 	\addcontentsline{toc}{section}{Abstract}
+The seagrasses, a unique group of aquatic plants, create complex, extremely diversified and productive ecological systems in the littoral coastal zones.
+The only flowering plant in the world that is able to live completely submerged, seagrasses play vital role in the marine ecosystems of the World Ocean.
+Seagrasses are the most important component in the environmental food chain of the coastal ecosystems, being a vital food source for various marine species (e.g.
+fish, dugongs, turtles, swans), and a producer of organic matter, which is the very basis of the food web.\\
+The \textit{P.oceanica} seagrass is an endemic for the Mediterranean region, and a main species in the marine coastal environment of Greece.
+Meadows of \textit{P.oceanica} are subjected to the human activities, because they occur in coastal areas, where they are affected both by anthropogenic and by climatic and environmental factors.
+Nowadays \textit{P.oceanica} is in the alarming state of regression, because of the deterioration of the environment in the Mediterranean Sea.
+Due to these reasons, \textit{P.oceanica} is a protected species since 1988 in some European countries (France).
+Monitoring \textit{P.oceanica} is therefore an important contribution to the saving and protecting the environment of Mediterranean region.\\
+The current MSc thesis focuses on the monitoring of seagrass \textit{P.oceanica} along the northern coasts of Crete Island, Greece, and investigates the application of the remote sensing techniques for the seagrass mapping.
+This research was articulated in two parts, where the first one involves an ecological approach to the seagrass distribution in various regions around the globe and the experience of seagrass monitoring nowadays.
+The second part of this work has technical character and investigates the application of the remote sensing techniques towards seagrass mapping.
+It, furthermore, focuses on the optical properties of the \textit{P.oceanica} and other seafloor cover types, and studies distinguishability of various seafloor cover types.
+Studies of the optical characteristics of separate seafloor cover types were made with purpose to clarify, whether their spectral properties change with varying environmental conditions.
+\\
+Special attention has been drawn on the role of environmental factors on the distribution of \textit{P.oceanica} along the coasts of Crete, and in particular, how the optical properties of the seafloor cover types, i.e.
+spectral reflectance, are being changed under varying external conditions, e.g.
+water column, amount of suspended particles and sediments in the seawater, and water temperature.
+For this purpose we studied differences in the spectral reflectance of \textit{P.oceanica} and other bottom cover types at three distinct depths.
+The diverse spectral values entail variations in optical properties of the seafloor cover types at changing environmental conditions.
+We applied WASI simulation techniques for the modelling of the optical parameters of various seafloor cover types by various spaceborne imaging spectrometers (MERIS, SeaWiFS, CZCS and MODIS), in order to understand their suitability and possible limitations for the seagrass mapping.\\
+Fieldwork research sites were presented by separate locations on the northern coast of Crete region (Ligaria, Agia Pelagia, Xerocampos).
+The additional measurements of the reflectance spectra of the seawater with and without sediments have been made in aquarium tank in 2009 by means of Trios-RAMSES spectroradiometer.
+Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and Landsat ETM+ scenes.
+
 \pagebreak
@@
 Alcoverro T., Romero J., Duarte C.M., Lopez N. 1997. \emph{Spatial and temporal variations
-in nutrient limitation of seagrass Posidonia oceanic growth in the NW Mediterranean}.
+in nutrient limitation of seagrass \textit{Posidonia oceanica} growth in the NW Mediterranean}.
 Marine ecology Progress Series.Vol.146: 155-161. \pageref{}
 \bibitem{Amoutzopoulou-Schina05}Amoutzopoulou-Schina H and S.Haritonidis. 2005. \emph{Distribution and phenology of the
-marine phanerogam \textit{	extit{Posidonia oceanica}} in the Pagassitikos Gulf, Greece}. Journal of
+marine phanerogam \textit{Posidonia oceanica} in the Pagassitikos Gulf, Greece}. Journal of
 Biological Research 4: 203 – 211, 2005 J. \pageref{Amoutzopoulou-Schina05}
@@
 Ardizzone G., Belluscio A. and Maiorano L. 2006. \emph{Long-term change in the structure of
-a \textit{	extit{Posidonia oceanica}} landscape and its reference for a monitoring plan}. Marine Ecology.
+a \textit{Posidonia oceanica} landscape and its reference for a monitoring plan}. Marine Ecology.
 ISSN 0173-9565. \pageref{Ardizzone06}
@@
 an introduction to monitoring and management}. The M and MS project. ISBN: 87-89143-21-3.
-Internet version: Available at http://www.seagrasses.org \pageref{Borum04}
+Internet version: Available at \href{http://www.seagrasses.org}{Seagrasses.org} \pageref{Borum04}
 \bibitem{Brown02}
@@
 using \textit{in-situ} hyperspectral reflectance}. The Geospatial Resource Portal. Remote Sensing.
-www.gisdevelopment.net \pageref{Bharathi03}
+\href{www.gisdevelopment.net}{gisdevelopment} \pageref{Bharathi03}
 \bibitem{Buia91}
@@
 Coverage on the Mississippi Barrier Islands, Northern Gulf of Mexico}.
-http://precedings.nature.com/documents/3669/version/1/files/npre20093669-1.pdf \pageref{Carter08}
+\href{http://precedings.nature.com/documents/3669/version/1/files/npre20093669-1.pdf}{precedings.nature.com} \pageref{Carter08}
 \bibitem{Calvo03}
@@
 \bibitem{Choo06}
-Choo, C.K. 2006. \emph{Save our Seahorses}. First Edition. www.sosmalaysia.org \pageref{Choo06}
+Choo, C.K. 2006. \emph{Save our Seahorses}. First Edition. \href{www.sosmalaysia.org}{sosmalaysia} \pageref{Choo06}
 \bibitem{Clark99}
@@
 \bibitem{Dehouck08}
-Dehouck A., Martiny N., Froidefond J.-M., Sénéchal N. 2008. \emph{In-Water Reflectance
+Dehouck A., Martiny N., Froidefond J.-M., S\'en\'echal N. 2008. \emph{In-Water Reflectance
 Spectra Measured On Board a Jet-Ski Across a Complex Nearshore Zone of Case-2 Waters
@@
 Argyrou M., Karakassis I, Duarte C.M. 2006. \emph{Benthic inputs as predictors of seagrass 1
-(\textit{Posidonia oceanica}) fish farm-induced decline}. http://arxiv.org/ftp/qbio/
-papers/0611/0611006.pdf \pageref{Diaz-Almela06}
+(\textit{Posidonia oceanica}) fish farm-induced decline}. \href{http://arxiv.org/ftp/qbio/
+papers/0611/0611006.pdf}{arxiv.org} \pageref{Diaz-Almela06}
 \bibitem{Dierssen03}
@@
 Edwards A.J. (ed.) 2000. \emph{Remote Sensing handbook for tropical coastal management}.
-http://www.unesco.org/csi/pub/source/rs.htm \pageref{Edwards00}
+\href{http://www.unesco.org/csi/pub/source/rs.htm}{UNESCO} \pageref{Edwards00}
 \bibitem{Ferguson97}
@@
 Monitoring Landscape Changes in Texas Seagrass Beds from Aerial Photography}. Coastal
-Education and Research Foundation doi: 10.2112/07-0882.1 \pageref{Fletcher09}
+Education and Research Foundation, doi: 10.2112/07-0882.1 \pageref{Fletcher09}
 \bibitem{Fornes06}
@@
 Francour P., Ganteaume A. and M. Poulain. 1999. \emph{Effects of boats anchoring in
-\textit{	extit{Posidonia oceanica}} seagrass beds in the Port-Cross National Park (north-western
+\textit{Posidonia oceanica} seagrass beds in the Port-Cross National Park (north-western
 Mediterranean Sea)}. Aquatic conservation: Marine and Freshwater Ecosystems: 9: 391-400. \pageref{Francour99}
@@
 the detection of light stress in Eelgrass Zostera capricorni. PhD Thesis}. The University of
-Wollongong. Research Online, http://ro.uow.edu.au/theses/472/ \pageref{Fyfe03}
+Wollongong. Research Online, \href{http://ro.uow.edu.au/theses/472/}{University of Wollongong Thesis Collection} \pageref{Fyfe03}
 \bibitem{Gege04}
@@
 Holmer M, Argyrou M, Dalsgaard T., Danovaro R, Diaz-Almela E, Duarte C,
-Frederiksen M, Grau A, Karakassis I, Marbà N, Mirto S, Pérez M, Pusceddu A, Tsapakis M.2008 \pageref{}
+Frederiksen M, Grau A, Karakassis I, Marbà N, Mirto S, Pérez M, Pusceddu A, Tsapakis M.2008
 \emph{Effects of fish farm waste on \textit{Posidonia oceanica} meadows: Synthesis and provision of
@@
 Lewis M. 2001. \emph{Discriminating Vegetation with Hyperspectral Imagery - What is
-Possible?} http://www.ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=00978199 Proceedings,
+Possible?} \href{http://www.ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=00978199}{ieeexplore.ieee.org} Proceedings,
 International Geoscience and Remote Sensing Symposium, 9-13 July 2001, Sydney,
@@
 sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}.
-http://www.igarss08.org/Abstracts/pdfs/3888.pdf \pageref{Matarrese06}
+\href{http://www.igarss08.org/Abstracts/pdfs/3888.pdf}{International Geoscience and Remote Sensing Symposium} \pageref{Matarrese06}
 \bibitem{McKenzie02}
@@
 Western Pacific}. Marine Plant Ecology Group, QDPI, Northern Fisheries Centre, Cairns,
-http://www.seagrasswatch.org \pageref{McKenzieetal03}
+\href{http://www.seagrasswatch.org}{seagrasswatch} \pageref{McKenzieetal03}
 \bibitem{McKenzie06}
@@
 Nursery for Important Coral Reef Fishes, Using a Visual Census Technique}. Estuarine,
-Coastal and Shelf Science 51, 31–44 doi:10.1006/ecss.2000.0617, online:
-http://www.idealibrary.com \pageref{Nagelkerken00}
+Coastal and Shelf Science 51, 31–44 doi:10.1006/ecss.2000.0617 \pageref{Nagelkerken00}
 \bibitem{Noralez10}
@@
 \bibitem{SPOT} 
-\emph{SPOT, Official Website} http://www.spotimage.com/web/en/172-spot-images.php \pageref{SPOT}
+\emph{SPOT, Official Website} \href{http://www.spotimage.com/web/en/172-spot-images.php}{SPOT} \pageref{SPOT}
 \bibitem{Orth06}
@@
 Island}, MS using hyperspectral imagery.
-ftp://ftp.ecn.purdue.edu/jshan/proceedings/asprs2005/Files/0134.pdf \pageref{Peneva07}
+\href{ftp://ftp.ecn.purdue.edu/jshan/proceedings/asprs2005/Files/0134.pdf}{ecn.purdue.edu} \pageref{Peneva07}
 \bibitem{Pergent90}
@@
 Rasib A.W. and Hashim M. 1997. \emph{Mapping Seagrass From Satellite Remote Sensing
-Data. Proceedings of Asian Conference on Remote Sensing}. http://eprints.utm.my/2161/ \pageref{Rasib97}
+Data. Proceedings of Asian Conference on Remote Sensing}. \href{http://eprints.utm.my/2161/}{eprints.utm.my} \pageref{Rasib97}
 \bibitem{Ribed02}
@@
 Villegas P. P. 2006. \emph{Magnitude of herbivory in \textit{Posidonia oceanica} (L.) Delile and
-factors responsible for spatial variation}. Available on web: http://www.tesisenxarxa.net/ \pageref{Villegas06}
+factors responsible for spatial variation}. Available on web: \href{http://www.tesisenxarxa.net/}{tesisenxarxa} \pageref{Villegas06}
 \bibitem{Wabnitz08}
-Wabnitz C.C., Andréfouët S., Torres-Pulliza D., Müller-Karger F. E. and Kramer P.A.
+Wabnitz C.C., Andréfouët S., Torres-Pulliza D., M\"uller-Karger F. E. and Kramer P.A.
 2008. \emph{Regional-scale seagrass habitat mapping in the Wider Caribbean region using

### 2011-01-31 MScDiplom_Lemenkova.tex
written Recommendations

@@
 This chapter proposes some final discussions to the reader, and comes to the main research outcomes.\\
-Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \textit{Chapter 6}: \nameref{sec:6}.\\
+Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \nameref{sec:6}.\\
 A lot of plotted graphs, statistical outcomes, tables, auxiliary (yet relevant to our research) pictures and images, illustrating our work - are collected in the \nameref{app}.
@@
 
-\section{Recommendations}\label{sec:6}
-
-\section{Keywords}
+\section*{Recommendations}\label{sec:6}
+\markright{Recommendations}
+\addcontentsline{toc}{section}{Recommendations}
+
+To make further studies of \textit{P.oceanica} more effective we would suggest the following recommendations to be considered by the future researchers:
+	\begin{description}
+		\item[First:] To extend the research area towards the eastern part of the Crete Island, in order to received more regular observations of the 	seagrass locations.
+		\item[Second:] To use different sources of imagery and thus, to increase the total collection of scenes covering the research area.
+		\item[Third:] To extend the temporal period of the imagery coverage, once the data are available.
+The current work only includes images covering short temporal period (ca 10-year); further estimation of the dynamics of \textit{P.oceanica} along the coasts of Crete would increase our understanding of the long-scale temporal variations of the seagrass distribution.
+		\item[Fourth:] To apply various classifications methods for the available imagery in order to compare the results received by means of various techniques
+		\item[Fifth:] To simulate various environmental conditions while modelling optical properties of different seafloor cover types.
+Not only the depths and 	the chemical content of the seawater should be considered, but also other factors determining the effect of the ecology and health of \textit{P.oceanica}.
+		\item[Sixth:] To consider seafloor geomorphology among other factors determining seagrass distribution.
+If possible, to find out bathymetric data for the research area, and to overlay them with existing images and maps, in order to analyse correlation between spatial distribution of seagrass \textit{P.oceanica} and underwater relief along Cretan coasts.
+		\item[Seventh:] Other RTM software may be tested and the modelling outcomes compared.
+		\item[Eighth:] Application of various open source GIS (ILWIS, GRASS) could be very useful for the validation of the cartographic results, assessment of accuracy and comparison of various classification methods.
+		\item[Ninth:] The analysis of the health indicators of the seagrass (such as number of leaves per shoot, biomass estimation within the single shot, etc) was not considered in the current work, as it would go beyond the scope of the MSc thesis.
+However, ecological investigations could be used for the assessment of the vulnerability of the seagrass meadows in various locations on Crete.
+		\item[Tenth:] A flexible combination of the multi-scale mapping and results of the fieldwork measurements with GPS-referenced underwater footage would enable more profound analysis of the coastal environment on Crete.
+	\end{description}
+
+\section*{Keywords}
 \markright{Keywords}
-Seagrass, \textit{P.oceanica}, analysis of spectral reflectance, remote sensing, monitoring, Google Earth, Crete
+\addcontentsline{toc}{section}{Keywords}
+\textit{P.oceanica}, seagrass monitoring, spectral reflectance, remote sensing, WASI, Google Earth, Crete
 \pagebreak

### 2011-01-31 MScDiplom_Lemenkova.tex
added 'Supervisory Committee', 'Course Title' as sections

@@
 \begin{document}
-
 \pagenumbering{roman}
 
+\begin{titlepage}
+	\centering
 	\title{Seagrass mapping and monitoring along the coasts of Crete, Greece}
@@
 	\maketitle	
-\pagebreak
-
-\textbf{Course Title:} \\
+\end{titlepage}
+\pagebreak
+
+\section*{}
+\addcontentsline{toc}{section}{Course Title}
+\textbf{Course Title} \\
 {Geo-Information Science and Earth Observation for Environmental Modelling and Management}\\
@@
 \textbf{Level} \hfill{Master of Science (MSc)}\\
+
+\vspace{1em}
+
 \textbf{Course Duration:} \hfill{September 2009 – March 2011}\\
+
+\vspace{1em}
+
 \textbf{Consortium partners:} \\
@@
 
+\section*{}
+\addcontentsline{toc}{section}{Supervisory Committee}
 \begin{center}
@@
 \begin{itemize}
-	\item Name Examiner 1 (Chair)
-	\item Name Examiner 2
-	\item External Examiner \hfill{Dr. \textit{Petter Pilesj\"o} (Lund University, GIS Centre)}
+	\item Name Examiner 1 (Chair)\dotfill
+	\item Name Examiner 2\dotfill
+	\item External Examiner \dotfill \\ \hfill{Dr. \textit{Petter Pilesj\"o} (Lund University, GIS Centre)}
 \end{itemize}
@@
 \begin{itemize}
-	\item[* ] Mr. \textit{Valentijn Venus} (Department of Natural Resource, ITC)
-	\item[* ] Dr. \textit{Albertus G. Toxopeus} (Department of Natural Resource, ITC)
+	\item[* ] Mr. \textit{Valentijn Venus} \dotfill \\ (Department of Natural Resource, ITC)
+	\item[* ] Dr. \textit{Albertus G. Toxopeus} \dotfill \\ (Department of Natural Resource, ITC)
 \end{itemize}
@@
 
-		Signed.................................(Lemenkova Polina) \\
+		Signed................................(Lemenkova Polina) \\
 

### 2011-02-01 MScDiplom_Lemenkova.tex
added Index

@@
 \usepackage{paralist}
+\usepackage{makeidx}
+\makeindex
 \usepackage{fancyhdr}
@@
 	\addcontentsline{toc}{section}{Table of Content}
+	\setlength{\textwidth}{6.5in}
+	\addtolength{\voffset}{+3pt}
+	\textheight = 620pt
 	\tableofcontents 
@@
 The \nameref{bib} section includes literature and internet resources used for the current work.
-The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.
+The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.\\
+An Index is placed in the end of the Document to help the reader find what he is looking for.
 
@@
 	\markright{List of Tools}
-	\addcontentsline{toc}{section}{List of Tools}
+	\addcontentsline{toc}{section}{List of Tools}\index{Software used}
 
@@
 (\textit{Liliaceae}) than to the true grass, despite their name \textit{seagrass}, caused by the ribbon-like, grassy leaves
-\cite{McKenzie08}\label{McKenzie08}. Seagrasses create unique, complex, extremely diversified and productive ecological systems in the
+\cite{McKenzie08}\label{McKenzie08}. Seagrasses\index{Seagrasses} create unique, complex, extremely diversified and productive ecological systems in the
 littoral coastal zones between 0-50 meters in shallow waters all over the world \cite{Hogarth07}\label{Hogarth07}, and
@@
 \subsubsection{Global distribution of the seagrasses}
-Globally, there are 58 recognized and described seagrass species (Fig. \ref{fig:2}), belonging to two orders
+Globally, there are 58 recognized and described seagrass species\index{Seagrasses!global species} (Fig. \ref{fig:2}), belonging to two orders
 (\textit{Hydrocharitales} and \textit{Najadales}), four families (\textit{Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
@@
 The distribution of the seagrasses is strongly influenced by several environmental factors, which
-include climate (mostly, tropical and temperate areas), bathymetry (shallow shelf zones), hydrological
+include climate (mostly, tropical and temperate areas\index{Seagrasses!distribution of}), bathymetry (shallow shelf zones), hydrological
 particularities (chemical content of water, nutrient availability and turbidity of waves), and geological 
 characteristics - sedimentation and cover types of the seafloor \cite{McKenzie06}\label{McKenzie06}.\\
-There are four European seagrass species in Mediterranean area \cite{Borum04}\label{Borum04}: \textit{Zostera marina,
+There are four European seagrass species \index{Seagrasses!Mediterranean species} in Mediterranean area \cite{Borum04}\label{Borum04}: \textit{Zostera marina,
 Zostera noltii, Cymodocea nodosa} and \textit{P.oceanica}. In Greece the common species are \textit{P.oceanica}
@@
 \subsubsection{Environmental vulnerability of the seagrasses}
-Meadows of \textit{P. oceanica} are subjected to the human activities, as they occur in coastal areas, where
+Meadows of \textit{P. oceanica} are subjected to the human activities\index{Seagrasses!ecological vulnerability}, as they occur in coastal areas, where
 they can be affected both directly \cite{Meinesz91}\label{Meinesz91} or indirectly, through the impact on the
@@
 affecting both growth and status of the seagrasses the environmental contaminants can be mentioned,
-e.g. thermal, sewage, dredging and chemical pollution as well as any other kind of maritime works,
+e.g. thermal, sewage, dredging and chemical pollution\index{Seagrasses!anthropogenic impacts on}  as well as any other kind of maritime works,
 e.g. trawling and anchoring of boats \cite{Ribed02}\label{Ribed02}. Other human activities that cause degrading of the
@@
 \cite{McKenzie06}\label{McKenzie06}.\\
-Besides anthropogenic factors, various biochemical, climatic and environmental processes can cause
+Besides anthropogenic factors, various biochemical, climatic and environmental processes\index{Seagrasses!natural impacts on} can cause
 negative influence on seagrass distribution. Seagrass is exposed to threats from the global climate and
@@
 seagrass globally \cite{Orth06}\label{Orth06}.
-In tropical areas, where most of seagrasses are located (Fig.\ref{Fig.4}), seagrasses are subject to catastrophic
+In tropical areas, where most of seagrasses are located (Fig.\ref{Fig.4}), seagrasses are subjectto catastrophic
 extinction and loss, due to the cyclones, typhoons, storms, regular floods and increased rainfalls. Recovery from such
@@
 
-\subsubsection{General characteristics of \textit{P.oceanica}}
+\subsubsection{General characteristics of \textit{P.oceanica}}\index{Seagrasses!\textit{P.oceanica}}
 The endemic Mediterranean seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}), is a main species in
@@
 The dominant and most productive coastal ecosystem of the Mediterranean, \textit{P.oceanica} is
-spatially restricted to the Mediterranean area (Fig.\ref{fig:5}), with its extension limited by the western part
+spatially restricted to the Mediterranean area\index{Seagrasses!\textit{P.oceanica}!distribution} (Fig.\ref{fig:5}), with its extension limited by the western part
 of the Mediterranean Sea where cold Atlantic waters enter Gibraltar and mix with warm
 Mediterranean waters, thus decreasing its temperature.
-Morphological \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots
+Morphologically\index{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots
 and short rhizomes (0.5-2.0 mm). The leaves are, perhaps, the most particular characteristics of
@@
 of the meadow health status. \\
-There are several environmental factors, determining the growth of the \textit{P.oceanica}. The adaptation
+There are several environmental factors, determining the growth of the \textit{P.oceanica}. The adaptation\index{Seagrasses!\textit{P.oceanica}!environmental adaptation}
 to dry-summer subtropical climate reduces its extension to Mediterranean area only (Fig.\ref{fig:5}).
@@
 
-The phenology of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in
+The phenology\index{Seagrasses!\textit{P.oceanica}!phenology} of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in
 shallow waters plants have shorter and falciform leaves, compared to ones in the deeper and central
@@
 
-\subsection{Research problem}
-Monitoring of the marine benthic ecosystems of seagrasses is essential for the environmental
+\subsection{Research problem}\index{Research}
+Monitoring\index{Research!problem} of the marine benthic ecosystems of seagrasses is essential for the environmental
 assessment of the coastal zones. It increases our knowledge of the seagrass ecology, highlights 
@@
 
-\subsection{Research objective}
+\subsection{Research objective}\index{Research!objective}
 The current MSc research aims to explore the environmental conditions for the spatial distribution of \textit{P.oceanica} seagrass along the northern coast of Crete Island, based on the remote sensing and GIS techniques, knowledge about the coastal environment in Crete and integration of various data
-from the following sources:
+from the following sources \index{Data}:
 	\begin{itemize}
@@
 
-\subsubsection{General objective}
+\subsubsection{General objective}\index{Research!objective!general}
 The main objective of this study is to analyse the optical properties of the seagrass \textit{P.oceanica} and other seafloor types, and to apply the remote sensing techniques towards the investigation of the seagrass distribution in the selected locations along the northern coast of Crete.
@@
 
-\subsubsection{Specific objectives}
+\subsubsection{Specific objectives}\index{Research!objective!specific}
 	\begin{itemize}	
@@
 
-\subsection{Research questions}
+\subsection{Research questions}\index{Research!questions}
 	\begin{itemize}	
@@
 
-\subsection{Hypotheses testing}
+\subsection{Hypotheses testing}\index{Research!questions!hypothesis}
 A statistical testing will be used to compare between the spectral responses of the different seafloor cover types (i.e. sand and \textit{P. oceanica}),
@@
 
-\subsection{Research approach}
+\subsection{Research approach}\index{Research!approach}
 Seagrass consistent monitoring and mapping is necessary and important for the sustainable coastal
@@
 mattes, is used for the improvement of the accuracy of large seagrass meadows and separate mattes
-within the meadows. The in situ underwater videometric measurements of the seafloor are collected
+within the meadows. The \textit{in situ} underwater videometric measurements of the seafloor are collected
 during the fieldwork in Crete, for the validation of the classification results and to determine the exact
@@
 
-\subsection{Seagrass global monitoring: history and perspectives}
+\subsection{Seagrass global monitoring: history and perspectives}\index{Seagrasses!global monitoring}
 Mapping and monitoring the seagrass is important for the environmental assessment of the marine
@@
 growing on the seafloor.\\
-Several global seagrass surveys organize and provide regular monitoring of the seagrass distribution,
+Several global seagrass surveys\index{Seagrasses!global surveys} organize and provide regular monitoring of the seagrass distribution,
 health and species sustainability, amongst which the most world-known are, perhaps, the following
@@
 
-The optical properties of the sea water vary with different environmental conditions and reflect
+The optical properties\index{Seawater optical properties} of the sea water vary with different environmental conditions and reflect
 current chemical content and physical specifics of the water, revealing the variability and distribution
@@
 properties incessantly. The effects of these processes influence application of the hyperspectral
-remote sensing and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
+remote sensing\index{Optical properties of the seafloor!hyperspectral measurements} and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
 sedimentation processes, which in turn, may change micro relief properties and topology. \\
@@
 characteristics, and is expressed in percentage.
-The irradiance and radiance of the water thus should be measured, in order to estimate spectral
+The irradiance\index{Seawater optical properties!irradiance} and radiance\index{Seawater optical properties!radiance} of the water thus should be measured, in order to estimate spectral
 reflectance of the various seafloor cover types. \\
 The optical measurements of the irradiance and radiance of the sea water and bottom cover types of the seafloor can be received by the means of the
-optical sensors spectroradiometers. There are several companies producing radiometers with various
+optical sensors spectroradiometers\index{Seawater optical properties!spectroradiometers}. There are several companies producing radiometers with various
 characteristics and adjusted for different purposes, e.g. portable and miniature spectrometers from
@@
 optical sensors, GER-Series Field portable spectroradiometers from
-\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. Trios-RAMSES hyperspectral radiometers
+\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. Trios-RAMSES\index{Seawater optical properties!spectroradiometers!Trios-RAMSES} hyperspectral radiometers
 (Trios-RAMSES Hyperspectral UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-
@@
 
-Understanding radiance distribution within a water column is necessary for the studies of the underwater visibility, because scattering properties of the water body naturally vary with changing depth, wavelength and environmental conditions.
+Understanding radiance distribution within a water colum\index{Seawater optical properties!RTM}n is necessary for the studies of the underwater visibility, because scattering properties of the water body naturally vary with changing depth, wavelength and environmental conditions.
 Simulation of the radiance quantities for natural water bodies enables to analyse seafloor color remote sensing properties.
@@
 	\begin{itemize}
-		\item [*] The \textit{Second Simulation of a Satellite Signal in the Solar Spectrum}, version 1 \href{http://6s.ltdri.org/}{(6SV1)} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \href{http://modis.gsfc.nasa.gov/}{NASA MODIS} satellite imagery lookup tables.
+		\item [*] The \textit{Second Simulation of a Satellite Signal in the Solar Spectrum}, version 1 \href{http://6s.ltdri.org/}{(6SV1)}\index{Seawater optical properties!RTM!6SV1} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \href{http://modis.gsfc.nasa.gov/}{NASA MODIS} satellite imagery lookup tables.
 The 6S algorithm has been implemented by the \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{GRASS} software for the atmospheric correction
-		\item [*] Another example of the RTM adjusted for a specific imagery is a German \href{http://www.imk-asf.kit.edu/english/312.php}{\textit{KOPRA RTM}}, fitted for \href{http://envisat.esa.int/instruments/meris/}{MIPAS/ENVISAT} imagery.
+		\item [*] Another example of the RTM adjusted for a specific imagery is a German \href{http://www.imk-asf.kit.edu/english/312.php}{\textit{KOPRA RTM}}\index{Seawater optical properties!RTM!KOPRA}, fitted for \href{http://envisat.esa.int/instruments/meris/}{MIPAS/ENVISAT} imagery.
 		\item [*]  The \href{http://www.sequoiasci.com/products/Hydrolight.aspx}{\textit{HydroLight RTM}} commercial software is an advanced model for oceanic radiative transfer calculations, developed by the Sequoia Scientific, Inc., USA.
 It is designed to solve a wide range of problems in optical oceanography and limnology.
-		\item [*] The British \href{http://research.metoffice.gov.uk/research/interproj/nwpsaf/rtm/}{\textit{RTTOV-9 RTM}} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers.
+		\item [*] The British \href{http://research.metoffice.gov.uk/research/interproj/nwpsaf/rtm/}{\textit{RTTOV-9 RTM}}\index{Seawater optical properties!RTM!RTTOV-9} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers.
 It includes a number of useful tolls, e.g.
@@
 This program runs under Unix/Linux platforms and is open source.
-		\item [*] The \href{http://rtweb.aer.com/}{\textit{Rapid Radiative Transfer Model (RRTM)}} is one more RTM freeware, designed in the USA, Massachusetts. 
-		\item [*] The \href{http://www.modtran.org/}{\textit{MODTRAN}}, a commercial software from Spectral Sciences Inc., is another example of the RTM.
+		\item [*] The \href{http://rtweb.aer.com/}{\textit{Rapid Radiative Transfer Model (RRTM)}}\index{Seawater optical properties!RTM!RRTM} is one more RTM freeware, designed in the USA, Massachusetts. 
+		\item [*] The \href{http://www.modtran.org/}{\textit{MODTRAN}}\index{Seawater optical properties!RTM!MODTRAN}, a commercial software from Spectral Sciences Inc., is another example of the RTM.
 Among its newest updates is generation of atmospheric correction data; it also has an option to write spectral output in binary, and convert it to ASCII.
-		\item [*] A \textit{Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium} \href{ftp://climate1.gsfc.nasa.gov/wiscombe/Multiple_Scatt/DISORT1.2/}{(DISORT)}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
-		\item [*] The \textit{Community Radiative Transfer Model}, \href{http://www.jcsda.noaa.gov/projects_crtm.php}{CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA, NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
-		\item [*] An \textit{Automatized Atmospheric Absorption Atlas Operational Release} \href{http://www.noveltis.fr/4AOP/}{(4AOP)} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{CNES} (the French Space Agency).
+		\item [*] A \textit{Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium} \href{ftp://climate1.gsfc.nasa.gov/wiscombe/Multiple_Scatt/DISORT1.2/}{(DISORT)}\index{Seawater optical properties!RTM!DISORT}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
+		\item [*] The \textit{Community Radiative Transfer Model}, \href{http://www.jcsda.noaa.gov/projects_crtm.php}{CRTM}\index{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA, NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
+		\item [*] An \textit{Automatized Atmospheric Absorption Atlas Operational Release} \href{http://www.noveltis.fr/4AOP/}{(4AOP)}\index{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{CNES} (the French Space Agency).
 Being a fine program which enables accurate computations of the radiance, defining other emissivity functions and other set up, it is however, particularly designed for the infrared waves (thus, not quite suitable in our case).
@@
 	\end{itemize}
-For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
+For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}}\index{Seawater optical properties!RTM!WASI} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
 
@@
 
-\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}
+\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}\index{Remote sensing}
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
@@
 unsuitability for the case of monitoring large areas of water for small-scale mapping. 
-The underwater videography with a GPS is a tool of seagrass monitoring which has certain advantages, i.e.
+The underwater videography\index{Remote sensing!underwater videography} with a GPS\index{Remote sensing!GPS} is a tool of seagrass monitoring which has certain advantages, i.e.
 high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
@@
 
-Spectral measurements of the target objects are made
+Spectral measurements\index{Remote sensing!spectral measurements} of the target objects are made
 by means of the radiometers, which receive and register the amounts of energy (radiance and
@@
 using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01}\label{Louchard03}\label{Dehouck08}\label{Werdell03}.
-Studies of spectral reflectance of the different seagrass species comparing to the spectra of sand and
+Studies of spectral reflectance\index{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and
 other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
@@
 
-Spectral reflectance of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
+Spectral reflectance\index{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
 depth where it is located, and is generally decreasing in values by increasing depths.
-The most important spectral diapason for marine mapping of submerged vegetation
+The most important spectral diapason\index{Seawater optical properties!spectral reflectance!spectral diapason} for marine mapping of submerged vegetation
 and, particularly, for seagrass, is 350-800 nm \cite{Barille09}\label{Barille09}.
@@
 
-\subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}
+\subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}\index{Remote sensing!multispectral imagery} 
 Seagrass mapping using remotely sensed data from multispectral sensors is based on the classification
@@
 
-The Landsat TM and Landsat ETM+ data, with the most recent from sensor Landsat 7 - an advanced
+The Landsat TM\index{Remote sensing!multispectral imagery!Landsat TM} and Landsat ETM+\index{Remote sensing!multispectral imagery!Landsat ETM+} data, with the most recent from sensor Landsat 7 - an advanced
 and multispectral scanning, launched in 1999, prove to be feasible and useful for the mapping of
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
-Another well-known multispectral sensor, SPOT provides multispectral imagery with a spatial
+Another well-known multispectral sensor, SPOT\index{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial
 resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
@@
 Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}. 
-The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
+The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}\index{Remote sensing!multispectral imagery!IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
 available high-resolution imagery at 1- and 4-meter resolution from \href{http://www.geoeye.com}{Geoeye}. IKONOS
@@
 the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
-made using IKONOS, Landsat TM, and CASI, show that in the blue part of the
+made using IKONOS, Landsat TM, and CASI\index{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
 spectrum, the best results are achieved by the IKONOS and CASI, while Landsat TM has not high
@@
  demonstrates high suitability of CASI images for the fine-scale mapping of the seagrass
-landscapes. Thus, CASI and Quickbird-2 images enable to identify even separate seagrass species
+landscapes. Thus, CASI and Quickbird-2\index{Remote sensing!multispectral imagery!Quickbird} images enable to identify even separate seagrass species
 with small width and heterogeneous nature of the seagrass patches, which could not be detected using
 Landsat TM images with their 30*30m resolution.
-Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}), launched in
-1999 on board Terra sensor, provides high-resolution images of the Earth in 15 different bands of the
+Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}\index{Remote sensing!multispectral imagery!ASTER}), launched in
+1999 on board Terra sensor\index{Remote sensing!multispectral imagery!Terra sensor}, provides high-resolution images of the Earth in 15 different bands of the
 electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:2.4}). It has diverse
@@
 
-\subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}
+\subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}\index{Remote sensing!hyperspectral imagery}
 The application of the hyperspectral sensors is most effective and provide more accurate classification
@@
 While comparing multispectral imagery with airborne hyperspectral, the last one showed higher
-overall accuracies \cite{Phinn08}\label{Phinn08}. Several, regular and narrow (10 nm) spectral bands, specific for
+overall accuracies \cite{Phinn08}\label{Phinn08}. Several, regular and narrow (10 nm) spectral bands\label{sec:2.4.2}\index{Remote sensing!hyperspectral imagery!spectral bands}, specific for
 hyperspectral imagery, are strong tool enabling to detect even slight and subtle differences in the
@@
 approaches requires application of high-spatial resolution (higher than 5 m) or hyperspectral imagery.
-Comparative analysis of the application of hyperspectral and multispectral imagery towards the
+Comparative analysis of the application of hyperspectral\label{sec:2.4.2}\index{Remote sensing!hyperspectral imagery!application} and multispectral imagery towards the
 seafloor types classification \cite{Hochberg03a}\label{Hochberg03a} demonstrates that coral, seagrasses and
@@
 
-\subsection{Fieldwork data collection}
+\subsection{Fieldwork data collection}\index{Data!collection}
 Seagrass sampling has been performed at two research stations at Crete Island - Ligaria beach
@@
 
-\subsection{Data pre-processing}
+\subsection{Data pre-processing}\index{Data!pre-processing}
 A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of \textit{P.oceanica} made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
@@
 
-\subsubsection{Water color spectral simulation using WASI}
+\subsubsection{Water color spectral simulation using WASI}\index{Data!processing}
 The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
@@
 based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
-types, where the spectral signatures were examined. WASI software (Fig.\ref{fig:4.7}) is used to simulate
+types, where the spectral signatures were examined. WASI software\index{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate
 spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
@@
 
-An analysis of the spectral reflectance of \textit{P.oceanica} is done using the WASI simulations in order to
+An analysis of the spectral reflectance of \textit{P.oceanica} is done using the WASI simulations\index{Seawater optical properties!RTM!WASI} in order to
 determine, which wavebands can be still used to identify \textit{P.oceanica}. The analysis of spectra shows
@@
 
-\subsubsection{Statistical analysis of the observational data and hypothesis testing}
+\subsubsection{Statistical analysis of the observational data and hypothesis testing}\index{Data!statistical analysis}
 The total amount of measured data was large and included following datasets made using
@@
 spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
-The statistical calculators were mainly made by means of Gretle and SPSS. 
+The statistical calculators were mainly made by means of Gretle\index{Data!statistical analysis!Gretl} and SPSS\index{Data!statistical analysis!SPSS}. 
 The Open Office was used for preliminary data view and preprocessing, and included following computations
@@
 seawater with and without sediment (e.g. Fig..\ref{fig:A.16}).
-In the statistical analysis of the raw observed data pool of the values of spectral reflectance (about 400 single measurements for \textit{P.oceanica}, e.g.
+In the statistical analysis of the raw observed data pool\index{Data!measurements pool} of the values of spectral reflectance (about 400 single measurements for \textit{P.oceanica}, e.g.
 Tab.\ref{tab:A.8}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
@@
 There is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
-Each time we observe the outcome of a random variable of spectral reflectance, we obtain an observation, which is hence no longer random. 
+Each time we observe the outcome of a random variable\index{Data!statistical analysis!variables} of spectral reflectance, we obtain an observation, which is hence no longer random. 
 We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the \nameref{app}: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyze the distribution of values of spectral reflectance of \textit{P.oceanica}.
-A probability distribution, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
-The \textit{least absolute deviations (LAD)}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
-Non-linear logistic analysis is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
+A probability distribution\index{Data!statistical analysis!probability distribution}, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
+The \textit{least absolute deviations (LAD)}\index{Data!statistical analysis!LAD}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
+Non-linear logistic analysis\index{Data!statistical analysis!non-linear logistic analysis} is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
 the value of spectral reflectance for each single observation) is not continuous, but discrete and independent for each case (Tab.\ref{tab:11}).
 \\
-As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{	extit{P.oceanica}} and Tab.\ref{tab:28} for carbonate sand.
-The \textit{autocorrelation}, a cross-correlation of an observed value with itself, made on two different days (14th and 15th October),
- shows the similarity between observations as a function of the time separation between them.
+As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand.
+The \textit{autocorrelation}\index{Data!statistical analysis!autocorrelation}, a cross-correlation\index{Data!statistical analysis!cross-correlation} of an observed value with itself, made on two different days (14th and 15th October), shows the similarity between observations as a function of the time separation between them.
 As can be seen on Fig.\ref{fig:A.21}, Fig.\ref{fig:A.22} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
@@
 \\
-The method of \textit{least squares} minimizes the sum of the squares of the errors made in solving every single observation made.
+The method of \textit{least squares}\index{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made.
 We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}.\\
@@
 The summary of the values of spectral reflectance properties on various depths is presented on the Tab.\ref{tab:A.7}.
-The \textit{k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
+The \textit{k-means cluster analysis}\index{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
 It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
-To estimate spectral density of a spectral reflectance measurements we used periodogram function (Tab.\ref{tab:30}), 
+To estimate spectral density of a spectral reflectance measurements we used periodogram\index{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30}), 
 which corresponds to the general spectrum of the observations with representation of a variable quantity.  
-Thus, the periodogram (Fig. \ref{fig:A.31}) of the series of observation highlights so-called “dead spots” of low power between the frequencies we wish to
+Thus, the periodogram (Fig.
+\ref{fig:A.31}) of the series of observation\index{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to
 exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:A.12} ) 
@@
 The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
-areas of quartiles (Q1, Q3, shaded areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution
+areas of quartiles\index{Data!statistical analysis!quartiles Q1-Q3} (Q1, Q3, shaded areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution
 (spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
@@
 
-A statistical hypothesis test has been applied for the making decision and controlling the wealth of the
+A statistical hypothesis\index{Data!statistical analysis!hypothesis testing} test has been applied for the making decision and controlling the wealth of the
 observational data of the hyperspectral measurements of the water reflectance.
@@
 
-\paragraph{ANOVA-testing}
+\paragraph{ANOVA-testing}\index{Data!statistical analysis!ANOVA}
 We applied ANOVA (ANalysis Of Variance) testing for the analysis of the probability of the
@@
 
-\paragraph{Student-t testing}
+\paragraph{Student-t testing}\index{Data!statistical analysis!Student-t testing}
 The Student-t test, one of the most commonly used techniques for testing a hypothesis on the basis
@@
 where the abbreviations stand for the following values:
-Mean: Average of the data to be plotted, AVERAGE {data}
-St Dev: Standard deviation, STDEV{data}
+Mean: Average of the data to be plotted, AVERAGE\index{Data!statistical analysis!average} {data}
+St Dev: Standard deviation\index{Data!statistical analysis!standard deviation}, STDEV{data}
 Median: Median of the data, MEDIAN{data}
@@
 25th Pct: Plotting value of first quartile = Q1
-50th Pct: Plotting value of median = Median-Q1
+50th Pct: Plotting value of median = Median\index{Data!statistical analysis!median}-Q1
 75th Pct: Plotting value of third quartile = Median-Q3
@@
 \subsubsection{Results of the remote sensing application}
-The in situ spectral reflectance data of \textit{	extit{P.oceanica}} and sand were used to model these seafloor cover
-types as the different sensors would percept them: MODIS, ASTER, MERIS, SeaWiFS and CZCS.\\
+The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
+types as the different sensors would percept them: MODIS\index{Remote sensing!multispectral imagery!MODIS}, ASTER\index{Remote sensing!multispectral imagery!ASTER}, MERIS\index{Remote sensing!multispectral imagery!MERIS}, SeaWiFS\index{Remote sensing!multispectral imagery!SeaWiFS} and CZCS\index{Remote sensing!multispectral imagery!CZCS}.\\
 These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
@@
 detailed examination of the imagery. \\
-Seagrass beds are clearly visible in color aerial Google Earth
+Seagrass beds are clearly visible in color aerial Google Earth\index{Remote sensing!aerial imagery!GoogleEarth}
 photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
@@
 
-\subsubsection{Data integration}
+\subsubsection{Data integration}\index{Data!integration}
 The integrated approach used in this research work has high potential as a means to monitor changes
@@
 high resolution aerial color Google Earth photography, spaceborne satellite imagery, assessment of
-spectral signatures using WASI software, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
+spectral signatures using WASI software\index{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
 based mapping. The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
@@
 
-\subsubsection{Upscale mapping of the seagrass landscapes}
+\subsubsection{Upscale mapping of the seagrass landscapes}\index{Seagrasses!spatial structure}
 Although ocean pelagic landscapes have a high degree of spatial variance and less structurally
@@
 \begin{thebibliography}{999}
-\addcontentsline{toc}{chapter}{\textbf{Bibliography}}\label{bib}
+\addcontentsline{toc}{section}{\textbf{Bibliography}}\label{bib}
 \bibliographystyle{plainnat}
@@
 
+\cleardoublepage
+\newpage
+\phantomsection
+\addcontentsline{toc}{section}{Index}
+\printindex
+
 \end{document}

### 2011-02-01 MScDiplom_Lemenkova.tex
changed Document general style (font size and layout), some cosmetic improvements in figures placement. Changed boxed links to colour ones

@@
-\documentclass[10pt, a4paper]{article}
+\documentclass[11pt, a4paper]{article}
 \usepackage{color}
 \usepackage{setspace}
+\usepackage[a4paper, left=30mm,right=25mm,top=30mm,bottom=35mm]{geometry}
 \usepackage{graphics}
@@
 \usepackage[a4paper=true, ]{hyperref}
+\hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue}
 
@@
 	\addcontentsline{toc}{section}{Table of Content}
-	\setlength{\textwidth}{6.5in}
-	\addtolength{\voffset}{+3pt}
-	\textheight = 620pt
+	%\setlength{\textwidth}{6in}
+%	\addtolength{\voffset}{+5pt}
+%	\marginparwidth = 30pt
+%	\textheight = 630pt
 	\tableofcontents 
@@
 	\addcontentsline{toc}{section}{Abstract}
+
 The seagrasses, a unique group of aquatic plants, create complex, extremely diversified and productive ecological systems in the littoral coastal zones.
@@
 The additional measurements of the reflectance spectra of the seawater with and without sediments have been made in aquarium tank in 2009 by means of Trios-RAMSES spectroradiometer.
-Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and Landsat ETM+ scenes.
+Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and ETM+ scenes.
 
@@
 It describes general background of the research problem, outlines the need and actuality of the problem (including the environmental vulnerability of the seagrasses), and highlights limitations and possibilities of the remote sensing application for the current work.
-In this chapter we also set up research objectives, put research questions for hypothesis testing and sketch research approach for the proposed work.\\
+In this chapter we also set up research objectives, put research questions for hypothesis testing and sketch research approach for the proposed work.
+\newline
 In \textit{Chapter 2}: \nameref{sec:2} we review the existing literature and reported research experience on the similar problem: studies of spectral separability of various seafloor types, limitations and advantages of the remote sensing techniques applied for seagrass mapping seagrass ecology and mapping seagrass environment.
@@
 	\addcontentsline{toc}{section}{List of Acronyms}
-\onehalfspace
-
+\doublespacing
 \textbf{4AOP}: Automatized Atmospheric Absorption Atlas Operational Release \\
@@
 \textbf{Open EV}: 	Open Earth View\\
-\textbf{P.oceanica}:	Posidonia oceanica\\
 \textbf{RAMSES}:	RAdiance Measurement SEnSor\\
@@
 \textbf{VNIR}:		Visible and Near-InfraRed\\
-\textbf{WASI}:		WAter colour SImulator\\
-
+\textbf{WASI}:		WAter colour SImulator
 \pagebreak
@@
 \addcontentsline{toc}{section}{List of Figures}
+\onehalfspacing
 \listoffigures
@@
 	\centering
-	\includegraphics[scale=0.25]{Fig-1-2.jpg}
+	\includegraphics[scale=0.30]{Fig-1-2.jpg}
 	\caption{Distribution of seagrasses in the world.\\ Source:\cite{Green03}\label{Green03} 
@@
 Seagrass plays vital role in the marine ecosystems of the world ocean. 
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-1-3.jpg}
-	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
-	\label{fig:3}
-\end{wrapfigure}
 Seagrasses are the only flowering plant in the world that is able to live completely submerged. Seagrass is a habitat for
@@
 Finally, seagrasses are an important component in the environmental "food chain" of the coastal
-ecosystems, being the food source for dugongs, turtles, swans and various fish \cite{Cappo95}\label{Cappo95}.\\
+ecosystems, being the food source for dugongs, turtles, swans and various fish \cite{Cappo95}\label{Cappo95}.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-1-3.jpg}
+	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
+	\label{fig:3}
+\end{wrapfigure}
+
 Due to their wide distribution, meadows size, easy collection and abundance, sensitivity to the
@@
 growth rate, the anthropogenic modifications of the coastal zone, happening more rapidly than the
-capacity of the plant to adapt to these changes, reduce its distribution area \cite{Micheli05}\label{Micheli05}. One
-of the main drivers of seagrass decline is, for instance, the location of the fish farming near the
+capacity of the plant to adapt to these changes, reduce its distribution area \cite{Micheli05}\label{Micheli05}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-1-4.jpg}
+	\caption{Distribution of seagrass in relation to mean ocean temperature. \\
+	Source: \cite{Orth06}\label{Orth06}}
+	\label{Fig.4}
+\end{wrapfigure}
+
+One of the main drivers of seagrass decline is, for instance, the location of the fish farming near the
 seagrass meadows. The negative effects of the sedimentation of waste particles in the farm vicinities
@@
 environmental conditions \cite{Leoni06,Leoni07}\label{Leoni06} \label{Leoni07}. 
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.20]{Fig-1-4.jpg}
-	\caption{Distribution of seagrass in relation to mean ocean temperature. \\
-	Source: \cite{Orth06}\label{Orth06}}
-	\label{Fig.4}
-\end{wrapfigure}
-
 The detailed research of the fish
@@
 fish farm organic and nutrient content in the sediments with dynamics of the key seagrass species (\textit{P.
-oceanica}) in the Mediterranean Sea. Nowadays P.oceanica is in the alarming state of regression due
+oceanica}) in the Mediterranean Sea. Nowadays \textit{P.oceanica} is in the alarming state of regression due
 to the deterioration of the environment in the Mediterranean Sea \cite{Ribed02}\label{Ribed02}. Due to these reasons,
@@
 \cite{Montefalcone10}\label{Montefalcone10}.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.20]{Fig-1-5.jpg}
-	\caption{Geographical distribution of
-	\textit{P.oceanica}. Source: \cite{Borum04}\label{Borum04}}
-	\label{fig:5}
-\end{wrapfigure}
-
 Seagrasses are vulnerable fragile species, important for the marine coastal ecosystems, especially for
@@
 fields per hour \cite{Mellors09b}\label{Mellors09b}.
+\pagebreak
 
@@
 homogeneous and dense meadows in the Mediterranean between 5 and 40 m depth \cite{DenHartog70}\label{DenHartog70}.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.20]{Fig-1-5.jpg}
+	\caption{Geographical distribution of
+	\textit{P.oceanica}. Source: \cite{Borum04}\label{Borum04}}
+	\label{fig:5}
+\end{wrapfigure}
+
 The dominant and most productive coastal ecosystem of the Mediterranean, \textit{P.oceanica} is
-spatially restricted to the Mediterranean area\index{Seagrasses!\textit{P.oceanica}!distribution} (Fig.\ref{fig:5}), with its extension limited by the western part
-of the Mediterranean Sea where cold Atlantic waters enter Gibraltar and mix with warm
+spatially restricted to the Mediterranean area\index{Seagrasses!\textit{P.oceanica}!distribution} (Fig.\ref{fig:5}), 
+with its extension limited by the western part of the Mediterranean Sea where cold Atlantic waters enter Gibraltar and mix with warm
 Mediterranean waters, thus decreasing its temperature.
-Morphologically\index{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots
-and short rhizomes (0.5-2.0 mm). The leaves are, perhaps, the most particular characteristics of
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-1-6.jpg}
-	\caption{\textit{P.oceanica}.\\
-	Source \cite{Luque04}\label{Luque04}}
-	\label{fig:6}
-\end{wrapfigure}
-
-\textit{P.oceanica}, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:6}): having
-usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}).
-Growing \textit{P.oceanica} make a meadows which, in turn, consist of smaller patches, called “matte”
-(Fig.\ref{fig:7}), a monumental construction made by the growth of rhizomes and leaves with entangled roots
+Morphologically\index{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots and short rhizomes (0.5-2.0 mm). 
+The leaves are, perhaps, the most particular characteristics of \textit{P.oceanica}, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:8}(a)): having usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}(c)).
+Growing \textit{P.oceanica} make a meadows which, in turn, consist of smaller patches, called \textit{matte}
+(Fig.\ref{fig:8}(b)), a monumental construction made by the growth of rhizomes and leaves with entangled roots
 and entrapped sediment \cite{Francour06}\label{Francour06}.
+
+\begin{figure}[H]
+	\centering
+	\subfloat [\textit{P.oceanica}. Source \cite{Luque04}\label{Luque04}]{\includegraphics[width=0.25\textwidth]{Fig-1-6.jpg}}
+	\subfloat [Scheme of matte structure of \textit{P.oceanica}. Source\cite{Pergent90}\label{Pergent90}]{\includegraphics[width=0.3\textwidth]{Fig-1-7.jpg}}
+	\subfloat [Structure and components of \textit{P.oceanica}. Source\cite{DiCarlo04}\label{DiCarlo04}]{\includegraphics[width=0.3\textwidth]{Fig-1-8.jpg}}
+	\caption{Morphology of \textit{P.oceanica}}\label{fig:8}
+\end{figure}
+
 Representing one of the most productive Mediterranean ecosystems, \textit{P.oceanica} usually serves as a
@@
 \cite{Boudouresque89}\label{Boudouresque89}.
-Some authors \cite{Guidetti08,Montefalcone09}\label{Guidetti08} \label{Montefalcone09} used status and population dynamics of \textit{P.oceanica} as indicators for the evaluation
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-1-7.jpg}
-	\caption{Scheme of matte structure of \textit{	P.oceanica}.\\ Source \cite{Pergent90}\label{Pergent90}}
-	\label{fig:7}
-\end{wrapfigure}
-
-of the meadow health status. \\
-There are several environmental factors, determining the growth of the \textit{P.oceanica}. The adaptation\index{Seagrasses!\textit{P.oceanica}!environmental adaptation}
+Some authors \cite{Guidetti08,Montefalcone09}\label{Guidetti08} \label{Montefalcone09} used status and population dynamics of \textit{P.oceanica} as indicators for the evaluation of the meadow health status.
+There are several environmental factors, determining the growth of the \textit{P.oceanica}. 
+The adaptation\index{Seagrasses!\textit{P.oceanica}!environmental adaptation}
 to dry-summer subtropical climate reduces its extension to Mediterranean area only (Fig.\ref{fig:5}).
@@
 the different maximum summer temperatures at those depths \cite{Buia91}\label{Buia91}.
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-1-8.jpg}
-	\caption{Structure and components of \textit{P.oceanica} \\ Source \cite{DiCarlo04}\label{DiCarlo04}}
-	\label{fig:8}
-\end{wrapfigure}
-
-The phenology\index{Seagrasses!\textit{P.oceanica}!phenology} of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in
-shallow waters plants have shorter and falciform leaves, compared to ones in the deeper and central
+The phenology\index{Seagrasses!\textit{P.oceanica}!phenology} of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in shallow waters plants have shorter and falciform leaves, compared to ones in the deeper and central
 areas \cite{Dural10}\label{Dural10}. In \textit{P.oceanica} flower abundance is related to the structure of the meadow with
@@
 properties incessantly. The effects of these processes influence application of the hyperspectral
-remote sensing\index{Optical properties of the seafloor!hyperspectral measurements} and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
+remote sensing\index{Seawater optical properties!hyperspectral measurements} and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
 sedimentation processes, which in turn, may change micro relief properties and topology. \\
@@
 \\ In such cases a retrieval of water optical parameters from the remote measurements should be tested and analyzed.
-\\
-\pagebreak
-\newpage
+\newline
 We briefly discuss below the most effective RTM software and algorithms from a range of various best-known up-to-date tools. 
@@
 		\item [*] The \textit{Second Simulation of a Satellite Signal in the Solar Spectrum}, version 1 \href{http://6s.ltdri.org/}{(6SV1)}\index{Seawater optical properties!RTM!6SV1} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \href{http://modis.gsfc.nasa.gov/}{NASA MODIS} satellite imagery lookup tables.
-The 6S algorithm has been implemented by the \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{GRASS} software for the atmospheric correction
+The 6S algorithm has been implemented by the \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{GRASS} software for the atmospheric correction 						\pagebreak
 		\item [*] Another example of the RTM adjusted for a specific imagery is a German \href{http://www.imk-asf.kit.edu/english/312.php}{\textit{KOPRA RTM}}\index{Seawater optical properties!RTM!KOPRA}, fitted for \href{http://envisat.esa.int/instruments/meris/}{MIPAS/ENVISAT} imagery.
@@
 For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}}\index{Seawater optical properties!RTM!WASI} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
+\pagebreak
 
@@
 the benthic habitat discrimination. The results of the \textit{in-situ} measurements and observation are usually
-managed and treated using integrated GIS approach, as, for example, reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
+managed and treated using integrated GIS approach, as, e.g. reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
 
@@
 available high-resolution imagery at 1- and 4-meter resolution from \href{http://www.geoeye.com}{Geoeye}. IKONOS
-imagery has been applied for the seagrass mapping due to its high resolution and accessibility. Thus,
-the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
+imagery has been applied for the seagrass mapping due to its high resolution and accessibility.
+
+\begin{wrapfigure}{r}{0.5\textwidth}[H]
+	\centering
+	\includegraphics[scale=0.30]{Fig-14.jpg}
+	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
+	\label{fig:2.5}
+\end{wrapfigure}
+
+Thus, the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
 made using IKONOS, Landsat TM, and CASI\index{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
@@
 than IKONOS \cite{Mumby02}\label{Mumby02}.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-14.jpg}
-	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
-	\label{fig:2.5}
-\end{wrapfigure}
-
 Another comparative analysis of the application of \href{http://www.itres.com/products/imagers/casi550}{CASI}, Landsat and \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird} imagery \cite{Phinn08}\label{Phinn08}
@@
 Landsat TM images with their 30*30m resolution.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.15]{Fig-15.jpg}
+	\caption{Reflectance spectra of seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
+		reflectance. Source:\cite{Yang10}\label{Yang10}}
+	\label{fig:2.6}
+\end{wrapfigure}
+
 Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}\index{Remote sensing!multispectral imagery!ASTER}), launched in
@@
 reported, for example, by Hirose \cite{Hirose04}\label{Hirose04}.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-15.jpg}
-	\caption{Reflectance spectra of seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
-		reflectance. Source:\cite{Yang10}\label{Yang10}}
-	\label{fig:2.6}
-\end{wrapfigure}
 
@@
 \end{figure}
-
-\begin{figure}[h]
+\begin{figure}[H]
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.43, angle=90]{UML_StatAn.jpg}
+	\includegraphics[scale=0.5, angle=90]{UML_StatAn.jpg}
 	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
@@
 
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.45]{GNU-13.jpg}
-		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
-			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.50]{GNU-13.jpg}
+		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
+Series 1-100.
+Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
+\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
 		\label{fig:4.13}
-	\end{center}
-\end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.5]{GNU-16.jpg}
-		\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown mid spread of the statistical quartiles Q1 and Q3 (vertical dashes) and
-			mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
-		\label{fig:4.14}
 	\end{center}
@@
 Answering the first research question, we suggest the following statement.
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{GNU-16.jpg}
+		\caption{Multiplot of spectral reflectance of carbonate sand.
+Series 1-75.
+Shown mid spread of the statistical quartiles Q1 and Q3 (vertical dashes) and	mean value within the range (red bold dots).
+\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
+		\label{fig:4.14}
+	\end{center}
+\end{figure}
+
 Assuming that the null hypothesis, Hypothesis Ho, is true, which is “the spectral distinguishability of
@@
 
-\begin{table}[H]
+\begin{table}[H]\footnotesize
 	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}\label{tab:A.1}
@@
 		     PositionLongitude & 25.01269=25°0'45.6834" & 25.0098 = 25°0'35.2794" \\ \hline
-		     Comment & 	extit{P.oceanica} & 	extit{P.oceanica} \\ \hline
+		     Comment & \textit{P.oceanica} & \textit{P.oceanica} \\ \hline
 		     CommentSub1 & 0.5m depth & 2.5m depth \\ \hline
@@
 
-\begin{table}[H]
+\begin{table}[H]\footnotesize
 	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}\label{tab:A.2}	\centering
@@
 	\centering
-	\includegraphics[scale=0.5]{Fig-24.jpg}
+	\includegraphics[scale=0.55]{Fig-24.jpg}
 	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
@@
 	\begin{center}
-		\includegraphics[scale=0.20]{GNU-401-420.jpg}
+		\includegraphics[scale=0.25]{GNU-401-420.jpg}
 		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
@@
 \end{figure}
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.20]{GNU-401-420-candles.jpg}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.25]{GNU-401-420-candles.jpg}
 		\caption{Statistical analysis of the measurement data: spectral reflectance of the \textit{P.oceanica}. 
@@
 	\begin{center}
-		\includegraphics[scale=0.20]{GNU-enlarged-M-401-420.jpg}
+		\includegraphics[scale=0.25]{GNU-enlarged-M-401-420.jpg}
 		\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
@@
 	\begin{center}
-		\includegraphics[scale=0.25]{GNU-Irr-Splines.jpg}
+		\includegraphics[scale=0.30]{GNU-Irr-Splines.jpg}
 		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}­}
@@
 	\centering
-	\includegraphics[scale=0.25]{GNU-Irr.jpg}
+	\includegraphics[scale=0.30]{GNU-Irr.jpg}
 	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}}
@@
 	\begin{center}
-		\includegraphics[scale=0.25]{G-period-sand30.jpg}
+		\includegraphics[scale=0.4]{G-period-sand30.jpg}
 		\caption{Plot illustrating periodogram for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
@@
 \end{figure}
-\pagebreak
 
@@
 	\begin{center}
-		\includegraphics[scale=0.32]{G-polynom-trend-wo.jpg}
+		\includegraphics[scale=0.4]{G-polynom-trend-wo.jpg}
 		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15.
@@
 	\begin{center}
-		\includegraphics[scale=0.32]{G-polynom-trend-with.jpg}
+		\includegraphics[scale=0.4]{G-polynom-trend-with.jpg}
 		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
@@
 	\begin{center}
-		\includegraphics[scale=0.32]{G-exp-mov-aver.jpg}
+		\includegraphics[scale=0.4]{G-exp-mov-aver.jpg}
 		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
@@
 	\centering
-	\includegraphics[scale=0.35]{Fig-26.jpg}
+	\includegraphics[scale=0.5]{Fig-26.jpg}
 	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES. \href{http://www.gnuplot.info/}{Gnuplot}}
@@
 \newpage
+\section*{List of People}
+\markright{List of People}
+\addcontentsline{toc}{section}{List of People}
+\pagebreak
+
+\singlespacing
+\cleardoublepage
+\newpage
 \phantomsection

### 2011-02-02 MScDiplom_Lemenkova.tex
finished 'List of Acronyms', added hyperreferences, logos, footnotes

@@
 \documentclass[11pt, a4paper]{article}
+\usepackage{authblk}
+%\usepackage{authorindex}
 \usepackage{color}
+\usepackage[stable]{footmisc }
 \usepackage{setspace}
+\usepackage[printonlyused,withpage]{acronym}
 \usepackage[a4paper, left=30mm,right=25mm,top=30mm,bottom=35mm]{geometry}
@@
 	\centering
-	\title{Seagrass mapping and monitoring along the coasts of Crete, Greece}
-	\author{by Lemenkova Polina}
+	\title{Seagrass mapping and monitoring \\ along the coasts of Crete, Greece}
+	\author[1]{by Lemenkova Polina}
+	\affil[1]{\textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
 	\date{\today}
@@
 \end{itemize}
+
+\begin{figure}[H]
+	\centering
+	\subfloat{\includegraphics [width=0.2\textwidth] {Logo-Sou.jpg}}
+	\hspace{3mm}
+	\subfloat{\includegraphics [width=0.1\textwidth] {Logo-Lun.jpg}}
+	\hspace{3mm}
+	\subfloat{\includegraphics [width=0.1\textwidth] {Logo-War.jpg}}
+	\hspace{3mm}
+	\subfloat{\includegraphics [width=0.5\textwidth] {ITC-logo.jpg}}	
+\end{figure}
+
 \pagebreak
@@
 	\item Name Examiner 2\dotfill
-	\item External Examiner \dotfill \\ \hfill{Dr. \textit{Petter Pilesj\"o} (Lund University, GIS Centre)}
+	\item External Examiner \dotfill \\ \hfill{Dr.{\textit{Petter Pilesj\"o}}(Lund University, GIS Centre)}
 \end{itemize}
@@
 \vspace{1mm}\\
-Many thanks for Dr.\textit{Dingtian Yang} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.
+Many thanks for Dr.
+\author{\textit{Dingtian Yang}} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.
 
@@
 The diverse spectral values entail variations in optical properties of the seafloor cover types at changing environmental conditions.
-We applied WASI simulation techniques for the modelling of the optical parameters of various seafloor cover types by various spaceborne imaging spectrometers (MERIS, SeaWiFS, CZCS and MODIS), in order to understand their suitability and possible limitations for the seagrass mapping.\\
+We applied \ac{WASI} simulation techniques for the modelling of the optical parameters of various seafloor cover types by various spaceborne imaging spectrometers (\ac{MERIS}, \ac{SeaWiFS}, \ac{CZCS} and \ac{MODIS}), in order to understand their suitability and possible limitations for the seagrass mapping.\\
 Fieldwork research sites were presented by separate locations on the northern coast of Crete region (Ligaria, Agia Pelagia, Xerocampos).
@@
 \textit{Chapter 4}: \nameref{sec:4} presents to the reader the main results of the current MSc work, obtained during the processing fieldwork and other collected data.
-It starts from the review of the collected data from different sources, then describes the modelling component of the research, namely WASI colour simulation and plotted resulting graphs.
+It starts from the review of the collected data from different sources, then describes the modelling component of the research, namely \ac{WASI} colour simulation and plotted resulting graphs.
 We also discuss here the particularities of various sensors, suitable enough for deriving radiometric information to discriminate sand from the seagrass.
@@
 
-\section*{List of Acronyms}
+\section*{List of Acronyms\footnote{The page is given where the glossary is first entered and defined}}
 	\markright{Acronyms}
 	\addcontentsline{toc}{section}{List of Acronyms}
-\doublespacing
-\textbf{4AOP}: Automatized Atmospheric Absorption Atlas Operational Release \\
-\textbf{6SV1}:  Second Simulation of a Satellite Signal in the Solar Spectrum, Version1\\
-\textbf{ANOVA}: ANalysis of VAriance\\
-\textbf{ASTER}: Advanced Spaceborne Thermal Emission and Reflection Radiometer\\
-\textbf{BRDF}:  Bidirectional Reflectance Distribution Function\\
-\textbf{CASI}:		Compact Airborne Spectrographic Imager\\
-\textbf{CNES}: 		Centre National d'Etudes Spatiales (of France)\\
-\textbf{CRTM}:		Community Radiative Transfer Model\\
-\textbf{CZCS}:		 Coastal Zone Color Scanner\\
-\textbf{DISORT}:	 Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium\\
-\textbf{ENVISAT}:	ENVIronmental SATellite\\
-\textbf{ESA}:		European Space Agency\\
-\textbf{ERS}: 		European Remote Sensing satellite\\
-\textbf{GIS}:		Geographic Information System\\
-\textbf{GNU}:		GNU's Not Unix \\
-\textbf{GPS}:		 Global Positioning System\\
-\textbf{GRASS}:	 Geographic Resources Analysis Support System\\
-\textbf{GRETL}:		GNU Regression, Econometrics and Time-series Library\\
-\textbf{HyMap}:	Hyperspectral Mapper\\
-\textbf{HyperOCR}:	Hyperspectral Ocean Colour Radiometer\\
-\textbf{IKONOS}: 	derived its name from the Greek term \textit{eikon} for image\\
-\textbf{ILWIS}:		Integrated Land and Water Information System\\
-\textbf{iPAQ}:	internet Pocket PC by CompAQ\\
-\textbf{KOPRA}:	 Karlsruhe Optimized and Precise Radiative transfer Algorithm\\
-\textbf{LaTeX}:		Lamport Tau Epsilon Chi (Greek for \textit{techne}), modified by L.Lamport\\
-\textbf{LAI}:	 Leaf Area Index\\
-\textbf{Landsat TM}:	 Landsat Thematic Mapper \\
-\textbf{Landsat ETM+}:	Landsat Enhanced Thematic Mapper Plus\\
-\textbf{LyX}: 		LyriX-files\\
-\textbf{MERIS}:		MEdium Resolution Imaging Spectrometer\\
-\textbf{MIPAS}:		Michelson Interferometer for Passive Atmospheric Sounding\\
-\textbf{MODIS}:	Moderate Resolution Imaging Spectroradiometer\\
-\textbf{NASA}:		National Aeronautics and Space Administration\\
-\textbf{NIR}:		 Near-Infra-Red\\
-\textbf{NOAA}:		 National Oceanic and Atmospheric Administration\\
-\textbf{OBC}:		On Board Calibration\\
-\textbf{Olympus ST}: 	 Olympus Stylus Tough\\
-\textbf{Open EV}: 	Open Earth View\\
-\textbf{RAMSES}:	RAdiance Measurement SEnSor\\
-\textbf{RRTM}:		Rapid Radiative Transfer Model\\
-\textbf{RTTOV}: 	Radiative Transfer for TIROS Operational Vertical Sounder\\
-\textbf{RTM}:		Radiative Transfer Model\\
-\textbf{SAM}:		Scanning Acoustic Microscope\\
-\textbf{SCUBA}:	Self Contained Underwater Breathing Apparatus\\
-\textbf{SeaWiFS}: 	Sea-viewing Wide Field-of-view Sensor\\
-\textbf{SPOT}:		Satellite Pour l'Observation de la Terre\\
-\textbf{SPSS	}:	Statistical Package for the Social Sciences\\
-\textbf{STDEV}:		Standard Deviation\\
-\textbf{SWIR}: 		Short Wave InfraRed\\
-\textbf{TIR}: 		Thermal InfraRed\\
-\textbf{TIROS}: 	Television and InfraRed Observation Satellite\\
-\textbf{UML}:		Unified Modelling Language\\
-\textbf{UV}:		UltraViolet (10 nm to 400 nm)\\
-\textbf{UVA}:		UltraViolet A, or long waves (315 nm-400 nm)\\
-\textbf{UVB}:		 UltraViolet B, or medium waves (280 nm-315 nm)\\
-\textbf{VIS}:		VISible\\
-\textbf{VNIR}:		Visible and Near-InfraRed\\
-\textbf{WASI}:		WAter colour SImulator
-\pagebreak
+%\singlespacing
+\begin{acronym}
+	\acro{4AOP}{\href{http://www.noveltis.fr/4AOP/interface/index.php}{Automatized Atmospheric Absorption Atlas Operational Release}} 
+	\acro{6SV1}{Second Simulation of a Satellite Signal in the Solar Spectrum Version1} 
+	\acro{ASCII}{\href{http://en.wikipedia.org/wiki/ASCII}{American Standard Code for Information Interchange}}
+	\acro{ANOVA}{\href{http://en.wikipedia.org/wiki/Anova}{ANalysis of VAriance}} 
+	\acro{ASTER}{\href{http://asterweb.jpl.nasa.gov/}{Advanced Spaceborne Thermal Emission and Reflection Radiometer}} 
+	%\acro{BRDF}{Bidirectional Reflectance Distribution Function}
+	\acro{CASI}{\href{http://www.borstad.com/casi.html}{Compact Airborne Spectrographic Imager}}
+	\acro{CNES}{\href{http://www.cnes.fr/}{Centre National d'Etudes Spatiales (of France)}}
+	\acro{CRTM}{\href{http://www.jcsda.noaa.gov/projects_crtm.php}{Community Radiative Transfer Model}}
+	\acro{CZCS}{\href{http://oceancolor.gsfc.nasa.gov/CZCS/}{Coastal Zone Color Scanner}}
+	\acro{DISORT}{\href{ftp://climate1.gsfc.nasa.gov/wiscombe/Multiple_Scatt/DISORT1.2/}{Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium}}
+	\acro{ENVISAT}{\href{http://envisat.esa.int/instruments/meris/}{ENVIronmental SATellite}}
+	\acro{ESA}{\href{http://www.esa.int/}{European Space Agency}}
+	%\acro{ERS}{European Remote Sensing satellite}
+	\acro{GIS}{Geographic Information System}
+	%\acro{GNU}{GNU's Not Unix}
+	\acro{GloVis}{\href{http://glovis.usgs.gov/}{Global Visualization Viewer}}
+	\acro{GPS}{Global Positioning System}
+	\acro{GRASS}{\href{http://grass.fbk.eu/}{Geographic Resources Analysis Support System}}
+	%\acro{GRETL}{GNU Regression, Econometrics and Time-series Library}
+	\acro{HyMap}{\href{http://www.hyvista.com/}{Hyperspectral Mapper}}
+	\acro{HyperOCR}{\href{http://www.satlantic.com/Hyperspectral_Radiometers/Hyperspectral_Radiometers.asp}{Hyperspectral Ocean Colour Radiometer}}
+	\acro{ILWIS}{\href{http://www.ilwis.org/}{Integrated Land and Water Information System}}
+	\acro{iPAQ}{\href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{Internet Pocket PC by CompAQ}}
+	\acro{KOPRA}{\href{http://www.imk-asf.kit.edu/english/312.php}{Karlsruhe Optimized and Precise Radiative transfer Algorithm}}
+	\acro{LAI}{\href{http://en.wikipedia.org/wiki/Leaf_Area_Index}{Leaf Area Index}}
+	\acro{Landsat TM}{\href{http://landsat.gsfc.nasa.gov/}{Landsat Thematic Mapper}}
+	\acro{Landsat ETM+}{\href{http://landsat.gsfc.nasa.gov/}{Landsat Enhanced Thematic Mapper Plus}}
+	\acro{MERIS}{\href{http://envisat.esa.int/instruments/meris/}{MEdium Resolution Imaging Spectrometer}}
+	\acro{MIPAS}{\href{http://envisat.esa.int/instruments/mipas/}{Michelson Interferometer for Passive Atmospheric Sounding}}
+	\acro{MODIS}{\href{http://modis.gsfc.nasa.gov/}{Moderate Resolution Imaging Spectroradiometer}}
+	\acro{NASA}{\href{http://www.nasa.gov/}{National Aeronautics and Space Administration}}
+	\acro{NIR}{Near-Infra-Red}
+	\acro{NOAA}{\href{http://www.osdpd.noaa.gov/ml/index.html}{National Oceanic and Atmospheric Administration}}
+	\acro{OBC}{On Board Calibration}
+	\acro{Olympus ST}{\href{http://www.olympusamerica.com/}{Olympus Stylus Tough}}
+	\acro{RRTM}{\href{http://rtweb.aer.com/}{Rapid Radiative Transfer Model}}
+	\acro{RTTOV-9}{\href{http://research.metoffice.gov.uk/research/interproj/nwpsaf/rtm/}{Radiative Transfer for TIROS (Television and InfraRed Observation Satellite) Operational Vertical Sounder}}
+	\acro{RTM}{Radiative Transfer Model}
+	\acro{SAM}{Scanning Acoustic Microscope}
+	\acro{SCUBA}{\href{http://www.padi.com/scuba/}{Self Contained Underwater Breathing Apparatus}}
+	\acro{SeaWiFS}{\href{http://oceancolor.gsfc.nasa.gov/SeaWiFS/}{Sea-viewing Wide Field-of-view Sensor}}
+	\acro{SPOT}{\href{http://www.spotimage.fr/}{Satellite Pour l'Observation de la Terre}}
+	\acro{SPSS}{ \href{http://www.spss.com/}{Statistical Package for the Social Sciences}}
+	\acro{STDEV}{Standard Deviation}
+	\acro{SWIR}{Short Wave InfraRed}
+	\acro{TIR}{Thermal InfraRed}
+	\acro{Trios-RAMSES}{\href{http://trios.de/index.php?option=com_content&view=category&layout=trios&id=47&Itemid=76&lang=en}{RAdiance Measurement SEnSor}}
+	\acro{UML}{Unified Modelling Language}
+	\acro{USGS}{\href{http://www.usgs.gov/}{U.S. Geological Survey}}
+	\acro{UV}{UltraViolet (10 nm to 400 nm)}
+	\acro{UVA}{UltraViolet A, or long waves (315 nm-400 nm)}
+	\acro{UVB}{UltraViolet B, or medium waves (280 nm-315 nm)}
+	\acro{VIS}{VISible}
+	\acro{VNIR}{Visible and Near-InfraRed}
+	\acro{WASI}{\href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}}
+\end{acronym}
+\pagebreak
+
 \singlespace
@@
 		\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
-		\item \href{http://www.ilwis.org/}{ILWIS} , \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
+		\item \href{http://www.ilwis.org/}{ILWIS}, \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
 		\item \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}, for modelling optical water 					properties under changing environment
-		\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and GPS for data capture during the fieldwork
+		\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and \ac{GPS} for data capture during the fieldwork
 		\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
@@
 		\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
-		\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data pre-processing and preliminary statistical evaluation
-
-		\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
+		\item \href{http://www.openoffice.org/}{Open Office}: document typesetting, data pre-processing and preliminary statistical evaluation
+
+		\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)}: statistical data edition and analysis 
 		\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
-		\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia} , \href{http://argouml.tigris.org/}						{ArgoUML}: diagramming flowcharts for work algorithms
+		\item \href{http://staruml.sourceforge.net/en/}{StarUML}, \href{http://live.gnome.org/Dia}{Dia}, \href{http://argouml.tigris.org/}						{ArgoUML}: \ac{UML} diagram tool for creating flowcharts of work algorithms
 		\item \href{http://inkscape.org/}{Inkscape}: drawing
@@
 		\item \href{http://www.lyx.org/}{LyX} for exporting \LaTeX document to the Word .doc format via HTML
-		\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
+		\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX}: converting spreadsheets into \LaTeX environment
 		\item \href{http://openev.sourceforge.net/}{OpenEV}: used for raster Google Earth imagery viewing, visualisation and preliminary analysis
@@
 provides sediments stability by reducing the degree of water movements, etc \cite{Francour99}\label{Francour99}.\\
-The purpose of current MSc research work aims to apply methods of remote sensing analysis, including Radiative Transfer Models (RTM), 
-GIS-based spatial analysis, processing and classification of satellite and aerial photos, as well as videometric underwater footage, towards mapping and environmental monitoring of seagrass \textit{P.
+The purpose of current MSc research work aims to apply methods of remote sensing analysis, including \ac{RTM}, 
+\ac{GIS}-based spatial analysis, processing and classification of satellite and aerial photos, as well as videometric underwater footage, towards mapping and environmental monitoring of seagrass \textit{P.
 oceanica} along the selected locations of the northern coasts of Crete island, Greece.
-The technical implementation is based on WASI RTM software, GIS (ILWIS, Erdas Imagine and ArcGIS), using aerial and satellite images and the results of the underwater videometric measurements.
+The technical implementation is based on \ac{WASI} RTM software, \ac{GIS} (ILWIS, Erdas Imagine and ArcGIS), using aerial and satellite images and the results of the underwater videometric measurements.
 
@@
 Other example of monitoring of \textit{P.oceanica} using remote sensing techniques \cite{Calvo03}\label{Calvo03}
-reports the application of the CZCS images towards the case study of the Italian coast and shows
+reports the application of the \ac{CZCS} images towards the case study of the Italian coast and shows
 successful results of the neural-based classification using Isodata method of supervised classification.
@@
 \subsection{Research objective}\index{Research!objective}
-The current MSc research aims to explore the environmental conditions for the spatial distribution of \textit{P.oceanica} seagrass along the northern coast of Crete Island, based on the remote sensing and GIS techniques, knowledge about the coastal environment in Crete and integration of various data
+The current MSc research aims to explore the environmental conditions for the spatial distribution of \textit{P.oceanica} seagrass along the northern coast of Crete Island, based on the remote sensing and \ac{GIS} techniques, knowledge about the coastal environment in Crete and integration of various data
 from the following sources \index{Data}:
@@
 	\end{itemize}
-The research aims to explore the limitations of the application of the Radiative Transfer Models (further RTMs) and remote sensing techniques 
+The research aims to explore the limitations of the application of the \ac{RTM} and remote sensing techniques 
 towards the study of the environmental properties of the \textit{P.oceanica} meadows distribution
@@
 of Crete (Ligaria beach in Agia Pelagia district and Xerocampos), using following methods of the
-remote sensing techniques: spectral modelling by means of the RTM Water Colour Simulator WASI,  underwater videometric measurements made by Olympus camera, aerial Google Earth and satellite images from different sources, spatial GIS and statistical analysis.
+remote sensing techniques: spectral modelling by means of the\ac{RTM} \ac{WASI},  underwater videometric measurements made by \ac{Olympus ST}camera, aerial Google Earth and satellite images from different sources, spatial \ac{GIS} and statistical analysis.
 
@@
 		\item[]1.	To study narrow-band spectral reflectance properties of \textit{P.oceanica} and other seafloor cover types (sand and silt) using WASI water colour simulation software 
-		\item[]2.	To use methods of the in situ diving observations and underwater videometric measurements by Olympus camera in order to receive large-scale imagery of the P.oceanica mattes 
-		\item[]3.	To apply remote sensing data (Google Earth aerial images, Landsat TM and ETM+ satellite images) for the monitoring of the seagrass meadows distribution 
+		\item[]2.	To use methods of the in situ diving observations and underwater videometric measurements by \ac{Olympus ST}camera in order to receive large-scale imagery of the P.oceanica mattes 
+		\item[]3.	To apply remote sensing data (Google Earth aerial images, \ac{Landsat TM} and ETM+ satellite images) for the monitoring of the seagrass meadows distribution 
 		\item[]4.	To perform supervised images classification for the thematic mapping of the \textit{P.oceanica} seagrass distribution along the selected locations over the coasts of northern Crete. 
@@
 the equality of the statistical variances.
-The hypothesis testing is suggested to be carried out using the ANOVA statistical test. The purpose of
+The hypothesis testing is suggested to be carried out using the \ac{ANOVA} statistical test. The purpose of
 ANOVA test is to visualize in an effective and quick way the spectral differences between seagrass
@@
 optical sensors, GER-Series Field portable spectroradiometers from
-\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. Trios-RAMSES\index{Seawater optical properties!spectroradiometers!Trios-RAMSES} hyperspectral radiometers
-(Trios-RAMSES Hyperspectral UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-
-VIS Radiance Sensor) are small-sized, low power-consuming, flexible for fieldwork yet with high
+\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. \ac{Trios-RAMSES}\index{Seawater optical properties!spectroradiometers!Trios-RAMSES} hyperspectral radiometers
+(\ac{Trios-RAMSES} Hyperspectral \ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor) are small-sized, low power-consuming, flexible for fieldwork yet with high
 level of precision, specially calibrated for air and for water application as well as color
@@
 Simulation of the radiance quantities for natural water bodies enables to analyse seafloor color remote sensing properties.
-Therefore, the artificial modelling of the seawater optical properties by means of the Radiative Transfer Modelling (further RTM) is used when water optics is studied under changing environment (e.g., depths, sun angle, suspended particles in water column).
+Therefore, the artificial modelling of the seawater optical properties by means of the \ac{RTM} is used when water optics is studied under changing environment (e.g., depths, sun angle, suspended particles in water column).
 \\ In such cases a retrieval of water optical parameters from the remote measurements should be tested and analyzed.
@@
 	\begin{itemize}
-		\item [*] The \textit{Second Simulation of a Satellite Signal in the Solar Spectrum}, version 1 \href{http://6s.ltdri.org/}{(6SV1)}\index{Seawater optical properties!RTM!6SV1} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \href{http://modis.gsfc.nasa.gov/}{NASA MODIS} satellite imagery lookup tables.
-The 6S algorithm has been implemented by the \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{GRASS} software for the atmospheric correction 						\pagebreak
-		\item [*] Another example of the RTM adjusted for a specific imagery is a German \href{http://www.imk-asf.kit.edu/english/312.php}{\textit{KOPRA RTM}}\index{Seawater optical properties!RTM!KOPRA}, fitted for \href{http://envisat.esa.int/instruments/meris/}{MIPAS/ENVISAT} imagery.
+		\item [*] The \ac{6SV1} \index{Seawater optical properties!RTM!6SV1} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \ac{MODIS} satellite imagery lookup tables.
+The 6S algorithm has been \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{implemented} by the \ac{GRASS} software for the atmospheric correction
+		\item [*] Another example of the RTM adjusted for a specific imagery is a German \ac{KOPRA}\index{Seawater optical properties!RTM!KOPRA}, fitted for \ac{MIPAS} and \ac{ENVISAT} imagery.
 		\item [*]  The \href{http://www.sequoiasci.com/products/Hydrolight.aspx}{\textit{HydroLight RTM}} commercial software is an advanced model for oceanic radiative transfer calculations, developed by the Sequoia Scientific, Inc., USA.
 It is designed to solve a wide range of problems in optical oceanography and limnology.
-		\item [*] The British \href{http://research.metoffice.gov.uk/research/interproj/nwpsaf/rtm/}{\textit{RTTOV-9 RTM}}\index{Seawater optical properties!RTM!RTTOV-9} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers.
+		\item [*] The British \ac{RTTOV-9}\index{Seawater optical properties!RTM!RTTOV-9} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers.
 It includes a number of useful tolls, e.g.
@@
 This program runs under Unix/Linux platforms and is open source.
-		\item [*] The \href{http://rtweb.aer.com/}{\textit{Rapid Radiative Transfer Model (RRTM)}}\index{Seawater optical properties!RTM!RRTM} is one more RTM freeware, designed in the USA, Massachusetts. 
+		\item [*] The \ac{RRTM}\index{Seawater optical properties!RTM!RRTM} is one more RTM freeware, designed in the USA, Massachusetts. 
 		\item [*] The \href{http://www.modtran.org/}{\textit{MODTRAN}}\index{Seawater optical properties!RTM!MODTRAN}, a commercial software from Spectral Sciences Inc., is another example of the RTM.
-Among its newest updates is generation of atmospheric correction data; it also has an option to write spectral output in binary, and convert it to ASCII.
-		\item [*] A \textit{Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium} \href{ftp://climate1.gsfc.nasa.gov/wiscombe/Multiple_Scatt/DISORT1.2/}{(DISORT)}\index{Seawater optical properties!RTM!DISORT}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
-		\item [*] The \textit{Community Radiative Transfer Model}, \href{http://www.jcsda.noaa.gov/projects_crtm.php}{CRTM}\index{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA, NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
-		\item [*] An \textit{Automatized Atmospheric Absorption Atlas Operational Release} \href{http://www.noveltis.fr/4AOP/}{(4AOP)}\index{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{CNES} (the French Space Agency).
+Among its newest updates is generation of atmospheric correction data; it also has an option to write spectral output in binary, and convert it to \ac{ASCII}.
+		\item [*] A \ac{DISORT}\index{Seawater optical properties!RTM!DISORT}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
+		\item [*] The \ac{CRTM} \index{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA}, \ac{NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
+		\item [*] An \ac{4AOP}\index{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{the French Space Agency} (\ac{CNES}).
 Being a fine program which enables accurate computations of the radiance, defining other emissivity functions and other set up, it is however, particularly designed for the infrared waves (thus, not quite suitable in our case).
@@
 	\end{itemize}
-For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}}\index{Seawater optical properties!RTM!WASI} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
-\pagebreak
+For our purpose we have chosen the \ac{WASI}\index{Seawater optical properties!RTM!WASI} \ac{RTM} software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
 
@@
 scheme \cite{McKenzieetal03} \label{McKenzieetal03}. The seagrass is being sampled on the selected sites using transect lines,
-quadrant frame, single point markers, markers, GPS and other equipment. The seagrass sampling is
+quadrant frame, single point markers, markers, \ac{GPS} and other equipment. The seagrass sampling is
 taken on the regular way with observation points covering the study area with normal distribution.
@@
 the benthic habitat discrimination. The results of the \textit{in-situ} measurements and observation are usually
-managed and treated using integrated GIS approach, as, e.g. reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
+managed and treated using integrated \ac{GIS} approach, as, e.g. reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
 
@@
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
-digitized aerial photographs, GPS data, remote sensing and SCUBA-based fieldwork measurements.
-SCUBA-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
+digitized aerial photographs, \ac{GPS} data, remote sensing and \ac{SCUBA}-based fieldwork measurements.
+\ac{SCUBA}-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
 mapping, is limited in application, because of their time consumption, weather-dependency and
 unsuitability for the case of monitoring large areas of water for small-scale mapping. 
-The underwater videography\index{Remote sensing!underwater videography} with a GPS\index{Remote sensing!GPS} is a tool of seagrass monitoring which has certain advantages, i.e.
+The underwater videography\index{Remote sensing!underwater videography} with a \ac{GPS}\index{Remote sensing!GPS} is a tool of seagrass monitoring which has certain advantages, i.e.
 high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
-collection in the field \cite{Schultz08}\label{Schultz08}. However, it cannot cover large areas for small-scale mapping.
-Remote sensing techniques offer clear advantages over other methods of \textit{in-situ} field measurements
-and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
-their weather-independence, cost-effectiveness, accuracy and spatial coverage, which enables
-periodic monitoring of the seagrass meadows and gives access to the distant and unapproachable
-areas. Integrated together with GIS vector layers and maps, remote sensing data enable historical
-mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.
+collection in the field \cite{Schultz08}\label{Schultz08}.
 
@@
 
+However, it cannot cover large areas for small-scale mapping.
+Remote sensing techniques offer clear advantages over other methods of \textit{in-situ} field measurements
+and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
+their weather-independence, cost-effectiveness, accuracy and spatial coverage, which enables
+periodic monitoring of the seagrass meadows and gives access to the distant and unapproachable
+areas. Integrated together with \ac{GIS} vector layers and maps, remote sensing data enable historical
+mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.
 However, application of the remote sensing techniques for mapping of submerged vegetation,
@@
 least, nation-specific, depending to large extent on the tools available for the researchers \cite{Mellors09a}\label{Mellors09a}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-12.jpg}
+	\caption{Difference between broadband multispectral and hyperspectral resolution of
+		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
+	\label{fig:2.3}
+\end{wrapfigure}
+
 Universal, international, standardized methods for seagrass directly for seagrasses
@@
 measurements of hyperspectral reflectance. 
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-12.jpg}
-	\caption{Difference between broadband multispectral and hyperspectral resolution of
-		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
-	\label{fig:2.3}
-\end{wrapfigure}
-
 Spectral measurements\index{Remote sensing!spectral measurements} of the target objects are made
@@
 spectrally based radiative transfer approach to quantitatively estimate shallow-water bathymetry and
-leaf area index (LAI) of the seagrass. The spectral reflectance in general is the
+leaf area index (\ac{LAI}) of the seagrass. The spectral reflectance in general is the
 result of the spectral absorption in different bands, typical for each target object.
@@
 susceptible by the means of satellite imagery from remote sensors, both with moderate resolution
-(e.g., \href{http://landsat.gsfc.nasa.gov/}{Landsat MSS, Landsat TM, Landsat ETM+}, \href{http://envisat.esa.int/instruments/meris/}{MERIS}, \href{http://asterweb.jpl.nasa.gov/}{ASTER}, \href{http://modis.gsfc.nasa.gov/}{MODIS}) and high resolution as
+(e.g., \href{http://landsat.gsfc.nasa.gov/}{Landsat MSS, Landsat TM, Landsat ETM+}, \href{http://envisat.esa.int/instruments/meris/}{MERIS} from \ac{ESA}, \href{http://asterweb.jpl.nasa.gov/}{ASTER}, \href{http://modis.gsfc.nasa.gov/}{MODIS}) and high resolution as
 well (e.g., \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird}, \href{http://www.spotimage.fr/}{SPOT}, \href{http://www.itres.com/products/imagers/casi550}{CASI}).
@@
 
-The Landsat TM\index{Remote sensing!multispectral imagery!Landsat TM} and Landsat ETM+\index{Remote sensing!multispectral imagery!Landsat ETM+} data, with the most recent from sensor Landsat 7 - an advanced
+The \ac{Landsat TM}\index{Remote sensing!multispectral imagery!Landsat TM} and \ac{Landsat ETM+}\index{Remote sensing!multispectral imagery!Landsat ETM+} data, with the most recent from sensor Landsat 7 - an advanced
 and multispectral scanning, launched in 1999, prove to be feasible and useful for the mapping of
 submerged vegetation, such as seagrasses or coral reefs. The successful applications of the imagery
-Landsat TM towards the seagrass mapping, were reported in numerous research works \cite{Palandro03,Gullstroom06,Wabnitz08,Bierwirth93,Ferguson97,Rasib97}\label{Palandro03}\label{Gullstroom06}\label{Wabnitz08}\label{Bierwirth93}\label{Ferguson97}\label{Rasib97}.\\ The Landsat data are particularly suitable for the
+\ac{Landsat TM} towards the seagrass mapping, were reported in numerous research works \cite{Palandro03,Gullstroom06,Wabnitz08,Bierwirth93,Ferguson97,Rasib97}\label{Palandro03}\label{Gullstroom06}\label{Wabnitz08}\label{Bierwirth93}\label{Ferguson97}\label{Rasib97}.\\ The Landsat data are particularly suitable for the
 case of change detection of seagrass landscapes at a decade scale, because being the main sensor
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
-Another well-known multispectral sensor, SPOT\index{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial
+Another well-known multispectral sensor, \ac{SPOT}\index{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial
 resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
 orbital repeat cycle for nadir viewing and imagery with a spatial resolution 20 - 2.5 m
-\cite{SPOT}\label{SPOT}. \href{http://www.spotimage.fr/}{SPOT} imagery was used for mapping beds of \textit{P.oceanica} in the
+\cite{SPOT}\label{SPOT}. \ac{SPOT} imagery was used for mapping beds of \textit{P.oceanica} in the
 Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}. 
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}[H]
+\begin{wrapfigure}{r}{0.5\textwidth}
 	\centering
@@
 Thus, the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
-made using IKONOS, Landsat TM, and CASI\index{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
-spectrum, the best results are achieved by the IKONOS and CASI, while Landsat TM has not high
+made using IKONOS, \ac{Landsat TM}, and \ac{CASI}\index{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
+spectrum, the best results are achieved by the IKONOS and \ac{CASI}, while \ac{Landsat TM} has not high
 enough resolution. 
@@
 Rayleigh scattering and defects of scattering. However, comparing between CASI and IKONOS, the
-same authors prove that CASI enable to receive still more accurate results of the classification
+same authors prove that \ac{CASI} enable to receive still more accurate results of the classification
 than IKONOS \cite{Mumby02}\label{Mumby02}.
 Another comparative analysis of the application of \href{http://www.itres.com/products/imagers/casi550}{CASI}, Landsat and \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird} imagery \cite{Phinn08}\label{Phinn08}
- demonstrates high suitability of CASI images for the fine-scale mapping of the seagrass
-landscapes. Thus, CASI and Quickbird-2\index{Remote sensing!multispectral imagery!Quickbird} images enable to identify even separate seagrass species
+ demonstrates high suitability of \ac{CASI} images for the fine-scale mapping of the seagrass
+landscapes. Thus, \ac{CASI} and Quickbird-2\index{Remote sensing!multispectral imagery!Quickbird} images enable to identify even separate seagrass species
 with small width and heterogeneous nature of the seagrass patches, which could not be detected using
 Landsat TM images with their 30*30m resolution.
+\ac{ASTER}\index{Remote sensing!multispectral imagery!ASTER}), launched in
+1999 on board Terra sensor\index{Remote sensing!multispectral imagery!Terra sensor}, provides high-resolution images of the Earth in 15 different bands of the
+electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:2.4}). It has diverse
+subsystems for the \ac{VNIR} with 15 m resolution, shortwave infrared \ac{SWIR}, and
+\ac{TIR} wavelength regions \cite{Kalinowski04}\label{Kalinowski04}. For each channel there is separate
+on board calibration \ac{OBC} system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}. 
+This makes \ac{ASTER} imagery especially suitable for the application towards detailed
+mapping of surface temperature, emissivity and reflectance of objects and bathymetric elevations as
+well. Successful application of the \ac{ASTER} imagery towards mapping of submerged vegetation
+reported, for example, by Hirose \cite{Hirose04}\label{Hirose04}. 
+Other multispectral images have also been used for the
+seagrass interpretation. For example, the success of using imagery from the multispectral airborne
+scanner Daedalus AADS1268 is reported by Heege et al \cite{Heege03}\label{Heege03} where they aim at classification of
+macrophytes in shallow waters of the Lake Constance.
 
@@
 
-Advanced Spectrometer for Thermal Emission and Reflection Radiometer (\href{http://asterweb.jpl.nasa.gov/}{ASTER}\index{Remote sensing!multispectral imagery!ASTER}), launched in
-1999 on board Terra sensor\index{Remote sensing!multispectral imagery!Terra sensor}, provides high-resolution images of the Earth in 15 different bands of the
-electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:2.4}). It has diverse
-subsystems for the visible near infrared (VNIR) with 15 m resolution, shortwave infrared (SWIR), and
-thermal infrared (TIR) wavelength regions \cite{Kalinowski04}\label{Kalinowski04}. For each channel there is separate
-on board calibration (OBC) system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}. 
-This makes ASTER imagery especially suitable for the application towards detailed
-mapping of surface temperature, emissivity and reflectance of objects and bathymetric elevations as
-well. Successful application of the ASTER imagery towards mapping of submerged vegetation
-reported, for example, by Hirose \cite{Hirose04}\label{Hirose04}.
-
-Other multispectral images have also been used for the
-seagrass interpretation. For example, the success of using imagery from the multispectral airborne
-scanner Daedalus AADS1268 is reported by Heege et al \cite{Heege03}\label{Heege03} where they aim at classification of
-macrophytes in shallow waters of the Lake Constance. 
 In regard to the methods chosen for the image interpretation, the supervised classifications proves to
@@
 wavelengths of hyperspectral imagery, while this information can be lost in the broader
-wavelengths of other sensors. With its 126 spectral bands, HyMap imagery enables to
+wavelengths of other sensors. With its 126 spectral bands, \ac{HyMap} imagery enables to
 distinguish features of interest, i.e. seagrass types \cite{Peneva07}\label{Peneva07}, which is the major
@@
 We mentioned in the \nameref{sec:0} the software that we used for logically separated research sections, which required various approach.
-Thus, for example,  spectral analysis and assessment of spectral signatures is technically based on WASI, statistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://gretl.sourceforge.net/}{Gretl} and \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
+Thus, for example,  spectral analysis and assessment of spectral signatures is technically based on \ac{WASI}, statistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://gretl.sourceforge.net/}{Gretl} and \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
 \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0} software is used for the spatial analysis, general mapping and cartographic layout
@@
 distribution using supervised classification.
-The research data include Google Earth aerial images and scenes from the \href{http://landsat.gsfc.nasa.gov/}{Landsat TM and ETM+}
-covering research period of 10 years in the same year time, taken from \href{http://www.usgs.gov/pubprod/}{USGS}, \href{http://glovis.usgs.gov/}{GloVis}, \href{http://www.osdpd.noaa.gov/ml/index.html}{NOAA} and \href{http://glcfapp.glcf.umd.edu:8080/esdi/index.jsp}{The Earth Science Data Interface}.
+The research data include Google Earth aerial images and scenes from the \ac{Landsat TM} and \ac{Landsat ETM+}
+covering research period of 10 years in the same year time, taken from \ac{USGS}, \ac{GloVis}, \ac{NOAA} and \href{http://glcfapp.glcf.umd.edu:8080/esdi/index.jsp}{The Earth Science Data Interface}. 
 The satellite imagery provides
@@
 above-ground seagrass patches, sediment seafloor cover types, species compositions, water depth and
-geographic locations recorded using GPS.
+geographic locations recorded using \ac{GPS}.
 
 \subsubsection{Fieldwork equipment}
-The research materials and equipment were provided by \href{http://www.nhmc.uoc.gr/index-2.htm}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:3.3}): 1) three iPAQs (Fig.\ref{fig:3.3a}); 2)
-three GPS (Fig.\ref{fig:3.3c}); 3) Three underwater video cameras, Olympus ST 8000  (Fig.\ref{fig:3.3d}), suitable for
+The research materials and equipment were provided by \href{http://147.52.130.151/MFIK_en/}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:3.3}): 1) three iPAQs (Fig.\ref{fig:3.3a}); 2)
+three GPS (Fig.\ref{fig:3.3c}); 3) Three underwater video cameras, \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), suitable for
 photographing up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
 Markers and cords for the depths measurements; 5) Waterproof plastic Otterbox  (Fig.\ref{fig:3.3b})for keeping the
-iPAQ dry; 6) \href{http://www.padi.com/scuba/}{SCUBA Diving equipment}, taken from the Ligaria Diving Center; 7) Boat (Fig.\ref{fig:3.4});
-
-\begin{figure}[h]
+iPAQ dry; 6) \ac{SCUBA} Diving equipment, taken from the Ligaria Diving Center; 7) Boat (Fig.\ref{fig:3.4});
+
+\begin{figure}[H]
 	\centering
@@
 
-A GPS and iPAQ have been used for detection of the geodetic coordinates and keeping the tracklogs
-along the boat route for GIS project.
+A \ac{GPS} and \ac{iPAQ} have been used for detection of the geodetic coordinates and keeping the tracklogs
+along the boat route for \ac{GIS} project.
 
@@
 boat path. The general locations of the sampling sites and routes were selected on the basis of the visual examination
-of the seagrass beds during snorkeling and SCUBA diving, recommendations from the Greek collaborators of the \textit{Natural History Museum of Crete}, and available maps covering the research area.
+of the seagrass beds during snorkeling and \ac{SCUBA} diving, recommendations from the Greek collaborators of the \textit{Natural History Museum of Crete}, and available maps covering the research area.
 To ensure even and objective selection of sampling sites
@@
 The videographic measurements and photos, enabling to detect types of the
-seafloor, were captured by means of three underwater digital video cameras Olympus ST 8000  (Fig.\ref{fig:3.3d}), at
+seafloor, were captured by means of three underwater digital video cameras \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), at
 five minutes interval along the tracklog path (Fig.\ref{fig:3.6}).
@@
 most determining factors for the seagrass locations. Measurements of depths were performed during
-the fieldwork using cord, iPAQs, and markers in order to assure that the videometric measurements
+the fieldwork using cord, \ac{iPAQ}, and markers in order to assure that the videometric measurements
 are taken at depths not more than four meters, in the shelf area.
 As a result, nine transects were established of one m wide and 20 m long track, to cover seagrass beds
-with videographic measurements. The GPS allowed to capture measurement locations on the iPAQ,
+with videographic measurements. The \ac{GPS} allowed to capture measurement locations on the \ac{iPAQ},
 encapsulated in a plastic waterproof Otterbox (Fig \ref{fig:3.3b}). The camera were adjusted horizontally by a leveler and
@@
 which located separately from broad meadows (Fig.\ref{fig:A.3} (c)).
-The results of the underwater videometric measurements of the Olympus cameras made during the
+The results of the underwater videometric measurements of the \ac{Olympus ST} cameras made during the
 ship route include nine total tracklog routes in the selected research area, including series of
@@
 Pelagia beach, Crete), sand (measured at Agia Pelagia beach, Crete), silt and default artificial
-spectrum of constant albedo at WASI (Fig.\ref{fig:A.9}). 
+spectrum of constant albedo at \ac{WASI} (Fig.\ref{fig:A.9}). 
 We included measurements of the Thalassia seagrass
@@
 	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:A.37})
-	\item[]4. Results of underwater videometric measurements of the Olympus cameras made during the ship route
+	\item[]4. Results of underwater videometric measurements of the \ac{Olympus ST} cameras made during the ship route
 \end{itemize}
@@
 
-The spatial resolution of Landsat ETM+ image is 30 m in the visible and near infrared bands (bands
-1-5 and 7); the spatial resolution of ASTER 15 m for the visible and near-infrared bands. IKONOS
-acquires data in 3 visible channels and NIR, with spatial resolution of 1-4 meters. ASTER and
+The spatial resolution of \ac{Landsat ETM+} image is 30 m in the visible and near infrared bands (bands
+1-5 and 7); the spatial resolution of \ac{ASTER} 15 m for the visible and near-infrared bands. IKONOS
+acquires data in 3 visible channels and \ac{NIR}, with spatial resolution of 1-4 meters. \ac{ASTER} and
 IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
@@
 The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
-the seafloor have been received by the means of the optical sensors Trios-RAMSES Hyperspectral
-UVA/UVB Irradiance Sensor and RAMSES-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.
+the seafloor have been received by the means of the optical sensors \ac{Trios-RAMSES} Hyperspectral
+UVA/UVB Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.
 \ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see \nameref{app}: Tab.
@@
 floor types with changing environmental conditions, using the differences in their spectral signatures
-on the graphs in a WASI, the Water Colour Simulator software. \\The Water Colour Simulator WASI, a
+on the graphs in a \ac{WASI} software. \\The \ac{WASI}, a
 software tool for analyzing and simulating the most common types of spectra \cite{Gege05}\label{Gege05}, is highly
 suitable for the seagrass spectral analysis.
-There are several environmental characteristics, included in WASI interface (Fig.\ref{fig:4.7}), which influence the
+There are several environmental characteristics, included in \ac{WASI} interface (Fig.\ref{fig:4.7}), which influence the
 results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
@@
 seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
-remote sensing sensors (MODIS, ASTER, MERIS and SeaWiFS), (Fig.\ref{fig:4.6}).
+remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}).
 
@@
 As can be seen on Fig.\ref{fig:A.21}, Fig.\ref{fig:A.22} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
-The autocorrelation analysis has been performed using both SPSS and Gretl software.
+The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{Gretl} software.
 \\
@@
 \paragraph{ANOVA-testing}\index{Data!statistical analysis!ANOVA}
-We applied ANOVA (ANalysis Of Variance) testing for the analysis of the probability of the
+We applied \ac{ANOVA} (ANalysis Of Variance) testing for the analysis of the probability of the
 spectral reflectance of \textit{P.oceanica} to be more or less spectrally distinct from other seafloor cover
@@
 The result P=.05 is (Tab.\ref{tab:8}), hence, very small, which makes the statement of Hypothesis Ho highly unlikely
-(less than 1 in a 10 chance). The one-way ANOVA highlighted a significant difference between data
+(less than 1 in a 10 chance). The one-way \ac{ANOVA} highlighted a significant difference between data
 of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
@@
 Mean: Average of the data to be plotted, AVERAGE\index{Data!statistical analysis!average} {data}
-St Dev: Standard deviation\index{Data!statistical analysis!standard deviation}, STDEV{data}
+St Dev: Standard deviation\index{Data!statistical analysis!standard deviation}, \ac{STDEV} {data}
 Median: Median of the data, MEDIAN{data}
@@
 The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
-types as the different sensors would percept them: MODIS\index{Remote sensing!multispectral imagery!MODIS}, ASTER\index{Remote sensing!multispectral imagery!ASTER}, MERIS\index{Remote sensing!multispectral imagery!MERIS}, SeaWiFS\index{Remote sensing!multispectral imagery!SeaWiFS} and CZCS\index{Remote sensing!multispectral imagery!CZCS}.\\
+types as the different sensors would percept them: MODIS\index{Remote sensing!multispectral imagery!MODIS}, ASTER\index{Remote sensing!multispectral imagery!ASTER}, MERIS\index{Remote sensing!multispectral imagery!MERIS}, SeaWiFS\index{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index{Remote sensing!multispectral imagery!CZCS}.\\
 These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
@@
 		\item[Seventh:] Other RTM software may be tested and the modelling outcomes compared.
-		\item[Eighth:] Application of various open source GIS (ILWIS, GRASS) could be very useful for the validation of the cartographic results, assessment of accuracy and comparison of various classification methods.
+		\item[Eighth:] Application of various open source \ac{GIS} (\ac{ILWIS}, \ac{GRASS}) could be very useful for the validation of the cartographic results, assessment of accuracy and comparison of various classification methods.
 		\item[Ninth:] The analysis of the health indicators of the seagrass (such as number of leaves per shoot, biomass estimation within the single shot, etc) was not considered in the current work, as it would go beyond the scope of the MSc thesis.
 However, ecological investigations could be used for the assessment of the vulnerability of the seagrass meadows in various locations on Crete.
-		\item[Tenth:] A flexible combination of the multi-scale mapping and results of the fieldwork measurements with GPS-referenced underwater footage would enable more profound analysis of the coastal environment on Crete.
+		\item[Tenth:] A flexible combination of the multi-scale mapping and results of the fieldwork measurements with \ac{GPS}-referenced underwater footage would enable more profound analysis of the coastal environment on Crete.
 	\end{description}
@@
 \begin{table}[H]\footnotesize
-	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}\label{tab:A.1}
+	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer and \ac{SAM} during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}\label{tab:A.1}
 	\centering
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.
-\href{http://www.spss.com/}{SPSS}}
+	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.\ac{SPSS}}
 	\begin{center}
@@
 	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.
-\href{http://www.spss.com/}{SPSS}}
+\ac{SPSS}}
 	\begin{center}
@@
 	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
-\href{http://www.spss.com/}{SPSS}}\label{tab:8}
+\ac{SPSS}}\label{tab:8}
 	\begin{center}
@@
 \bibitem{Arai05}
-Arai K. and Tonooka H. 2005. \emph{Radiometric Performance Evaluation of ASTER VNIR, SWIR and TIR} 
+Arai K. and Tonooka H. 2005. \emph{Radiometric Performance Evaluation of \ac{ASTER} VNIR, \ac{SWIR} and \ac{TIR}} 
 IEEE Transactions on Geoscience and Remote Sensing, Vol. 43, No. 12, Dec. 2005. \pageref{Arai05}

### 2011-02-04 MScDiplom_Lemenkova.tex
added 'Index of People'

@@
 \usepackage[a4paper, left=30mm,right=25mm,top=30mm,bottom=35mm]{geometry}
+%\usepackage{glossary}
 \usepackage{graphics}
@@
 \usepackage{paralist}
-\usepackage{makeidx}
-\makeindex
+%\usepackage{makeidx}\makeindex
+%\usepackage[makeindex,protected]{splitidx}
+%\usepackage{index}
+\usepackage{imakeidx}
+\makeindex[name=concepts,title=Index of Concepts,intoc]
+\makeindex[name=people,title=Index of People,intoc,columns=3,columnseprule]
 \usepackage{fancyhdr}
@@
 
+%\newcommand{\listpeoplename}{}
+%\newlistof{people}{peo}{\listpeoplename}
+%\newcommand{\people}[1]{\refstepcounter{people}}
+
+%\newindex[List of People]{person}
+%\setindexpreamble[person]{...}
+%\newcommand{\person}[2]{\sindex[person]{#2, #1}#2}% "Given Surname" in text
+
+\phantomsection	
+\addcontentsline{toc}{section}{Front Page}
 \begin{titlepage}
@@
 	\author[1]{by Lemenkova Polina}
-	\affil[1]{\textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
+	\affil[1]{MSc student at the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
 	\date{\today}
@@
 \begin{itemize}
-	\item[*] University of Southampton (UK)\\
-	\item[*] Lund University (Sweden)\\
-	\item[*] University of Warsaw (Poland)\\
-	\item[*] University of Twente, Faculty ITC (The Netherlands)
+	\item[*] \href{http://www.southampton.ac.uk/}{University of Southampton}, \href{http://www.southampton.ac.uk/geography/}{School of Geography} (UK)\\
+	\item[*] \href{http://www.lunduniversity.lu.se/}{Lund University}, \href{http://www.giscentrum.lu.se/english/index.htm}{GIS Centre} (Sweden)\\
+	\item[*] \href{http://www.uw.edu.pl/en/}{University of Warsaw} (Poland)\\
+	\item[*] \href{http://www.utwente.nl/en}{University of Twente}, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)} (The Netherlands)
 \end{itemize}
@@
 \vspace{1em}
- {by Lemenkova Polina}
+by Lemenkova Polina\index[people]{Lemenkova, P.}
 \end{center}
@@
 	\item Name Examiner 2\dotfill
-	\item External Examiner \dotfill \\ \hfill{Dr.{\textit{Petter Pilesj\"o}}(Lund University, GIS Centre)}
+	\item External Examiner \dotfill \\ \hfill{Dr.{\textit{Petter Pilesj\"o}\index[people]{Pilesj\"o, P.}} (Lund University, GIS Centre)}
 \end{itemize}
@@
 \begin{itemize}
-	\item[* ] Mr. \textit{Valentijn Venus} \dotfill \\ (Department of Natural Resource, ITC)
-	\item[* ] Dr. \textit{Albertus G. Toxopeus} \dotfill \\ (Department of Natural Resource, ITC)
+	\item[* ] Mr. \textit{Valentijn Venus}\index[people]{Venus, V.} \dotfill \\ (Department of Natural Resource, ITC)
+	\item[* ] Dr. \textit{Albertus G. Toxopeus}\index[people]{Toxopeus, A.G.} \dotfill \\ (Department of Natural Resource, ITC)
 \end{itemize}
@@
 I thank my scientific supervisors, Dr.
-\textit{Albertus G.Toxopeus}  and Mr.
-\textit{Valentijn Venus} from the ITC, Department of Natural Resources, for our successful scientific collaboration, help and discussions of my research work.
+\textit{Albertus G.Toxopeus}\index[people]{Toxopeus, A.G.}  and Mr.
+\textit{Valentijn Venus}\index[people]{Venus, V.} from the ITC, Department of Natural Resources, for our successful scientific collaboration, help and discussions of my research work.
 \\
@@
 \\ Personally, I would like to mention that I have visited Greece the first time in my life, and I will always remember my Cretan period-2010 and Ligaria beach.\vspace{1mm}\\
-I extend my gratitude to the staff of the Museum of Natural History, the University of Crete, especially to Dr.\textit{Petros Limberakis} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g.
+I extend my gratitude to the staff of the Museum of Natural History, the University of Crete, especially to Dr.\textit{Petros Limberakis}\index[people]{Limberakis, Petros} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g.
 SCUBA gear diving, instruments for underwater measurements, etc).
@@
 \vspace{1mm}\\
-My sincere gratitude goes again to Mr.\textit{Valentijn Venus} for his generous providing me with a lot of spatial data   relevant to my research (e.g.
+My sincere gratitude goes again to Mr.\textit{Valentijn Venus}\index[people]{Venus, V.} for his generous providing me with a lot of spatial data   relevant to my research (e.g.
 such as digital vector layers of Crete island, raster georeferenced maps for the whole island for my GIS project, Excel spreadsheets, etc).
@@
 \\ My special thank to his personal, friendly support in various situations on Crete.\vspace{1mm}\\
-I am thankful to Dr.rer.nat.\textit{Peter Gege}, the software developer from the Earth Observation Centre of the German Space Agency, for his generous consultancies concerning the technical question about WASI that arose during my work.
+I am thankful to Dr.rer.nat.\textit{Peter Gege}\index[people]{Gege P.}, the software developer from the Earth Observation Centre of the German Space Agency, for his generous consultancies concerning the technical question about WASI that arose during my work.
 I am thankful for all his explanations and articles given to me via e-mail \vspace{1mm}\\
-My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throughout my career.
+My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}\index[people]{Lemenkov, P.}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throughout my career.
 He encouraged me to use various very useful software and tools, always gave me necessary recommendations and precious help, thus highly contributing to my professional growth.
@@
 \vspace{1mm}\\
-I am grateful to Mr.\textit{Willem Nieuwenhuis}, the software developer from the ITC, Department of Natural Resources, for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery.
+I am grateful to Mr.\textit{Willem Nieuwenhuis}\index[people]{Nieuwenhuis, W.}, the software developer from the ITC, Department of Natural Resources, for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery.
 \vspace{1mm}\\
@@
 \textit{Cornelis A.J.M.
-de Bie} from the Department of Natural Resource, ITC, for his patient sharing of iPAQ knowledge during my Polish fieldwork in summer 2010.
+de Bie}\index[people]{de Bie C.A.J.M.} from the Department of Natural Resource, ITC, for his patient sharing of iPAQ knowledge during my Polish fieldwork in summer 2010.
 I used this skills on Crete.
@@
 Many thanks for Dr.
-\author{\textit{Dingtian Yang}} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.
+\textit{Dingtian Yang}\index[people]{Yang D.} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.
 
@@
 	\markright{List of Tools}
-	\addcontentsline{toc}{section}{List of Tools}\index{Software used}
+	\addcontentsline{toc}{section}{List of Tools}\index[concepts]{Software used}
 
@@
 (\textit{Liliaceae}) than to the true grass, despite their name \textit{seagrass}, caused by the ribbon-like, grassy leaves
-\cite{McKenzie08}\label{McKenzie08}. Seagrasses\index{Seagrasses} create unique, complex, extremely diversified and productive ecological systems in the
+\cite{McKenzie08}\label{McKenzie08}. Seagrasses\index[concepts]{Seagrasses} create unique, complex, extremely diversified and productive ecological systems in the
 littoral coastal zones between 0-50 meters in shallow waters all over the world \cite{Hogarth07}\label{Hogarth07}, and
@@
 \subsubsection{Global distribution of the seagrasses}
-Globally, there are 58 recognized and described seagrass species\index{Seagrasses!global species} (Fig. \ref{fig:2}), belonging to two orders
+Globally, there are 58 recognized and described seagrass species\index[concepts]{Seagrasses!global species} (Fig. \ref{fig:2}), belonging to two orders
 (\textit{Hydrocharitales} and \textit{Najadales}), four families (\textit{Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
@@
 The distribution of the seagrasses is strongly influenced by several environmental factors, which
-include climate (mostly, tropical and temperate areas\index{Seagrasses!distribution of}), bathymetry (shallow shelf zones), hydrological
+include climate (mostly, tropical and temperate areas\index[concepts]{Seagrasses!distribution of}), bathymetry (shallow shelf zones), hydrological
 particularities (chemical content of water, nutrient availability and turbidity of waves), and geological 
 characteristics - sedimentation and cover types of the seafloor \cite{McKenzie06}\label{McKenzie06}.\\
-There are four European seagrass species \index{Seagrasses!Mediterranean species} in Mediterranean area \cite{Borum04}\label{Borum04}: \textit{Zostera marina,
+There are four European seagrass species \index[concepts]{Seagrasses!Mediterranean species} in Mediterranean area \cite{Borum04}\label{Borum04}: \textit{Zostera marina,
 Zostera noltii, Cymodocea nodosa} and \textit{P.oceanica}. In Greece the common species are \textit{P.oceanica}
@@
 \subsubsection{Environmental vulnerability of the seagrasses}
-Meadows of \textit{P. oceanica} are subjected to the human activities\index{Seagrasses!ecological vulnerability}, as they occur in coastal areas, where
+Meadows of \textit{P. oceanica} are subjected to the human activities\index[concepts]{Seagrasses!ecological vulnerability}, as they occur in coastal areas, where
 they can be affected both directly \cite{Meinesz91}\label{Meinesz91} or indirectly, through the impact on the
@@
 affecting both growth and status of the seagrasses the environmental contaminants can be mentioned,
-e.g. thermal, sewage, dredging and chemical pollution\index{Seagrasses!anthropogenic impacts on}  as well as any other kind of maritime works,
+e.g. thermal, sewage, dredging and chemical pollution\index[concepts]{Seagrasses!anthropogenic impacts on}  as well as any other kind of maritime works,
 e.g. trawling and anchoring of boats \cite{Ribed02}\label{Ribed02}. Other human activities that cause degrading of the
@@
 \cite{McKenzie06}\label{McKenzie06}.\\
-Besides anthropogenic factors, various biochemical, climatic and environmental processes\index{Seagrasses!natural impacts on} can cause
+Besides anthropogenic factors, various biochemical, climatic and environmental processes\index[concepts]{Seagrasses!natural impacts on} can cause
 negative influence on seagrass distribution. Seagrass is exposed to threats from the global climate and
@@
 
-\subsubsection{General characteristics of \textit{P.oceanica}}\index{Seagrasses!\textit{P.oceanica}}
+\subsubsection{General characteristics of \textit{P.oceanica}}\index[concepts]{Seagrasses!\textit{P.oceanica}}
 The endemic Mediterranean seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}), is a main species in
@@
 The dominant and most productive coastal ecosystem of the Mediterranean, \textit{P.oceanica} is
-spatially restricted to the Mediterranean area\index{Seagrasses!\textit{P.oceanica}!distribution} (Fig.\ref{fig:5}), 
+spatially restricted to the Mediterranean area\index[concepts]{Seagrasses!\textit{P.oceanica}!distribution} (Fig.\ref{fig:5}), 
 with its extension limited by the western part of the Mediterranean Sea where cold Atlantic waters enter Gibraltar and mix with warm
 Mediterranean waters, thus decreasing its temperature.
-Morphologically\index{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots and short rhizomes (0.5-2.0 mm). 
+Morphologically\index[concepts]{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots and short rhizomes (0.5-2.0 mm).
 The leaves are, perhaps, the most particular characteristics of \textit{P.oceanica}, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:8}(a)): having usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}(c)).
@@
 There are several environmental factors, determining the growth of the \textit{P.oceanica}. 
-The adaptation\index{Seagrasses!\textit{P.oceanica}!environmental adaptation}
+The adaptation\index[concepts]{Seagrasses!\textit{P.oceanica}!environmental adaptation}
 to dry-summer subtropical climate reduces its extension to Mediterranean area only (Fig.\ref{fig:5}).
@@
 the different maximum summer temperatures at those depths \cite{Buia91}\label{Buia91}.
-The phenology\index{Seagrasses!\textit{P.oceanica}!phenology} of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in shallow waters plants have shorter and falciform leaves, compared to ones in the deeper and central
+The phenology\index[concepts]{Seagrasses!\textit{P.oceanica}!phenology} of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in shallow waters plants have shorter and falciform leaves, compared to ones in the deeper and central
 areas \cite{Dural10}\label{Dural10}. In \textit{P.oceanica} flower abundance is related to the structure of the meadow with
@@
 
-\subsection{Research problem}\index{Research}
-Monitoring\index{Research!problem} of the marine benthic ecosystems of seagrasses is essential for the environmental
+\subsection{Research problem}\index[concepts]{Research}
+Monitoring\index[concepts]{Research!problem} of the marine benthic ecosystems of seagrasses is essential for the environmental
 assessment of the coastal zones. It increases our knowledge of the seagrass ecology, highlights 
@@
 
-\subsection{Research objective}\index{Research!objective}
+\subsection{Research objective}\index[concepts]{Research!objective}
 The current MSc research aims to explore the environmental conditions for the spatial distribution of \textit{P.oceanica} seagrass along the northern coast of Crete Island, based on the remote sensing and \ac{GIS} techniques, knowledge about the coastal environment in Crete and integration of various data
-from the following sources \index{Data}:
+from the following sources \index[concepts]{Data}:
 	\begin{itemize}
@@
 
-\subsubsection{General objective}\index{Research!objective!general}
+\subsubsection{General objective}\index[concepts]{Research!objective!general}
 The main objective of this study is to analyse the optical properties of the seagrass \textit{P.oceanica} and other seafloor types, and to apply the remote sensing techniques towards the investigation of the seagrass distribution in the selected locations along the northern coast of Crete.
@@
 
-\subsubsection{Specific objectives}\index{Research!objective!specific}
+\subsubsection{Specific objectives}\index[concepts]{Research!objective!specific}
 	\begin{itemize}	
@@
 
-\subsection{Research questions}\index{Research!questions}
+\subsection{Research questions}\index[concepts]{Research!questions}
 	\begin{itemize}	
@@
 
-\subsection{Hypotheses testing}\index{Research!questions!hypothesis}
+\subsection{Hypotheses testing}\index[concepts]{Research!questions!hypothesis}
 A statistical testing will be used to compare between the spectral responses of the different seafloor cover types (i.e. sand and \textit{P. oceanica}),
@@
 
-\subsection{Research approach}\index{Research!approach}
+\subsection{Research approach}\index[concepts]{Research!approach}
 Seagrass consistent monitoring and mapping is necessary and important for the sustainable coastal
@@
 
-\subsection{Seagrass global monitoring: history and perspectives}\index{Seagrasses!global monitoring}
+\subsection{Seagrass global monitoring: history and perspectives}\index[concepts]{Seagrasses!global monitoring}
 Mapping and monitoring the seagrass is important for the environmental assessment of the marine
@@
 growing on the seafloor.\\
-Several global seagrass surveys\index{Seagrasses!global surveys} organize and provide regular monitoring of the seagrass distribution,
+Several global seagrass surveys\index[concepts]{Seagrasses!global surveys} organize and provide regular monitoring of the seagrass distribution,
 health and species sustainability, amongst which the most world-known are, perhaps, the following
@@
 
-The optical properties\index{Seawater optical properties} of the sea water vary with different environmental conditions and reflect
+The optical properties\index[concepts]{Seawater optical properties} of the sea water vary with different environmental conditions and reflect
 current chemical content and physical specifics of the water, revealing the variability and distribution
@@
 properties incessantly. The effects of these processes influence application of the hyperspectral
-remote sensing\index{Seawater optical properties!hyperspectral measurements} and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
+remote sensing\index[concepts]{Seawater optical properties!hyperspectral measurements} and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
 sedimentation processes, which in turn, may change micro relief properties and topology. \\
@@
 characteristics, and is expressed in percentage.
-The irradiance\index{Seawater optical properties!irradiance} and radiance\index{Seawater optical properties!radiance} of the water thus should be measured, in order to estimate spectral
+The irradiance\index[concepts]{Seawater optical properties!irradiance} and radiance\index[concepts]{Seawater optical properties!radiance} of the water thus should be measured, in order to estimate spectral
 reflectance of the various seafloor cover types. \\
 The optical measurements of the irradiance and radiance of the sea water and bottom cover types of the seafloor can be received by the means of the
-optical sensors spectroradiometers\index{Seawater optical properties!spectroradiometers}. There are several companies producing radiometers with various
+optical sensors spectroradiometers\index[concepts]{Seawater optical properties!spectroradiometers}. There are several companies producing radiometers with various
 characteristics and adjusted for different purposes, e.g. portable and miniature spectrometers from
@@
 optical sensors, GER-Series Field portable spectroradiometers from
-\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. \ac{Trios-RAMSES}\index{Seawater optical properties!spectroradiometers!Trios-RAMSES} hyperspectral radiometers
+\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. \ac{Trios-RAMSES}\index[concepts]{Seawater optical properties!spectroradiometers!Trios-RAMSES} hyperspectral radiometers
 (\ac{Trios-RAMSES} Hyperspectral \ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor) are small-sized, low power-consuming, flexible for fieldwork yet with high
@@
 
-Understanding radiance distribution within a water colum\index{Seawater optical properties!RTM}n is necessary for the studies of the underwater visibility, because scattering properties of the water body naturally vary with changing depth, wavelength and environmental conditions.
+Understanding radiance distribution within a water colum\index[concepts]{Seawater optical properties!RTM}n is necessary for the studies of the underwater visibility, because scattering properties of the water body naturally vary with changing depth, wavelength and environmental conditions.
 Simulation of the radiance quantities for natural water bodies enables to analyse seafloor color remote sensing properties.
@@
 	\begin{itemize}
-		\item [*] The \ac{6SV1} \index{Seawater optical properties!RTM!6SV1} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \ac{MODIS} satellite imagery lookup tables.
+		\item [*] The \ac{6SV1} \index[concepts]{Seawater optical properties!RTM!6SV1} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \ac{MODIS} satellite imagery lookup tables.
 The 6S algorithm has been \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{implemented} by the \ac{GRASS} software for the atmospheric correction
-		\item [*] Another example of the RTM adjusted for a specific imagery is a German \ac{KOPRA}\index{Seawater optical properties!RTM!KOPRA}, fitted for \ac{MIPAS} and \ac{ENVISAT} imagery.
+		\item [*] Another example of the RTM adjusted for a specific imagery is a German \ac{KOPRA}\index[concepts]{Seawater optical properties!RTM!KOPRA}, fitted for \ac{MIPAS} and \ac{ENVISAT} imagery.
 		\item [*]  The \href{http://www.sequoiasci.com/products/Hydrolight.aspx}{\textit{HydroLight RTM}} commercial software is an advanced model for oceanic radiative transfer calculations, developed by the Sequoia Scientific, Inc., USA.
 It is designed to solve a wide range of problems in optical oceanography and limnology.
-		\item [*] The British \ac{RTTOV-9}\index{Seawater optical properties!RTM!RTTOV-9} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers.
+		\item [*] The British \ac{RTTOV-9}\index[concepts]{Seawater optical properties!RTM!RTTOV-9} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers.
 It includes a number of useful tolls, e.g.
@@
 This program runs under Unix/Linux platforms and is open source.
-		\item [*] The \ac{RRTM}\index{Seawater optical properties!RTM!RRTM} is one more RTM freeware, designed in the USA, Massachusetts. 
-		\item [*] The \href{http://www.modtran.org/}{\textit{MODTRAN}}\index{Seawater optical properties!RTM!MODTRAN}, a commercial software from Spectral Sciences Inc., is another example of the RTM.
+		\item [*] The \ac{RRTM}\index[concepts]{Seawater optical properties!RTM!RRTM} is one more RTM freeware, designed in the USA, Massachusetts. 
+		\item [*] The \href{http://www.modtran.org/}{\textit{MODTRAN}}\index[concepts]{Seawater optical properties!RTM!MODTRAN}, a commercial software from Spectral Sciences Inc., is another example of the RTM.
 Among its newest updates is generation of atmospheric correction data; it also has an option to write spectral output in binary, and convert it to \ac{ASCII}.
-		\item [*] A \ac{DISORT}\index{Seawater optical properties!RTM!DISORT}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
-		\item [*] The \ac{CRTM} \index{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA}, \ac{NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
-		\item [*] An \ac{4AOP}\index{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{the French Space Agency} (\ac{CNES}).
+		\item [*] A \ac{DISORT}\index[concepts]{Seawater optical properties!RTM!DISORT}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
+		\item [*] The \ac{CRTM} \index[concepts]{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA}, \ac{NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
+		\item [*] An \ac{4AOP}\index[concepts]{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{the French Space Agency} (\ac{CNES}).
 Being a fine program which enables accurate computations of the radiance, defining other emissivity functions and other set up, it is however, particularly designed for the infrared waves (thus, not quite suitable in our case).
@@
 	\end{itemize}
-For our purpose we have chosen the \ac{WASI}\index{Seawater optical properties!RTM!WASI} \ac{RTM} software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
+For our purpose we have chosen the \ac{WASI}\index[concepts]{Seawater optical properties!RTM!WASI} \ac{RTM} software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.
 
@@
 
-\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}\index{Remote sensing}
+\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
@@
 unsuitability for the case of monitoring large areas of water for small-scale mapping. 
-The underwater videography\index{Remote sensing!underwater videography} with a \ac{GPS}\index{Remote sensing!GPS} is a tool of seagrass monitoring which has certain advantages, i.e.
+The underwater videography\index[concepts]{Remote sensing!underwater videography} with a \ac{GPS}\index[concepts]{Remote sensing!GPS} is a tool of seagrass monitoring which has certain advantages, i.e.
 high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
@@
 measurements of hyperspectral reflectance. 
-Spectral measurements\index{Remote sensing!spectral measurements} of the target objects are made
+Spectral measurements\index[concepts]{Remote sensing!spectral measurements} of the target objects are made
 by means of the radiometers, which receive and register the amounts of energy (radiance and
@@
 using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01}\label{Louchard03}\label{Dehouck08}\label{Werdell03}.
-Studies of spectral reflectance\index{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and
+Studies of spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and
 other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
@@
 
-Spectral reflectance\index{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
+Spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
 depth where it is located, and is generally decreasing in values by increasing depths.
-The most important spectral diapason\index{Seawater optical properties!spectral reflectance!spectral diapason} for marine mapping of submerged vegetation
+The most important spectral diapason\index[concepts]{Seawater optical properties!spectral reflectance!spectral diapason} for marine mapping of submerged vegetation
 and, particularly, for seagrass, is 350-800 nm \cite{Barille09}\label{Barille09}.
@@
 
-\subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}\index{Remote sensing!multispectral imagery} 
+\subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}\index[concepts]{Remote sensing!multispectral imagery} 
 Seagrass mapping using remotely sensed data from multispectral sensors is based on the classification
@@
 
-The \ac{Landsat TM}\index{Remote sensing!multispectral imagery!Landsat TM} and \ac{Landsat ETM+}\index{Remote sensing!multispectral imagery!Landsat ETM+} data, with the most recent from sensor Landsat 7 - an advanced
+The \ac{Landsat TM}\index[concepts]{Remote sensing!multispectral imagery!Landsat TM} and \ac{Landsat ETM+}\index[concepts]{Remote sensing!multispectral imagery!Landsat ETM+} data, with the most recent from sensor Landsat 7 - an advanced
 and multispectral scanning, launched in 1999, prove to be feasible and useful for the mapping of
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
-Another well-known multispectral sensor, \ac{SPOT}\index{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial
+Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial
 resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
@@
 Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}. 
-The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}\index{Remote sensing!multispectral imagery!IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
+The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}\index[concepts]{Remote sensing!multispectral imagery!IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
 available high-resolution imagery at 1- and 4-meter resolution from \href{http://www.geoeye.com}{Geoeye}. IKONOS
@@
 Thus, the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
-made using IKONOS, \ac{Landsat TM}, and \ac{CASI}\index{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
+made using IKONOS, \ac{Landsat TM}, and \ac{CASI}\index{concepts}{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
 spectrum, the best results are achieved by the IKONOS and \ac{CASI}, while \ac{Landsat TM} has not high
@@
  demonstrates high suitability of \ac{CASI} images for the fine-scale mapping of the seagrass
-landscapes. Thus, \ac{CASI} and Quickbird-2\index{Remote sensing!multispectral imagery!Quickbird} images enable to identify even separate seagrass species
+landscapes. Thus, \ac{CASI} and Quickbird-2\index[concepts]{Remote sensing!multispectral imagery!Quickbird} images enable to identify even separate seagrass species
 with small width and heterogeneous nature of the seagrass patches, which could not be detected using
 Landsat TM images with their 30*30m resolution.
-\ac{ASTER}\index{Remote sensing!multispectral imagery!ASTER}), launched in
-1999 on board Terra sensor\index{Remote sensing!multispectral imagery!Terra sensor}, provides high-resolution images of the Earth in 15 different bands of the
+\ac{ASTER}\index[concepts]{Remote sensing!multispectral imagery!ASTER}), launched in
+1999 on board Terra sensor\index[concepts]{Remote sensing!multispectral imagery!Terra sensor}, provides high-resolution images of the Earth in 15 different bands of the
 electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:2.4}). It has diverse
@@
 
-\subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}\index{Remote sensing!hyperspectral imagery}
+\subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery}
 The application of the hyperspectral sensors is most effective and provide more accurate classification
@@
 While comparing multispectral imagery with airborne hyperspectral, the last one showed higher
-overall accuracies \cite{Phinn08}\label{Phinn08}. Several, regular and narrow (10 nm) spectral bands\label{sec:2.4.2}\index{Remote sensing!hyperspectral imagery!spectral bands}, specific for
+overall accuracies \cite{Phinn08}\label{Phinn08}. Several, regular and narrow (10 nm) spectral bands\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery!spectral bands}, specific for
 hyperspectral imagery, are strong tool enabling to detect even slight and subtle differences in the
@@
 approaches requires application of high-spatial resolution (higher than 5 m) or hyperspectral imagery.
-Comparative analysis of the application of hyperspectral\label{sec:2.4.2}\index{Remote sensing!hyperspectral imagery!application} and multispectral imagery towards the
+Comparative analysis of the application of hyperspectral\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery!application} and multispectral imagery towards the
 seafloor types classification \cite{Hochberg03a}\label{Hochberg03a} demonstrates that coral, seagrasses and
@@
 
-\subsection{Fieldwork data collection}\index{Data!collection}
+\subsection{Fieldwork data collection}\index[concepts]{Data!collection}
 Seagrass sampling has been performed at two research stations at Crete Island - Ligaria beach
@@
 
-\subsection{Data pre-processing}\index{Data!pre-processing}
+\subsection{Data pre-processing}\index[concepts]{Data!pre-processing}
 A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of \textit{P.oceanica} made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
@@
 
-\subsubsection{Water color spectral simulation using WASI}\index{Data!processing}
+\subsubsection{Water color spectral simulation using WASI}\index[concepts]{Data!processing}
 The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
@@
 based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
-types, where the spectral signatures were examined. WASI software\index{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate
+types, where the spectral signatures were examined. WASI software\index[concepts]{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate
 spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
@@
 
-An analysis of the spectral reflectance of \textit{P.oceanica} is done using the WASI simulations\index{Seawater optical properties!RTM!WASI} in order to
+An analysis of the spectral reflectance of \textit{P.oceanica} is done using the WASI simulations\index[concepts]{Seawater optical properties!RTM!WASI} in order to
 determine, which wavebands can be still used to identify \textit{P.oceanica}. The analysis of spectra shows
@@
 
-\subsubsection{Statistical analysis of the observational data and hypothesis testing}\index{Data!statistical analysis}
+\subsubsection{Statistical analysis of the observational data and hypothesis testing}\index[concepts]{Data!statistical analysis}
 The total amount of measured data was large and included following datasets made using
@@
 spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
-The statistical calculators were mainly made by means of Gretle\index{Data!statistical analysis!Gretl} and SPSS\index{Data!statistical analysis!SPSS}. 
+The statistical calculators were mainly made by means of Gretle\index[concepts]{Data!statistical analysis!Gretl} and SPSS\index[concepts]{Data!statistical analysis!SPSS}. 
 The Open Office was used for preliminary data view and preprocessing, and included following computations
@@
 seawater with and without sediment (e.g. Fig..\ref{fig:A.16}).
-In the statistical analysis of the raw observed data pool\index{Data!measurements pool} of the values of spectral reflectance (about 400 single measurements for \textit{P.oceanica}, e.g.
+In the statistical analysis of the raw observed data pool\index[concepts]{Data!measurements pool} of the values of spectral reflectance (about 400 single measurements for \textit{P.oceanica}, e.g.
 Tab.\ref{tab:A.8}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e.
@@
 There is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
-Each time we observe the outcome of a random variable\index{Data!statistical analysis!variables} of spectral reflectance, we obtain an observation, which is hence no longer random. 
+Each time we observe the outcome of a random variable\index[concepts]{Data!statistical analysis!variables} of spectral reflectance, we obtain an observation, which is hence no longer random. 
 We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the \nameref{app}: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyze the distribution of values of spectral reflectance of \textit{P.oceanica}.
-A probability distribution\index{Data!statistical analysis!probability distribution}, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
-The \textit{least absolute deviations (LAD)}\index{Data!statistical analysis!LAD}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
-Non-linear logistic analysis\index{Data!statistical analysis!non-linear logistic analysis} is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
+A probability distribution\index[concepts]{Data!statistical analysis!probability distribution}, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
+The \textit{least absolute deviations (LAD)}\index[concepts]{Data!statistical analysis!LAD}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
+Non-linear logistic analysis\index[concepts]{Data!statistical analysis!non-linear logistic analysis} is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
 the value of spectral reflectance for each single observation) is not continuous, but discrete and independent for each case (Tab.\ref{tab:11}).
@@
 As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand.
-The \textit{autocorrelation}\index{Data!statistical analysis!autocorrelation}, a cross-correlation\index{Data!statistical analysis!cross-correlation} of an observed value with itself, made on two different days (14th and 15th October), shows the similarity between observations as a function of the time separation between them.
+The \textit{autocorrelation}\index[concepts]{Data!statistical analysis!autocorrelation}, a cross-correlation\index[concepts]{Data!statistical analysis!cross-correlation} of an observed value with itself, made on two different days (14th and 15th October), shows the similarity between observations as a function of the time separation between them.
 As can be seen on Fig.\ref{fig:A.21}, Fig.\ref{fig:A.22} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
@@
 \\
-The method of \textit{least squares}\index{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made.
+The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made.
 We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}.\\
@@
 The summary of the values of spectral reflectance properties on various depths is presented on the Tab.\ref{tab:A.7}.
-The \textit{k-means cluster analysis}\index{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
+The \textit{k-means cluster analysis}\index[concepts]{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
 It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
-To estimate spectral density of a spectral reflectance measurements we used periodogram\index{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30}), 
+To estimate spectral density of a spectral reflectance measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30}), 
 which corresponds to the general spectrum of the observations with representation of a variable quantity.  
 Thus, the periodogram (Fig.
-\ref{fig:A.31}) of the series of observation\index{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to
+\ref{fig:A.31}) of the series of observation\index[concepts]{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to
 exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:A.12} ) 
@@
 The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
-areas of quartiles\index{Data!statistical analysis!quartiles Q1-Q3} (Q1, Q3, shaded areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution
+areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1, Q3, shaded areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution
 (spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
@@
 
-A statistical hypothesis\index{Data!statistical analysis!hypothesis testing} test has been applied for the making decision and controlling the wealth of the
+A statistical hypothesis\index[concepts]{Data!statistical analysis!hypothesis testing} test has been applied for the making decision and controlling the wealth of the
 observational data of the hyperspectral measurements of the water reflectance.
@@
 
-\paragraph{ANOVA-testing}\index{Data!statistical analysis!ANOVA}
+\paragraph{ANOVA-testing}\index[concepts]{Data!statistical analysis!ANOVA}
 We applied \ac{ANOVA} (ANalysis Of Variance) testing for the analysis of the probability of the
@@
 
-\paragraph{Student-t testing}\index{Data!statistical analysis!Student-t testing}
+\paragraph{Student-t testing}\index[concepts]{Data!statistical analysis!Student-t testing}
 The Student-t test, one of the most commonly used techniques for testing a hypothesis on the basis
@@
 where the abbreviations stand for the following values:
-Mean: Average of the data to be plotted, AVERAGE\index{Data!statistical analysis!average} {data}
-St Dev: Standard deviation\index{Data!statistical analysis!standard deviation}, \ac{STDEV} {data}
+Mean: Average of the data to be plotted, AVERAGE\index[concepts]{Data!statistical analysis!average} {data}
+St Dev: Standard deviation\index[concepts]{Data!statistical analysis!standard deviation}, \ac{STDEV} {data}
 Median: Median of the data, MEDIAN{data}
@@
 25th Pct: Plotting value of first quartile = Q1
-50th Pct: Plotting value of median = Median\index{Data!statistical analysis!median}-Q1
+50th Pct: Plotting value of median = Median\index[concepts]{Data!statistical analysis!median}-Q1
 75th Pct: Plotting value of third quartile = Median-Q3
@@
 The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
-types as the different sensors would percept them: MODIS\index{Remote sensing!multispectral imagery!MODIS}, ASTER\index{Remote sensing!multispectral imagery!ASTER}, MERIS\index{Remote sensing!multispectral imagery!MERIS}, SeaWiFS\index{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index{Remote sensing!multispectral imagery!CZCS}.\\
+types as the different sensors would percept them: MODIS\index[concepts]{Remote sensing!multispectral imagery!MODIS}, ASTER\index[concepts]{Remote sensing!multispectral imagery!ASTER}, MERIS\index[concepts]{Remote sensing!multispectral imagery!MERIS}, SeaWiFS\index[concepts]{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index[concepts]{Remote sensing!multispectral imagery!CZCS}.\\
 These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
@@
 detailed examination of the imagery. \\
-Seagrass beds are clearly visible in color aerial Google Earth\index{Remote sensing!aerial imagery!GoogleEarth}
+Seagrass beds are clearly visible in color aerial Google Earth\index[concepts]{Remote sensing!aerial imagery!GoogleEarth}
 photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
@@
 
-\subsubsection{Data integration}\index{Data!integration}
+\subsubsection{Data integration}\index[concepts]{Data!integration}
 The integrated approach used in this research work has high potential as a means to monitor changes
@@
 high resolution aerial color Google Earth photography, spaceborne satellite imagery, assessment of
-spectral signatures using WASI software\index{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
+spectral signatures using WASI software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
 based mapping. The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
@@
 
-\subsubsection{Upscale mapping of the seagrass landscapes}\index{Seagrasses!spatial structure}
+\subsubsection{Upscale mapping of the seagrass landscapes}\index[concepts]{concepts}{Seagrasses!spatial structure}
 Although ocean pelagic landscapes have a high degree of spatial variance and less structurally
@@
 \bibliographystyle{plainnat}
-\pagenumbering{roman}
+\pagenumbering{alph}
 \markright{Bibliography}
 
-\bibitem{Ackleson2003}Ackleson S. 2003. \emph{Light in shallow waters: A brief research review}. Limnol. Oceanogr.,
+\bibitem{Ackleson2003}Ackleson S.\index[people]{} 2003. \emph{Light in shallow waters: A brief research review}. Limnol. Oceanogr.,
 48(1, part 2), 2003, 323–328. American Society of Limnology and Oceanography, Inc. \pageref{Ackleson, 2003}
 \bibitem{Alcoverro1997} %
-Alcoverro T., Romero J., Duarte C.M., Lopez N. 1997. \emph{Spatial and temporal variations
+Alcoverro T.\index[people]{Alcoverro, T.}, Romero J.\index[people]{Romero J.}, Duarte C.M.\index[people]{Duarte C.M.}, Lopez N.\index[people]{Lopez, N.} 1997. \emph{Spatial and temporal variations
 in nutrient limitation of seagrass \textit{Posidonia oceanica} growth in the NW Mediterranean}.
 Marine ecology Progress Series.Vol.146: 155-161. \pageref{}
-\bibitem{Amoutzopoulou-Schina05}Amoutzopoulou-Schina H and S.Haritonidis. 2005. \emph{Distribution and phenology of the
-marine phanerogam \textit{Posidonia oceanica} in the Pagassitikos Gulf, Greece}. Journal of
+\bibitem{Amoutzopoulou-Schina05}Amoutzopoulou-Schina H.\index[people]{Amoutzopoulou-Schina, H.} and Haritonidis S.\index[people]{Haritonidis, S.} 2005.
+\emph{Distribution and phenology of the marine phanerogam \textit{Posidonia oceanica} in the Pagassitikos Gulf, Greece}.
+Journal of
 Biological Research 4: 203 – 211, 2005 J. \pageref{Amoutzopoulou-Schina05}
 \bibitem{Arai05}
-Arai K. and Tonooka H. 2005. \emph{Radiometric Performance Evaluation of \ac{ASTER} VNIR, \ac{SWIR} and \ac{TIR}} 
+Arai K.\index[people]{Arai, K.} and Tonooka H.\index[people]{Tonooka, H.} 2005. \emph{Radiometric Performance Evaluation of \ac{ASTER} VNIR, \ac{SWIR} and \ac{TIR}} 
 IEEE Transactions on Geoscience and Remote Sensing, Vol. 43, No. 12, Dec. 2005. \pageref{Arai05}
 \bibitem{Ardizzone06}
-Ardizzone G., Belluscio A. and Maiorano L. 2006. \emph{Long-term change in the structure of
+Ardizzone G.D.\index[people]{Ardizzone G.D.}, Belluscio A.\index[people]{Belluscio A.} and Maiorano L.\index[people]{Maiorano, L.} 2006. \emph{Long-term change in the structure of
 a \textit{Posidonia oceanica} landscape and its reference for a monitoring plan}. Marine Ecology.
@@
 \bibitem{Ball09}
-Ball D., Soto-Berelov M., Young P. and Coots A. 2009. \emph{Baywide Seagrass Monitoring
-Program - Historical Seagrass Mapping}. Fisheries Victoria Technical Report Series No. 70. \pageref{Ball09}
+Ball D.\index[people]{Ball, D.}, Soto-Berelov M.\index[people]{Soto-Berelov, M.}, Young P.\index[people]{Young, P.} and Coots A.\index[people]{Coots, A.} 2009.
+\emph{Baywide Seagrass Monitoring Program - Historical Seagrass Mapping}.
+Fisheries Victoria Technical Report Series No.
+70.
+\pageref{Ball09}
 \bibitem{Barille09}
-Barille L., Robin M., Harin N., Bargain A., Patrick L. 2009. \emph{Increase in seagrass
-distribution at Bourgneuf Bay (France) detected by spatial remote sensing}. Aquatic Botany. \pageref{Barille09}
+Barille L.\index[people]{Barille, L.}, Robin M.\index[people]{Robin, M.}, Harin N.\index[people]{Harin, N.}, Bargain A.\index[people]{Bargain, A.}, Patrick L.\index[people]{Patrick, L.} 2009.
+\emph{Increase in seagrass distribution at Bourgneuf Bay (France) detected by spatial remote sensing}.
+Aquatic Botany.
+\pageref{Barille09}
 \bibitem{Bakran-Petricioli06} %
-Bakran-Petricioli T., Antonic O., Bukovec D., Petricioli D., Janekovic I., Krizan J.,
-Kusan V. and Dujmovic S. 2006. \emph{Modelling spatial distribution of the Croatian marine
+Bakran-Petricioli T.\index[people]{Bakran-Petricioli, T.}, Antonic O.\index[people]{Antonic, O.}, Bukovec D.\index[people]{Bukovec, D.}, Petricioli D.\index[people]{Petricioli, D.}, Janekovic I.\index[people]{Janekovic, I.}, Krizan J.\index[people]{Krizan, J.}, Kusan V.\index[people]{Kusan, V.} and Dujmovic S.\index[people]{Dujmovic, S.} 2006.
+\emph{Modelling spatial distribution of the Croatian marine
 benthic habitats}. Ecological Modelling 191, 96–105. \pageref{Bakran-Petricioli06}
 \bibitem{Bell99}
-Bell S.S., Robbins B.D. and Jensen S.L. 1999. \emph{Gap Dynamics in a Seagrass Landscape}.
+Bell S.S.\index[people]{Bell S.S.}, Robbins B.D.\index[people]{Robbins B.D.} and Jensen S.L.\index[people]{Jensen S.L.} 1999. \emph{Gap Dynamics in a Seagrass Landscape}.
 Springer-Verlag. Ecosystems (1999) 2: 493–504. \pageref{Bell99}
 \bibitem{Bierwirth93}
-Bierwirth P. N., Lee T. J. and Burne R. V. 1993. \emph{Shalow sea-floor reflectance and water
+Bierwirth P. N.\index[people]{Bierwirth, P. N.}, Lee T. J.\index[people]{Lee, T. J.} and Burne R. V.\index[people]{Burne, R.V.} 1993. \emph{Shalow sea-floor reflectance and water
 depth derived by unmixing multispectral imagery}. Photogrammetric Engineering and Remote
@@
 \bibitem{Blake00}
-Blake S., Roob R. and Patterson E. 2000. \emph{Seagrass mapping of Victoria's Minor Inlets}.
+Blake S.\index[people]{Blake, S.}, Roob R.\index[people]{Roob, R.} and Patterson E.\index[people]{Patterson, E.} 2000. \emph{Seagrass mapping of Victoria's Minor Inlets}.
 Marine and Freshwater Resources Institute Report No. 28. \pageref{Blake00}
 \bibitem{Borum04}
-Borum J., Duarte C.M., Krause-Jensen D. and Greve T.M. 2004. \emph{European seagrasses:
-an introduction to monitoring and management}. The M and MS project. ISBN: 87-89143-21-3.
+Borum J.\index[people]{Borum, J.}, Duarte C.M.\index[people]{Duarte C.M.}, Krause-Jensen D.\index[people]{Krause-Jensen, D.} and Greve T.M.\index[people]{Greve, T.M} 2004.
+\emph{European seagrasses: an introduction to monitoring and management}.
+The M and MS project.
+ISBN: 87-89143-21-3.
 Internet version: Available at \href{http://www.seagrasses.org}{Seagrasses.org} \pageref{Borum04}
 \bibitem{Brown02}
-Brown C.J., Cooper K.M., Meadows W.J., Limpenny D.S., Rees H.L. 2002. \emph{Small scale
-mapping of sea bed assemblages in the eastern English Channel using sonar and remote
+Brown C.J.\index[people]{Brown, C.J.}, Cooper K.M.\index[people]{Cooper, K.M.}, Meadows W.J.\index[people]{Meadows, W.J.}, Limpenny D.S.\index[people]{Limpenny, D.S.}, Rees H.L.\index[people]{Rees, H.L.} 2002.
+\emph{Small scale mapping of sea bed assemblages in the eastern English Channel using sonar and remote
 sampling techniques}. Estuarine Coastal and Shelf Science, 54, 263-278. \pageref{Brown02}
 \bibitem{Boudouresque89}
-Boudouresque, C.F., Meinesz, A., Fresi, E. and Gravez, V., 1989. \emph{2nd International
-Workshop on \textit{Posidonia oceanica} Beds}. G.I.S. Posidonie, Oceanica, 2, 322 pp. \pageref{Boudouresque89}
+Boudouresque C.F.\index[people]{Boudouresque C.F.}, Meinesz, A.\index[people]{Meinesz A.}, Fresi, E.\index[people]{Fresi, E.} and Gravez, V.\index[people]{Gravez, V}, 1989.
+\emph{2nd International Workshop on \textit{Posidonia oceanica} Beds}.
+G.I.S.
+Posidonie, Oceanica, 2, 322 pp.
+\pageref{Boudouresque89}
 \bibitem{Bharathi03}
-Bharathi A.M, Kumar S.C, Kalliappan S. 2003. \emph{Target Identification in multispectral images
-using \textit{in-situ} hyperspectral reflectance}. The Geospatial Resource Portal. Remote Sensing.
+Bharathi A.M.\index[people]{Bharathi A.M.}, Kumar S.C.\index[people]{Kumar S.C.}, Kalliappan S.\index[people]{Kalliappan S.} 2003.
+\emph{Target Identification in multispectral images using \textit{in-situ} hyperspectral reflectance}.
+The Geospatial Resource Portal.
+Remote Sensing.
 \href{www.gisdevelopment.net}{gisdevelopment} \pageref{Bharathi03}
 \bibitem{Buia91}
-Buia M.C. and Mazzella L. 1991. \emph{Reproductive phenology of the Mediterranean
+Buia M.C.\index[people]{Buia M.C.} and Mazzella L.\index[people]{Mazzella L.} 1991. \emph{Reproductive phenology of the Mediterranean
 seagrasses \textit{Posidonia oceanica} (L.) Delile, Cymodocea nodosa ( Ucria ) Aschers., and
@@
 \bibitem{Butler87}
-Butler 1987. \emph{Marine resource mapping: an introductory manual}. \pageref{Butler87}
+Butler\index[people]{Butler} 1987. \emph{Marine resource mapping: an introductory manual}. \pageref{Butler87}
 \bibitem{Cappo95}
-Cappo, M., Alongi, D.M., Williams, D. and Duke, N. 1995. \emph{A Review and Synthesis of
-Australian Fisheries Habitat Research: Major threats, issues and gaps in knowledge of
+Cappo M.\index[people]{Cappo, M.}, Alongi D.M.\index[people]{Alongi D.M.}, Williams D.\index[people]{Williams D.} and Duke N.\index[people]{Duke N.} 1995.
+\emph{A Review and Synthesis of Australian Fisheries Habitat Research: Major threats, issues and gaps in knowledge of
 coastal and marine fisheries habitats}. Fisheries Research and Development Corporation. \pageref{Cappo95}
 \bibitem{Carter08}
-Carter G.A., Criss G.A., Blossom G.A. and Biber P.D. 2008. \emph{Decadal-Scale Changes in
-Coverage on the Mississippi Barrier Islands, Northern Gulf of Mexico}.
+Carter G.A.\index[people]{Carter G.A.}, Criss G.A.\index[people]{Criss G.A.}, Blossom G.A.\index[people]{Blossom G.A.} and Biber P.D.\index[people]{Biber P.D.} 2008.
+\emph{Decadal-Scale Changes in Coverage on the Mississippi Barrier Islands, Northern Gulf of Mexico}.
 \href{http://precedings.nature.com/documents/3669/version/1/files/npre20093669-1.pdf}{precedings.nature.com} \pageref{Carter08}
 \bibitem{Calvo03}
-Calvo S., Ciraolo G. and Loggia G. la. 2003. \emph{Monitoring \textit{Posidonia oceanica} meadows
-in a Mediterranean coastal lagoon (Stagnone, Italy) by means of neural network and
+Calvo S.\index[people]{Calvo S.}, Ciraolo G.\index[people]{Ciraolo G.} and Loggia G.
+la\index[people]{Loggia G.
+la}.
+2003.
+\emph{Monitoring \textit{Posidonia oceanica} meadows in a Mediterranean coastal lagoon (Stagnone, Italy) by means of neural network and
 ISODATA classification methods}. INT. J. REMOTE SENSING, 2003, VOL. 24, NO. 13,
@@
 \bibitem{Cambridge09}
-Cambridge M.L., Kendrick G.A. 2009. \emph{Contrasting responses of seagrass transplants
+Cambridge M.L.\index[people]{Cambridge M.L.}, Kendrick G.A.\index[people]{Kendrick G.A.} 2009. \emph{Contrasting responses of seagrass transplants
 (Posidonia australis) to nitrogen, phosphorus and iron addition in an estuary and a coastal
@@
 \bibitem{Choo06}
-Choo, C.K. 2006. \emph{Save our Seahorses}. First Edition. \href{www.sosmalaysia.org}{sosmalaysia} \pageref{Choo06}
+Choo, C.K.\index[people]{Choo, C.K.} 2006. \emph{Save our Seahorses}. First Edition. \href{www.sosmalaysia.org}{sosmalaysia} \pageref{Choo06}
 \bibitem{Clark99}
-Clark, R.N., 1999. \emph{Spectroscopy of rocks and minerals and principles of spectroscopy}.
-In: A.N. Rencz (ed.). Manual of Remote Sensing, Chapter 1, John Wiley and Sons, New
-York. 3-58. \pageref{Choo06}
+Clark, R.N.\index[people]{Clark, R.N.}, 1999. \emph{Spectroscopy of rocks and minerals and principles of spectroscopy}.
+In: A.N. Rencz (ed.). Manual of Remote Sensing, Chapter 1, John Wiley and Sons, New York. 3-58. \pageref{Choo06}
 \bibitem{Connell78}
-Connell J.H. 1978. \emph{Diversity in tropical rain forests and coral reefs}. Science 199:1302–
+Connell J.H.\index[people]{Connell J.H.} 1978. \emph{Diversity in tropical rain forests and coral reefs}. Science 199:1302–
 10. \pageref{Connell78}
 \bibitem{Dehouck08}
-Dehouck A., Martiny N., Froidefond J.-M., S\'en\'echal N. 2008. \emph{In-Water Reflectance
-Spectra Measured On Board a Jet-Ski Across a Complex Nearshore Zone of Case-2 Waters
-during the ECORS Experiment}. Manuscript published in "Ocean Optics XIX, Barga: Italie
+Dehouck A.\index[people]{Dehouck A.}, Martiny N.\index[people]{Martiny N.}, Froidefond J.-M.\index[people]{Froidefond J.-M.}, S\'en\'echal N.\index[people]{S\'en\'echal N.} 2008.
+\emph{In-Water Reflectance Spectra Measured On Board a Jet-Ski Across a Complex Nearshore Zone of Case-2 Waters during the ECORS Experiment}.
+Manuscript published in "Ocean Optics XIX, Barga: Italie
 (2008)" \pageref{Dehouck08}
 \bibitem{Dekker05a}
-Dekker A.G., Brando V.E., Anstee J.M. 2005. \emph{Retrospective seagrass change detection
-in a shallow coastal tidal Australian lake}. Remote Sensing of Environment 97:415-433. \pageref{Dekker05a}
+Dekker A.G.\index[people]{Dekker A.G.}, Brando V.E.\index[people]{Brando V.E.}, Anstee J.M.\index[people]{Anstee J.M.} 2005.
+\emph{Retrospective seagrass change detection in a shallow coastal tidal Australian lake}.
+Remote Sensing of Environment 97:415-433.
+\pageref{Dekker05a}
 \bibitem{Dekker05b}
-Dekker A., Brando V., Anstee J. 2005. \emph{Remote Sensing of Seagrass Ecosystems: Use of
-Spaceborne and Airborne Sensors}. SVNL083-Larkum. \pageref{Dekker05b}
+Dekker A.\index[people]{Dekker A.G.}, Brando V.E.\index[people]{Brando V.E.}, Anstee J.M.\index[people]{Anstee J.M.} 2005.
+\emph{Remote Sensing of Seagrass Ecosystems: Use of Spaceborne and Airborne Sensors}.
+SVNL083-Larkum.
+\pageref{05b}
 \bibitem{DiCarlo04}
-Di Carlo G. 2004. \emph{The natural recolonisation process of the seagrass Posidonia
+Di Carlo G.\index[people]{Di Carlo G.} 2004. \emph{The natural recolonisation process of the seagrass Posidonia
 oceanica (L.) Delile after the introduction of the Italo-Algerian methane pipeline in the SW
-Mediterranean sea}. 
-\pageref{DiCarlo04}
+Mediterranean sea}.\pageref{DiCarlo04}
 \bibitem{Diaz-Almela06}
-Diaz-Almela E., Marba N., Alvarez E., Santiago R., Holmer M., Grau A., Danovaro R.,
-Argyrou M., Karakassis I, Duarte C.M. 2006. \emph{Benthic inputs as predictors of seagrass 1
-(\textit{Posidonia oceanica}) fish farm-induced decline}. \href{http://arxiv.org/ftp/qbio/
+Diaz-Almela E.\index[people]{Diaz-Almela E.}, Marb\`a N.\index[people]{Marb\`a N.}, Alvarez E.\index[people]{Alvarez E.}, Santiago R.\index[people]{Santiago R.}, Holmer M.\index[people]{Holmer M.}, Grau A.\index[people]{Grau A.}, Danovaro R.\index[people]{Danovaro R.}, Argyrou M.\index[people]{Argyrou M.}, Karakassis I.\index[people]{Karakassis I.}, Duarte C.M.\index[people]{Duarte C.M.} 2006.
+\emph{Benthic inputs as predictors of seagrass (\textit{Posidonia oceanica}) fish farm-induced decline}.
+\href{http://arxiv.org/ftp/qbio/
 papers/0611/0611006.pdf}{arxiv.org} \pageref{Diaz-Almela06}
 \bibitem{Dierssen03}
-Dierssen H and Zimmerman R.C. 2003. \emph{Ocean color remote sensing of seagrass and
+Dierssen H.\index[people]{} and Zimmerman R.C.\index[people]{} 2003. \emph{Ocean color remote sensing of seagrass and
 bathymetry in the Bahamas Banks by high-resolution airborne imagery}. Limnol. Oceanogr.,
@@
 \bibitem{Dixon05}
-Dixon R, editor. 2005. \emph{Watching the Seagrass Grow – a guide for community seagrass
+Dixon R.\index[people]{}, editor. 2005. \emph{Watching the Seagrass Grow – a guide for community seagrass
 monitoring in NSW (2nd Ed)}. Reference manual . The Community Environment Network,
@@
 \bibitem{Duarte02}
-Duarte, C.M. 2002. \emph{The future of seagrass meadows}. Environmental Conservation 29
-(2), 192–206. \pageref{Duarte02}
+Duarte, C.M.\index[people]{Duarte C.M.} 2002. \emph{The future of seagrass meadows}. Environmental Conservation 29 (2), 192–206. \pageref{Duarte02}
 \bibitem{Dumay02}
-Dumay O., Fernandez C. and Pergent G. 2002. \emph{Primary production and vegetative cycle
-in \textit{Posidonia oceanica} when in competition with the green algae Caulerpa taxifolia nd
+Dumay O.\index[people]{Dumay O.}, Fernandez C.\index[people]{Fernandez C.} and Pergent G.\index[people]{Pergent G.} 2002.
+\emph{Primary production and vegetative cycle in \textit{Posidonia oceanica} when in competition with the green algae Caulerpa taxifolia nd
 Caulerpa racemosa}. J.Mar. Biol. Ass. UK, 82, 379-387. \pageref{Dumay02}
 \bibitem{Dural10}
-Dural B. 2010. \emph{Phenological observations on \textit{Posidonia oceanica} (L.) Delile medows
+Dural B.\index[people]{Dural B.} 2010. \emph{Phenological observations on \textit{Posidonia oceanica} (L.) Delile medows
 along the coast of Akkum (Sıgacık Bay, Aegean Sea, Turkey)}. J. Black Sea/Mediterranean
@@
 \bibitem{Edwards00}
-Edwards A.J. (ed.) 2000. \emph{Remote Sensing handbook for tropical coastal management}.
+Edwards A.J.\index[people]{Edwards A.J.} (ed.) 2000. \emph{Remote Sensing handbook for tropical coastal management}.
 \href{http://www.unesco.org/csi/pub/source/rs.htm}{UNESCO} \pageref{Edwards00}
 \bibitem{Ferguson97}
-Ferguson R.L. and Karl Korfmacher K. 1997 \emph{Remote sensing and GIS analysis of seagrass
-meadows in North Carolina, USA}. Aquatic Botany 58 (1997) 241-258. \pageref{Ferguson97}
+Ferguson R.L.\index[people]{Ferguson R.L.} and Karl Korfmacher K.\index[people]{Korfmacher Karl K.} 1997 \emph{Remote sensing and GIS analysis of seagrass meadows in North Carolina, USA}.
+Aquatic Botany 58 (1997) 241-258.
+\pageref{Ferguson97}
 \bibitem{Fletcher09}
-Fletcher R.S., Pulich Jr. W. and Hardegree B. 2009. \emph{A Semiautomated Approach for
-Monitoring Landscape Changes in Texas Seagrass Beds from Aerial Photography}. Coastal
+Fletcher R.S.\index[people]{Fletcher R.S.}, Pulich Jr.
+W.\index[people]{Pulich Jr.
+W.} and Hardegree B.\index[people]{Hardegree B.} 2009.
+\emph{A Semiautomated Approach for Monitoring Landscape Changes in Texas Seagrass Beds from Aerial Photography}.
+Coastal
 Education and Research Foundation, doi: 10.2112/07-0882.1 \pageref{Fletcher09}
 \bibitem{Fornes06}
-Fornes A., Basterretxea G., Orfila A., Jordi A., Alvarez A. and Tintore J. 2006. \emph{Mapping
+Fornes A.\index[people]{Fornes A.}, Basterretxea G.\index[people]{Basterretxea G.}, Orfila A.\index[people]{Orfila A.}, Jordi A.\index[people]{Jordi A.}, Alvarez A.\index[people]{Alvarez A.} and Tintore J.\index[people]{Tintore J.} 2006.
+\emph{Mapping
 \textit{Posidonia oceanica} from IKONOS}. ISPRS Journal of Photogrammetry and Remote Sensing
@@
 \bibitem{Francour06}
-Francour, P., Magreau, J.F., Mannoni, P.A., Cottalorda, J.M., Gratiot, J., 2006.
-\emph{Management guide for Marine Protected Areas of the Mediterranean Sea, Permanent
-Ecological Moorings}. Universite de Nice-Sophia Antipolis and Parc National de Port-Cros,
+Francour, P.\index[people]{Francour P.}, Magreau, J.F.\index[people]{Magreau, J.F.}, Mannoni, P.A.\index[people]{Mannoni, P.A.}, Cottalorda, J.M.\index[people]{Cottalorda, J.M}, Gratiot, J.\index[people]{Gratiot, J.}, 2006.\emph{Management guide for Marine Protected Areas of the Mediterranean Sea, Permanent Ecological Moorings}.
+Universite de Nice-Sophia Antipolis and Parc National de Port-Cros,
 Nice, 68 pp. \pageref{Francour06}
 \bibitem{Francour99}
-Francour P., Ganteaume A. and M. Poulain. 1999. \emph{Effects of boats anchoring in
-\textit{Posidonia oceanica} seagrass beds in the Port-Cross National Park (north-western
+Francour P.\index[people]{Francour P.}, Ganteaume A.\index[people]{Ganteaume A.} and M.
+Poulain\index[people]{Poulain M.}.
+1999.
+\emph{Effects of boats anchoring in \textit{Posidonia oceanica} seagrass beds in the Port-Cross National Park (north-western
 Mediterranean Sea)}. Aquatic conservation: Marine and Freshwater Ecosystems: 9: 391-400. \pageref{Francour99}
 \bibitem{Fyfe03}
-Fyfe S.K. 2003. \emph{Spatial and temporal variation in spectral reflectance: Are seagrass
+Fyfe S.K.\index[people]{Fyfe S.K.} 2003. \emph{Spatial and temporal variation in spectral reflectance: Are seagrass
 species spectrally distinct?} Limnology and Oceanography 48: 464–479. \pageref{Fyfe03}
 \bibitem{Fyfe04}
-Fyfe S.K. 2004. \emph{Hyperspectral studies of New South Wales with particular emphasis on
+Fyfe S.K.\index[people]{Fyfe S.K.} 2004. \emph{Hyperspectral studies of New South Wales with particular emphasis on
 the detection of light stress in Eelgrass Zostera capricorni. PhD Thesis}. The University of
@@
 \bibitem{Gege04}
-Gege P. 2004. \emph{The water color simulator WASI: an integrating software tool for analysis
+Gege P.\index[people]{Gege P.} 2004. \emph{The water color simulator WASI: an integrating software tool for analysis
 and simulation of optical in situ spectra}. Computers and Geosciences 30 (2004) 523–532. \pageref{Gege04}
 \bibitem{Gege05}
-Gege, P. 2005. \emph{The Water Colour Simulator WASI. User manual for version 3}. DLR
+Gege, P.\index[people]{Gege P.} 2005. \emph{The Water Colour Simulator WASI. User manual for version 3}. DLR
 Internal Report IB 564-01/05, 83 pp. \pageref{Gege05}
 \bibitem{Gobert06}
-Gobert S., Cambridge, M. L. Velimirov, B. Pergent, G. Lepoint, G. Bouquegneau, J. M.
-Dauby, P. Pergent-Martini, C. Walker, D. I. 2006. Biology of Posidonia. In Larkum A.W.D.,
-Orth, R.J., Duarte C.M. (eds.), \emph{Seagrasses: Biology, ecology and conservation}. Springer
-Verlag, Berlin, Pages 387–408. \pageref{Gobert06}
+Gobert S.\index[people]{Gobert S.}, Cambridge, M.
+L.\index[people]{Cambridge M.L.}, Velimirov, B.\index[people]{Velimirov B.}, Pergent, G.\index[people]{Pergent G.}, Lepoint, G.\index[people]{Lepoint G.}, Bouquegneau J.-M.\index[people]{Bouquegneau J.-M.}, Dauby, P.\index[people]{Dauby, P.}, Pergent-Martini, C.\index[people]{Pergent-Martini C.}, Walker, D.
+I.\index[people]{Walker, D.I.}.
+2006.
+Biology of Posidonia.
+In Larkum A.W.D.\index[people]{Larkum A.W.D}, Orth, R.J.\index[people]{Orth R.J.}, Duarte C.M.\index[people]{Duarte C.M.} (eds.), \emph{Seagrasses: Biology, ecology and conservation}.
+Springer Verlag, Berlin, Pages 387–408.
+\pageref{Gobert06}
 \bibitem{Gobert01}
-Gobert S., Defawe O, Lepoint G, Demoulin V and Bouquegneau J-M. 2001. \emph{Anthesis
-effects on \textit{Posidonia oceanica} (L.) Delile phenology in the Bay of Calvi (Corsica,
-Mediterranean Sea)}. Hydrobiologia 455: 121–125. \pageref{Gobert01}
+Gobert S.\index[people]{Gobert S.}, Defawe O.\index[people]{Defawe O.}, Lepoint G.\index[people]{Lepoint G.}, Demoulin V.\index[people]{Demoulin V.} and Bouquegneau J.-M.\index[people]{Bouquegneau J.-M.} 2001.
+\emph{Anthesis effects on \textit{Posidonia oceanica} (L.) Delile phenology in the Bay of Calvi (Corsica, Mediterranean Sea)}.
+Hydrobiologia 455: 121–125.
+\pageref{Gobert01}
 \bibitem{Green96}
-Green, E.P., P.J. Mumby, A.J. Edwards, and C.D. Clark, 1996. \emph{A review of remote
-sensing for the assessment and management of tropical coastal resources}. Coast. Manag.,
+Green, E.P.\index[people]{Green E.P.}, Mumby P.J.\index[people]{Mumby P.J.}, Edwards  A.J.\index[people]{Edwards A.J.}, and Clark C.D.\index[people]{Clark C.D.}, 1996.
+\emph{A review of remote sensing for the assessment and management of tropical coastal resources}.
+Coast.
+Manag.,
 24(1):1-40. \pageref{Green96}
 \bibitem{Green03}
-Green E.P., Short F.T. 2003. \emph{World Atlas of Seagrasses}. University of California Press,
+Green E.P.\index[people]{Green E.P.}, Short F.T.\index[people]{Short F.T.} 2003. \emph{World Atlas of Seagrasses}. University of California Press,
 Berkeley, USA. \pageref{Green03}
 \bibitem{Guidetti08}
-Guidetti P. 2008. \emph{Population dynamics of the seagrass \textit{Posidonia oceanica} for
+Guidetti P.\index[people]{Guidetti P.} 2008. \emph{Population dynamics of the seagrass \textit{Posidonia oceanica} for
 evaluating meadow health status: a study case using reconstruction techniques}. Laboratory
@@
 \bibitem{Gullstroom06}
-Gullstroom M.,, Lunden B., Bodin M., Kangwe J., Ohman M.C., Mtolera M.S.P. and
-Bjork M. 2006. \emph{Assessment of changes in the seagrass-dominated submerged vegetation of
-tropical Chwaka Bay (Zanzibar) using satellite remote sensing}. Estuarine, Coastal and Shelf
+Gullstroom M.\index[people]{Gullstroom M.}, Lunden B.\index[people]{Lunden B.}, Bodin M.\index[people]{Bodin M.}, Kangwe J.\index[people]{Kangwe J.}, Ohman M.C.\index[people]{Ohman M.C.}, Mtolera M.S.P.\index[people]{Mtolera M.S.P.} and Bjork M.\index[people]{Bjork M.} 2006.
+\emph{Assessment of changes in the seagrass-dominated submerged vegetation of tropical Chwaka Bay (Zanzibar) using satellite remote sensing}.
+Estuarine, Coastal and Shelf
 Science 67 (2006) 399-408. \pageref{Gullstroom06}
 \bibitem{Haag08}
-Haag S.M., Kennish M., Sakowicz G.P. 2008. \emph{Seagrass Habitat Characterization in
-Estuarine Waters of the Jacques Cousteau National Estuarine Research Reserve Using
+Haag S.M.\index[people]{Haag S.M.}, Kennish M.\index[people]{Kennish M.}, Sakowicz G.P.\index[people]{Sakowicz G.P.} 2008.
+\emph{Seagrass Habitat Characterization in Estuarine Waters of the Jacques Cousteau National Estuarine Research Reserve Using
 Underwater Videographic Imaging Techniques}. Journal of Coastal Research, SI(55), 171–
@@
 \bibitem{DenHartog70}
-Den Hartog, C. 1970. \emph{The Seagrasses of the World}. North Holland Publishing Company,
+Den Hartog, C.\index[people]{Den Hartog C.} 1970. \emph{The Seagrasses of the World}. North Holland Publishing Company,
 Amsterdam, 275 pp. 31 plates. \pageref{DenHartog70}
 \bibitem{DenHartog77}
-Den Hartog, C., 1977. \emph{Structure, function, and classification in seagrass communities}.
-In: C.P. McRoy and C. Helfferich (Editors), Seagrass Ecosystems: a Scientific Perspective.
-Marcel Dekker, New York, pp. 89-121. \pageref{DenHartog77}
+Den Hartog, C.\index[people]{Den Hartog C.}, 1977. \emph{Structure, function, and classification in seagrass communities}.
+In: C.P. McRoy\index[people]{McRoy, C.P.} and C. Helfferich\index[people]{Helfferich C.} (Editors), Seagrass Ecosystems: a Scientific Perspective.
+Marcel, New York, pp. 89-121. \pageref{DenHartog77}
 \bibitem{Heege03}
-Heege T., Bogner A., Pinnel N. 2003. \emph{Mapping of submerged aquatic vegetation with a
-physically based process chain}. Proc. Remote Sensing, September 8-12 2--3/ Barcelona,
-Spain. Ed. by Kramer E. The International Society for Optical Engineering, Vol.5233. \pageref{Heege03}
+Heege T.\index[people]{Heege T.}, Bogner A.\index[people]{Bogner A.}, Pinnel N.\index[people]{Pinnel N.} 2003.
+\emph{Mapping of submerged aquatic vegetation with a physically based process chain}.
+Proc.
+Remote Sensing, September 8-12 2--3/ Barcelona,
+Spain. Ed. by Kramer E.\index[people]{Kramer E.} The International Society for Optical Engineering, Vol.5233. \pageref{Heege03}
 \bibitem{Hirose04}
-Hirose K., Syoji M., Hang H.T.M., Anh N.H., Triet T. and Nam V.N. 2004. \emph{Satellite
-data application for mangrove management}. International Symposium on Geoinfomatics for
-Spatial Infrastructure Development in Earth and Allied Sciences. \pageref{Hirose04} 
+Hirose K.\index[people]{Hirose K.}, Syoji M.\index[people]{Syoji M.}, Hang H.T.M.\index[people]{Hang H.T.M.}, Anh N.H.\index[people]{Anh N.H.}, Triet T.\index[people]{Triet T} and Nam V.N.\index[people]{Nam V.N.} 2004.
+\emph{Satellite data application for mangrove management}.
+International Symposium on Geoinfomatics for Spatial Infrastructure Development in Earth and Allied Sciences.
+\pageref{Hirose04}
 \bibitem{Hochberg03a}
-Hochberg, E.J., Atkinson, M.J. 2003a. \emph{Capabilities of remote sensors to classify coral,
+Hochberg, E.J.\index[people]{Hochberg E.J.}, Atkinson, M.J.\index[people]{Atkinson, M.J.} 2003a. \emph{Capabilities of remote sensors to classify coral,
 algae, and sand as pure and mixed spectra}. Remote Sensing of Environment 85, 174–189. \pageref{Hochberg03a}
 \bibitem{Hochberg03b}
-Hochberg, E.J., Atkinson, M.J., Andrefouet, S., 2003b. \emph{Spectral reflectance of coral reef
-bottom-types worldwide and implications for coral reef remote sensing}. Remote Sensing of
+Hochberg, E.J.\index[people]{Hochberg E.J.}, Atkinson, M.J.\index[people]{Atkinson, M.J.}, Andr\'efou\"et, S.\index[people]{Andr\'efou\"et S.}, 2003b.
+\emph{Spectral reflectance of coral reef bottom-types worldwide and implications for coral reef remote sensing}.
+Remote Sensing of
 Environment 85, 159–173. \pageref{Hochberg03b}
 \bibitem{Hogarth07}
-Hogarth P. 2007. \emph{The Biology of Mangroves and Seagrasses}. Oxford University Press. \pageref{Hogarth07}
+Hogarth P.\index[people]{Hogarth P.} 2007. \emph{The Biology of Mangroves and Seagrasses}. Oxford University Press. \pageref{Hogarth07}
 \bibitem{Holmer09}
-Holmer M., Marba N, Lamote M and Duarte C. 2009\emph{Deterioration of Sediment Quality in
-Seagrass Meadows (\textit{Posidonia oceanica}) Invaded by Macroalgae (Caulerpa sp.)}. Estuaries
+Holmer M.\index[people]{Holmer M.}, Marb\`a N.\index[people]{Marb\`a N.}, Lamote M.\index[people]{Lamote M.} and Duarte C.\index[people]{Duarte C.M.} 2009\emph{Deterioration of Sediment Quality in Seagrass Meadows (\textit{Posidonia oceanica}) Invaded by Macroalgae (Caulerpa sp.)}.
+Estuaries
 and Coasts (2009) 32:456–466 DOI 10.1007/s12237-009-9133-4 \pageref{Hogarth07}
 \bibitem{Holmer08}
-Holmer M, Argyrou M, Dalsgaard T., Danovaro R, Diaz-Almela E, Duarte C,
-Frederiksen M, Grau A, Karakassis I, Marbà N, Mirto S, Pérez M, Pusceddu A, Tsapakis M.2008
+Holmer M.\index[people]{Holmer M.}, Argyrou M.\index[people]{Argyrou M.}, Dalsgaard T.\index[people]{Dalsgaard T.}, Danovaro R.\index[people]{Danovaro R.}, Diaz-Almela E.\index[people]{Diaz-Almela E.}, Duarte C.\index[people]{Duarte C.M.}, Frederiksen M.\index[people]{Frederiksen M.}, Grau A.\index[people]{Grau A.}, Karakassis I.\index[people]{Karakassis I.}, Marb\`a N.\index[people]{Marb\`a N.}, Mirto S.\index[people]{Mirto S.}, P\'erez M.\index[people]{P\'erez M}, Pusceddu A.\index[people]{Pusceddu A.}, Tsapakis M.\index[people]{Tsapakis M.}, 2008
 \emph{Effects of fish farm waste on \textit{Posidonia oceanica} meadows: Synthesis and provision of
@@
 \bibitem{Iverson86}
-Iverson R.L. and Bittaker H.F. 1986. \emph{Seagrass Distribution and Abundance in Eastern
+Iverson R.L.\index[people]{Iverson R.L.} and Bittaker H.F.\index[people]{Bittaker H.F.} 1986. \emph{Seagrass Distribution and Abundance in Eastern
 Gulf of Mexico Coastal Waters}. Estuarine, Coastal and Shelf Science, 22, 577-602. \pageref{Iverson86}
 \bibitem{Jackson07}
-Jackson J., Nemeth D.J. 2007. \emph{A New Method to Describe Seagrass Habitat Sampled
-During Fisheries-Independent Monitoring}. Estuaries and Coasts Vol. 30, No. 1,
-p. 171–178 February 2007 \pageref{Jackson07}
+Jackson J.\index[people]{Jackson J.}, Nemeth D.J.\index[people]{Nemeth D.J.} 2007. \emph{A New Method to Describe Seagrass Habitat Sampled
+During Fisheries-Independent Monitoring}. Estuaries and Coasts Vol. 30, No. 1, p. 171–178 February 2007 \pageref{Jackson07}
 \bibitem{Jensen01}
-Jensen S. and Bell S. 2001. \emph{Seagrass growth and patch dynamics: cross-scale
+Jensen S.\index[people]{Jensen S.L.} and Bell S.S.\index[people]{Bell S.S.} 2001. \emph{Seagrass growth and patch dynamics: cross-scale
 morphological plasticity}. Plant Ecology 155: 201–217, 2001. \pageref{Jensen01}
 \bibitem{Jensen95}
-Jensen J.R., Rutchey K, Koch M.S, Narumalani S. 1995. \emph{Inland wetland change
-detection in the everglades water conservation area 2a using a time-series of normalized
+Jensen J.R.\index[people]{Jensen J.R.}, Rutchey K.\index[people]{Rutchey K.}, Koch M.S.\index[people]{Koch M.S.}, Narumalani S.\index[people]{Narumalani S.} 1995.
+\emph{Inland wetland change detection in the everglades water conservation area 2a using a time-series of normalized
 remotely-sensed data}. Photogrammetric Eng Remote Sensing 61: 199–209 \pageref{Jensen95}
 \bibitem{Jerlov52}
-Jerlov N.G. 1952. \emph{Influence of suspended and dissolved matter on the transparency of
+Jerlov N.G.\index[people]{Jerlov N.G.} 1952. \emph{Influence of suspended and dissolved matter on the transparency of
 sea water}. Wiley online Library. TellusVolume 5, Issue 1, Published online: 2010. \pageref{Jerlov52}
 \bibitem{Kalinowski04}
-Kalinowski A. and Oliver S. \emph{2004 ASTER Mineral Index Processing Manual}. Remote
+Kalinowski A.\index[people]{Kalinowski A.} and Oliver S.\index[people]{Oliver S.} \emph{2004 ASTER Mineral Index Processing Manual}. Remote
 Sensing Applications. Geoscience Australia \pageref{Kalinowski04}
 \bibitem{Kelly80}
-Kelly, M.G., 1980. \emph{Remote sensing of seagrass beds, Handbook of Seagrass Biology: An
-Ecosystem Perspective} (R.C. Phillips and C.P. Mc Roy, editors), Garland STPM Press, New
+Kelly, M.G.\index[people]{Kelly, M.G.}, 1980. \emph{Remote sensing of seagrass beds, Handbook of Seagrass Biology: An
+Ecosystem Perspective} (R.C. Phillips\index[people]{Phillips, R.C.} and C.P. McRoy\index[people]{McRoy, C.P.}, editors), Garland STPM Press, New
 York, N.Y., pp. 69-86. \pageref{Kelly80}
 \bibitem{Kendrick00}
-Kendrick GA, Hegge BJ, Wyllie A, Davidson A, Lord DA. 2000. \emph{Changes in Seagrass
-Cover on Success and Parmelia Banks, Western Australia Between 1965 and 1995}.
-Estuarine, Coastal and Shelf Science, 50:341-353. \pageref{Kendrick00}
+Kendrick G.A.\index[people]{Kendrick G.A.}, Hegge B.J.\index[people]{Hegge B.J.}, Wyllie A.\index[people]{Wyllie A.}, Davidson A.\index[people]{Davidson A.}, Lord D.A.\index[people]{Lord D.A.}.
+2000.
+\emph{Changes in Seagrass Cover on Success and Parmelia Banks, Western Australia Between 1965 and 1995}.
+Estuarine, Coastal and Shelf Science, 50:341-353.
+\pageref{Kendrick00}
 \bibitem{Kirk94}
-Kirk J.T.O. 1994. \emph{Light and photosynthesis in aquatic ecosystems}, pp 1–509. University
+Kirk J.T.O.\index[people]{Kirk J.T.O.} 1994. \emph{Light and photosynthesis in aquatic ecosystems}, pp 1–509. University
 Press, Cambridge, UK. \pageref{Kirk94}
 \bibitem{Kirkman96}
-Kirkman H. 1996. \emph{Baseline and Monitoring Methods for Seagrass Meadows}. Journal of
+Kirkman H.\index[people]{Kirkman H.} 1996. \emph{Baseline and Monitoring Methods for Seagrass Meadows}. Journal of
 Environmental Management (1996) 47, 191–201. \pageref{Kirkman96}
 \bibitem{Kuo89}
-Kuo J, McComb AJ, 1989. \emph{Seagrass taxonomy, structure and development. In: Larkum
-AWD, McComb AJ, Shepherd, eds. Biology of seagrasses: a treatise on the biology of
-seagrasses with special reference to the Australian region}. Elsevier, The Netherlands: 6-73. \pageref{Kuo89}
+Kuo J.\index[people]{Kuo J.}, McComb A.J.\index[people]{McComb A.J.}, 1989. \emph{Seagrass taxonomy, structure and development. In: Larkum
+A.W.D.\index[people]{Larkum A.W.D.}, McComb A.J.\index[people]{McComb A.J.}, Shepherd J.\index[people]{Shepherd J.}, eds.
+Biology of seagrasses: a treatise on the biology of seagrasses with special reference to the Australian region}.
+Elsevier, The Netherlands: 6-73.
+\pageref{Kuo89}
 \bibitem{Leoni06}
-Leoni V., Pasqualini V., Pergent-Martini Ch., Vela A., Pergent G. 2006. \emph{Morphological
-responses of \textit{	extit{Posidonia oceanica}} to experimental nutrient enrichment of the canopy water}.
-Journal of Experimental Marine Biology and Ecology xx (2006) xxx–xxx. \pageref{Leoni06}
+Leoni V.\index[people]{Leoni V.}, Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Vela A.\index[people]{Vela A.}, Pergent G.\index[people]{Pergent G.} 2006.
+\emph{Morphological responses of \textit{Posidonia oceanica} to experimental nutrient enrichment of the canopy water}.
+Journal of Experimental Marine Biology and Ecology xx (2006) xxx–xxx.
+\pageref{Leoni06}
 \bibitem{Leoni07}
-Leoni V., Pasqualini V., Pergent-Martini Ch., Vela A., Pergent G. 2007. \emph{Physiological
-responses of \textit{{Posidonia oceanica}} to experimental nutrient enrichment of the canopy water}.
-Journal of Experimental Marine Biology and Ecology 349, 73–83. \pageref{Leoni07}
+Leoni V.\index[people]{Leoni V.}, Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Vela A.\index[people]{Vela A.}, Pergent G.\index[people]{Pergent G.} 2007.
+\emph{Physiological responses of \textit{{Posidonia oceanica}} to experimental nutrient enrichment of the canopy water}.
+Journal of Experimental Marine Biology and Ecology 349, 73–83.
+\pageref{Leoni07}
 \bibitem{Leoni08}
-Leoni V., Vela A., Pasqualini V., Pergent-Martini Ch., Pergent G. 2008. \emph{Effects of
-experimental reduction of light and nutrient enrichments (N and P) on seagrasses: a review}.
+Leoni V.\index[people]{Leoni V.}, Vela A.\index[people]{Vela A.}, Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pergent G.\index[people]{Pergent G.} 2008.
+\emph{Effects of experimental reduction of light and nutrient enrichments (N and P) on seagrasses: a review}.
 Aquatic Conservation: Marine and Freshwater Ecosystems, 18: 202–220. \pageref{Leoni08}
 \bibitem{Lewis01}
-Lewis M. 2001. \emph{Discriminating Vegetation with Hyperspectral Imagery - What is
+Lewis M.\index[people]{Lewis M.} 2001. \emph{Discriminating Vegetation with Hyperspectral Imagery - What is
 Possible?} \href{http://www.ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=00978199}{ieeexplore.ieee.org} Proceedings,
-International Geoscience and Remote Sensing Symposium, 9-13 July 2001, Sydney,
-Australia. ISBN 0-7803-7031-7 \pageref{Lewis01}
+International Geoscience and Remote Sensing Symposium, 9-13 July 2001, Sydney, Australia. ISBN 0-7803-7031-7 \pageref{Lewis01}
 \bibitem{Louchard03}
-Louchard E.M., Reid R.P., Stephens F.C., Davis C.O., Leathers R.A., Downes T.V.
-2003. \emph{Optical remote sensing of benthic habitats and bathymetry in coastal environments at
-Lee Stocking Island, Bahamas: A comparative spectral classification approach}. Limnology
+Louchard E.M.\index[people]{Louchard E.M.}, Reid R.P.\index[people]{Reid R.P.}, Stephens F.C.\index[people]{Stephens F.C.}, Davis C.O.\index[people]{Davis C.O.}, Leathers R.A.\index[people]{Leathers R.A.}, Downes T.V.\index[people]{Downes T.V.} 2003.
+\emph{Optical remote sensing of benthic habitats and bathymetry in coastal environments at Lee Stocking Island, Bahamas: A comparative spectral classification approach}.
+Limnology
 and Oceanography (2003) 48:511-521. \pageref{Louchard03}
 \bibitem{Luque04}
-Luque A. and Templado J.A. 2004. \emph{Praderas y bosques marinos de Andalucia
+Luque A.\index[people]{Luque A.} and Templado J.A.\index[people]{Templado J.A.} 2004. \emph{Praderas y bosques marinos de Andalucia
 Consejeria de Medio Ambiente, Junta de Andalucia} (in press). \pageref{Luque04}
 \bibitem{Lyzenga81}
-Lyzenga D.R. 1981. \emph{Remote sensing of bottom reflectance and water attenuation in
+Lyzenga D.R.\index[people]{Lyzenga D.R.} 1981. \emph{Remote sensing of bottom reflectance and water attenuation in
 shallow water using aircraft and Landsat data}. International Journal of Remote Sensing 2,
@@
 \bibitem{Maffione97}
-Maffione R.A. and Dana D.R. 1997 \emph{Instruments and methods for measuring the
+Maffione R.A.\index[people]{Maffione R.A.} and Dana D.R.\index[people]{Dana D.R.} 1997 \emph{Instruments and methods for measuring the
 backward-scattering coefficient of ocean waters}. Optical Society of America \pageref{Maffione97}
 \bibitem{Maffione01a}
-Maffione R.A. 2001a. \emph{Evolution and Revolution in Measuring Ocean Optical Properties}.
+Maffione R.A.\index[people]{Maffione R.A.} 2001a. \emph{Evolution and Revolution in Measuring Ocean Optical Properties}.
 Oceanography, Vol.14, No.3. \pageref{Maffione01a}
 \bibitem{Maffione01b}
-Maffione R.A. 2001b.\emph{Hyperspectral Optical Properties, Remote Sensing, and
+Maffione R.A.\index[people]{Maffione R.A.} 2001b.\emph{Hyperspectral Optical Properties, Remote Sensing, and
 Underwater Visibility}. Hydro-Optics, Biology, and Instrumentation Laboratories. \pageref{Maffione01b}
 \bibitem{Malthus03}
-Malthus T.J. and Karpouzli E. 2003. \emph{Integrating field and high spatial resolution
-satellite based methods for monitoring shallow submersed aquatic habitats in the Sound of
-Eriskay, Scotland, UK}. Int Journal Remote Sensing 24: 2585–2593. \pageref{Malthus03}
+Malthus T.J.\index[people]{Malthus T.J.} and Karpouzli E.\index[people]{Karpouzli E.} 2003. \emph{Integrating field and high spatial resolution
+satellite based methods for monitoring shallow submersed aquatic habitats in the Sound of Eriskay, Scotland, UK}. Int Journal Remote Sensing 24: 2585–2593. \pageref{Malthus03}
 \bibitem{Matarrese08}
-Matarrese R., Acquaro M., Morea A., Tijani K., Chiaradia M.T. 2008. \emph{Applications of
-remote sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}. Geoscience and
-Remote Sensing Symposium. ISBN:978-1-4244-2807-6. \pageref{Matarrese08}
+Matarrese R.\index[people]{Matarrese R.}, Acquaro M.\index[people]{Acquaro M.}, Morea A.\index[people]{Morea A.}, Tijani K.\index[people]{Tijani K.}, Chiaradia M.T.\index[people]{Chiaradia M.T.} 2008.
+\emph{Applications of remote sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}.
+Geoscience and Remote Sensing Symposium.
+ISBN:978-1-4244-2807-6.
+\pageref{Matarrese08}
 \bibitem{Matarrese06}
-Matarrese R., Acquaro M., Morea A., Khalid T., Chiaradia M.T. 2006 \emph{Applications of remote
-sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}.
+Matarrese R.\index[people]{Matarrese R.}, Acquaro M.\index[people]{Acquaro M.}, Morea A.\index[people]{Morea A.}, Khalid T.\index[people]{Khalid T.}, Chiaradia M.T.\index[people]{Chiaradia M.T.} 2006 \emph{Applications of remote sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}.
 \href{http://www.igarss08.org/Abstracts/pdfs/3888.pdf}{International Geoscience and Remote Sensing Symposium} \pageref{Matarrese06}
 \bibitem{McKenzie02}
-McKenzie L.J., Campbell S.J. 2002. \emph{Manual for Community (citizen) Monitoring of
-Seagrass Habitat}. Western Pacific Edition. Marine Plant Ecology Group, QDPI, Northern
-Fisheries Centre, Cairns. 43 pp. \pageref{McKenzie02}
+McKenzie L.J.\index[people]{McKenzie L.J.}, Campbell S.J.\index[people]{Campbell S.J.} 2002. \emph{Manual for Community (citizen) Monitoring of
+Seagrass Habitat}. Western Pacific Edition. Marine Plant Ecology Group, QDPI, Northern Fisheries Centre, Cairns. 43 pp. \pageref{McKenzie02}
 \bibitem{McKenzie03}
-McKenzie L.J., Campbell S.J. and Roder C.A. 2003. \emph{Seagrass-Watch: Manual for
-Mapping and Monitoring Seagrass Resources by Community (citizen) Volunteers}. 2nd
-Edition. QFS, NFC, Cairns. 100pp. \pageref{McKenzie03}
+McKenzie L.J.\index[people]{McKenzie L.J.}, Campbell S.J.\index[people]{Campbell S.J.} and Roder C.A.\index[people]{Roder C.A.} 2003.
+\emph{Seagrass-Watch: Manual for Mapping and Monitoring Seagrass Resources by Community (citizen) Volunteers}.
+2nd Edition.
+QFS, NFC, Cairns.
+100pp.
+\pageref{McKenzie03}
 \bibitem{McKenzieetal03}
-McKenzie L.J. 2003. \emph{Guidelines for the rapid assessment of seagrass habitats in the
-Western Pacific}. Marine Plant Ecology Group, QDPI, Northern Fisheries Centre, Cairns,
-\href{http://www.seagrasswatch.org}{seagrasswatch} \pageref{McKenzieetal03}
+McKenzie L.J.\index[people]{McKenzie L.J.} 2003. \emph{Guidelines for the rapid assessment of seagrass habitats in the
+Western Pacific}. Marine Plant Ecology Group, QDPI, Northern Fisheries Centre, Cairns, \href{http://www.seagrasswatch.org}{seagrasswatch} \pageref{McKenzieetal03}
 \bibitem{McKenzie06}
-McKenzie L.J., Campbell S.J. and F.Lasi 2006. \emph{Seagrasses and Mangroves}. In: Green
-A.P. Lokani W., Atu P., Ramonia P., Thomas P. and Almanmy J (eds). 2006 \emph{Solomon
-Islands Marine Assessment. Technical report of survey conducted May 13 to June 17 2004}.
-The Nature Conservancy, Indo-Pacific Resource Centre. Report No 1/06 \pageref{McKenzie06}
+McKenzie L.J.\index[people]{McKenzie L.J.}, Campbell S.J.\index[people]{Campbell S.J.} and F.Lasi\index[people]{Lasi, F.} 2006.
+\emph{Seagrasses and Mangroves}.
+In: Green A.P.\index[people]{Green A.P.}, Lokani W.\index[people]{Lokani W.}, Atu P.\index[people]{Atu P.}, Ramonia P.\index[people]{Ramonia P.}, Thomas P.\index[people]{Thomas P.} and Almanmy J.\index[people]{Almanmy J.} (eds).
+2006 \emph{Solomon Islands Marine Assessment.
+Technical report of survey conducted May 13 to June 17 2004}.
+The Nature Conservancy, Indo-Pacific Resource Centre.
+Report No 1/06 \pageref{McKenzie06}
 \bibitem{McKenzie07}
-McKenzie L.J. and Yoshida R. 2007. \emph{Seagrass-Watch: Guidelines for Monitoring
-Seagrass Habitats in the Fiji Islands. Proceeding of a training workshop. Corpus Christi
-Teachers College Laucala Bay, Suva, Fiji, 16th June 2007}. Seagrass-Watch HQ Cairns, 42pp. \pageref{McKenzie07}
+McKenzie L.J.\index[people]{McKenzie L.J.} and Yoshida R.\index[people]{Yoshida R.} 2007. \emph{Seagrass-Watch: Guidelines for Monitoring
+Seagrass Habitats in the Fiji Islands. Proceeding of a training workshop. Corpus Christi Teachers College Laucala Bay, Suva, Fiji, 16th June 2007}. Seagrass-Watch HQ Cairns, 42pp. \pageref{McKenzie07}
 \bibitem{McKenzie08}
-McKenzie L.J. 2008. \emph{Seagrass-Watch. Proceedings of a Workshop for Mapping and
-Monitoring Seagrass Habitats in North East Arnhem Land, Northern Territory. 18-20
-October 2008}. Seagrass-Watch HQ, Cairns, 49pp. \pageref{McKenzie08}
+McKenzie L.J.\index[people]{McKenzie L.J.} 2008. \emph{Seagrass-Watch. Proceedings of a Workshop for Mapping and
+Monitoring Seagrass Habitats in North East Arnhem Land, Northern Territory. 18-20 October 2008}. Seagrass-Watch HQ, Cairns, 49pp. \pageref{McKenzie08}
 \bibitem{McKenzie09}
-McKenzie L.J., Yoshida R. 2009. Seagrass Watch. Proceedings of a Workshop for
-Monitoring Seagrass Habitats in Indonesia. The Nature Conservancy, Coral Triangle
-Centre, Sanur, Bali, Indonesia. \pageref{McKenzie09}
+McKenzie L.J.\index[people]{McKenzie L.J.}, Yoshida R.\index[people]{Yoshida R.} 2009. Seagrass Watch. Proceedings of a Workshop for
+Monitoring Seagrass Habitats in Indonesia. The Nature Conservancy, Coral Triangle Centre, Sanur, Bali, Indonesia. \pageref{McKenzie09}
 \bibitem{Meinesz91}
-Meinesz, A., Lefevre, J.R., Astier, J.M., 1991. \emph{Impact of coastal development on the
-infralittoral zone along the south-eastern Mediterranean shore of continental France}.
-Marine Pollution Bullettin 23, 343–347. \pageref{Meinesz91}
+Meinesz, A.\index[people]{Meinesz A.}, Lefevre, J.R.\index[people]{Lefevre, J.R.}, Astier, J.M.\index[people]{Astier, J.M.}, 1991.
+\emph{Impact of coastal development on the infralittoral zone along the south-eastern Mediterranean shore of continental France}.
+Marine Pollution Bullettin 23, 343–347.
+\pageref{Meinesz91}
 \bibitem{Mellors09a}
-Mellors J. and McKenzie L. 2009(a). \emph{Seagrass-Watch: Guidelines for Monitoring
-Seagrass Habitats. Proceedings of a workshop held on Thursday Island, Torres Strait 4-
-8.03.2009}. \pageref{Mellors09a}
+Mellors J.\index[people]{Mellors J.} and McKenzie L.\index[people]{McKenzie L.J.} 2009(a). \emph{Seagrass-Watch: Guidelines for Monitoring
+Seagrass Habitats. Proceedings of a workshop held on Thursday Island, Torres Strait 4-8.03.2009}. \pageref{Mellors09a}
 \bibitem{Mellors09b}
-Mellors J and McLenzie L. 2009(b). \emph{Seagrass-Watch: Seagrass Monitoring Guidelines
-for Torres Strait Communities. Meer Island. Proceedings of a workshop held on Meer
-Island, Torres Strait 14-15 September 2009}. 58 pp. \pageref{Mellors09b}
+Mellors J.\index[people]{Mellors J.} and McKenzie L.\index[people]{McKenzie L.J.} 2009(b). \emph{Seagrass-Watch: Seagrass Monitoring Guidelines
+for Torres Strait Communities. Meer Island. Proceedings of a workshop held on Meer Island, Torres Strait 14-15 September 2009}. 58 pp. \pageref{Mellors09b}
 \bibitem{Micheli05}
-Micheli C., Paganin P, Peirano A., G. Caye, Meinesz A. and Bianchi C.N. 2005. \emph{Genetic
-variability of \textit{{Posidonia oceanica}} (L.) Delile in relation to local factors and biogeographic
-patterns}. Aquatic Botany 82 (2005) 210–221 \pageref{Micheli05}
+Micheli C.\index[people]{Micheli C.}, Paganin P.\index[people]{Paganin P.}, Peirano A.\index[people]{Peirano A.}, Caye, G.\index[people]{Caye, G.}, Meinesz A.\index[people]{Meinesz A.} and Bianchi C.N.\index[people]{Bianchi C.N.} 2005.
+\emph{Genetic variability of \textit{{Posidonia oceanica}} (L.) Delile in relation to local factors and biogeographic patterns}.
+Aquatic Botany 82 (2005) 210–221 \pageref{Micheli05}
 \bibitem{Montefalcone09}
-Montefalcone M. 2009.\emph{Ecosystem health assessment using the Mediterranean seagrass
-\textit{{Posidonia oceanica}}: A review}. Ecological indicators 595 – 604 \pageref{Montefalcone09}
+Montefalcone M.\index[people]{Montefalcone M.} 2009.\emph{Ecosystem health assessment using the Mediterranean seagrass
+\textit{Posidonia oceanica}: A review}. Ecological indicators 595 – 604 \pageref{Montefalcone09}
 \bibitem{Montefalcone10}
-Montefalcone M., Albertelli G., Morri C. and Bianchi C.N. 2010. \emph{Patterns of wide-scale
-substitution within meadows of the seagrass \textit{{Posidonia oceanica}} in NW Mediterranean Sea:
+Montefalcone M.\index[people]{Montefalcone M.}, Albertelli G.\index[people]{Albertelli G.}, Morri C.\index[people]{Morri C.} and Bianchi C.N.\index[people]{Bianchi C.N.} 2010.
+\emph{Patterns of wide-scale substitution within meadows of the seagrass \textit{{Posidonia oceanica}} in NW Mediterranean Sea:
 invaders are stronger than natives}. Aquatic Conserv: Mar. Freshw. Ecosyst. 20: 507–515. \pageref{Montefalcone10}
 \bibitem{Montefalcone06}
-Montefalcone M., Albertelli G., Bianchi C.N., Mariani M., Morri C. 2006. \emph{A new
-synthetic index and a protocol for monitoring the status of \textit{{Posidonia oceanica}} meadows: a
-case study at Sanremo (Ligurian Sea, NW Mediterranean)}. Aquatic Conservation: Marine and
-Freshwater Ecosystems, 16, 29–42. \pageref{Montefalcone06}
+Montefalcone M.\index[people]{Montefalcone M.}, Albertelli G.\index[people]{Albertelli G.}, Bianchi C.N.\index[people]{Bianchi C.N.}, Mariani M.\index[people]{Mariani M.}, Morri C.\index[people]{Morri C.} 2006.
+\emph{A new synthetic index and a protocol for monitoring the status of \textit{{Posidonia oceanica}} meadows: a case study at Sanremo (Ligurian Sea, NW Mediterranean)}.
+Aquatic Conservation: Marine and Freshwater Ecosystems, 16, 29–42.
+\pageref{Montefalcone06}
 \bibitem{Mount03}
-Mount R. 2003. \emph{The application of digital aerial photography to shallow water seabed
-mapping and monitoring – how deep can you see? Coastal GIS 2003: and integrated
-approach to Australian coastal issues, 7th - 8th July 2003}. Wollongong, University of
-Wollongong, pp. 139-156. \pageref{Mount03}
+Mount R.\index[people]{Mount R.} 2003. \emph{The application of digital aerial photography to shallow water seabed
+mapping and monitoring – how deep can you see?
+Coastal GIS 2003: and integrated approach to Australian coastal issues, 7th - 8th July 2003}.
+Wollongong, University of Wollongong, pp.
+139-156.
+\pageref{Mount03}
 \bibitem{Mumby99}
-Mumby P.G., Green E.P., Edwards A.J. and Clark C.D. 1999. \emph{The cost-effectiveness of
-remote sensing for tropical coastal resources assessment and management}. Journal of
-Environmental Management Volume 55, Issue 3, March 1999, Pages 157-166 \pageref{Mumby99}
+Mumby P.J.\index[people]{Mumby P.J.}, Green E.P.\index[people]{Green E.P.}, Edwards A.J.\index[people]{Edwards A.J.} and Clark C.D.\index[people]{Clark C.D.} 1999.
+\emph{The cost-effectiveness of remote sensing for tropical coastal resources assessment and management}.
+Journal of Environmental Management Volume 55, Issue 3, March 1999, Pages 157-166 \pageref{Mumby99}
 \bibitem{Mumby02}
-Mumby, P. J. and Edwards, A. J. 2002. \emph{Mapping marine environments with IKONOS
-imagery: Enhanced spatial resolution can deliver greater thematic accuracy}. Remote
-Sensing of Environment, 82, 248–257 \pageref{Mumby02}
+Mumby, P.J.\index[people]{Mumby P.J.} and Edwards A.J.\index[people]{Edwards A.J.} 2002. \emph{Mapping marine environments with IKONOS
+imagery: Enhanced spatial resolution can deliver greater thematic accuracy}. Remote Sensing of Environment, 82, 248–257 \pageref{Mumby02}
 \bibitem{Mumby04}
-Mumby P.J. Skirving W., Strong A.E., Hardy J.T., LeDrew E.F., Hochberg E.J., Stumpf
-R.P. and David L.T. 2004. \emph{Remote sensing of coral reefs and their physical environment}.
-Marine Pollution Bulletin, 48:219-228. \pageref{Mumby04}
+Mumby P.J.\index[people]{Mumby P.J.}, Skirving W.\index[people]{Skirving W.}, Strong A.E.\index[people]{Strong A.E.}, Hardy J.T.\index[people]{Hardy J.T.}, LeDrew E.F.\index[people]{LeDrew E.F.}, Hochberg E.J.\index[people]{Hochberg E.J.}, Stumpf R.P.\index[people]{Stumpf R.P.} and David L.T.\index[people]{David L.T.} 2004.
+\emph{Remote sensing of coral reefs and their physical environment}.
+Marine Pollution Bulletin, 48:219-228.
+\pageref{Mumby04}
 \bibitem{Nagelkerken00}
-Nagelkerken I., van der Velde G., Gorissen M.W., Meijer G.J., van’t Hofc T. and den
-Hartog C. 2000. \emph{Importance of Mangroves, Seagrass Beds and the Shallow Coral Reef as a
-Nursery for Important Coral Reef Fishes, Using a Visual Census Technique}. Estuarine,
+Nagelkerken I.\index[people]{Nagelkerken I.}, van der Velde G.\index[people]{van der Velde G.}, Gorissen M.W.\index[people]{Gorissen M.W.}, Meijer G.J.\index[people]{Meijer G.J}, van’t Hofc T.\index[people]{van’t Hofc T.} and Den Hartog C.\index[people]{Den Hartog C.} 2000.
+\emph{Importance of Mangroves, Seagrass Beds and the Shallow Coral Reef as a Nursery for Important Coral Reef Fishes, Using a Visual Census Technique}.
+Estuarine,
 Coastal and Shelf Science 51, 31–44 doi:10.1006/ecss.2000.0617 \pageref{Nagelkerken00}
 \bibitem{Noralez10}
-Noralez S. 2010. \emph{Exploring the Limits of Mapping Seagrass. A case study for P.
-oceanica}. MSc Thesis. ITC. Netherlands. \pageref{Noralez10}
+Noralez S.\index[people]{Noralez S.} 2010. \emph{Exploring the Limits of Mapping Seagrass. A case study for P. oceanica}. MSc Thesis. ITC. Netherlands. \pageref{Noralez10}
 \bibitem{Norris97}
-Norris J.G., Wyllie-Echeverria S., Mumford T., Bailey A. and Turner T. 1997.
+Norris J.G.\index[people]{Norris J.G.}, Wyllie-Echeverria S.\index[people]{Wyllie-Echeverria S.}, Mumford T.\index[people]{Mumford T.}, Bailey A.\index[people]{Bailey A.} and Turner T.\index[people]{Turner T.} 1997.
 \emph{Estimating basal area coverage of subtidal seagrass beds using underwater videography}.
@@
 \bibitem{Orth06}
-Orth R. J., Carruthers T.J.B., Dennison W.C., Duarte C. M., Fourqurean J.W. , Heck Jr.
-K. L. , Hughes A. R., Kendrick G. A., Kenworthy W. J., Olyarnik S., Short F.T., Waycott
-M. and Williams S. L. 2006. \emph{A Global Crisis for Seagrass Ecosystems}. Vol. 56 No. 12,
-BioScience 987. \pageref{Orth06}
+Orth R.J.\index[people]{Orth R.J.}, Carruthers T.J.B.\index[people]{Carruthers T.J.B.}, Dennison W.C.\index[people]{Dennison W.C.}, Duarte C.M.\index[people]{Duarte C.M.}, Fourqurean J.W.\index[people]{Fourqurean J.W.}, Heck Jr.K.
+L.\index[people]{Heck Jr.K.
+L.}, Hughes A.
+R.\index[people]{Hughes A.
+R.}, Kendrick G.
+A.\index[people]{Kendrick G.A.}, Kenworthy W.
+J.\index[people]{Kenworthy W.
+J.}, Olyarnik S.\index[people]{Olyarnik S.}, Short F.T.\index[people]{Short F.T}, Waycott M.\index[people]{Waycott M.} and Williams S.
+L.\index[people]{Williams S.L.} 2006.
+\emph{A Global Crisis for Seagrass Ecosystems}.
+Vol.
+56 No.
+12, BioScience 987.
+\pageref{Orth06}
 \bibitem{Palandro03}
-Palandro D., Andréfouët S., Muller-Karger F. E., Dustan Ph., Hu Ch., and Hallock P.
-2003. \emph{Detection of changes in coral reef communities using Landsat-5 TM and Landsat-7
-ETM+ data}. Can. J. Remote Sensing, Vol. 29, No. 2, pp. 201–209. \pageref{Palandro03}
+Palandro D.\index[people]{Palandro D.}, Andr\'efou\"et S.\index[people]{Andr\'efou\"et S.}, Muller-Karger F.
+E.\index[people]{Muller-Karger F.
+E.}, Dustan Ph.\index[people]{Dustan Ph}, Hu Ch.\index[people]{Hu Ch.}, and Hallock P.\index[people]{Hallock P.} 2003.
+\emph{Detection of changes in coral reef communities using Landsat-5 TM and Landsat-7 ETM+ data}.
+Can.
+J.
+Remote Sensing, Vol.
+29, No.
+2, pp.
+201–209.
+\pageref{Palandro03}
 \bibitem{Pasqualini98a}
-Pasqualini V., Pergent-Martini C. and Pergent, G. 1998(a). \emph{Use of remote sensing for the
-characterization of the Mediterranean coastal environment – the case of Posidonia
+Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.} and Pergent, G.\index[people]{Pergent G.} 1998(a).
+\emph{Use of remote sensing for the characterization of the Mediterranean coastal environment – the case of Posidonia
 oceanica}. Journal of Coastal Conservation 4: 59-66, 1998, EUCC; Opulus Press Uppsala. \pageref{Pasqualini98a}
 \bibitem{Pasqualini98b}
-Pasqualini V., Pergent-Martini C., Clabaut P., Pergent G. 1998(b). \emph{Mapping of
-\textit{{Posidonia oceanica}} using aerial photographs and side scan sonar: Application off the
+Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Clabaut P.\index[people]{Clabaut P.}, Pergent G.\index[people]{Pergent G.} 1998(b).
+\emph{Mapping of \textit{{Posidonia oceanica}} using aerial photographs and side scan sonar: Application off the
 island of Corsica (France)}. Estuarine Coastal Shelf Sci 47:359–367. \pageref{Pasqualini98b}
 \bibitem{Pasqualini99}
-Pasqualini V, Pergent-Martini C, Pergent G. 1999. \emph{Environmental impact identification
-along the Corsican coast (Mediterranean Sea) using image processing}. Aquatic Botany,
+Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pergent G.\index[people]{Pergent G.} 1999.
+\emph{Environmental impact identification along the Corsican coast (Mediterranean Sea) using image processing}.
+Aquatic Botany,
 65:311-320. \pageref{Pasqualini99}
 \bibitem{Pasqualini01}
-Pasqualini V., Pergent-Martini C, Clabaut P, Marteel H., and Pergent G. 2001.
+Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Clabaut P.\index[people]{Clabaut P.}, Marteel H.\index[people]{Marteel H.}, and Pergent G.\index[people]{Pergent G.} 2001.
 \emph{Integration of Aerial Remote Sensing, Photogrammetry, and GIS Technologies in Seagrass
-Mapping}. Photogrammetric Engineering and Remote Sensing Vol. 67, No. 1, January 2001,
-pp. 99-105. \pageref{Pasqualini01}
+Mapping}. Photogrammetric Engineering and Remote Sensing Vol. 67, No. 1, January 2001, pp. 99-105. \pageref{Pasqualini01}
 \bibitem{Pasqualini05}
-Pasqualini V., Pergent-Martini C, Pergent G., Agreila M., Skoufas G., Sourbes L. and
-Tsirika A. 2005. \emph{Use of SPOT 5 for mapping seagrasses: An application to Posidonia
-oceanica}. Remote Sensing of Environment, 94:39-45. \pageref{Pasqualini05}
+Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pergent G.\index[people]{Pergent G.}, Agreila M.\index[people]{Agreila M.}, Skoufas G.\index[people]{Skoufas G.}, Sourbes L.\index[people]{Sourbes L.} and Tsirika A.\index[people]{Tsirika A.} 2005.
+\emph{Use of SPOT 5 for mapping seagrasses: An application to Posidonia oceanica}.
+Remote Sensing of Environment, 94:39-45.
+\pageref{Pasqualini05}
 \bibitem{Peneva08}
-Peneva E., Griffith J.A. and Carter G.A. 2008. \emph{Seagrass Mapping in the Northern Gulf
-of Mexico using Airborne Hyperspectral Imagery: A Comparison of Classification Methods}.
-Journal of Coastal Research, Number 244:850-856. 2008. Published By: Coastal Education
-and Research Foundation DOI: 10.2112/06-0764.1 \pageref{Peneva08}
+Peneva E.\index[people]{Peneva E.I.}, Griffith J.A.\index[people]{Griffith J.A.} and Carter G.A.\index[people]{Carter G.A.} 2008.
+\emph{Seagrass Mapping in the Northern Gulf of Mexico using Airborne Hyperspectral Imagery: A Comparison of Classification Methods}.
+Journal of Coastal Research, Number 244:850-856.
+2008.
+Published By: Coastal Education and Research Foundation DOI: 10.2112/06-0764.1 \pageref{Peneva08}
 \bibitem{Peneva07}
-Peneva E.I., Griffith J.A. and Carter G.A. 2007. \emph{Mapping seagrass distribution on Horn
-Island}, MS using hyperspectral imagery.
+Peneva E.I.\index[people]{Peneva E.I.}, Griffith J.A.\index[people]{Griffith J.A.} and Carter G.A.\index[people]{Carter G.A.} 2007.
+\emph{Mapping seagrass distribution on Horn Island}, MS using hyperspectral imagery.
 \href{ftp://ftp.ecn.purdue.edu/jshan/proceedings/asprs2005/Files/0134.pdf}{ecn.purdue.edu} \pageref{Peneva07}
 \bibitem{Pergent90}
-Pergent G. 1990. \emph{Lepidochronological analysis of the seagrass \textit{{Posidonia oceanica}} (L.)
-Delile: a standardized approach}. Aquatic Botany, 37 (1990) 39-54 39 Elsevier Science
-Publishers B.V., Amsterdam -- Printed in The Netherlands \pageref{Pergent90}
+Pergent G.\index[people]{Pergent G.} 1990. \emph{Lepidochronological analysis of the seagrass \textit{{Posidonia oceanica}} (L.)
+Delile: a standardized approach}. Aquatic Botany, 37 (1990) 39-54 39 Elsevier Science Publishers B.V., Amsterdam -- Printed in The Netherlands \pageref{Pergent90}
 \bibitem{Pergent-Martini06}
-Pergent-Martini C., Pasqualini V., Ferrat L and Pergent G. 2006. \emph{Ecological Data in
-Integrated Coastal Zone Management: Case Study of \textit{{Posidonia oceanica}} Meadows Along
+Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pasqualini V.\index[people]{Pasqualini V.}, Ferrat L.\index[people]{Ferrat L.} and Pergent G.\index[people]{Pergent G.} 2006.
+\emph{Ecological Data in Integrated Coastal Zone Management: Case Study of \textit{{Posidonia oceanica}} Meadows Along
 the Corsican Coastline (Mediterranean Sea)}. Environ Manage (2006) 38:889–895 DOI
@@
 \bibitem{Pergent-Martini05}
-Pergent-Martini C, Leoni V., Pasqualini V., Ardizzone G.D., Balestri E., Bedini R.,
-Belluscio A., Belsher T., Borg J., Boudouresque C.F., Boumaza S., Bouquegneau J.M., Buia
-M.C., Calvo S., Cebrian J., Charbonnel E., Cinelli F., Cossu A., Di Maida G., Dural B.,
-Francour P., Gobert S., Lepoint G., Meinesz A., Molenaar H., Mansour H.M., Panayotidis
-P., Peirano A., Pergent G., Piazzi L, Pirrotta M, Relini G, Romero G, Sanchez-Lizaso J.L,
-Semroud R., Shembri P., Shili A., Tomasello A., Velimirov B. 2005. \emph{Descriptors of
-\textit{{Posidonia oceanica}} meadows: Use and application}. Ecological Indicators 5, 213–230. \pageref{Pergent-Martini05}
+Pergent-Martini C.\index[people]{Pergent-Martini C.}, Leoni V.\index[people]{Leoni V.}, Pasqualini V.\index[people]{Pasqualini V.}, Ardizzone G.D.\index[people]{Ardizzone G.D.}, Balestri E.\index[people]{Balestri E.}, Bedini R.\index[people]{Bedini R.}, Belluscio A.\index[people]{Belluscio A.}, Belsher T.\index[people]{Belsher T.}, Borg J.\index[people]{Borg J.}, Boudouresque C.F.\index[people]{Boudouresque C.F.}, Boumaza S.\index[people]{Boumaza S.}, Bouquegneau J.-M.\index[people]{Bouquegneau J.-M.}, Buia M.C.\index[people]{Buia M.C.}, Calvo S.\index[people]{Calvo S.}, Cebrian J.\index[people]{Cebrian J.}, Charbonnel E.\index[people]{Charbonnel E.}, Cinelli F.\index[people]{Cinelli F.}, Cossu A.\index[people]{Cossu A.}, Di Maida G.\index[people]{Di Maida G.}, Dural B.\index[people]{Dural B.}, Francour P.\index[people]{Francour P.}, Gobert S.\index[people]{Gobert S.}, Lepoint G.\index[people]{Lepoint G.}, Meinesz A.\index[people]{Meinesz A.}, Molenaar H.\index[people]{Molenaar H.}, Mansour H.M.\index[people]{Mansour H.M.}, Panayotidis P.\index[people]{Panayotidis P.}, Peirano A.\index[people]{Peirano A.}, Pergent G.\index[people]{Pergent G.}, Piazzi L.\index[people]{Piazzi L.}, Pirrotta M.\index[people]{Pirrotta M.}, Relini G.\index[people]{Relini G}, Romero J.\index[people]{Romero J.}, Sanchez-Lizaso J.L.\index[people]{Sanchez-Lizaso J.L.}, Semroud R.\index[people]{Semroud R.}, Shembri P.\index[people]{Shembri P.}, Shili A.\index[people]{Shili A.}, Tomasello A.\index[people]{Tomasello A.}, Velimirov B.\index[people]{Velimirov B.} 2005.
+\emph{Descriptors of \textit{{Posidonia oceanica}} meadows: Use and application}.
+Ecological Indicators 5, 213–230.
+\pageref{Pergent-Martini05}
 \bibitem{Phinn08}
-Phinn S, Roelfsema C, Dekker A, Brando V, Anstee J. 2008. \emph{Mapping seagrass species,
-cover and biomass in shallow waters: An assessment of satellite multi-spectral and airborne
-hyperspectral imaging systems in Moreton Bay (Australia)}. Remote Sensing of
-Environment, 112:3413-3425. \pageref{Phinn08}
+Phinn S.\index[people]{Phinn S.}, Roelfsema C.\index[people]{Roelfsema C.}, Dekker A.\index[people]{Dekker A.G.}, Brando V.E.\index[people]{Brando V.E.}, Anstee J.M.\index[people]{Anstee J.M.} 2008.
+\emph{Mapping seagrass species, cover and biomass in shallow waters: An assessment of satellite multi-spectral and airborne
+hyperspectral imaging systems in Moreton Bay (Australia)}. Remote Sensing of Environment, 112:3413-3425. \pageref{Phinn08}
 \bibitem{Ralph05}
-Ralph P.J., Macinnis-Ng C.M.O., Frankart C. 2005. \emph{Fluorescence imaging application:
-effect of leaf age on seagrass photokinetics}. Aquatic Botany 81 (2005) 69–84. \pageref{Ralph05}
+Ralph P.J.\index[people]{Ralph P.J.}, Macinnis-Ng C.M.O.\index[people]{Macinnis-Ng C.M.O.}, Frankart C.\index[people]{Frankart C.} 2005.
+\emph{Fluorescence imaging application: effect of leaf age on seagrass photokinetics}.
+Aquatic Botany 81 (2005) 69–84.
+\pageref{Ralph05}
 \bibitem{Rasib97}
-Rasib A.W. and Hashim M. 1997. \emph{Mapping Seagrass From Satellite Remote Sensing
+Rasib A.W.\index[people]{Rasib A.W.} and Hashim M.\index[people]{Hashim M.} 1997. \emph{Mapping Seagrass From Satellite Remote Sensing
 Data. Proceedings of Asian Conference on Remote Sensing}. \href{http://eprints.utm.my/2161/}{eprints.utm.my} \pageref{Rasib97}
 \bibitem{Ribed02}
-Ribed P.S. 2002. \emph{Submarine vegetation studies using remote sensing: importance in
-beach measurements}. Journal of Environmental Protection and Ecology, 3 No 4, 953-958
-(2002) \pageref{Ribed02}
+Ribed P.S.\index[people]{Ribed P.S.} 2002.
+\emph{Submarine vegetation studies using remote sensing: importance in beach measurements}.
+Journal of Environmental Protection and Ecology, 3 No 4, 953-958 (2002) \pageref{Ribed02}
 \bibitem{Richardson06}
-Richardson L.L., LeDrew Ellesworth F. 2006. \emph{Remote Sensing of Aquatic Costal
+Richardson L.L.\index[people]{Richardson L.L.}, LeDrew Ellesworth F.\index[people]{LeDrew E.F.} 2006. \emph{Remote Sensing of Aquatic Costal
 Ecosystems Processes}. Springer. Science and Management Applications. \pageref{Richardson06}
 \bibitem{Robbins94}
-Robbins B.D., Bell S.S. 1994. \emph{Seagrass landscapes: a terrestrial approach to the marine
+Robbins B.D.\index[people]{Robbins B.D.}, Bell S.S.\index[people]{Bell S.S.} 1994. \emph{Seagrass landscapes: a terrestrial approach to the marine
 subtidal environment}. TREE vol.9, no 8, doi:10.1016/0169-5347(94)90041-8 \pageref{Robbins94}
 \bibitem{Sabins97}
-Sabins F. 1997. \emph{Remote Sensing, Principles and Interpretation}. W.H. Freeman and
+Sabins F.\index[people]{Sabins F.} 1997. \emph{Remote Sensing, Principles and Interpretation}. W.H. Freeman and
 Company, New York. \pageref{Sabins97}
 \bibitem{Schmidt03}
-Schmidt K. and Skidmore A.K. 2003. \emph{Spectral discrimination of vegetation types in a
+Schmidt K.\index[people]{Schmidt K.} and Skidmore A.K.\index[people]{Skidmore A.K.} 2003. \emph{Spectral discrimination of vegetation types in a
 coastal wetland}. Remote Sensing of Environment. 85:92-108. \pageref{Schmidt03}
 \bibitem{Schmieder97}
-Schmieder K. 1997. \emph{Littoral zone - GIS of Lake Constance: a useful tool in lake
+Schmieder K.\index[people]{Schmieder K.} 1997. \emph{Littoral zone - GIS of Lake Constance: a useful tool in lake
 monitoring and autecological studies with submersed macrophytes}. Aquatic Botany 58
@@
 \bibitem{Schultz08}
-Schultz S.T. 2008. \emph{Seagrass monitoring by underwater videography: Disturbance
+Schultz S.T.\index[people]{Schultz S.T.} 2008. \emph{Seagrass monitoring by underwater videography: Disturbance
 regimes, sampling design, and statistical power}. Aquatic Botany 88 (2008) 228–238. \pageref{Schultz08}
 \bibitem{Short01}
-Short F.T. and Coles R.G. (eds). 2001. \emph{Global Seagrass Research Methods}, p 482.
+Short F.T.\index[people]{Short F.T.} and Coles R.G.\index[people]{Coles R.G.} (eds). 2001. \emph{Global Seagrass Research Methods}, p 482.
 Elsevier Publishing Co., Amsterdam \pageref{Short01}
 \bibitem{Shortis07}
-Shortis M., Harvey E., Seager J. 2007. \emph{A Review of the Status and Trends in Underwater
-Videometric Measurement}. SPIE Conference 6491, Videometrics IX,IS and T/SPIE Electronic
-Imaging, San Jose, California, U.S.A., 28 January–1 February 2007 \pageref{Shortis07}
+Shortis M.\index[people]{Shortis M.}, Harvey E.\index[people]{Harvey E.}, Seager J.\index[people]{Seager J.} 2007.
+\emph{A Review of the Status and Trends in Underwater Videometric Measurement}.
+SPIE Conference 6491, Videometrics IX,IS and T/SPIE Electronic Imaging, San Jose, California, U.S.A., 28 January–1 February 2007 \pageref{Shortis07}
 \bibitem{Stephens03}
-Stephens F.C., Louchard E.M., Reid R.P., Maffione R.A. 2003. \emph{Effects of microalgal
-communities on reflectance spectra of carbonate sediments in subtidal optically shallow
+Stephens F.C.\index[people]{Stephens F.C.}, Louchard E.M.\index[people]{Louchard E.M.}, Reid R.P.\index[people]{Reid R.P.}, Maffione R.A.\index[people]{Maffione R.A.} 2003.
+\emph{Effects of microalgal communities on reflectance spectra of carbonate sediments in subtidal optically shallow
 marine environments}. Limnology and Oceanography, 48:535-546. \pageref{Stephens03}
 \bibitem{Thorhaug07}
-Thorhaug, A., Richardson, A. D. and Berlyn, G. P. 2007. \emph{Spectral reflectance of the
-seagrasses: Thalassia testudinum, Halodule wrightii, Syringodium filiforme and five marine
+Thorhaug, A.\index[people]{Thorhaug, A.}, Richardson, A.
+D.\index[people]{Richardson, A.D.} and Berlyn, G.P.\index[people]{Berlyn, G.P.} 2007.
+\emph{Spectral reflectance of the seagrasses: Thalassia testudinum, Halodule wrightii, Syringodium filiforme and five marine
 algae}. International Journal of Remote Sensing, 28: 7, 1487 — 1501. \pageref{Thorhaug07}
 \bibitem{Toole00}
-Toole D.A., Siegel D.A., Menzies D.W., Neumann M.J. and Smith R.C. 2000. 
-\emph{Remote-sensing reflectance determinations in the coastal ocean
-environment: impact of instrumental characteristics and environmental variability}. Applied
-Optics 39(3), 456-469. \pageref{Toole00}
+Toole D.A.\index[people]{Toole D.A.}, Siegel D.A.\index[people]{Siegel D.A.}, Menzies D.W.\index[people]{Menzies D.W.}, Neumann M.J.\index[people]{Neumann M.J.} and Smith R.C.\index[people]{Smith R.C.} 2000.
+\emph{Remote-sensing reflectance determinations in the coastal ocean environment: impact of instrumental characteristics and environmental variability}.
+Applied Optics 39(3), 456-469.
+\pageref{Toole00}
 \bibitem{Vahtmae06}
-Vahtmae E, Kutser T, Martin G, Kotta J. 2006. \emph{Feasibility of hyperspectral remote
-sensing for mapping benthic macroalgal cover in turbid coastal waters - a Baltic Sea case
+Vahtmae E.\index[people]{Vahtmae E.}, Kutser T.\index[people]{Kutser T.}, Martin G.\index[people]{Martin G.}, Kotta J.\index[people]{Kotta J.}, 2006.
+\emph{Feasibility of hyperspectral remote sensing for mapping benthic macroalgal cover in turbid coastal waters - a Baltic Sea case
 study. Remote Sensing of Environment} (2006) 101:342-351. \pageref{Vahtmae06}
 \bibitem{Vahtmae07}
-Vahtmae E. and Kutser T. 2007. \emph{Mapping bottom type and water depth in shallow
-coastal waters with satellite remote sensing}. Journal of Coastal Research, SI 50 (Proceedings
-of the 9th International Coastal Symposium), 185 – 189. Gold Coast, Australia, ISSN
-0749.0208 \pageref{Vahtmae07}
+Vahtmae E.\index[people]{Vahtmae E.} and Kutser T.\index[people]{Kutser T.} 2007. \emph{Mapping bottom type and water depth in shallow
+coastal waters with satellite remote sensing}.
+Journal of Coastal Research, SI 50 (Proceedings of the 9th International Coastal Symposium), 185 – 189.
+Gold Coast, Australia, ISSN 0749.0208 \pageref{Vahtmae07}
 \bibitem{Villegas06}
-Villegas P. P. 2006. \emph{Magnitude of herbivory in \textit{Posidonia oceanica} (L.) Delile and
+Villegas P. P.\index[people]{Villegas P. P.} 2006. \emph{Magnitude of herbivory in \textit{Posidonia oceanica} (L.) Delile and
 factors responsible for spatial variation}. Available on web: \href{http://www.tesisenxarxa.net/}{tesisenxarxa} \pageref{Villegas06}
 \bibitem{Wabnitz08}
-Wabnitz C.C., Andréfouët S., Torres-Pulliza D., M\"uller-Karger F. E. and Kramer P.A.
-2008. \emph{Regional-scale seagrass habitat mapping in the Wider Caribbean region using
-Landsat sensors: Applications to conservation and ecology}. Remote Sensing of
-Environment Vol. 112, Issue 8, 15 August 2008, Pages 3455-3467 Earth Observations for
-Marine and Coastal Biodiversity and Ecosystems Special Issue. \pageref{Wabnitz08}
+Wabnitz C.C.\index[people]{Wabnitz C.C.}, Andr\'efou\"et S.\index[people]{Andr\'efou\"et S.}, Torres-Pulliza D.\index[people]{Torres-Pulliza D.}, M\"uller-Karger F.
+E.\index[people]{M\"uller-Karger F.
+E.} and Kramer P.A.\index[people]{Kramer P.A.}, 2008.
+\emph{Regional-scale seagrass habitat mapping in the Wider Caribbean region using Landsat sensors: Applications to conservation and ecology}.
+Remote Sensing of Environment Vol.
+112, Issue 8, 15 August 2008, Pages 3455-3467 Earth Observations for Marine and Coastal Biodiversity and Ecosystems Special Issue.
+\pageref{Wabnitz08}
 \bibitem{Walker89}
-Walker, D.I., 1989. \emph{Methods for monitoring seagrass habitat}. Proceedings Rep.
-Workshop, 20-22 June 1988, Melbourne, Australia, VIMS Working Paper 18, pp. 1-26. \pageref{Walker89}
+Walker, D.I.\index[people]{Walker, D.I.}, 1989.
+\emph{Methods for monitoring seagrass habitat}.
+Proceedings Rep.
+Workshop, 20-22 June 1988, Melbourne, Australia, VIMS Working Paper 18, pp.
+1-26.
+\pageref{Walker89}
 \bibitem{Werdell03}
-Werdell P.J. and Roesler C.S. 2003. \emph{Remote assessment of benthic substrate composition
-in shallow waters using multispectral reflectance}. Limnol. Oceanogr., 48 (1, part 2), 557–
-567 \pageref{Werdell03}
+Werdell P.J.\index[people]{Werdell P.J.} and Roesler C.S.\index[people]{Roesler C.S.} 2003. \emph{Remote assessment of benthic substrate composition
+in shallow waters using multispectral reflectance}. Limnol. Oceanogr., 48 (1, part 2), 557–567 \pageref{Werdell03}
 \bibitem{Yang10}
-Yang Ch., Yang D., Cao W., Zhao J., Wang G., Sun Zh., Xu Zh., Kumar M. S. R. 2010. 
-\emph{Analysis of seagrass reflectivity by using a water column correction algorithm}. International
-Journal of Remote Sensing. 31: 17, 4595 — 4608. \pageref{Yang10}
+ Yang Ch.\index[people]{Yang Ch.}, Yang D.\index[people]{Yang D.}, Cao W.\index[people]{Cao W.}, Zhao J.\index[people]{Zhao J.}, Wang G.\index[people]{Wang G.}, Sun Zh.\index[people]{Sun Zh.}, Xu Zh.\index[people]{Xu Zh.}, Kumar M.
+S.
+R.\index[people]{Kumar M.
+S.
+R.} 2010.
+\emph{Analysis of seagrass reflectivity by using a water column correction algorithm}. International Journal of Remote Sensing. 31: 17, 4595 — 4608. \pageref{Yang10}
 
 \end{thebibliography}
+
+%\cleardoublepage
+\newpage
+\singlespacing
+%\section*{List of People}
+%\markright{List of People}
+%\addcontentsline{toc}{section}{List of People}
+%\indexsetup{level=\section,toclevel=section}
+\printindex[people]
+\pagebreak
 
@@
 \newpage
-\section*{List of People}
-\markright{List of People}
-\addcontentsline{toc}{section}{List of People}
-\pagebreak
-
-\singlespacing
-\cleardoublepage
-\newpage
-\phantomsection
-\addcontentsline{toc}{section}{Index}
-\printindex
+%\phantomsection
+%\addcontentsline{toc}{section}{Index}
+\printindex[concepts]
 

### 2011-02-05 MScDiplom_Lemenkova.tex
changed page layout to ISO-216 PA5 (140*210mm)

@@
-\documentclass[11pt, a4paper]{article}
+\documentclass[11pt]{article}
+\special{papersize=140mm,210mm}
 \usepackage{authblk}
@@
 \usepackage[printonlyused,withpage]{acronym}
-\usepackage[a4paper, left=30mm,right=25mm,top=30mm,bottom=35mm]{geometry}
+\usepackage[left=30mm,right=20mm,top=30mm,bottom=35mm]{geometry}
 %\usepackage{glossary}
@@
 \restylefloat{figure}
-\usepackage[a4paper=true, ]{hyperref}
+\usepackage{hyperref}
 \hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue}
@@
 
-\textbf{Course Duration:} \hfill{September 2009 – March 2011}\\
+\textbf{Course Duration:} \hfill{September 2009 $-$ March 2011}\\
 
@@
 	\item[*] \href{http://www.uw.edu.pl/en/}{University of Warsaw} (Poland)\\
-	\item[*] \href{http://www.utwente.nl/en}{University of Twente}, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)} (The Netherlands)
+	\item[*] \href{http://www.utwente.nl/en}{University of Twente}, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)} \\ (The Netherlands)
 \end{itemize}
@@
 \textit{Valentijn Venus}\index[people]{Venus, V.} from the ITC, Department of Natural Resources, for our successful scientific collaboration, help and discussions of my research work.
-\\
 I also express them my most sincere appreciations for our fieldwork on Crete.
@@
 	\acro{GRASS}{\href{http://grass.fbk.eu/}{Geographic Resources Analysis Support System}}
-	%\acro{GRETL}{GNU Regression, Econometrics and Time-series Library}
+	\acro{GRETL}{\href{http://gretl.sourceforge.net/}{GNU Regression, Econometrics and Time-series Library}}
 	\acro{HyMap}{\href{http://www.hyvista.com/}{Hyperspectral Mapper}}
@@
 
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-1-2.jpg}
-	\caption{Distribution of seagrasses in the world.\\ Source:\cite{Green03}\label{Green03} 
+\begin{wrapfigure}{r}{0.45\textwidth}
+	\centering
+	\includegraphics[scale=0.40]{Fig-1-2.jpg}
+	\caption{Distribution of seagrasses in the world. Source:\cite{Green03}\label{Green03} 
 	\label{fig:2}} 
@@
 
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-1-3.jpg}
+\begin{wrapfigure}{r}{0.35\textwidth}
+	\centering
+	\includegraphics[scale=0.3]{Fig-1-3.jpg}
 	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
@@
 capacity of the plant to adapt to these changes, reduce its distribution area \cite{Micheli05}\label{Micheli05}. 
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-1-4.jpg}
-	\caption{Distribution of seagrass in relation to mean ocean temperature. \\
-	Source: \cite{Orth06}\label{Orth06}}
-	\label{Fig.4}
-\end{wrapfigure}
-
 One of the main drivers of seagrass decline is, for instance, the location of the fish farming near the
@@
 seagrass are recreational boating, commercial overexploitation of coastal resources, eutrophication
-\cite{McKenzie06}\label{McKenzie06}.\\
+\cite{McKenzie06}\label{McKenzie06}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-1-4.jpg}
+	\caption{Distribution of seagrass in relation to mean ocean temperature. \\
+	Source: \cite{Orth06}\label{Orth06}}
+	\label{Fig.4}
+\end{wrapfigure}
+
 Besides anthropogenic factors, various biochemical, climatic and environmental processes\index[concepts]{Seagrasses!natural impacts on} can cause
@@
 the evaluating of the seagrass current distribution, analysis of its dynamics and changes over time, as
-well as estimations of the degrading of seagrass meadows for the purpose of the coastal management.\\
+well as estimations of the degrading of seagrass meadows for the purpose of the coastal management.
+
+\begin{wrapfigure}{r}{0.3\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-1-9.jpg}
+	\caption{Spectra of the seagrass on different depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}}
+	\label{fig:9}
+\end{wrapfigure}
+
 Precise, correct and up-to-date information about the distribution of \textit{P.oceanica} is necessary for the
@@
 monitoring. 
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-1-9.jpg}
-	\caption{Spectra of the seagrass on different depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}}
-	\label{fig:9}
-\end{wrapfigure}
-
 The general overview of the application of various remote sensing data types (colour,
@@
 using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01}\label{Louchard03}\label{Dehouck08}\label{Werdell03}.
-Studies of spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and
-other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
-macroalgae differ from each other, as well as from sand and deep water reflectance spectra. These
-differences are well detectable by the means of the remote sensing research methods. Comparing to
-the terrestrial plants, aquatic vegetation, \textit{inter alia} seagrass, cannot be detected using red edge of the
-spectrum, as these wavelengths are significantly absorbed by water \cite{Kirk94}\label{Kirk94}, as well as by
-scattering and absorption by phytoplankton. Some authors \cite{Dierssen03}\label{Dierssen03} also use
-spectrally based radiative transfer approach to quantitatively estimate shallow-water bathymetry and
-leaf area index (\ac{LAI}) of the seagrass. The spectral reflectance in general is the
-result of the spectral absorption in different bands, typical for each target object.
 
@@
 
+Studies of spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and
+other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
+macroalgae differ from each other, as well as from sand and deep water reflectance spectra. These
+differences are well detectable by the means of the remote sensing research methods. Comparing to
+the terrestrial plants, aquatic vegetation, \textit{inter alia} seagrass, cannot be detected using red edge of the
+spectrum, as these wavelengths are significantly absorbed by water \cite{Kirk94}\label{Kirk94}, as well as by
+scattering and absorption by phytoplankton. Some authors \cite{Dierssen03}\label{Dierssen03} also use
+spectrally based radiative transfer approach to quantitatively estimate shallow-water bathymetry and
+leaf area index (\ac{LAI}) of the seagrass. The spectral reflectance in general is the
+result of the spectral absorption in different bands, typical for each target object.
 Spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
@@
 and discrimination of the seafloor types using their spectral characteristics in different wavebands.
+
+\begin{wrapfigure}{r}{0.45\textwidth}
+	\centering
+	\includegraphics[scale=0.35]{Fig-13.jpg}
+ 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
+	\label{fig:2.4}
+\end{wrapfigure}
+
 Perhaps, the most known imagery, widely used in the remote sensing mapping, is received from the
@@
 landscapes.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.35]{Fig-13.jpg}
- 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
-	\label{fig:2.4}
-\end{wrapfigure}
-
 The \ac{Landsat TM}\index[concepts]{Remote sensing!multispectral imagery!Landsat TM} and \ac{Landsat ETM+}\index[concepts]{Remote sensing!multispectral imagery!Landsat ETM+} data, with the most recent from sensor Landsat 7 - an advanced
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{wrapfigure}{r}{0.4\textwidth}
 	\centering
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.5]{GNU-17.jpg}
-	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}				{Gnuplot}}\label{fig:4.2}
-\end{figure}
-\pagebreak
-
 %
@@
 matter), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small
-particle \cite{Gege04}\label{Gege04}. The backward-scattering coefficient (bb), also included in WASI, is a
+particle \cite{Gege04}\label{Gege04}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{GNU-17.jpg}
+	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}				{Gnuplot}}\label{fig:4.2}
+\end{figure}
+
+The backward-scattering coefficient (bb), also included in WASI, is a
 fundamental optical property which plays a central role in the ocean-color remote sensing, providing
@@
 \\
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5, angle=90]{UML_StatAn.jpg}
-	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
-	\label{fig:4.12}
-\end{figure}
-\pagebreak
-
 Thus, we collect a number of realizations (a sample of 25 measurements for both sand and \textit{P.oceanica}) and then estimate the statistical parameters (\nameref{app}). \\
@@
 As can be seen on Fig.\ref{fig:A.21}, Fig.\ref{fig:A.22} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
-The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{Gretl} software.
-\\
+The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{GRETL} software.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5, angle=90]{UML_StatAn.jpg}
+	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
+	\label{fig:4.12}
+\end{figure}
+
 The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made.
@@
 
-\subsubsection{Results of the remote sensing application}
+\subsubsection{Remote sensing application}
 The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
@@
 
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.40]{Fig-36.jpg}
-		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }
-		\label{fig:39}
-	\end{center}
-\end{figure}
-
 \subsection{Seagrass mapping using ArcGIS}
@@
 high resolution aerial color Google Earth photography, spaceborne satellite imagery, assessment of
-spectral signatures using WASI software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
-based mapping. The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
+spectral signatures using WASI software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS based mapping.
+The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
 enables to analyze environmental changes within seagrass landscapes based on data from various
@@
 
-\subsubsection{Upscale mapping of the seagrass landscapes}\index[concepts]{concepts}{Seagrasses!spatial structure}
+\subsubsection{Upscale GIS mapping of the seagrass landscapes}\index[concepts]{Seagrasses!spatial structure}
 Although ocean pelagic landscapes have a high degree of spatial variance and less structurally
@@
 openings in either floral or faunal cover” \cite{Connell78}\label{Connell78}.
-
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.25]{Fig-39.jpg}
-		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project}
-		\label{fig:42}
-	\end{center}
-\end{figure}
-
 Formation and increase of these
@@
 
-\subsubsection{Results of GIS mapping}
-This research part results in series of thematic maps showing seagrass spatial distribution over the Island of
-Crete. The research is based on the aerial images classification and analysis using ArcGIS and Erdas Imagine
-software. The map of the seagrass locations within the research area of Ligaria beach illustrates the seagrass
-distribution at a more detailed scale.
-
-\pagebreak
 \section{Discussion and conclusions}\label{sec:5}
-The suggested methodology of the analysis of seagrass adaptation towards different environments of
-the coastal zones can be applied to other marine coastal areas with growing seagrass.\\
-More than 50 \% of the world population lives within one km of the coast which results in continued
-anthropogenic pressures in coastal regions. \\Therefore, management of coastal resources and shelf
-zone protection in the Mediterranean Sea become increasingly important and require large-scale
-monitoring and mapping of the shelf zone as vital instrument for the environmental assessment.
-\pagebreak
+More than 50 \% of the world population lives within one km of the coast, which results in continued anthropogenic pressure on the coastal regions.
+Therefore, management of coastal resources and shelf zone protection become increasingly important nowadays, and require large-scale monitoring and mapping of the shelf areas as vital instrument for the environmental assessment.
+Besides Mediterranean area, the suggested methodology of the analysis of seagrass monitoring and analysis of itsadaptation towards different environmental conditions within the coastal zones can be applied to other marine coastal areas with growing seagrass.
 
@@
 \end{figure}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.25]{Fig-39.jpg}
+		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project}
+		\label{fig:42}
+	\end{center}
+\end{figure}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.40]{Fig-36.jpg}
+		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }
+		\label{fig:39}
+	\end{center}
+\end{figure}
 \pagebreak
@@
 \emph{Benthic inputs as predictors of seagrass (\textit{Posidonia oceanica}) fish farm-induced decline}.
-\href{http://arxiv.org/ftp/qbio/
-papers/0611/0611006.pdf}{arxiv.org} \pageref{Diaz-Almela06}
+\href{http://arxiv.org/ftp/qbio/papers/0611/0611006.pdf}{arxiv.org} \pageref{Diaz-Almela06}
 \bibitem{Dierssen03}
-Dierssen H.\index[people]{} and Zimmerman R.C.\index[people]{} 2003. \emph{Ocean color remote sensing of seagrass and
+Dierssen H.\index[people]{Dierssen H.} and Zimmerman R.C.\index[people]{Zimmerman R.C.} 2003. \emph{Ocean color remote sensing of seagrass and
 bathymetry in the Bahamas Banks by high-resolution airborne imagery}. Limnol. Oceanogr.,
@@
 \bibitem{Dixon05}
-Dixon R.\index[people]{}, editor. 2005. \emph{Watching the Seagrass Grow – a guide for community seagrass
+Dixon R.\index[people]{Dixon R.}, editor. 2005. \emph{Watching the Seagrass Grow – a guide for community seagrass
 monitoring in NSW (2nd Ed)}. Reference manual . The Community Environment Network,

### 2011-02-05 MScDiplom_Lemenkova.tex
comment 3rd of Valentijn: updated 'Figure 1.8 UML diagram'; changed document text font: Times

@@
 \usepackage[left=30mm,right=20mm,top=30mm,bottom=35mm]{geometry}
+\usepackage{pslatex}
+%\usepackage{times}
+%\usepackage{psnfss}
 %\usepackage{glossary}
@@
 \begin{document}
+%\usefont{T1}
 \pagenumbering{roman}
@@
 The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.\\
-An Index is placed in the end of the Document to help the reader find what he is looking for.
+All persons mentioned in the current document are listed in the Index of People. \\
+An Index of Concepts is placed in the end of the Document to help the reader find what he is looking for.
 

### 2011-02-05 MScDiplom_Lemenkova.tex
comment 10 of Valentijn: added legend to GNUplot-18.

@@
 		\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and \ac{GPS} for data capture during the fieldwork
+		\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
 		\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
-		\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
-		\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
-		\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
+		\item \href{http://www.gnuplot.info/}{GNUplot} for plotting graphs
+		\item \href{http://gretl.sourceforge.net/}{Gretl: GNU Regression, Econometrics and Time-series Library} for statistic analysis
 		\item \href{http://www.openoffice.org/}{Open Office}: document typesetting, data pre-processing and preliminary statistical evaluation		
-
 		\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)}: statistical data edition and analysis
-		\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
+
 		\item \href{http://staruml.sourceforge.net/en/}{StarUML}, \href{http://live.gnome.org/Dia}{Dia}, \href{http://argouml.tigris.org/}						{ArgoUML}: \ac{UML} diagram tool for creating flowcharts of work algorithms
-		\item \href{http://inkscape.org/}{Inkscape}: drawing
 	\end{itemize}
@@
 		\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX}: converting spreadsheets into \LaTeX environment
+		\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
 		\item \href{http://openev.sourceforge.net/}{OpenEV}: used for raster Google Earth imagery viewing, visualisation and preliminary analysis
 		\item \href{http://www.notepad-plus-plus.org/}{Notepad++}: auxiliary text editing and raw data formatting
+		\item \href{http://inkscape.org/}{Inkscape} and \href{http://www.gimp.org/}{Gimp: GNU Image Manipulation Program} for auxiliary drawing
 	\end{itemize}
@@
 	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
-	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}}
+	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}}
 	\label{fig:1.10}
@@
 	\centering
+	\includegraphics[scale=0.35]{GNU-20-new.jpg}
+	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
+	\label{fig:28}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
 	\includegraphics[scale=0.5]{GNU-17.jpg}
@@
 and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.27]{GNU-18.jpg}
-	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis.}
-	\label{fig:4.4}
-\end{figure}
-
 Seagrass \textit{P.oceanica} can be mapped using remotely acquired spectral imagery, but only if it has distinctive
@@
 \textit{P.oceanica}: 0.5, 2.0 and 3.5 meters with an interval of 1.5 meters. 
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.3]{GNU-20-new.jpg}
-	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
-	\label{fig:28}
-\end{figure}
-
 The 3.5 m depth limit was chosen based on the analysis of the separability of seagrass reflectance signatures, received by the means of
@@
 broadband remote sensing.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.3]{GNU-18-legend.png}
+	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
+\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis.}
+	\label{fig:4.4}
+\end{figure}
 
@@
 	\subfloat [Sand]{\label{fig:34a}\includegraphics[width=0.4\textwidth]{Fig-27a.jpg}}
-	\subfloat [\textit{	extit{P.oceanica}}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
+	\subfloat [\textit{P.oceanica}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
 	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
@@
 \rowcolors{1}{Khaki}{LightYellow}
-\caption{Statistical analysis of the measurements of spectral reflectance of 	extit{P.oceanica} (fragment)}
+\caption{Statistical analysis of the measurements of spectral reflectance of \textit{P.oceanica} (fragment)}
 \centering

### 2011-02-05 MScDiplom_Lemenkova.tex
Comment 13 of Valentijn: combined plots of MODIS, MERIS, CZCS and SeaWiFS sensors

@@
 seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
-remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}).
-
-\begin{figure}[h]
-	\centering
-	\subfloat [CZCS]{\label{fig:29a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
-	\subfloat [MERIS]{\label{fig:29b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
-	\subfloat [MODIS]{\label{fig:29c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
-	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
-	\caption{Simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
+remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}, Fig.\ref{fig:4.6W}).
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.45]{Sensors.png}
+	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: MODIS, MERIS, SeaWiFS and CZCS.}
 	\label{fig:4.6}
-\end{figure}
+\end{wrapfigure}
 
@@
 	\end{center}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\subfloat [CZCS]{\label{fig:29a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
+	\subfloat [MERIS]{\label{fig:29b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
+	\subfloat [MODIS]{\label{fig:29c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
+	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
+	\caption{WASI-simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
+	\label{fig:4.6W}
 \end{figure}

### 2011-02-06 MScDiplom_Lemenkova.tex
Comment 14 of Valentijn: Fig.4.5: variable iterated over: depth

@@
 	\includegraphics[scale=0.45]{Sensors.png}
-	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: MODIS, MERIS, SeaWiFS and CZCS.}
+	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: MODIS, MERIS, SeaWiFS and CZCS, iterated over three depths.}
 	\label{fig:4.6}

### 2011-02-06 MScDiplom_Lemenkova.tex
Comment 15 of Valentijn: MERIS sensor (Fig.4.8)
Comment 8 of Valentijn: discussed spectra of P.oceanica with Thalassia and other seagrasses (p.2.5.3)

@@
 The tradition of global seagrass mapping though has not a very long history comparing to the
-terrestrial cartography, due to the technical difficulties of underwater observations, yet nowadays is
-has become a successful, rapidly developing and increasingly popular and challenging research branch.\\
-Seagrasses can be mapped using various approaches, with the most popular method of the remote
-sensing application, enabling to detect and discriminate the substratum and various vegetation types
-growing on the seafloor.\\
-Several global seagrass surveys\index[concepts]{Seagrasses!global surveys} organize and provide regular monitoring of the seagrass distribution,
-health and species sustainability, amongst which the most world-known are, perhaps, the following
-ones: \href{http://www.seagrassnet.org/}{Global Seagrass Monitoring project}, \href{http://www.seagrasswatch.org/}{Seagrasswatch}, 
-located in Australia, two European-located institutions - \href{http://mediterranean.seagrassonline.org/}{the Mediterranean association Seagrass-2000} and \href{http://www.imedea.uib.es/index.php}{the Mediterranean Institute for Advanced Studies}, US American
-\href{http://www2.fiu.edu/~seagrass/}{Seagrass Ecosystems Research Laboratory in South Florida} and
-\href{http://www.flseagrass.org/}{Florida Seagrass organisation}, the \href{http://www.seagrasses.org/}{Seagrasses.org} and the \href{http://wsa.seagrassonline.org/}{World Seagrass Association}.
-These organizations aim at the global seagrass monitoring, providing with
-research results and reporting guidelines and manuals with standardized methods and
-recommendations, specific for the seagrass research and monitoring. There are also university marine
-centres and research institutes conducting seagrass monitoring and as a particular part of their
-research. Their reports and guidelines were used for references in the current research.\\
+terrestrial cartography, due to the technical difficulties of underwater observations. However, nowadays is
+has become a rapidly developing, increasingly popular and challenging research branch.\vspace{1ex}
 Regular observations and monitoring of the seagrasses are known since 1960s, mainly in tropical
@@
 other side, development of the remote sensing methods and data acquisition from space contributed to
-the new methods of seagrass mapping, using distance approach and generally based on images
-classification.
+the new methods of seagrass mapping, using distance approach and generally based on images classification.\vspace{1ex}\\
+Various global seagrass survey organizations \index[concepts]{Seagrasses!global surveys} organize and provide regular monitoring of the seagrass species distribution, health and environmental sustainability.
+\\ We list below the most known seagrass research institutions:
+\begin{itemize}
+	\item Global-scaled: \href{http://www.seagrassnet.org/}{Global Seagrass Monitoring Network} and the \href{http://wsa.seagrassonline.org/}{World Seagrass Association}; \\ The \href{http://www.unep-wcmc.org/marine/seagrassatlas/contents.htm}{World Atlas of Seagrasses} is published by the \href{http://www.unep.org/}{UNEP}.
+	\item Australian \href{http://www.seagrasswatch.org/}{Seagrasswatch} (perhaps, the best organization, regularly publishing informative reports)
+	\item European: \href{http://mediterranean.seagrassonline.org/}{the Mediterranean association Seagrass-2000}, \href{http://www.imedea.uib.es/index.php}{the Mediterranean Institute for Advanced Studies} and \href{http://www.seagrasses.org/}{Seagrasses.org};
+	\item  US American \href{http://www.columbia.com/Seagrass-Recovery/About_Us_Giving_Back_Seagrass,default,pg.html}{seagrass recovery campaign} by the \href{http://www.seagrassgrow.org/}{Seagrassgrow}, \href{http://www2.fiu.edu/~seagrass/}{Seagrass Ecosystems Research Laboratory in South Florida}, \href{http://counties.cce.cornell.edu/suffolk/habitat_restoration/seagrassli/index.html}{Seagrass.LI} and
+\href{http://www.flseagrass.org/}{Florida Seagrass organisation}; 
+	\item Asian: \href{http://www.unepscs.org/}{UNEP/GEF South China Sea Project}, \href{http://www.marineconservationcambodia.org/Marine-Species-Photographic-Database/Cambodian-Marine-Life/Cambodian-Marine-Plants/Seagrasses.html}{Marine Conservation Cambodia} and \href{http://www.sosmalaysia.org/home.html}{Sosmalaysia.org}.
+\end{itemize}
+All these organizations aim at the global seagrass monitoring, providing with
+research results and reporting guidelines and manuals with standardized methods and
+recommendations, specific for the seagrass research and monitoring. There are also university marine
+centres and research institutes conducting seagrass monitoring and as a particular part of their
+research and reporting various approaches for the monitoring and mapping of the seagrasses, including remote
+sensing applications. Their reports and guidelines were used for references in the current research.
 
 \subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}
+Among various research approaches for the seagrass mapping the most popular methods are based on application of the remote sensing techniques,
+ enabling to detect and discriminate the substratum and various vegetation types growing on the seafloor.
 The application of the remote sensing data for seagrass mapping is based on our knowledge of the
@@
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 		\item [*] The \ac{CRTM} \index[concepts]{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA}, \ac{NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
-		\item [*] An \ac{4AOP}\index[concepts]{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{the French Space Agency} (\ac{CNES}).
+		\item [*] An \ac{4AOP}\index[concepts]{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{the French Space Agency}, \ac{CNES}.
 Being a fine program which enables accurate computations of the radiance, defining other emissivity functions and other set up, it is however, particularly designed for the infrared waves (thus, not quite suitable in our case).
@@
 	\centering
-	\includegraphics[scale=0.30]{Fig-11.jpg}
+	\includegraphics[scale=0.35]{Fig-11.jpg}
 	\caption{\textit{In situ} optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
@@
 
-Studies of spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and
-other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
+Studies of spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
 macroalgae differ from each other, as well as from sand and deep water reflectance spectra. These
@@
 depth where it is located, and is generally decreasing in values by increasing depths.
-The most important spectral diapason\index[concepts]{Seawater optical properties!spectral reflectance!spectral diapason} for marine mapping of submerged vegetation
-and, particularly, for seagrass, is 350-800 nm \cite{Barille09}\label{Barille09}.
+The most important spectral diapason\index[concepts]{Seawater optical properties!spectral reflectance!spectral diapason} for marine mapping of submerged vegetation and, particularly, for seagrass, is 350-800 nm \cite{Barille09}\label{Barille09}.
 Using airborne imagery for retrospective data (before 1970s) together with the most
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
-Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial
-resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
+Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
 orbital repeat cycle for nadir viewing and imagery with a spatial resolution 20 - 2.5 m
@@
 on board calibration \ac{OBC} system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.15]{Fig-15.jpg}
+	\caption{Reflectance spectra of \textit{Thalassia} seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
+		reflectance. Source:\cite{Yang10}\label{Yang10}}
+	\label{fig:2.6}
+\end{wrapfigure}
+
 This makes \ac{ASTER} imagery especially suitable for the application towards detailed
@@
 macrophytes in shallow waters of the Lake Constance. 
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-15.jpg}
-	\caption{Reflectance spectra of seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
-		reflectance. Source:\cite{Yang10}\label{Yang10}}
-	\label{fig:2.6}
-\end{wrapfigure}
-
 In regard to the methods chosen for the image interpretation, the supervised classifications proves to
@@
 high spectral resolution, desirable for the case of seagrass monitoring. 
-The most suitable
-scanner fit for detailed seagrass mapping would cover bands of 550-750 nm and have a spectrum
+The most suitable scanner fit for detailed seagrass mapping would cover bands of 550-750 nm and have a spectrum
 resolution of 5-15 nm \cite{Fyfe04}\label{Fyfe04}, which exactly characterizes typical airborne and
@@
 mapping benthic vegetation in areas where the spatial heterogeneity is very high \cite{Vahtmae07}\label{Vahtmae07}.
+
+\subsubsection{Comparison of seagrass spectra }\label{sec:2.5.3}
+
+The reflectance spectra from different seagrass species reported in various research works demonstrate diverse spectral signatures which are dependent on the mixture of environmental conditions as well as individual characteristics of the seagrass.
+In the current work we have compiled an extensive measurements dataset (up to 400 single spectral profiles) of the reflectance spectra from various seafloor cover types (P.oceanica and carbonate sand) in the coastal waters of Crete Island, which is partly visualized on the plotted graphs (Fig.\ref{fig:4.2}, Fig.\ref{fig:4.10}, \ref{fig:A.12}, \ref{fig:A.13}, \ref{fig:A.15}, \ref{fig:A.25}, \ref{fig:A.26}).
+Having analyzed reflectance spectra received in previous works by other researchers and compared them with our results, obtained in the current work, we noticed some general trends in the character of reflectance spectra of various seafloor cover types.
+We mention below some general conclusions and observed tendencies in the profiles and patterns of spectra of \textit{P.oceanica} and carbonate sand (as for Crete Island).
+Measured remote sensing spectral reflectance is highly variable with changing environmental conditions, such as water physical content (amounts of suspended particles and organic matter, salinity, chemical content), water column depth, atmospheric conditions (e.g.
+sun zenith angle) and individual features on the plant, e.g.
+morphology and colour pigmentation of the leaves, differed in various plant's health and age conditions.
+The maximal values of the spectral reflectance of \textit{P.oceanica}, received both in the current work and by previous researchers, lies in wavelength interval of 500-600 nm, as, for example, shown on Fig.\ref{fig:13}, Fig.\ref{fig:2.6}, and can be clearly seen in our results as well (Fig.\ref{fig:4.2}, Fig.\ref{fig:4.10}, Fig.\ref{fig:4.13}, \ref{fig:A.25}).
+Comparison of the remote sensing reflectance spectra of various seagrass species, e.g.
+\textit{P.oceanica} (Fig.
+\ref{fig:4.6}) and \textit{Thalassia testudinum} (Fig.\ref{fig:2.6}), shows that \textit{P.oceanica} is in generally brighter in the wavelength interval 500-600 nm which can be caused by various colour pigmentation.
+The plotted graphs with spectra of the seagrasses \textit{P.oceanica} and \textit{Thalassia testudinum} demonstrate that the dip in the spectral reflectance after ca 600 nm is more pronounced in case of \textit{P.oceanica} seagrass beds (Fig.\ref{fig:4.6}) than in those with \textit{Thalassia testudinum} (Fig.\ref{fig:2.6}), due to absorption caused by pigment of green leaves.
+The remote sensing of the bottom albedo of \textit{P.oceanica} and carbonate sand (Fig.\ref{fig:4.11}) shows clear difference between the carbonate sand and \textit{P.oceanica}, which is evidently much darker.
+To estimate and further analyze particularities of the \textit{P.oceanica} remote sensing reflectance spectra over regions with different bathymetry (0.5, 1.5 and 2.5 meters column depth) we used \ac{WASI} \ac{RTM} software, which incorporates measurements of the seafloor reflectance and water column optical properties.
 
@@
 initial data of measurements of the spectral reflectance have been preprocessed and statistically analyzed thereafter with graphs visualizing mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:4.10}, \ref{fig:28}).
-These data
-have been measured using different step of the wavebands: some measurements
+These data have been measured using different step of the wavebands: some measurements
 were made with 3 nm interval, while others – using 4 and 1 nm step.
@@
 comparing to splines (see \nameref{app} for more results). 
-Therefore, after several experiments with
-various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have chosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contains convex
-hulls made on its control points, and therefore is best suitable for our case: analysis of optical
+Therefore, after several experiments with various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have chosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contains convex hulls made on its control points, and therefore is best suitable for our case: analysis of optical
 properties of seawater.
@@
 nm, which is suitable for characteristics of seagrass reflectance.
-Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:A.12}, or \ref{fig:A.14}, enlarged) represent several series of
-the measurements. The values of the spectral reflectance are received from the
+Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:A.12}, or \ref{fig:A.14}, enlarged) represent several series of the measurements.
+The values of the spectral reflectance are received from the
 computations of these values using mathematical formulae.
@@
 	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d.jpg}}
-	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths}
+	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths. MERIS sensor.}
 	\label{fig:4.9}
@@
 %%%Gretl end
-\pagebreak
 
@@
 		\includegraphics[scale=0.23]{OpenEV-3.jpg}
-	\caption{\textit{Logarithmic Enhancement to Raster}, applied to the aerial Google Earth image: the seagrass meadow can now be easily seen as a bright spot of purple color.
+	\caption{\textit{Logarithmic Enhancement to Raster}, applied to the aerial Google Earth image: the seagrass meadow seen as a bright spot of purple color.
 \href{http://openev.sourceforge.net/}{OpenEV}}\label{fig:A.40}
@@
 	\begin{center}
-		\includegraphics[scale=0.40]{Fig-36.jpg}
+		\includegraphics[scale=0.25]{Fig-36.jpg}
 		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }

### 2011-02-07 MScDiplom_Lemenkova.tex
Valentijn's comments of 6Feb: 2(chart with buoys),4(broadband sensors); written 'Conclusions', updated 'Supervisory Committee'

@@
 	\item[*] \href{http://www.lunduniversity.lu.se/}{Lund University}, \href{http://www.giscentrum.lu.se/english/index.htm}{GIS Centre} (Sweden)\\
-	\item[*] \href{http://www.uw.edu.pl/en/}{University of Warsaw} (Poland)\\
+	\item[*] \href{http://www.uw.edu.pl/en/}{University of Warsaw}, \href{http://www.wgsr.uw.edu.pl/index.php?option=com_content&task=category&sectionid=10&id=32&Itemid=103}{Faculty of Geography and Regional Studies} (Poland)\\
 	\item[*] \href{http://www.utwente.nl/en}{University of Twente}, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)} \\ (The Netherlands)
@@
 
-Thesis submitted to the University of Twente, Faculty ITC, in partial fulfillment of the requirements
+Thesis submitted to the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, in partial fulfillment of the requirements
 for the degree of Master of Science in Geo-information Science and Earth Observation for
@@
 \begin{itemize}
-	\item Name Examiner 1 (Chair)\dotfill
-	\item Name Examiner 2\dotfill
-	\item External Examiner \dotfill \\ \hfill{Dr.{\textit{Petter Pilesj\"o}\index[people]{Pilesj\"o, P.}} (Lund University, GIS Centre)}
+	\item Name Examiner 1 (Chair) \hfill{Prof. Dr. \textit{Andrew K. Skidmore}\index[people]{Skidmore A.K.} (ITC, Department of Natural Resource)}
+	\item Name Examiner 2 \hfill{Ms. Ir. \textit{Louise M. van Leeuwen}\index[people]{Leeuwen, L. M. van} (ITC, Department of Natural Resource)}
+	\item External Examiner \hfill{Dr. {\textit{Ma\l{}gorzata Roge-Wi\'sniewska}\index[people]{Roge-Wi\'sniewska, M.}} (University of Warsaw, Faculty of Geography and Regional Studies)}
 \end{itemize}
@@
 \begin{itemize}
-	\item[* ] Mr. \textit{Valentijn Venus}\index[people]{Venus, V.} \dotfill \\ (Department of Natural Resource, ITC)
-	\item[* ] Dr. \textit{Albertus G. Toxopeus}\index[people]{Toxopeus, A.G.} \dotfill \\ (Department of Natural Resource, ITC)
+	\item[* ] Mr. \textit{Valentijn Venus}\index[people]{Venus, V.}  \hfill{(University of Twente, ITC, Department of Natural Resource)}
+	\item[* ] Dr. \textit{Albertus G. Toxopeus}\index[people]{Toxopeus, A.G.}  \hfill{(University of Twente, ITC, Department of Natural Resource)}
 \end{itemize}
@@
 Furthermore, this chapter includes analysis of the spectral signatures of various seafloor cover types in conditions of changing environment, possibility and limitations for \textit{P.oceanica} spectral discrimination, statistical analysis of the data sets and mapping based on images processing.\\
-\textit{Chapter 5}: \nameref{sec:5} summarizes and briefly discusses again the main results, reported in detail in a previous chapter.
-This chapter proposes some final discussions to the reader, and comes to the main research outcomes.\\
-Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \nameref{sec:6}.\\
+\textit{Cahpter 5}: \nameref{sec:5} discusses research results, reported in detail in a previous chapter, comes to the research insights and discusses relevance of this work in scientific context.\\
+\textit{Chapter 6}: \nameref{sec:6} summarizes and highlights the final research outcomes and main results.\\
+Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \nameref{sec:7}.\\
 A lot of plotted graphs, statistical outcomes, tables, auxiliary (yet relevant to our research) pictures and images, illustrating our work - are collected in the \nameref{app}.
@@
 The research materials and equipment were provided by \href{http://147.52.130.151/MFIK_en/}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:3.3}): 1) three iPAQs (Fig.\ref{fig:3.3a}); 2)
-three GPS (Fig.\ref{fig:3.3c}); 3) Three underwater video cameras, \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), suitable for
-photographing up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
+three GPS (Fig.\ref{fig:3.3c}); 3) Three underwater video cameras, \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), suitable for photographing
+
+\begin{figure}[H]
+	\centering
+	\subfloat [iPAQ, HP]{\label{fig:3.3a}\includegraphics[width=0.2\textwidth]{Fig-19a.jpg}}
+	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:3.3b}\includegraphics[width=0.2\textwidth]{Fig-19b.jpg}}
+	\subfloat [GPS]{\label{fig:3.3c}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
+	\subfloat [Olympus waterproof camera]{\label{fig:3.3d}\includegraphics[width=0.2\textwidth]{Fig-19d.jpg}}
+	\caption{Fieldwork equipment}
+	\label{fig:3.3}
+\end{figure}
+
+up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
 Markers and cords for the depths measurements; 5) Waterproof plastic Otterbox  (Fig.\ref{fig:3.3b})for keeping the
@@
 	\centering
-	\subfloat [iPAQ, HP]{\label{fig:3.3a}\includegraphics[width=0.3\textwidth]{Fig-19a.jpg}}
-	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:3.3b}\includegraphics[width=0.3\textwidth]{Fig-19b.jpg}}
-	\subfloat [GPS]{\label{fig:3.3c}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
-	\subfloat [Olympus waterproof camera]{\label{fig:3.3d}\includegraphics[width=0.25\textwidth]{Fig-19d.jpg}}
-	\caption{Fieldwork equipment}
-	\label{fig:3.3}
+	\includegraphics[scale=0.60]{Fig-18.jpg}
+	\caption{Boat used for the fieldwork measurements on Crete Island. 2010}
+	\label{fig:3.4}
 \end{figure}
@@
 course plot to cover the area of growing seagrass: shelf areas not deeper than four meters.
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.35]{Sampling_design.jpg}
+		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}\label{fig:A.4}
+	\end{center}	
+\end{figure}
+
 The measurements of the seafloor cover types have been made using underwater video cameras
@@
 we used transect sampling method, i.e. photographs were taken along the research path. 
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.70]{Fig-18.jpg}
-	\caption{Boat used for the fieldwork measurements on Crete Island. 2010}
-	\label{fig:3.4}
-\end{figure}
-
 Transect method is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
@@
 without location bias.
+
+\begin{figure}[H]
+	\centering
+	\subfloat [view from below]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
+	\subfloat [view \textit{en-face}]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
+	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
+	\caption{Scheme of placements of the Olimpus cameras during the measurements}
+	\label{fig:3.5}
+\end{figure}
+
 The videographic measurements and photos, enabling to detect types of the
@@
 were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state. 
-
-\begin{figure}[H]
-	\centering
-	\subfloat [view from below]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
-	\subfloat [view \textit{en-face}]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
-	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
-	\caption{Scheme of placements of the Olimpus cameras during the measurements}
-	\label{fig:3.5}
-\end{figure}
 The locations of the route were randomly selected in the areas of the Ligaria beach, to ensure most dense
@@
 	\centering
-	\includegraphics[scale=0.40]{Fig-21.jpg}
+	\includegraphics[scale=0.50]{Fig-21.jpg}
 	\caption{Locations of the video measurements and GPS tracklogs, Ligaria}
@@
 	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d.jpg}}
-	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths. MERIS sensor.}
+	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths, simulated for broadband sensors.}
 	\label{fig:4.9}
@@
 
-\section{Discussion and conclusions}\label{sec:5}
+\section{Discussion}\label{sec:5}
+
+\pagebreak
+
+\section{Conclusion}\label{sec:6}
+The goal of this MSc research was to explore the perspectives, advantages and limitations of the narrow-band and broadband sensors for the environmental mapping and monitoring of \textit{P.oceanica} seagrass along the coasts of Crete Island.
+The research outcome demonstrated that the application of the remote sensing data from the broadband sensors is highly advantageous for the seagrass mapping, the spectral discrimination of \textit{P.oceanica} from other seafloor cover types is possible at diverse and changing environmental conditions, and that \textit{P.oceanica} is spectrally distinct from other seagrass species (\textit{Thalassia testudinum}).\vspace{1ex}\\
+The \ac{RTM} software is a powerful means for analyzing spectral signatures of various seafloor types and enabling simulations of data received from broadband and narrowband remote sensors.\\ The example of application of \ac{WASI} \ac{RTM}, given in this work, is an achievement of the research insight towards the spectral properties of \textit{P.oceanica} and other bottom cover types, enabling to discriminate them from each other with changing environmental conditions.
+The research shows that spectral signatures of \textit{P.oceanica} are distinct at various depths.\vspace{1ex}\\
+The methodology of the spectral discrimination of seafloor cover types is designed in the frame of this research and is based on the application of the remote sensing \ac{RTM} techniques, data from broadband sensors, hyperspectral radiometers for measurements of optical properties of the seawater, categorical and continuous statistical analysis for the data processing and \ac{GIS} raster based software for images visualization, classification and analysis.
+Technically, we used different software, adjusted for diverse research purposes, to manage, integrate and process data from various origin and resources, and finally to receive  accurate research results.\vspace{1ex}\\
+The marine coastal ecosystems are complex, constantly changing and developing.
+Using flexibility of \ac{GIS} combined with remote sensing methods and application of data from broadband sensors is therefore advantageous for the monitoring of coastal areas.
+Besides Mediterranean area, the methodology of the seagrass environmental studies can be applied towards other shelf areas with dominating seagrass landscapes.\vspace{1ex}\\
 More than 50 \% of the world population lives within one km of the coast, which results in continued anthropogenic pressure on the coastal regions.
-Therefore, management of coastal resources and shelf zone protection become increasingly important nowadays, and require large-scale monitoring and mapping of the shelf areas as vital instrument for the environmental assessment.
-Besides Mediterranean area, the suggested methodology of the analysis of seagrass monitoring and analysis of itsadaptation towards different environmental conditions within the coastal zones can be applied to other marine coastal areas with growing seagrass.
-
-\section*{Recommendations}\label{sec:6}
+Therefore, management of coastal resources and shelf zone protection become increasingly important nowadays, and require large-scale monitoring and mapping of the shelf areas as a vital instrument for the environmental assessment.\vspace{1ex}\\
+This research is a contribution to the development of the methodology of seagrass mapping with the aim of the environmental monitoring, and a case study of \textit{P.oceanica}, an endemic Mediterranean seagrass species, dominating in underwater landscapes and ecosystems along the coasts of Crete Island.
+\pagebreak
+
+\section*{Recommendations}\label{sec:7}
 \markright{Recommendations}
@@
 
-\subsection{\textit{Appendix-4}. Sampling design of the depths measurements. Ligaria beach, Crete}
-%\appendixname{\textbf{4: \\Sampling design of the depths measurements. Ligaria beach, Crete}}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.35]{Sampling_design.jpg}
-		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}\label{fig:A.4}
-	\end{center}	
-\end{figure}
-\pagebreak
-
-\subsection{\textit{Appendix-5}. Results of videographic measurements: seafloor types on Crete island}
+\subsection{\textit{Appendix-4}. Results of videographic measurements: seafloor types on Crete island}
 %\appendixname{\textbf{5: \\Results of videographic measurements: seafloor types on Crete island}}
@@
 
-\subsection{\textit{Appendix-6}. Data pre-processing}
+\subsection{\textit{Appendix-5}. Data pre-processing}
 %\appendixname{\textbf{6: \\Data pre-processing}}
@@
 
-\subsection{\textit{Appendix-7}. Data processing and statistical evaluation: Results of the measurements of the seafloor optical properties.}
+\subsection{\textit{Appendix-6}. Data processing and statistical evaluation: Results of the measurements of the seafloor optical properties.}
 %\appendixname{\textbf{7:\\ Data processing and statistical evaluation: \\ Results of the measurements of the seafloor optical properties.}}
@@
 
-\subsection{\textit{Appendix-8}. Satellite and aerial images covering research area of Crete island: selected examples}
+\subsection{\textit{Appendix-7}. Satellite and aerial images covering research area of Crete island: selected examples}
 %\appendixname{\textbf{8:\\ Satellite and aerial images covering research area of Crete island: selected examples}}
@@
 
-\subsection{\textit{Appendix-9}. Analysis and classification of the satellite and aerial images}
+\subsection{\textit{Appendix-8}. Analysis and classification of the satellite and aerial images}
 
@@
 
-\subsection{\textit{Appendix-10}. Research general workflow}
+\subsection{\textit{Appendix-9}. Research general workflow}
 \begin{figure}[H]

### 2011-02-08 MScDiplom_Lemenkova.tex
started 'Chapter5. Discussion'

@@
 \usepackage{authblk}
+\usepackage[english]{babel}
 %\usepackage{authorindex}
@@
 Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and ETM+ scenes.
-
 \pagebreak
@@
 	\acro{MIPAS}{\href{http://envisat.esa.int/instruments/mipas/}{Michelson Interferometer for Passive Atmospheric Sounding}}
-	\acro{MODIS}{\href{http://modis.gsfc.nasa.gov/}{Moderate Resolution Imaging Spectroradiometer}}
+	\acro{MODIS}{\href{http://modis.gsfc.nasa.gov/}{MODerate resolution Imaging Spectroradiometer}}
 	\acro{NASA}{\href{http://www.nasa.gov/}{National Aeronautics and Space Administration}}
@@
 		\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX}: converting spreadsheets into \LaTeX environment
-		\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
+		\item \href{http://www.python.org/}{Python language}, for writing ancillary scripts (e.g. raw data interpolation)
 		\item \href{http://openev.sourceforge.net/}{OpenEV}: used for raster Google Earth imagery viewing, visualisation and preliminary analysis
-		\item \href{http://www.notepad-plus-plus.org/}{Notepad++}: auxiliary text editing and raw data formatting
-		\item \href{http://inkscape.org/}{Inkscape} and \href{http://www.gimp.org/}{Gimp: GNU Image Manipulation Program} for auxiliary drawing
+		\item \href{http://www.notepad-plus-plus.org/}{Notepad++}: ancillary text editing and raw data formatting
+		\item \href{http://inkscape.org/}{Inkscape} and \href{http://www.gimp.org/}{Gimp: GNU Image Manipulation Program} for drawing
 	\end{itemize}
@@
 oceanica} along the selected locations of the northern coasts of Crete island, Greece.
-The technical implementation is based on \ac{WASI} RTM software, \ac{GIS} (ILWIS, Erdas Imagine and ArcGIS), using aerial and satellite images and the results of the underwater videometric measurements.
+The technical implementation is based on \ac{WASI} \ac{RTM} software, \ac{GIS} (\ac{ILWIS}, Erdas Imagine and ArcGIS), using aerial and satellite images and the results of the underwater videometric measurements.
 
@@
 constitutes the basis of the food web both within and outside the ecosystem \cite{Gobert06}\label{Gobert06}.
-Finally, seagrasses are an important component in the environmental "food chain" of the coastal
-ecosystems, being the food source for dugongs, turtles, swans and various fish \cite{Cappo95}\label{Cappo95}.
+Finally, seagrasses are an important component in the environmental ``food chain" of the coastal
 
@@
 
+ecosystems, being the food source for dugongs, turtles, swans and various fish \cite{Cappo95}\label{Cappo95}.
 Due to their wide distribution, meadows size, easy collection and abundance, sensitivity to the
@@
 ecosystems, and is highly suitable for the environmental monitoring \cite{Pergent-Martini05}\label{Pergent-Martini05}.\\
-Being often confused with marine “algae”, “seagrasses” are vastly different from them. There are
-fundamental differences between both marine organisms, the major of them should be briefly
-mentioned: first, seagrasses are true plants with root system and leaves which photosynthesis, have
+Being often confused with marine ``\textit{algae}"\index[concepts]{Algae}, ``\textit{seagrasses}" are vastly different from them.
+There are fundamental differences between both marine organisms, which we briefly mention below: 
+\begin{itemize}
+	\item[*] first, \textit{seagrasses} are true plants with root system and leaves which photosynthesis, have
 complex structure and create landscape-similar vast formations on the colonized seafloor areas with
-soft sediment, whereas algae are simple organisms that can only holdfast; secondly, seagrasses are
-complex vascular plants with reproductive mechanism such as fruits, seeds and spores, while algae
-have simple few cell structure with spores and gametes; finally, seagrasses uptake nutrient through
-root system, while algae nourish directly from the water column \cite{Dixon05}\label{Dixon05}. There are other
-differences between “seagrass” and “algae” but they go beyond the scope of this work.
+soft sediment, whereas \textit{algae} are simple organisms that can only holdfast; 
+	\item[*] secondly, \textit{seagrasses} are complex vascular plants with reproductive mechanism such as fruits, seeds and spores, while \textit{algae} are primitive organisms with simple cell structure, spores and gametes;
+	\item[*] finally, \textit{seagrasses} uptake nutrient through the root system, while \textit{algae} nourish directly from the water column \cite{Dixon05}\label{Dixon05}. 
+\end{itemize}
+There are other nuances and details of difference between the ``\textit{seagrasses}" and ``\textit{algae}", but their discussion would go beyond the scope of this work.
 
@@
 	\end{itemize}
+\pagebreak
 
@@
 		\item Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying \textit{in-situ} conditions ?
-		\item Whether broadband and hyperspectral remote sensing data can be used for the mapping of \textit{P.oceanica} ?
+		\item Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types ?
 	\end{itemize}
@@
 The hypothesis testing is suggested to be carried out using the \ac{ANOVA} statistical test. The purpose of
-ANOVA test is to visualize in an effective and quick way the spectral differences between seagrass
+\ac{ANOVA} test is to visualize in an effective and quick way the spectral differences between seagrass
 species and their spatial distribution. Thus, the key hypotheses of the research will be tested to prove
@@
 images, GIS-based spatial analysis and ground \textit{in-situ} surveys.
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
-	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}}
-	\label{fig:1.10}
-\end{figure}
 The current study is based on the application of the remote sensing data, broadband satellite imagery,
@@
 shallow waters keeps information on the optical properties of the seawater components and benthic
-substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. \\
+substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
+	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}}
+	\label{fig:1.10}
+\end{figure}
+
 The spectral reflectance, distinct for each seafloor cover type, enables their discrimination on the image. The pre-processing
@@
 Supervised, Maximum Likelihood).
+\pagebreak
 
@@
 Various global seagrass survey organizations \index[concepts]{Seagrasses!global surveys} organize and provide regular monitoring of the seagrass species distribution, health and environmental sustainability.
-\\ We list below the most known seagrass research institutions:
+\\ We list below the most important and world-known scientific institutions focusing on seagrass research:
 \begin{itemize}
@@
 and up-to-date source of imagery.
-Perhaps, the clearest advantage of the Google
-Earth imagery is its high resolution (15 m in land areas and lower in the oceans).
+Perhaps, the clearest advantage of the Google Earth imagery is its high resolution (15 m in land areas and lower in the oceans).
 Obtained from the airborne platforms, Google Earth images have general spatial
@@
 the seafloor have been received by the means of the optical sensors \ac{Trios-RAMSES} Hyperspectral
-UVA/UVB Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral UV-VIS Radiance Sensor (Fig.
+\ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor (Fig.
 \ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see \nameref{app}: Tab.
@@
 
-The backward-scattering coefficient (bb), also included in WASI, is a
+The backward-scattering coefficient (bb), also included in \ac{WASI}, is a
 fundamental optical property which plays a central role in the ocean-color remote sensing, providing
 the remotely sensed optical signal, as well as suspended particle distributions, water clarity, and
-underwater visibility \cite{Maffione97}\label{Maffione97}. WASI enables simulation of backscattering of pure
+underwater visibility \cite{Maffione97}\label{Maffione97}. \ac{WASI} enables simulation of backscattering of pure
 water, large and small particles. The values of all these parameters can be redacted and changed
@@
 bay, which indicated that \textit{P.oceanica} seagrass could be only well separated within depths of 3.5 m.
-A statistical analysis of WASI-simulated spectral reflectance has been used in order to answer the first
-research question: whether the 	extit{P.oceanica} spectra is spectrally distinct at varying environmental \textit{in-situ}
+A statistical analysis of \ac{WASI}-simulated spectral reflectance has been used in order to answer the first
+research question: whether the \textit{P.oceanica} spectra is spectrally distinct at varying environmental \textit{in-situ}
 conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
@@
 	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis.}
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Two complimentary graphs below show the results of the statistical analysis.}
 	\label{fig:4.4}
@@
 	\includegraphics[scale=0.45]{Sensors.png}
-	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: MODIS, MERIS, SeaWiFS and CZCS, iterated over three depths.}
+	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths.}
 	\label{fig:4.6}
@@
 i.e. sun angle, concentration of the suspected particles in the water column, content of gelbstoff, etc.
-For these factors default values of WASI are accepted.
+For these factors default values of \ac{WASI} are accepted.
 Under normal conditions by independent water colour sampling, values of the remote-sensing
@@
 based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
-types, where the spectral signatures were examined. WASI software\index[concepts]{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate
+types, where the spectral signatures were examined. \ac{WASI} software\index[concepts]{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate
 spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
@@
 cover fraction of the seafloor, in order to assess spectral signatures of the seagrass and carbonate sand
-as major seafloor types. WASI enables to use forward or inverse calculation for the spectrum types at
+as major seafloor types. \ac{WASI} enables to use forward or inverse calculation for the spectrum types at
 a diapason of 350-800 nm with a 1nm spectral resolution. For the spectral analysis we applied
@@
 
-An analysis of the spectral reflectance of \textit{P.oceanica} is done using the WASI simulations\index[concepts]{Seawater optical properties!RTM!WASI} in order to
-determine, which wavebands can be still used to identify \textit{P.oceanica}. The analysis of spectra shows
+An analysis of the spectral reflectance of \textit{P.oceanica} is done using the \ac{WASI} simulations\index[concepts]{Seawater optical properties!RTM!WASI} in order to determine, which wavebands can be still used to identify \textit{P.oceanica}.
+The analysis of spectra shows
 that the appropriate wavebands for seagrass mapping lay between 500 and 600 nm and has
@@
 The patches of white sandy bottoms of the seafloor, studied in the fieldwork in Ligaria beach, are
-much brighter than mattes of P.Oceanica (Fig.\ref{fig:A.6}, \ref{fig:A.5}): seafloor types in Ligaria), which can be clearly seen
-at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:A.35}) and that of P.
-oceanica (Fig.\ref{fig:A.28}).
+much brighter than mattes of \textit{P.oceanica} (Fig.\ref{fig:A.6}, \ref{fig:A.5}): seafloor types in Ligaria), which can be clearly seen
+at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:A.35}) and that of \textit{P.oceanica} (Fig.\ref{fig:A.28}).
 The graph is received in excel spreadsheet using mean values of spectral reflectance of
@@
 These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
-These models simulate spectral views of the chosen sensors, how these sensors will "see" seagrass (Fig.\ref{fig:28}) and sand as
+These models simulate spectral views of the chosen sensors, how these sensors will ``see" seagrass (Fig.\ref{fig:28}) and sand as
 pixels, with accepted default atmosphere and water column effects (given by WASI software). In such
@@
 high resolution aerial color Google Earth photography, spaceborne satellite imagery, assessment of
-spectral signatures using WASI software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS based mapping.
-The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
+spectral signatures using \ac{WASI} software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc \ac{GIS} based mapping.
+The use of \ac{GIS} for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
 enables to analyze environmental changes within seagrass landscapes based on data from various
@@
 The final mapping has been supported in ArcGIS through the data exporting, conversion and
-integration of various data in one GIS-project. Data collected during the fieldwork, imagery
-of the seagrass distribution are added into a GIS dataset for the assessment and spatial analysis.
+integration of various data in one \ac{GIS}-project. Data collected during the fieldwork, imagery
+of the seagrass distribution are added into a \ac{GIS} dataset for the assessment and spatial analysis.
 
@@
 \section{Discussion}\label{sec:5}
+An approach of the seagrass spectral analysis, monitoring and mapping has been taken in this work, which integrates various research techniques and tools, combining remote sensing methods of spectral analysis of the seafloor cover types, and knowledge of the ecology of \textit{P.oceanica}, with the aim to develop a method of seagrass spectral optical discrimination for the seagrass mapping based on imagery classification.\vspace{1ex}\\
+The ecological variables, specific to the field environmental conditions, were factored into the \ac{WASI}-based simulation models.
+Through the \ac{WASI} simulation process, imitating spectral properties of \textit{P.oceanica} and carbonate sand for various broadband and narrowband sensors, models were created that accounted for not only atmospheric conditions (i.e.
+sun zenith angle), but also height of water column, thus approaching it to the Mediterranean conditions, and chemical content of the seawater (i.e.
+amount of suspended particles, Gelbstoff, etc), which results in models of optical properties of ``seawater with sediments" and ``seawater without sediments".\vspace{1ex}\\
+The \textit{in-situ} field large-scale matte-level level of seafloor monitoring was then upscaled to airborne Google Earth aerial imagery interpretation, to provide a meadow-level view of seagrass landscapes.
+An attempt of the small-scale mapping is designed on an example of Landsat satellite imagery.
+However, in upscaling to this third, small-scale mapping level further environmental variables need to be considered: health conditions of the seagrass, presence of other underwater vegetation (e.g.
+other seagrass species), hydrological specifications (e.g.
+direction and speed of currents, amplitude of tidal waves, etc), season, date and times of the image taken.
+Therefore, we focused on the first two levels in the current work.
+These different levels have been individually considered in terms of the seagrass spectral discernibility for monitoring and mapping, from which the first two levels have been brought together, to provide a roundup of the achieved results and an overview of what still has to be done in \textit{P.oceanica} seagrass mapping by future researchers (see Recommendations).
+\vspace{1ex}\\
+The finding of \textit{Chapter 4}, \nameref{sec:4} showed that the relationship between the spectral reflectance of various seafloor cover types was tied to depth, i.e.
+water column height.
+Thus, the results of the in-situ fieldwork measurements revealed that spectral reflectance of P.oceanica undergo alterations at depths of 0.5, 2.0 and 3.5m (Fig.\ref{fig:4.9}), which positively answers the first research question of this thesis (``Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ conditions ?").
+\vspace{1ex}\\
+Further in \textit{Chapter 4}, \nameref{sec:4}, studies of the broadband versus narrowband sensors demonstrate that simulated spectra of the seagrass, made using \ac{WASI} modeller, have the best results at \ac{CZCS} scanner, especially devoted to the measurement of ocean color.
+The spectrum of \textit{P.oceanica} reflectance, simulated for \ac{CZCS}, covers the wavelength interval of 400-800 nm, and is distinctive for various depths.
+Other remote sensors (\ac{MODIS}, \ac{SeaWiFS}) may also be used for the seagrass mapping, because their technical characteristics enable to spectrally discriminate \textit{P.oceanica} seagrass from other seafloor cover types (Fig.\ref{fig:4.6}), particularly carbonate sand as tested in the current work.
+Therefore, the second research question of this MSc thesis (``Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\
+The graphs showing optical properties of seawater with and without sediments (Fig.\ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:4.2}) focused on spectral variability of the water with changed physical and chemical content.
+The alterations in the individual spectral signatures of single measurements (e.g.
+on Fig.\ref{fig:4.13}, Fig.\ref{fig:A.25}, Fig.\ref{fig:A.27}, Fig.\ref{fig:A.28}) reflect individual health properties of leaves: different nitrogen and chlorophyll content causing diverse colour pigmentation and light absorption, water content in leaves, plant physiological conditions, which vary across seagrass meadow, shoot morphology, etc.
+The differences in spectral reflectance values of the measurements taken on various days might have been caused by the impact of atmospheric conditions, such as solar radiation and sun illumination by different zenith angle.
 
@@
 The goal of this MSc research was to explore the perspectives, advantages and limitations of the narrow-band and broadband sensors for the environmental mapping and monitoring of \textit{P.oceanica} seagrass along the coasts of Crete Island.
-The research outcome demonstrated that the application of the remote sensing data from the broadband sensors is highly advantageous for the seagrass mapping, the spectral discrimination of \textit{P.oceanica} from other seafloor cover types is possible at diverse and changing environmental conditions, and that \textit{P.oceanica} is spectrally distinct from other seagrass species (\textit{Thalassia testudinum}).\vspace{1ex}\\
+The research outcome demonstrated that the application of the remote sensing data from the broadband sensors is highly advantageous for the seagrass mapping, the spectral discrimination of \textit{P.oceanica} from other seafloor cover types is possible at diverse and changing environmental conditions, and that \textit{P.oceanica} is spectrally distinct from other seagrass species as well (\textit{Thalassia testudinum}).\vspace{1ex}\\
 The \ac{RTM} software is a powerful means for analyzing spectral signatures of various seafloor types and enabling simulations of data received from broadband and narrowband remote sensors.\\ The example of application of \ac{WASI} \ac{RTM}, given in this work, is an achievement of the research insight towards the spectral properties of \textit{P.oceanica} and other bottom cover types, enabling to discriminate them from each other with changing environmental conditions.
@@
 This research is a contribution to the development of the methodology of seagrass mapping with the aim of the environmental monitoring, and a case study of \textit{P.oceanica}, an endemic Mediterranean seagrass species, dominating in underwater landscapes and ecosystems along the coasts of Crete Island.
-\pagebreak
 
@@
 If possible, to find out bathymetric data for the research area, and to overlay them with existing images and maps, in order to analyse correlation between spatial distribution of seagrass \textit{P.oceanica} and underwater relief along Cretan coasts.
-		\item[Seventh:] Other RTM software may be tested and the modelling outcomes compared.
-		\item[Eighth:] Application of various open source \ac{GIS} (\ac{ILWIS}, \ac{GRASS}) could be very useful for the validation of the cartographic results, assessment of accuracy and comparison of various classification methods.
-		\item[Ninth:] The analysis of the health indicators of the seagrass (such as number of leaves per shoot, biomass estimation within the single shot, etc) was not considered in the current work, as it would go beyond the scope of the MSc thesis.
+		\item[Seventh:] In upscaling to the small-scale mapping level further environmental variables need to be considered: health conditions of the seagrass (usually, indicated by the number of leaves per shoot), presence of other underwater vegetation, hydrology (e.g.
+direction and speed of currents, amplitude of tides and waves), season, date and times of the image taken.
+		\item[Eighth:] Other \ac{RTM} software may be tested and the modelling outcomes compared.
+		\item[Ninth] Application of various open source \ac{GIS} (\ac{ILWIS}, \ac{GRASS}) could be very useful for the validation of the cartographic results, assessment of accuracy and comparison of various classification methods.
+		\item[Tenth:] The analysis of the health indicators of the seagrass (such as number of leaves per shoot, biomass estimation within the single shot, etc) was not considered in the current work, as it would go beyond the scope of the MSc thesis.
 However, ecological investigations could be used for the assessment of the vulnerability of the seagrass meadows in various locations on Crete.
-		\item[Tenth:] A flexible combination of the multi-scale mapping and results of the fieldwork measurements with \ac{GPS}-referenced underwater footage would enable more profound analysis of the coastal environment on Crete.
+		\item[Eleventh:] A flexible combination of the multi-scale mapping and results of the fieldwork measurements with \ac{GPS}-referenced underwater footage would enable more profound analysis of the coastal environment on Crete.
 	\end{description}
@@
 
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.45]{Fig-3-14.jpg}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.67]{Fig-3-14.jpg}
 		\caption{Ligaria beach, Crete: seafloor types}\label{fig:A.5}
@@
 
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.25]{Fig-3-15.jpg}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.3]{Fig-3-15.jpg}
 		\caption{Various seafloor types; Ligaria beach, Crete}\label{fig:A.6}
@@
 \end{figure}
-\pagebreak
-
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[scale=0.35]{Fig-3-16.jpg}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.3]{Fig-3-16.jpg}
 		\caption{Measurement underwater equipment}\label{fig:A.7}		
@@
 \end{figure}
-\begin{figure}[h]
+\begin{figure}[H]
 	\begin{center}

### 2011-02-09 MScDiplom_Lemenkova.tex
Written 'Discussion', added Appendix 2 (illustrations of some RS concepts), Appendix 10 (snapshots of the working process)

@@
 \usepackage{hyperref}
-\hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue}
+\hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue,pdfkeywords={P.oceanica, seagrass, spectral reflectance, remote sensing, WASI, RTM, Google Earth, mapping, Crete}}
 
@@
 
+\clearpage
+\section*{Dedication}
+	\markright{Dedication}
+	\addcontentsline{toc}{section}{Dedication}
+%\begin{flushright}
+\newenvironment{dedication}
+    {\vspace{16em}\begin{quotation}\begin{center}\begin{em}}
+    {\par\end{em}\end{center}\end{quotation}}
+\begin{dedication}
+\large{\textsc{To my elder brother and best friend, \textbf{Peter Lemenkov}\index[people]{Lemenkov, P.}}}
+\end{dedication}
+%\end{flushright}
+\pagebreak
+
 \section*{Acknowledgements}
@@
 \\ Personally, I would like to mention that I have visited Greece the first time in my life, and I will always remember my Cretan period-2010 and Ligaria beach.\vspace{1mm}\\
-I extend my gratitude to the staff of the Museum of Natural History, the University of Crete, especially to Dr.\textit{Petros Limberakis}\index[people]{Limberakis, Petros} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g.
+I extend my gratitude to the staff of the Museum of Natural History, Department of Zoology of the University of Crete, especially to Dr.\textit{Petros Lymberakis}\index[people]{Lymberakis, P.} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g.
 SCUBA gear diving, instruments for underwater measurements, etc).
@@
 Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and ETM+ scenes.
+
+\section*{Keywords}
+\markright{Keywords}
+\addcontentsline{toc}{section}{Keywords}
+\textit{P.oceanica}, seagrass monitoring, spectral reflectance, remote sensing, WASI, Google Earth, Crete
+
 \pagebreak
@@
 	\acro{CASI}{\href{http://www.borstad.com/casi.html}{Compact Airborne Spectrographic Imager}}
+	\acro{CDOM}{\href{http://en.wikipedia.org/wiki/CDOM}{Colored Dissolved Organic Matter}}
 	\acro{CNES}{\href{http://www.cnes.fr/}{Centre National d'Etudes Spatiales (of France)}}
@@
 	\acro{ESA}{\href{http://www.esa.int/}{European Space Agency}}
+	\acro{FOV}{\href{http://en.wikipedia.org/wiki/FOV}{Field Of View}}
 	%\acro{ERS}{European Remote Sensing satellite}
@@
 		\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)}: statistical data edition and analysis 
-
+		\item \href{http://fwtools.maptools.org/}{FWTools}: for operating images formats	
 		\item \href{http://staruml.sourceforge.net/en/}{StarUML}, \href{http://live.gnome.org/Dia}{Dia}, \href{http://argouml.tigris.org/}						{ArgoUML}: \ac{UML} diagram tool for creating flowcharts of work algorithms
@@
 Supervised, Maximum Likelihood).
+
+\subsection{Research assumptions}\index[concepts]{Research!assumptions}
+The wind roughening the water surface causes sun glitter, and the values spectral reflectance may contain disturbances and noisy if the solar zenith angle is too high, say approaching 60-70 degrees.
+Therefore, the measurements should be made at midday (noon) with as low zenith angle (Fig.\ref{fig:A.2-1})\label{zenith} as possible.
 \pagebreak
@@
 
-\subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}
+\subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}\label{sec:2-2}
 Among various research approaches for the seagrass mapping the most popular methods are based on application of the remote sensing techniques,
@@
 
-\subsection[Radiative Transfer Models (RTM)]{Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing tools}
+\subsection[Radiative Transfer Models (RTM)]{Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing tools}\label{sec:2-3}
 
@@
 well: while in general providing smaller area coverage than satellites images, the Google Earth
-images can be stitched to the composite maps of the acceptable spatial extent, using Google grabbing
-process which allows multiple overlapping of single images over the flight paths, and generates
-mosaics.
+images can be stitched to the composite maps of the acceptable spatial extent, using script written on Python for the \textit{Google Earth} aerial imagery grabbing process (Fig.\ref{fig:A.1} and Fig.\ref{fig:A.2})\label{gdal} which allows multiple overlapping of single images over the flight paths into one swath, and generates imagery mosaics.
 
@@
 meadows. Finally, increasing the degree of fragmentation of the landscapes of 	extit{P.oceanica} meadows
-can be caused by the invasion of alien species, such as Cymodocea nodosa, Caulerpa prolifera,
-Caulerpa taxifolia. Invaders are in general strong colonizers comparing to native \textit{P.oceanica}: they
+can be caused by the invasion of alien species, such as \textit{Cymodocea nodosa, Caulerpa prolifera,
+Caulerpa taxifolia}. Invaders are in general strong colonizers comparing to native \textit{P.oceanica}: they
 occupy much greater habitat space within the regressed meadows of stressed native seagrass
@@
 \section{Discussion}\label{sec:5}
-An approach of the seagrass spectral analysis, monitoring and mapping has been taken in this work, which integrates various research techniques and tools, combining remote sensing methods of spectral analysis of the seafloor cover types, and knowledge of the ecology of \textit{P.oceanica}, with the aim to develop a method of seagrass spectral optical discrimination for the seagrass mapping based on imagery classification.\vspace{1ex}\\
+An approach of the seagrass spectral analysis, monitoring and mapping has been taken in this work, which integrates various research techniques and tools, combining remote sensing methods of spectral analysis of the seafloor cover types, and knowledge of the ecology of \textit{P.oceanica}, with the aim to develop a method of seagrass spectral optical discrimination for the seagrass mapping based on the aerial imagery classification.\vspace{1ex}\\
+In \textit{Chapter 1:} \nameref{sec:1} we discusses main objective of this MSc thesis, which was to study possibilities of seagrass mapping, based on the application of the remote sensing measurement of the seawater optical properties using hyperspectral radiometers.
+The relationship between the optical properties (spectral reflectance) of the seafloor cover types and hydrological parameters of the environment has been studied in order to analyse limitations and capabilities of broadband and narrowband sensors under the conditions of altering environmental parameters.
+For the retrieval of hydrological parameters - seawater radiance, irradiance and spectral reflectance of various seafloor cover types, - spectral optical field measurements were carried out in 2009 at the testing sites in Agia Pelagia, Heraklion and Ligaria beach, Crete Island using \ac{Trios-RAMSES} spectroradiometers.\\
+Further in \textit{Chapter 1:} 1.8 Research assumptions, we assume the constant values of the optical properties of the seawater, phytoplankton, total amount of suspended particles and solids, atmospheric conditions, as well as \ac{CDOM}, which have been set up in modelling part of this work, during \ac{WASI} simulations of various remote sensors.\vspace{1ex}\\
+The second \textit{Chapter 2:} \nameref{sec:2} starts from the review of the available research resources and then discusses various \nameref{sec:2-3} and concludes with reported experience of the remote sensing application towards seagrass mapping: \nameref{sec:2.4.1}.
+In section 2.2 of \textit{Chapter 2:} \nameref{sec:2}, \nameref{sec:2-2}, Measuring water optical properties: hyperspectral radiometers, the \ac{Trios-RAMSES}-ACC-\ac{UV} and \ac{Trios-RAMSES}-ARC spectroradiometers of \ac{Trios-RAMSES} Hyperspectral Sensor series are described.
+\\
+The instruments \ac{Trios-RAMSES} have been used during the fieldwork measurements-2009 for the collection of the reflectance spectra.
+The \ac{Trios-RAMSES}-ACC-\ac{UV} measures spectra in the wavelength domain between 280 nm and 500nm, the \ac{Trios-RAMSES}-ARC, suitable for UV and visible spectra, covers diapason of 320-950 nm with spectral accuracy of 0.3nm (better than 6\%), typical saturation (at 200nm) of  $1Wm^{-2}*nm^{-1}*sr^{-1}$ in 256 channels with a sampling interval of 3.3 nm/pixel and a \ac{FOV} is 7 degrees.
+The spectral reflectance of the seawater with and without sediments, respectively, was calculated by the ratio of the radiance and irradiance values.
+\\
+The reflectance spectra of \textit{P.oceanica} show (Fig.\ref{fig:4.10}, Fig.\ref{fig:A.12}, Fig.
+\ref{fig:A.12}, Fig.\ref{fig:A.13}, Fig.\ref{fig:A.25}, Fig.
+\ref{fig:A.26}) a values maximum between 450 nm and 600 nm, first, because of the chlorophyll absorption peak at 465 and 665nm (Fig.
+\ref{fig:A.2-2})\label{chlorophyll}, secondly, because of the weakening of \ac{CDOM} (or Gelbstoff) in the blue part of the \ac{VIS} spectrum, as it most strongly absorbs short wavelength light in blue to ultraviolet range, and finally, because the absorption of the seawater increases in the red part of the \ac{VIS} spectra.
+The decrease in spectral reflectance values of \textit{P.oceanica} after 660 nm (Fig.\ref{fig:A.13}) is caused by the second absorption peak of phytoplankton.
+The magnitude of the reflectance maximum slightly varies at single variables between about 8 \% and 12 \% (as on Fig.\ref{fig:A.12}) and is probably related to the individual pigmentation and colour composition of single leaves, their structure and geometric orientation, which naturally causes variations in radiance values.
+\\
 The ecological variables, specific to the field environmental conditions, were factored into the \ac{WASI}-based simulation models.
@@
 sun zenith angle), but also height of water column, thus approaching it to the Mediterranean conditions, and chemical content of the seawater (i.e.
-amount of suspended particles, Gelbstoff, etc), which results in models of optical properties of ``seawater with sediments" and ``seawater without sediments".\vspace{1ex}\\
+amount of suspended particles, Gelbstoff, etc), which results in models of optical properties of ``seawater with sediments" and ``seawater without sediments".\\
 The \textit{in-situ} field large-scale matte-level level of seafloor monitoring was then upscaled to airborne Google Earth aerial imagery interpretation, to provide a meadow-level view of seagrass landscapes.
@@
 Therefore, we focused on the first two levels in the current work.
-These different levels have been individually considered in terms of the seagrass spectral discernibility for monitoring and mapping, from which the first two levels have been brought together, to provide a roundup of the achieved results and an overview of what still has to be done in \textit{P.oceanica} seagrass mapping by future researchers (see Recommendations).
+These different scale levels have been individually considered in terms of the seagrass spectral discernibility for monitoring and mapping, and have been brought together, to provide a roundup of the achieved results and an overview of what still has to be done in \textit{P.oceanica} seagrass mapping by future researchers (see \nameref{sec:7}).
 \vspace{1ex}\\
+In \textit{Chapter 3}: \nameref{sec:3}, describing data collection, a videographic approach tested in previous works, has been applied during the summer fieldwork, when we captured imagery and video footage of the seafloor on several routes of the boat in the Ligaria beach.\\
 The finding of \textit{Chapter 4}, \nameref{sec:4} showed that the relationship between the spectral reflectance of various seafloor cover types was tied to depth, i.e.
 water column height.
-Thus, the results of the in-situ fieldwork measurements revealed that spectral reflectance of P.oceanica undergo alterations at depths of 0.5, 2.0 and 3.5m (Fig.\ref{fig:4.9}), which positively answers the first research question of this thesis (``Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ conditions ?").
+Thus, the results of the in-situ fieldwork measurements revealed that spectral reflectance of \textit{P.oceanica} undergo alterations at depths of 0.5, 2.0 and 3.5m (Fig.\ref{fig:4.9}).
+These results indicate that seagrass \textit{P.oceanica} can in general be detected and discriminated from other seafloor cover types with varying environmental conditions, i.e.
+water column height, by hyperspectral spectroradiometers (\ac{Trios-RAMSES}), which positively answers the \textbf{first research question} of this thesis (``Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ conditions ?").
 \vspace{1ex}\\
@@
 Other remote sensors (\ac{MODIS}, \ac{SeaWiFS}) may also be used for the seagrass mapping, because their technical characteristics enable to spectrally discriminate \textit{P.oceanica} seagrass from other seafloor cover types (Fig.\ref{fig:4.6}), particularly carbonate sand as tested in the current work.
-Therefore, the second research question of this MSc thesis (``Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\
+Therefore, the \textbf{second research question} of this MSc thesis (``Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\
 The graphs showing optical properties of seawater with and without sediments (Fig.\ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:4.2}) focused on spectral variability of the water with changed physical and chemical content.
@@
 on Fig.\ref{fig:4.13}, Fig.\ref{fig:A.25}, Fig.\ref{fig:A.27}, Fig.\ref{fig:A.28}) reflect individual health properties of leaves: different nitrogen and chlorophyll content causing diverse colour pigmentation and light absorption, water content in leaves, plant physiological conditions, which vary across seagrass meadow, shoot morphology, etc.
-The differences in spectral reflectance values of the measurements taken on various days might have been caused by the impact of atmospheric conditions, such as solar radiation and sun illumination by different zenith angle.
-
+The differences in spectral reflectance values of the measurements taken on various days might have been caused by the impact of atmospheric conditions, such as solar radiation and sun illumination by different zenith angle.\vspace{1ex}\\
+For further development of the  remote sensing based monitoring and mapping of the seagrass and other seafloor cover types it is desirable to consider the substratum and underwater relief in a more detailed way, i.e.
+bathymetric properties of the testing sizes, because changes in relief directly cause and reinforce the colonisation process of \textit{P.oceanica} meadows.
+Thus, the depression and valleys in relief in general stimulate increase of the sedimentation process.
+In turn, it enables accumulation of necessary nutrient for seagrass plant growth, increases resource allocation to the seagrass roots for better exploitation of pore-water nutrients by the \textit{P.oceanica} shoots.
+As for patchiness in seagrass landscapes, besides natural reasons, the anthropogenic disturbances (e.g.
+ocean trawling) still remain a main cause for landscapes fragmentation.
 \pagebreak
@@
 
-\section*{Keywords}
-\markright{Keywords}
-\addcontentsline{toc}{section}{Keywords}
-\textit{P.oceanica}, seagrass monitoring, spectral reflectance, remote sensing, WASI, Google Earth, Crete
 \pagebreak
@@
 \section{appendices}\label{app}
-
-\subsection{\textit{Appendix-1}. Capturing aerial imagery from the Google Earth: grabbing process}
+\subsection{\textit{Appendix-1} \footnote{The page where the illustration is explained in the main text of MSc thesis is given in angle brackets (e.g., $<$10$>$).}.
+Capturing aerial imagery from the Google Earth}
 %\appendixname{\textbf{1: \\Capturing aerial imagery from the Google Earth: grabbing process}}
@@
 	\centering
-	\includegraphics[scale=0.6]{Fig-22.jpg}
-	\caption{Capturing aerial imagery from the Google Earth: grabbing process}
-	\label{fig:A.1}
+	\includegraphics[scale=0.55]{Fig-22.jpg}
+	\caption{Capturing aerial imagery from the Google Earth: grabbing process. Page: $<$\pageref{gdal}$>$}\label{fig:A.1}
 \end{figure}
@@
 		\includegraphics[scale=0.4]{Script_Gdal.jpg}
-		\caption{Script command of FWTools2.4.7 enabling to reduce the size of the aerial images, from .tif to .ecw format.}
-	\end{center}
-	\label{fig:A.2}
-\end{figure}
-\pagebreak
-
-\subsection[\textit{Appendix-2}. Instrumental adjustment and tuning]{\textit{Appendix-2}. Instrumental adjustment and tuning \\ (Trios-RAMSES setup)}
+		\caption{Script command of \href{http://fwtools.maptools.org/}{FWTools} enabling to reduce the size of the aerial images, from .tif to .ecw format. Page: $<$\pageref{gdal}$>$}\label{fig:A.2}
+	\end{center}	
+\end{figure}
+
+\subsection{\textit{Appendix-2}. Illustrations of some concepts and principles of the remote sensing, relevant for this work}
+%\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.65]{Zenith.jpg}
+		\caption{Schematic illustration of the \textbf{solar zenith angle} and viewing zenith angle for observations from satellite-based instrument.
+Source: \href{http://sacs.aeronomie.be/}{Support to Aviation Control Service (SACS)}.
+Page: $<$\pageref{zenith}$>$}\label{fig:A.2-1}
+	\end{center}
+\end{figure}
+\pagebreak
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.6]{Chlorophyll.jpg}
+		\caption{Absorbance spectra of free chlorophyll a (green) and b (red) in a solvent.
+The spectra of chlorophyll molecules are slightly modified in vivo depending on specific pigment-protein interactions.
+\\ Source: \href{http://en.wikipedia.org/wiki/Portal:Contents/Quick_index}{Wikipedia.org}.
+Page: $<$\pageref{chlorophyll}$>$}\label{fig:A.2-2}
+	\end{center}
+\end{figure}
+\pagebreak
+
+\subsection[\textit{Appendix-3}. Instrumental adjustment and tuning]{\textit{Appendix-3}. Instrumental adjustment and tuning \\ (Trios-RAMSES setup)}
 %\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}
@@
 \subsection{\textit{Appendix-8}. Analysis and classification of the satellite and aerial images}
+%\appendixname{\textbf{8:\\ Satellite and aerial images covering research area of Crete island: selected examples}}
 
@@
 \subsection{\textit{Appendix-9}. Research general workflow}
+%%%%%%%%%%%%%%%%%%%\appendixname{\textbf{9}}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[H]
@@
 	\label{fig:1.11}
+\end{figure}
+\pagebreak
+
+\subsection{\textit{Appendix-10}. Some snapshots of the working process}
+%%%%%%%%%%%%%%%%%%%\appendixname{\textbf{10}}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{photo-1.jpg}
+	\caption{Google Earth aerial imagery grabbing, Heraklion, the University of Crete}
+	\label{fig:A.47}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{photo-2.jpg}
+	\caption{Adjusting waterproof Olympus cameras for underwater seafloor videometric measurements.\\Dr.
+\textit{Petros Lymberakis}\index[people]{Lymberakis, P.} (left) and Dr.
+\textit{Bert Toxopeus}\index[people]{Toxopeus, A.G.} (right)}
+	\label{fig:A.48}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{photo-3.jpg}
+	\caption{Sticking the marker into the seafloor bottom in mattte of \textit{P.oceanica} for depth measurements}
+	\label{fig:A.48}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{photo-5.jpg}
+	\caption{Placing the 0.5m circle and depth marker in the mattte of \textit{P.oceanica} for photo capture}
+	\label{fig:A.48}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{photo-4.jpg}
+	\caption{Monitoring different seafloor cover types: mattte of \textit{P.oceanica} vs carbonate sand. Ligaria beach.}
+	\label{fig:A.48}
 \end{figure}

### 2011-02-10 MScDiplom_Lemenkova.tex
added hyperreferences from pictures to text; changes in 'In-situ observations', p.15, added pictuers of SCUBA and sonar

@@
 	\acro{4AOP}{\href{http://www.noveltis.fr/4AOP/interface/index.php}{Automatized Atmospheric Absorption Atlas Operational Release}} 
-	\acro{6SV1}{Second Simulation of a Satellite Signal in the Solar Spectrum Version1} 
+	\acro{6SV1}{\href{http://6s.ltdri.org/}{Second Simulation of a Satellite Signal in the Solar Spectrum Version1}} 
 	\acro{ASCII}{\href{http://en.wikipedia.org/wiki/ASCII}{American Standard Code for Information Interchange}}
@@
 	\acro{GloVis}{\href{http://glovis.usgs.gov/}{Global Visualization Viewer}}
-	\acro{GPS}{Global Positioning System}
+	\acro{GPS}{\href{http://en.wikipedia.org/wiki/GPS}{Global Positioning System}}
 	\acro{GRASS}{\href{http://grass.fbk.eu/}{Geographic Resources Analysis Support System}}
@@
 	\acro{KOPRA}{\href{http://www.imk-asf.kit.edu/english/312.php}{Karlsruhe Optimized and Precise Radiative transfer Algorithm}}
+	\acro{LAD}{\href{http://en.wikipedia.org/wiki/Least_absolute_deviations}{Least Absolute Deviations}}
 	\acro{LAI}{\href{http://en.wikipedia.org/wiki/Leaf_Area_Index}{Leaf Area Index}}
@@
 	\acro{Trios-RAMSES}{\href{http://trios.de/index.php?option=com_content&view=category&layout=trios&id=47&Itemid=76&lang=en}{RAdiance Measurement SEnSor}}
-	\acro{UML}{Unified Modelling Language}
+	\acro{UML}{\href{http://en.wikipedia.org/wiki/Unified_Modeling_Language}{Unified Modelling Language}}
 	\acro{USGS}{\href{http://www.usgs.gov/}{U.S. Geological Survey}}
@@
 their structure and functioning.\\
-The adaptation to the salt waters is evidently influenced the global distribution of the seagrasses,
+The adaptation to the salt waters is evidently influenced the global distribution of the seagrasses (Fig.\ref{fig:2}),
 limiting it to shallow coastal areas. A number of critical conditions determine growth of the seagrass,
@@
 The seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}) is a key species to inhabit littoral of the
-Mediterranean Sea, see Fig. \ref{fig:1}
-
-\begin{figure}[h]
+Mediterranean Sea, see Fig.\ref{fig:1}\footnote{The page where the illustration is explained in the main text of MSc thesis is given in angle brackets (e.g., Page: $<$\pageref{page-40}$>$}\label{page-1}
+
+\begin{figure}[H]
 	\centering
@@
 	\subfloat {\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-1-1b.jpg}}
-	\caption{Seagrass \textit{Posidonia Oceanica}}
+	\caption{Seagrass \textit{Posidonia Oceanica}. Pages: $<$\pageref{page-1}$>$, $<$\pageref{page-3}$>$, $<$\pageref{page-5}$>$}
 	\label{fig:1}
@@
 \ac{GIS}-based spatial analysis, processing and classification of satellite and aerial photos, as well as videometric underwater footage, towards mapping and environmental monitoring of seagrass \textit{P.
-oceanica} along the selected locations of the northern coasts of Crete island, Greece.
+oceanica} along the selected locations of the northern coasts of Crete island, Greece (Fig.\ref{fig:3.1})\label{page-2}.
 The technical implementation is based on \ac{WASI} \ac{RTM} software, \ac{GIS} (\ac{ILWIS}, Erdas Imagine and ArcGIS), using aerial and satellite images and the results of the underwater videometric measurements.
 
-\subsection{Background}
+\subsection{Background}\label{page-2}
 \subsubsection{Global distribution of the seagrasses}
-Globally, there are 58 recognized and described seagrass species\index[concepts]{Seagrasses!global species} (Fig. \ref{fig:2}), belonging to two orders
+Globally, there are 58 recognized and described seagrass species\index[concepts]{Seagrasses!global species} (Fig.\ref{fig:2}, Fig.\ref{Fig.4})\label{page-2}, belonging to two orders
 (\textit{Hydrocharitales} and \textit{Najadales}), four families (\textit{Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
@@
 	\includegraphics[scale=0.40]{Fig-1-2.jpg}
-	\caption{Distribution of seagrasses in the world. Source:\cite{Green03}\label{Green03} 
-	\label{fig:2}} 
+	\caption{Distribution of seagrasses in the world. Source:\cite{Green03}\label{Green03}. Pages: $<$\pageref{page-1}$>$, $<$\pageref{page-2}$>$}\label{fig:2}
 \end{wrapfigure}
@@
 The distribution of the seagrasses is strongly influenced by several environmental factors, which
-include climate (mostly, tropical and temperate areas\index[concepts]{Seagrasses!distribution of}), bathymetry (shallow shelf zones), hydrological
+include climate (mostly, tropical and temperate areas, Fig.\ref{Fig.4}\index[concepts]{Seagrasses!distribution of}), bathymetry (shallow shelf zones), hydrological
 particularities (chemical content of water, nutrient availability and turbidity of waves), and geological 
@@
 Finally, seagrasses are an important component in the environmental ``food chain" of the coastal
-
-\begin{wrapfigure}{r}{0.35\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{Fig-1-3.jpg}
-	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
-	\label{fig:3}
-\end{wrapfigure}
-
 ecosystems, being the food source for dugongs, turtles, swans and various fish \cite{Cappo95}\label{Cappo95}.
@@
 \subsubsection{Environmental vulnerability of the seagrasses}
-Meadows of \textit{P. oceanica} are subjected to the human activities\index[concepts]{Seagrasses!ecological vulnerability}, as they occur in coastal areas, where
+Meadows of \textit{P. oceanica} (Fig.\ref{fig:1})\label{page-3} are subjected to the human activities\index[concepts]{Seagrasses!ecological vulnerability}, as they occur in coastal areas, where
 they can be affected both directly \cite{Meinesz91}\label{Meinesz91} or indirectly, through the impact on the
@@
 physiological (e.g. chlorophyll and nitrogen content of the leaves) responses towards changed
-environmental conditions \cite{Leoni06,Leoni07}\label{Leoni06} \label{Leoni07}. 
-The detailed research of the fish
-farm-induced decline of the seagrass meadows \cite{Diaz-Almela06}\label{Diaz-Almela06}reports the relationships of
+environmental conditions \cite{Leoni06,Leoni07}\label{Leoni06}\label{Leoni07}. 
+The detailed research of the fish farm-induced decline of the seagrass meadows \cite{Diaz-Almela06}\label{Diaz-Almela06}reports the relationships of
 fish farm organic and nutrient content in the sediments with dynamics of the key seagrass species (\textit{P.
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.25]{Fig-1-4.jpg}
-	\caption{Distribution of seagrass in relation to mean ocean temperature. \\
-	Source: \cite{Orth06}\label{Orth06}}
-	\label{Fig.4}
+\begin{wrapfigure}{r}{0.35\textwidth}
+	\centering
+	\includegraphics[scale=0.35]{Fig-1-3.jpg}
+	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}. Page: $<$\pageref{page-5}$>$}
+	\label{fig:3}
 \end{wrapfigure}
@@
 negative influence on seagrass distribution. Seagrass is exposed to threats from the global climate and
-environmental change, i.e. increases in sea surface temperature; sea level rise; increased frequency
+environmental change, i.e. increases in sea surface temperature\label{page-4}; sea level rise; increased frequency
 and intensity of storms and waves; local decrease of water quality, increased sedimentation,
@@
 seagrass globally \cite{Orth06}\label{Orth06}.
-In tropical areas, where most of seagrasses are located (Fig.\ref{Fig.4}), seagrasses are subjectto catastrophic
+In tropical areas, where most of seagrasses are located (Fig.\ref{Fig.4}), seagrasses are subject to catastrophic
 extinction and loss, due to the cyclones, typhoons, storms, regular floods and increased rainfalls. Recovery from such
@@
 eutrophication and invasive macroalgae. These processes have strong affect on the status of the
-seagrass meadows, and often lead to their complete disappearance \cite{Holmer09}\label{Holmer09}.\\
+seagrass meadows, and often lead to their complete disappearance \cite{Holmer09}\label{Holmer09}.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.25]{Fig-1-4.jpg}
+	\caption{Distribution of seagrass in relation to mean ocean temperature. 	Source: \cite{Orth06}\label{Orth06}. Pages: $<$\pageref{page-2}$>$, $<$\pageref{page-4}$>$}
+	\label{Fig.4}
+\end{wrapfigure}
+
 Other environmental threat for the seagrasses arises as a result of the environmental struggle and
-competition for existence among species. Thus, meadows of P.oceanica in the Mediterranean Sea are
+competition for existence among species. Thus, meadows of \textit{P.oceanica} in the Mediterranean Sea (Fig.\ref{fig:5})\label{page-4} are
 presently facing invasion by alien algal species, particularly in areas where \textit{P.oceanica} is already
@@
 marine coastal environment of Greece, forming, despite its slow growth, the largest, most widespread,
-homogeneous and dense meadows in the Mediterranean between 5 and 40 m depth \cite{DenHartog70}\label{DenHartog70}.
+homogeneous and dense meadows (Fig.\ref{fig:1}) in the Mediterranean Sea between 5 and 40 m depth \cite{DenHartog70}\label{DenHartog70}.
 
@@
 	\includegraphics[scale=0.20]{Fig-1-5.jpg}
-	\caption{Geographical distribution of
-	\textit{P.oceanica}. Source: \cite{Borum04}\label{Borum04}}
+	\caption{Geographical distribution of \textit{P.oceanica}. Source: \cite{Borum04}\label{Borum04}. Pages: $<$\pageref{page-4}$>$, $<$\pageref{page-5}$>$}
 	\label{fig:5}
@@
 Mediterranean waters, thus decreasing its temperature.
-Morphologically\index[concepts]{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots and short rhizomes (0.5-2.0 mm).
+Morphologically\index[concepts]{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots and short rhizomes (0.5-2.0 mm), Fig.\ref{fig:8}, Fig.\ref{fig:3}.\label{page-5}
 The leaves are, perhaps, the most particular characteristics of \textit{P.oceanica}, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:8}(a)): having usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}(c)).
@@
 	\subfloat [Structure and components of \textit{P.oceanica}. Source\cite{DiCarlo04}\label{DiCarlo04}]{\includegraphics[width=0.3\textwidth]{Fig-1-8.jpg}}
-	\caption{Morphology of \textit{P.oceanica}}\label{fig:8}
+	\caption{Morphology of \textit{P.oceanica}. Page: $<$\pageref{page-5}$>$}\label{fig:8}
 \end{figure}
@@
 	\includegraphics[scale=0.25]{Fig-1-9.jpg}
-	\caption{Spectra of the seagrass on different depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}}
+	\caption{Spectra of the seagrass on different depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}. Page: $<$\pageref{page-7}$>$}
 	\label{fig:9}
@@
 However, mapping the seagrass has limitations due to the specific location and characteristics of the
-research object. The remote sensing techniques have traditionally been widely used for the seagrass
-monitoring. 
-The general overview of the application of various remote sensing data types (colour,
+research object. \\
+The remote sensing techniques have traditionally been widely used for the seagrass
+monitoring. The general overview of the application of various remote sensing data types (colour,
 infrared, and black and white images) for the seagrass monitoring shows its high suitability and
@@
 processing of colour aerial photographs for the monitoring of littoral environment with seagrass beds
-have been reported by several authors \cite{Kelly80,Walker89,Green96}\label{Kelly80} \label{Walker89} \label{Green96}.
+have been reported by several authors \cite{Kelly80,Walker89,Green96}\label{Kelly80}\label{Walker89}\label{Green96}.
 Satellite imagery processing has also being used for the seagrass monitoring, due to their accuracy,
@@
 Various research papers report successful application of the image processing for the
-seagrass mapping \cite{Calvo03,Dekker05a,Fornes06,Green96,Jackson07,Jensen95,Kendrick00,Lyzenga81,Malthus03,Matarrese08,Mount03,Pasqualini01,Pasqualini99,Pasqualini98a,Pergent-Martini06,Ralph05,Ribed02,Short01,Walker89}\label{Calvo03}\label{Dekker05a}\label{Fornes06}\label{Green96}\label{Jackson07}\label{Jensen95}\label{Kendrick00} \label{Lyzenga81}\label{Malthus03}\label{Matarrese08}\label{Mount03}\label{Pasqualini01}\label{Pasqualini99}\label{Pasqualini98a}\label{Pergent-Martini06}\label{Ralph05}\label{Ribed02}\label{Short01}\label{Walker89}.
-The application of the remote sensing data towards seagrass mapping is based on the spectral reflectance
- characteristics of the \textit{P.oceanica} seagrass, which enable its spectral discrimination from spectra of other seafloor types. 
+seagrass mapping \cite{Calvo03,Dekker05a,Fornes06,Green96,Jackson07,Jensen95,Kendrick00,Lyzenga81,Malthus03,Matarrese08,Mount03,Pasqualini01,Pasqualini99,Pasqualini98a,Pergent-Martini06,Ralph05,Ribed02,Short01,Walker89}\label{Calvo03}\label{Dekker05a}\label{Fornes06}\label{Green96}\label{Jackson07}\label{Jensen95}\label{Kendrick00}\label{Lyzenga81}\label{Malthus03}\label{Matarrese08}\label{Mount03}\label{Pasqualini01}\label{Pasqualini99}\label{Pasqualini98a}\label{Pergent-Martini06}\label{Ralph05}\label{Ribed02}\label{Short01}\label{Walker89}.
+\\The application of the remote sensing data towards seagrass mapping is based on the spectral reflectance  characteristics of the \textit{P.oceanica} seagrass, which enable its spectral discrimination from spectra of other seafloor types.
 It is proved \cite{Thorhaug07} \label{Thorhaug07} that the spectral signatures of different species of tropical seagrasses are well
@@
 mapping is based on the classifying the pixels on the image according to their spectral reflectance
-values \cite{Fornes06} \label{Fornes06}, so that the seafloor can be divided into several types: sand, rock, P.
-oceanica, other vegetation, etc.
+values (Fig.\ref{fig:9})\cite{Fornes06}\label{Fornes06}, so that the seafloor can be divided into several types: sand, rock, \textit{P.oceanica}, other vegetation, etc\label{page-7}.
 Other example of monitoring of \textit{P.oceanica} using remote sensing techniques \cite{Calvo03}\label{Calvo03}
@@
 successful results of the neural-based classification using Isodata method of supervised classification.
-In case of \textit{P.oceanica} meadows aerial and satellite images are particularly suited for the surveying
+In case of \textit{P.oceanica} meadows aerial and satellite images are particularly suitable for the surveying
 shallow waters \cite{Pasqualini98b}\label{Pasqualini98b} enabling to distinguish seagrass formations and dynamics of
@@
 However, using space borne satellite imagery for the seagrass mapping has certain limitations, due to
-the uncertainties of the spectral signature of the seagrass at higher depths  (Fig.\ref{fig:9}), as well as some
+the uncertainties of the spectral signature of the seagrass at higher depths  (Fig.\ref{fig:9})\label{page-7}, as well as some
 optical particularities, e.g. light refraction under water, unevenness of the water surface, depths, etc.
@@
 		\item[] i) spectra of \textit{P.oceanica}, carbonate sand, silt and other seafloor types
-		\item[] ii) satellite imagery: Landsat TM, Landsat ETM+
+		\item[] ii) satellite imagery: \ac{Landsat TM}, \ac{Landsat ETM+}
 		\item[] iii) aerial photos: Google Earth
@@
 of Crete (Ligaria beach in Agia Pelagia district and Xerocampos), using following methods of the
-remote sensing techniques: spectral modelling by means of the\ac{RTM} \ac{WASI},  underwater videometric measurements made by \ac{Olympus ST}camera, aerial Google Earth and satellite images from different sources, spatial \ac{GIS} and statistical analysis.
+remote sensing techniques: spectral modelling by means of the \ac{RTM} \ac{WASI}, underwater videometric measurements made by \ac{Olympus ST} camera, aerial Google Earth and satellite images from different sources, spatial \ac{GIS} and statistical analysis.
 
@@
 	\begin{itemize}
-		\item[*]Analysing spectral reflectance of the \textit{P.oceanica} and other seafloor cover types means of radiative transfer model tools (RTMs), using WASI.
+		\item[*]Analysing spectral reflectance of the \textit{P.oceanica} and other seafloor cover types means of radiative transfer model tools (\ac{RTM}s), using \ac{WASI}.
 		\item[*]Mapping spatial distribution of the seagrass \textit{P.oceanica} over selected locations along the northern coasts of Crete Island 
@@
 	\begin{itemize}	
-		\item[]1.	To study narrow-band spectral reflectance properties of \textit{P.oceanica} and other seafloor cover types (sand and silt) using WASI water colour simulation software 
-		\item[]2.	To use methods of the in situ diving observations and underwater videometric measurements by \ac{Olympus ST}camera in order to receive large-scale imagery of the P.oceanica mattes 
-		\item[]3.	To apply remote sensing data (Google Earth aerial images, \ac{Landsat TM} and ETM+ satellite images) for the monitoring of the seagrass meadows distribution 
+		\item[]1.	To study narrow-band spectral reflectance properties of \textit{P.oceanica} and other seafloor cover types (sand and silt) using \ac{WASI} water colour simulation software 
+		\item[]2.	To use methods of the in situ diving observations and underwater videometric measurements by \ac{Olympus ST} camera in order to receive large-scale imagery of the \textit{P.oceanica} mattes
+		\item[]3.	To apply remote sensing data (Google Earth aerial images, \ac{Landsat TM} and \ac{Landsat ETM+} satellite images) for the monitoring of the seagrass meadows distribution 
 		\item[]4.	To perform supervised images classification for the thematic mapping of the \textit{P.oceanica} seagrass distribution along the selected locations over the coasts of northern Crete. 
@@
 band.
-Hypothesis Ho: seagrass types are not spectrally distinct from other seafloor types with
+Hypothesis $H_o$: seagrass types are not spectrally distinct from other seafloor types with
 varying \textit{in-situ} conditions, which means
-Ho: μ1 =μ2 =μ3=...= μn.
-The alternative Hypothesis Ha claims the opposite statement: seagrass is spectrally
-distinct with varying \textit{in-situ} conditions, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn.
+$H_o$: μ1 =μ2 =μ3=...= μn.
+The alternative Hypothesis $H_a$ claims the opposite statement: seagrass is spectrally
+distinct with varying \textit{in-situ} conditions, $H_o$: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn.
 The distribution of the spectral responses at every spectral band is assumed to be normal as well as
@@
 aerial images and the results of the underwater videographic measurements towards seagrass
-mapping  (Fig.\ref{fig:1.11}). Image classification is based on the principle of the differentiation between the spectral
-signatures of various seafloor cover types  (Fig.\ref{fig:1.10}). The spectrum of light coming up from the ocean surface in
+mapping  (Fig.\ref{fig:1.11})\label{page-9}. Image classification is based on the principle of the differentiation between the spectral
+signatures of various seafloor cover types  (Fig.\ref{fig:1.10})\label{page-9}. The spectrum of light coming up from the ocean surface in
 shallow waters keeps information on the optical properties of the seawater components and benthic
@@
 	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
-	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}}
+	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}. \\ Page: $<$\pageref{page-9}$>$}
 	\label{fig:1.10}
@@
 regions (Australia). Since that time traditional methods of the seagrass monitoring and common
-recommendations are being elaborated. The development of the underwater SCUBA diving
-equipment and devices enabled to conduct underwater detailed measurements and observations
+recommendations are being elaborated. The development of the underwater \ac{SCUBA} diving
+equipment (Fig.\ref{fig:A.52}) and devices enabled to conduct underwater detailed measurements and observations
 largely contributed to the improvement of the traditional \textit{in-situ} observations of seagrass. From the
@@
 
-\subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}\label{sec:2-2}
+\subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}\label{sec:2-2}\label{page-12}
 Among various research approaches for the seagrass mapping the most popular methods are based on application of the remote sensing techniques,
@@
 	\centering
-	\subfloat [RAMSES-ACC-UV - Hyperspectral UVA/UVB Irradiance Sensor: 280-500 nm]{\label{fig:1-1a}\includegraphics[width=0.4\textwidth]{Fig-2-1.jpg}}
-	\subfloat [RAMSES-ARC - Hyperspectral UV-VIS Radiance Sensor: 320-950 nm.]{\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-2-2.jpg}}
-	\caption{RAMSES Hyperspectral Sensors. Source: \href{http://www.trios.de}{Trios}}
+	\subfloat [\ac{Trios-RAMSES}-ACC-UV - Hyperspectral UVA/UVB Irradiance Sensor: 280-500 nm]{\label{fig:1-1a}\includegraphics[width=0.4\textwidth]{Fig-2-1.jpg}}
+	\subfloat [\ac{Trios-RAMSES}-ARC - Hyperspectral UV-VIS Radiance Sensor: 320-950 nm.]{\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-2-2.jpg}}
+	\caption{\ac{Trios-RAMSES} Hyperspectral Sensors. Source: \href{http://www.trios.de}{Trios}. Pages: $<$\pageref{page-12}$>$, $<$\pageref{page-13}$>$, $<$\pageref{page-17}$>$}
 	\label{fig:2.1}
@@
 \href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. \ac{Trios-RAMSES}\index[concepts]{Seawater optical properties!spectroradiometers!Trios-RAMSES} hyperspectral radiometers
-(\ac{Trios-RAMSES} Hyperspectral \ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor) are small-sized, low power-consuming, flexible for fieldwork yet with high
+(\ac{Trios-RAMSES} Hyperspectral \ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor), Fig.\ref{fig:2.1} are small-sized, low power-consuming, flexible for fieldwork yet with high
 level of precision, specially calibrated for air and for water application as well as color
-measurements  (Fig.\ref{fig:2.1}). These products have been used for the radiance and irradiance
+measurements  (Fig.\ref{fig:2.1})\label{page-13}. These products have been used for the radiance and irradiance
 measurements in Agia Pelagia bay, Crete Island, 2009.
@@
 The traditional methods of \textit{in-situ} seagrass monitoring include in general the following standard
-scheme \cite{McKenzieetal03} \label{McKenzieetal03}. The seagrass is being sampled on the selected sites using transect lines,
-quadrant frame, single point markers, markers, \ac{GPS} and other equipment. The seagrass sampling is
-taken on the regular way with observation points covering the study area with normal distribution.
+scheme \cite{McKenzieetal03}\label{McKenzieetal03}. The seagrass is being sampled on the selected sites using transect lines,
+quadrant frame, single point markers, \ac{GPS}, laptops with installed software (\ac{GIS}), \ac{SCUBA} equipment, \ac{iPAQ} and other auxiliary tools\label{page-15} (Fig.\ref{fig:3.3}).
+The seagrass sampling is taken on the regular way with observation points covering the study area with normal distribution.
 During the measurement process, the vertical photograph of the measurements frame is taken, and the
@@
 reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. \\
-Other methods of seagrass \textit{in-situ} monitoring are based on
-using the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then
-collect the return echoes for the analysis, yet this requires specialized equipment and is mostly used in
-the depth waters, combined with bathymetric measurements. Another limitation of the acoustic
-techniques is that they are not effective at identifying the composition or even the presence of
-submerged aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}, because
-initially acoustic methods were intended for the bathymetric surveying, i.e. they are not adjusted for
-the benthic habitat discrimination. The results of the \textit{in-situ} measurements and observation are usually
+Other methods of seagrass \textit{in-situ} monitoring are based on the application of the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then collect the return echoes for the analysis (Fig.\ref{fig:sonar}).
+One of the examples of the acoustic sonar sensor designed for seabed classification is the British \href{http://sonavision.co.uk/products.asp?cat_id=1}{RoxAnn} system, which application for the seafloor mapping is \href{http://www.ozcoasts.org.au/geom_geol/toolkit/Tech_CA_asc.jsp}{very well described} on the \href{http://www.ozcoasts.org.au/}{OzCoasts website}.\\ However, the applications of the acoustic methods requires specialized expensive equipment and is mostly used in the deep open waters, combined with bathymetric measurements.
+Another limitation of the acoustic
+techniques is that, initially intended for the bathymetric surveying, acoustic equipment is mostly designed for the geomorphological and geological studies of the underwater substratum and are, therefore, more adjusted for the benthic habitat discrimination, and they are not effective for the identifying of the biological species composition or even the presence of aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}.
+The results of the \textit{in-situ} measurements and observation are usually
 managed and treated using integrated \ac{GIS} approach, as, e.g. reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
@@
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
-digitized aerial photographs, \ac{GPS} data, remote sensing and \ac{SCUBA}-based fieldwork measurements.
+digitized aerial photographs, \ac{GPS} data, remote sensing and \ac{SCUBA}-based (Fig.\ref{fig:A.52}) fieldwork measurements.
 \ac{SCUBA}-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
@@
 	\caption{\textit{In situ} optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
-		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
+		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}. Page: $<$\pageref{page-21}$>$}
 	\label{fig:13}
@@
 	\caption{Difference between broadband multispectral and hyperspectral resolution of
-		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
+		spectral signatures. Source: \cite{Clark99}\label{Clark99}. Page: $<$\pageref{page-20}$>$}
 	\label{fig:2.3}
@@
 Spectral measurements\index[concepts]{Remote sensing!spectral measurements} of the target objects are made
-by means of the radiometers, which receive and register the amounts of energy (radiance and
+by means of the radiometers (Fig.\ref{fig:2.1})\label{page-17}, which receive and register the amounts of energy (radiance and
 irradiance) from the objects. Measuring optical properties of of the seawater allows to calculate spectra of the
@@
 \begin{table}[htbp]
-	\caption{Characteristics of selected ocean-colour sensors}
+	\caption{Characteristics of selected ocean-colour sensors. Page: $<$\pageref{page-47}$>$}
 	\begin{center}
@@
 \subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}\index[concepts]{Remote sensing!multispectral imagery} 
-Seagrass mapping using remotely sensed data from multispectral sensors is based on the classification
+Seagrass mapping using remotely sensed data\label{page-18} from multispectral sensors is based on the classification
 and discrimination of the seafloor types using their spectral characteristics in different wavebands.
@@
 	\includegraphics[scale=0.35]{Fig-13.jpg}
- 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
+ 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}. Pages: $<$\pageref{page-18}$>$, $<$\pageref{page-19}$>$}
 	\label{fig:2.4}
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
-Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
+Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial resolution of 10 m (Fig.\ref{fig:2.4}), covering covers a surface area of 3600 km² (60*60 km swath), 26-day
 orbital repeat cycle for nadir viewing and imagery with a spatial resolution 20 - 2.5 m
@@
 	\includegraphics[scale=0.30]{Fig-14.jpg}
-	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
+	\caption{Multispectral vs. hyperspectral band coverage. \\ Source: \cite{Sabins97}\label{Sabins97}. Pages: $<$\pageref{page-20}$>$,  $<$\pageref{page-18}$>$}
 	\label{fig:2.5}
@@
 Thus, the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
-made using IKONOS, \ac{Landsat TM}, and \ac{CASI}\index{concepts}{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
+made using IKONOS, \ac{Landsat TM}, and \ac{CASI}\index[concepts]{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
 spectrum, the best results are achieved by the IKONOS and \ac{CASI}, while \ac{Landsat TM} has not high
@@
 on board calibration \ac{OBC} system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}. 
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-15.jpg}
-	\caption{Reflectance spectra of \textit{Thalassia} seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
-		reflectance. Source:\cite{Yang10}\label{Yang10}}
-	\label{fig:2.6}
-\end{wrapfigure}
-
-This makes \ac{ASTER} imagery especially suitable for the application towards detailed
+This makes \ac{ASTER}\label{page-19} imagery especially suitable for the application towards detailed
 mapping of surface temperature, emissivity and reflectance of objects and bathymetric elevations as
-well. Successful application of the \ac{ASTER} imagery towards mapping of submerged vegetation
+well (Fig.\ref{fig:2.4}). Successful application of the \ac{ASTER} imagery towards mapping of submerged vegetation
 reported, for example, by Hirose \cite{Hirose04}\label{Hirose04}. 
@@
 macrophytes in shallow waters of the Lake Constance.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.15]{Fig-15.jpg}
+	\caption{Reflectance spectra of \textit{Thalassia} seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
+		reflectance. Source:\cite{Yang10}\label{Yang10}. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-54}$>$}
+	\label{fig:2.6}
+\end{wrapfigure}
+
 In regard to the methods chosen for the image interpretation, the supervised classifications proves to
@@
 mapping the submerged landscapes on the basis of pixels classification, the application of data from
-hyperspectral sensors has better potential due to their higher resolution (Fig.\ref{fig:2.5}).
+hyperspectral sensors has better potential due to their higher resolution (Fig.\ref{fig:2.5})\label{page-20}.
 
 \subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery}
-The application of the hyperspectral sensors is most effective and provide more accurate classification
+The application of the hyperspectral sensors \label{page-20} is most effective and provide more accurate classification
 results  (Fig.\ref{fig:2.5}), due to their higher spectral resolution \cite{Bharathi03}\label{Bharathi03} with interval
@@
 \textit{P.oceanica} (Fig.
-\ref{fig:4.6}) and \textit{Thalassia testudinum} (Fig.\ref{fig:2.6}), shows that \textit{P.oceanica} is in generally brighter in the wavelength interval 500-600 nm which can be caused by various colour pigmentation.
+\ref{fig:4.6}) and \textit{Thalassia testudinum}\label{page-21} (Fig.\ref{fig:2.6}), shows that \textit{P.oceanica} is in generally brighter in the wavelength interval 500-600 nm which can be caused by various colour pigmentation.
 The plotted graphs with spectra of the seagrasses \textit{P.oceanica} and \textit{Thalassia testudinum} demonstrate that the dip in the spectral reflectance after ca 600 nm is more pronounced in case of \textit{P.oceanica} seagrass beds (Fig.\ref{fig:4.6}) than in those with \textit{Thalassia testudinum} (Fig.\ref{fig:2.6}), due to absorption caused by pigment of green leaves.
 The remote sensing of the bottom albedo of \textit{P.oceanica} and carbonate sand (Fig.\ref{fig:4.11}) shows clear difference between the carbonate sand and \textit{P.oceanica}, which is evidently much darker.
-To estimate and further analyze particularities of the \textit{P.oceanica} remote sensing reflectance spectra over regions with different bathymetry (0.5, 1.5 and 2.5 meters column depth) we used \ac{WASI} \ac{RTM} software, which incorporates measurements of the seafloor reflectance and water column optical properties.
+To estimate and further analyze particularities of the \textit{P.oceanica} remote sensing reflectance spectra over regions with different bathymetry (0.5, 1.5 and 2.5 meters column depth) we used \ac{WASI} \ac{RTM} software, which incorporates measurements of the seafloor reflectance and water column optical properties.\label{page-22}
 
@@
 \subsection{Study area}
-General area: Island of Crete, Greece (Fig.\ref{fig:3.1}).\\
+General area: Island of Crete, Greece (Fig.\ref{fig:3.1})\label{page-23}.\\
 The study area of the current MSc research is located in the shallow areas of the Ligaria beach, Agia
@@
 	\includegraphics[scale=0.10]{Fig-16.jpg}
-	\caption{Study area: Crete island}
+	\caption{Study area: Crete island. Pages: $<$\pageref{page-2}$>$, $<$\pageref{page-23}$>$}
 	\label{fig:3.1}
@@
 	\includegraphics[scale=0.25]{Fig-17.jpg}
-	\caption{Study area: Ligaria beach, Crete island}
+	\caption{Study area: Ligaria beach, Crete island. Pages: $<$\pageref{page-23}$>$, $<$\pageref{page-24}$>$}
 	\label{fig:3.2}
@@
 
-\subsubsection{Fieldwork equipment}
-The research materials and equipment were provided by \href{http://147.52.130.151/MFIK_en/}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:3.3}): 1) three iPAQs (Fig.\ref{fig:3.3a}); 2)
+\subsubsection{Fieldwork equipment}\label{page-24}
+The research materials and equipment were provided by \href{http://147.52.130.151/MFIK_en/}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items\label{page-24} (Fig.\ref{fig:3.3}): 1) three iPAQs (Fig.\ref{fig:3.3a}); 2)
 three GPS (Fig.\ref{fig:3.3c}); 3) Three underwater video cameras, \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), suitable for photographing
@@
 	\subfloat [Olympus waterproof camera]{\label{fig:3.3d}\includegraphics[width=0.2\textwidth]{Fig-19d.jpg}}
-	\caption{Fieldwork equipment}
+	\caption{Fieldwork equipment. Pages: $<$\pageref{page-15}$>$, $<$\pageref{page-24}$>$, $<$\pageref{page-25}$>$}
 	\label{fig:3.3}
@@
 up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
-Markers and cords for the depths measurements; 5) Waterproof plastic Otterbox  (Fig.\ref{fig:3.3b})for keeping the
+Markers and cords for the depths measurements (Fig.\ref{fig:A.54}); 5) Waterproof plastic Otterbox  (Fig.\ref{fig:3.3b}) for keeping the
 iPAQ dry; 6) \ac{SCUBA} Diving equipment, taken from the Ligaria Diving Center; 7) Boat (Fig.\ref{fig:3.4});
@@
 	\includegraphics[scale=0.60]{Fig-18.jpg}
-	\caption{Boat used for the fieldwork measurements on Crete Island. 2010}
+	\caption{Boat used for the fieldwork measurements on Crete Island. 2010. Page: $<$\pageref{page-24}$>$}
 	\label{fig:3.4}
@@
 
-\subsubsection{Sampling design}
+\subsubsection{Sampling design}\label{page-25}
 The sampling design of the fieldwork was aimed at surveying of the spatial distribution of the
@@
 course plot to cover the area of growing seagrass: shelf areas not deeper than four meters.
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.35]{Sampling_design.jpg}
-		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}\label{fig:A.4}
-	\end{center}	
-\end{figure}
-
 The measurements of the seafloor cover types have been made using underwater video cameras
-Olympus ST 8010, mounted under the boat to capture video footage and imagery. The data records
+Olympus ST 8010 (Fig.\ref{fig:3.3}(d)), mounted under the boat to capture video footage and imagery (Fig.\ref{fig:A.51}). The data records
 were made along each path using these cameras. A videographic approach, tested previously \cite{Norris97}\label{Norris97},
@@
 distribution of the seagrass patches from photo transects, in order to use for the calibration of mapping approaches.
-Three underwater video cameras, located on the bottom of the boat, provided videometric measurements of the
-seafloor during the track, and resulted in a series of consequent overlapping images of the sea bottom under the
+Three underwater video cameras, adjusted on the bottom of the boat (Fig.\ref{fig:A.51}), provided videometric measurements of the
+seafloor during the track (Fig.\ref{fig:3.6}), and resulted in a series of consequent overlapping images of the sea bottom under the
 boat path. The general locations of the sampling sites and routes were selected on the basis of the visual examination
-of the seagrass beds during snorkeling and \ac{SCUBA} diving, recommendations from the Greek collaborators of the \textit{Natural History Museum of Crete}, and available maps covering the research area.
+of the seagrass beds during snorkeling and \ac{SCUBA} diving (Fig.\ref{fig:A.52}, Fig.\ref{fig:A.53}), recommendations from the Greek collaborators of the \textit{Natural History Museum of Crete}, and available maps covering the research area.
 To ensure even and objective selection of sampling sites
 we used transect sampling method, i.e. photographs were taken along the research path.
-Transect method is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.35]{Sampling_design.jpg}
+		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach. Page: $<$\pageref{page-26}$>$}\label{fig:A.4}
+	\end{center}	
+\end{figure}
+
+Transect method (Fig.\ref{fig:A.4}) is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
 enabling the occurrences of the seagrass meadows to be recorded and counted in a systematic and
@@
 	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
-	\caption{Scheme of placements of the Olimpus cameras during the measurements}
+	\caption{Scheme of placements of the Olimpus cameras during the measurements. Page: $<$\pageref{page-26}$>$}
 	\label{fig:3.5}
@@
 The videographic measurements and photos, enabling to detect types of the
-seafloor, were captured by means of three underwater digital video cameras \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), at
+seafloor, were captured by means of three underwater digital video cameras \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d})\label{page-26}, at
 five minutes interval along the tracklog path (Fig.\ref{fig:3.6}).
@@
 encapsulated in a plastic waterproof Otterbox (Fig \ref{fig:3.3b}). The camera were adjusted horizontally by a leveler and
- mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:3.5}). The data
+mounted under the bottom of the boat Fig.\ref{fig:A.51}\label{page-27} to capture video footage and imagery (Fig.\ref{fig:3.5}). The data
 were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state. 
@@
 coverage of the seagrass meadows in the research area. The underwater measurements of the seagrass
-coverage were carried out by taking video footage and photos of ca 0.5 m2 size each (Fig \ref{fig:A.7}).
+coverage were carried out by taking video footage and photos of ca 0.5 m2 size each (Fig.\ref{fig:A.7}, Fig.\ref{fig:A.55}).
 The results of the underwater videometric measurements include series of digital images helping to
-classify seafloor cover types and seagrass meadows, according to the differences in the structure,
+classify seafloor cover types (Fig.\ref{fig:A.56}) and seagrass meadows, according to the differences in the structure,
 colour, texture and shapes of the depicted objects.
@@
 	\includegraphics[scale=0.50]{Fig-21.jpg}
-	\caption{Locations of the video measurements and GPS tracklogs, Ligaria}
+	\caption{Locations of the video measurements and GPS tracklogs, Ligaria. Page: $<$\pageref{page-25}$>$}
 	\label{fig:3.6}
@@
 continuously covering vast areas (Fig.\ref{fig:A.3} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:A.3} (b)) and isolated seagrass patches, or mattes,
-which located separately from broad meadows (Fig.\ref{fig:A.3} (c)).
+which located separately from broad meadows (Fig.\ref{fig:A.3} (c), Fig.\ref{fig:A.56}).
 The results of the underwater videometric measurements of the \ac{Olympus ST} cameras made during the
@@
 The values of their spectral reflectance are accepted as constant. The results of
-spectral measurements enable to analyze, whether P.Oceanica is spectrally distinct from other sea
+spectral measurements enable to analyze, whether \textit{P.oceanica} is spectrally distinct from other sea
 floor types with changing environmental conditions, using the differences in their spectral signatures
@@
 results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
-in water column, water temperature, sun angle, concentration of Gelbstoff (coloured dissolved organic
-matter), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small
+in water column, water temperature, sun angle, concentration of Gelbstoff (\ac{CDOM}), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small
 particle \cite{Gege04}\label{Gege04}.
@@
 	\includegraphics[scale=0.35]{GNU-20-new.jpg}
-	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
+	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand.
+Selected measurements sets.
+\href{http://www.gnuplot.info/}{Gnuplot} graph.
+Page: $<$\pageref{page-47}$>$}
 	\label{fig:28}
@@
 	\includegraphics[scale=0.5]{GNU-17.jpg}
-	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}				{Gnuplot}}\label{fig:4.2}
+	\caption{Optical properties of the sea water with sediments, measured in aquarium tank.
+Agia Pelagia district, Crete.
+\href{http://www.gnuplot.info/}				{Gnuplot}.
+Page: $<$\pageref{page-21}$>$}\label{fig:4.2}
 \end{figure}
@@
 	\rowcolors{1}{LightSteelBlue}{Honeydew}
-	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, with average values (for sets 1 - 350). Generalisation up to step 20 nm}\label{tab:4.1}
+	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, with average values (for sets 1 - 350).
+Generalisation up to step 20 nm.
+Page: $<$\pageref{page-40}$>$}\label{tab:4.1}
 	\begin{center}
@@
 	\includegraphics[scale=0.45]{Sensors.png}
-	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths.}
+	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths.
+Page: $<$\pageref{page-21}$>$}
 	\label{fig:4.6}
@@
 Under normal conditions by independent water colour sampling, values of the remote-sensing
-reflectance can vary by 12–24 \% \cite{Toole00}\label{Toole00}, and these variations in the radiometric
+reflectance can vary by 12 - 24 \% \cite{Toole00}\label{Toole00}, and these variations in the radiometric
 determinations are mainly caused by the variety of the environmental factors.
@@
 depths, concentration of suspended particles in water column, water temperature, sun angle,
-concentration of Gelbstoff (coloured dissolved organic matter), concentration of phytoplankton,
+concentration of \ac{CDOM}, concentration of phytoplankton,
 aerosol scattering, exponent of backscattering by small particles, cloudiness, viewing geometry and
@@
 
-\subsubsection{Water color spectral simulation using WASI}\index[concepts]{Data!processing}
+\subsubsection{Water color spectral simulation using WASI}\index[concepts]{Data!processing}\label{page-35}
 The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
@@
 	\includegraphics[scale=0.30]{Fig-30.jpg}
-	\caption{Main interface of WASI water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m}
+	\caption{Main interface of \ac{WASI} water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m. Page: $<$\pageref{page-35}$>$}
 	\label{fig:4.7}
@@
 from 4.4 \%  minimal up to 29 \% \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as $35-45\,^{\circ}$
-and reflection factor 0.0201. \textit{P.oceanica}. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
+and reflection factor 0.0201. \textit{P.oceanica}\label{page-36}. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
 The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:4.8}).
@@
 	\subfloat [Bottom reflectance at depth 0.5-4.0 meters]{\label{fig:31b}\includegraphics[width=0.4\textwidth]{Fig-31b.jpg}}
-	\caption{Sea water physical properties, modeled by WASI}
+	\caption{Sea water physical properties, modeled by \ac{WASI}. Page: $<$\pageref{page-36}$>$}
 	\label{fig:4.8}
@@
 	\rowcolors{1}{Aquamarine}{Azure}
-	\caption{Model-specific parameters of water WASI adjusted to simulate environment of the Mediterranean Sea along Crete}
+	\caption{Model-specific parameters of water WASI adjusted to simulate environment of the Mediterranean Sea along Crete. Page: $<$\pageref{page-36}$>$}
 	\centering
@@
 the seagrass. Therefore, the most healthy and suitable areas for the seagrass grow are located at depths
-lesser than 4 meters (Fig.\ref{fig:4.9}).
+lesser than 4 meters (Fig.\ref{fig:4.9})\label{page-37}.
 
@@
 	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d.jpg}}
-	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths, simulated for broadband sensors.}
+	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths, simulated for broadband sensors. Page: $<$\pageref{page-37}$>$}
 	\label{fig:4.9}
@@
 
-\paragraph{Spectral discrimination of \textit{P.oceanica}}
+\paragraph{Spectral discrimination of \textit{P.oceanica}}\label{page-39}
 The patches of white sandy bottoms of the seafloor, studied in the fieldwork in Ligaria beach, are
@@
 	\caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and 	measured values (dotted): multiplot of measurement sets 200-300.
-\href{http://www.gnuplot.info/}{Gnuplot}}
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-37}$>$}
 	\label{fig:4.10}
@@
 spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
-The statistical calculators were mainly made by means of Gretle\index[concepts]{Data!statistical analysis!Gretl} and SPSS\index[concepts]{Data!statistical analysis!SPSS}. 
+The statistical calculators were mainly made by means of \ac{GRETL}\index[concepts]{Data!statistical analysis!Gretl} and SPSS\index[concepts]{Data!statistical analysis!SPSS}. 
 The Open Office was used for preliminary data view and preprocessing, and included following computations
 at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). 
-The statistical pre-processing and analysis were made to display the mean values of spectral reflectance,
+The statistical pre-processing\label{page-40} and analysis were made to display the mean values of spectral reflectance,
 which were used for comparison with reflectance data of various seafloor cover types as well as
-seawater with and without sediment (e.g. Fig..\ref{fig:A.16}).
+seawater with and without sediment (e.g. Fig.\ref{fig:A.16}).
 In the statistical analysis of the raw observed data pool\index[concepts]{Data!measurements pool} of the values of spectral reflectance (about 400 single measurements for \textit{P.oceanica}, e.g.
@@
 However, in the statistical analysis of the observed data our goal is to estimate these parameters using a finite amount of information available to us: fieldwork observations.
-\\
-Thus, we collect a number of realizations (a sample of 25 measurements for both sand and \textit{P.oceanica}) and then estimate the statistical parameters (\nameref{app}). \\
+Thus, we collected a number of realizations (a sample of 25 measurements for both sand and \textit{P.oceanica}) and then estimate the statistical parameters (\nameref{app}). \\
 Finally we analyzed about 400 datasets totally (example of carbonate sand on Fig.\ref{fig:4.14}, \textit{P.oceanica} - on  Fig.\ref{fig:4.13}) made during different days.
 Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random.
-Thus, we analyzed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implemented in Gretl software (Tab.\ref{tab:19}). 
+Thus, we analyzed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implemented in \ac{GRETL} software (Tab.\ref{tab:19}).
 Their probability expresses uncertainty about the possible values of the spectral reflectance of \textit{P.oceanica} and sand, respectively.
@@
 Each time we observe the outcome of a random variable\index[concepts]{Data!statistical analysis!variables} of spectral reflectance, we obtain an observation, which is hence no longer random. 
-We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the \nameref{app}: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyze the distribution of values of spectral reflectance of \textit{P.oceanica}.
+\label{statistics-1}
+We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the \nameref{app}\label{page-41}: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyze the distribution of values of spectral reflectance of \textit{P.oceanica}.
 A probability distribution\index[concepts]{Data!statistical analysis!probability distribution}, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
-The \textit{least absolute deviations (LAD)}\index[concepts]{Data!statistical analysis!LAD}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
+The \ac{LAD}\index[concepts]{Data!statistical analysis!LAD}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
 Non-linear logistic analysis\index[concepts]{Data!statistical analysis!non-linear logistic analysis} is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
@@
 The \textit{autocorrelation}\index[concepts]{Data!statistical analysis!autocorrelation}, a cross-correlation\index[concepts]{Data!statistical analysis!cross-correlation} of an observed value with itself, made on two different days (14th and 15th October), shows the similarity between observations as a function of the time separation between them.
-As can be seen on Fig.\ref{fig:A.21}, Fig.\ref{fig:A.22} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
+As can be seen on Fig.\ref{fig:A.21}\label{A-23}, Fig.\ref{fig:A.22}\label{A-24} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
 The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{GRETL} software.
@@
 	\includegraphics[scale=0.5, angle=90]{UML_StatAn.jpg}
-	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
+	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML}).
+Page: $<$\pageref{page-40}$>$}
 	\label{fig:4.12}
@@
 It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
-To estimate spectral density of a spectral reflectance measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30}), 
-which corresponds to the general spectrum of the observations with representation of a variable quantity.  
-Thus, the periodogram (Fig.
-\ref{fig:A.31}) of the series of observation\index[concepts]{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to
-exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:A.12} ) 
+To estimate spectral density of a spectral reflectance measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30}), which corresponds to the general spectrum of the observations with representation of a variable quantity.
+Thus, the periodogram (Fig.\ref{fig:A.31}) of the series of observation\index[concepts]{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to exclude (can be seen clearly at the end of the graphs, e.g.
+Fig.\ref{fig:A.12} )
 and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:A.14}, enlarged part of 500-660nm).\\
@@
 The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
-areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1, Q3, shaded areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution
-(spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
+areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1, Q3, shaded areas on graph, Fig.
+\ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution (spectral reflectance) for each spectral band.
+In x-axis displayed are the areas of 400 – 950 nm in
 spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
@@
 Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
+\href{http://www.gnuplot.info/}{Gnuplot}­.
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-41}$>$}
 		\label{fig:4.13}
@@
 
-A statistical hypothesis\index[concepts]{Data!statistical analysis!hypothesis testing} test has been applied for the making decision and controlling the wealth of the
-observational data of the hyperspectral measurements of the water reflectance.
+A statistical hypothesis\index[concepts]{Data!statistical analysis!hypothesis testing} test has been applied for the making decision and controlling the wealth of the observational data of the hyperspectral measurements of the water reflectance.
 The received results are statistically significant if they are unlikely to have occurred by chance alone,
@@
 Shown mid spread of the statistical quartiles Q1 and Q3 (vertical dashes) and	mean value within the range (red bold dots).
-\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-41}$>$­}
 		\label{fig:4.14}
@@
 
-Assuming that the null hypothesis, Hypothesis Ho, is true, which is “the spectral distinguishability of
+Assuming that the null hypothesis, Hypothesis $H_o$, is true, which is “the spectral distinguishability of
 the seagrass \textit{P.oceanica} from other seafloor types is not changing with varying \textit{in-situ} conditions,
-Ho: μ1 =μ2 =μ3=...= μn.”, what is the probability of observing the values of spectral reflectance of
+$H_o$: μ1 =μ2 =μ3=...= μn.”, what is the probability of observing the values of spectral reflectance of
 \textit{P.oceanica} at various depths, for the test statistic, that is at least as extreme as the values of spectral
-reflectance of \textit{P.oceanica} at these depths, that was actually observed? The alternative Hypothesis Ha
+reflectance of \textit{P.oceanica} at these depths, that was actually observed? The alternative Hypothesis $H_a$
 Fig.15. Multiplot display of spectral reflectance of \textit{P.oceanica}. Measurement sets 1-100, 15th
@@
 claims the opposite statement: “the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly
-changing with varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
+changing with varying \textit{in-situ} conditions, i.e. increasing depth, $H_o$: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
 Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.
@@
 spectral reflectance of \textit{P.oceanica} to be more or less spectrally distinct from other seafloor cover
-types with changing depth. W are checking the statement Hypothesis Ho, is true, which is “the
+types with changing depth. W are checking the statement Hypothesis $H_o$, is true, which is “the
 spectral distinguishability of the seagrass \textit{P.oceanica} from other seafloor types is not changing with
-varying \textit{in-situ} conditions, Ho: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
-P(reject H0|H0 is valid)= P(X$>$c|p=)=.05
+varying \textit{in-situ} conditions, $H_o$: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
+P(reject $H_o$|$H_o$ is valid)= P(X$>$c|p=)=.05
 where c comes for critical values and p=probability.
-The result P=.05 is (Tab.\ref{tab:8}), hence, very small, which makes the statement of Hypothesis Ho highly unlikely
+The result P=.05 is (Tab.\ref{tab:8}), hence, very small, which makes the statement of Hypothesis $H_o$ highly unlikely
 (less than 1 in a 10 chance). The one-way \ac{ANOVA} highlighted a significant difference between data
 of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
-Hypothesis Ha=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with
-varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
+Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with
+varying \textit{in-situ} conditions, i.e. increasing depth, $H_o$: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
 
@@
 As a result of the statistical testing, we came to the following conclusion.
-The Hypothesis Ha is true, which claims that the spectral discernibility of the seagrass 	\textit{P.oceanica} is
-distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ1≠μ2≠μ3≠...≠ μn .
+The Hypothesis $H_a$ is true, which claims that the spectral discernibility of the seagrass \textit{P.oceanica} is
+distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, $H_o$: μ1≠μ2≠μ3≠...≠ μn .
 Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
@@
 \rowcolors{1}{Khaki}{LightYellow}
-\caption{Statistical analysis of the measurements of spectral reflectance of \textit{P.oceanica} (fragment)}
+\caption{Statistical analysis of the measurements of spectral reflectance of \textit{P.oceanica} (fragment). Page: $<$\pageref{page-40}$>$}
 \centering
@@
 The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
-types as the different sensors would percept them: MODIS\index[concepts]{Remote sensing!multispectral imagery!MODIS}, ASTER\index[concepts]{Remote sensing!multispectral imagery!ASTER}, MERIS\index[concepts]{Remote sensing!multispectral imagery!MERIS}, SeaWiFS\index[concepts]{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index[concepts]{Remote sensing!multispectral imagery!CZCS}.\\
+types as the different sensors would percept them: \ac{MODIS}\index[concepts]{Remote sensing!multispectral imagery!MODIS}, \ac{ASTER}\index[concepts]{Remote sensing!multispectral imagery!ASTER}, \ac{MERIS}\index[concepts]{Remote sensing!multispectral imagery!MERIS}, \ac{SeaWiFS}\index[concepts]{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index[concepts]{Remote sensing!multispectral imagery!CZCS}.\\
 These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
 These models simulate spectral views of the chosen sensors, how these sensors will ``see" seagrass (Fig.\ref{fig:28}) and sand as
-pixels, with accepted default atmosphere and water column effects (given by WASI software). In such
+pixels, with accepted default atmosphere and water column effects (given by \ac{WASI} software). In such
 a way we defined an empirical upper limit to the discriminative potential of these sensors. To create
@@
 
-\subsection{Image processing using Erdas Imagine}
+\subsection{Image processing using Erdas Imagine}\label{page-47}
 The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
@@
 the areas. The classification is based on the spectral properties of the \textit{P.oceanica}, such as brightness,
-colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping
+colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping\label{page-48}
 includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
@@
 	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
-	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
+	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images. Page: $<$\pageref{page-48}$>$}
 	\label{fig:38}
@@
 seagrass meadows provided on the basis of the available and collected field data. The classification
-has been completed using ENVI and Erdas Imagine software.
+has been completed using Erdas Imagine software.
 
@@
 high resolution aerial color Google Earth photography, spaceborne satellite imagery, assessment of
-spectral signatures using \ac{WASI} software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc \ac{GIS} based mapping.
+spectral signatures using \ac{WASI} software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}, Fig.\ref{fig:42}) and Arc \ac{GIS} based mapping.
 The use of \ac{GIS} for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
@@
 
-\subsubsection{Upscale GIS mapping of the seagrass landscapes}\index[concepts]{Seagrasses!spatial structure}
+\subsubsection{Upscale GIS mapping of the seagrass landscapes}\index[concepts]{Seagrasses!spatial structure}\label{page-48}
 Although ocean pelagic landscapes have a high degree of spatial variance and less structurally
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%% Discussion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}\label{sec:5}
@@
 The goal of this MSc research was to explore the perspectives, advantages and limitations of the narrow-band and broadband sensors for the environmental mapping and monitoring of \textit{P.oceanica} seagrass along the coasts of Crete Island.
-The research outcome demonstrated that the application of the remote sensing data from the broadband sensors is highly advantageous for the seagrass mapping, the spectral discrimination of \textit{P.oceanica} from other seafloor cover types is possible at diverse and changing environmental conditions, and that \textit{P.oceanica} is spectrally distinct from other seagrass species as well (\textit{Thalassia testudinum}).\vspace{1ex}\\
+The research outcome demonstrated that the application of the remote sensing data from the broadband sensors is highly advantageous for the seagrass mapping, the spectral discrimination of \textit{P.oceanica} from other seafloor cover types is possible at diverse and changing environmental conditions, and that \textit{P.oceanica} is spectrally distinct from other seagrass species as well (\textit{Thalassia testudinum}), Fig.\ref{fig:2.6}\label{page-54}.\vspace{1ex}\\
 The \ac{RTM} software is a powerful means for analyzing spectral signatures of various seafloor types and enabling simulations of data received from broadband and narrowband remote sensors.\\ The example of application of \ac{WASI} \ac{RTM}, given in this work, is an achievement of the research insight towards the spectral properties of \textit{P.oceanica} and other bottom cover types, enabling to discriminate them from each other with changing environmental conditions.
@@
 \section{appendices}\label{app}
-\subsection{\textit{Appendix-1} \footnote{The page where the illustration is explained in the main text of MSc thesis is given in angle brackets (e.g., $<$10$>$).}.
-Capturing aerial imagery from the Google Earth}
+\subsection{\footnote{The page where the illustration is explained in the main text of MSc thesis is given in angle brackets (e.g., $<$10$>$).} Capturing aerial imagery from the Google Earth}
 %\appendixname{\textbf{1: \\Capturing aerial imagery from the Google Earth: grabbing process}}
@@
 
-\subsection{\textit{Appendix-2}. Illustrations of some concepts and principles of the remote sensing, relevant for this work}
-%\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}
+\subsection{Illustrations of some concepts and principles of the remote sensing, relevant for this work}
+%%%%%%%%%%  \appendixname{Illustrations of some concepts and principles of the remote sensing}  %%%%%%%%%%%%%%
 
@@
 \end{figure}
-\pagebreak
-
-\subsection[\textit{Appendix-3}. Instrumental adjustment and tuning]{\textit{Appendix-3}. Instrumental adjustment and tuning \\ (Trios-RAMSES setup)}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.6]{sonar.jpg}
+		\caption{Example of sonar beam acoustic systems used for
+mapping seagrasses habitat boundaries.
+\\ Source: \href{http://kurrawa.gbrmpa.gov.au/corp_site/info_services/publications/reef_research/issue1_98/1wot.html}{Reef Research}.
+Page: $<$\pageref{page-15}$>$}.\label{fig:sonar}
+	\end{center}
+\end{figure}
+
+\pagebreak
+
+\subsection[Instrumental adjustment and tuning]{Instrumental adjustment and tuning. (Trios-RAMSES setup)}
 %\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}
@@
 
-\subsection[\textit{Appendix-3}. Types of seagrass structural patterns]{\textit{Appendix-3}. Types of seagrass structural patterns, Ligaria beach, Crete}
+\subsection[Types of seagrass structural patterns]{Types of seagrass structural patterns, Ligaria beach, Crete}
 %\appendixname{\textbf{3: \\Types of seagrass structural patterns, Ligaria beach, Crete}}
@@
 
-\subsection{\textit{Appendix-4}. Results of videographic measurements: seafloor types on Crete island}
+\subsection{Results of videographic measurements: seafloor types on Crete island}
 %\appendixname{\textbf{5: \\Results of videographic measurements: seafloor types on Crete island}}
@@
 		\includegraphics[scale=0.67]{Fig-3-14.jpg}
-		\caption{Ligaria beach, Crete: seafloor types}\label{fig:A.5}
+		\caption{Ligaria beach, Crete: seafloor types. Page: $<$\pageref{page-39}$>$}\label{fig:A.5}
 	\end{center}
@@
 
-\subsection{\textit{Appendix-5}. Data pre-processing}
+\subsection{Data pre-processing}
 %\appendixname{\textbf{6: \\Data pre-processing}}
@@
 	\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
-		seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
+		seagrasses (various species), sand, silt and artificial spectrum of constant albedo. Page: }
 	\label{fig:A.9}
@@
 
-\subsection{\textit{Appendix-6}. Data processing and statistical evaluation: Results of the measurements of the seafloor optical properties.}
+\subsection[Data processing and statistical evaluation]{Data processing and statistical evaluation: measurements of the seafloor optical properties.}
 %\appendixname{\textbf{7:\\ Data processing and statistical evaluation: \\ Results of the measurements of the seafloor optical properties.}}
@@
 	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis}
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Two complimentary graphs below show the results of the statistical analysis.
+Page: $<$\pageref{page-35}$>$}
 	\label{fig:27}
@@
 		\includegraphics[scale=0.20]{App-1.jpg}
-		\caption{Values of spectral reflectances of seagrass \textit{P.oceanica} (fragment)}
+		\caption{Values of spectral reflectances of seagrass \textit{P.oceanica} (fragment). Page: $<$\pageref{page-40}$>$}
 		\label{fig:A.11}
@@
 		\includegraphics[scale=0.25]{GNU-401-420.jpg}
-		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
+		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-21}$>$}
 		\label{fig:A.12}
@@
 Shown mid-spread of  statistical quartiles Q1 and Q3, min and max values within the range.
-\href{http://www.gnuplot.info/}{Gnuplot}}
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Page: $<$\pageref{page-21}$>$}
 		\label{fig:A.13}
@@
 
-\begin{figure}[h]
-	\begin{center}
-	\includegraphics[scale=0.20]{GNU-Stat-M-401-420.jpg}
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[scale=0.27]{GNU-Stat-M-401-420.jpg}
 	\caption{Fragment of the statistical analysis of the \textit{P.oceanica} reflectance. Example of data set 401-420. 
-		Visualization of the measured data together with statistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
+		Visualization of the measured data together with statistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-21}$>$}
 	\label{fig:A.15}
@@
 	\centering
-	\includegraphics[scale=0.25]{GNU-Radiance_Bezier.jpg}
-	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}}
+	\includegraphics[scale=0.35]{GNU-Radiance_Bezier.jpg}
+	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-40}$>$}
 	\label{fig:A.16}
@@
 \begin{longtable}{|c|c|c|c|c|c|c|}
-	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October}\label{tab:A.8} \\
+	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, sets 1-350).
+Wavelength step: 3 nm.
+Measured on Agia Pelagia beach, 15th October.
+Page: $<$\pageref{page-40}$>$}\label{tab:A.8} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 \begin{table}
-	\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16. \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16.
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page $<$\pageref{page-41}$>$}
 	\begin{center}
@@
 
-\begin{table}[htbp]
+\begin{table}[H]
 	\rowcolors{1}{Ivory}{GhostWhite}
@@
 \end{table}
-
 \begin{figure}[H]
@@
 	\includegraphics[scale=0.45]{Autocorr.jpg}
-	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
+	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+Visualization in \href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{A-23}$>$}
 	\label{fig:A.21}
 \end{figure}
-\pagebreak
-\newpage
 
@@
 	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3.
-\href{http://www.spss.com/}{SPSS}}\label{tab:14}
+\href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{page-41}$>$}\label{tab:14}
 	\begin{center}
@@
 	\includegraphics[scale=0.45]{Partial_Corr.jpg}
-	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
+	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+Visualization in \href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{A-24}$>$}
 	\label{fig:A.22}
@@
 \pagebreak
-\newpage
 
@@
 	\begin{center}
-		\includegraphics[scale=0.25]{GNU-11.jpg}
-		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
+		\includegraphics[scale=0.35]{GNU-11.jpg}
+		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display. Page: $<$\pageref{page-21}$>$­}
 		\label{fig:A.25}
@@
 	\begin{center}
-		\includegraphics[scale=0.25]{GNU-12.jpg}
+		\includegraphics[scale=0.35]{GNU-12.jpg}
 		\caption{Remote sensing reflectance of \textit{P.oceanica}.
@@
 \\Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range.
-\href{http://www.gnuplot.info/}{Gnuplot}­}
+\\ \href{http://www.gnuplot.info/}{Gnuplot}­.
+Page: $<$\pageref{page-21}$>$}
 		\label{fig:A.26}
@@
 \end{figure}
+%%% Gretl start
+\begin{table}[htbp]
+	\rowcolors{1}{Ivory}{GhostWhite}
+	\caption{Basic mathematical formulae used for statistical analysis of the measurement set of the spectral reflectance. Page: $<$\pageref{page-41}$>$}
+	\label{tab:13}
+	\begin{center}
+	\begin{tabular}{|c|c|}
+		\hline\hline
+		Basic Statistics & Math Formulae \\ \hline\hline
+		\hline
+
+		\vspace{1ex}
+
+		Mean & $\Sigma(x_i/n=\bar x)$ \\ \hline
+		Variance & $1 \over n-1$ $\Sigma(x_i-\bar x)=s^2_x$  \\ \hline
+		Standard Deviation & $s=s^2$ \\ \hline
+		Coefficient of Variation & $s/\bar x$ \\ \hline
+		Skewness & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^3$  \\ \hline
+		Excess Kurtosis & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^4-3$  \\ \hline
+	\end{tabular}
+	\end{center}
+	\label{tab:19}
+\end{table}
+%%% Gretl end
 
@@
 	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textit{P.oceanica} (15.X.).
-\href{http://gretl.sourceforge.net/}{Gretl}}
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
 	\begin{center}
@@
 	\begin{center}
-		\includegraphics[scale=0.4]{GNU-14.jpg}
+		\includegraphics[scale=0.5]{GNU-14.jpg}
 		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
@@
 Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
+\href{http://www.gnuplot.info/}{Gnuplot} visualization.
+Page: $<$\pageref{page-39}$>$­}
 		\label{fig:A.27}
@@
 
-%%% Gretl start
-\begin{table}[htbp]
-	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Basic mathematical formulae used for statistical analysis of the measurement set of the spectral reflectance.}
-	\label{tab:13}
-	\begin{center}
-	\begin{tabular}{|c|c|}
-		\hline\hline
-		Basic Statistics & Math Formulae \\ \hline\hline
-		\hline
-
-		\vspace{1ex}
-
-		Mean & $\Sigma(x_i/n=\bar x)$ \\ \hline
-		Variance & $1 \over n-1$ $\Sigma(x_i-\bar x)=s^2_x$  \\ \hline
-		Standard Deviation & $s=s^2$ \\ \hline
-		Coefficient of Variation & $s/\bar x$ \\ \hline
-		Skewness & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^3$  \\ \hline
-		Excess Kurtosis & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^4-3$  \\ \hline
-	\end{tabular}
-	\end{center}
-	\label{tab:19}
-\end{table}
-%%% Gretl end
-\pagebreak
-
-\begin{table}[htbp]
+\pagebreak
+
+\begin{table}[H]
 	\rowcolors{1}{Ivory}{GhostWhite}
 	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).
-\href{http://www.spss.com/}{SPSS}}
+\href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{page-41}$>$}
 	\label{tab:13}
@@
 
-\begin{table}[htbp]
+\begin{table}[H]
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
+	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+\href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{page-41}$>$}
 	\begin{center}
@@
 	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).
-\href{http://www.spss.com/}{SPSS}}
+\href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{page-41}$>$}
 		\begin{center}
@@
 \end{table}
-\pagebreak
-
-\begin{table}[htbp]
+
+\begin{table}[H]
 	\rowcolors{1}{Ivory}{GhostWhite}
@@
 	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica} (15.X.).
-\href{http://gretl.sourceforge.net/}{Gretl}}
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
 	\begin{center}
@@
 		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textit{P.oceanica}.
-\href{http://gretl.sourceforge.net/}{Gretl}}
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
 			Model 2: Tobit, using observations 1:01--8:23 ($T$ = 191)\\
@@
 Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-\href{http://www.gnuplot.info/}{Gnuplot} visualization­}
+\href{http://www.gnuplot.info/}{Gnuplot} visualization­.
+Page: $<$\pageref{page-39}$>$}
 		\label{fig:A.28}
@@
 	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textit{P.oceanica}.
-\href{http://gretl.sourceforge.net/}{Gretl}}
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
 	\label{tab:26}
@@
 	\includegraphics[scale=0.4]{G-autocorr-sand51-75.jpg}
-	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand. Series 51-75, variable 27. \href{http://gretl.sourceforge.net/}{Gretl} visualization­}
+	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand.
+Series 51-75, variable 27.
+\href{http://gretl.sourceforge.net/}{Gretl} visualization.
+Page: $<$\pageref{page-41}$>$­}
 	\label{fig:A.29}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of carbonate sand, \href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
 	\begin{center}
@@
 \begin{table}[H]
-	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
 	\begin{center}
@@
 Wavelength step: 3 nm.
-Measured on Agia Pelagia beach, 14th October}\label{tab:A.29} \\
+Measured on Agia Pelagia beach, 14th October.
+Page: $<$\pageref{page-40}$>$}\label{tab:A.29} \\
  		 \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 	\includegraphics[scale=0.5]{Fig-26.jpg}
-	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES. \href{http://www.gnuplot.info/}{Gnuplot}}
+	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach.
+Results of single measurement set made by spectroradiometer Trios-RAMSES.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Page: $<$\pageref{page-39}$>$}
 	\label{fig:A.35}
@@
 
-\subsection{\textit{Appendix-7}. Satellite and aerial images covering research area of Crete island: selected examples}
+\subsection{Satellite and aerial images covering research area of Crete island: selected examples}
 %\appendixname{\textbf{8:\\ Satellite and aerial images covering research area of Crete island: selected examples}}
@@
 		\includegraphics[scale=0.20]{Locations_Meas.jpg}
-		\caption{Locations of selected measurements, visualization on Google Earth­}\label{fig:A.36}
-	\end{center}	
-\end{figure}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.25]{App-8.jpg}
-		\caption{Random mosaic of selected aerial Google Earth images}\label{fig:A.37}	
+		\caption{Locations of selected measurements, visualization on Google Earth­. Page: $<$\pageref{page-47}$>$}\label{fig:A.36}
+	\end{center}	
+\end{figure}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.30]{App-8.jpg}
+		\caption{Random mosaic of selected aerial Google Earth images. Page: $<$\pageref{page-47}$>$}\label{fig:A.37}	
 	\end{center}	
@@
 
-\subsection{\textit{Appendix-8}. Analysis and classification of the satellite and aerial images}
+\subsection{Analysis and classification of the satellite and aerial images}
 %\appendixname{\textbf{8:\\ Satellite and aerial images covering research area of Crete island: selected examples}}
@@
 		\includegraphics[scale=0.25]{Fig-39.jpg}
-		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project}
+		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project.
+Page: $<$\pageref{page-48}$>$}
 		\label{fig:42}
@@
 		\includegraphics[scale=0.25]{Fig-36.jpg}
-		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }
+		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete.
+Page: $<$\pageref{page-48}$>$}
 		\label{fig:39}
@@
 
-\subsection{\textit{Appendix-9}. Research general workflow}
+\subsection{Research general workflow}
 %%%%%%%%%%%%%%%%%%%\appendixname{\textbf{9}}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\includegraphics[scale=0.57]{UML_GIS.png}
-	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
+	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}. Page: $<$\pageref{page-9}$>$}
 	\label{fig:1.11}
@@
 
-\subsection{\textit{Appendix-10}. Some snapshots of the working process}
+\subsection{Some snapshots of the working process}
 %%%%%%%%%%%%%%%%%%%\appendixname{\textbf{10}}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\caption{Google Earth aerial imagery grabbing, Heraklion, the University of Crete}
-	\label{fig:A.47}
+	\label{fig:A.50}
 \end{figure}
@@
 \textit{Petros Lymberakis}\index[people]{Lymberakis, P.} (left) and Dr.
-\textit{Bert Toxopeus}\index[people]{Toxopeus, A.G.} (right)}
-	\label{fig:A.48}
-\end{figure}
+\textit{Bert Toxopeus}\index[people]{Toxopeus, A.G.} (right).
+Pages: $<$\pageref{page-25}$>$, $<$\pageref{page-27}$>$}
+	\label{fig:A.51}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.7]{SCUBA-gear.jpg}
+	\caption{SCUBA gear diving equipment necessary for seagrass monitoring. Source: \href{http://divinginstructortraining.com/}{Aquanauts.com}. Page: $<$\pageref{page-25}$>$}
+	\label{fig:A.52}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.6]{Diving.jpg}
+	\caption{...and it's me, learning diving skills on Ligaria beach, Crete, 2010. On the photo: left. Page: $<$\pageref{page-25}$>$}
+	\label{fig:A.53}
+\end{figure}
+
 \begin{figure}[H]
@@
 	\includegraphics[scale=0.4]{photo-3.jpg}
-	\caption{Sticking the marker into the seafloor bottom in mattte of \textit{P.oceanica} for depth measurements}
-	\label{fig:A.48}
+	\caption{Sticking marker into the sea bottom in matte of \textit{P.oceanica} for depth measurements. Page: $<$\pageref{page-24}$>$}
+	\label{fig:A.54}
 \end{figure}
@@
 	\caption{Placing the 0.5m circle and depth marker in the mattte of \textit{P.oceanica} for photo capture}
-	\label{fig:A.48}
+	\label{fig:A.55}
 \end{figure}
@@
 	\caption{Monitoring different seafloor cover types: mattte of \textit{P.oceanica} vs carbonate sand. Ligaria beach.}
-	\label{fig:A.48}
+	\label{fig:A.56}
 \end{figure}

### 2011-02-11 MScDiplom_Lemenkova.tex
written '1.8. Research Assumptions'p.11

@@
 	\acro{ASTER}{\href{http://asterweb.jpl.nasa.gov/}{Advanced Spaceborne Thermal Emission and Reflection Radiometer}} 
-	%\acro{BRDF}{Bidirectional Reflectance Distribution Function}
+	\acro{BRDF}{Bidirectional Reflectance Distribution Function}
 	\acro{CASI}{\href{http://www.borstad.com/casi.html}{Compact Airborne Spectrographic Imager}}
@@
 	\centering
-	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
+	\includegraphics[scale=0.63]{UML_ClassDiagram.jpg}
 	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}. \\ Page: $<$\pageref{page-9}$>$}
@@
 Supervised, Maximum Likelihood).
+\pagebreak
 
 \subsection{Research assumptions}\index[concepts]{Research!assumptions}
-The wind roughening the water surface causes sun glitter, and the values spectral reflectance may contain disturbances and noisy if the solar zenith angle is too high, say approaching 60-70 degrees.
-Therefore, the measurements should be made at midday (noon) with as low zenith angle (Fig.\ref{fig:A.2-1})\label{zenith} as possible.
+The general research assumption that is used for further applications of the remote sensing methods of this work is that optical properties of the spectral reflectance of the seawater with different bottom substrata enable to discriminate seafloor cover types using remote sensing products from broadband sensors.\vspace{1ex}\\
+The accepted model-specific parameters of the \ac{WASI} simulator are described in Table 4.2.
+The weather conditions for the measurements are assumed to be perfect: clear, sunny, windless days, because wind roughen water surface causing sun glitter, therefore, values of spectral reflectance may contain disturbances.
+\\ Sky radiance might be influenced by multiple reflectance between the sea surface albedo and the atmosphere and in general, sky radiance increases rapidly while viewing zenith angle is near to $90\,^{\circ}$ (i.e.
+in the evening hours).
+Therefore, ideally measurements should be done at noon, with as low zenith angle (Fig.\ref{fig:A.2-1})\label{zenith} as possible; otherwise, if the solar zenith angle is too high (e.g.
+approaching $60-70\,^{\circ}$), received data may contain noise.
+For \ac{WASI} simulations reflection factor of sky radiance is taken as 0.0201 with simulated ideal conditions: viewing angle ($0\,^{\circ}$ = nadir).
+The interval of water temperature taken as default lies in the diapason $17-25\,^{\circ}$ as to simulate the conditions of the Mediterranean Sea.
+However, in real time conditions the sun zenith angle of $45\,^{\circ}$ has been accepted as suitable.\vspace{1ex}\\
+The anisotropy factor of upwelling radiation or the quality (Q-) factor, showing the directionally dependency of the radiance, is taken as 5.
+We accept some values of optical parameters as default values at \ac{WASI}.
+Thus, concentration of phytoplankton is accepted at the interval of $0.035-0.089 mg^{-1}$ and concentration of large suspended particles is given to 8.
+Reference wavelength for \ac{CDOM} (Gelbstoff) absorption is equal to 440.
+The coefficient of attenuation remains equal to 1.0546, as set up by default at \ac{WASI}.\vspace{1ex}\\
+We also assume that concentration of non-chlorophyll particles (absorption at 0) as well as concentration of small suspended particles is equal to zero, so we do not count them in this work.
+Exponent of \ac{CDOM} (Gelbstoff) absorption is accepted as 0.0140.
+Finally, the \ac{BRDF} of bottom reflectance (sand), which defines the reflection of light  is at an opaque surface (Fig.A.2.5), is assumed to be 0.318 $sr^{-1}$.
+
 \pagebreak
@@
 
-\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}
+\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}\label{page-15}
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
@@
 	\caption{\textit{In situ} optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
-		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}. Page: $<$\pageref{page-21}$>$}
+		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}. \\ Pages: $<$\pageref{page-17}$>$, $<$\pageref{page-21}$>$}
 	\label{fig:13}
@@
 result of the spectral absorption in different bands, typical for each target object.
-Spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
+Spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13})\label{page-17} is largely influenced by the water
 depth where it is located, and is generally decreasing in values by increasing depths.
@@
 %%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and methods}\label{sec:3}
+\section{Materials and methods}\label{sec:3}\label{page-22}
 \counterwithin{figure}{section}
@@
 	\subfloat [Olympus waterproof camera]{\label{fig:3.3d}\includegraphics[width=0.2\textwidth]{Fig-19d.jpg}}
-	\caption{Fieldwork equipment. Pages: $<$\pageref{page-15}$>$, $<$\pageref{page-24}$>$, $<$\pageref{page-25}$>$}
+	\caption{Fieldwork equipment. Pages: $<$\pageref{page-15}$>$, $<$\pageref{page-24}$>$, $<$\pageref{page-25}$>$, $<$\pageref{page-26}$>$, $<$\pageref{page-27}$>$}
 	\label{fig:3.3}
@@
 	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
-	\caption{Scheme of placements of the Olimpus cameras during the measurements. Page: $<$\pageref{page-26}$>$}
+	\caption{Scheme of placements of the Olimpus cameras during the measurements. Pages: $<$\pageref{page-26}$>$, $<$\pageref{page-27}$>$}
 	\label{fig:3.5}
@@
 classify seafloor cover types (Fig.\ref{fig:A.56}) and seagrass meadows, according to the differences in the structure,
-colour, texture and shapes of the depicted objects.
+colour, texture and shapes of the depicted objects\label{page-27}.
 
@@
 	\includegraphics[scale=0.50]{Fig-21.jpg}
-	\caption{Locations of the video measurements and GPS tracklogs, Ligaria. Page: $<$\pageref{page-25}$>$}
+	\caption{Locations of the video measurements and GPS tracklogs, Ligaria. Pages: $<$\pageref{page-25}$>$, $<$\pageref{page-26}$>$}
 	\label{fig:3.6}
@@
 information on seagrass presence within the study area, distribution of seagrass \textit{P.oceanica} meadows
-and nature of the seafloor cover types: rocks, sandy, mixed (Fig. \ref{fig:A5}). Seagrass species on the Ligaria and Agia
-Pelagia beach consist of 	extit{P.oceanica}. The types of sediment on the Ligaria vary from coarse sand,
+and nature of the seafloor cover types: rocks, sandy, mixed (Fig.\ref{fig:A.6}, Fig.\ref{fig:A.5}, Fig.\ref{fig:A.3}).
+Seagrass species on the Ligaria and Agia Pelagia beach consist of \textit{P.oceanica}.
+The types of sediment on the Ligaria vary from coarse sand,
 \textit{P.oceanica} patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:A.6}).
@@
 A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of \textit{P.oceanica} made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
-\ref{tab:A.8}), 1-445 for 14th October for 	\textit{P.oceanica}; 1-106 for water without sediments, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.
-\ref{tab:17}).
+\ref{tab:A.8}), 1-445 for 14th October for \textit{P.oceanica}; 1-106 for water without sediments, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.\ref{tab:17}).
 All these data have been analyzed and statistically tested.
@@
 We visualized the behavior of different spectra on relevant graphs (Fig.
-\ref{fig:4.4}, \ref{fig:A.35}, \ref{fig:A.28}).
+\ref{fig:4.4}, \ref{fig:A.35}, \ref{fig:A.28}).\label{page-28}
 The available collected data were tested for the spectral variability and separability under varying conditions of different environmental
@@
 column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:A.11}). The next step included
-calculation of the median, mean, quartiles and other statistical values at every data set (see \nameref{app}, Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
+calculation of the median, mean, quartiles and other statistical values at every data set (see \nameref{app},\label{dataprep} Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
 After the preliminary analysis, the measured data were visualized using Gnuplot program, which
@@
 \end{itemize}
-The available satellite imagery (Fig.\ref{fig:A.38}) and aerial images are read into the ArcGIS project.
+The available satellite imagery (Fig.\ref{fig:A.38}) \label{page-30}and aerial images are read into the ArcGIS project.
 The available broadband and hyperspectral remote sensing data are used for the mapping of the
@@
 The distinguishing spectral signatures for various seafloor types (e.g. seagrass species, coral reefs,
-various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})) exist in well-defined and narrow (10–20 nm)
+various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})\label{page-31} exist in well-defined and narrow (10–20 nm)
 wavelength ranges. 
@@
 \href{http://www.gnuplot.info/}{Gnuplot} graph.
-Page: $<$\pageref{page-47}$>$}
+Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-47}$>$, $<$\pageref{dataprep}$>$}
 	\label{fig:28}
@@
 \href{http://www.gnuplot.info/}				{Gnuplot}.
-Page: $<$\pageref{page-21}$>$}\label{fig:4.2}
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-31}$>$}\label{fig:4.2}
 \end{figure}
@@
 water, large and small particles. The values of all these parameters can be redacted and changed
-manually. However, the most important, major factors affecting the \textit{in-situ} conditions are water depth (Fig.\ref{fig:4.9})
+manually. However, the most important, major factors affecting the \textit{in-situ} conditions are water depth (Fig.\ref{fig:4.9})\label{page-32}
 and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand.
@@
 \href{http://www.gnuplot.info/}{Gnuplot}.
-Two complimentary graphs below show the results of the statistical analysis.}
+Two complimentary graphs below show the results of the statistical analysis.\\ Pages: $<$\pageref{page-28}$>$, $<$\pageref{dataprep}$>$, $<$\pageref{page-31}$>$}
 	\label{fig:4.4}
@@
 seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
-remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}, Fig.\ref{fig:4.6W}).
+remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}, Fig.\ref{fig:4.6W}).\label{page-33}
 
@@
 	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths.
-Page: $<$\pageref{page-21}$>$}
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$}
 	\label{fig:4.6}
@@
 	\includegraphics[scale=0.30]{Fig-30.jpg}
-	\caption{Main interface of \ac{WASI} water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m. Page: $<$\pageref{page-35}$>$}
+	\caption{Main interface of \ac{WASI} water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m. Pages: $<$\pageref{page-35}$>$, $<$\pageref{page-31}$>$}
 	\label{fig:4.7}
@@
 		    C-L & Concentration of large suspended particles & 8 \\ \hline
-		   C(i), i=0..5 & Concentration of Phytoplankton & 0.035 -0.089 ug -l \\ \hline
-		    bbS & Specific backscattering for small particles & 0.005 m2g-1 \\ \hline
-		   T-W & Temperature of water & 17-25 °C \\ \hline
-		    n & Exponent of Backscattering by small particles & 0.005 m2g-1 \\ \hline
+		   C(i), i=0..5 & Concentration of Phytoplankton & 0.035 -0.089 $ug^{-1}$ \\ \hline
+		    bbS & Specific backscattering for small particles & 0.005 $m^2g^{-1}$ \\ \hline
+		   T-W & Temperature of water & $17-25\,^{\circ}C$ \\ \hline
+		    n & Exponent of Backscattering by small particles & 0.005 $m^2g^{-1}$ \\ \hline
 		    Q &Anisotropy factor of upwelling radiation ("Q-factor") & 5.00 \\ \hline
 		    sigma-L & Reflection factor of sky radiance & 0.0201 \\ \hline
-		    b1 & Backscattering coefficient of saline waters & 0.00144 m-1 \\ \hline
-		    λ0 & Reference Wavelength for Gelbstoff absorption & 440 \\ \hline
+		    b1 & Backscattering coefficient of saline waters & 0.00144 $m^{-1}$ \\ \hline
+		    λ0 & Reference Wavelength for \ac{CDOM} (Gelbstoff) absorption & 440 \\ \hline
 		    sun & Sun zenith angle & 45.0 \\ \hline
@@
 		   f(i), i=0...5 & Areal fraction of bottom surface type number n & 01/10/10 \\ \hline
-		    K0 & Coefficient of Attentuation & 1.0546 \\ \hline
+		    K0 & Coefficient of Attenuation & 1.0546 \\ \hline
 		   view & Viewing angle (0 = nadir) & 0\\ \hline
-		    C-X & Concentration of non-chlorophyllous particles (absorption at λ0) & 0 \\ \hline
+		    C-X & Concentration of non-chlorophyll particles (absorption at $\lambda_0$) & 0 \\ \hline
 		   λS & Reference wavelength for scattering of small particles & 500 \\ \hline
 		    C-S & Concentration of small suspended particles & 0 \\ \hline
-		   S & Exponent of Gelbstoff absorption & 0.0140 \\ \hline
-		    C-Y & Concentration of Gelbstoff (absorption at λ0) & 0.400 \\ \hline
-		    Bn & BDRF of bottom reflectance (sand) & 0.318 sr -1\\
+		   S & Exponent of \ac{CDOM} absorption & 0.0140 \\ \hline
+		    C-Y & Concentration of \ac{CDOM} Gelbstoff (absorption at $\lambda_0$) & 0.400 \\ \hline
+		    Bn & BDRF of bottom reflectance (sand) & 0.318 $sr^{-1}$\\
 	    \hline
@@
 	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d.jpg}}
-	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths, simulated for broadband sensors. Page: $<$\pageref{page-37}$>$}
+	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths, simulated for broadband sensors. Pages: $<$\pageref{page-37}$>$, $<$\pageref{page-32}$>$}
 	\label{fig:4.9}
@@
 \href{http://www.gnuplot.info/}{Gnuplot}.
-Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-37}$>$}
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-28}$>$, $<$\pageref{page-37}$>$}
 	\label{fig:4.10}
@@
 	\subfloat [\textit{P.oceanica}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
-	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
+	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia. Page: $<$\pageref{page-22}$>$}
 	\label{fig:4.11}
@@
 ocean trawling) still remain a main cause for landscapes fragmentation.
-\pagebreak
 
@@
 \addcontentsline{toc}{section}{Recommendations}
+\singlespacing
 
@@
 \section{appendices}\label{app}
-\subsection{\footnote{The page where the illustration is explained in the main text of MSc thesis is given in angle brackets (e.g., $<$10$>$).} Capturing aerial imagery from the Google Earth}
+\subsection{\footnote{The page where the illustration is explained in the main text of MSc thesis is given in angle brackets (e.g., Page: $<$\pageref{page-40}$>$).} Capturing aerial imagery from the Google Earth}
 %\appendixname{\textbf{1: \\Capturing aerial imagery from the Google Earth: grabbing process}}
@@
 
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.6]{backscattering.jpg}
+		\caption{Example of wave backscattering from the vegetation (here: rice).
+Source: \href{http://cse.niaes.affrc.go.jp/miwa/esid/highlight/microwave-backscatter.html}{Yoshio Inoue}}.\label{fig:backscat}
+	\end{center}
+\end{figure}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.4]{BRDF.jpg}
+		\caption{BRDF is a ratio of reflected radiance along $\omega_o$ to the irradiance from direction $\omega_i$, all parameterized by azimuth angle $\phi$ and zenith angle $\theta$.
+Source: \href{http://en.wikipedia.org/wiki/Portal:Contents/Quick_index}{Wikipedia.org}}.\label{fig:BRDF}
+	\end{center}
+\end{figure}
+
 \pagebreak
@@
 \begin{table}[H]\footnotesize
-	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer and \ac{SAM} during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}\label{tab:A.1}
+	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer and \ac{SAM} during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4). Page: $<$\pageref{page-30}$>$}\label{tab:A.1}
 	\centering
@@
 		     DateTime & 2009-10-15 08:56:22 & 10/14/2009, 15:15:59 \\ \hline
-		     PositionLatitude & 35.2451=35°14'29.4354” & 35.24752 = 35° 14' 51.072" \\ \hline
-		     PositionLongitude & 25.01269=25°0'45.6834" & 25.0098 = 25°0'35.2794" \\ \hline
+		     PositionLatitude & 35.2451=$35\,^{\circ}$14'29.4354” & 35.24752 = $35\,^{\circ}$14' 51.072" \\ \hline
+		     PositionLongitude & 25.01269=$25\,^{\circ}$0'45.6834" & 25.0098 = $25\,^{\circ}$0'35.2794" \\ \hline
 		     Comment & \textit{P.oceanica} & \textit{P.oceanica} \\ \hline
@@
 \begin{table}[H]\footnotesize
-	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}\label{tab:A.2}	\centering
+	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4). Page: $<$\pageref{page-30}$>$}\label{tab:A.2}	\centering
 	\rowcolors{1}{Lavender}{GhostWhite}
@@
 	      DateTime & 2009-10-14 15:32:02 & 2009-10-15 10:36:02 \\ \hline
-	     PositionLatitude & 35.24752= 35°14'51.072" & 35.245071=35°14'42.2556" \\ \hline
-	     PositionLongitude & 25.0098= 25°0'35.2794" & 25.012661=25°0'45.5796" \\ \hline
+	     PositionLatitude & 35.24752= $35\,^{\circ}14\textquoteright51.072"$ & 35.245071=$35\,^{\circ}14\textquoteright42.2556"$ \\ \hline
+	     PositionLongitude & 25.0098= $25\,^{\circ}0'35.2794"$ & 25.012661=$25\,^{\circ}0'45.5796"$ \\ \hline
 	     Comment & \textit{P.oceanica} & \textit{P.oceanica} \\ \hline
@@
 	\subfloat [Isolated patch of seagrass]{\includegraphics[width=0.3\textwidth]{isolated_patch.jpg}}
-	\caption{Types of seagrass structural patterns, Ligaria beach, Crete}
+	\caption{Types of seagrass structural patterns, Ligaria beach, Crete.  Pages: $<$\pageref{page-27}$>$, $<$\pageref{page-28}$>$}
 	\label{fig:A.3}
@@
 		\includegraphics[scale=0.67]{Fig-3-14.jpg}
-		\caption{Ligaria beach, Crete: seafloor types. Page: $<$\pageref{page-39}$>$}\label{fig:A.5}
+		\caption{Ligaria beach, Crete: seafloor types. Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-39}$>$}\label{fig:A.5}
 	\end{center}
@@
 		\includegraphics[scale=0.3]{Fig-3-15.jpg}
-		\caption{Various seafloor types; Ligaria beach, Crete}\label{fig:A.6}
+		\caption{Various seafloor types; Ligaria beach, Crete. Page: $<$\pageref{page-28}$>$}\label{fig:A.6}
 	\end{center}
@@
 	\includegraphics[scale=0.55]{Fig-24.jpg}
-	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
+	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements. \\ Page: $<$\pageref{page-28}$>$}
 	\label{fig:A.8}
@@
 	\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
-		seagrasses (various species), sand, silt and artificial spectrum of constant albedo. Page: }
+		seagrasses (various species), sand, silt and artificial spectrum of constant albedo. Page: $<$\pageref{page-28}$>$}
 	\label{fig:A.9}
@@
 		\includegraphics[scale=0.35]{Interpolation.jpg}
-		\caption{Interpolation of the spectral measurements by means of Open Office}
+		\caption{Interpolation of the spectral measurements by means of Open Office. Page: $<$\pageref{page-28}$>$}
 		\label{fig:A.10}
@@
 Two complimentary graphs below show the results of the statistical analysis.
-Page: $<$\pageref{page-35}$>$}
+Pages: $<$\pageref{page-28}$>$, $<$\pageref{dataprep}$>$, $<$\pageref{page-35}$>$}
 	\label{fig:27}
@@
 		\includegraphics[scale=0.20]{App-1.jpg}
-		\caption{Values of spectral reflectances of seagrass \textit{P.oceanica} (fragment). Page: $<$\pageref{page-40}$>$}
+		\caption{Values of spectral reflectances of seagrass \textit{P.oceanica} (fragment). Pages: $<$\pageref{dataprep}$>$, $<$\pageref{page-40}$>$}
 		\label{fig:A.11}
@@
 		\includegraphics[scale=0.25]{GNU-401-420.jpg}
-		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-21}$>$}
+		\caption{Spectral reflectance of \textit{P.oceanica}.
+Measurement series 401-420.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages:  $<$\pageref{page-21}$>$, $<$\pageref{dataprep}$>$, $<$\pageref{page-31}$>$}
 		\label{fig:A.12}
@@
 \href{http://www.gnuplot.info/}{Gnuplot}.
-Page: $<$\pageref{page-21}$>$}
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-31}$>$}
 		\label{fig:A.13}
@@
 		\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
-		Visualization of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
+		Visualization of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-31}$>$}
 		\label{fig:A.14}
@@
 	\caption{Fragment of the statistical analysis of the \textit{P.oceanica} reflectance. Example of data set 401-420. 
-		Visualization of the measured data together with statistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-21}$>$}
+		Visualization of the measured data together with statistical values: interquartile ranging, medians, means.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{dataprep}$>$, $<$\pageref{page-30}$>$}
 	\label{fig:A.15}
@@
 	\includegraphics[scale=0.35]{GNU-Radiance_Bezier.jpg}
-	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-40}$>$}
+	\caption{Radiance of the seawater with sediments.
+Measured in aquarium tank.
+Bezier Interpolation.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{dataprep}$>$, $<$\pageref{page-40}$>$, $<$\pageref{page-30}$>$}
 	\label{fig:A.16}
@@
 	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
-\ac{SPSS}}\label{tab:8}
+\ac{SPSS}.
+Page: $<$\pageref{page-28}$>$}\label{tab:8}
 	\begin{center}
@@
 	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters.
-\href{http://www.spss.com/}{SPSS}}\label{tab:A.7}
+\href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{page-28}$>$}\label{tab:A.7}
 	\begin{center}
@@
 Measured on Agia Pelagia beach, 15th October.
-Page: $<$\pageref{page-40}$>$}\label{tab:A.8} \\
+Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-40}$>$}\label{tab:A.8} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 		\includegraphics[scale=0.30]{GNU-Irr-Splines.jpg}
-		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}­}
+		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-31}$>$­}
 		\label{fig:A.17}
@@
 	\includegraphics[scale=0.30]{GNU-Irr.jpg}
-	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}}
+	\caption{Irradiance of the seawater, measured in aquarium tank.
+Smooth Bezier interpolation.
+Visualization in \href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{dataprep}$>$, $<$\pageref{page-30}$>$}
 	\label{fig:A.18}
@@
 		\includegraphics[scale=0.25]{GNU-09.jpg}
-		\caption{Spectral reflectance of the seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle. \href{http://www.gnuplot.info/}{Gnuplot}­}
+		\caption{Spectral reflectance of the seawater with sediments.
+Bezier interpolation.
+Mean value is shown by the vertical impulses linestyle.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{dataprep}$>$­, $<$\pageref{page-31}$>$}
 		\label{fig:A.23}
@@
 		\includegraphics[scale=0.22]{GNU-10.jpg}
-		\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. \href{http://www.gnuplot.info/}{Gnuplot} display­}
+		\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. \href{http://www.gnuplot.info/}{Gnuplot} display. \\
+Page: $<$\pageref{page-31}$>$­}
 		\label{fig:A.24}
@@
 	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
-	\caption{WASI-simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
+	\caption{WASI-simulated remote sensing reflectance of \textit{P.oceanica} at various sensors. Page: $<$\pageref{page-33}$>$}
 	\label{fig:4.6W}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Summary of the table with values of the spectral reflectance of \textit{P.oceanica} measured on three different depths, R}
+	\caption{Summary of the table with values of the spectral reflectance of \textit{P.oceanica} measured on three different depths, R. Page: $<$\pageref{page-28}$>$}
 	\begin{center}
@@
 		\includegraphics[scale=0.35]{GNU-11.jpg}
-		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display. Page: $<$\pageref{page-21}$>$­}
+		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display. Pages: $<$\pageref{page-21}$>$, $<$\pageref{dataprep}$>$­}
 		\label{fig:A.25}
@@
 \href{http://www.spss.com/}{SPSS}.
-Page: $<$\pageref{page-41}$>$}
+Pages: $<$\pageref{page-30}$>$, $<$\pageref{page-41}$>$}
 	\label{tab:13}
@@
 
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.38]{GNU-15.jpg}
-		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
-Series 151-200.
-Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-\href{http://www.gnuplot.info/}{Gnuplot} visualization­.
-Page: $<$\pageref{page-39}$>$}
-		\label{fig:A.28}
-	\end{center}
-\end{figure}
-\pagebreak
-
 %%% one more Gretl start
@@
 %%%one more Gretl end
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.38]{GNU-15.jpg}
+		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
+Series 151-200.
+Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
+\href{http://www.gnuplot.info/}{Gnuplot} visualization­.
+Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-39}$>$}
+		\label{fig:A.28}
+	\end{center}
+\end{figure}
+\pagebreak
 
@@
 \href{http://www.gnuplot.info/}{Gnuplot}.
-Page: $<$\pageref{page-39}$>$}
+Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-39}$>$}
 	\label{fig:A.35}
@@
 		\includegraphics[scale=0.30]{App-8.jpg}
-		\caption{Random mosaic of selected aerial Google Earth images. Page: $<$\pageref{page-47}$>$}\label{fig:A.37}	
+		\caption{Random mosaic of selected aerial Google Earth images. Pages: $<$\pageref{page-30}$>$, $<$\pageref{page-47}$>$}\label{fig:A.37}	
 	\end{center}	
@@
 \begin{table}[H]
-	\caption{Available broadband Landsat satellite images covering the research area of Crete island}\label{tab:34}
+	\caption{Available broadband Landsat satellite images covering the research area of Crete island.}\label{tab:34}
 	\centering
@@
 	\hspace{1mm}
-	\subfloat [Mosaic-Landsat-ETM-1999-08-08]{\includegraphics[width=0.3\textwidth]{Landsat10.jpg}}
-	\caption{Landsat imagery, Crete island. Previews}
+	\subfloat [Landsat-ETM-99-08-08]{\includegraphics[width=0.3\textwidth]{Landsat10.jpg}}
+	\caption{Landsat imagery, Crete island. Previews. Page: $<$\pageref{page-30}$>$}
 	\label{fig:A.38}
@@
 	\includegraphics[scale=0.7]{SCUBA-gear.jpg}
-	\caption{SCUBA gear diving equipment necessary for seagrass monitoring. Source: \href{http://divinginstructortraining.com/}{Aquanauts.com}. Page: $<$\pageref{page-25}$>$}
+	\caption{SCUBA gear diving equipment necessary for seagrass monitoring. Source: \href{http://divinginstructortraining.com/}{Aquanauts.com}. \\ Pages: $<$\pageref{page-15}$>$, $<$\pageref{page-25}$>$}
 	\label{fig:A.52}
@@
 	\includegraphics[scale=0.4]{photo-4.jpg}
-	\caption{Monitoring different seafloor cover types: mattte of \textit{P.oceanica} vs carbonate sand. Ligaria beach.}
+	\caption{Monitoring different seafloor cover types: mattte of \textit{P.oceanica} vs carbonate sand. Ligaria beach. \\ Page: $<$\pageref{page-27}$>$}
 	\label{fig:A.56}

### 2011-02-11 MScDiplom_Lemenkova.tex
Valentijn's new comments (11 Feb). From total 36 finished 14, 15, 20, 23, 29, 34. Partly done: 18, 19.

@@
 	\end{itemize}
+Although there are a variety of environmental factors that contribute to the spectral
+reflectance, the most important ones are water column height and seafloor fraction.
+It is because spectra of the seagrass \textit{P.oceanica} vary qualitatively over the depths interval of 0.5–4 m,
+and secondly, the content and cover fraction of the seafloor have the most distinctive effect on the
+spectral reflectance of the water. \\
+Therefore, in the current research we focus on these two major factors, and study the response of the
+water reflectance towards changing conditions of the water column depth and seafloor bottom cover
+fraction (seagrass \textit{P.oceanica}, carbonate sand and silt).\\
 The research aims to explore the limitations of the application of the \ac{RTM} and remote sensing techniques 
@@
 \textit{P.oceanica} patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:A.6}).
+
+\subsection{Review of the collected data}
+The remote sensing imagery was collected for the Crete island area covering the research period and
+area, enabling the field data to be used for the calibration and validation. The available aerial and
+satellite imagery are commonly used for the mapping of seagrass landscapes and their application is
+proven by various research papers. The imagery includes satellite multi-spectral imagery (Landsat-TM,
+ETM+) and aerial imagery from the Google Earth.
+The overview of the collected data enables to summarize their following types:
+\begin{itemize}
+	\item[] 1.
+Optical spectra of \textit{P.oceanica}, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig.
+\ref{fig:A.23})
+	\item[]2. Aerial imagery from the \href{http://www.google.com/earth/index.html}{Google Earth}
+	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:A.37})
+	\item[]4. Results of underwater videometric measurements of the \ac{Olympus ST} cameras made during the ship route
+\end{itemize}
+The available satellite imagery (Fig.\ref{fig:A.38}) \label{page-30}and aerial images are read into the ArcGIS project.
+The available broadband and hyperspectral remote sensing data are used for the mapping of the
+seagrass in shelf areas no deeper than 4.0 meters, for the environmental monitoring in order to detect
+the dynamics in changes of the seagrass distribution along Crete during the past 10 years using
+different satellite images for 2000-2010.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.40]{UML_Dia.jpg}
+	\caption{Work flow for the data acquisition, \href{http://live.gnome.org/Dia}{Dia} display}
+	\label{fig:24}
+\end{figure}
+
+The spatial resolution of \ac{Landsat ETM+} image is 30 m in the visible and near infrared bands (bands
+1-5 and 7); the spatial resolution of \ac{ASTER} 15 m for the visible and near-infrared bands. IKONOS
+acquires data in 3 visible channels and \ac{NIR}, with spatial resolution of 1-4 meters. \ac{ASTER} and
+IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
+research includes fieldwork \textit{in-situ} measurements of the seagrass distribution along the northern and
+south-eastern coasts of Crete in chosen locations.\\
+The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
+the seafloor have been received by the means of the optical sensors \ac{Trios-RAMSES} Hyperspectral
+\ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor (Fig.
+\ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see \nameref{app}: Tab.
+\ref{tab:A.1}, Tab.
+\ref{tab:A.2}).
+\\
+The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:A.18}) and 430-650 for the water radiance (Fig.\ref{fig:A.16}). 
+Afterwards, the measured values of the radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
+properties of the sea water and bottom cover types (Fig.\ref{fig:A.15}). The spectral range of radiance
+cover diapason of 320-950 nm, and irradiance measurements are covered in the interval of 280-500
+nm, which is suitable for characteristics of seagrass reflectance.
+Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:A.12}, or \ref{fig:A.14}, enlarged) represent several series of the measurements.
+The values of the spectral reflectance are received from the
+computations of these values using mathematical formulae.
+The graphs shown on Fig.
+\ref{fig:4.2}, Fig.
+\ref{fig:A.23}, Fig.
+\ref{fig:A.24}, Fig.
+\ref{fig:A.17} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively.
+Graph on Fig \ref{fig:A.13} displays statistical analysis of the measured sets of observations, i.e.
+shows the mid-spread of the statistical quartiles (Q1-Q3), mean and extreme values.
 
@@
 Therefore, after several experiments with various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have chosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contains convex hulls made on its control points, and therefore is best suitable for our case: analysis of optical
-properties of seawater.
+properties of seawater.\\
+We used several statistical methods to asses data received as a result of measurements. 
+The Student-t test\index[concepts]{Data!statistical analysis!Student-t testing},
+one of the most commonly used techniques for testing a hypothesis on the basis
+of a difference between sample means, is used for the data analysis. In our case the Student t-test demonstrates, if the variation
+between two analyzed groups – values of spectral reflectance of seagrass \textit{P.oceanica} and sand - is
+significant. Therefore, we use Student t-test to compare two sets of quantitative data of spectral
+reflectance of \textit{P.oceanica} and sand, respectively, with samples collected independently of one
+another. The Student-t test can be performed knowing just the means, standard deviation, and number
+of data points. Therefore, we used the data (\nameref{app}) of means of spectral reflectance of the both
+ cover types within data sets, their standard deviation, and the number of data points.
 
@@
 
+\subsection{Image processing using Erdas Imagine}\label{page-47}
+The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
+investigate the distribution of the seagrass \textit{P.oceanica} within the research area.
+In the image processing part of the research supervised classification has been applied to the aerial
+and satellite images. Seagrass meadows and other seafloor cover types were evaluated through a
+detailed examination of the imagery. \\
+Seagrass beds are clearly visible in color aerial Google Earth\index[concepts]{Remote sensing!aerial imagery!GoogleEarth}
+photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
+detected using different bands combinations, masked and studied for the estimation of the changes in
+the areas. The classification is based on the properties of the \textit{P.oceanica}, such as brightness,
+colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping\label{page-48}
+includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
+for each photograph by classification a series of polygons characteristic of each of the sea floor
+bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
+seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).
+
+\begin{figure}
+	\centering
+	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
+	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
+	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images. Page: $<$\pageref{page-48}$>$}
+	\label{fig:38}
+\end{figure}
+
+On the basis of the image processing and classification, applied to the northern coats of Crete, the
+marine bottom types between 10 m depth has been mapped, including the limit depths of the of the
+seagrass meadows provided on the basis of the available and collected field data. The classification
+has been completed using Erdas Imagine software.
+
 %%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
@@
 \counterwithin{table}{section}
-
-\subsection{Review of the collected data}
-The remote sensing imagery was collected for the Crete island area covering the research period and
-area, enabling the field data to be used for the calibration and validation. The available aerial and
-satellite imagery are commonly used for the mapping of seagrass landscapes and their application is
-proven by various research papers. The imagery includes satellite multi-spectral imagery (Landsat-TM,
-ETM+) and aerial imagery from the Google Earth.
-The overview of the collected data enables to summarize their following types:
-\begin{itemize}
-	\item[] 1.
-Optical spectra of \textit{P.oceanica}, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig.
-\ref{fig:A.23})
-	\item[]2. Aerial imagery from the \href{http://www.google.com/earth/index.html}{Google Earth}
-	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:A.37})
-	\item[]4. Results of underwater videometric measurements of the \ac{Olympus ST} cameras made during the ship route
-\end{itemize}
-The available satellite imagery (Fig.\ref{fig:A.38}) \label{page-30}and aerial images are read into the ArcGIS project.
-The available broadband and hyperspectral remote sensing data are used for the mapping of the
-seagrass in shelf areas no deeper than 4.0 meters, for the environmental monitoring in order to detect
-the dynamics in changes of the seagrass distribution along Crete during the past 10 years using
-different satellite images for 2000-2010.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.40]{UML_Dia.jpg}
-	\caption{Work flow for the data acquisition, \href{http://live.gnome.org/Dia}{Dia} display}
-	\label{fig:24}
-\end{figure}
-
-The spatial resolution of \ac{Landsat ETM+} image is 30 m in the visible and near infrared bands (bands
-1-5 and 7); the spatial resolution of \ac{ASTER} 15 m for the visible and near-infrared bands. IKONOS
-acquires data in 3 visible channels and \ac{NIR}, with spatial resolution of 1-4 meters. \ac{ASTER} and
-IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
-research includes fieldwork \textit{in-situ} measurements of the seagrass distribution along the northern and
-south-eastern coasts of Crete in chosen locations.\\
-The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
-the seafloor have been received by the means of the optical sensors \ac{Trios-RAMSES} Hyperspectral
-\ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor (Fig.
-\ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see \nameref{app}: Tab.
-\ref{tab:A.1}, Tab.
-\ref{tab:A.2}).
-\\
-The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:A.18}) and 430-650 for the water radiance (Fig.\ref{fig:A.16}). 
-Afterwards, the measured values of the radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
-properties of the sea water and bottom cover types (Fig.\ref{fig:A.15}). The spectral range of radiance
-cover diapason of 320-950 nm, and irradiance measurements are covered in the interval of 280-500
-nm, which is suitable for characteristics of seagrass reflectance.
-Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:A.12}, or \ref{fig:A.14}, enlarged) represent several series of the measurements.
-The values of the spectral reflectance are received from the
-computations of these values using mathematical formulae.
-The graphs shown on Fig.
-\ref{fig:4.2}, Fig.
-\ref{fig:A.23}, Fig.
-\ref{fig:A.24}, Fig.
-\ref{fig:A.17} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively.
-Graph on Fig \ref{fig:A.13} displays statistical analysis of the measured sets of observations, i.e.
-shows the mid-spread of the statistical quartiles (Q1-Q3), mean and extreme values.
 
@@
 Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}				{Gnuplot}.
+\href{http://www.gnuplot.info/}{Gnuplot}.
 Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-31}$>$}\label{fig:4.2}
@@
 relationship between their concentrations and the radiance of water that finally influence its spectral
-reflectance.\\
-However, although there are a variety of environmental factors that contribute to the spectral
-reflectance, the most important ones from the mentioned above are the water column (depth) and seafloor fraction.
-It is because spectra of the seagrass \textit{P.oceanica} vary qualitatively over the depths interval of 0.5–4 m,
-and secondly, the content and cover fraction of the seafloor have the most distinctive effect on the
-spectral reflectance of the water. \\
-Therefore, in the current research we focus on these two major factors, and study the response of the
-water reflectance towards changing conditions of the water column depth and seafloor bottom cover
-fraction (seagrass \textit{P.oceanica}, carbonate sand and silt).\\
-The analysis of the spectral signatures of the seagrass \textit{P.oceanica} and sand clearly shows (Fig.\ref{fig:27}) that seagrass
-has spectral reflectance much lesser than that of a carbonate sand, in general not increasing values of
-10 \% reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33\% in its
-highest values.
+reflectance.
 
@@
 environmental conditions which influence its colour.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.30]{Fig-30.jpg}
-	\caption{Main interface of \ac{WASI} water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m. Pages: $<$\pageref{page-35}$>$, $<$\pageref{page-31}$>$}
-	\label{fig:4.7}
-\end{figure}
-
 The data provided with the model was determined at freshwater of the Boddensee \cite{Gege05}\label{Gege05}, 
@@
 forward calculations, i.e. a computing and plotting of series of spectra according to specified
-parameter settings, with exactly defined depths and cover fraction. \\
+parameter settings, with exactly defined depths and cover fraction.
+
+\begin{figure}[H]
+	\centering
+	\subfloat [Concentration of Gelbstoff at depths 0.5 – 8.0 m]{\label{fig:31a}\includegraphics[width=0.4\textwidth]{Fig-31a.jpg}}
+	\subfloat [Bottom reflectance at depth 0.5-4.0 meters]{\label{fig:31b}\includegraphics[width=0.4\textwidth]{Fig-31b.jpg}}
+	\caption{Sea water physical properties, modeled by \ac{WASI}. Page: $<$\pageref{page-36}$>$}
+	\label{fig:4.8}
+\end{figure}
+
 The specific parameters have been chosen for the simulation of the environmental conditions where seagrass
@@
 
-\begin{figure}[h]
-	\centering
-	\subfloat [Concentration of Gelbstoff at depths 0.5 – 8.0 m]{\label{fig:31a}\includegraphics[width=0.4\textwidth]{Fig-31a.jpg}}
-	\subfloat [Bottom reflectance at depth 0.5-4.0 meters]{\label{fig:31b}\includegraphics[width=0.4\textwidth]{Fig-31b.jpg}}
-	\caption{Sea water physical properties, modeled by \ac{WASI}. Page: $<$\pageref{page-36}$>$}
-	\label{fig:4.8}
-\end{figure}
-
 \paragraph{Model parameters: depth and bottom cover fraction}
-Although seagrass \textit{P.oceanica} can be found until depth limits down to 40 m depth (Hartog
-1970), the most preferable limits of its distribution in the Mediterranean Sea are shallow waters until
+Although seagrass \textit{P.oceanica} can be found until depth limits down to 40 m depth \cite{DenHartog70}\label{DenHartog70}, the most preferable limits of its distribution in the Mediterranean Sea are shallow waters until
 4 meters of depth.
 
-\begin{table}
+\begin{table}[H]\footnotesize
 	\rowcolors{1}{Aquamarine}{Azure}
@@
 could still be discriminated are lesser than 2.5 meters.
-The remote sensing properties of the broad-band satellite sensors are often limited and do not allow
-easy and unique discrimination of seagrass from other seafloor types. Therefore, we focus on narrow-band
-spectra with the resolution of 1 nm.
 
@@
 at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:A.35}) and that of \textit{P.oceanica} (Fig.\ref{fig:A.28}).
-The graph is received in excel spreadsheet using mean values of spectral reflectance of
-sand (Fig.\ref{fig:A.27}) and seagrass (Fig.\ref{fig:A.35}), respectively, which have been calculated from measurements of radiance and
-irradiance received in Agia Pelagia bay.
 
@@
 	\centering
-	\subfloat [Sand]{\label{fig:34a}\includegraphics[width=0.4\textwidth]{Fig-27a.jpg}}
-	\subfloat [\textit{P.oceanica}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
+	\subfloat [Sand]{\label{fig:34a}\includegraphics[width=0.3\textwidth]{Fig-27a.jpg}}
+	\subfloat [\textit{P.oceanica}]{\label{fig:34b}\includegraphics[width=0.3\textwidth]{Fig-27b.jpg}}
 	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia. Page: $<$\pageref{page-22}$>$}
@@
 \end{figure}
+The graph is received in excel spreadsheet using mean values of spectral reflectance of
+sand (Fig.\ref{fig:A.27}) and seagrass (Fig.\ref{fig:A.35}), respectively, which have been calculated from measurements of radiance and
+irradiance received in Agia Pelagia bay.
 
@@
 varying \textit{in-situ} conditions, i.e. increasing depth, $H_o$: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
-
-\paragraph{Student-t testing}\index[concepts]{Data!statistical analysis!Student-t testing}
-The Student-t test, one of the most commonly used techniques for testing a hypothesis on the basis
-of a difference between sample means, is used for the data analysis. In our case the Student t-test demonstrates, if the variation
-between two analyzed groups – values of spectral reflectance of seagrass \textit{P.oceanica} and sand - is
-significant. Therefore, we use Student t-test to compare two sets of quantitative data of spectral
-reflectance of \textit{P.oceanica} and sand, respectively, with samples collected independently of one
-another. The Student-t test can be performed knowing just the means, standard deviation, and number
-of data points. Therefore, we used the data (\nameref{app}) of means of spectral reflectance of the both
- cover types within data sets, their standard deviation, and the number of data points.
-The results show the following outcome.
 As a result of the statistical testing, we came to the following conclusion.
@@
 
-\subsection{Image processing using Erdas Imagine}\label{page-47}
-The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
-investigate the distribution of the seagrass \textit{P.oceanica} within the research area.
-In the image processing part of the research supervised classification has been applied to the aerial
-and satellite images. Seagrass meadows and other seafloor cover types were evaluated through a
-detailed examination of the imagery. \\
-Seagrass beds are clearly visible in color aerial Google Earth\index[concepts]{Remote sensing!aerial imagery!GoogleEarth}
-photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
-detected using different bands combinations, masked and studied for the estimation of the changes in
-the areas. The classification is based on the spectral properties of the \textit{P.oceanica}, such as brightness,
-colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping\label{page-48}
-includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
-for each photograph by classification a series of polygons characteristic of each of the sea floor
-bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
-seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).
-
-\begin{figure}
-	\centering
-	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
-	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
-	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images. Page: $<$\pageref{page-48}$>$}
-	\label{fig:38}
-\end{figure}
-
-On the basis of the image processing and classification, applied to the northern coats of Crete, the
-marine bottom types between 10 m depth has been mapped, including the limit depths of the of the
-seagrass meadows provided on the basis of the available and collected field data. The classification
-has been completed using Erdas Imagine software.
-
-\subsection{Seagrass mapping using ArcGIS}
-
-\subsubsection{Data integration}\index[concepts]{Data!integration}
+\subsection{Integrated mapping of seagrass}
+
+\index[concepts]{Data!integration}
 The integrated approach used in this research work has high potential as a means to monitor changes
@@
 
-\subsubsection{Upscale GIS mapping of the seagrass landscapes}\index[concepts]{Seagrasses!spatial structure}\label{page-48}
+%%%%%%%%%%%%%%%%%%%%%%%%% Discussion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}\label{sec:5}
+
+\subsection{Remote sensing for seagrass mapping}
+An approach of the seagrass spectral analysis, monitoring and mapping has been taken in this work, which integrates various research techniques and tools, combining remote sensing methods of spectral analysis of the seafloor cover types, and knowledge of the ecology of \textit{P.oceanica}, with the aim to develop a method of seagrass spectral optical discrimination for the seagrass mapping based on the aerial imagery classification.\vspace{1ex}\\
+In \textit{Chapter 1:} \nameref{sec:1} we discusses main objective of this MSc thesis, which was to study possibilities of seagrass mapping, based on the application of the remote sensing measurement of the seawater optical properties using hyperspectral radiometers.
+The relationship between the optical properties (spectral reflectance) of the seafloor cover types and hydrological parameters of the environment has been studied in order to analyse limitations and capabilities of broadband and narrowband sensors under the conditions of altering environmental parameters.
+For the retrieval of hydrological parameters - seawater radiance, irradiance and spectral reflectance of various seafloor cover types, - spectral optical field measurements were carried out in 2009 at the testing sites in Agia Pelagia, Heraklion and Ligaria beach, Crete Island using \ac{Trios-RAMSES} spectroradiometers.\\
+Further in \textit{Chapter 1:} 1.8 Research assumptions, we assume the constant values of the optical properties of the seawater, phytoplankton, total amount of suspended particles and solids, atmospheric conditions, as well as \ac{CDOM}, which have been set up in modelling part of this work, during \ac{WASI} simulations of various remote sensors.\vspace{1ex}\\
+The second \textit{Chapter 2:} \nameref{sec:2} starts from the review of the available research resources and then discusses various \nameref{sec:2-3} and concludes with reported experience of the remote sensing application towards seagrass mapping: \nameref{sec:2.4.1}.
+In section 2.2 of \textit{Chapter 2:} \nameref{sec:2}, \nameref{sec:2-2}, Measuring water optical properties: hyperspectral radiometers, the \ac{Trios-RAMSES}-ACC-\ac{UV} and \ac{Trios-RAMSES}-ARC spectroradiometers of \ac{Trios-RAMSES} Hyperspectral Sensor series are described.
+\\
+The instruments \ac{Trios-RAMSES} have been used during the fieldwork measurements-2009 for the collection of the reflectance spectra.
+The \ac{Trios-RAMSES}-ACC-\ac{UV} measures spectra in the wavelength domain between 280 nm and 500nm, the \ac{Trios-RAMSES}-ARC, suitable for UV and visible spectra, covers diapason of 320-950 nm with spectral accuracy of 0.3nm (better than 6\%), typical saturation (at 200nm) of  $1Wm^{-2}*nm^{-1}*sr^{-1}$ in 256 channels with a sampling interval of 3.3 nm/pixel and a \ac{FOV} is 7 degrees.
+The spectral reflectance of the seawater with and without sediments, respectively, was calculated by the ratio of the radiance and irradiance values.
+\\
+The reflectance spectra of \textit{P.oceanica} show (Fig.\ref{fig:4.10}, Fig.\ref{fig:A.12}, Fig.
+\ref{fig:A.12}, Fig.\ref{fig:A.13}, Fig.\ref{fig:A.25}, Fig.
+\ref{fig:A.26}) a values maximum between 450 nm and 600 nm, first, because of the chlorophyll absorption peak at 465 and 665nm (Fig.
+\ref{fig:A.2-2})\label{chlorophyll}, secondly, because of the weakening of \ac{CDOM} (or Gelbstoff) in the blue part of the \ac{VIS} spectrum, as it most strongly absorbs short wavelength light in blue to ultraviolet range, and finally, because the absorption of the seawater increases in the red part of the \ac{VIS} spectra.
+The decrease in spectral reflectance values of \textit{P.oceanica} after 660 nm (Fig.\ref{fig:A.13}) is caused by the second absorption peak of phytoplankton.
+The magnitude of the reflectance maximum slightly varies at single variables between about 8 \% and 12 \% (as on Fig.\ref{fig:A.12}) and is probably related to the individual pigmentation and colour composition of single leaves, their structure and geometric orientation, which naturally causes variations in radiance values.
+\\
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.30]{Fig-30.jpg}
+	\caption{\ac{WASI} water color spectral simulation modeller: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m. Pages: $<$\pageref{page-35}$>$, $<$\pageref{page-31}$>$}
+	\label{fig:4.7}
+\end{figure}
+
+The ecological variables, specific to the field environmental conditions, were factored into the \ac{WASI}-based simulation models.
+Through the \ac{WASI} simulation process, imitating spectral properties of \textit{P.oceanica} and carbonate sand for various broadband and narrowband sensors, models were created that accounted for not only atmospheric conditions (i.e.
+sun zenith angle), but also height of water column, thus approaching it to the Mediterranean conditions, and chemical content of the seawater (i.e.
+amount of suspended particles, Gelbstoff, etc), which results in models of optical properties of ``seawater with sediments" and ``seawater without sediments".\\
+The \textit{in-situ} field large-scale matte-level level of seafloor monitoring was then upscaled to airborne Google Earth aerial imagery interpretation, to provide a meadow-level view of seagrass landscapes.
+An attempt of the small-scale mapping is designed on an example of Landsat satellite imagery.
+However, in upscaling to this third, small-scale mapping level further environmental variables need to be considered: health conditions of the seagrass, presence of other underwater vegetation (e.g.
+other seagrass species), hydrological specifications (e.g.
+direction and speed of currents, amplitude of tidal waves, etc), season, date and times of the image taken.
+Therefore, we focused on the first two levels in the current work.
+These different scale levels have been individually considered in terms of the seagrass spectral discernibility for monitoring and mapping, and have been brought together, to provide a roundup of the achieved results and an overview of what still has to be done in \textit{P.oceanica} seagrass mapping by future researchers (see \nameref{sec:7}).
+\vspace{1ex}\\
+In \textit{Chapter 3}: \nameref{sec:3}, describing data collection, a videographic approach tested in previous works, has been applied during the summer fieldwork, when we captured imagery and video footage of the seafloor on several routes of the boat in the Ligaria beach.\\
+The finding of \textit{Chapter 4}, \nameref{sec:4} showed that the relationship between the spectral reflectance of various seafloor cover types was tied to depth, i.e.
+water column height.
+Thus, the results of the in-situ fieldwork measurements revealed that spectral reflectance of \textit{P.oceanica} undergo alterations at depths of 0.5, 2.0 and 3.5m (Fig.\ref{fig:4.9}).
+The analysis of the spectral signatures of the seagrass \textit{P.oceanica} and sand clearly shows (Fig.\ref{fig:27}) that seagrass has spectral reflectance much lesser than that of a carbonate sand, in general not increasing values of
+10 \% reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33\% in its highest values.
+These results indicate that seagrass \textit{P.oceanica} can in general be detected and discriminated from other seafloor cover types with varying environmental conditions, i.e.
+water column height, by hyperspectral spectroradiometers (\ac{Trios-RAMSES}), which positively answers the \textbf{first research question} of this thesis (``Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ conditions ?").
+\vspace{1ex}\\
+Further in \textit{Chapter 4}, \nameref{sec:4}, studies of the broadband versus narrowband sensors demonstrate that simulated spectra of the seagrass, made using \ac{WASI} modeller, have the best results at \ac{CZCS} scanner, especially devoted to the measurement of ocean color.
+The spectrum of \textit{P.oceanica} reflectance, simulated for \ac{CZCS}, covers the wavelength interval of 400-800 nm, and is distinctive for various depths.
+Other remote sensors (\ac{MODIS}, \ac{SeaWiFS}) may also be used for the seagrass mapping, because their technical characteristics enable to spectrally discriminate \textit{P.oceanica} seagrass from other seafloor cover types (Fig.\ref{fig:4.6}), particularly carbonate sand as tested in the current work.
+Therefore, the \textbf{second research question} of this MSc thesis (``Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\
+The graphs showing optical properties of seawater with and without sediments (Fig.\ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:4.2}) focused on spectral variability of the water with changed physical and chemical content.
+The alterations in the individual spectral signatures of single measurements (e.g.
+on Fig.\ref{fig:4.13}, Fig.\ref{fig:A.25}, Fig.\ref{fig:A.27}, Fig.\ref{fig:A.28}) reflect individual health properties of leaves: different nitrogen and chlorophyll content causing diverse colour pigmentation and light absorption, water content in leaves, plant physiological conditions, which vary across seagrass meadow, shoot morphology, etc.
+The differences in spectral reflectance values of the measurements taken on various days might have been caused by the impact of atmospheric conditions, such as solar radiation and sun illumination by different zenith angle.\vspace{1ex}\\
+For further development of the remote sensing based monitoring and mapping of the seagrass and other seafloor cover types it is desirable to consider upscale mapping with concern to bathymetry.
+Studies of the substratum and underwater relief in a more detailed way, i.e.
+bathymetric properties of the testing sizes, gives information about seagrass landscape distribution, because changes in relief directly cause and reinforce the colonisation process of \textit{P.oceanica} meadows.
+Thus, the depression and valleys in relief in general stimulate increase of the sedimentation process.
+In turn, it enables accumulation of necessary nutrient for seagrass plant growth, increases resource allocation to the seagrass roots for better exploitation of pore-water nutrients by the \textit{P.oceanica} shoots.
+As for patchiness in seagrass landscapes, besides natural reasons, the anthropogenic disturbances (e.g.
+ocean trawling) still remain a main cause for landscapes fragmentation.
+
+\subsection{Upscale mapping of the seagrass landscapes}\index[concepts]{Seagrasses!spatial structure}\label{page-48}
 Although ocean pelagic landscapes have a high degree of spatial variance and less structurally
@@
 Besides spatial structure of the seagrass meadows, there are strong and complex patterns of depth zoning, specific to individual
-seagrass species, but such detailed classification goes beyond the scope of the current MSc research.
+seagrass species, but such detailed classification goes beyond the scope of the current study.
 Unique feature of seagrass meadows, most characteristic for shallow areas, is dynamics of their
@@
 performed using underwater videometric measurements, aerial and satellite imagery.
-
-%%%%%%%%%%%%%%%%%%%%%%%%% Discussion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}\label{sec:5}
-An approach of the seagrass spectral analysis, monitoring and mapping has been taken in this work, which integrates various research techniques and tools, combining remote sensing methods of spectral analysis of the seafloor cover types, and knowledge of the ecology of \textit{P.oceanica}, with the aim to develop a method of seagrass spectral optical discrimination for the seagrass mapping based on the aerial imagery classification.\vspace{1ex}\\
-In \textit{Chapter 1:} \nameref{sec:1} we discusses main objective of this MSc thesis, which was to study possibilities of seagrass mapping, based on the application of the remote sensing measurement of the seawater optical properties using hyperspectral radiometers.
-The relationship between the optical properties (spectral reflectance) of the seafloor cover types and hydrological parameters of the environment has been studied in order to analyse limitations and capabilities of broadband and narrowband sensors under the conditions of altering environmental parameters.
-For the retrieval of hydrological parameters - seawater radiance, irradiance and spectral reflectance of various seafloor cover types, - spectral optical field measurements were carried out in 2009 at the testing sites in Agia Pelagia, Heraklion and Ligaria beach, Crete Island using \ac{Trios-RAMSES} spectroradiometers.\\
-Further in \textit{Chapter 1:} 1.8 Research assumptions, we assume the constant values of the optical properties of the seawater, phytoplankton, total amount of suspended particles and solids, atmospheric conditions, as well as \ac{CDOM}, which have been set up in modelling part of this work, during \ac{WASI} simulations of various remote sensors.\vspace{1ex}\\
-The second \textit{Chapter 2:} \nameref{sec:2} starts from the review of the available research resources and then discusses various \nameref{sec:2-3} and concludes with reported experience of the remote sensing application towards seagrass mapping: \nameref{sec:2.4.1}.
-In section 2.2 of \textit{Chapter 2:} \nameref{sec:2}, \nameref{sec:2-2}, Measuring water optical properties: hyperspectral radiometers, the \ac{Trios-RAMSES}-ACC-\ac{UV} and \ac{Trios-RAMSES}-ARC spectroradiometers of \ac{Trios-RAMSES} Hyperspectral Sensor series are described.
-\\
-The instruments \ac{Trios-RAMSES} have been used during the fieldwork measurements-2009 for the collection of the reflectance spectra.
-The \ac{Trios-RAMSES}-ACC-\ac{UV} measures spectra in the wavelength domain between 280 nm and 500nm, the \ac{Trios-RAMSES}-ARC, suitable for UV and visible spectra, covers diapason of 320-950 nm with spectral accuracy of 0.3nm (better than 6\%), typical saturation (at 200nm) of  $1Wm^{-2}*nm^{-1}*sr^{-1}$ in 256 channels with a sampling interval of 3.3 nm/pixel and a \ac{FOV} is 7 degrees.
-The spectral reflectance of the seawater with and without sediments, respectively, was calculated by the ratio of the radiance and irradiance values.
-\\
-The reflectance spectra of \textit{P.oceanica} show (Fig.\ref{fig:4.10}, Fig.\ref{fig:A.12}, Fig.
-\ref{fig:A.12}, Fig.\ref{fig:A.13}, Fig.\ref{fig:A.25}, Fig.
-\ref{fig:A.26}) a values maximum between 450 nm and 600 nm, first, because of the chlorophyll absorption peak at 465 and 665nm (Fig.
-\ref{fig:A.2-2})\label{chlorophyll}, secondly, because of the weakening of \ac{CDOM} (or Gelbstoff) in the blue part of the \ac{VIS} spectrum, as it most strongly absorbs short wavelength light in blue to ultraviolet range, and finally, because the absorption of the seawater increases in the red part of the \ac{VIS} spectra.
-The decrease in spectral reflectance values of \textit{P.oceanica} after 660 nm (Fig.\ref{fig:A.13}) is caused by the second absorption peak of phytoplankton.
-The magnitude of the reflectance maximum slightly varies at single variables between about 8 \% and 12 \% (as on Fig.\ref{fig:A.12}) and is probably related to the individual pigmentation and colour composition of single leaves, their structure and geometric orientation, which naturally causes variations in radiance values.
-\\
-The ecological variables, specific to the field environmental conditions, were factored into the \ac{WASI}-based simulation models.
-Through the \ac{WASI} simulation process, imitating spectral properties of \textit{P.oceanica} and carbonate sand for various broadband and narrowband sensors, models were created that accounted for not only atmospheric conditions (i.e.
-sun zenith angle), but also height of water column, thus approaching it to the Mediterranean conditions, and chemical content of the seawater (i.e.
-amount of suspended particles, Gelbstoff, etc), which results in models of optical properties of ``seawater with sediments" and ``seawater without sediments".\\
-The \textit{in-situ} field large-scale matte-level level of seafloor monitoring was then upscaled to airborne Google Earth aerial imagery interpretation, to provide a meadow-level view of seagrass landscapes.
-An attempt of the small-scale mapping is designed on an example of Landsat satellite imagery.
-However, in upscaling to this third, small-scale mapping level further environmental variables need to be considered: health conditions of the seagrass, presence of other underwater vegetation (e.g.
-other seagrass species), hydrological specifications (e.g.
-direction and speed of currents, amplitude of tidal waves, etc), season, date and times of the image taken.
-Therefore, we focused on the first two levels in the current work.
-These different scale levels have been individually considered in terms of the seagrass spectral discernibility for monitoring and mapping, and have been brought together, to provide a roundup of the achieved results and an overview of what still has to be done in \textit{P.oceanica} seagrass mapping by future researchers (see \nameref{sec:7}).
-\vspace{1ex}\\
-In \textit{Chapter 3}: \nameref{sec:3}, describing data collection, a videographic approach tested in previous works, has been applied during the summer fieldwork, when we captured imagery and video footage of the seafloor on several routes of the boat in the Ligaria beach.\\
-The finding of \textit{Chapter 4}, \nameref{sec:4} showed that the relationship between the spectral reflectance of various seafloor cover types was tied to depth, i.e.
-water column height.
-Thus, the results of the in-situ fieldwork measurements revealed that spectral reflectance of \textit{P.oceanica} undergo alterations at depths of 0.5, 2.0 and 3.5m (Fig.\ref{fig:4.9}).
-These results indicate that seagrass \textit{P.oceanica} can in general be detected and discriminated from other seafloor cover types with varying environmental conditions, i.e.
-water column height, by hyperspectral spectroradiometers (\ac{Trios-RAMSES}), which positively answers the \textbf{first research question} of this thesis (``Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ conditions ?").
-\vspace{1ex}\\
-Further in \textit{Chapter 4}, \nameref{sec:4}, studies of the broadband versus narrowband sensors demonstrate that simulated spectra of the seagrass, made using \ac{WASI} modeller, have the best results at \ac{CZCS} scanner, especially devoted to the measurement of ocean color.
-The spectrum of \textit{P.oceanica} reflectance, simulated for \ac{CZCS}, covers the wavelength interval of 400-800 nm, and is distinctive for various depths.
-Other remote sensors (\ac{MODIS}, \ac{SeaWiFS}) may also be used for the seagrass mapping, because their technical characteristics enable to spectrally discriminate \textit{P.oceanica} seagrass from other seafloor cover types (Fig.\ref{fig:4.6}), particularly carbonate sand as tested in the current work.
-Therefore, the \textbf{second research question} of this MSc thesis (``Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\
-The graphs showing optical properties of seawater with and without sediments (Fig.\ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:4.2}) focused on spectral variability of the water with changed physical and chemical content.
-The alterations in the individual spectral signatures of single measurements (e.g.
-on Fig.\ref{fig:4.13}, Fig.\ref{fig:A.25}, Fig.\ref{fig:A.27}, Fig.\ref{fig:A.28}) reflect individual health properties of leaves: different nitrogen and chlorophyll content causing diverse colour pigmentation and light absorption, water content in leaves, plant physiological conditions, which vary across seagrass meadow, shoot morphology, etc.
-The differences in spectral reflectance values of the measurements taken on various days might have been caused by the impact of atmospheric conditions, such as solar radiation and sun illumination by different zenith angle.\vspace{1ex}\\
-For further development of the  remote sensing based monitoring and mapping of the seagrass and other seafloor cover types it is desirable to consider the substratum and underwater relief in a more detailed way, i.e.
-bathymetric properties of the testing sizes, because changes in relief directly cause and reinforce the colonisation process of \textit{P.oceanica} meadows.
-Thus, the depression and valleys in relief in general stimulate increase of the sedimentation process.
-In turn, it enables accumulation of necessary nutrient for seagrass plant growth, increases resource allocation to the seagrass roots for better exploitation of pore-water nutrients by the \textit{P.oceanica} shoots.
-As for patchiness in seagrass landscapes, besides natural reasons, the anthropogenic disturbances (e.g.
-ocean trawling) still remain a main cause for landscapes fragmentation.
 
@@
 
-\subsection{Satellite and aerial images covering research area of Crete island: selected examples}
+\subsection{Satellite and aerial images covering research area of Crete Island: selected examples}
 %\appendixname{\textbf{8:\\ Satellite and aerial images covering research area of Crete island: selected examples}}

### 2011-02-12 MScDiplom_Lemenkova.tex
Valentijn's comments: 1, 2, 3, 4, 5, 7: Updated Hypotheses, Assumtions, Research Questions. Added long names of Chapters to ToC.

@@
 \makeindex[name=people,title=Index of People,intoc,columns=3,columnseprule]
+\usepackage[fit]{truncate}
 \usepackage{fancyhdr}
 \pagestyle{fancy}
+\fancyhead[RO,LE]{\truncate{.95\headwidth}{\leftmark}}
 \usepackage{fancyheadings}
@@
 	\begin{itemize}	
-		\item Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying \textit{in-situ} conditions ?
-		\item Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types ?
+		\item Is \textit{P.oceanica} spectrally distinct from carbonate sand with varying \textit{in-situ} conditions ?
+		\item Do broadband and hyperspectral sensors provide enough radiometric information for spectral discrimination of seagrass, and therefore, can be used for mapping of \textit{P.oceanica} ?
 	\end{itemize}
 
-\subsection{Hypotheses testing}\index[concepts]{Research!questions!hypothesis}
+\subsection{Hypotheses}\index[concepts]{Research!questions!hypothesis}
 A statistical testing will be used to compare between the spectral responses of the different seafloor cover types (i.e. sand and \textit{P. oceanica}),
  whether it is spectrally distinct and at least one pair is statistically different at every spectral
-band.
-Hypothesis $H_o$: seagrass types are not spectrally distinct from other seafloor types with
-varying \textit{in-situ} conditions, which means
-$H_o$: μ1 =μ2 =μ3=...= μn.
+band.\vspace{1ex}\\
+For the \emph{Research Question 1} the Hypothesis $H_o$ claims: seagrass types are not spectrally distinct from other seafloor types with
+varying \textit{in-situ} conditions, which means $H_o: \mu1 =\mu2 =\mu3=...= \mu n$.
 The alternative Hypothesis $H_a$ claims the opposite statement: seagrass is spectrally
-distinct with varying \textit{in-situ} conditions, $H_o$: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn.
+distinct with varying \textit{in-situ} conditions, $H_o:\mu1\neq\mu2\neq\mu3\neq...\neq\mu n$.\vspace{1ex}\\
+For the \emph{Research Question 2} the Hypothesis $H_o$ claims: broadband and hyperspectral remote sensing data can be used for the mapping of \textit{P.oceanica}, because they do not provide enough radiometric information to discriminate sand from seagrass, which means $H_o:\mu1=\mu2=\mu3=...=\mu n$.
+The alternative Hypothesis $H_a$ claims the opposite statement: broadband and hyperspectral sensors do prove to provide enough radiometric information to discriminate sand from seagrass,
+$H_o: \mu1\neq\mu2\neq\mu3\neq ... \neq\mu n$.\vspace{1ex}\\
 The distribution of the spectral responses at every spectral band is assumed to be normal as well as
@@
 whether the results of the research are accurate, reasonable and correct.
+
+\subsection{Assumptions}\index[concepts]{Research!assumptions}
+The general research assumption, used in this work, in order to make feasible application of the remote sensing methods, includes some statements about the \textit{in-situ} atmospheric conditions, viewing angle, wave backscattering and other optical properties of the environment.
+They all definitely play a certain role and impact final results of optical measurements, but for practical reasons we have chosen to ignore their contribution in the spectral separability of seafloor cover types.
+We briefly list below general research assumptions for the optical properties of the environmental variables used in this study.
+\vspace{1ex}\\
+Weather conditions for the measurements are assumed to be ideal: clear, sunny, windless day.
+Otherwise wind roughens water surface causing sun glitter, and values of spectral reflectance may contain disturbances.
+\\ Sky radiance might be influenced by multiple reflectance between the sea surface albedo and the atmosphere and in general, sky radiance increases rapidly while viewing zenith angle is near to $90\,^{\circ}$ (i.e.
+in the evening hours).
+Therefore, ideally measurements should be done at noon, with as low zenith angle (Fig.\ref{fig:A.2-1})\label{zenith} as possible; otherwise, if the solar zenith angle is too high (e.g.
+approaching $60-70\,^{\circ}$), received data may contain noise.
+For \ac{WASI} simulations reflection factor of sky radiance is taken as 0.0201 with simulated ideal conditions: viewing angle ($0\,^{\circ}$ = nadir).
+The interval of water temperature taken as default lies in the diapason $17-25\,^{\circ}$ as to simulate the conditions of the Mediterranean Sea.
+However, in real time conditions the sun zenith angle of $45\,^{\circ}$ has been accepted as suitable.
+The accepted model-specific parameters of the \ac{WASI} simulator are shown in the Table \ref{tab:3}.\label{page-10} \vspace{1ex}\\
+The anisotropy factor of upwelling radiation or the quality (Q-) factor, showing the directionally dependency of the radiance, is taken as 5.
+We accepted some values of model-specific optical parameters as default values at \ac{WASI} simulator, which are shown in the Table \ref{tab:3}.
+Thus, concentration of phytoplankton is accepted at the interval of $0.035-0.089 mg^{-1}$ and concentration of large suspended particles is given to 8.
+Reference wavelength for \ac{CDOM}, or Gelbstoff, absorption is equal to 440.
+The backscattering (Fig.\ref{fig:backscat}) is accepted to be $0.00144m^{-1}$.
+The coefficient of attenuation remains equal to 1.0546, as set up by default at \ac{WASI}.\vspace{1ex}\\
+We also assume that concentration of non-chlorophyll particles (absorption at 0) as well as concentration of small suspended particles is equal to zero, so we do not count them in this work.
+Exponent of \ac{CDOM} (Gelbstoff) absorption is accepted as 0.0140.
+Finally, the \ac{BRDF} of bottom reflectance (sand), which defines the reflection of light  is at an opaque surface (Fig.\ref{fig:BRDF})\label{page-11}, is assumed to be 0.318 $sr^{-1}$.
 
@@
 
-The spectral reflectance, distinct for each seafloor cover type, enables their discrimination on the image. The pre-processing
-of the images includes imagery corrections for atmospheric noises and effects of the water
+The pre-processing of the images includes imagery corrections for atmospheric noises and effects of the water
 column. Reflectance spectra of the seagrass canopy at different depths of the water-column are
@@
 
-\subsection{Research assumptions}\index[concepts]{Research!assumptions}
-The general research assumption that is used for further applications of the remote sensing methods of this work is that optical properties of the spectral reflectance of the seawater with different bottom substrata enable to discriminate seafloor cover types using remote sensing products from broadband sensors.\vspace{1ex}\\
-The accepted model-specific parameters of the \ac{WASI} simulator are described in Table 4.2.
-The weather conditions for the measurements are assumed to be perfect: clear, sunny, windless days, because wind roughen water surface causing sun glitter, therefore, values of spectral reflectance may contain disturbances.
-\\ Sky radiance might be influenced by multiple reflectance between the sea surface albedo and the atmosphere and in general, sky radiance increases rapidly while viewing zenith angle is near to $90\,^{\circ}$ (i.e.
-in the evening hours).
-Therefore, ideally measurements should be done at noon, with as low zenith angle (Fig.\ref{fig:A.2-1})\label{zenith} as possible; otherwise, if the solar zenith angle is too high (e.g.
-approaching $60-70\,^{\circ}$), received data may contain noise.
-For \ac{WASI} simulations reflection factor of sky radiance is taken as 0.0201 with simulated ideal conditions: viewing angle ($0\,^{\circ}$ = nadir).
-The interval of water temperature taken as default lies in the diapason $17-25\,^{\circ}$ as to simulate the conditions of the Mediterranean Sea.
-However, in real time conditions the sun zenith angle of $45\,^{\circ}$ has been accepted as suitable.\vspace{1ex}\\
-The anisotropy factor of upwelling radiation or the quality (Q-) factor, showing the directionally dependency of the radiance, is taken as 5.
-We accept some values of optical parameters as default values at \ac{WASI}.
-Thus, concentration of phytoplankton is accepted at the interval of $0.035-0.089 mg^{-1}$ and concentration of large suspended particles is given to 8.
-Reference wavelength for \ac{CDOM} (Gelbstoff) absorption is equal to 440.
-The coefficient of attenuation remains equal to 1.0546, as set up by default at \ac{WASI}.\vspace{1ex}\\
-We also assume that concentration of non-chlorophyll particles (absorption at 0) as well as concentration of small suspended particles is equal to zero, so we do not count them in this work.
-Exponent of \ac{CDOM} (Gelbstoff) absorption is accepted as 0.0140.
-Finally, the \ac{BRDF} of bottom reflectance (sand), which defines the reflection of light  is at an opaque surface (Fig.A.2.5), is assumed to be 0.318 $sr^{-1}$.
-
-\pagebreak
-
 %%%%%%%%%%%%%%%%%  Section2   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Overview of literature]{Seagrass monitoring: overview of literature and research resources} \label{sec:2}
-\markright{II. Seagrass monitoring: overview of literature and research resources}
+\section[Seagrass monitoring: overview of literature and research resources]{Seagrass monitoring: overview of literature and research resources} \label{sec:2}
+\markright{Seagrass monitoring: overview of literature and research resources}
 \renewcommand{\headrulewidth}{0.4pt}
@@
 
-\subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}\label{sec:2-2}\label{page-12}
+\subsection[Measuring optical properties of benthic vegetation: hyperspectral radiometers]{Measuring optical properties of benthic vegetation: hyperspectral radiometers%
+\sectionmark{Measuring optical properties}}
+\sectionmark{Overview of literature...}\label{sec:2-2}\label{page-12}
+
 Among various research approaches for the seagrass mapping the most popular methods are based on application of the remote sensing techniques,
@@
 
-\subsection[Radiative Transfer Models (RTM)]{Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing tools}\label{sec:2-3}
+%\subsection[Radiative Transfer Models (RTM)]{Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing software tools}\label{sec:2-3}
+
+\subsection[Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing software tools]{Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing software tools%
+\subsectionmark{Radiative Transfer Models (RTM)}}
+\sectionmark{Overview of literature...}\label{sec:2-3}
 
@@
 
-\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}\label{page-15}
+\subsection[Application of the remote sensing data towards seagrass mapping]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}\label{page-15}
+
 Various methods and approaches have been applied towards mapping of the seagrasses, based on
@@
 result of the spectral absorption in different bands, typical for each target object.
-Spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13})\label{page-17} is largely influenced by the water
-depth where it is located, and is generally decreasing in values by increasing depths.
+Spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13})\label{page-17} is largely influenced by the water depth where it is located, and is generally decreasing in values by increasing depths.
 The most important spectral diapason\index[concepts]{Seawater optical properties!spectral reflectance!spectral diapason} for marine mapping of submerged vegetation and, particularly, for seagrass, is 350-800 nm \cite{Barille09}\label{Barille09}.
@@
 
-\subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}\index[concepts]{Remote sensing!multispectral imagery} 
+\subsubsection[Review of multispectral imagery used for seagrass mapping]{Review of multispectral imagery used for seagrass mapping}\subsectionmark{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}\index[concepts]{Remote sensing!multispectral imagery}
+
 Seagrass mapping using remotely sensed data\label{page-18} from multispectral sensors is based on the classification
@@
 
-\subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery}
+\subsubsection[Review of hyperspectral imagery used for seagrass mapping]{Review of hyperspectral imagery used for seagrass mapping}\subsectionmark{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery}
+
 The application of the hyperspectral sensors \label{page-20} is most effective and provide more accurate classification
@@
 
-\subsubsection{Comparison of seagrass spectra }\label{sec:2.5.3}
+\subsubsection[Comparison of seagrass spectra]{Comparison of seagrass spectra}\subsectionmark{Comparison of seagrass spectra}\label{sec:2.5.3}
 
@@
 coverage of the seagrass meadows in the research area. The underwater measurements of the seagrass
-coverage were carried out by taking video footage and photos of ca 0.5 m2 size each (Fig.\ref{fig:A.7}, Fig.\ref{fig:A.55}).
+coverage were carried out by taking video footage and photos of ca 0.5 $m^2$ size each (Fig.\ref{fig:A.7}, Fig.\ref{fig:A.55}).
 The results of the underwater videometric measurements include series of digital images helping to
@@
 results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
-in water column, water temperature, sun angle, concentration of Gelbstoff (\ac{CDOM}), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small
-particle \cite{Gege04}\label{Gege04}.
+in water column, water temperature, sun angle, concentration of Gelbstoff (\ac{CDOM}), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small particles \cite{Gege04}\label{Gege04}.
 
@@
 	\rowcolors{1}{Aquamarine}{Azure}
-	\caption{Model-specific parameters of water WASI adjusted to simulate environment of the Mediterranean Sea along Crete. Page: $<$\pageref{page-36}$>$}
+	\caption{Model-specific parameters of water WASI adjusted to simulate environment of the Mediterranean Sea along Crete. Pages: $<$\pageref{page-10}$>$, $<$\pageref{page-36}$>$} \label{tab:3}
 	\centering
@@
 	  \end{tabular}	  
-	   \label{tab:3}
 \end{table}
@@
 
-\subsection{Illustrations of some concepts and principles of the remote sensing, relevant for this work}
+\subsection{Illustrations of remote sensing principles used for seagrass monitoring}
 %%%%%%%%%%  \appendixname{Illustrations of some concepts and principles of the remote sensing}  %%%%%%%%%%%%%%
@@
 		\includegraphics[scale=0.6]{backscattering.jpg}
-		\caption{Example of wave backscattering from the vegetation (here: rice).
-Source: \href{http://cse.niaes.affrc.go.jp/miwa/esid/highlight/microwave-backscatter.html}{Yoshio Inoue}}.\label{fig:backscat}
+		\caption{Example of wave backscattering from the vegetation.
+Source: \href{http://cse.niaes.affrc.go.jp/miwa/esid/highlight/microwave-backscatter.html}{Yoshio Inoue}.
+Page: $<$\pageref{page-11}$>$}.\label{fig:backscat}
 	\end{center}
@@
 		\caption{BRDF is a ratio of reflected radiance along $\omega_o$ to the irradiance from direction $\omega_i$, all parameterized by azimuth angle $\phi$ and zenith angle $\theta$.
-Source: \href{http://en.wikipedia.org/wiki/Portal:Contents/Quick_index}{Wikipedia.org}}.\label{fig:BRDF}
-	\end{center}
-\end{figure}
-
-\pagebreak
-
-\subsection[Instrumental adjustment and tuning]{Instrumental adjustment and tuning. (Trios-RAMSES setup)}
+Source: \href{http://en.wikipedia.org/wiki/Portal:Contents/Quick_index}{Wikipedia.org}.
+Page: $<$\pageref{page-11}$>$}.\label{fig:BRDF}
+	\end{center}
+\end{figure}
+
+\pagebreak
+
+\subsection[Instrumental adjustment and tuning: Trios-RAMSES setup]{Instrumental adjustment and tuning: Trios-RAMSES setup}
 %\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}
@@
 
-\subsection[Types of seagrass structural patterns]{Types of seagrass structural patterns, Ligaria beach, Crete}
+\subsection[Types of seagrass structural patterns: Ligaria beach, Crete]{Types of seagrass structural patterns: Ligaria beach, Crete}
 %\appendixname{\textbf{3: \\Types of seagrass structural patterns, Ligaria beach, Crete}}

### 2011-02-13 MScDiplom_Lemenkova.tex
Comments of Valentijn: 8, 9, 11, 36 (RTM, upscale mapping, discussion of spectra, optical measurement techniques)

@@
 \newline
-We briefly discuss below the most effective RTM software and algorithms from a range of various best-known up-to-date tools. 
+We briefly discuss below the most effective RTM software and algorithms, suitable for underwater radiance simulations, from a range of various best-known up-to-date tools. 
 	\begin{itemize}
@@
 Among its newest updates is generation of atmospheric correction data; it also has an option to write spectral output in binary, and convert it to \ac{ASCII}.
-		\item [*] A \ac{DISORT}\index[concepts]{Seawater optical properties!RTM!DISORT}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.
 		\item [*] The \ac{CRTM} \index[concepts]{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA}, \ac{NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
-		\item [*] An \ac{4AOP}\index[concepts]{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{the French Space Agency}, \ac{CNES}.
-Being a fine program which enables accurate computations of the radiance, defining other emissivity functions and other set up, it is however, particularly designed for the infrared waves (thus, not quite suitable in our case).
-To the advantages of this freeware we would count its user friendly interface and open source accessibility for the academic purposes.
 	\end{itemize}
@@
 involves intensive and time-consuming \textit{in-situ} observations during the fieldwork, as, for example,
-reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. \\
+reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}.  The results of the \textit{in-situ} measurements and observation are usually
+managed and treated using integrated \ac{GIS} approach, as, e.g. reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.\\
 Other methods of seagrass \textit{in-situ} monitoring are based on the application of the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then collect the return echoes for the analysis (Fig.\ref{fig:sonar}).
@@
 Another limitation of the acoustic
-techniques is that, initially intended for the bathymetric surveying, acoustic equipment is mostly designed for the geomorphological and geological studies of the underwater substratum and are, therefore, more adjusted for the benthic habitat discrimination, and they are not effective for the identifying of the biological species composition or even the presence of aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}.
-The results of the \textit{in-situ} measurements and observation are usually
-managed and treated using integrated \ac{GIS} approach, as, e.g. reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.
+techniques is that, initially intended for the bathymetric surveying, acoustic equipment is mostly designed for the geomorphological and geological studies of the underwater substratum and are, therefore, more adjusted for the benthic habitat discrimination, and they are not effective for the identifying of the biological species composition or even the presence of aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}.\\
+The current study is based on the application of the remote sensing optical measurement techniques, due to their effectiveness, non-destructive nature and availability of necessary tools: spectral radiometers and \ac{RTM}.
 
@@
 digitized aerial photographs, \ac{GPS} data, remote sensing and \ac{SCUBA}-based (Fig.\ref{fig:A.52}) fieldwork measurements.
+
+\begin{wrapfigure}{r}{0.4\textwidth}[H]
+	\centering
+	\includegraphics[scale=0.35]{Fig-11.jpg}
+	\caption{\textit{In situ} optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
+		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}. \\ Pages: $<$\pageref{page-17}$>$, $<$\pageref{page-21}$>$}
+	\label{fig:13}
+\end{wrapfigure}
+
 \ac{SCUBA}-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
@@
 collection in the field \cite{Schultz08}\label{Schultz08}. 
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.35]{Fig-11.jpg}
-	\caption{\textit{In situ} optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
-		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}. \\ Pages: $<$\pageref{page-17}$>$, $<$\pageref{page-21}$>$}
-	\label{fig:13}
-\end{wrapfigure}
-
 However, it cannot cover large areas for small-scale mapping.
@@
 periodic monitoring of the seagrass meadows and gives access to the distant and unapproachable
-areas. Integrated together with \ac{GIS} vector layers and maps, remote sensing data enable historical
+areas. 
+Integrated together with \ac{GIS} vector layers and maps, remote sensing data enable historical
 mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.
-However, application of the remote sensing techniques for mapping of submerged vegetation,
-seafloor cover types and benthic vegetation, \textit{inter alia} seagrasses, are still in their development.
-Approaches and methods for the seagrass protection and monitoring still remain location-specific or, at
-least, nation-specific, depending to large extent on the tools available for the researchers \cite{Mellors09a}\label{Mellors09a}.
-
-\begin{wrapfigure}{r}{0.5\textwidth}
+
+\begin{wrapfigure}{r}{0.4\textwidth}[H]
 	\centering
@@
 
+However, application of the remote sensing techniques for mapping of submerged vegetation,
+seafloor cover types and benthic vegetation, \textit{inter alia} seagrasses, are still in their development.
+Approaches and methods for the seagrass protection and monitoring still remain location-specific or, at
+least, nation-specific, depending to large extent on the tools available for the researchers \cite{Mellors09a}\label{Mellors09a}.
 Universal, international, standardized methods for seagrass directly for seagrasses
@@
 
-\begin{table}[htbp]
+\begin{table}[htbp]\footnotesize
 	\caption{Characteristics of selected ocean-colour sensors. Page: $<$\pageref{page-47}$>$}
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.30]{Fig-14.jpg}
+	\caption{Multispectral vs. hyperspectral band coverage. \\ Source: \cite{Sabins97}\label{Sabins97}. Pages: $<$\pageref{page-20}$>$,  $<$\pageref{page-18}$>$}
+	\label{fig:2.5}
+\end{wrapfigure}
+
 Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial resolution of 10 m (Fig.\ref{fig:2.4}), covering covers a surface area of 3600 km² (60*60 km swath), 26-day
@@
 imagery has been applied for the seagrass mapping due to its high resolution and accessibility. 
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-14.jpg}
-	\caption{Multispectral vs. hyperspectral band coverage. \\ Source: \cite{Sabins97}\label{Sabins97}. Pages: $<$\pageref{page-20}$>$,  $<$\pageref{page-18}$>$}
-	\label{fig:2.5}
-\end{wrapfigure}
-
 Thus, the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
@@
 than IKONOS \cite{Mumby02}\label{Mumby02}.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.15]{Fig-15.jpg}
+	\caption{Reflectance spectra of \textit{Thalassia} seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
+		reflectance. Source:\cite{Yang10}\label{Yang10}. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-54}$>$}
+	\label{fig:2.6}
+\end{wrapfigure}
+
 Another comparative analysis of the application of \href{http://www.itres.com/products/imagers/casi550}{CASI}, Landsat and \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird} imagery \cite{Phinn08}\label{Phinn08}
@@
 macrophytes in shallow waters of the Lake Constance. 
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.15]{Fig-15.jpg}
-	\caption{Reflectance spectra of \textit{Thalassia} seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
-		reflectance. Source:\cite{Yang10}\label{Yang10}. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-54}$>$}
-	\label{fig:2.6}
-\end{wrapfigure}
-
 In regard to the methods chosen for the image interpretation, the supervised classifications proves to
@@
 pure reflectance from the benthos; as a result, it may cause significant errors \cite{Dierssen03}\label{Dierssen03}.
-Therefore, for the accurate assessments of various seafloor cover types, water
+For the accurate assessments of various seafloor cover types, water
 column depth and its optical properties, methods of supervised classification should be preferably
 used as being more suitable in classification and interpretations of imagery. 
+However, it is only true if spectrally distinct regions of the spectrum are covered by a space-born sensor
+ and if the atmospheric distortion and viewing geometry is not degrading the radiometric quality at essential wavelengths.
 While this and other studies \cite{Phinn08}\label{Phinn08} demonstrated the advantages and success of the

### 2011-02-13 MScDiplom_Lemenkova.tex
Comments of Valentijn: 12, 13, 26, 27 - done, 25 - partly done. Discussion of advantages of WASI, moved pictures, updated methods. restructures Appendices, added paragraphs to discussion of P.oceanica.

@@
 	\acro{GRETL}{\href{http://gretl.sourceforge.net/}{GNU Regression, Econometrics and Time-series Library}}
+	\acro{GUI}{\href{http://en.wikipedia.org/wiki/Graphical_user_interface}{Graphical User Interface}}
 	\acro{HyMap}{\href{http://www.hyvista.com/}{Hyperspectral Mapper}}
@@
 \subsubsection{General characteristics of \textit{P.oceanica}}\index[concepts]{Seagrasses!\textit{P.oceanica}}
+\paragraph{Morphology of \textit{P.oceanica}}
 The endemic Mediterranean seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}), is a main species in
@@
 Some authors \cite{Guidetti08,Montefalcone09}\label{Guidetti08} \label{Montefalcone09} used status and population dynamics of \textit{P.oceanica} as indicators for the evaluation of the meadow health status.
+\paragraph{Phenology of \textit{P.oceanica}}
 There are several environmental factors, determining the growth of the \textit{P.oceanica}. 
@@
 	\begin{center}
-		\includegraphics[scale=0.35]{Sampling_design.jpg}
+		\includegraphics[scale=0.30]{Sampling_design.jpg}
 		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach. Page: $<$\pageref{page-26}$>$}\label{fig:A.4}
@@
 classify seafloor cover types (Fig.\ref{fig:A.56}) and seagrass meadows, according to the differences in the structure,
-colour, texture and shapes of the depicted objects\label{page-27}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.50]{Fig-21.jpg}
-	\caption{Locations of the video measurements and GPS tracklogs, Ligaria. Pages: $<$\pageref{page-25}$>$, $<$\pageref{page-26}$>$}
-	\label{fig:3.6}
-\end{figure}
-
-Along the Ligaria beach there are several types of the seagrass landscapes, namely, seagrass meadows
-continuously covering vast areas (Fig.\ref{fig:A.3} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:A.3} (b)) and isolated seagrass patches, or mattes,
-which located separately from broad meadows (Fig.\ref{fig:A.3} (c), Fig.\ref{fig:A.56}).
+colour, texture and shapes of the depicted objects\label{page-27}.\\
+There are several types of the seagrass landscapes alon Ligaria beach, namely, seagrass meadows
+continuously covering vast areas (Fig.\ref{fig:A.3} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:A.3} (b)) and isolated seagrass patches, or mattes, located separately from broad meadows (Fig.\ref{fig:A.3} (c), Fig.\ref{fig:A.56}).
 The results of the underwater videometric measurements of the \ac{Olympus ST} cameras made during the
@@
 \subsection{Data pre-processing}\index[concepts]{Data!pre-processing}
+
+\subsubsection{Auxiliary data: spectral dataset}
 A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of \textit{P.oceanica} made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab.
 \ref{tab:A.8}), 1-445 for 14th October for \textit{P.oceanica}; 1-106 for water without sediments, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab.\ref{tab:17}).
+Measurements of 2009 has been used for the database and statistical analysis, because 2010 data pool was not available.
 All these data have been analyzed and statistically tested.
@@
 spectrum of constant albedo at \ac{WASI} (Fig.\ref{fig:A.9}).
-We included measurements of the Thalassia seagrass
-(measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectance of
+
+\subsubsection{Modelling method: WASI water colour simulator}
+To estimate and further analyze particularities of the P.oceanica remote sensing reflectance spectra over regions with different bathymetry (0.5, 1.5 and 2.5 meters column depth) we used \ac{WASI} \ac{RTM} software, which incorporates measurements of the seafloor reflectance and water column optical properties.
+We have chosen \ac{WASI} \ac{RTM} software among other aquatic \ac{RTM} (2.3.\nameref{sec:2-3}) due to its effective cognitive approach and because it is specifically adjusted and developed for the purpose of aquatic optical modelling.
+Consistency of simple and logical \ac{GUI} (Fig.\ref{fig:4.7})\label{GUI} let the researcher to easily define model parameters, which can be changed directly in the main window tools menu, and to choose between forward or inverse
+calculations.
+Besides other important advantages of \ac{WASI} we would mention its learnability: available documentation, supporting materials, learning curve with hints and tips, as well as its user-friendly \ac{GUI}, enabling to learn this product in a most effective and prompt way in a tight schedule of MSc studies.
+Technically, \ac{WASI} need for system memory is minimised and the installation is simple and easy.\\
+We included measurements of the \textit{Thalassia} seagrass
+(measured at Southern Chinese Sea by C.Yang and D.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectance of
 different seagrass species under various environmental conditions.
@@
 further research experiment. The results of the linear interpolation (Fig.\ref{fig:A.35}) demonstrate values of the
-sand spectral reflectance with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
+sand spectral reflectance with 1 nm interval covering the wavelength diapason of 400-750 nm.
+
+\subsubsection{Implementation of statistical analysis}
 The initial measured data were stored in raw-oriented format, so that re-formatting them into the
@@
 properties of seawater.\\
+To analyze average values of spectral reflectance of \textit{P.oceanica} by means of the measured sets we calculated 
+mean values for the total set of measurements 1 – 350 (15th October): mean of min and max values, 
+mean of average for each waveband, mean for statistical quartiles for each wavelength and mean for median, respectively. 
+The adjusted averages evaluate mean values, needed for statistical analysis of the datasets. 
+The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (\nameref{app}). \\
+As a result, we received mean values of spectral reflectance of \textit{P.oceanica} for extreme, average, median and both quartiles,
+independent of the individual sets 1-350. \\
 We used several statistical methods to asses data received as a result of measurements. 
@@
 
-\begin{table}[htbp]
-	\rowcolors{1}{LightSteelBlue}{Honeydew}
-	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, with average values (for sets 1 - 350).
-Generalisation up to step 20 nm.
-Page: $<$\pageref{page-40}$>$}\label{tab:4.1}
-	\begin{center}
-	\begin{tabular}{|c|c|c|c|c|c|c|}
-	\hline
-	Wavelength, nm & mean & Q1 & min & Q3 & max & median \\ \hline
-		\textbf{318.23354} & 0.030467 & 0.038907 & 0.029764 & 0.069762 & 0.116100 & 0.048984 \\ \hline
-		\textbf{338.29900} & 0.033560 & 0.043030 & 0.032847 & 0.076252 & 0.128290 & 0.054334 \\ \hline
-		\textbf{358.37932} & 0.039897 & 0.051320 & 0.039131 & 0.091143 & 0.154409 & 0.066354 \\ \hline
-		\textbf{378.47247} & 0.045786 & 0.058849 & 0.045039 & 0.105993 & 0.177150 & 0.078339 \\ \hline
-		\textbf{398.57640} & 0.052346 & 0.066859 & 0.051625 & 0.122328 & 0.201038 & 0.091905 \\ \hline
-		\textbf{418.68905} & 0.058942 & 0.074842 & 0.058240 & 0.138984 & 0.226028 & 0.106207 \\ \hline
-		\textbf{438.80838} & 0.070690 & 0.089761 & 0.069999 & 0.167825 & 0.267712 & 0.130580 \\ \hline
-		\textbf{458.93235} & 0.080107 & 0.102089 & 0.079392 & 0.191424 & 0.304933 & 0.149762 \\ \hline
-		\textbf{479.05891} & 0.086278 & 0.110348 & 0.085546 & 0.207940 & 0.332934 & 0.162881 \\ \hline
-		\textbf{499.18600} & 0.098498 & 0.126592 & 0.097689 & 0.239161 & 0.380603 & 0.188192 \\ \hline
-		\textbf{519.31159} & 0.110222 & 0.143976 & 0.109022 & 0.265948 & 0.421370 & 0.210237 \\ \hline
-		\textbf{539.43363} & 0.125694 & 0.166156 & 0.124310 & 0.302678 & 0.475842 & 0.240492 \\ \hline
-		\textbf{559.55006} & 0.131844 & 0.175865 & 0.130257 & 0.319385 & 0.503907 & 0.253726 \\ \hline
-		\textbf{579.65885} & 0.121915 & 0.166084 & 0.120186 & 0.302078 & 0.485661 & 0.237873 \\ \hline
-		\textbf{599.75794} & 0.077236 & 0.111113 & 0.075669 & 0.200557 & 0.343876 & 0.154016 \\ \hline
-		\textbf{619.84529} & 0.057038 & 0.085386 & 0.055588 & 0.154781 & 0.281737 & 0.117003 \\ \hline
-		\textbf{639.91885} & 0.050841 & 0.077438 & 0.049432 & 0.141810 & 0.265181 & 0.106188 \\ \hline
-		\textbf{659.97657} & 0.038118 & 0.059971 & 0.036997 & 0.110774 & 0.223366 & 0.082411 \\ \hline
-		\textbf{680.01641} & 0.032892 & 0.052265 & 0.032012 & 0.095789 & 0.198389 & 0.071686 \\ \hline
-		\textbf{700.03633} & 0.034756 & 0.058426 & 0.033643 & 0.101504 & 0.215413 & 0.075789 \\ \hline
-		\textbf{720.03426} & 0.026287 & 0.047084 & 0.025162 & 0.082940 & 0.193535 & 0.061184 \\ \hline
-		\textbf{740.00817} & 0.009740 & 0.022355 & 0.010145 & 0.042861 & 0.132244 & 0.029414 \\ \hline
-		\textbf{759.95601} & 0.008651 & 0.020234 & 0.009191 & 0.039371 & 0.128165 & 0.026558 \\ \hline
-		\textbf{779.87573} & 0.006019 & 0.014157 & 0.006397 & 0.028320 & 0.100314 & 0.019322 \\ \hline
-		\textbf{799.76528} & 0.008607 & 0.020030 & 0.009050 & 0.037323 & 0.118123 & 0.026535 \\ \hline
-		\textbf{819.62263} & 0.008777 & 0.020295 & 0.009288 & 0.037598 & 0.118055 & 0.026799 \\ \hline
-		\textbf{839.44571} & 0.002562 & 0.008010 & 0.002864 & 0.017472 & 0.076515 & 0.011432 \\ \hline
-		\textbf{859.23249} & 0.002653 & 0.006921 & 0.002897 & 0.016121 & 0.070377 & 0.009889 \\ \hline
-		\textbf{878.98091} & 0.002327 & 0.008080 & 0.002821 & 0.018625 & 0.082462 & 0.011617 \\ \hline
-		\textbf{898.68893} & 0.004077 & 0.013098 & 0.004709 & 0.031217 & 0.101146 & 0.019767 \\ \hline
-		\textbf{918.35451} & 0.006215 & 0.020162 & 0.007229 & 0.045213 & 0.199711 & 0.029950 \\ \hline
-		\textbf{937.97559} & 0.005783 & 0.026132 & 0.007307 & 0.065475 & 0.195072 & 0.040848 \\ \hline
-		\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
-	\end{tabular}
-	\end{center}
-\end{table}
-
 The results enable to study reflectance properties of the seagrass and other
@@
 
-\subsubsection{Water color spectral simulation using WASI}\index[concepts]{Data!processing}\label{page-35}
+\subsubsection{Spectral simulation of aquatic objects}\index[concepts]{Data!processing}\label{page-35}
 The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
@@
 \end{figure}
-
-Assuming that the null hypothesis, Hypothesis $H_o$, is true, which is “the spectral distinguishability of
-the seagrass \textit{P.oceanica} from other seafloor types is not changing with varying \textit{in-situ} conditions,
-$H_o$: μ1 =μ2 =μ3=...= μn.”, what is the probability of observing the values of spectral reflectance of
-\textit{P.oceanica} at various depths, for the test statistic, that is at least as extreme as the values of spectral
-reflectance of \textit{P.oceanica} at these depths, that was actually observed? The alternative Hypothesis $H_a$
-Fig.15. Multiplot display of spectral reflectance of \textit{P.oceanica}. Measurement sets 1-100, 15th
-October. Bold red dots show the mean values within each dataset; vertical areas: quartiles Q1-Q2.
-claims the opposite statement: “the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly
-changing with varying \textit{in-situ} conditions, i.e. increasing depth, $H_o$: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
+%
+Answering the first research question, we suggest the following statement.
+If the Hypothesis $H_o$ is true, then the spectral distinguishability of the seagrass \textit{P.oceanica} from other seafloor types (carbonate sand) is not changing with varying \textit{in-situ} environmental conditions, $H_o: \mu1 =\mu2 =\mu3=...= \mu n$.
+The alternative Hypothesis $H_a$ claims the opposite statement: “the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with varying in-situ conditions, i.e.
+increasing depth, $H_o: \mu1\neq\mu2\neq\mu3\neq...\neq\mu n$.
 Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.
-To analyze average values of spectral reflectance of \textit{P.oceanica} by means of the measured sets we calculated 
-mean values for the total set of measurements 1 – 350 (15th October): mean of min and max values, 
-mean of average for each waveband, mean for statistical quartiles for each wavelength and mean for median, respectively. 
-The adjusted averages evaluate mean values, needed for statistical analysis of the datasets. 
-The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (\nameref{app}). 
-As a result, we have mean values of spectral reflectance of \textit{P.oceanica} for extreme, average, median and both quartiles,
-independent of the individual sets 1-350.
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.5]{GNU-14.jpg}
+		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
+Series 126-150.
+Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
+\href{http://www.gnuplot.info/}{Gnuplot} visualization.
+Page: $<$\pageref{page-39}$>$­}
+		\label{fig:A.27}
+	\end{center}
+\end{figure}
 
@@
 spectral distinguishability of the seagrass \textit{P.oceanica} from other seafloor types is not changing with
-varying \textit{in-situ} conditions, $H_o$: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
-P(reject $H_o$|$H_o$ is valid)= P(X$>$c|p=)=.05
+varying \textit{in-situ} conditions, $H_o: \mu1 =\mu2 =\mu3=...= \mu n$, which gives the following outcome.
+P(reject $H_o\mid H_o$ is valid)= P(X$>c\mid p$=.05).
 where c comes for critical values and p=probability.
@@
 Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with
-varying \textit{in-situ} conditions, i.e. increasing depth, $H_o$: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
+varying \textit{in-situ} conditions, i.e. increasing depth, $H_o: \mu\neq\mu2\neq\mu3\neq ... \neq\mu n$.
 As a result of the statistical testing, we came to the following conclusion.
 The Hypothesis $H_a$ is true, which claims that the spectral discernibility of the seagrass \textit{P.oceanica} is
-distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, $H_o$: μ1≠μ2≠μ3≠...≠ μn .
+distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, $H_o: \mu1\neq\mu2\neq\mu3\neq...\neq\mu n$ .
 Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
@@
 	\includegraphics[scale=0.30]{Fig-30.jpg}
-	\caption{\ac{WASI} water color spectral simulation modeller: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m. Pages: $<$\pageref{page-35}$>$, $<$\pageref{page-31}$>$}
+	\caption{\ac{WASI} water color spectral simulation modeller: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m.
+Pages: $<$\pageref{page-35}$>$, $<$\pageref{page-31}$>$, $<$\pageref{GUI}$>$}
 	\label{fig:4.7}
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.50]{Fig-21.jpg}
+	\caption{Locations of the video measurements and GPS tracklogs, Ligaria. Pages: $<$\pageref{page-25}$>$, $<$\pageref{page-26}$>$}
+	\label{fig:3.6}
+\end{figure}
+
 \subsection{Results of videographic measurements: seafloor types on Crete island}
@@
 %\appendixname{\textbf{7:\\ Data processing and statistical evaluation: \\ Results of the measurements of the seafloor optical properties.}}
+
+%%% Gretl start
+\begin{table}[htbp]
+	\rowcolors{1}{Ivory}{GhostWhite}
+	\caption{Basic mathematical formulae used for statistical analysis of the measurement set of the spectral reflectance. Page: $<$\pageref{page-41}$>$}
+	\label{tab:13}
+	\begin{center}
+	\begin{tabular}{|c|c|}
+		\hline\hline
+		Basic Statistics & Math Formulae \\ \hline\hline
+		\hline
+
+		\vspace{1ex}
+
+		Mean & $\Sigma(x_i/n=\bar x)$ \\ \hline
+		Variance & $1 \over n-1$ $\Sigma(x_i-\bar x)=s^2_x$  \\ \hline
+		Standard Deviation & $s=s^2$ \\ \hline
+		Coefficient of Variation & $s/\bar x$ \\ \hline
+		Skewness & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^3$  \\ \hline
+		Excess Kurtosis & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^4-3$  \\ \hline
+	\end{tabular}
+	\end{center}
+	\label{tab:19}
+\end{table}
+%%% Gretl end
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.35]{GNU-Radiance_Bezier.jpg}
+	\caption{Radiance of the seawater with sediments.
+Measured in aquarium tank.
+Bezier Interpolation.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{dataprep}$>$, $<$\pageref{page-40}$>$, $<$\pageref{page-30}$>$}
+	\label{fig:A.16}
+\end{figure}
 
@@
 \end{figure}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.30]{GNU-Irr-Splines.jpg}
+		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-31}$>$­}
+		\label{fig:A.17}
+	\end{center}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.30]{GNU-Irr.jpg}
+	\caption{Irradiance of the seawater, measured in aquarium tank.
+Smooth Bezier interpolation.
+Visualization in \href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{dataprep}$>$, $<$\pageref{page-30}$>$}
+	\label{fig:A.18}
+\end{figure}
+
+%%%
+\begin{table}
+	\rowcolors{1}{Lavender}{LightYellow}
+	\caption{Robust estimation of the seawater radiance (measured in aquarium tank, Heraklion): results of the \textit{least absolute deviation} (LAD), Series:V16.
+\href{http://gretl.sourceforge.net/}{Gretl}}
+	\begin{center}
+		Model 6: LAD, using observations 1:01--8:23 ($T$ = 191)\\
+		Dependent variable: v16\\
+	\vspace{1em}
+
+	\begin{tabular}{|c c c c c|}
+		  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
+		const &   0.738911 &     0.0862011 &       8.5719 &         0.0000 \\
+		v1 &   $-$0.000775186 &     0.000114010 &       $-$6.7993 &         0.0000 \\
+	\end{tabular}
+
+	\vspace{1ex}
+	\begin{tabular}{lrlr}
+		Median depend. var &  0.227857 & S.D. dependent var &  0.205031 \\
+		Sum absolute resid &  30.22342 & Sum squared resid &  8.059632 \\
+		Log-likelihood &  28.74727 & Akaike criterion & $-$53.49454 \\
+		Schwarz criterion & $-$46.99000 & Hannan--Quinn & $-$50.85990 \\
+	\end{tabular}
+	\label{tab:12}
+	\end{center}
+\end{table}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.25]{GNU-09.jpg}
+		\caption{Spectral reflectance of the seawater with sediments.
+Bezier interpolation.
+Mean value is shown by the vertical impulses linestyle.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{dataprep}$>$­, $<$\pageref{page-31}$>$}
+		\label{fig:A.23}
+	\end{center}
+\end{figure}
+
+%%%Begin Plot Gretl
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.4]{G-polynom-trend-wo.jpg}
+		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15.
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
+	\end{center}
+	\label{fig:A.32}
+\end{figure}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.4]{G-polynom-trend-with.jpg}
+		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
+	\end{center}
+	\label{fig:A.33}
+\end{figure}
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.4]{G-exp-mov-aver.jpg}
+		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
+	\end{center}
+	\label{fig:A.34}
+\end{figure}
+\pagebreak
+%%%end Plot Gretl
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.4]{G-gamma-rad.jpg}
+		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}}
+		\label{fig:A.19}
+	\end{center}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.4]{G-normal-dist.jpg}
+	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}­}
+	\label{fig:A.20}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection[Data processing and statistical evaluation: \textit{P.oceanica}]{Data processing and statistical evaluation: \textit{P.oceanica}}
+%\appendixname{\textbf{7:\\ Data processing and statistical evaluation: \\ Results of the measurements of the seafloor optical properties.}}
 \begin{figure}[H]
@@
 \end{figure}
+\begin{table}[htbp]
+	\rowcolors{1}{LightSteelBlue}{Honeydew}
+	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, with average values (for sets 1 - 350).
+Generalisation up to step 20 nm.
+Page: $<$\pageref{page-40}$>$}\label{tab:4.1}
+	\begin{center}
+	\begin{tabular}{|c|c|c|c|c|c|c|}
+	\hline
+	Wavelength, nm & mean & Q1 & min & Q3 & max & median \\ \hline
+		\textbf{318.23354} & 0.030467 & 0.038907 & 0.029764 & 0.069762 & 0.116100 & 0.048984 \\ \hline
+		\textbf{338.29900} & 0.033560 & 0.043030 & 0.032847 & 0.076252 & 0.128290 & 0.054334 \\ \hline
+		\textbf{358.37932} & 0.039897 & 0.051320 & 0.039131 & 0.091143 & 0.154409 & 0.066354 \\ \hline
+		\textbf{378.47247} & 0.045786 & 0.058849 & 0.045039 & 0.105993 & 0.177150 & 0.078339 \\ \hline
+		\textbf{398.57640} & 0.052346 & 0.066859 & 0.051625 & 0.122328 & 0.201038 & 0.091905 \\ \hline
+		\textbf{418.68905} & 0.058942 & 0.074842 & 0.058240 & 0.138984 & 0.226028 & 0.106207 \\ \hline
+		\textbf{438.80838} & 0.070690 & 0.089761 & 0.069999 & 0.167825 & 0.267712 & 0.130580 \\ \hline
+		\textbf{458.93235} & 0.080107 & 0.102089 & 0.079392 & 0.191424 & 0.304933 & 0.149762 \\ \hline
+		\textbf{479.05891} & 0.086278 & 0.110348 & 0.085546 & 0.207940 & 0.332934 & 0.162881 \\ \hline
+		\textbf{499.18600} & 0.098498 & 0.126592 & 0.097689 & 0.239161 & 0.380603 & 0.188192 \\ \hline
+		\textbf{519.31159} & 0.110222 & 0.143976 & 0.109022 & 0.265948 & 0.421370 & 0.210237 \\ \hline
+		\textbf{539.43363} & 0.125694 & 0.166156 & 0.124310 & 0.302678 & 0.475842 & 0.240492 \\ \hline
+		\textbf{559.55006} & 0.131844 & 0.175865 & 0.130257 & 0.319385 & 0.503907 & 0.253726 \\ \hline
+		\textbf{579.65885} & 0.121915 & 0.166084 & 0.120186 & 0.302078 & 0.485661 & 0.237873 \\ \hline
+		\textbf{599.75794} & 0.077236 & 0.111113 & 0.075669 & 0.200557 & 0.343876 & 0.154016 \\ \hline
+		\textbf{619.84529} & 0.057038 & 0.085386 & 0.055588 & 0.154781 & 0.281737 & 0.117003 \\ \hline
+		\textbf{639.91885} & 0.050841 & 0.077438 & 0.049432 & 0.141810 & 0.265181 & 0.106188 \\ \hline
+		\textbf{659.97657} & 0.038118 & 0.059971 & 0.036997 & 0.110774 & 0.223366 & 0.082411 \\ \hline
+		\textbf{680.01641} & 0.032892 & 0.052265 & 0.032012 & 0.095789 & 0.198389 & 0.071686 \\ \hline
+		\textbf{700.03633} & 0.034756 & 0.058426 & 0.033643 & 0.101504 & 0.215413 & 0.075789 \\ \hline
+		\textbf{720.03426} & 0.026287 & 0.047084 & 0.025162 & 0.082940 & 0.193535 & 0.061184 \\ \hline
+		\textbf{740.00817} & 0.009740 & 0.022355 & 0.010145 & 0.042861 & 0.132244 & 0.029414 \\ \hline
+		\textbf{759.95601} & 0.008651 & 0.020234 & 0.009191 & 0.039371 & 0.128165 & 0.026558 \\ \hline
+		\textbf{779.87573} & 0.006019 & 0.014157 & 0.006397 & 0.028320 & 0.100314 & 0.019322 \\ \hline
+		\textbf{799.76528} & 0.008607 & 0.020030 & 0.009050 & 0.037323 & 0.118123 & 0.026535 \\ \hline
+		\textbf{819.62263} & 0.008777 & 0.020295 & 0.009288 & 0.037598 & 0.118055 & 0.026799 \\ \hline
+		\textbf{839.44571} & 0.002562 & 0.008010 & 0.002864 & 0.017472 & 0.076515 & 0.011432 \\ \hline
+		\textbf{859.23249} & 0.002653 & 0.006921 & 0.002897 & 0.016121 & 0.070377 & 0.009889 \\ \hline
+		\textbf{878.98091} & 0.002327 & 0.008080 & 0.002821 & 0.018625 & 0.082462 & 0.011617 \\ \hline
+		\textbf{898.68893} & 0.004077 & 0.013098 & 0.004709 & 0.031217 & 0.101146 & 0.019767 \\ \hline
+		\textbf{918.35451} & 0.006215 & 0.020162 & 0.007229 & 0.045213 & 0.199711 & 0.029950 \\ \hline
+		\textbf{937.97559} & 0.005783 & 0.026132 & 0.007307 & 0.065475 & 0.195072 & 0.040848 \\ \hline
+		\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
+	\end{tabular}
+	\end{center}
+\end{table}
+
 \begin{figure}[H]
@@
 \end{figure}
-\pagebreak
 %%%
 
-\begin{table}[htbp]
+\begin{table}[H]\footnotesize
 	\rowcolors{1}{Ivory}{GhostWhite}
@@
 
-\begin{table}[htbp]
+\begin{table}[H]\footnotesize
 	\rowcolors{1}{Ivory}{GhostWhite}
@@
 \end{table}
-\pagebreak
 
@@
 	\end{center}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.35]{GNU-Radiance_Bezier.jpg}
-	\caption{Radiance of the seawater with sediments.
-Measured in aquarium tank.
-Bezier Interpolation.
-\href{http://www.gnuplot.info/}{Gnuplot}.
-Pages: $<$\pageref{dataprep}$>$, $<$\pageref{page-40}$>$, $<$\pageref{page-30}$>$}
-	\label{fig:A.16}
 \end{figure}
@@
 
-\rowcolors{1}{GhostWhite}{LightCyan}
+\rowcolors{1}{GhostWhite}{LightCyan}\footnotesize
 \begin{longtable}{|c|c|c|c|c|c|c|}
@@
 \end{longtable}
-\pagebreak
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.30]{GNU-Irr-Splines.jpg}
-		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}. Page: $<$\pageref{page-31}$>$­}
-		\label{fig:A.17}
-	\end{center}
-\end{figure}
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.30]{GNU-Irr.jpg}
-	\caption{Irradiance of the seawater, measured in aquarium tank.
-Smooth Bezier interpolation.
-Visualization in \href{http://www.gnuplot.info/}{Gnuplot}.
-Pages: $<$\pageref{dataprep}$>$, $<$\pageref{page-30}$>$}
-	\label{fig:A.18}
-\end{figure}
-\pagebreak
-\newpage
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.4]{G-gamma-rad.jpg}
-		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}}
-		\label{fig:A.19}
-	\end{center}
-\end{figure}
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.4]{G-normal-dist.jpg}
-	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}­}
-	\label{fig:A.20}
-\end{figure}
 
 %%% Gretle start
-\begin{table}
+\begin{table}[H]
 	\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16.
@@
 
-%%%
-\begin{table}
-	\rowcolors{1}{Lavender}{LightYellow}
-	\caption{Robust estimation of the seawater radiance (measured in aquarium tank, Heraklion): results of the \textit{least absolute deviation} (LAD), Series:V16.
-\href{http://gretl.sourceforge.net/}{Gretl}}
-	\begin{center}
-		Model 6: LAD, using observations 1:01--8:23 ($T$ = 191)\\
-		Dependent variable: v16\\
-	\vspace{1em}
-
-	\begin{tabular}{|c c c c c|}
-		  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
-		const &   0.738911 &     0.0862011 &       8.5719 &         0.0000 \\
-		v1 &   $-$0.000775186 &     0.000114010 &       $-$6.7993 &         0.0000 \\
-	\end{tabular}
-
-	\vspace{1ex}
-	\begin{tabular}{lrlr}
-		Median depend. var &  0.227857 & S.D. dependent var &  0.205031 \\
-		Sum absolute resid &  30.22342 & Sum squared resid &  8.059632 \\
-		Log-likelihood &  28.74727 & Akaike criterion & $-$53.49454 \\
-		Schwarz criterion & $-$46.99000 & Hannan--Quinn & $-$50.85990 \\
-	\end{tabular}
-	\label{tab:12}
-	\end{center}
-\end{table}
-%%%
 \pagebreak
 \newpage
-
-\begin{table}[H]
+%%%%%%%
+\begin{table}[H]\footnotesize
 	\rowcolors{1}{Ivory}{GhostWhite}
@@
 	\centering
+	\includegraphics[scale=0.45]{Partial_Corr.jpg}
+	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+Visualization in \href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{A-24}$>$}
+	\label{fig:A.22}
+\end{figure}
+\begin{figure}[H]
+	\centering
 	\includegraphics[scale=0.45]{Autocorr.jpg}
@@
 \end{table}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.45]{Partial_Corr.jpg}
-	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.).
-Visualization in \href{http://www.spss.com/}{SPSS}.
-Page: $<$\pageref{A-24}$>$}
-	\label{fig:A.22}
-\end{figure}
 \pagebreak
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.25]{GNU-09.jpg}
-		\caption{Spectral reflectance of the seawater with sediments.
-Bezier interpolation.
-Mean value is shown by the vertical impulses linestyle.
-\href{http://www.gnuplot.info/}{Gnuplot}.
-Pages: $<$\pageref{dataprep}$>$­, $<$\pageref{page-31}$>$}
-		\label{fig:A.23}
-	\end{center}
-\end{figure}
 
@@
 \end{figure}
+
 %%% Gretl start
-\begin{table}[htbp]
-	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Basic mathematical formulae used for statistical analysis of the measurement set of the spectral reflectance. Page: $<$\pageref{page-41}$>$}
-	\label{tab:13}
-	\begin{center}
-	\begin{tabular}{|c|c|}
-		\hline\hline
-		Basic Statistics & Math Formulae \\ \hline\hline
-		\hline
-
-		\vspace{1ex}
-
-		Mean & $\Sigma(x_i/n=\bar x)$ \\ \hline
-		Variance & $1 \over n-1$ $\Sigma(x_i-\bar x)=s^2_x$  \\ \hline
-		Standard Deviation & $s=s^2$ \\ \hline
-		Coefficient of Variation & $s/\bar x$ \\ \hline
-		Skewness & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^3$  \\ \hline
-		Excess Kurtosis & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^4-3$  \\ \hline
-	\end{tabular}
-	\end{center}
-	\label{tab:19}
-\end{table}
-%%% Gretl end
-
-%%% Gretl start
-\begin{table}[htbp]
+\begin{table}[H]
 	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textit{P.oceanica} (15.X.).
@@
 
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.5]{GNU-14.jpg}
-		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
-Series 126-150.
-Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots).
-\href{http://www.gnuplot.info/}{Gnuplot} visualization.
-Page: $<$\pageref{page-39}$>$­}
-		\label{fig:A.27}
-	\end{center}
-\end{figure}
-
-\pagebreak
-
-\begin{table}[H]
+\begin{table}[H]\footnotesize
 	\rowcolors{1}{Ivory}{GhostWhite}
@@
 
-\begin{table}[H]
+\begin{table}[H]\footnotesize
 	\rowcolors{1}{Ivory}{GhostWhite}
@@
 \end{table}
-
-%%%Gretl begin
-\begin{table}[H]
-	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica} (15.X.).
-\href{http://gretl.sourceforge.net/}{Gretl}.
-Page: $<$\pageref{page-41}$>$}
-	\begin{center}
-	Eigenanalysis of the Correlation Matrix\\
-	\vspace{1em}
-
-		\begin{tabular}{|c|c|c|c|}
-			{Component } &  {Eigenvalue } & {Proportion} & {Cumulative} \\
-			1 &   6.4848    &   0.9264    &   0.9264 \\
-			2 & 0.4949 &      0.0707  &     0.9971 \\
-			3 &  0.0140    &   0.0020    &   0.9991 \\
-			4 &    0.0052  &     0.0007    &   0.9998 \\
-			5 &   0.0008   &    0.0001     &  1.0000 \\
-			6 &  0.0002    &   0.0000      & 1.0000 \\
-			7 &   0.0001   &    0.0000     &  1.0000 \\
-		\end{tabular}
-
-	Eigenvectors (component loadings)
-
-	\vspace{1ex}
-
-		\begin{tabular}{|c|c|c|c|c|c|c|c|}
-			v1        &	     -0.308  &	  -0.880  &	   0.334   &	 -0.126  &	   0.033  &	  -0.042   &	  0.006\\ \hline
-			v2      &	      0.383  &	  -0.295  &	  -0.512   &	 -0.230  &	  -0.365  &	  -0.370   &	 -0.425\\ \hline
-			v3      &	      0.383  &	  -0.300  &	  -0.465   &	 -0.010  &	   0.562  &	   0.307   &	  0.370\\ \hline
-			v4      &	      0.390  &	  -0.145  &	   0.203   &	  0.714  &	  -0.129  &	  -0.415   &	  0.293\\ \hline
-			v5      &	      0.392  &	  -0.091  &	   0.300   &	  0.262  &	   0.020  &	   0.568   &	 -0.597\\ \hline
-			v6      &	      0.392  &	  -0.011  &	   0.259   &	 -0.410  &	  -0.543  &	   0.299   &	  0.476\\ \hline
-			v7      &	      0.390  &	   0.138  &	   0.460   &	 -0.429  &	   0.488  &	  -0.427   &	 -0.113\\ \hline
-		\end{tabular}
-	\end{center}
-	\label{tab:24}
-\end{table}
-%%%Gretl end
 \pagebreak
@@
 	\begin{center}
-		\includegraphics[scale=0.38]{GNU-15.jpg}
+		\includegraphics[scale=0.35]{GNU-15.jpg}
 		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}.
@@
 \end{figure}
+%%%Gretl begin
+\begin{table}[H]\footnotesize
+	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
+	\begin{center}
+	Eigenanalysis of the Correlation Matrix\\
+	\vspace{1em}
+
+		\begin{tabular}{|c|c|c|c|}
+			{Component } &  {Eigenvalue } & {Proportion} & {Cumulative} \\
+			1 &   6.4848    &   0.9264    &   0.9264 \\
+			2 & 0.4949 &      0.0707  &     0.9971 \\
+			3 &  0.0140    &   0.0020    &   0.9991 \\
+			4 &    0.0052  &     0.0007    &   0.9998 \\
+			5 &   0.0008   &    0.0001     &  1.0000 \\
+			6 &  0.0002    &   0.0000      & 1.0000 \\
+			7 &   0.0001   &    0.0000     &  1.0000 \\
+		\end{tabular}
+
+	Eigenvectors (component loadings)
+
+	\vspace{1ex}
+
+		\begin{tabular}{|c|c|c|c|c|c|c|c|}
+			v1        &	     -0.308  &	  -0.880  &	   0.334   &	 -0.126  &	   0.033  &	  -0.042   &	  0.006\\ \hline
+			v2      &	      0.383  &	  -0.295  &	  -0.512   &	 -0.230  &	  -0.365  &	  -0.370   &	 -0.425\\ \hline
+			v3      &	      0.383  &	  -0.300  &	  -0.465   &	 -0.010  &	   0.562  &	   0.307   &	  0.370\\ \hline
+			v4      &	      0.390  &	  -0.145  &	   0.203   &	  0.714  &	  -0.129  &	  -0.415   &	  0.293\\ \hline
+			v5      &	      0.392  &	  -0.091  &	   0.300   &	  0.262  &	   0.020  &	   0.568   &	 -0.597\\ \hline
+			v6      &	      0.392  &	  -0.011  &	   0.259   &	 -0.410  &	  -0.543  &	   0.299   &	  0.476\\ \hline
+			v7      &	      0.390  &	   0.138  &	   0.460   &	 -0.429  &	   0.488  &	  -0.427   &	 -0.113\\ \hline
+		\end{tabular}
+	\end{center}
+	\label{tab:24}
+\end{table}
+%%%Gretl end
 \pagebreak
@@
 
-\begin{figure}[H]
-	\begin{center}
-	\includegraphics[scale=0.4]{G-autocorr-sand51-75.jpg}
-	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand.
-Series 51-75, variable 27.
-\href{http://gretl.sourceforge.net/}{Gretl} visualization.
-Page: $<$\pageref{page-41}$>$­}
-	\label{fig:A.29}
-	\end{center}
-\end{figure}
-
+\subsubsection[Data processing and statistical evaluation: \textit{carbonate sand}]{Data processing and statistical evaluation: \textit{carbonate sand}}
+%\appendixname{\textbf{7:\\ Data processing and statistical evaluation: \\ Results of the measurements of the seafloor optical properties.}}
 %%% Gretl start
-\begin{table}[htbp]
+
+\begin{table}[H]\footnotesize
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of carbonate sand, \href{http://gretl.sourceforge.net/}{Gretl}.
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}.
 Page: $<$\pageref{page-41}$>$}
@@
 %%% Gretl end
-\pagebreak
 
@@
 		\includegraphics[scale=0.4]{G-QQplot-sand.jpg}
-		\caption{Normal Q-Q plot: estimated versus observed values of the measurement of carbonate sand, variable 27. Series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
+		\caption{Normal Q-Q plot: estimated versus observed values of the measurement of \textit{carbonate sand}, variable 27.
+Series 51-75.
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
 	\label{fig:A.30}
@@
 \end{figure}
-
-%%% Gretl start
-\begin{table}[H]
-	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}.
-Page: $<$\pageref{page-41}$>$}
-	\begin{center}
-		Model 2: Quantile estimates, using observations 1:01--5:09 ($T$ = 105)\\
-		Dependent variable: v27\\
-		$\tau$ = 0.5\\
-		Asymptotic standard errors assuming IID errors\\
-
-\vspace{1em}
-
-\begin{tabular}{c c c c c}
-	***   & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
-	const &  0.837827 &    0.0386417 &      21.6820 &         0.0000 \\
-	v1 &   $-$0.00111120 &     6.60949\textrm{e--005} &       $-$16.8122 &         0.0000 \\ \hline \hline
-\end{tabular}
-
-\vspace{1ex}
-	\begin{tabular}{lrlr}
-		Median depend. var &  0.191024 & S.D. dependent var &  0.135439 \\
-		Sum absolute resid &  8.625039 & Sum squared resid &  1.370360 \\
-		Log-likelihood &  84.64508 & Akaike criterion & $-$165.2902 \\
-		Schwarz criterion & $-$159.9822 & Hannan--Quinn & $-$163.1393 \\ \hline \hline
-	\end{tabular}
-	\end{center}
-\label{tab:29}
-\end{table}
-%%% Gretl end
-\pagebreak
-
-%%%Gretl begin
-\begin{table}[H]
-	\begin{small}
-	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
-	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
-	\label{tab:30}
-	\begin{center}
-	\begin{tabular}{|p{15mm} p{15mm} p{15mm} p{15mm} || p{15mm} p{15mm}p{15mm} p{15mm}|}
-		omega & scaled frequency & periods & log spectral density & omega & scaled frequency & periods & log spectral density \\ \\ \hline\hline
-		0.05984      & 1    &      105.00  &      -1.2466  &  1.61568   &   27    &        3.89     &    -11.111\\ 
-		 0.11968     &  2   &        52.50   &      -5.8761 &   1.67552   &   28    &        3.75     &    -11.170\\ 
-		 0.17952     &  3   &        35.00   &      -4.0657 &  1.73536   &   29    &        3.62     &    -11.208\\ 
-		 0.23936     &  4   &        26.25   &      -7.1031 &  1.79520   &   30    &        3.50     &    -11.461\\ 
-		 0.29920     &  5   &        21.00   &      -6.1742 &  1.85504   &   31    &        3.39     &    -11.380\\ 
-		 0.35904     &  6   &        17.50   &      -7.2772  & 1.91488    & 32     &       3.28      &   -11.686\\ 
-		 0.41888     &  7     &      15.00   &      -8.7802 &  1.97472    &  33    &        3.18     &    -11.624\\ 
-		 0.47872     &  8     &      13.13   &      -7.6806 & 2.03456    &  34    &        3.09         &-11.724\\ 
-		 0.53856     &  9     &      11.67   &      -8.4848  & 2.09440    &  35    &        3.00    &     -11.818\\ 
-		 0.59840     & 10    &       10.50  &       -8.5764 &    2.15423    &  36    &        2.92    &     -11.805\\ 
-		 0.65824     & 11    &        9.55   &      -9.4643 &  2.21407    &  37    &        2.84    &     -12.002\\ 
-		 0.71808     & 12    &        8.75   &      -9.0868 &  2.27391    &  38    &        2.76    &     -11.930\\ 
-		 0.77792     & 13    &        8.08   &      -9.3608 &  2.33375    &  39    &        2.69    &     -12.010\\ 
-		 0.83776     & 14    &        7.50     &    -9.9332 &  2.39359    &  40    &        2.63    &     -12.124\\ 
-		 0.89760     & 15    &        7.00     &    -9.6836 &  2.45343    &  41    &        2.56    &     -12.190\\ 
-		 0.95744     & 16    &        6.56     &    -10.00  & 2.51327    &  42           & 2.50    &     -12.248\\ 
-		 1.01728     & 17    &        6.18     &    -9.8994 & 2.57311    &  43      &      2.44    &     -12.306\\ 
-		 1.07712     & 18    &        5.83     &    -10.188 & 2.63295    &  44      &      2.39    &     -12.323\\ 
-		 1.13696     & 19    &        5.53     &    -10.197 & 2.69279    &  45      &      2.33    &     -12.353\\ 
-		 1.19680     & 20    &        5.25     &    -10.467 &  2.75263    &  46      &      2.28    &     -12.420\\ 
-		 1.25664     & 21    &        5.00     &    -10.505 &  2.81247    &  47      &      2.23    &     -12.382\\ 
-		 1.31648     & 22    &        4.77    &    -10.641 &  2.87231    &  48      &      2.19    &     -12.487\\ 
-		 1.37632     & 23    &        4.57     &    -10.906 &  2.93215    &  49      &    2.14      &   -12.433\\ 
-		 1.43616     & 24    &        4.38     &    -10.957 &  2.99199    &  50        &    2.10    &     -12.609\\ 
-		 1.49600      &25    &        4.20     &    -10.785 &   3.05183    &  51      &      2.06    &     -12.410\\ 
-		 3.11167    &  52      &      2.02    &     -12.477 & & & & \\   \hline \hline
-	\end{tabular}
-	\end{center}
-	\end{small}
-	Number of observations = 105
-\end{table}
-%%% Gretl end
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.4]{G-period-sand30.jpg}
-		\caption{Plot illustrating periodogram for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
-	\end{center}	
-\end{figure}
-
-%%%Begin Plot Gretl
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.4]{G-polynom-trend-wo.jpg}
-		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15.
-\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\end{center}
-	\label{fig:A.32}
-\end{figure}
-
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.4]{G-polynom-trend-with.jpg}
-		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\end{center}
-	\label{fig:A.33}
-\end{figure}
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[scale=0.4]{G-exp-mov-aver.jpg}
-		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\end{center}
-	\label{fig:A.34}
-\end{figure}
-\pagebreak
-%%%end Plot Gretl
 
@@
 \begin{longtable}{|r|r|r|r|r|r|r|}
-	\caption{Results of the statistical analysis of spectral reflectance of \textbf{carbonate sand}, with average values (for sets 1-3).
+	\caption{Results of the statistical analysis of spectral reflectance of \textit{carbonate sand}, with average values (for sets 1-3).
 Wavelength step: 3 nm.
@@
 	\includegraphics[scale=0.5]{Fig-26.jpg}
-	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach.
-Results of single measurement set made by spectroradiometer Trios-RAMSES.
+	\caption{Spectral reflectance of \textit{carbonate sand} on A.Pelagia beach.
+Results of single measurement set made by spectroradiometer \ac{Trios-RAMSES}.
 \href{http://www.gnuplot.info/}{Gnuplot}.
@@
 \end{figure}
+%%% Gretl start
+\begin{table}[H]\footnotesize
+	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-41}$>$}
+	\begin{center}
+		Model 2: Quantile estimates, using observations 1:01--5:09 ($T$ = 105)\\
+		Dependent variable: v27\\
+		$\tau$ = 0.5\\
+		Asymptotic standard errors assuming IID errors\\
+
+\vspace{1em}
+
+\begin{tabular}{c c c c c}
+	***   & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
+	const &  0.837827 &    0.0386417 &      21.6820 &         0.0000 \\
+	v1 &   $-$0.00111120 &     6.60949\textrm{e--005} &       $-$16.8122 &         0.0000 \\ \hline \hline
+\end{tabular}
+
+\vspace{1ex}
+	\begin{tabular}{lrlr}
+		Median depend. var &  0.191024 & S.D. dependent var &  0.135439 \\
+		Sum absolute resid &  8.625039 & Sum squared resid &  1.370360 \\
+		Log-likelihood &  84.64508 & Akaike criterion & $-$165.2902 \\
+		Schwarz criterion & $-$159.9822 & Hannan--Quinn & $-$163.1393 \\ \hline \hline
+	\end{tabular}
+	\end{center}
+\label{tab:29}
+\end{table}
+%%% Gretl end
+
+%%%Gretl begin
+\begin{table}[H]
+	\begin{small}
+	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
+	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\label{tab:30}
+	\begin{center}
+	\begin{tabular}{|p{15mm} p{15mm} p{15mm} p{15mm} || p{15mm} p{15mm}p{15mm} p{15mm}|}
+		omega & scaled frequency & periods & log spectral density & omega & scaled frequency & periods & log spectral density \\ \\ \hline\hline
+		0.05984      & 1    &      105.00  &      -1.2466  &  1.61568   &   27    &        3.89     &    -11.111\\ 
+		 0.11968     &  2   &        52.50   &      -5.8761 &   1.67552   &   28    &        3.75     &    -11.170\\ 
+		 0.17952     &  3   &        35.00   &      -4.0657 &  1.73536   &   29    &        3.62     &    -11.208\\ 
+		 0.23936     &  4   &        26.25   &      -7.1031 &  1.79520   &   30    &        3.50     &    -11.461\\ 
+		 0.29920     &  5   &        21.00   &      -6.1742 &  1.85504   &   31    &        3.39     &    -11.380\\ 
+		 0.35904     &  6   &        17.50   &      -7.2772  & 1.91488    & 32     &       3.28      &   -11.686\\ 
+		 0.41888     &  7     &      15.00   &      -8.7802 &  1.97472    &  33    &        3.18     &    -11.624\\ 
+		 0.47872     &  8     &      13.13   &      -7.6806 & 2.03456    &  34    &        3.09         &-11.724\\ 
+		 0.53856     &  9     &      11.67   &      -8.4848  & 2.09440    &  35    &        3.00    &     -11.818\\ 
+		 0.59840     & 10    &       10.50  &       -8.5764 &    2.15423    &  36    &        2.92    &     -11.805\\ 
+		 0.65824     & 11    &        9.55   &      -9.4643 &  2.21407    &  37    &        2.84    &     -12.002\\ 
+		 0.71808     & 12    &        8.75   &      -9.0868 &  2.27391    &  38    &        2.76    &     -11.930\\ 
+		 0.77792     & 13    &        8.08   &      -9.3608 &  2.33375    &  39    &        2.69    &     -12.010\\ 
+		 0.83776     & 14    &        7.50     &    -9.9332 &  2.39359    &  40    &        2.63    &     -12.124\\ 
+		 0.89760     & 15    &        7.00     &    -9.6836 &  2.45343    &  41    &        2.56    &     -12.190\\ 
+		 0.95744     & 16    &        6.56     &    -10.00  & 2.51327    &  42           & 2.50    &     -12.248\\ 
+		 1.01728     & 17    &        6.18     &    -9.8994 & 2.57311    &  43      &      2.44    &     -12.306\\ 
+		 1.07712     & 18    &        5.83     &    -10.188 & 2.63295    &  44      &      2.39    &     -12.323\\ 
+		 1.13696     & 19    &        5.53     &    -10.197 & 2.69279    &  45      &      2.33    &     -12.353\\ 
+		 1.19680     & 20    &        5.25     &    -10.467 &  2.75263    &  46      &      2.28    &     -12.420\\ 
+		 1.25664     & 21    &        5.00     &    -10.505 &  2.81247    &  47      &      2.23    &     -12.382\\ 
+		 1.31648     & 22    &        4.77    &    -10.641 &  2.87231    &  48      &      2.19    &     -12.487\\ 
+		 1.37632     & 23    &        4.57     &    -10.906 &  2.93215    &  49      &    2.14      &   -12.433\\ 
+		 1.43616     & 24    &        4.38     &    -10.957 &  2.99199    &  50        &    2.10    &     -12.609\\ 
+		 1.49600      &25    &        4.20     &    -10.785 &   3.05183    &  51      &      2.06    &     -12.410\\ 
+		 3.11167    &  52      &      2.02    &     -12.477 & & & & \\   \hline \hline
+	\end{tabular}
+	\end{center}
+	\end{small}
+	Number of observations = 105
+\end{table}
+%%% Gretl end
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[scale=0.4]{G-period-sand30.jpg}
+		\caption{Plot illustrating periodogram for v30 from measurements of \textit{carbonate sand}, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
+	\end{center}	
+\end{figure}
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[scale=0.4]{G-autocorr-sand51-75.jpg}
+	\caption{Graph of the autocorrelation analysis of the measurements of \textit{carbonate sand}.
+Series 51-75, variable 27.
+\href{http://gretl.sourceforge.net/}{Gretl} visualization.
+Page: $<$\pageref{page-41}$>$­}
+	\label{fig:A.29}
+	\end{center}
+\end{figure}
+
 \pagebreak

### 2011-02-14 MScDiplom_Lemenkova.tex
Comments of Valentijn: 6,10,16,17. Updated image of sensors ASTER, MODIS, MERIS, SeaWiFS (now with spectral bands)

@@
 	%\acro{ERS}{European Remote Sensing satellite}
+	\acro{GDAL}{\href{http://www.gdal.org/}{Geospatial Data Abstraction Library}}
 	\acro{GIS}{Geographic Information System}
@@
 	\acro{GRETL}{\href{http://gretl.sourceforge.net/}{GNU Regression, Econometrics and Time-series Library}}
+	\acro{HCMR}{\href{http://www.hcmr.gr/}{Hellenic Centre for Marine Research }}
 	\acro{GUI}{\href{http://en.wikipedia.org/wiki/Graphical_user_interface}{Graphical User Interface}}
@@
 of organic matter and seagrass decline \cite{Holmer08}\label{Holmer08}. Seagrasses are subject to anthropogenic
-nutrient (N and P) loading, which may occasionally casue morphological (e.g. leaf length) and
-physiological (e.g. chlorophyll and nitrogen content of the leaves) responses towards changed
+nutrient (N and P) loading \cite{Cambridge09}\label{Cambridge09}, \cite{Leoni08}\label{Leoni08}, which may occasionally casue morphological (e.g.
+leaf length) and physiological (e.g.
+chlorophyll and nitrogen content of the leaves) responses towards changed
 environmental conditions \cite{Leoni06,Leoni07}\label{Leoni06}\label{Leoni07}. 
@@
 events can take up to several years and often it is only possible by means of the seed reserves from the local
-environmental surveys \cite{McKenzie07}\label{McKenzie07}. Other threats for seagrasses in tropical areas are 
-increased nutrient availability in the coastal zones, increased
+environmental surveys \cite{McKenzie07}\label{McKenzie07}. Other threats for seagrasses are 
+increased nutrient availability in the coastal zones \cite{Alcoverro1997}\label{Alcoverro1997}, increased
 eutrophication and invasive macroalgae. These processes have strong affect on the status of the
@@
 and secondly, the content and cover fraction of the seafloor have the most distinctive effect on the
-spectral reflectance of the water. \\
+spectral reflectance of the water. 
+In the environmental conditions of Crete Island, \textit{P.oceanica} may grow at depths up to six meters; however, the most usual depth is four meters, 
+which caused our decision to focus on depths up to four meters, and use these depths for further WASI modelling. 
 Therefore, in the current research we focus on these two major factors, and study the response of the
 water reflectance towards changing conditions of the water column depth and seafloor bottom cover
-fraction (seagrass \textit{P.oceanica}, carbonate sand and silt).\\
+fraction (seagrass \textit{P.oceanica} and carbonate sand).\\
 The research aims to explore the limitations of the application of the \ac{RTM} and remote sensing techniques 
@@
 dynamics in \textit{P.oceanica} seagrass distribution along the northern coasts of Crete. Aerial imagery from
-Google Earth with high spatial resolution, suitable for the large-scale detailed mapping of seagrass
+Google Earth with high spatial resolution (Fig.\ref{fig:A.50})\label{GG-2}, suitable for the large-scale detailed mapping of seagrass
 mattes, is used for the improvement of the accuracy of large seagrass meadows and separate mattes
@@
 Other methods of seagrass \textit{in-situ} monitoring are based on the application of the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then collect the return echoes for the analysis (Fig.\ref{fig:sonar}).
-One of the examples of the acoustic sonar sensor designed for seabed classification is the British \href{http://sonavision.co.uk/products.asp?cat_id=1}{RoxAnn} system, which application for the seafloor mapping is \href{http://www.ozcoasts.org.au/geom_geol/toolkit/Tech_CA_asc.jsp}{very well described} on the \href{http://www.ozcoasts.org.au/}{OzCoasts website}.\\ However, the applications of the acoustic methods requires specialized expensive equipment and is mostly used in the deep open waters, combined with bathymetric measurements.
-Another limitation of the acoustic
-techniques is that, initially intended for the bathymetric surveying, acoustic equipment is mostly designed for the geomorphological and geological studies of the underwater substratum and are, therefore, more adjusted for the benthic habitat discrimination, and they are not effective for the identifying of the biological species composition or even the presence of aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}.\\
-The current study is based on the application of the remote sensing optical measurement techniques, due to their effectiveness, non-destructive nature and availability of necessary tools: spectral radiometers and \ac{RTM}.
-
-\subsection[Application of the remote sensing data towards seagrass mapping]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}\label{page-15}
-
-Various methods and approaches have been applied towards mapping of the seagrasses, based on
-digitized aerial photographs, \ac{GPS} data, remote sensing and \ac{SCUBA}-based (Fig.\ref{fig:A.52}) fieldwork measurements.
 
@@
 
+One of the examples of the acoustic sonar sensor designed for seabed classification is the British \href{http://sonavision.co.uk/products.asp?cat_id=1}{RoxAnn} system, which application for the seafloor mapping is \href{http://www.ozcoasts.org.au/geom_geol/toolkit/Tech_CA_asc.jsp}{very well described} on the \href{http://www.ozcoasts.org.au/}{OzCoasts website}.\\ However, the applications of the acoustic methods requires specialized expensive equipment and is mostly used in the deep open waters, combined with bathymetric measurements.
+Another limitation of the acoustic
+techniques is that, initially intended for the bathymetric surveying, acoustic equipment is mostly designed for the geomorphological and geological studies of the underwater substratum and are, therefore, more adjusted for the benthic habitat discrimination, and they are not effective for the identifying of the biological species composition or even the presence of aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}.\\
+The current study is based on the application of the remote sensing optical measurement techniques, due to their effectiveness, non-destructive nature and availability of necessary tools: spectral radiometers and \ac{RTM}.
+
+\subsection[Application of the remote sensing data towards seagrass mapping]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}\label{page-15}
+
+Various methods and approaches have been applied towards mapping of the seagrasses, based on
+digitized aerial photographs, \ac{GPS} data, remote sensing and \ac{SCUBA}-based (Fig.\ref{fig:A.52}) fieldwork measurements.
 \ac{SCUBA}-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
@@
 Approaches and methods for the seagrass protection and monitoring still remain location-specific or, at
-least, nation-specific, depending to large extent on the tools available for the researchers \cite{Mellors09a}\label{Mellors09a}.
+least, nation-specific, depending to large extent on the tools available for the researchers \cite{Bakran-Petricioli06}\label{Bakran-Petricioli06}, \cite{Mellors09a}\label{Mellors09a}.
 Universal, international, standardized methods for seagrass directly for seagrasses
@@
 generally based on the assumption that various types of the seafloor bottom have different
-characteristics of the reflectivity, which is visually expressed in distinct colors of the objects. In its
+characteristics of the reflectivity, which is visually expressed in distinct colors of the objects \cite{Richardson06}\label{Richardson06}. In its
 turn, reflectivity of the sediments is affected by the water optical properties and content. For
@@
 available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
-
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\centering
-	\includegraphics[scale=0.30]{Fig-14.jpg}
-	\caption{Multispectral vs. hyperspectral band coverage. \\ Source: \cite{Sabins97}\label{Sabins97}. Pages: $<$\pageref{page-20}$>$,  $<$\pageref{page-18}$>$}
-	\label{fig:2.5}
-\end{wrapfigure}
-
 Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial resolution of 10 m (Fig.\ref{fig:2.4}), covering covers a surface area of 3600 km² (60*60 km swath), 26-day
@@
 	\centering
-	\includegraphics[scale=0.15]{Fig-15.jpg}
-	\caption{Reflectance spectra of \textit{Thalassia} seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
-		reflectance. Source:\cite{Yang10}\label{Yang10}. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-54}$>$}
-	\label{fig:2.6}
+	\includegraphics[scale=0.30]{Fig-14.jpg}
+	\caption{Multispectral vs. hyperspectral band coverage. \\ Source: \cite{Sabins97}\label{Sabins97}. Pages: $<$\pageref{page-20}$>$,  $<$\pageref{page-18}$>$}
+	\label{fig:2.5}
 \end{wrapfigure}
@@
 subsystems for the \ac{VNIR} with 15 m resolution, shortwave infrared \ac{SWIR}, and
-\ac{TIR} wavelength regions \cite{Kalinowski04}\label{Kalinowski04}. For each channel there is separate
-on board calibration \ac{OBC} system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}. 
+\ac{TIR} wavelength regions \cite{Kalinowski04}\label{Kalinowski04}.
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\centering
+	\includegraphics[scale=0.15]{Fig-15.jpg}
+	\caption{Reflectance spectra of \textit{Thalassia} seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
+		reflectance. Source:\cite{Yang10}\label{Yang10}. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-54}$>$}
+	\label{fig:2.6}
+\end{wrapfigure}
+
+For each channel there is separate \ac{OBC} system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}. 
 This makes \ac{ASTER}\label{page-19} imagery especially suitable for the application towards detailed
@@
 To estimate and further analyze particularities of the \textit{P.oceanica} remote sensing reflectance spectra over regions with different bathymetry (0.5, 1.5 and 2.5 meters column depth) we used \ac{WASI} \ac{RTM} software, which incorporates measurements of the seafloor reflectance and water column optical properties.\label{page-22}
+\pagebreak
 
@@
 field campaign has been carried out during the September-October period 2010. 
-The information about the location of the seagrass (mostly represented by the 	extit{P.oceanica} species) is useful for the
+The information about the location of the seagrass (mostly represented by the \textit{P.oceanica} species) is useful for the
 understanding of the relationship between the spatial distribution of the seagrass and the environment
@@
 \ref{tab:A.2}).
-\\
+These products have been used for the radiance and irradiance measurements in at the \ac{HCMR} at the Institute of Aquaculture, Crete Island, 2009 by Ms Sylvia Noralez using following workflow.
+The spectrometer was adjusted for automatic measurements mode, with measurements taken as fast as possible.
+The spectrometer head was held submerged, and the sampling was controlled by an operator on the surface boat.
+The head of the sensor was pointed downward at an angle of $0\,^{\circ}$ (nadir) in order to capture the spatial discernability in the radiance for the benthic cover types.
+The frame was held at $45\,^{\circ}$ angle in order to keep sensor looking down at $0\,^{\circ}$ (nadir view).
+A waterproof camera was attached to the platform to assist with the identification of the target object  being measured (\cite{Noralez10}\label{Noralez10}).
 The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:A.18}) and 430-650 for the water radiance (Fig.\ref{fig:A.16}). 
@@
 well: while in general providing smaller area coverage than satellites images, the Google Earth
-images can be stitched to the composite maps of the acceptable spatial extent, using script written on Python for the \textit{Google Earth} aerial imagery grabbing process (Fig.\ref{fig:A.1} and Fig.\ref{fig:A.2})\label{gdal} which allows multiple overlapping of single images over the flight paths into one swath, and generates imagery mosaics.
+images can be stitched to the composite maps of the acceptable spatial extent, using script written on Python (Fig.\ref{fig:A.1}) and \ac{GDAL} technologies (Fig.\ref{fig:A.2})\label{gdal} for the \textit{Google Earth} aerial imagery grabbing process (Fig.\ref{fig:A.50}\label{GG-1}) which allows multiple overlapping of single images over the flight paths into one swath, and generates imagery mosaics.
 
@@
 
-\begin{figure}
-	\centering
-	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
-	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
+\begin{figure}[H]
+	\centering
+	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.45\textwidth]{Fig35a.jpg}}
+	\hspace{1ex}
+	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.45\textwidth]{Fig35b.jpg}}
 	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images. Page: $<$\pageref{page-48}$>$}
@@
 has been completed using Erdas Imagine software.
+\pagebreak
 
@@
 	\centering
-	\includegraphics[scale=0.3]{GNU-18-legend.png}
+	\includegraphics[scale=0.35]{GNU-18-legend.png}
 	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.45]{Sensors.png}
-	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths.
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{Sensors.jpg}
+	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths, as stripes shown spectral bands covered by these sensors.
 Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$}
 	\label{fig:4.6}
-\end{wrapfigure}
+\end{figure}
 
@@
 	\begin{center}
-		\includegraphics[scale=0.25]{Fig-37.jpg}
+		\includegraphics[scale=0.27]{Fig-37.jpg}
 		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout}
@@
 of the seagrass distribution are added into a \ac{GIS} dataset for the assessment and spatial analysis.
+\pagebreak
 
@@
 The magnitude of the reflectance maximum slightly varies at single variables between about 8 \% and 12 \% (as on Fig.\ref{fig:A.12}) and is probably related to the individual pigmentation and colour composition of single leaves, their structure and geometric orientation, which naturally causes variations in radiance values.
-\\
 
@@
 	\includegraphics[scale=0.4]{photo-1.jpg}
-	\caption{Google Earth aerial imagery grabbing, Heraklion, the University of Crete}
+	\caption{Google Earth aerial imagery grabbing, Heraklion, the University of Crete. Pages: $<$\pageref{GG-1}$>$, $<$\pageref{GG-2}$>$}
 	\label{fig:A.50}
@@
 in nutrient limitation of seagrass \textit{Posidonia oceanica} growth in the NW Mediterranean}.
-Marine ecology Progress Series.Vol.146: 155-161. \pageref{}
+Marine ecology Progress Series.Vol.146: 155-161. \pageref{Alcoverro1997}
 \bibitem{Amoutzopoulou-Schina05}Amoutzopoulou-Schina H.\index[people]{Amoutzopoulou-Schina, H.} and Haritonidis S.\index[people]{Haritonidis, S.} 2005.
@@
 SVNL083-Larkum.
-\pageref{05b}
+\pageref{Dekker05b}
 \bibitem{DiCarlo04}

### 2011-02-14 MScDiplom_Lemenkova.tex
Comments 21 and 24 of Valentijn: updated 'Methods' chapter

@@
 The spectrometer head was held submerged, and the sampling was controlled by an operator on the surface boat.
-The head of the sensor was pointed downward at an angle of $0\,^{\circ}$ (nadir) in order to capture the spatial discernability in the radiance for the benthic cover types.
+The head of the sensor was pointed downward at an angle of $0\,^{\circ}$ (nadir) in order to capture the spatial discernibility in the radiance for the benthic cover types.
 The frame was held at $45\,^{\circ}$ angle in order to keep sensor looking down at $0\,^{\circ}$ (nadir view).
@@
 independent of the individual sets 1-350. \\
-We used several statistical methods to asses data received as a result of measurements. 
+We used several statistical methods (Fig.\ref{fig:4.12})\label{stan} to asses data received as a result of measurements. 
 The Student-t test\index[concepts]{Data!statistical analysis!Student-t testing},
@@
  cover types within data sets, their standard deviation, and the number of data points.
-
-\subsection[Google Earth aerial imagery]{Google Earth aerial imagery for the seagrass mapping}
-Apart from the satellite imagery, the aerial photographs from the Google Earth provide
-a powerful tool for seagrass mapping, because they are important, reliable, detailed
-and up-to-date source of imagery.
-Perhaps, the clearest advantage of the Google Earth imagery is its high resolution (15 m in land areas and lower in the oceans).
-Obtained from the airborne platforms, Google Earth images have general spatial
-resolution of several meters (though varying in different areas), which allows very
-detailed habitat and seafloor types discrimination, comparing with images
-acquired from the space-borne satellite platforms.\\
-The spatial coverage of the Google imagery is lesser comparing to spaceborne images, but this can be solved as
-well: while in general providing smaller area coverage than satellites images, the Google Earth
-images can be stitched to the composite maps of the acceptable spatial extent, using script written on Python (Fig.\ref{fig:A.1}) and \ac{GDAL} technologies (Fig.\ref{fig:A.2})\label{gdal} for the \textit{Google Earth} aerial imagery grabbing process (Fig.\ref{fig:A.50}\label{GG-1}) which allows multiple overlapping of single images over the flight paths into one swath, and generates imagery mosaics.
-
-\subsection{Image processing using Erdas Imagine}\label{page-47}
-The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
-investigate the distribution of the seagrass \textit{P.oceanica} within the research area.
-In the image processing part of the research supervised classification has been applied to the aerial
-and satellite images. Seagrass meadows and other seafloor cover types were evaluated through a
-detailed examination of the imagery. \\
-Seagrass beds are clearly visible in color aerial Google Earth\index[concepts]{Remote sensing!aerial imagery!GoogleEarth}
-photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
-detected using different bands combinations, masked and studied for the estimation of the changes in
-the areas. The classification is based on the properties of the \textit{P.oceanica}, such as brightness,
-colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping\label{page-48}
-includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
-for each photograph by classification a series of polygons characteristic of each of the sea floor
-bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
-seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).
-
-\begin{figure}[H]
-	\centering
-	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.45\textwidth]{Fig35a.jpg}}
-	\hspace{1ex}
-	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.45\textwidth]{Fig35b.jpg}}
-	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images. Page: $<$\pageref{page-48}$>$}
-	\label{fig:38}
-\end{figure}
-
-On the basis of the image processing and classification, applied to the northern coats of Crete, the
-marine bottom types between 10 m depth has been mapped, including the limit depths of the of the
-seagrass meadows provided on the basis of the available and collected field data. The classification
-has been completed using Erdas Imagine software.
 \pagebreak
 
-%%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}\label{sec:4}
-\markright{IV. Results}
-\counterwithin{figure}{section}
-\counterwithin{table}{section}
-
-%
-\subsection{Analysis of spectral signatures}
-The distinguishing spectral signatures for various seafloor types (e.g. seagrass species, coral reefs,
-various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})\label{page-31} exist in well-defined and narrow (10–20 nm)
-wavelength ranges. 
-The values of their spectral reflectance are accepted as constant. The results of
-spectral measurements enable to analyze, whether \textit{P.oceanica} is spectrally distinct from other sea
-floor types with changing environmental conditions, using the differences in their spectral signatures
-on the graphs in a \ac{WASI} software. \\The \ac{WASI}, a
-software tool for analyzing and simulating the most common types of spectra \cite{Gege05}\label{Gege05}, is highly
-suitable for the seagrass spectral analysis.
-There are several environmental characteristics, included in \ac{WASI} interface (Fig.\ref{fig:4.7}), which influence the
-results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
-in water column, water temperature, sun angle, concentration of Gelbstoff (\ac{CDOM}), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small particles \cite{Gege04}\label{Gege04}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.35]{GNU-20-new.jpg}
-	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand.
-Selected measurements sets.
-\href{http://www.gnuplot.info/}{Gnuplot} graph.
-Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-47}$>$, $<$\pageref{dataprep}$>$}
-	\label{fig:28}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{GNU-17.jpg}
-	\caption{Optical properties of the sea water with sediments, measured in aquarium tank.
-Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot}.
-Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-31}$>$}\label{fig:4.2}
-\end{figure}
-
-The backward-scattering coefficient (bb), also included in \ac{WASI}, is a
-fundamental optical property which plays a central role in the ocean-color remote sensing, providing
-the remotely sensed optical signal, as well as suspended particle distributions, water clarity, and
-underwater visibility \cite{Maffione97}\label{Maffione97}. \ac{WASI} enables simulation of backscattering of pure
-water, large and small particles. The values of all these parameters can be redacted and changed
-manually. However, the most important, major factors affecting the \textit{in-situ} conditions are water depth (Fig.\ref{fig:4.9})\label{page-32}
-and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand.
-Seagrass \textit{P.oceanica} can be mapped using remotely acquired spectral imagery, but only if it has distinctive
-reflectance signatures at different depths. Therefore, the depths of the shelf area are the first variable
-condition for this research question. The depths values chosen for the current research are lesser than
-3.5 meters, covering shelf zone, and providing the best environmental conditions for the seagrass
-\textit{P.oceanica}: 0.5, 2.0 and 3.5 meters with an interval of 1.5 meters. 
-The 3.5 m depth limit was chosen based on the analysis of the separability of seagrass reflectance signatures, received by the means of
-previous \textit{in-situ} measurements (year 2009) of the radiance and irradiance of water in Agia Pelagia
-bay, which indicated that \textit{P.oceanica} seagrass could be only well separated within depths of 3.5 m.
-A statistical analysis of \ac{WASI}-simulated spectral reflectance has been used in order to answer the first
-research question: whether the \textit{P.oceanica} spectra is spectrally distinct at varying environmental \textit{in-situ}
-conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
-To answer this research question, different seafloor cover types are discriminated using data of the
-broadband remote sensing.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.35]{GNU-18-legend.png}
-	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
-\href{http://www.gnuplot.info/}{Gnuplot}.
-Two complimentary graphs below show the results of the statistical analysis.\\ Pages: $<$\pageref{page-28}$>$, $<$\pageref{dataprep}$>$, $<$\pageref{page-31}$>$}
-	\label{fig:4.4}
-\end{figure}
-
-The results enable to study reflectance properties of the seagrass and other
-seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
-remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}, Fig.\ref{fig:4.6W}).\label{page-33}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5]{Sensors.jpg}
-	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths, as stripes shown spectral bands covered by these sensors.
-Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$}
-	\label{fig:4.6}
-\end{figure}
-
-In order to focus on the factors of primary importance, other and less influencing factors are excluded,
-i.e. sun angle, concentration of the suspected particles in the water column, content of gelbstoff, etc.
-For these factors default values of \ac{WASI} are accepted.
-Under normal conditions by independent water colour sampling, values of the remote-sensing
-reflectance can vary by 12 - 24 \% \cite{Toole00}\label{Toole00}, and these variations in the radiometric
-determinations are mainly caused by the variety of the environmental factors.
-Different factors influence colour and spectral reflectance of water, among which different bottom
-depths, concentration of suspended particles in water column, water temperature, sun angle,
-concentration of \ac{CDOM}, concentration of phytoplankton,
-aerosol scattering, exponent of backscattering by small particles, cloudiness, viewing geometry and
-wind speed (which is, however, not the major source of uncertainty). All these environmental
-components increase the absorption and scattering of light which, in its turn, results in a complex
-relationship between their concentrations and the radiance of water that finally influence its spectral
-reflectance.
-
-\subsubsection{Spectral simulation of aquatic objects}\index[concepts]{Data!processing}\label{page-35}
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5, angle=90]{UML_StatAn.jpg}
+	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML}).
+Page: $<$\pageref{page-40}$>$, $<$\pageref{stan}$>$}
+	\label{fig:4.12}
+\end{figure}
+
+\subsection{Spectral simulation of aquatic objects}\index[concepts]{Data!processing}\label{page-35}
 The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
@@
 
-\paragraph{Model parameters: depth and bottom cover fraction}
+\subsubsection{Model parameters: depth and bottom cover fraction}
 Although seagrass \textit{P.oceanica} can be found until depth limits down to 40 m depth \cite{DenHartog70}\label{DenHartog70}, the most preferable limits of its distribution in the Mediterranean Sea are shallow waters until
@@
 
+\subsection[Google Earth aerial imagery]{Google Earth aerial imagery for the seagrass mapping}
+Apart from the satellite imagery, the aerial photographs from the Google Earth provide
+a powerful tool for seagrass mapping, because they are important, reliable, detailed
+and up-to-date source of imagery.
+Perhaps, the clearest advantage of the Google Earth imagery is its high resolution (15 m in land areas and lower in the oceans).
+Obtained from the airborne platforms, Google Earth images have general spatial
+resolution of several meters (though varying in different areas), which allows very
+detailed habitat and seafloor types discrimination, comparing with images
+acquired from the space-borne satellite platforms.\\
+The spatial coverage of the Google imagery is lesser comparing to spaceborne images, but this can be solved as
+well: while in general providing smaller area coverage than satellites images, the Google Earth
+images can be stitched to the composite maps of the acceptable spatial extent, using script written on Python (Fig.\ref{fig:A.1}) and \ac{GDAL} technologies (Fig.\ref{fig:A.2})\label{gdal} for the \textit{Google Earth} aerial imagery grabbing process (Fig.\ref{fig:A.50}\label{GG-1}) which allows multiple overlapping of single images over the flight paths into one swath, and generates imagery mosaics.
+
+\subsection{Image processing using Erdas Imagine}\label{page-47}
+The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
+investigate the distribution of the seagrass \textit{P.oceanica} within the research area.
+In the image processing part of the research supervised classification has been applied to the aerial
+and satellite images. Seagrass meadows and other seafloor cover types were evaluated through a
+detailed examination of the imagery. \\
+Seagrass beds are clearly visible in color aerial Google Earth\index[concepts]{Remote sensing!aerial imagery!GoogleEarth}
+photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
+detected using different bands combinations, masked and studied for the estimation of the changes in
+the areas. The classification is based on the properties of the \textit{P.oceanica}, such as brightness,
+colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping\label{page-48}
+includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
+for each photograph by classification a series of polygons characteristic of each of the sea floor
+bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
+seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).
+
+\begin{figure}[H]
+	\centering
+	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.45\textwidth]{Fig35a.jpg}}
+	\hspace{1ex}
+	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.45\textwidth]{Fig35b.jpg}}
+	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images. Page: $<$\pageref{page-48}$>$}
+	\label{fig:38}
+\end{figure}
+
+On the basis of the image processing and classification, applied to the northern coats of Crete, the
+marine bottom types between 10 m depth has been mapped, including the limit depths of the of the
+seagrass meadows provided on the basis of the available and collected field data. The classification
+has been completed using Erdas Imagine software.
+\pagebreak
+
+%%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}\label{sec:4}
+\markright{IV. Results}
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+
+%
+\subsection{Analysis of spectral signatures}
+The distinguishing spectral signatures for various seafloor types (e.g. seagrass species, coral reefs,
+various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})\label{page-31} exist in well-defined and narrow (10–20 nm)
+wavelength ranges. 
+The values of their spectral reflectance are accepted as constant. The results of
+spectral measurements enable to analyze, whether \textit{P.oceanica} is spectrally distinct from other sea
+floor types with changing environmental conditions, using the differences in their spectral signatures
+on the graphs in a \ac{WASI} software. \\The \ac{WASI}, a
+software tool for analyzing and simulating the most common types of spectra \cite{Gege05}\label{Gege05}, is highly
+suitable for the seagrass spectral analysis.
+There are several environmental characteristics, included in \ac{WASI} interface (Fig.\ref{fig:4.7}), which influence the
+results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
+in water column, water temperature, sun angle, concentration of Gelbstoff (\ac{CDOM}), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small particles \cite{Gege04}\label{Gege04}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.35]{GNU-20-new.jpg}
+	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand.
+Selected measurements sets.
+\href{http://www.gnuplot.info/}{Gnuplot} graph.
+Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-47}$>$, $<$\pageref{dataprep}$>$}
+	\label{fig:28}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{GNU-17.jpg}
+	\caption{Optical properties of the sea water with sediments, measured in aquarium tank.
+Agia Pelagia district, Crete.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-31}$>$}\label{fig:4.2}
+\end{figure}
+
+The backward-scattering coefficient (bb), also included in \ac{WASI}, is a
+fundamental optical property which plays a central role in the ocean-color remote sensing, providing
+the remotely sensed optical signal, as well as suspended particle distributions, water clarity, and
+underwater visibility \cite{Maffione97}\label{Maffione97}. \ac{WASI} enables simulation of backscattering of pure
+water, large and small particles. The values of all these parameters can be redacted and changed
+manually. However, the most important, major factors affecting the \textit{in-situ} conditions are water depth (Fig.\ref{fig:4.9})\label{page-32}
+and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand.
+Seagrass \textit{P.oceanica} can be mapped using remotely acquired spectral imagery, but only if it has distinctive
+reflectance signatures at different depths. Therefore, the depths of the shelf area are the first variable
+condition for this research question. The depths values chosen for the current research are lesser than
+3.5 meters, covering shelf zone, and providing the best environmental conditions for the seagrass
+\textit{P.oceanica}: 0.5, 2.0 and 3.5 meters with an interval of 1.5 meters. 
+The 3.5 m depth limit was chosen based on the analysis of the separability of seagrass reflectance signatures, received by the means of
+previous \textit{in-situ} measurements (year 2009) of the radiance and irradiance of water in Agia Pelagia
+bay, which indicated that \textit{P.oceanica} seagrass could be only well separated within depths of 3.5 m.
+A statistical analysis of \ac{WASI}-simulated spectral reflectance has been used in order to answer the first
+research question: whether the \textit{P.oceanica} spectra is spectrally distinct at varying environmental \textit{in-situ}
+conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
+To answer this research question, different seafloor cover types are discriminated using data of the
+broadband remote sensing.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.35]{GNU-18-legend.png}
+	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete.
+\href{http://www.gnuplot.info/}{Gnuplot}.
+Two complimentary graphs below show the results of the statistical analysis.\\ Pages: $<$\pageref{page-28}$>$, $<$\pageref{dataprep}$>$, $<$\pageref{page-31}$>$}
+	\label{fig:4.4}
+\end{figure}
+
+The results enable to study reflectance properties of the seagrass and other
+seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
+remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}, Fig.\ref{fig:4.6W}).\label{page-33}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.5]{Sensors.jpg}
+	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths, as stripes shown spectral bands covered by these sensors.
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$}
+	\label{fig:4.6}
+\end{figure}
+
+In order to focus on the factors of primary importance, other and less influencing factors are excluded,
+i.e. sun angle, concentration of the suspected particles in the water column, content of gelbstoff, etc.
+For these factors default values of \ac{WASI} are accepted.
+Under normal conditions by independent water colour sampling, values of the remote-sensing
+reflectance can vary by 12 - 24 \% \cite{Toole00}\label{Toole00}, and these variations in the radiometric
+determinations are mainly caused by the variety of the environmental factors.
+Different factors influence colour and spectral reflectance of water, among which different bottom
+depths, concentration of suspended particles in water column, water temperature, sun angle,
+concentration of \ac{CDOM}, concentration of phytoplankton,
+aerosol scattering, exponent of backscattering by small particles, cloudiness, viewing geometry and
+wind speed (which is, however, not the major source of uncertainty). All these environmental
+components increase the absorption and scattering of light which, in its turn, results in a complex
+relationship between their concentrations and the radiance of water that finally influence its spectral
+reflectance.
+
+\subsection{Spectral discrimination of \textit{P.oceanica}}
+
 \begin{figure}[H]
@@
 could still be discriminated are lesser than 2.5 meters.
-
-\paragraph{Spectral discrimination of \textit{P.oceanica}}\label{page-39}
 The patches of white sandy bottoms of the seafloor, studied in the fieldwork in Ligaria beach, are
-much brighter than mattes of \textit{P.oceanica} (Fig.\ref{fig:A.6}, \ref{fig:A.5}): seafloor types in Ligaria), which can be clearly seen
+much brighter than mattes of \textit{P.oceanica} (Fig.\ref{fig:A.6}, \ref{fig:A.5})\label{page-39}: seafloor types in Ligaria), which can be clearly seen
 at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:A.35}) and that of \textit{P.oceanica} (Fig.\ref{fig:A.28}).
@@
 The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{GRETL} software.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.5, angle=90]{UML_StatAn.jpg}
-	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML}).
-Page: $<$\pageref{page-40}$>$}
-	\label{fig:4.12}
-\end{figure}
-
 The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made.

### 2011-02-14 MScDiplom_Lemenkova.tex
Comments of Valentijn: 31, 33. Updated 'Methods', 'Results'

@@
 \begin{table}[htbp]\footnotesize
-	\caption{Characteristics of selected ocean-colour sensors. Page: $<$\pageref{page-47}$>$}
+	\caption{Characteristics of selected ocean-colour sensors. Pages: $<$\pageref{page-intmap}$>$, $<$\pageref{page-47}$>$}
 	\begin{center}
@@
 Technically, \ac{WASI} need for system memory is minimised and the installation is simple and easy.\\
-We included measurements of the \textit{Thalassia} seagrass
-(measured at Southern Chinese Sea by C.Yang and D.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectance of
-different seagrass species under various environmental conditions.
+To create sensor-specific reflectance spectra using \ac{WASI}, spectral responses of \ac{MERIS}, \ac{MODIS}, \ac{CZCS} and \ac{SeaWiFS} sensors were applied for simulation of seagrass spectral signatures in full-resolution spectra.\\
+We also included measurements of other seagrass species - \textit{Thalassia} seagrass
+(measured at Southern Chinese Sea by C.Yang and D.Yang \cite{Yang10} \label{Yang10}) in order to compare spectral reflectance of
+different seagrasses under various environmental conditions.
 The most suitable wavebands for the seagrass monitoring usually lay between 400 and 700 nm,
@@
 types as the different sensors would percept them: \ac{MODIS}\index[concepts]{Remote sensing!multispectral imagery!MODIS}, \ac{ASTER}\index[concepts]{Remote sensing!multispectral imagery!ASTER}, \ac{MERIS}\index[concepts]{Remote sensing!multispectral imagery!MERIS}, \ac{SeaWiFS}\index[concepts]{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index[concepts]{Remote sensing!multispectral imagery!CZCS}.\\
-These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
+These sensors vary in technical characteristics and therefore, have different spectral sensitivity, 
+which we briefly illustrated in a small summary table (Tab.\ref{tab:1})\label{page-intmap}
 These models simulate spectral views of the chosen sensors, how these sensors will ``see" seagrass (Fig.\ref{fig:28}) and sand as
 pixels, with accepted default atmosphere and water column effects (given by \ac{WASI} software). In such
-a way we defined an empirical upper limit to the discriminative potential of these sensors. To create
-these sensor-specific reflectance spectra, spectral responses of each of these sensors were applied for
-the spectral signature of seafloor cover types in full-resolution spectra.
+a way we defined an empirical upper limit to the discriminative potential of these sensors.
 

### 2011-02-14 MScDiplom_Lemenkova.tex
Comment 22 of Valentijn: Fig.4.5: added MERIS spectral bands as stripes

@@
 	\centering
-	\subfloat [depth of 0.5 m]{\label{fig:32a}\includegraphics[width=0.4\textwidth]{Fig-32a.jpg}}
-	\subfloat [depth of 2.0 m]{\label{fig:32b}\includegraphics[width=0.4\textwidth]{Fig-32b.jpg}}\\
-	\subfloat [depth of 3.5 m]{\label{fig:32c}\includegraphics[width=0.4\textwidth]{Fig-32c.jpg}}
-	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d.jpg}}
-	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths, simulated for broadband sensors. Pages: $<$\pageref{page-37}$>$, $<$\pageref{page-32}$>$}
+	\subfloat [depth of 0.5 m]{\label{fig:32a}\includegraphics[width=0.4\textwidth]{Fig-32a-SB.jpg}}
+	\subfloat [depth of 2.0 m]{\label{fig:32b}\includegraphics[width=0.4\textwidth]{Fig-32b-SB.jpg}}\\
+	\subfloat [depth of 3.5 m]{\label{fig:32c}\includegraphics[width=0.4\textwidth]{Fig-32c-SB.jpg}}
+	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d-SB.jpg}}
+	\caption{Simulated remote sensing reflectance of \textit{P.oceanica} at various depths.
+Pages: $<$\pageref{page-37}$>$, $<$\pageref{page-32}$>$.\\ Vertical stripes show coverage of spectral bands by MERIS sensor}
 	\label{fig:4.9}

### 2011-02-16 MScDiplom_Lemenkova.tex
Comment 35 of Valentijn: Reported Classification accuracy and printed Kappa confusion matrix. Added two paragraphs: 'Data inregration' and 'Accuracy assessment'

@@
 \usepackage{color}
-\usepackage[stable]{footmisc }
+\usepackage[stable]{footmisc}
 \usepackage{setspace}
@@
 \usepackage{longtable}
+\usepackage{rotating}
 \usepackage[svgnames,table]{xcolor}
@@
 
-\subsection{Integrated mapping of seagrass}
+\subsection{GIS mapping of seagrass}
 
 \index[concepts]{Data!integration}
+\paragraph{Data integration}
 The integrated approach used in this research work has high potential as a means to monitor changes
@@
 of the seagrass distribution are added into a \ac{GIS} dataset for the assessment and spatial analysis.
+\paragraph{Accuracy assessment}\label{page-50acc}
+We prepared the error confusion matrices (Fig.\ref{tab:kappa-1} and Fig.\ref{tab:kappa-2}) using $\kappa$ (Kappa-) statistics to assess and evaluate accuracy of the classification.
+The accuracy assessment has been done using Erdas Imaging function Classifier/Accuracy Assessment.
+In the Accuracy Assessment viewer we have chosen utility Edit in order to generate random points throughout the classified image, and then chosen the Create or add random points dialog.
+After the points were generated, we entered the class values for the reference points from the supervised classification.
+In order to perform a proper accuracy assessment we needed about 300 points so that we defined the number of points for the selection.
+From the option parameters for points distribution we have chosen "stratified random" as better representing the variables.
+Then we evaluated the location of the points and determined their class value, which was done with ground truth points from Google Earth aerial photos for various data and seafloor imagery and with reference to the signature file, to find the numeric value that has been assigned to each class previously.
+Finally, to print the errors matrix, we selected Accuracy Assessment viewer again and chosen there Report/ Options with turned on Error Matrix, Kappa Statistics and Accuracy Total.
+The error matrix is just comparing reference points of various seafloor cover types to the classified points (i.e.
+seagrass, carbonate sand, various land cover types on the coast).
+The Kappa coefficient takes into account chance agreement and thus, shows the reduction in errors generated by a classification process compared with the errors which could be received by a completely random classification – in other words, it evaluates quality of the classification.
+The results of the classification can be seen on the Fig.
+(see appendices).
+The overall map accuracy is 68\%, which means that 68\% of the pixels are classified to the correctly chosen seafloor cover type.
 \pagebreak
@@
 performed using underwater videometric measurements, aerial and satellite imagery.
+\pagebreak
 
@@
 This research is a contribution to the development of the methodology of seagrass mapping with the aim of the environmental monitoring, and a case study of \textit{P.oceanica}, an endemic Mediterranean seagrass species, dominating in underwater landscapes and ecosystems along the coasts of Crete Island.
+\pagebreak
 
@@
 		\includegraphics[scale=0.25]{Fig-39.jpg}
-		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project.
+		\caption{Results of the image unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project.
 Page: $<$\pageref{page-48}$>$}
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[H]
 	\begin{center}
 		\includegraphics[scale=0.25]{Fig-36.jpg}
-		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete.
-Page: $<$\pageref{page-48}$>$}
+		\caption{Results of the image supervised classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete.
+Pages: $<$\pageref{page-48}$>$, $<$\pageref{page-liii}$>$}
 		\label{fig:39}
@@
 \pagebreak
+
+\subsection{Accuracy assessment}
+
+\begin{table}[H]\footnotesize\label{page-liii}
+	\caption{Confusion Matrix-1, between the classified Google Earth aerial image and fieldwork data, for Fig.\ref{fig:42}. Page: $<$\pageref{page-50acc}$>$}\label{tab:kappa-1}
+	\centering
+	\rowcolors{1}{Honeydew}{White}
+	 \begin{tabular}{| p{1.7cm} | p{0.4cm}| p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}| p{0.4cm}| p{0.4cm}|| p{0.4cm}| p{0.4cm}|}
+	    \hline
+		     \textbf{Correctly classified} & \begin{sideways}\textbf{Rocky bottom}\end{sideways} &  \begin{sideways}\textbf{Shelf, $<3$m}\end{sideways} &  \begin{sideways}\textbf{Healthy seagrass}\end{sideways} &  \begin{sideways}\textbf{Seagrass \textit{P.oceanica}}\end{sideways} &  \begin{sideways}\textbf{Carbonate sand}\end{sideways} &  \begin{sideways}\textbf{Shelf waters, 0-3m}\end{sideways} &  \begin{sideways}\textbf{Shelf waters, 3-7m}\end{sideways} & \begin{sideways}\textbf{Deep waters, $>7$m}\end{sideways} &  \begin{sideways}\textbf{Field: corn, greens}\end{sideways} & \begin{sideways}\textbf{Roads: asphalt+ground}\end{sideways} & \begin{sideways}\textbf{Seagrass, other$>3$m}\end{sideways} & \begin{sideways}\textbf{Ground}\end{sideways}& \begin{sideways}\textbf{Buildings (roofs)}\end{sideways} &\begin{sideways}\textbf{Trees}\end{sideways} & \begin{sideways}\textbf{Bushes}\end{sideways} & \begin{sideways}\textbf{Total}\end{sideways}& \begin{sideways}\textbf{$\kappa$ producer accuracy}\end{sideways} \\ \hline \hline
+		\textbf{Rocky bottom}& \cellcolor{GreenYellow}\textbf{14} & 2 & 0 & 0 & 3 & 0 & 0 & 0 & 0 & 1 & 0 & 2 & 0 & 1 & 1 & \cellcolor{LightGoldenrodYellow}24 & 0.58 \\ \hline
+		\textbf{Shelf, $<3$m} & 0 & \cellcolor{GreenYellow}\textbf{12} & 0 & 3 & 1 & 2 & 0& 0 & 1 & 0 & 0 & 0 & 1 & 0 & 0 & \cellcolor{LightGoldenrodYellow}20 & 0.60 \\ \hline
+		\textbf{Healthy seagrass} & 0 & 1 & \cellcolor{GreenYellow}\textbf{10} & 3 & 2 & 0 & 0 & 0 & 1 & 0 & 2 & 0 & 0 & 0 & 0 & \cellcolor{LightGoldenrodYellow}19 & 0.52 \\ \hline
+             	\textbf{Seagrass \textit{P.oceanica}} & 0 & 1 & 1 & \cellcolor{GreenYellow}\textbf{13} & 0 & 1 & 0 & 0 & 1 & 0 & 2 & 0 & 0 & 1 & 1 & \cellcolor{LightGoldenrodYellow}21 & 0.62 \\ \hline
+		\textbf{Carbonate sand} & 2 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{16} & 0 & 0 & 0 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & \cellcolor{LightGoldenrodYellow}20 & 0.80 \\ \hline
+		\textbf{Shelf waters, 0-3m} & 0 & 3 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{9} & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & \cellcolor{LightGoldenrodYellow}14 & 0.64 \\ \hline
+		\textbf{Shelf waters, 3-7m} & 1 & 2 & 0 & 0 & 0 & 3 & \cellcolor{GreenYellow}\textbf{8} & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & \cellcolor{LightGoldenrodYellow}15 & 0.53 \\ \hline
+		\textbf{Deep waters, $>7$m} & 1 & 1 & 0 & 0 & 1 & 2 & 3 & \cellcolor{GreenYellow}\textbf{16} & 0 & 0 & 1 & 0 & 0 & 0 & 0 & \cellcolor{LightGoldenrodYellow}25 & 0.64 \\ \hline
+		\textbf{Fields: corn, greens} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{14} & 0 & 0 & 0 & 0 & 3 & 4 & \cellcolor{LightGoldenrodYellow}21 & 0.66 \\ \hline
+		\textbf{Roads: asphalt+ground} & 1 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{5} & 2 & 0 & 0 & 0 & 0 & \cellcolor{LightGoldenrodYellow}9 & 0.55 \\ \hline
+		\textbf{Seagrass, other$>3$m} & 0 & 0 & 0 & 3 & 0 & 0 & 0 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{24} & 2 & 0 & 1 & 1 & \cellcolor{LightGoldenrodYellow}31 & 0.77 \\ \hline		
+		\textbf{Ground} & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{11} & 3 & 1 & 1 & \cellcolor{LightGoldenrodYellow}17 & 0.65 \\ \hline
+		\textbf{Buildings (roofs)} & 0 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{9} & 2 & 3 &\cellcolor{LightGoldenrodYellow} 16 & 0.56 \\ \hline
+		\textbf{Trees} & 0 & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & \cellcolor{GreenYellow}\textbf{7} & 3 & \cellcolor{LightGoldenrodYellow}12 & 0.58 \\ \hline
+		\textbf{Bushes} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 2 & \cellcolor{GreenYellow}\textbf{4}& \cellcolor{LightGoldenrodYellow}6 & 0.66 \\ \hline	\hline	
+		\rowcolor{LightGoldenrodYellow}\textbf{Total} & 20 & 22 & 13 & 22 & 26 & 17 & 13 & 17 & 17 & 6 & 31 & 17 & 13 & 18 & 18 & \cellcolor{LightGoldenrodYellow}\textbf{270} & \\ \hline
+		\textbf{$\kappa$ user`s accuracy} & 0.70 & 0.55 & 0.77 & 0.59 & 0.62 & 0.53 & 0.62 & 0.94 & 0.82 & 0.83 & 0.77 & 0.65 & 0.69 & 0.39 & 0.22 & - & \cellcolor{Chartreuse}\textbf{0.64} \\ \hline
+	\end{tabular}	
+\end{table}
+\normalsize
+Overall Kappa ($\kappa$) accuracy is calculated using the formula:  $\sum{A/N}$, where A is number of correctly mapped points (172) and N is the total number of points (270).
+Thus, according to the results we received overall accuracy= 172/270= 0.6370, which is 64\%.
+Overall $\kappa$ accuracy for unsupervised classification =\textbf{64\%}.
+\\
+ User’s accuracy (Reliability of classes) varies between 0.22 and 0.94 depending on class, which proves that supervised classification (see next table: Tab.\ref{tab:kappa-2} ) has better results for seagrass mapping than the unsupervsied classification.
+Producer accuracy lies in interval between 0.52-0.77 according to class as well. 
+\pagebreak
+
+\begin{table}[H]\footnotesize\label{page-liii}\label{tab:matrix-2}
+	\caption{Confusion Matrix-2, between the classified Google Earth aerial image and fieldwork data, for Fig.\ref{fig:39}. Page: $<$\pageref{page-50acc}$>$}\label{tab:kappa-2}
+	\centering
+	\rowcolors{1}{Lavender}{White}
+	 \begin{tabular}{| p{1.7cm} | p{0.7cm}| p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|| p{0.7cm}||p{0.7cm}|}
+	    \hline
+		     \begin{sideways}\textbf{Correctly classified}\end{sideways} & \begin{sideways}\textbf{Seagrass-1}\end{sideways} &  \begin{sideways}\textbf{Roads}\end{sideways} &  \begin{sideways}\textbf{Fields}\end{sideways} &  \begin{sideways}\textbf{Earth}\end{sideways} &  \begin{sideways}\textbf{Forest}\end{sideways} &  \begin{sideways}\textbf{Buildings}\end{sideways} &  \begin{sideways}\textbf{Seagrass-2}\end{sideways} & \begin{sideways}\textbf{Seagrass-3}\end{sideways} &  \begin{sideways}\textbf{Terrace}\end{sideways} & \begin{sideways}\textbf{Seagrass-4}\end{sideways} &  \begin{sideways}\textbf{Water}\end{sideways} & \cellcolor{LightBlue}\begin{sideways}\textbf{Total}\end{sideways} & \begin{sideways}\textbf{$\kappa$ producer accuracy}\end{sideways} \\ \hline \hline
+		\textbf{Seagrass \textit{P.oceanica}}& \cellcolor{LightSkyBlue}\textbf{22} & 0 & 0 & 0 & 0 & 0 & 2 & 1 & 0 & 1 & 1 & \cellcolor{LightBlue}27 & 0.81 \\ \hline
+		\textbf{Roads} & 0 & \cellcolor{LightSkyBlue}\textbf{18} & 0 & 3 & 1 & 2 & 0& 0 & 0 & 0 & 0 & \cellcolor{LightBlue}24 & 0.75 \\ \hline
+		\textbf{Fields} & 0 & 1 & \cellcolor{LightSkyBlue}\textbf{10} & 0 & 3 & 0 & 0 & 0 & 1 & 0 & 0 & \cellcolor{LightBlue}15 & 0.66 \\ \hline
+             	\textbf{Earth} & 0 & 1 & 1 & \cellcolor{LightSkyBlue}\textbf{13} & 0 & 1 & 0 & 0 & 1 & 0 & 0 & \cellcolor{LightBlue}17 & 0.76 \\ \hline
+		\textbf{Forest} & 0 & 1 & 1 & 1 & \cellcolor{LightSkyBlue}\textbf{9} & 0 & 0 & 0 & 1 & 0 & 0 & \cellcolor{LightBlue}13 & 0.69 \\ \hline
+		\textbf{Buildings} & 0 & 0 & 2 & 0 & 1 & \cellcolor{LightSkyBlue}\textbf{24} & 0 & 0 & 1 & 0 & 0 & \cellcolor{LightBlue}28 & 0.85 \\ \hline
+		\textbf{Seagrass-2} & 3 & 0 & 0 & 0 & 0 & 0 & \cellcolor{LightSkyBlue}\textbf{33} & 2 & 0 & 2 & 2 & \cellcolor{LightBlue}42 & 0.78 \\ \hline
+		\textbf{Seagrass-3} & 1 & 0 & 0 & 0 & 0 & 0 & 1 & \cellcolor{LightSkyBlue}\textbf{27} & 0 & 1 & 1 & \cellcolor{LightBlue}31 & 0.87 \\ \hline
+		\textbf{Terrace} & 0 & 2 & 3 & 1 & 0 & 1 & 0 & 0 & \cellcolor{LightSkyBlue}\textbf{19} & 0 & 0 & \cellcolor{LightBlue}26 & 0.73 \\ \hline
+		\textbf{Seagrass-4} & 1 & 0 & 0 & 0 & 0 & 0 & 2 & 2 & 0 & \cellcolor{LightSkyBlue}\textbf{37} & 1 & \cellcolor{LightBlue}43 & 0.86 \\ \hline
+		\textbf{Water} & 2 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & \cellcolor{LightSkyBlue}\textbf{14} & \cellcolor{LightBlue}19 & 0.74 \\ \hline\hline
+		\rowcolor{LightBlue}\textbf{Total} & 29 & 23 & 17 & 18 & 14 & 28 & 39 & 33 & 23 & 42 & \cellcolor{LightBlue}19 & \textbf{285} & - \\ \hline
+		\textbf{$\kappa$ user`s accuracy} & 0.76 & 0.78 & 0.59 & 0.72 & 0.64 & 0.86 & 0.85 & 0.82 & 0.83 & 0.88 & 0.74 & - & \cellcolor{MediumTurquoise}\textbf{0.72} \\ \hline
+	\end{tabular}	
+\end{table}
+\normalsize
+Overall Kappa ($\kappa$) accuracy is calculated using the formula:  $\sum{A/N}$, where A is number of correctly mapped points (226) and N is the total number of points (285).
+Thus, according to the results we received overall accuracy= 226/285= 0.79298, which is 72\%.
+Overall $\kappa$ accuracy for supervised classification=\textbf{72\%}.
+\\  User’s accuracy (Reliability of classes) varies between 0.59 and 0.88 depending on class.
+\\
+Producer accuracy lies in interval between 0.66-0.87 according to class as well.
+
+\pagebreak
+\newpage
 

### 2011-02-16 MScDiplom_Lemenkova.tex
finished references of Figures (links to text); updated Tables

@@
 \hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue,pdfkeywords={P.oceanica, seagrass, spectral reflectance, remote sensing, WASI, RTM, Google Earth, mapping, Crete}}
+%\hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=black, citecolor=black, filecolor=black, urlcolor=black,pdfkeywords={P.oceanica, seagrass, spectral reflectance, remote sensing, WASI, RTM, Google Earth, mapping, Crete}}
 
@@
 \begin{itemize}
-	\item Name Examiner 1 (Chair) \hfill{Prof. Dr. \textit{Andrew K. Skidmore}\index[people]{Skidmore A.K.} (ITC, Department of Natural Resource)}
-	\item Name Examiner 2 \hfill{Ms. Ir. \textit{Louise M. van Leeuwen}\index[people]{Leeuwen, L. M. van} (ITC, Department of Natural Resource)}
-	\item External Examiner \hfill{Dr. {\textit{Ma\l{}gorzata Roge-Wi\'sniewska}\index[people]{Roge-Wi\'sniewska, M.}} (University of Warsaw, Faculty of Geography and Regional Studies)}
+	\item Chair: Prof. Dr. \textit{Andrew K. Skidmore}\index[people]{Skidmore A.K.}
+	\item External Examiner: Dr. {\textit{Ma\l{}gorzata Roge-Wi\'sniewska}\index[people]{Roge-Wi\'sniewska, M.}}
+	\item Internal Examiner: Ms. Ir. \textit{Louise M. van Leeuwen}\index[people]{Leeuwen, L. M. van}
 \end{itemize}
 
-\vspace{2em}
+\vspace{1em}
 
@@
 \begin{itemize}
-	\item[* ] Mr. \textit{Valentijn Venus}\index[people]{Venus, V.}  \hfill{(University of Twente, ITC, Department of Natural Resource)}
-	\item[* ] Dr. \textit{Albertus G. Toxopeus}\index[people]{Toxopeus, A.G.}  \hfill{(University of Twente, ITC, Department of Natural Resource)}
+	\item[* ] First supervisor: Mr. \textit{Valentijn Venus}\index[people]{Venus, V.} 
+	\item[* ] Second supervisor: Dr. \textit{Albertus G. Toxopeus}\index[people]{Toxopeus, A.G.}  
 \end{itemize}
@@
 	\acro{OBC}{On Board Calibration}
+	\acro{OLS}{\href{http://en.wikipedia.org/wiki/Ordinary_least_squares}{Ordinary Least Squares}}
 	\acro{Olympus ST}{\href{http://www.olympusamerica.com/}{Olympus Stylus Tough}}
@@
 quadrant frame, single point markers, \ac{GPS}, laptops with installed software (\ac{GIS}), \ac{SCUBA} equipment, \ac{iPAQ} and other auxiliary tools\label{page-15} (Fig.\ref{fig:3.3}).
-The seagrass sampling is taken on the regular way with observation points covering the study area with normal distribution.
+The seagrass sampling is taken on the regular way with observation points covering the study area with normal distribution (Fig.\ref{fig:A.20}, Fig.\ref{fig:A.19}\label{page-norm}).
 During the measurement process, the vertical photograph of the measurements frame is taken, and the
@@
 accurate way, detect spatial distribution of the single seagrass mattes to be properly identified,
-without location bias.
+without location bias (Fig.\ref{fig:A.7})\label{page-A12}.
 
@@
 most determining factors for the seagrass locations. Measurements of depths were performed during
-the fieldwork using cord, \ac{iPAQ}, and markers in order to assure that the videometric measurements
+the fieldwork using cord, \ac{iPAQ}, and markers (Fig.\ref{fig:A.55})\label{page-A55} in order to assure that the videometric measurements
 are taken at depths not more than four meters, in the shelf area.
@@
 \end{itemize}
-The available satellite imagery (Fig.\ref{fig:A.38}) \label{page-30}and aerial images are read into the ArcGIS project.
+The available satellite imagery (Fig.\ref{fig:A.38}, Tab.\ref{tab:34}) \label{page-30}and aerial images are read into the ArcGIS project.
 The available broadband and hyperspectral remote sensing data are used for the mapping of the
-seagrass in shelf areas no deeper than 4.0 meters, for the environmental monitoring in order to detect
-the dynamics in changes of the seagrass distribution along Crete during the past 10 years using
-different satellite images for 2000-2010.
+seagrass in shelf areas no deeper than 4.0 meters (Fig.\ref{fig:A.39}, Fig.\ref{fig:A.40}, Fig.\ref{fig:A.41}, Fig.\ref{fig:A.42})\label{page-A45-47}, for the environmental monitoring in order to detect spatial distribution of the seagrass along Crete during the past ten years using
+different satellite images for 2000-2010 (Fig.\ref{fig:24})\label{page-fig37}.
 
@@
 	\includegraphics[scale=0.40]{UML_Dia.jpg}
-	\caption{Work flow for the data acquisition, \href{http://live.gnome.org/Dia}{Dia} display}
+	\caption{Work flow for the data acquisition, \href{http://live.gnome.org/Dia}{Dia} display. Page: $<$\pageref{page-fig37}$>$}
 	\label{fig:24}
@@
 mean of average for each waveband, mean for statistical quartiles for each wavelength and mean for median, respectively. 
-The adjusted averages evaluate mean values, needed for statistical analysis of the datasets. 
+The adjusted averages evaluate mean values of trend, needed for statistical analysis of the datasets (Fig.\ref{fig:A.32}). 
 The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (\nameref{app}). \\
-As a result, we received mean values of spectral reflectance of \textit{P.oceanica} for extreme, average, median and both quartiles,
-independent of the individual sets 1-350. \\
+As a result, we received mean values of spectral reflectance of \textit{P.oceanica} for extreme, average (Fig.\ref{fig:A.34}, Fig.\ref{fig:A.33})\label{page-ave}, median and both quartiles, independent of the individual sets 1-350.
+\\
 We used several statistical methods (Fig.\ref{fig:4.12})\label{stan} to asses data received as a result of measurements. 
@@
 A probability distribution\index[concepts]{Data!statistical analysis!probability distribution}, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed.
-The \ac{LAD}\index[concepts]{Data!statistical analysis!LAD}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}).
+The \ac{LAD}\index[concepts]{Data!statistical analysis!LAD}, a popular optimization technique, was used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12})\label{page-tab12}.
 Non-linear logistic analysis\index[concepts]{Data!statistical analysis!non-linear logistic analysis} is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e.
@@
 As can be seen on Fig.\ref{fig:A.21}\label{A-23}, Fig.\ref{fig:A.22}\label{A-24} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
-The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{GRETL} software.
+The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{GRETL} software (Fig.\ref{tab:28}\label{page-a-c}).
 The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made.
-We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}.\\
-To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectance at several depths: Tab.\ref{tab:5}, Tab.\ref{tab:6}, Tab.\ref{tab:7}.
+We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}, \ac{OLS}: Fig.\ref{tab:15}\label{page-OLS}\\
+To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectance at several depths: Tab.\ref{tab:5}\label{page-reg}, Tab.\ref{tab:6}, Tab.\ref{tab:7}.
 The summary of the values of spectral reflectance properties on various depths is presented on the Tab.\ref{tab:A.7}.
@@
 It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
-To estimate spectral density of a spectral reflectance measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30}), which corresponds to the general spectrum of the observations with representation of a variable quantity.
-Thus, the periodogram (Fig.\ref{fig:A.31}) of the series of observation\index[concepts]{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to exclude (can be seen clearly at the end of the graphs, e.g.
+To estimate spectral density of a spectral reflectance measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30})\label{page-period}, which corresponds to the general spectrum of the observations with representation of a variable quantity.
+Thus, the periodogram (Fig.\ref{fig:A.31})\label{page-A31} of the series of observation\index[concepts]{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to exclude (can be seen clearly at the end of the graphs, e.g.
 Fig.\ref{fig:A.12} )
@@
 The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
-areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1, Q3, shaded areas on graph, Fig.
+areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1 - Q3, Fig.\ref{fig:A.30}\label{page-A30}, and shaded Q-Q areas on graph, Fig.
 \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution (spectral reflectance) for each spectral band.
@@
 		\includegraphics[scale=0.27]{Fig-37.jpg}
-		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout}
+		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout. Page: $<$\pageref{fig4-17}$>$}
 		\label{fig:4.17}
@@
 spectral signatures using \ac{WASI} software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}, Fig.\ref{fig:42}) and Arc \ac{GIS} based mapping.
-The use of \ac{GIS} for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
-enables to analyze environmental changes within seagrass landscapes based on data from various
+The use of \ac{GIS} for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping enables to analyze environmental changes within seagrass landscapes based on data from various
 sources: aerial and satellite images, geographically referenced maps of Crete island and results of
@@
 The final mapping has been supported in ArcGIS through the data exporting, conversion and
-integration of various data in one \ac{GIS}-project. Data collected during the fieldwork, imagery
+integration of various data in one \ac{GIS}-project (Fig.\ref{fig:4.17})\label{fig4-17}. Data collected during the fieldwork, imagery
 of the seagrass distribution are added into a \ac{GIS} dataset for the assessment and spatial analysis.
 \paragraph{Accuracy assessment}\label{page-50acc}
-We prepared the error confusion matrices (Fig.\ref{tab:kappa-1} and Fig.\ref{tab:kappa-2}) using $\kappa$ (Kappa-) statistics to assess and evaluate accuracy of the classification.
+We prepared the error confusion matrices (Fig.\ref{tab:kappa-1} and Fig.\ref{tab:kappa-2}) using $\kappa$ (Kappa) statistics to assess and evaluate accuracy of the classification.
 The accuracy assessment has been done using Erdas Imaging function Classifier/Accuracy Assessment.
@@
 ecosystems in biodiversity and complexity; however, seagrass landscapes show variation in spatial
-patterns over different special scales. The complexity of the landscape of seagrass meadow is shown by the
+patterns over different special scales (Fig.\ref{fig:41})\label{page-fig41}. The complexity of the landscape of seagrass meadow is shown by the
 measure of patches in size and shape, expressed in ratio of patch perimeter to area. The fragmentation
@@
 	\subfloat [Patches (mattes) of the seagrass. Source: \textit{in-situ} videometric measurements]{\label{fig:41b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
-	\caption{Variations in spatial structure of the seagrass landscapes}
+	\caption{Variations in spatial structure of the seagrass landscapes. Page: $<$\pageref{page-fig41}$>$}
 	\label{fig:41}
@@
 Unique feature of seagrass meadows, most characteristic for shallow areas, is dynamics of their
-landscapes. Homogeneous, continuous seagrass meadows can reach in size up to 40 m2 (Fig.\ref{fig:38} a),
+landscapes. Homogeneous, continuous seagrass meadows can reach in size up to $40 m^2$ (Fig.\ref{fig:38} a),
 however they are often interrupted by gaps and channels of open spaces generated by the complex,
@@
 potential mechanism for increasing patch dynamics and morphological plasticity within seagrass
-meadows. Finally, increasing the degree of fragmentation of the landscapes of 	extit{P.oceanica} meadows
+meadows. Finally, increasing the degree of fragmentation of the landscapes of \textit{P.oceanica} meadows
 can be caused by the invasion of alien species, such as \textit{Cymodocea nodosa, Caulerpa prolifera,
@@
 		\includegraphics[scale=0.3]{Fig-3-16.jpg}
-		\caption{Measurement underwater equipment}\label{fig:A.7}		
+		\caption{Measurement underwater equipment.  Page: $<$\pageref{page-A12}$>$}\label{fig:A.7}		
 	\end{center}
@@
 	\caption{Robust estimation of the seawater radiance (measured in aquarium tank, Heraklion): results of the \textit{least absolute deviation} (LAD), Series:V16.
-\href{http://gretl.sourceforge.net/}{Gretl}}
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-tab12}$>$}
 	\begin{center}
@@
 		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15.
-\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\end{center}
-	\label{fig:A.32}
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­.
+Page: $<$\pageref{page-ave}$>$}\label{fig:A.32}
+	\end{center}	
 \end{figure}
@@
 		\includegraphics[scale=0.4]{G-polynom-trend-with.jpg}
-		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\end{center}
-	\label{fig:A.33}
+		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15.
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization.
+Page: $<$\pageref{page-ave}$>$­}\label{fig:A.33}
+	\end{center}	
 \end{figure}
@@
 		\includegraphics[scale=0.4]{G-exp-mov-aver.jpg}
-		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
-	\end{center}
-	\label{fig:A.34}
+		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15.
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization.
+Page: $<$\pageref{page-ave}$>$­}\label{fig:A.34}
+	\end{center}	
 \end{figure}
@@
 		\includegraphics[scale=0.4]{G-gamma-rad.jpg}
-		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}}
+		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank.
+Visualization in \href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-norm}$>$}
 		\label{fig:A.19}
@@
 	\includegraphics[scale=0.4]{G-normal-dist.jpg}
-	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}­}
+	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank.
+Visualization in \href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-norm}$>$­}
 	\label{fig:A.20}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.\ac{SPSS}}
+	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.
+\ac{SPSS}.
+Page: $<$\pageref{page-reg}$>$}
 	\begin{center}
@@
 	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.
-\ac{SPSS}}
+\ac{SPSS}.
+Page: $<$\pageref{page-reg}$>$}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textit{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
+	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textit{P.oceanica}.
+\href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{page-reg}$>$}
 	\begin{center}
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}
+	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3.
+\href{http://www.spss.com/}{SPSS}.
+Page $<$\pageref{page-a-c}$>$}
 	\begin{center}
@@
 		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textit{P.oceanica} (15.X.).
-\href{http://gretl.sourceforge.net/}{Gretl}}
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-OLS}$>$}
 		Model 2: OLS, using observations 326--350\\
@@
 Series 51-75.
-\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­.
+Page: $<$\pageref{page-A30}$>$}
 	\label{fig:A.30}
@@
 	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
-	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
+	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-period}$>$}
 	\label{tab:30}
@@
 		\includegraphics[scale=0.4]{G-period-sand30.jpg}
-		\caption{Plot illustrating periodogram for v30 from measurements of \textit{carbonate sand}, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
+		\caption{Plot illustrating periodogram for v30 from measurements of \textit{carbonate sand}, series 51-75.
+\href{http://gretl.sourceforge.net/}{Gretl} modelling visualization.
+Page: $<$\pageref{page-A31}$>$­­}\label{fig:A.31}
 	\end{center}	
@@
 
-\begin{table}[H]
-	\caption{Available broadband Landsat satellite images covering the research area of Crete island.}\label{tab:34}
+\begin{table}[H]\footnotesize
+	\caption{Available broadband Landsat satellite images covering the research area of Crete island. Page: $<$\pageref{page-30}$>$}\label{tab:34}
 	\centering
@@
 
-\begin{figure}[H]
+\begin{figure}[H]\footnotesize
 	\centering
@@
 	\subfloat [Landsat-ETM-99-08-08]{\includegraphics[width=0.3\textwidth]{Landsat10.jpg}}
-	\caption{Landsat imagery, Crete island. Previews. Page: $<$\pageref{page-30}$>$}
+	\caption{Landsat imagery, Crete Island. Previews. Page: $<$\pageref{page-30}$>$}
 	\label{fig:A.38}
@@
 		\includegraphics[scale=0.23]{OpenEV-2.jpg}
-		\caption{The point querying shows the selected points and their coordinates within the area of seagrass meadow (green). \href{http://openev.sourceforge.net/}{OpenEV}}\label{fig:A.39}	
+		\caption{The point querying shows the selected points and their coordinates within the area of seagrass meadow (green).
+\href{http://openev.sourceforge.net/}{OpenEV}.Page: $<$\pageref{page-A45-47}$>$}\label{fig:A.39}
 	\end{center}	
@@
 	\caption{\textit{Logarithmic Enhancement to Raster}, applied to the aerial Google Earth image: the seagrass meadow seen as a bright spot of purple color.
-\href{http://openev.sourceforge.net/}{OpenEV}}\label{fig:A.40}
+\href{http://openev.sourceforge.net/}{OpenEV}.
+Page: $<$\pageref{page-A45-47}$>$}\label{fig:A.40}
 	\end{center}	
@@
 		\includegraphics[scale=0.25]{OpenEV-1.jpg}
-		\caption{Raster properties dialog: visualization and spatial info about the image (projection UTM, zone 35, datum WGS-84, etc.)  \href{http://openev.sourceforge.net/}{OpenEV})}\label{fig:A.41}	
+		\caption{Raster properties dialog: visualization and spatial info about the image (projection UTM, zone 35, datum WGS-84, etc.)  \href{http://openev.sourceforge.net/}{OpenEV}).
+Page: $<$\pageref{page-A45-47}$>$}\label{fig:A.41}
 	\end{center}	
@@
 		\includegraphics[scale=0.25]{Ilwis-2.jpg}
-		\caption{Color composite image composed of 3 images of Cretan shelf, Google Earth. \href{http://www.ilwis.org/}{ILWIS}}\label{fig:A.42}	
+		\caption{Color composite image composed of 3 images of Cretan shelf, Google Earth. \href{http://www.ilwis.org/}{ILWIS}. Page: $<$\pageref{page-A45-47}$>$}\label{fig:A.42}	
 	\end{center}	
@@
 	\includegraphics[scale=0.4]{photo-5.jpg}
-	\caption{Placing the 0.5m circle and depth marker in the mattte of \textit{P.oceanica} for photo capture}
+	\caption{Placing the 0.5m circle and depth marker in the mattte of \textit{P.oceanica} for photo capture. Page: $<$\pageref{page-A55}$>$}
 	\label{fig:A.55}

### 2011-02-17 MScDiplom_Lemenkova.tex
30 Comment, 1 Research Question updated

@@
 	\author[1]{by Lemenkova Polina}
-	\affil[1]{MSc student at the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
+	%\affil[1]{MSc student at the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
 	\date{\today}
@@
 	\acro{WASI}{\href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}}
+	\acro{WLS}{\href{http://en.wikipedia.org/wiki/Weighted_least_squares}{Weighted Least Squares}}
 \end{acronym}
@@
 For the \emph{Research Question 1} the Hypothesis $H_o$ claims: seagrass types are not spectrally distinct from other seafloor types with
-varying \textit{in-situ} conditions, which means $H_o: \mu1 =\mu2 =\mu3=...= \mu n$.
+varying \textit{in-situ} conditions, which means $H_o: \mu1 =\mu2 =\mu3=...= \mu n$.\\
 The alternative Hypothesis $H_a$ claims the opposite statement: seagrass is spectrally
 distinct with varying \textit{in-situ} conditions, $H_o:\mu1\neq\mu2\neq\mu3\neq...\neq\mu n$.\vspace{1ex}\\
-For the \emph{Research Question 2} the Hypothesis $H_o$ claims: broadband and hyperspectral remote sensing data can be used for the mapping of \textit{P.oceanica}, because they do not provide enough radiometric information to discriminate sand from seagrass, which means $H_o:\mu1=\mu2=\mu3=...=\mu n$.
-The alternative Hypothesis $H_a$ claims the opposite statement: broadband and hyperspectral sensors do prove to provide enough radiometric information to discriminate sand from seagrass,
-$H_o: \mu1\neq\mu2\neq\mu3\neq ... \neq\mu n$.\vspace{1ex}\\
-The distribution of the spectral responses at every spectral band is assumed to be normal as well as
-the equality of the statistical variances.
+For the \emph{Research Question 2} the Hypothesis $H_o$ claims: broadband and hyperspectral remote sensing data cannot be used for the mapping of \textit{P.oceanica}, because they do not provide enough radiometric information to discriminate sand from seagrass, which means $H_o:\mu1=\mu2=\mu3=...=\mu n$.
+\\ The alternative Hypothesis $H_a$ claims the opposite statement: broadband and hyperspectral sensors do prove to provide enough radiometric information to discriminate sand from seagrass, $H_o: \mu1\neq\mu2\neq\mu3\neq ...
+\neq\mu n$.\vspace{1ex}, and can therefore, be used for the seagrass mapping.\\
+The distribution of the spectral responses at every spectral band is assumed to be normal as well as the equality of the statistical variances.
 The hypothesis testing is suggested to be carried out using the \ac{ANOVA} statistical test. The purpose of
@@
 The Open Office was used for preliminary data view and preprocessing, and included following computations
-at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). 
+at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}, Tab.\ref{tab-16}).\label{page-sumstat} 
 The statistical pre-processing\label{page-40} and analysis were made to display the mean values of spectral reflectance,
@@
 The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made.
-We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}, \ac{OLS}: Fig.\ref{tab:15}\label{page-OLS}\\
+We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, \ac{WLS}: Tab.\ref{tab:27}\label{page-WLS}, \ac{OLS}: Fig.\ref{tab:15}\label{page-OLS}\\
 To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectance at several depths: Tab.\ref{tab:5}\label{page-reg}, Tab.\ref{tab:6}, Tab.\ref{tab:7}.
 The summary of the values of spectral reflectance properties on various depths is presented on the Tab.\ref{tab:A.7}.
-The \textit{k-means cluster analysis}\index[concepts]{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
+The \textit{k-means cluster analysis}\index[concepts]{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23})\label{page-fincc} in which each observation belongs to the cluster with the nearest mean.
 It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
@@
 \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution (spectral reflectance) for each spectral band.
-In x-axis displayed are the areas of 400 – 950 nm in
-spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
+In x-axis displayed are the areas of 400 – 950 nm in spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
 \textit{P.oceanica} spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
@@
 of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
-Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with
-varying \textit{in-situ} conditions, i.e. increasing depth, $H_o: \mu\neq\mu2\neq\mu3\neq ... \neq\mu n$.
+Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing, 
+and seagrass can be spectrally discriminated from carbonate sand with varying \textit{in-situ} conditions,
+ i.e. increasing depth, $H_o: \mu\neq\mu2\neq\mu3\neq ... \neq\mu n$.
 As a result of the statistical testing, we came to the following conclusion.
 The Hypothesis $H_a$ is true, which claims that the spectral discernibility of the seagrass \textit{P.oceanica} is
-distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, $H_o: \mu1\neq\mu2\neq\mu3\neq...\neq\mu n$ .
+distinctly changing and can be discriminated from from other seafloor cover types (carbonate sand) with varying \textit{in-situ} conditions, 
+i.e. increasing depth, $H_o: \mu1\neq\mu2\neq\mu3\neq...\neq\mu n$, which positively answers first research question.
 Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
@@
 	\begin{center}
-		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. \href{http://gretl.sourceforge.net/}{Gretl}­}\label{tab:16}
+		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. \href{http://gretl.sourceforge.net/}{Gretl}. Page: $<$\pageref{page-sumstat}$>$­}\label{tab:16}
 		Summary Statistics, using the observations 301--325\\
@@
 	\rowcolors{1}{Ivory}{GhostWhite}
-	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
+	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).
+\href{http://www.spss.com/}{SPSS}.
+Page: $<$\pageref{page-fincc}$>$}\label{tab:23}
 	\begin{center}
@@
 	\end{center}	
-	\label{tab:23}
 \end{table}
@@
 	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica}.
-\href{http://gretl.sourceforge.net/}{Gretl}}
+\href{http://gretl.sourceforge.net/}{Gretl}.
+Page: $<$\pageref{page-WLS}$>$}\label{tab:27}
 		Model 1: WLS, using observations 1--95\\
@@
 	\end{tabular}
-	\label{tab:27}
 	\end{center}
@@
 \begin{table}[H]\footnotesize
-	\caption{Available broadband Landsat satellite images covering the research area of Crete island. Page: $<$\pageref{page-30}$>$}\label{tab:34}
+	\caption{Available broadband Landsat satellite images covering the research area of Crete Island. Page: $<$\pageref{page-30}$>$}\label{tab:34}
 	\centering

### 2011-02-17 MScDiplom_Lemenkova.tex
Comment 28 of Valentijn: updated Research Question 1

@@
 (less than 1 in a 10 chance). The one-way \ac{ANOVA} highlighted a significant difference between data
-of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
-Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing, 
+of spectral reflectances of \textit{P.oceanica} and carbonate sand at different depths, and and proved spectral discernibility of seagrass \textit{P.oceanica} from carbonate sand (Fig.\ref{fig:4.7})\label{firstq}, i.e.
+true is the opposite statement: Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing,
 and seagrass can be spectrally discriminated from carbonate sand with varying \textit{in-situ} conditions,
@@
 i.e. increasing depth, $H_o: \mu1\neq\mu2\neq\mu3\neq...\neq\mu n$, which positively answers first research question.
-Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
-(Box Plots).
+Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts (Box Plots).
 The graphs were plotted on the basis of the following statistical data calculated from the sampling
@@
 	\includegraphics[scale=0.30]{Fig-30.jpg}
-	\caption{\ac{WASI} water color spectral simulation modeller: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m.
-Pages: $<$\pageref{page-35}$>$, $<$\pageref{page-31}$>$, $<$\pageref{GUI}$>$}
+	\caption{Spectral reflectance of \textit{P.oceanica} and carbonate sand at different depths of 0.5-4.0m.
+\\ Modelled by \ac{WASI} water color spectral simulation.
+Pages: $<$\pageref{page-35}$>$, $<$\pageref{page-31}$>$, $<$\pageref{GUI}$>$, $<$\pageref{firstq}$>$}.
 	\label{fig:4.7}

### 2011-02-17 MScDiplom_Lemenkova.tex
Comment 32 of Valentijn (final): updated Research Question 2, proved hypothesis

@@
 based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
-types, where the spectral signatures were examined. \ac{WASI} software\index[concepts]{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate
-spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
+types, where the spectral signatures were examined.
+\ac{WASI} software\index[concepts]{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
 environmental conditions which influence its colour.
@@
 	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths, as stripes shown spectral bands covered by these sensors.
-Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$}
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$, $<$\pageref{secondq}$>$}
 	\label{fig:4.6}
@@
 \href{http://www.gnuplot.info/}{Gnuplot}­.
-Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-41}$>$}
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-41}$>$, $<$\pageref{secondq}$>$}
 		\label{fig:4.13}
@@
 \href{http://www.gnuplot.info/}{Gnuplot} visualization.
-Page: $<$\pageref{page-39}$>$­}
+Pages: $<$\pageref{page-39}$>$­, $<$\pageref{secondq}$>$­}
 		\label{fig:A.27}
@@
 a way we defined an empirical upper limit to the discriminative potential of these sensors. 
+The analysis of the remote sensing reflectance simulated by these sensors shows that the measured spectra of seagrass \textit{P.oceanica} (Fig.\ref{fig:4.13}, Fig.\ref{fig:A.27}) were statistically different at most of the spectral bands (Fig.\ref{fig:A.25}, Fig.\ref{fig:A.26}), with a 95\% confidence level (p value $< 0.
+05$).
+The F values of the test (F=8.477 as on Tab.\ref{tab:6}) are greater than F critical value, which is 2.64 at 0.05 confidence level.
+This proves that broadband and hyperspectral sensors enable spectral discrimination of seagrass, i.e.
+$H_o: \mu1\neq\mu2\neq\mu3\neq ...
+\neq\mu n$.
+Therefore, we answer the second research question positively, i.e.
+broadband and hyperspectral sensors (Fig.\ref{fig:4.6} and Fig.\ref{fig:4.6W}) provide enough radiometric information for spectral discrimination of seagrass and can therefore be used for mapping of seagrass \textit{P.oceanica}.\label{secondq}
 
@@
 \vspace{1ex}\\
-In \textit{Chapter 3}: \nameref{sec:3}, describing data collection, a videographic approach tested in previous works, has been applied during the summer fieldwork, when we captured imagery and video footage of the seafloor on several routes of the boat in the Ligaria beach.\\
+In \textit{Chapter 3}: \nameref{sec:3}, describing data collection, a videographic approach tested in previous works, has been applied during the summer fieldwork, when we captured imagery and video footage of the seafloor on several routes of the boat in the Ligaria beach.\vspace{1ex}\\
 The finding of \textit{Chapter 4}, \nameref{sec:4} showed that the relationship between the spectral reflectance of various seafloor cover types was tied to depth, i.e.
@@
 10 \% reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33\% in its highest values.
-These results indicate that seagrass \textit{P.oceanica} can in general be detected and discriminated from other seafloor cover types with varying environmental conditions, i.e.
-water column height, by hyperspectral spectroradiometers (\ac{Trios-RAMSES}), which positively answers the \textbf{first research question} of this thesis (``Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ conditions ?").
-\vspace{1ex}\\
-Further in \textit{Chapter 4}, \nameref{sec:4}, studies of the broadband versus narrowband sensors demonstrate that simulated spectra of the seagrass, made using \ac{WASI} modeller, have the best results at \ac{CZCS} scanner, especially devoted to the measurement of ocean color.
+These results indicate that seagrass \textit{P.oceanica} can in general be detected and discriminated from other seafloor cover types (carbonate sand) with varying environmental conditions, i.e.
+water column height, by hyperspectral spectroradiometers (\ac{Trios-RAMSES}), which positively answers the \textbf{first research question} of this thesis (``Is \textit{P.oceanica} spectrally distinct from carbonate sand with varying \textit{in-situ} conditions ?").
+Further in \textit{Chapter 4}, \nameref{sec:4}, studies of the broadband and narrowband sensors demonstrate that simulated spectra of the seagrass, made using \ac{WASI} modeller, have the best results at \ac{CZCS} scanner, especially devoted to the measurement of ocean color.
 The spectrum of \textit{P.oceanica} reflectance, simulated for \ac{CZCS}, covers the wavelength interval of 400-800 nm, and is distinctive for various depths.
 Other remote sensors (\ac{MODIS}, \ac{SeaWiFS}) may also be used for the seagrass mapping, because their technical characteristics enable to spectrally discriminate \textit{P.oceanica} seagrass from other seafloor cover types (Fig.\ref{fig:4.6}), particularly carbonate sand as tested in the current work.
-Therefore, the \textbf{second research question} of this MSc thesis (``Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\
+Therefore, the \textbf{second research question} of this MSc thesis (``Do broadband and hyperspectral sensors provide enough radiometric information for spectral discrimination of seagrass,
+and therefore, can be used for mapping of \textit{P.oceanica} ?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\
 The graphs showing optical properties of seawater with and without sediments (Fig.\ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:4.2}) focused on spectral variability of the water with changed physical and chemical content.
@@
 \ac{SPSS}.
-Page: $<$\pageref{page-reg}$>$}
+Page: $<$\pageref{page-reg}$>$}\label{tab:5}
 	\begin{center}
@@
 	The independent variable is wavelength. 	
-	\label{tab:5}
 \end{table}
@@
 \ac{SPSS}.
-Page: $<$\pageref{page-reg}$>$}
+Pages: $<$\pageref{page-reg}$>$, $<$\pageref{secondq}$>$}\label{tab:6}
 	\begin{center}
@@
 	The independent variable is wavelength. 	
-	\label{tab:6}
 \end{table}
@@
 \href{http://www.spss.com/}{SPSS}.
-Page: $<$\pageref{page-reg}$>$}
+Page: $<$\pageref{page-reg}$>$}\label{tab:7}
 	\begin{center}
@@
 	 Based on the asymptotic chi-square approximation.	
-	\label{tab:7}
 \end{table}
@@
 	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
-	\caption{WASI-simulated remote sensing reflectance of \textit{P.oceanica} at various sensors. Page: $<$\pageref{page-33}$>$}
+	\caption{WASI-simulated remote sensing reflectance of \textit{P.oceanica} at various sensors. Pages: $<$\pageref{page-33}$>$, $<$\pageref{secondq}$>$}
 	\label{fig:4.6W}
@@
 		\includegraphics[scale=0.35]{GNU-11.jpg}
-		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display. Pages: $<$\pageref{page-21}$>$, $<$\pageref{dataprep}$>$­}
+		\caption{Spectral reflectance of \textit{P.oceanica}.
+Measurement series 25-50.
+\\ \href{http://www.gnuplot.info/}{Gnuplot} display.
+Pages: $<$\pageref{page-21}$>$, $<$\pageref{dataprep}$>$­, $<$\pageref{secondq}$>$}
 		\label{fig:A.25}
@@
 \\ \href{http://www.gnuplot.info/}{Gnuplot}­.
-Page: $<$\pageref{page-21}$>$}
+Page: $<$\pageref{page-21}$>$, $<$\pageref{secondq}$>$}
 		\label{fig:A.26}

### 2011-02-17 MScDiplom_Lemenkova.tex
Some final improvements

@@
 	%\affil[1]{MSc student at the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
-	\date{\today}
+	\date{March 2011}
 	\maketitle	
@@
 equipment have been conducted for the measurements of depths.
+\pagebreak
 
@@
 remote sensing techniques: spectral modelling by means of the \ac{RTM} \ac{WASI}, underwater videometric measurements made by \ac{Olympus ST} camera, aerial Google Earth and satellite images from different sources, spatial \ac{GIS} and statistical analysis.
+\pagebreak
 
@@
 	\end{itemize}
-\pagebreak
 
@@
 The coefficient of attenuation remains equal to 1.0546, as set up by default at \ac{WASI}.\vspace{1ex}\\
-We also assume that concentration of non-chlorophyll particles (absorption at 0) as well as concentration of small suspended particles is equal to zero, so we do not count them in this work.
+We also assume that concentration of non-chlorophyll particles (absorption at $\lambda_0$) as well as concentration of small suspended particles is equal to zero, so we do not count them in this work.
 Exponent of \ac{CDOM} (Gelbstoff) absorption is accepted as 0.0140.
-Finally, the \ac{BRDF} of bottom reflectance (sand), which defines the reflection of light  is at an opaque surface (Fig.\ref{fig:BRDF})\label{page-11}, is assumed to be 0.318 $sr^{-1}$.
+Finally, the \ac{BRDF} of bottom reflectance (sand), which defines reflection of light  is at an opaque surface (Fig.\ref{fig:BRDF})\label{page-11}, is assumed to be 0.318 $sr^{-1}$.
 
@@
 substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. 
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[scale=0.63]{UML_ClassDiagram.jpg}
-	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}. \\ Page: $<$\pageref{page-9}$>$}
-	\label{fig:1.10}
-\end{figure}
-
 The pre-processing of the images includes imagery corrections for atmospheric noises and effects of the water
@@
 Supervised, Maximum Likelihood).
+\begin{figure}[H]
+	\centering
+	\includegraphics[scale=0.63]{UML_ClassDiagram.jpg}
+	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}, \href{http://www.gimp.org/}{Gimp}. \\ Page: $<$\pageref{page-9}$>$}
+	\label{fig:1.10}
+\end{figure}
 \pagebreak
@@
 The distinguishing spectral signatures for various seafloor types (e.g. seagrass species, coral reefs,
-various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})\label{page-31} exist in well-defined and narrow (10–20 nm)
+various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})\label{page-31} exist in well-defined and narrow ($10 - 20$ nm)
 wavelength ranges. 
@@
 The Open Office was used for preliminary data view and preprocessing, and included following computations
-at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}, Tab.\ref{tab-16}).\label{page-sumstat} 
+at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}, Tab.\ref{tab:A.8}).\label{page-sumstat} 
 The statistical pre-processing\label{page-40} and analysis were made to display the mean values of spectral reflectance,
@@
 As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand.
-The \textit{autocorrelation}\index[concepts]{Data!statistical analysis!autocorrelation}, a cross-correlation\index[concepts]{Data!statistical analysis!cross-correlation} of an observed value with itself, made on two different days (14th and 15th October), shows the similarity between observations as a function of the time separation between them.
-As can be seen on Fig.\ref{fig:A.21}\label{A-23}, Fig.\ref{fig:A.22}\label{A-24} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
+The \textit{autocorrelation}\index[concepts]{Data!statistical analysis!autocorrelation}, a cross-correlation\index[concepts]{Data!statistical analysis!cross-correlation} of an observed value with itself, made on two different days (14th and 15th October), shows the similarity between observations as a function of the time separation between them,
+as can be seen on Fig.\ref{fig:A.21}\label{A-23}, Fig.\ref{fig:A.22}\label{A-24} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise.
 The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{GRETL} software (Fig.\ref{tab:28}\label{page-a-c}).
@@
 It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
-To estimate spectral density of a spectral reflectance measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30})\label{page-period}, which corresponds to the general spectrum of the observations with representation of a variable quantity.
+To estimate spectral density of optical measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30})\label{page-period}, which corresponds to the general spectrum of the observations with representation of a variable quantity.
 Thus, the periodogram (Fig.\ref{fig:A.31})\label{page-A31} of the series of observation\index[concepts]{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to exclude (can be seen clearly at the end of the graphs, e.g.
-Fig.\ref{fig:A.12} )
-and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:A.14}, enlarged part of 500-660nm).\\
+Fig.\ref{fig:A.12}) and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:A.14}, enlarged part of 500-660nm).\\
 The visualization of the data plotting was made by means of Gnuplot software, which enables fine
@@
 \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution (spectral reflectance) for each spectral band.
-In x-axis displayed are the areas of 400 – 950 nm in spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
+In x-axis displayed are the areas of 400-950 nm in spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
 \textit{P.oceanica} spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
@@
 The graphs were plotted on the basis of the following statistical data calculated from the sampling
-measurements data: Table 2.
-
-\begin{table}
+measurements data: Tab.\ref{tab:A.8}, Tab.\ref{tab:A.29}, Tab\ref{tab:4.1}.\label{last}
+
+\begin{table}[H]
 \rowcolors{1}{Khaki}{LightYellow}
@@
 
-where the abbreviations stand for the following values:
+The abbreviations stand for the following values:
 Mean: Average of the data to be plotted, AVERAGE\index[concepts]{Data!statistical analysis!average} {data}
@@
 The Kappa coefficient takes into account chance agreement and thus, shows the reduction in errors generated by a classification process compared with the errors which could be received by a completely random classification – in other words, it evaluates quality of the classification.
-The results of the classification can be seen on the Fig.
-(see appendices).
-The overall map accuracy is 68\%, which means that 68\% of the pixels are classified to the correctly chosen seafloor cover type.
+The overall map accuracy by supervised classificationis 72\% (Fig.\ref{tab:kappa-2}), which means that 72\% of the pixels are classified to the correctly chosen seafloor cover type in case, and in case of unsupervised classification we received result of 64\% (Fig.\ref{tab:kappa-1})  which proves that supervised classification is preferable method for seagrass mapping.
 \pagebreak
@@
 \\
-The reflectance spectra of \textit{P.oceanica} show (Fig.\ref{fig:4.10}, Fig.\ref{fig:A.12}, Fig.
-\ref{fig:A.12}, Fig.\ref{fig:A.13}, Fig.\ref{fig:A.25}, Fig.
-\ref{fig:A.26}) a values maximum between 450 nm and 600 nm, first, because of the chlorophyll absorption peak at 465 and 665nm (Fig.
+The reflectance spectra of \textit{P.oceanica} show (Fig.\ref{fig:4.10}, Fig.\ref{fig:A.12}, Fig.\ref{fig:A.13}, Fig.\ref{fig:A.25}, Fig.\ref{fig:A.26}) a values maximum between 450 nm and 600 nm, first, because of the chlorophyll absorption peak at 465 and 665nm (Fig.
 \ref{fig:A.2-2})\label{chlorophyll}, secondly, because of the weakening of \ac{CDOM} (or Gelbstoff) in the blue part of the \ac{VIS} spectrum, as it most strongly absorbs short wavelength light in blue to ultraviolet range, and finally, because the absorption of the seawater increases in the red part of the \ac{VIS} spectra.
@@
 The alterations in the individual spectral signatures of single measurements (e.g.
-on Fig.\ref{fig:4.13}, Fig.\ref{fig:A.25}, Fig.\ref{fig:A.27}, Fig.\ref{fig:A.28}) reflect individual health properties of leaves: different nitrogen and chlorophyll content causing diverse colour pigmentation and light absorption, water content in leaves, plant physiological conditions, which vary across seagrass meadow, shoot morphology, etc.
+on Fig.\ref{fig:4.13}, Fig.\ref{fig:A.25}, Fig.\ref{fig:A.27}, Fig.\ref{fig:A.28}) reflect individual health properties of leaves: different nitrogen and chlorophyll content causing diverse colour pigmentation and light absorption, water content in leaves and plant physiological conditions, which vary across seagrass meadow, shoot morphology, etc.
 The differences in spectral reflectance values of the measurements taken on various days might have been caused by the impact of atmospheric conditions, such as solar radiation and sun illumination by different zenith angle.\vspace{1ex}\\
@@
 \section{appendices}\label{app}
-\subsection{\footnote{The page where the illustration is explained in the main text of MSc thesis is given in angle brackets (e.g., Page: $<$\pageref{page-40}$>$).} Capturing aerial imagery from the Google Earth}
+\subsection{Capturing aerial imagery from the Google Earth}
 %\appendixname{\textbf{1: \\Capturing aerial imagery from the Google Earth: grabbing process}}
@@
 Generalisation up to step 20 nm.
-Page: $<$\pageref{page-40}$>$}\label{tab:4.1}
+Pages: $<$\pageref{page-40}$>$, $<$\pageref{last}$>$}\label{tab:4.1}
 	\begin{center}
@@
 Measured on Agia Pelagia beach, 15th October.
-Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-40}$>$}\label{tab:A.8} \\
+Pages: $<$\pageref{page-28}$>$, $<$\pageref{page-40}$>$, $<$\pageref{last}$>$}\label{tab:A.8} \\
 	  \multicolumn{1}{|l|}{\textbf{wl}} &
@@
 Measured on Agia Pelagia beach, 14th October.
-Page: $<$\pageref{page-40}$>$}\label{tab:A.29} \\
+Pages: $<$\pageref{page-40}$>$, $<$\pageref{last}$>$}\label{tab:A.29} \\
  		 \multicolumn{1}{|l|}{\textbf{wl}} &
